### 2023-01-20 my_paper/FinalDraft.tex
push project to github

\documentclass{wptemp}
\usepackage{adjustbox}
\usepackage{dcolumn}    % aligning decimals
    \newcolumntype{d}[1]{D{.}{.}{#1}}
% *****************************************************************
% Estout LaTeX wrapper
% *****************************************************************

%%Original code developed by Jörg Weber: see
%% https://www.jwe.cc/2012/03/stata-latex-tables-estout/
%% and
%% https://www.jwe.cc/blog/

\let\estinput=\input % define a new input command so that we can still flatten the document

\newcommand{\estwide}[3]{
		\vspace{.75ex}{
			%\textsymbols% Note the added command here
			\begin{tabular*}
			{\textwidth}{@{\hskip\tabcolsep\extracolsep\fill}l*{#2}{#3}}
			\toprule
			\estinput{#1}
			\bottomrule
			\addlinespace[.75ex]
			\end{tabular*}
			}
		}

\newcommand{\estauto}[3]{
		\vspace{.75ex}{
			%\textsymbols% Note the added command here
			\begin{tabular}{l*{#2}{#3}}
			\toprule
			\estinput{#1}
			\bottomrule
			\addlinespace[.75ex]
			\end{tabular}
			}
		}

% Allow line breaks with \\ in specialcells
\newcommand{\specialcell}[2][c]{%
    \begin{tabular}[#1]{@{}c@{}}#2\end{tabular}
}

\newcommand{\sym}[1]{\rlap{#1}}% Thanks David Carlisle

%%%%%%%%%%% End of wrapper %%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%% TiKz %%%%%%%%%%%%%%%%%%%%%
\usepackage{tikz}
\usetikzlibrary{shapes.geometric, arrows}

\tikzstyle{startstop} = [rectangle, rounded corners, minimum width=3cm, minimum height=1cm,text centered, draw=black, fill=red!30]

\tikzstyle{io} = [trapezium, trapezium left angle=70, trapezium right angle=110, minimum width=3cm, minimum height=1cm, text centered, text width =5cm, draw=black, fill=blue!30]

\tikzstyle{process} = [rectangle, minimum width=3cm, minimum height=1cm, text centered, text width =5cm, draw=black, fill=orange!30]

\tikzstyle{decision} = [diamond, minimum width=3cm, minimum height=1cm, text centered, draw=black, fill=green!30]

\tikzstyle{arrow} = [thick,->,>=stealth]

\begin{document}

\linespread{1}{\title{{\fontfamily{ptm}
\selectfont{\Large{\MakeUppercase{The Impact of Hispanic Last Names and Identity on Labor Market Outcomes}}}}}}
	\author{{\fontfamily{pbk}\selectfont\large{\textsc{Hussain Hadah}}\thanks{Department of Economics, University of Houston, 206A McElhinney Hall, 3623 Cullen Boulevard, Houston, TX 77204-5019, United States (e-mail: \email{hhadah@uh.edu}).}}}
	\date{\fontfamily{pbk}\selectfont\normalsize{\textsc{\today}}}
	\maketitle

%\begin{center}
%\href{https://hussainhadah.com/worksinprogress/hispanicdiscrimnation/}{\textcolor{green!15!black!30!blue}{\footnotesize{\textsc{Click Here to Get the Most Updated Version}}}}
%\end{center}

\pagenumbering{arabic}

\begin{abstract}
In this paper, I compare the children of inter-ethnic marriages to study the impact of having a Hispanic last name. While men born to Hispanic father-White mothers earn less than those born to White father-Hispanic mothers, the gap could be explained entirely by educational differences. I also study the effect of identifying as Hispanic on earnings. I find that men that identify as Hispanic earn significantly less than those that do not, even after controlling for educational differences. 
\end{abstract}

\pagebreak

%\section{Questions}
%
%In this paper, I will answer the following questions:
%
%\begin{enumerate}
%\item Do Hispanics face discrimination in the labor market due to an ethnic signal (last name)?
%
%\item Does the marriage market penalize or reward Hispanics for their ethnicity? 
%
%\item How do Hispanic men and women fare in inter-ethnic marriages? 
%
%\item Is there assortative matching in inter-ethnic marriages?
%
%\item Is there a difference in identity cost for those that identify as Hispanics and those that I induce are Hispanic through their parents' place of birth?
%
%\item What are the minority groups that faced a marriage market penalty? (Historical analysis)
%
%\item How did the marriage market penalty change over time for the different minority groups?
%
%\end{enumerate}
\section{Introduction}\label{sec:intro}

There is a plethora of evidence of ethnic and racial discrimination in the labor market \citep{bayer2018divergent, charles2008prejudice, card1992school, fryer2004causes, rubinstein2014pride, bertrand2004emily, juhn1991accounting}. Hispanics constitute a large and growing portion of the population in the United States. As the number of Hispanics keeps increasing in the labor force, identifying whether they face ethnic discrimination grows in importance. Consequently, questions on assimilation and mobility grow in importance \citep{chettyUnitedStatesStill2014, chettyEffectsExposureBetter2016,chettyFadingAmericanDream2017,abramitzkyImmigrantsAssimilateMore2020a, abramitzkyNationImmigrantsAssimilation2014,abramitzkyCulturalAssimilationAge2016,chettyWhereLandOpportunity2014}, how identity could shift public opinions toward trade \citep{grossmanIdentityPoliticsTrade2021}, and how racial and gender attitudes affect the racial and gender earnings gaps \citep{charlesPrejudiceWagesEmpirical2008,charlesEffectsSexismAmerican2018}. Thus, it is important to understand whether a person's ethnicity affects their labor market outcomes.

In this paper, I answer the following questions. Does having a Hispanic last name affect labor market outcomes? What is the effect of identifying as Hispanic on earnings? Moreover, I aim to show that comparing Hispanic Whites to non-Hispanic Whites might create an artificially higher earnings gap since the two groups differ on many observable characteristics. Others have attempted to compare how native-born White Hispanics fare to non-Hispanic Whites and foreign-born Hispanics. In \citet{antman2020ethnic,antmanEthnicAttritionObserved2016,antmanEthnicAttritionObserved2016a,antmanEthnicAttritionAssimilation2020}, the authors compare the health and educational outcomes of Hispanic Whites to non-Hispanic Whites and native-born Hispanics to foreign-born Hispanics. They find gaps in education and health between Hispanics and Whites. They also find that native-born Hispanics are more likely than their foreign-born counterparts to report poor health. \citet{davilaChangesRelativeEarnings2008}, the authors documents many gaps in labor market outcomes between Hispanics and Whites. They attribute a big part of this gap to differences in education, experience, immigration status, and regional differences.

The US population is growing in diversity. The proportion of non-Whites has increased by more than 10 percentage points from 1995 to 2019 (see figure \ref{fig:1}), and the number of Hispanics has grown by 9 percentage points over the same period (see figure \ref{fig:2}). Native-born White Hispanic men earn 21\% less than White men \citep{duncan2018identifying}. However, a substantial portion of the earnings gap is due to educational differences between Hispanics and Whites \citep{duncan2006hispanics, duncan2018socioeconomic}. Therefore, understanding discrimination against Hispanics and their children is an important task. Discrimination against Hispanics has a couple of negative consequences. First, if Hispanics face a substantial labor market penalty for their ethnic identity, it could hinder assimilation and integration. Second, a significant labor market penalty contributes to the growing wage gap between Whites and other minority groups. In this paper, I examine the role of having a Hispanic last name and identifying as Hispanic on labor market outcomes.

Identifying discrimination is difficult because of factors that affect labor market outcomes that are unobservable to economists---like innate ability. Also, I should be able to disentangle skills differentials between Hispanics and non-Hispanic Whites. To causally identify ethnic discrimination against Hispanics and its effects on labor market outcomes, a researcher should compare two identical people that only differ in their last name. Since this approach is unfeasible, researchers took different approaches. \citet{bertrand2004emily} and \citet{fryer2004causes} conducted audit studies to examine the effect of Black sounding names on employment discrimination. This approach, however, has its drawbacks. Audit studies only observe callbacks, not wages.

I will use a method developed by \citet{rubinstein2014pride} in this research. I will compare children from intermarriages. More precisely, I will reach children of Hispanic fathers and White mothers (henceforth HW) to children of White fathers and Hispanic mothers (henceforth  WH). This approach stems from the fact that marriages are not random. Couples match on several observable characteristics like income, schooling, socio-economic background, etc. \citep{averettBetterWorseRelationship2008, averettEconomicRealityBeauty1996, beckerTheoryMarriagePart1973, beckerTheoryMarriagePart1974, beckerTreatiseFamily1993, browningCollectiveUnitaryModels2006, chiapporiFatterAttractionAnthropometric2012}. Children of HW and WH marriages have more similar observable characteristics than children of homogeneous marriages--- i.e., White fathers-White mothers and Hispanic fathers-Hispanic mothers. Moreover, children from a Hispanic father and White mother household will have a Hispanic last name from their fathers, allowing me to test for ethnic signals on annual log earnings.

Using the Current Population Survey (CPS) from 1994 to 2019 and the 1970-1990 US censuses, I will estimate the effect of having a Hispanic last name on the labor market outcomes. In other words, I will approximate the variation in labor market perception of ethnic signals, in this case having a Hispanic last name, on annual log earnings. I also study the effect of identifying as Hispanic on labor market outcomes.

To retrieve the effect of having a Hispanic last name, I compare the children of inter-ethnic marriages between Whites and Hispanics. I find that while men born to White father-Hispanic mothers earn less than men born to Hispanic father-White mothers, this gap is entirely explained by educational differences. Thus, I do not find a significant effect of having a Hispanic last name. Finally, I find that men identifying as Hispanic earn significantly less than those who do not, even when I control ancestry and education.

A long list of economic papers investigated taste-based discrimination, statistical discrimination, and Black-White and gender earnings gaps that date back to the seminal work of \citet{becker2010economics}. \citet{bayer2018divergent} investigated the Black-White wage gap's narrowing and then divergence. \citet{card1992school} argued that the Black-White wage gap narrowed from 1960 to 1980 due to improvement in the quality of Black schools, and \citet{juhn1991accounting} showed that skills play a big part in the Black-White wage gap. Moreover,  \citet{charles2008prejudice} provided evidence that 25\% of the Black-White wage gap is due to prejudice. \citet{black1995discrimination} introduced a search model showing that with a taste for discrimination, minorities earn less than the majority group even when hired by an unprejudiced firm. In comparison, \citet{altonji2001employer} found that firms do not statistically discriminate based on race and ethnicity. Little has been done, however, about discrimination against Hispanics, which offers an opportunity to exploit the observable racial similarities between Hispanics and Whites.

This is not the first attempt to use names as signals for race and ethnicity \citep{fryer2004causes, rubinstein2014pride, bertrand2004emily}. \citet{fryer2004causes} point out that names can be a predictor of a person's race. Specifically, they provide a rising pattern among Blacks having different names than Whites. They did, however, find that having a Black name, after controlling for the home environment at birth, does not affect their labor market outcomes. \citet{rubinstein2014pride} compared the children of mixed marriages between Sephardic and Ashkenazis in Israel. They found that workers with Sephardic last names earn substantially less than those with an Ashkenazi last name. \citet{bertrand2004emily} conducted an audit study by sending employers identical resumes that differ in the ethnic and racial signal of a name (Black sounding name versus a White sounding one). They found that resumes with Black-sounding names received substantially fewer callbacks than their White counterparts.

The rest of the paper is organized as follows. Section \ref{sec:data} provides an overview of the data and summary statistics of the sample I am using. Section \ref{sec:emp_model} introduces a formal empirical model. I go over the results in section \ref{sec:results}. Finally, I offer a brief conclusion in section \ref{sec:con}.

\section{Data}\label{sec:data}

For this paper, I will use two data sets. The Integrated Public Use Microdata Series (IPUMS) Current Population Survey (CPS) Annual Social and Economic (ASEC) \citep{cps2019} and the 1970 to 1990 US censuses \citep{acs2019}.

I use the CPS' rich data set to study the effect of having a Hispanic last name on a person's labor market outcomes. I take advantage of the fact that the CPS asks for parents' place of birth, ethnicity, and race. The data spans the period between 1994--- the earliest sample to ask about a parent's place of birth--- to 2019. Moreover, since the CPS does not provide data on parents' characteristics, essential to determine the family background, I use the census to construct synthetic parents. The census offers a more significant sample of potential parents. Similar to the CPS, the 1970 to 1990 censuses ask about the person's place of birth and the individual's race and ethnicity. I employ this information to construct "synthetic parents" using a method developed by \citet{rubinstein2014pride}.

I construct the synthetic parents by linking husbands and wives in the census data to each other. I assume that parents have children between the ages of 25 and 40. I then link these synthetic parents using the year they were surveyed to the year of birth of the "children" in the CPS sample.

\subsection{Children of the four parents' types}
I will use the CPS for my primary analysis of the effect of having a Hispanic last name on earnings. I restrict my sample to Whites, United States-born citizens aged 25 to 40 and born between 1970 to 1990. Taking advantage of data on parents' place of birth, I divide the sample into four groups depending on their parent's ethnicity. Mothers or fathers are Hispanic if they were born in a Spanish-speaking country and Puerto Rico, and White if they were born in the United States.\footnote{The list of Spanish Speaking countries include: Argentina, Bolivia, Chile, Colombia, Costa Rica, Cuba, Dominican Republic, Ecuador, El Salvador, Guatemala, Equatorial Guinea, Honduras, Mexico, Nicaragua, Panama, Paraguay, Peru, Spain, Uruguay, Venezuela.} Therefore, an observation can be the product of four types of parents: 
\begin{enumerate}
\item White father and White mother (hereafter WW) 
\item White father and Hispanic mother (hereafter WH)
\item Hispanic father and White mother (hereafter HW)
\item Hispanic father and Hispanic mother (hereafter HH).
\end{enumerate}
The distribution of the four types of children is presented in figure \ref{fig:dist}. The majority of the sample, 95.14\%, is WW children. The second biggest group is HH, which constitutes 3.36\% of the sample. Inter-ethnic children, WH and HW, make up 1.78\% of the sample with 60,147 observations. Even though WH and HW are only 1.78\% of the sample, I have plenty of observations to carry out an analysis. The summary statistics for the children of the four types of marriages are presented in table \ref{tab:c&p1}. Children of WW marriages (column i) do better on every measure while children of HH parents (column iv) do worse than other children on every measure.

On the one hand, WW children are more educated, with approximately 14 years of education for men and women. man and woman WW children have an employment rate of 95\%. They have log hourly earnings of 2.557 for men and 2.435 for women. Also, their annual log earnings are 10.529 for men and 10.297 for women.

On the other hand, children of HH marriages have an average of 12.898 and 13.204 years of education for men and women, respectively. HH men and women have an employment rate of 93\%. Their annual log earnings are 10.365 and 10.175 for men and women, respectively.

Inter-ethnic children, WH and HW, have lower educational attainment and income than WW children, but they are better than HH children. Also, WH and HW are comparable, unlike WW and HH children. These two groups are important to my analysis. Children of HW are going to be the treatment group. HW's children have a Hispanic father, who will carry his Hispanic last name. At the same time, the WH children have White fathers and will take his non-Hispanic, or White, last name. A first glance at the summary statistics of these two groups, table \ref{tab:c&p1} columns (ii) and (iii) can show that they are different from the children of non-intermarriages.

WH children have an average of 13.366 and 13.741 years of education for men and women. The employment rate for men is 94.5\% and for women 95.0\%. HW man and woman children obtained 13.072 and 13.270 years of schooling. Men's employment rate is 92.7\% and women's 93.1\%.

In panel D of table \ref{tab:c&p1}, I present the rates at which members of each group self-identify as Hispanic \footnote{Hispanic self-identification is observed when an individual notes that they are Hispanic on the CPS questionnaire}. The vast majority of HH children, 97\%, identify as Hispanic. There is considerable attrition in Hispanic identification of inter-ethnic children. Among WH children, 82.1\% of men and 85.0\% of women identified as Hispanic. Among HW children, 87.9\% of men and 85.9\% of women identified as Hispanic. These results align with the findings of a substantial body of literature that noted this attrition among children of immigrants, especially those coming from inter-ethnic marriages \citep{duncan2017complexity, duncan2018identifying, duncan2020new, antman2020ethnic}. Finally, a small portion of WW children, 6\%, identified as Hispanic. These 6\% of WW children are most likely third generation--- or more--- immigrants. For my analysis, I dropped the WW observations identified as Hispanic to compare with non-Hispanic Whites.

In columns (v) and (vi), I calculate the differences in means between HH-WW (column v) and HW-WH (column vi). Overall, HH children do worse than WW children. HH has one less year of education compared to WW, 2.1 percentage points less likely to be employed, HH men earn 16.4\% less than WW men, and HH women make 12.3\% less than WW women. Therefore, these two groups, HH and WW, are different in many observable.

The differentials between WH and HW are not as severe as the ones between HH and WW. HW children are still worse off than WH but are way better than HH. HW men attain 0.294 fewer years of education than WH men. HW women achieve 0.471 fewer years of schooling than WH women. HW children are 1.8-1.9 percentage points less likely to be employed. HW men earn the same annual earnings as WH men, and HW women earn 10.3\% less. These differences show that HW and WH children are different from children of homogeneous marriages and, thus, are more comparable.

Moreover, in table \ref{tab:c&p2}, I present the summary statistics for the four types but limit the sample to those that only identify as Hispanic\footnote{Limiting the sample to those that identify as Hispanic only affects WH, HW, and HH. Members of the WW group are not affected by this restriction.}. The same pattern discussed above between HH--WW and HW--WH persists. The one difference is that those who identify as Hispanic do slightly worse.
 %In columns five and six of table \ref{tab:c1}, I present the gaps between HH and WW children and HW and WH children, respectively. The gap between HH and WW children (column five) shows a consistent story, HH children are worse in education and work more. HH children have more kids. While the gap between HW and WH does not have a clear-cut direction. Even though WH children obtain more years of education and have a higher employment rate, HW and WH men have the same fertility rates. HW women have a higher fertility rate than WH women. WH and HW women have the same employment rate.

\subsection{Synthetic parents}

Using the 1970 and 1990 censuses, I constructed a panel of synthetic parents. The sample includes married White men and women. Even though the census asks a person whether they are Hispanic or not, I took advantage of the questions on place of birth to create a proxy for ethnicity. I consider a Hispanic person as White and born in a Spanish-speaking country. Consequently, Whites in the sample are people who are White and native-born. Using the information provided in the census, I can link husbands and wives with each other. I assume that parents have children between the ages of 25 and 45. Therefore, my sample consists of married White men and women that were born in the 1930 to 1965 cohorts\footnote{The construction of "synthetic parents" follows the method used by \citet{rubinstein2014pride}.}.

I show the distribution of the four types of couples in table \ref{tab:mat1}. White husbands and White wives (WW) make up the majority of couples in the sample, 98.18\% (681,206 couples). Hispanic husbands and wives (HH) are the second largest group making up 1.04\% (7,193 couples)of all couples. White husbands and Hispanic wives (WH) couples are 0.40\% (2,750 couples) of the sample while Hispanic husbands and White wives (HW) are 0.39\% (2,710 couples).

I present the summary statistics of the four types of synthetic parents in table \ref{tab:c&p1}'s parent's panel. WW couples have higher education, 11.9 years for husbands and 11.8 for wives. As a household, WW couples have 25.194 years of schooling. Husbands in HH marriages have 8.5 years of education, while women have 8.2 years of schooling. As a household, HH couples have 17.539 years of education.

WH husbands have 10.3 years of education, while wives have an average of 9.6 years, the lowest of any group. WH household attained a total of 22.308 years of schooling in total. HW husbands and wives obtained 9.7 and 9.8 years of education, respectively. HW household attained a total of 21.473 years of schooling in total.

In columns (v) and (vi) of table \ref{tab:c&p1}, I calculated the differences in means between HH-WW households and HW-WH households. An HH Hispanic husband attained 3.779 fewer years of education than a WW husband. While an HH wife acquired 3.876 fewer years of schooling than a White WW wife. The total educational difference between HH and WW households equals 7.655 years of education. The educational differences between HW and WH are less severe than those between HH and WW. A Hispanic husband in intermarriage attained 1.560 fewer years of schooling than a White husband in intermarriage. A White wife in intermarriage attained 0.726 years of education, more than a Hispanic wife in intermarriage. The difference in total years of education between HW and WH households is 0.835, which is an 89\% decrease from the HH--WW gap. These differences show that couples that intermarry are better and different than those that do not.

%In table \ref{tab:p1} columns five and six, I provide the  HH-WW and HW-WH gaps. WW couples are more educated and earn more than HH couples. Also, WH husbands are more educated and earn more than HW husbands. In comparison, WH wives are more educated than HW wives. Even though White husbands in intermarriages are more educated and earn more than husbands in HW marriages, they are more comparable to each other than the non-intermarriage couples.

\section{Empirical Strategy}\label{sec:emp_model}
In this section, I will present two empirical strategies. The first empirical strategy will estimate the effect of having a Hispanic last name on annual log earnings. The second empirical strategy will estimate the effect of identifying as Hispanic on log yearly earnings.

The difference in means between Hispanics and non-Hispanic Whites could result from discrimination. It can also be caused by differences in innate abilities, skills, and parental investments during infancy. To avoid these threats to identification, I compare children of intermarriages, HW and WH. WH and HW children are equally Hispanic physically but provide employers/labor market with different signals.\footnote{WH and HW children are both half White, half Hispanic.} WH children will have a non-Hispanic last name, while HW children will have a Hispanic last name. This is a method developed by \citet{rubinstein2014pride}.

\subsection{Estimating the effect of having a Hispanic last name}

Let $Y_{it}$ be the annual log earnings of person $i$ at time $t$. $WH_{i}$, $HW_{i}$ and $HH_{i}$ are indicator variables for the type of parents person $i$ has. $X_{it}$ is a vector of controls, $\gamma_{t}$ are year fixed effects and $\phi_{it}$ represents the error term. The equation for this strategy is written as follows:
\begin{equation} \label{eq:1a}
Y_{it} = \beta_{1} WH_{i} +\beta_{2} HW_{i} + \beta_{3} HH_{i} + X_{it} \pi + \gamma_{t}+\phi_{it}
\end{equation}

$\beta_{1}, \beta_{2}$ and $\beta_{3}$ are the coefficients of interest in this specification. They estimate the earnings gaps between the three groups and children of White-White parents. The difference between $\beta_{1}$ and $\beta_{2}$ captures the effect of having a Hispanic last name. This comes from the fact that the WH and HW children are similar. On the one hand, WH children will have a White father and carry his White last name. On the other hand, HW children will have a Hispanic father and carry his Hispanic last name. Therefore, \textit{ceteris paribus}, the difference will capture the effect of having a Hispanic last name.

\subsection{Estimating the effect of identifying as Hispanic}

In this section, I will present an estimation strategy that would allow me to capture the effect of identifying as a Hispanic. The empiricism in the section follows the theoretical identity model developed by \citet{akerlof2000economics}.

Let $Y_{it}$ be the annual log earnings of person $i$ at time $t$. $WH_{i}$, $HW_{i}$ and $WW_{i}$ are indicator variables for the type of parents person $i$ has. $X_{it}$ is a vector of controls, $\gamma_{t}$ are year fixed effects and $\phi_{it}$ represents the error term. The addition to the previous model is the $Hispanic_{i}$ indicator variable. $Hispanic_{i}$ is equal to one if a person identifies as Hispanic and zero otherwise. I interact this term with the $WH_{i}$, $HW_{i}$ and $HH_{i}$.  The regression is presented in the following equation.

\begin{align} \label{eq:iden}
Y_{it} &= \beta_{1} HW_{i} +  \beta_{2} WH_{i} + \beta_{3} HH_{i} +\\
& \beta_{4} WH_{i} \cdot Hispanic_{i} + \beta_{5} HW_{i}\cdot Hispanic_{i} +  \beta_{6} HH_{i} \cdot Hispanic_{i}+ \nonumber \\
&X_{it}^{\prime} + \gamma_{t}+\phi_{it} \nonumber
\end{align}

The coefficients of interest from equation \ref{eq:iden} are $\beta_{4}$, $\beta_{5}$ and $\beta_{6}$. The interaction terms capture the earnings gaps between members of $WH_{i}$, $HW_{i}$, and $WW_{i}$ that identify as Hispanic and White WW children.

\section{Results and Discussion}\label{sec:results}

In this section, I present the results from estimating the two specifications in equations \ref{eq:1a} and \ref{eq:iden}. The results are shown in tables \ref{tab:lastnamereg} and \ref{tab:identreg}. All estimations are done on a sample of White, native-born men between the ages of 25 and 40 employed for full-time full-year (FTFY) and are waged and salaried workers.

I find that the earnings gaps between WH and WW are 8.8\%, 14.1\% between HW and WW, and 16.1\% between HH and WW. Moreover, the difference in earnings between HW and WH children, which captures the effect of having a Hispanic last name, equals a statistically significant 5.3 percentage points. Meaning by comparing inter-ethnic children, a person with a Hispanic last name earns 5.3 percentage points less than someone with a White last name. However, more than half of the earnings gaps could be explained by educational differences between WW children and WH, HW, and HH. When I control for education, the gap between the three groups and WW is cut between 50-64.0\%. Furthermore, the last name effect decreases to a statistically insignificant 1.6 percentage points earnings gap.

I also find a significant earnings gap between those that identify as Hispanic and WW children. The gaps between WH, HW, and HH that identify as Hispanic and WW are, respectively, 13.4\%, 16.4\%, and 10.6\%. The difference between WH and HW that identify as Hispanic---captures the last name effect of those that identify as Hispanic--- is equal to 3 percentage points and is statistically insignificant. Controlling for education explains a big part of the gap. The gaps between WH, HW, and HH that identify as Hispanic and WW after controlling for education are 6.4\%, 10.8\%, and 6.2\%. The WH--WW and HH--WW gaps for those identifying as Hispanic become statistically insignificant after controlling for education. Finally, the difference between WH and HW that identify as Hispanic that identifies as Hispanic remains statistically insignificant but increases to 4.3 percentage points.

\subsection{The effect of having a Hispanic last name on labor market outcomes}

I provide the results to the estimation of equation \ref{eq:1a} in table \ref{tab:lastnamereg}. The sample in this analysis includes full-time, full-year, and waged and salaried men, and I control for hours worked and age and include year fixed effects (FE). Column one in table \ref{tab:lastnamereg} is the results without controlling for education. Overall, there is big Hispanic--White earnings gaps. Children of White father-Hispanic mothers (WH) earn 8.8\% less than children of White fathers-White mothers (WW) families. The gap between children of Hispanic father-White mothers (HW) and WW children is 14.1\%. The gap between children of Hispanic father-Hispanic mothers (HH) and WW children is 16.1\%. Therefore, Hispanic children, inter-ethnic or children of homogeneous HH marriages, face a substantial earnings gap.

The difference between the estimates of WH and HW would capture the effect of having a Hispanic last name. A WH child will have a White last name since their father is White, while an HW child will have a Hispanic last name like their Hispanic father. When I take the difference between HW and WH, I find a statistically significant gap of 5.3 percentage points, which means that an inter-ethnic person with a Hispanic last name earns 5.3 percentage points less than a similar inter-ethnic person but with a White last name.

The literature on the earnings gaps shows that educational differences could explain a big part of these gaps \citep{duncan2006hispanics, duncan2017complexity, duncan2018identifying, duncan2020new}. Therefore, I run the equation \ref{eq:1a}, but this time I control for education. The regression results are in column two of table \ref{tab:lastnamereg}. The sample includes full-time, full-year, and waged and salaried men, and I control for hours worked and age and include year fixed effects (FE). The gap between WH and WW is 4.4\%, a reduction of 50\% from the baseline after controlling for education. The earning gaps between HW and HH, and WW were respectively reduced by 58\% and 64\%. After controlling for education, an HW man earns 5.9\% less than a WW man, and an HH man earns 5.8\% less. The effect of having a Hispanic last name on earnings decreased to 1.6 percentage points and became statistically insignificant after controlling for education.

Therefore, Hispanics face big earnings gaps when compared to Whites. By comparing the children of intermarriages, I could study the impact of having a Hispanic last name on earnings. Having a Hispanic last name carries a cost in the labor market. A person with a Hispanic last name earns 5.3 percentage points less than someone with a White name. However, educational differences between Hispanics and Whites could explain more than 50\% of the earning gaps and all of the Hispanic last name costs.

\subsection{The effect of identifying as Hispanic on labor market outcomes}

In this section, I present the results of equation \ref{eq:iden} that estimates the effect of identifying as Hispanic on annual log earnings. I show the results in table \ref{tab:identreg}. The sample is restricted to full-time full-year (FTFY) salaried and waged men. I control for age and hours worked and include year fixed effects.

I run the estimating equation without controlling for education in table \ref{tab:identreg} column 1. I find that those that identify as Hispanic face a big earnings gap compared to those that do not. A WH man that identifies as Hispanic earns 13.4\% less than a WW man. An HW man who identifies as Hispanic makes 16.4\% less than a WW man, and an HH man who is determined as Hispanic earns 10.6\% less than a WW man. Comparing the coefficients of the interaction terms to those of the $WH_{i}$, $HW_{i}$, and $HH_{i}$ shows that members of the three groups that do not identify as Hispanic earn about the same as WW.
Moreover, the difference between $HW_{i} \cdot Hispanic_{i}$  and $WH_{i} \cdot Hispanic_{i}$ captures the effect of having a Hispanic last name and identifying as Hispanic compared to having a White last name and identifying as Hispanic. A person with a Hispanic last name that identifies as Hispanic earns 3 percentage points less than a person that identifies as Hispanic but has a White last name. The difference, however, is statistically insignificant.

In table  \ref{tab:identreg} column 2, I show the results of equation \ref{eq:iden} while controlling for education. Education explains all the identity costs for members of WH and HH. Also, controlling for education decreases the earnings gap for identifying as Hispanic and being a member of HW by 34.1\%. An HW member that identifies as Hispanic earns 10.8\% less than a WW man. Furthermore, the difference between $HW_{i} \cdot Hispanic_{i}$  and $WH_{i} \cdot Hispanic_{i}$ increases to 4.3 percentage points but remains statistically insignificant.
\subsection{Discussion}

 The results presented in tables \ref{tab:lastnamereg} and \ref{tab:identreg} have an implication that affects how Hispanic-White earnings gaps should be calculated. First, A big portion of the earnings gaps could be explained by education. However, more unobservable characteristics, like innate ability, could affect earnings outcomes. Therefore, comparing Hispanics to Whites could yield inflated gaps since the two groups are different. The reduced form regression of this approach could be written as:

 \begin{equation} \label{eq:naive}
 Y_{it} = \beta_{1} Hispanic_{i} + X_{it}^{\prime} + \gamma_{t}+\phi_{it}
 \end{equation}

 Where $Y_{it}$ in equation \ref{eq:naive} is the annual log earnings of person $i$ at time $t$. $Hispanic_{i}$ is a dummy variable equal to one if person $i$ is Hispanic and zero otherwise. $X_{it}$ is a vector of controls, $\gamma_{t}$ is time fixed effects and $\phi_{it}$ is the error term.

 From equation \ref{eq:naive}, $\beta_{1}$ estimates the earnings gap between Hispanics and Whites. When calculating the equation, before controlling for education, I find an earnings gap that is equal to 15.6\%. This gap is close to the gap between HH and WW. After controlling for education, the gap decreases by more than half and becomes 6\%, which is also similar to the gap between HH and WW after controlling for education. Since differences between Whites and Hispanics, on the observable and unobservables are large, estimating the gap HW-WH could provide us with a bound on how much of the discrimination against Hispanics is taste-based.

The point estimate of the difference between HW and WH from equation \ref{eq:1a} is equal to 1.3 percentage points. The upper bound of the confidence interval is equivalent to 4.5 percentage points. Therefore, I can put an upper bound on the portion of the Hispanic-White earnings gap that could be explained by taste based discrimination. In this case, the upper bound is equal to 4.5 percentage points.

\section{Conclusion}\label{sec:con}

As the Hispanic population keeps growing in the United States, studying discrimination against this group is also important. In this paper, I investigate discrimination against Hispanics in the labor market. More specifically, I explore the effect of having a Hispanic last name and identifying as Hispanic on annual log earnings.

I compare the children of inter-ethnic marriages to study the labor market impact of having a Hispanic last name. The earning gap between WH and WW is 8.8\%, 14.1\% between HW and WW, and 16.1\% between HH and WW. When I compare the earnings of HW and WH children, which captures the effect of having a Hispanic last name, HW children earn 5.3\% less than WH children. Thus, by comparing inter-ethnic children, a person with a Hispanic last name makes 5.3\% less than someone with a White last name. Controlling for education, however, explains more than half of the earnings gaps between WW children and WH, HW, and HH. When I control for education, the gap between the three groups and WW is cut between 50-64.0\%. Furthermore, the last name effect decreases to a statistically insignificant 1.6\% earnings gap.

I also find that men who identify as Hispanic earn significantly less than those who do not, even when I control ancestry and education. The gaps between WH, HW, and HH that identify as Hispanic and WW are, respectively, 13.4\%, 16.4\%, and 10.6\%. Controlling for education explains a big part of the gap. The gaps between WH, HW, and HH that identify as Hispanic and WW after controlling for education are 6.4\%, 10.8\%, and 6.2\%.

% \nocite{*}
\pagebreak
\bibliography{HispanicDiscriminationMarriageMarket}
\newpage
\pagebreak
\begin{appendices}

\section{Figures}\label{appendix:figures}

\begin{center}
\begin{figure}[H]
\caption{proportion of non-Whites in the United States from 1995 to 2019}
\includegraphics[width=\textwidth]{GraphNonWhites.png} 
\label{fig:1}
\end{figure}
\end{center}

\newpage

\begin{center}
\begin{figure}
\caption{Hispanics as a percentage of the total population in \\
the United States from 1995 to 2019.}
\includegraphics[width=\textwidth]{HispanicUSA.png} \label{fig:2}
\end{figure}
\end{center}

\newpage

\begin{center}
\begin{figure}
\caption{Inter-racial marriages as a percent of all marriages \\
 in the United States from 1995 to 2020.}
\includegraphics[width=\textwidth]{interracialovertime.png} 
\label{fig:3}
\end{figure}
\end{center}

\newpage

\begin{center}
\begin{figure}
\caption{Distribution of the four types of children.}
\includegraphics[width=\textwidth]{PirChart2.png} 
\label{fig:dist}
\end{figure}
\end{center}

\newpage
\begin{center}
\begin{figure}
\caption{Chart explaining which synthetic parents and children \\
and when we observe them.}
\begin{tikzpicture}[node distance =2cm]
\node (start) [startstop] {Synthetic Parents: Observed in 1970 to 1990. $25 \leq age \leq 40$};

\node (pro1) [process, below of = start, yshift = -2cm] {Children: Observed between 1994 and 2019. Born between 1970 to 1990 and $25 \leq age \leq 40$};

\draw [arrow] (start) -- (pro1);
\end{tikzpicture}
\label{flowchart1}
\end{figure}
\end{center}

\newpage

\section{tables}\label{appendix:tables}

\input{tables/tab01-observations-by-parents}

% \begin{table}[H]
%   \centering
%   \renewcommand*\TPTnoteLabel[1]{\parbox[b]{3em}{\hfill#1\,}}
%   \begin{adjustbox}{max width=\textwidth}
%   \begin{threeparttable}
% \caption{\textsc{Number of children from the different types of parents.}}\label{tab:mat1}
% \footnotesize
% \begin{tabular}{cccccccccccccc}
% \hline 
%  &  &  & \multicolumn{10}{c}{Perent's Type} & \tabularnewline
% \cline{4-13} 
%  &  &  & White Father &  &  & White Father &  &  & Hispanic father &  &  & Hispanic father & \tabularnewline
%  &  &  & White Mother &  &  & Hispanic Mother &  &  & White Mother &  &  & Hispanic Mother & \tabularnewline
% \hline 
%  &  &  &  &  &  &  &  &  &  &  &  &  & \tabularnewline
%  & Number &  & 3,828,024 &  &  & 24,306 &  &  & 35,841 &  &  & 135,300 & \tabularnewline
%  & (\%) &  & (95.14\%) &  &  & (0.89\%) &  &  & (0.89\%) &  &  & (3.36\%) & \tabularnewline
%  &  &  &  &  &  &  &  &  &  &  &  &  & \tabularnewline
% \hline 
% \end{tabular}
%  \begin{tablenotes}
%    \item [\textit{Notes.}] The data is restricted to people interviewed between 1994 and 2019, also White, married, and between the ages of 18 and 65. I identify the ethnicity of a person's parents through the parent's place of birth. A parent is Hispanic if both her parents were born in a Spanish-speaking country. A parent is White if born parents were born in the United States.
%   \end{tablenotes}
%   \end{threeparttable}
%   \end{adjustbox}
% \end{table}

\newpage

\input{tables/tab02-observations-by-couple}

% \begin{table}[H]
% \begin{adjustbox}{max width=\textwidth}
%   \centering
%   \renewcommand*\TPTnoteLabel[1]{\parbox[b]{3em}{\hfill#1\,}}
%   \begin{threeparttable}
% \caption{\textsc{Parents' types}}\label{tab:mat2}
% \footnotesize
% \begin{tabular}{cccccccccccccc}
% \hline 
%  &  &  & \multicolumn{10}{c}{Couples' Type} & \tabularnewline
% \cline{4-13} 
%  &  &  & White Husband &  &  & White Husband &  &  & Hispanic Husband &  &  & Hispanic Husband & \tabularnewline
%  &  &  & White Wife &  &  & Hispanic Wife &  &  & White Wife &  &  & Hispanic Wife & \tabularnewline
% \hline 
%  &  &  &  &  &  &  &  &  &  &  &  &  & \tabularnewline
%  & Number &  & 1,198,446 &  &  & 3,819 &  &  & 3,413 &  &  & 6,920 & \tabularnewline
%  & (\%) &  & (98.83\%) &  &  & (0.31\%) &  &  & (0.28\%) &  &  & (0.57\%) & \tabularnewline
%  &  &  &  &  &  &  &  &  &  &  &  &  & \tabularnewline
% \hline 
% \end{tabular}

%   \begin{tablenotes}
%     \item [\textit{Notes.}] The data is restricted to people interviewed in 1970 and 1960 and also White and married. I identify the ethnicity of a person through their place of birth. A parent is Hispanic if they were born in a Spanish-speaking country. A parent is White if they were born in the United States.
%     \item [\textit{Notes.}] The table includes information on the proportion of the four types of synthetic parents that I have constructed.
%   \end{tablenotes}
%   \end{threeparttable}
% \end{adjustbox}
% \end{table}

\newpage

\input{tables/tab03-summary-stats}

% \newpage

% \begin{table}[H]
%   \centering
%   \renewcommand*\TPTnoteLabel[1]{\parbox[b]{3em}{\hfill#1\,}}
%   \begin{adjustbox}{max width=\textwidth}
%   \begin{threeparttable}
% \caption{\textsc{Children's outcome using parent's place of birth}}\label{tab:c&p1}
% \footnotesize
% \begin{tabular}{llccccccccr@{\extracolsep{0pt}.}l}
% \toprule 
%  &  & \multicolumn{4}{c}{Father's and Mother's Ethnicities} &  & \multicolumn{5}{c}{Differences}\tabularnewline
% \cmidrule{3-6} \cmidrule{8-12} 
%  &  &  &  &  &  &  &  &  &  & \multicolumn{2}{c}{}\tabularnewline
% Variables &  & White Father & White Father & Hispanic Father & Hispanic Father &  & HH - WW &  & HW-WH & \multicolumn{2}{c}{}\tabularnewline
%  &  & White Mother & Hispanic Mother & White Mother & Hispanic Mother &  &  &  &  & \multicolumn{2}{c}{}\tabularnewline
%  &  & (WW) & (WH) & (HW) & (HH) &  &  &  &  & \multicolumn{2}{c}{}\tabularnewline
%  &  & (i) & (ii) & (iii) & (iv) &  & (v) &  & (vi) & \multicolumn{2}{c}{}\tabularnewline
% \midrule
% \textbf{Parent's Panel:} &  &  &  &  &  &  &  &  &  & \multicolumn{2}{c}{}\tabularnewline
% Husband's education (total years) &  & 12.727 & 11.847 & 10.286 & 8.949 &  & -3.779{*}{*}{*} &  & -1.560{*}{*}{*} & \multicolumn{2}{c}{}\tabularnewline
%  &  & (2.616) & (3.658) & (4.602) & (4.422) &  & (0.018) &  & (0.041) & \multicolumn{2}{c}{}\tabularnewline
% Wife's education (total years) &  & 12.467 & 10.462 & 11.188 & 8.590 &  & -3.876{*}{*}{*} &  & 0.726{*}{*}{*} & \multicolumn{2}{c}{}\tabularnewline
%  &  & (2.182) & (4.139) & (3.394) & (4.120) &  & (0.015) &  & (0.035) & \multicolumn{2}{c}{}\tabularnewline
% Total Household education (years) &  & 25.194 & 22.308 & 21.473 & 17.539 &  & -7.655{*}{*}{*} &  & -0.835{*}{*}{*} & \multicolumn{2}{c}{}\tabularnewline
%  &  & (4.300) & (7.034) & (7.202) & (7.836) &  & (0.029) &  & (0.068) & \multicolumn{2}{c}{}\tabularnewline
% \textbf{Panel A:} &  &  &  &  &  &  &  &  &  & \multicolumn{2}{c}{}\tabularnewline
% \textbf{age and education} &  &  &  &  &  &  &  &  &  & \multicolumn{2}{c}{}\tabularnewline
% man's Education &  & 13.861 & 13.366 & 13.072 & 12.898 &  & -0.962{*}{*}{*} &  & -0.294{*}{*}{*} & \multicolumn{2}{c}{}\tabularnewline
%  &  & (2.421) & (2.317) & (2.270) & (2.356) &  &  &  &  & \multicolumn{2}{c}{}\tabularnewline
% woman's Education &  & 14.200 & 13.741 & 13.270 & 13.204 &  & -0.996{*}{*}{*} &  & -0.471{*}{*}{*} & \multicolumn{2}{c}{}\tabularnewline
%  &  & (2.424) & (2.467) & (2.342) & (2.453) &  &  &  &  & \multicolumn{2}{c}{}\tabularnewline
% \textbf{Panel B: Employment \& Earnings} &  &  &  &  &  &  &  &  &  & \multicolumn{2}{c}{}\tabularnewline
% men' employment rate &  & 0.952 & 0.945 & 0.927 & 0.931 &  & -0.021{*}{*}{*} &  & -0.018{*}{*}{*} & \multicolumn{2}{c}{}\tabularnewline
%  &  & (0.215) & (0.227) & (0.260) & (0.254) &  &  &  &  & \multicolumn{2}{c}{}\tabularnewline
% women' employment rate &  & 0.956 & 0.950 & 0.931 & 0.936 &  & -0.020{*}{*}{*} &  & -0.019{*}{*}{*} & \multicolumn{2}{c}{}\tabularnewline
%  &  & (0.205) & (0.218) & (0.253) & (0.244) &  &  &  &  & \multicolumn{2}{c}{}\tabularnewline
% men' log hourly earnings &  & 2.557 & 2.524 & 2.468 & 2.505 &  & -0.041{*}{*}{*} &  & -0.056{*}{*}{*} & \multicolumn{2}{c}{}\tabularnewline
%  &  & (0.399) & (0.421) & (0.423) & (0.401) &  &  &  &  & \multicolumn{2}{c}{}\tabularnewline
% women' log hourly earnings &  & 2.435 & 2.444 & 2.385 & 2.389 &  & -0.016{*}{*} &  & -0.061{*}{*}{*} & \multicolumn{2}{c}{}\tabularnewline
%  &  & (0.415) & (0.426) & (0.375) & (0.376) &  &  &  &  & \multicolumn{2}{c}{}\tabularnewline
% men' log annual earnings &  & 10.529 & 10.435 & 10.386 & 10.365 &  & -0.164{*}{*}{*} &  & -0.049 & \multicolumn{2}{c}{}\tabularnewline
%  &  & (0.703) & (0.637) & (0.872) & (0.622) &  &  &  &  & \multicolumn{2}{c}{}\tabularnewline
% women' log annual earnings &  & 10.297 & 10.262 & 10.159 & 10.175 &  & -0.123{*}{*}{*} &  & -0.103{*}{*}{*} & \multicolumn{2}{c}{}\tabularnewline
%  &  & (0.633) & (0.613) & (0.541) & (0.650) &  &  &  &  & \multicolumn{2}{c}{}\tabularnewline
% \textbf{Panel D: Identity} &  &  &  &  &  &  &  &  &  & \multicolumn{2}{c}{}\tabularnewline
% Proportion of men &  & 0.060 & 0.821 & 0.879 & 0.969 &  &  &  &  & \multicolumn{2}{c}{}\tabularnewline
% identifying as Hispanic &  &  &  &  &  &  &  &  &  & \multicolumn{2}{c}{}\tabularnewline
% Proportion of women &  & 0.059 & 0.850 & 0.859 & 0.970 &  &  &  &  & \multicolumn{2}{c}{}\tabularnewline
% identifying as Hispanic &  &  &  &  &  &  &  &  &  & \multicolumn{2}{c}{}\tabularnewline
%  &  &  &  &  &  &  &  &  &  & \multicolumn{2}{c}{}\tabularnewline
% \bottomrule
% \end{tabular}
%   \begin{tablenotes}
%    \item [\textit{Notes.}] The data is restricted to native-born United States citizens between 1994 and 2019 who are also White and between the ages of 25 and 40. I identify the ethnicity of a person's parents through the parent's place of birth. A parent is Hispanic if they were born in a Spanish-speaking country. A parent is White if they were born in the United States.
%    \item [\textit{Notes.}] In each column, I present the average statistics of the different types of people based on the ethnicities of their parents. In column one, I show the summary statistics of children of White fathers and White mothers. In column two, I present the summary statistics of children of White fathers and Hispanic mothers. In column three, I show the summary statistics of children of Hispanic fathers and White mothers. In column four, I present the summary statistics of children of Hispanic fathers and mothers.
%    \item [\textit{Notes.}] Columns five and six have data on the HH--WW gaps (column five) and the HW--WH gaps (column six).

%   \end{tablenotes}
%   \end{threeparttable}
%   \end{adjustbox}
% \end{table}

\newpage

\input{tables/tab04-summary-stats}

% \begin{table}[H]
%   \centering
%   \renewcommand*\TPTnoteLabel[1]{\parbox[b]{3em}{\hfill#1\,}}
%   \begin{adjustbox}{max width=\textwidth}
%   \begin{threeparttable}
% \caption{\textsc{Children's outcomes using parent's place of birth only for those that identify as Hispanic}}\label{tab:c&p2}
% \footnotesize
% \begin{tabular}{llccccccccr@{\extracolsep{0pt}.}l}
% \hline 
%  &  & \multicolumn{4}{c}{Father's and Mother's Ethnicities} &  & \multicolumn{5}{c}{Differences}\tabularnewline
% \cline{3-6} \cline{8-12} 
%  &  &  &  &  &  &  &  &  &  & \multicolumn{2}{c}{}\tabularnewline
% Variables &  & White Father & White Father & Hispanic Father & Hispanic Father &  & HH - WW &  & HW-WH & \multicolumn{2}{c}{}\tabularnewline
%  &  & White Mother & Hispanic Mother & White Mother & Hispanic Mother &  &  &  &  & \multicolumn{2}{c}{}\tabularnewline
%  &  & (WW) & (WH) & (HW) & (HH) &  &  &  &  & \multicolumn{2}{c}{}\tabularnewline
%  &  & (i) & (ii) & (iii) & (iv) &  & (v) &  & (vi) & \multicolumn{2}{c}{}\tabularnewline
% \hline 
% \textbf{Parent's Panel:} &  &  &  &  &  &  &  &  &  & \multicolumn{2}{c}{}\tabularnewline
% Husband's education (total years) &  & 11.856 & 10.211 & 9.665 & 8.427 &  & -3.429{*}{*}{*} &  & -0.546{*}{*}{*} & \multicolumn{2}{c}{}\tabularnewline
%  &  & (2.922) & (3.876) & (4.095) & (4.228) &  & (0.035) &  & (0.079) & \multicolumn{2}{c}{}\tabularnewline
% Wife's education (total years) &  & 11.793 & 9.556 & 9.791 & 8.137 &  & -3.656{*}{*}{*} &  & 0.235{*}{*}{*} & \multicolumn{2}{c}{}\tabularnewline
%  &  & (2.329) & (3.751) & (3.524) & (3.899) &  & (0.028) &  & (0.063) & \multicolumn{2}{c}{}\tabularnewline
% Total Household education (years) &  & 23.650 & 19.767 & 19.456 & 16.564 &  & -7.085{*}{*}{*} &  & -0.311{*}{*}{*} & \multicolumn{2}{c}{}\tabularnewline
%  &  & (4.716) & (6.843) & (6.896) & (7.338) &  & (0.057) &  & (0.128) & \multicolumn{2}{c}{}\tabularnewline
% \textbf{Panel A:} &  &  &  &  &  &  &  &  &  & \multicolumn{2}{c}{}\tabularnewline
% \textbf{age and education} &  &  &  &  &  &  &  &  &  & \multicolumn{2}{c}{}\tabularnewline
% man's Education &  & 13.861 & 13.237 & 13.008 & 12.874 &  & -0.986{*}{*}{*} &  & -0.230{*}{*}{*} & \multicolumn{2}{c}{}\tabularnewline
%  &  & (2.421) & (2.313) & (2.217) & (2.347) &  &  &  &  & \multicolumn{2}{c}{}\tabularnewline
% woman's Education &  & 14.200 & 13.603 & 13.133 & 13.182 &  & -1.018{*}{*}{*} &  & -0.470{*}{*}{*} & \multicolumn{2}{c}{}\tabularnewline
%  &  & (2.424) & (2.391) & (2.325) & (2.453) &  &  &  &  & \multicolumn{2}{c}{}\tabularnewline
% \textbf{Panel D: Employment \& Earnings} &  &  &  &  &  &  &  &  &  & \multicolumn{2}{c}{}\tabularnewline
% men' employment rate &  & 0.952 & 0.943 & 0.925 & 0.930 &  & -0.021{*}{*}{*} &  & -0.018{*}{*}{*} & \multicolumn{2}{c}{}\tabularnewline
%  &  & (0.215) & (0.232) & (0.264) & (0.254) &  &  &  &  & \multicolumn{2}{c}{}\tabularnewline
% women' employment rate &  & 0.956 & 0.948 & 0.928 & 0.936 &  & -0.020{*}{*}{*} &  & -0.020{*}{*}{*} & \multicolumn{2}{c}{}\tabularnewline
%  &  & (0.205) & (0.222) & (0.258) & (0.245) &  &  &  &  & \multicolumn{2}{c}{}\tabularnewline
% men' log hourly earnings &  & 2.501 & 2.470 & 2.408 & 2.460 &  & -0.041{*}{*}{*} &  & -0.062{*}{*}{*} & \multicolumn{2}{c}{}\tabularnewline
%  &  & (0.454) & (0.458) & (0.463) & (0.434) &  &  &  &  & \multicolumn{2}{c}{}\tabularnewline
% women' log hourly earnings &  & 2.339 & 2.358 & 2.288 & 2.322 &  & -0.017{*}{*} &  & -0.071{*}{*}{*} & \multicolumn{2}{c}{}\tabularnewline
%  &  & (0.494) & (0.499) & (0.464) & (0.423) &  &  &  &  & \multicolumn{2}{c}{}\tabularnewline
% men' log annual earnings &  & 10.529 & 10.414 & 10.394 & 10.364 &  & -0.165{*}{*}{*} &  & -0.020 & \multicolumn{2}{c}{}\tabularnewline
%  &  & (0.703) & (0.610) & (0.704) & (0.625) &  &  &  &  & \multicolumn{2}{c}{}\tabularnewline
% women' log annual earnings &  & 10.297 & 10.240 & 10.136 & 10.172 &  & -0.126{*}{*}{*} &  & -0.104{*}{*}{*} & \multicolumn{2}{c}{}\tabularnewline
%  &  & (0.634) & (0.635) & (0.537) & (0.652) &  &  &  &  & \multicolumn{2}{c}{}\tabularnewline
%  &  &  &  &  &  &  &  &  &  & \multicolumn{2}{c}{}\tabularnewline
% \hline 
% \end{tabular}
%   \begin{tablenotes}
%    \item [\textit{Notes.}] The data is restricted to native-born United States citizens between 1994 and 2019 who are also White and between the ages of 25 and 40. I identify the ethnicity of a person's parents through the parent's place of birth. A parent is Hispanic if they were born in a Spanish-speaking country. A parent is White if they were born in the United States.
%    \item [\textit{Notes.}] In each column, I present the average statistics of those that identify as Hispanic by the different types of people based on the ethnicities of their parents. In column one, I present the summary statistics of children of White fathers and White mothers. In column two, I show the summary statistics of children of White fathers and Hispanic mothers identifying as Hispanic. In column three, I present the summary statistics of children of Hispanic fathers and White mothers that identify as Hispanic. In column four, I show the summary statistics of children of Hispanic fathers and Hispanic mothers that identify as Hispanic.
%    \item [\textit{Notes.}] Columns five and six have data on the HH--WW gaps (column five) and the HW--WH gaps (column six).

%   \end{tablenotes}
%   \end{threeparttable}
%   \end{adjustbox}
% \end{table}

\newpage

\input{tables/tab05-regression}

% \begin{table}[H]
%   \centering
%   \renewcommand*\TPTnoteLabel[1]{\parbox[b]{3em}{\hfill#1\,}}
%     \begin{adjustbox}{max width=\textwidth}
%   \begin{threeparttable}
% \caption{\textsc{Effect of having a Hispanic last name}}\label{tab:lastnamereg}
% \footnotesize

% \begin{tabular}{llcccc}
% \toprule 
%  &  &  &  &  & \tabularnewline
% Variables &  & Log annual earnings &  & Log annual earnings & \tabularnewline
% \midrule

% $WH_{i}$ &  & -0.088{*}{*}{*} &  & -0.044{*}{*}{*} & \tabularnewline
%  &  & (0.018) &  & (0.017) & \tabularnewline
%  &  &  &  &  & \tabularnewline
% $HW_{i}$ &  & -0.141{*}{*}{*} &  & -0.059{*}{*}{*} & \tabularnewline
%  &  & (0.015) &  & (0.014) & \tabularnewline
%  &  &  &  &  & \tabularnewline
% $HH_{i}$ &  & -0.161{*}{*}{*} &  & -0.058{*}{*}{*} & \tabularnewline
%  &  & (0.008) &  & (0.007) & \tabularnewline
%  &  &  &  &  & \tabularnewline
% \midrule
% $HW_{i}$ - $WH_{i}$ &  & -0.053{*}{*} &  & -0.016 & \tabularnewline
%  &  & (0.024) &  & (0.022) & \tabularnewline
%  &  &  &  &  & \tabularnewline
% \midrule
% \textit{Controlling for:} &  &  &  &  & \tabularnewline
% Education &  & No &  & Yes & \tabularnewline
% Hours worked &  & Yes &  & Yes & \tabularnewline
% Age &  & Yes &  & Yes & \tabularnewline
% Year FE &  & Yes &  & Yes & \tabularnewline
% \bottomrule
% \end{tabular}

%   \begin{tablenotes}
%    \item [\textit{Notes.}] This table includes the estimation results of equation \ref{eq:1a}.
%    \item [\textit{Notes.}] The four groups stand for White Husband White Wife (WW), White Husband Hispanic Wife (WH), Hispanic Husband White (HW), and Hispanic Husband Hispanic Wife (HH).
%    \item [\textit{Notes.}] The sample is restricted to men working full-time full-year and are waged and salaried workers.
%     \item [\textit{Notes.}] Column one has the regression results when controlling for hours worked, age, and fixed effects. Column two has the results after controlling for education.
%   \end{tablenotes}
%   \end{threeparttable}
%   \end{adjustbox}
% \end{table}

\newpage

\input{tables/tab06-regression}

% \begin{table}[H]
%   \centering
%   \renewcommand*\TPTnoteLabel[1]{\parbox[b]{3em}{\hfill#1\,}}
%     \begin{adjustbox}{max width=\textwidth}
%   \begin{threeparttable}
% \caption{\textsc{Last name effect}}\label{tab:identreg}
% \footnotesize

% \begin{tabular}{lccccc}
% \hline 
%  &  & Log annual earnings &  & Log annual earnings & \tabularnewline
% Variables &  &  &  &  & \tabularnewline
% \hline 
%  &  &  &  &  & \tabularnewline
% $WH_{i}\times Hispanic_{i}$ &  & -0.134{*}{*} &  & -0.064 & \tabularnewline
%  &  & (0.053) &  & (0.049) & \tabularnewline
% $HW_{i}\times Hispanic_{i}$ &  & -0.164{*}{*}{*} &  & -0.108{*}{*} & \tabularnewline
%  &  & (0.052) &  & (0.048) & \tabularnewline
% $HH_{i}\times Hispanic_{i}$ &  & -0.106{*}{*} &  & -0.062 & \tabularnewline
%  &  & (0.049) &  & (0.045) & \tabularnewline
% $WH_{i}$ &  & 0.027 &  & 0.011 & \tabularnewline
%  &  & (0.048) &  & (0.045) & \tabularnewline
% $HW_{i}$ &  & 0.007 &  & 0.038 & \tabularnewline
%  &  & (0.050) &  & (0.046) & \tabularnewline
% $HH_{i}$ &  & -0.057 &  & 0.003 & \tabularnewline
%  &  & (0.057) &  & (0.044) & \tabularnewline
% \hline 
% $HW_{i}\times Hispanic_{i}$ - $WH_{i}\times Hispanic_{i}$ &  & -0.030 &  & -0.043 & \tabularnewline
%  &  & (0.074) &  & (0.068) & \tabularnewline
% \hline 
% \textit{Controlling for:} &  &  &  &  & \tabularnewline
% Education &  & No &  & Yes & \tabularnewline
% Hours worked &  & Yes &  & Yes & \tabularnewline
% Age &  & Yes &  & Yes & \tabularnewline
% Year FE &  & Yes &  & Yes & \tabularnewline
% \hline 
% \end{tabular}

%   \begin{tablenotes}
%    \item [\textit{Notes.}] This table includes the estimation results of equation \ref{eq:iden}.
%    \item [\textit{Notes.}] The four groups stand for White Husband White Wife (WW), White Husband Hispanic Wife (WH), Hispanic Husband White (HW), and Hispanic Husband Hispanic Wife (HH). $Hispanic_{i}$ is a dummy variable equal to one if a person identifies as Hispanic and zero otherwise.
%    \item [\textit{Notes.}] The sample is restricted to men working full-time full-year and are waged and salaried workers.
%     \item [\textit{Notes.}] Column one has the regression results when controlling for hours worked, age, and years of fixed effects. Column two has the results after controlling for education.
%   \end{tablenotes}
%   \end{threeparttable}
%   \end{adjustbox}
% \end{table}

\newpage

\input{tables/tab07-regression}

% \begin{table}[H]
%   \centering
%   \renewcommand*\TPTnoteLabel[1]{\parbox[b]{3em}{\hfill#1\,}}
%     \begin{adjustbox}{max width=\textwidth}
%   \begin{threeparttable}
% \caption{\textsc{Hsiapnic--White earnings gap using Hispanic variable only}}\label{tab:hispwhitegap}
% \footnotesize

% \begin{tabular}{lccccc}
% \hline 
%  &  &  &  &  & \tabularnewline
% Variables &  & Log annual earnings &  & Log annual earnings & \tabularnewline
% \hline 
% $Hispanic_{i}$ &  & -0.156{*}{*}{*} &  & -0.060{*}{*}{*} & \tabularnewline
%  &  & (0.006) &  & (0.006) & \tabularnewline
% \textit{Controlling for:} &  &  &  &  & \tabularnewline
% Education &  & No &  & Yes & \tabularnewline
% Hours worked &  & Yes &  & Yes & \tabularnewline
% Age &  & Yes &  & Yes & \tabularnewline
% Year FE &  & Yes &  & Yes & \tabularnewline
% \hline 
% \end{tabular}

%   \begin{tablenotes}
%    \item [\textit{Notes.}] This table includes the estimation results of equation \ref{eq:naive}.
%    \item [\textit{Notes.}] $Hispanic_{i}$ is a dummy variable equal to one if a person identifies as Hispanic and zero otherwise.
%    \item [\textit{Notes.}] The sample is restricted to men working full time full-year and are waged and salaried workers.
%     \item [\textit{Notes.}] Column one has the regression results when controlling for hours worked, age, and fixed effects. Column two has the results after controlling for education.
%   \end{tablenotes}
%   \end{threeparttable}
%   \end{adjustbox}
% \end{table}

\end{appendices}

%\begin{table}[!htbp]
%\caption{T-tests across  groups by several variables} \label{tab:table8}
%\resizebox{0.95\textwidth}{!}{
%	  \estauto{sumstat_parents2.tex}{8}{c}
%	  }
%\small{Note:  *** p$<$0.01, ** p$<$0.05, * p$<$0.10.}  
%\end{table}

\end{document}

### 2023-01-20 my_paper/tables/tab01-observations-by-parents.tex
push project to github

\begin{table}[!h]

\caption{Number of Children from the Different Types of Parents \label{tab:mat1}}
\centering
\resizebox{\linewidth}{!}{
\begin{threeparttable}
\begin{tabular}[t]{>{}lcccc}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{4}{c}{Perent's Type} \\
\cmidrule(l{3pt}r{3pt}){2-5}
Parents' Type & \specialcell{White Father \\ White Mother} & \specialcell{White Father \\ Hispanic Mother} & \specialcell{Hispanic Father \\ White Mother} & \specialcell{Hispanic Father \\ Hispanic Mother}\\
\midrule
\textbf{Observations} & \specialcell{3,828,024\\(0.95)} & \specialcell{24,306\\(0.01)} & \specialcell{35,841\\(0.01)} & \specialcell{135,300\\(0.03)}\\
\bottomrule
\end{tabular}
\begin{tablenotes}
\item[1] The data is restricted to people interviewed between 1994 and 2019, also White, married, and between the ages of 18 and 65. I identify the ethnicity of a person's parents through the parent's place of birth. A parent is Hispanic if both her parents were born in a Spanish-speaking country. A parent is White if born parents were born in the United States.
\end{tablenotes}
\end{threeparttable}}
\end{table}

### 2023-01-20 my_paper/tables/tab02-observations-by-couple.tex
push project to github

\begin{table}[!h]

\caption{Couples' Type \label{tab:mat2}}
\centering
\resizebox{\linewidth}{!}{
\begin{threeparttable}
\begin{tabular}[t]{>{}lcccc}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{4}{c}{Couples' Type} \\
\cmidrule(l{3pt}r{3pt}){2-5}
  & \specialcell{White Husband \\ White Wife} & \specialcell{White Husband \\ Hispanic Wife} & \specialcell{Hispanic Husband \\ White Wife} & \specialcell{Hispanic Husband \\ Hispanic Wife}\\
\midrule
\textbf{Observations} & \specialcell{1,286,731\\(0.97)} & \specialcell{7,178\\(0.01)} & \specialcell{7,606\\(0.01)} & \specialcell{20,911\\(0.02)}\\
\bottomrule
\end{tabular}
\begin{tablenotes}
\item[1] The data is restricted to people interviewed in 1970 and 1960 and also White and married. I identify the ethnicity of a person through their place of birth. A parent is Hispanic if they were born in a Spanish-speaking country. A parent is White if they were born in the United States.
\item[2] The table includes information on the proportion of the four types of synthetic parents that I have constructed.
\end{tablenotes}
\end{threeparttable}}
\end{table}

### 2023-01-20 my_paper/tables/tab03-summary-stats.tex
push project to github

\begin{table}[!h]

\caption{Children's outcome using parent's place of birth \label{tab:c&p1}}
\centering
\resizebox{\linewidth}{!}{
\begin{threeparttable}
\begin{tabular}[t]{lcccccc}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{4}{c}{Father's and Mother's Ethnicities} & \multicolumn{2}{c}{Differences} \\
\cmidrule(l{3pt}r{3pt}){2-5} \cmidrule(l{3pt}r{3pt}){6-7}
Variables & \specialcell{White Father \\ White Mother \\ (WW) \\ (i)} & \specialcell{White Father \\ Hispanic Mother \\ (WH) \\ (ii)} & \specialcell{Hispanic Father \\ White Mother \\ (HW) \\ (iii)} & \specialcell{Hispanic Father \\ Hispanic Mother \\ (HH) \\ (iv)} & \specialcell{HH - WW \\ (v)} & \specialcell{HW - WH \\ (vi)}\\
\midrule
\textbf{Panel A: Parent’s} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
\hspace{1em}Husband’seducation (Total Years) & \specialcell{13.05\\(2.44)} & \specialcell{12.32\\(3.33)} & \specialcell{10.65\\(4.39)} & \specialcell{8.93\\(4.41)} & \specialcell{-4.11\\(0.02)} & \specialcell{-1.67\\(0.04)}\\
\hspace{1em}Wife’seducation (Total Years) & \specialcell{12.74\\(2.12)} & \specialcell{11.03\\(3.92)} & \specialcell{11.54\\(3.12)} & \specialcell{8.6\\(4.13)} & \specialcell{-4.13\\(0.02)} & \specialcell{0.51\\(0.04)}\\
\hspace{1em}Total Household seducation (Total Years) & \specialcell{25.78\\(4.08)} & \specialcell{23.35\\(6.51)} & \specialcell{22.19\\(6.69)} & \specialcell{17.54\\(7.83)} & \specialcell{-4.13\\(0.02)} & \specialcell{0.51\\(0.04)}\\
\textbf{Panel B: Education} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
\addlinespace
\hspace{1em}Men’s education (Total Years) & \specialcell{13.91\\(2.39)} & \specialcell{13.58\\(2.35)} & \specialcell{13.21\\(2.32)} & \specialcell{12.91\\(2.26)} & \specialcell{-1.00\\(0.01)} & \specialcell{-0.36\\(0.03)}\\
\hspace{1em}Women’s education (Total Years) & \specialcell{14.29\\(2.41)} & \specialcell{13.87\\(2.47)} & \specialcell{13.42\\(2.35)} & \specialcell{13.27\\(2.37)} & \specialcell{-1.01\\(0.01)} & \specialcell{-0.46\\(0.03)}\\
\textbf{Panel C: Employment and Earnings} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
\hspace{1em}Men’s Employment Rate & \specialcell{0.95\\(0.22)} & \specialcell{0.94\\(0.23)} & \specialcell{0.92\\(0.26)} & \specialcell{0.93\\(0.26)} & \specialcell{-0.02\\(0)} & \specialcell{-0.02\\(0)}\\
\hspace{1em}Women’s Employment Rate & \specialcell{0.96\\(0.2)} & \specialcell{0.95\\(0.22)} & \specialcell{0.93\\(0.25)} & \specialcell{0.94\\(0.24)} & \specialcell{-0.02\\(0)} & \specialcell{-0.02\\(0)}\\
\addlinespace
\hspace{1em}Men’s Log Hourly Earnings & \specialcell{2.48\\(0.45)} & \specialcell{2.41\\(0.45)} & \specialcell{2.4\\(0.44)} & \specialcell{2.41\\(0.43)} & \specialcell{-0.07\\(0.01)} & \specialcell{-0.00\\(0.02)}\\
\hspace{1em}Women’s Hourly Earnings & \specialcell{2.33\\(0.49)} & \specialcell{2.33\\(0.46)} & \specialcell{2.27\\(0.45)} & \specialcell{2.3\\(0.41)} & \specialcell{-0.02\\(0.01)} & \specialcell{-0.06\\(0.02)}\\
\hspace{1em}Men’s Log Annual Earnings & \specialcell{10.25\\(1.01)} & \specialcell{10.08\\(1.05)} & \specialcell{10.04\\(1.06)} & \specialcell{10.01\\(1.04)} & \specialcell{-0.25\\(0.01)} & \specialcell{-0.04\\(0.03)}\\
\hspace{1em}Women’s Hourly Earnings & \specialcell{9.46\\(1.78)} & \specialcell{9.54\\(1.64)} & \specialcell{9.47\\(1.57)} & \specialcell{9.53\\(1.52)} & \specialcell{0.07\\(0.02)} & \specialcell{-0.07\\(0.06)}\\
\textbf{Panel D: Hispanic Identity} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
\addlinespace
\hspace{1em}Men & \specialcell{0.05} & \specialcell{0.81} & \specialcell{0.88} & \specialcell{0.97} &  & \\
\hspace{1em}Women & \specialcell{0.06} & \specialcell{0.85} & \specialcell{0.87} & \specialcell{0.97} &  & \\
\bottomrule
\end{tabular}
\begin{tablenotes}
\item[1] The data is restricted to native-born United States citizens between 1994 and 2019 who are also White and between the ages of 25 and 40. I identify the ethnicity of a person's parents through the parent's place of birth. A parent is Hispanic if they were born in a Spanish-speaking country. A parent is White if they were born in the United States.
\item[2] In each column, I present the average statistics of the different types of people based on the ethnicities of their parents. In column one, I show the summary statistics of children of White fathers and White mothers. In column two, I present the summary statistics of children of White fathers and Hispanic mothers. In column three, I show the summary statistics of children of Hispanic fathers and White mothers. In column four, I present the summary statistics of children of Hispanic fathers and mothers.
\item[3] Columns five and six have data on the HH--WW gaps (column five) and the HW--WH gaps (column six).
\end{tablenotes}
\end{threeparttable}}
\end{table}

### 2023-01-20 my_paper/tables/tab06-regression.tex
push project to github

\begin{table}[!h]

\caption{Effect of Having Hispanic Last Name \label{tab:identreg}}
\centering
\resizebox{\linewidth}{!}{
\begin{threeparttable}
\begin{tabular}[t]{lcc}
\toprule
  & \specialcell{(1) \\ Log annual earnings} & \specialcell{(2) \\  Log annual earnings}\\
\midrule
$WH_{i} \times Hispanic_{i}$ & \num{-0.13}** & \num{-0.06}\\
 & (\num{0.05}) & \vphantom{2} (\num{0.05})\\
$HW_{i} \times Hispanic_{i}$ & \num{-0.16}*** & \num{-0.11}**\\
 & (\num{0.05}) & \vphantom{1} (\num{0.05})\\
$HH_{i} \times Hispanic_{i}$ & \num{-0.11}** & \num{-0.06}\\
 & (\num{0.05}) & \vphantom{2} (\num{0.04})\\
$WH_{i}$ & \num{0.03} & \num{0.01}\\
 & (\num{0.05}) & \vphantom{1} (\num{0.04})\\
$HW_{i}$ & \num{0.01} & \num{0.04}\\
 & (\num{0.05}) & (\num{0.05})\\
$HH_{i}$ & \num{-0.06} & \num{0.00}\\
 & (\num{0.05}) & (\num{0.04})\\
\midrule
$HW_{i} \times Hispanic_{i} - WH_{i} \times Hispanic_{i}$ & -0.03 & -0.04\\
 & (0.07) & (0.07)\\
\midrule
Observations & \num{129359} & \num{129359}\\
\textit{Controlling for:} &  & \\
Hours Worked & X & X\\
Age & X & X\\
Year FE & X & X\\
Education &  & X\\
\bottomrule
\multicolumn{3}{l}{\rule{0pt}{1em}* p $<$ 0.1, ** p $<$ 0.05, *** p $<$ 0.01}\\
\end{tabular}
\begin{tablenotes}
\item[1] \footnotesize{This table includes the estimation results of equation (\ref{eq:iden}).}
\item[2] \footnotesize{The four groups stand for White Husband White Wife (WW), White Husband Hispanic Wife (WH), Hispanic Husband White (HW), and Hispanic Husband Hispanic Wife (HH). $Hispanic_{i}$ is a dummy variable equal to one if a person identifies as Hispanic and zero otherwise.}
\item[3] \footnotesize{The sample is restricted to men working full-time full-year and are waged and salaried workers.}
\item[4] \footnotesize{Column one has the regression results when controlling for hours worked, age, and years of fixed effects. Column two has the results after controlling for education.}
\end{tablenotes}
\end{threeparttable}}
\end{table}

### 2023-01-20 output/tables/tab05-regression.tex
push project to github

--- 2023-01-20 my_paper/tables/tab06-regression.tex
@@
 
-\caption{Effect of Having Hispanic Last Name \label{tab:identreg}}
+\caption{Effect of Having Hispanic Last Name \label{tab:lastnamereg}}
 \centering
@@
 \midrule
-$WH_{i} \times Hispanic_{i}$ & \num{-0.13}** & \num{-0.06}\\
- & (\num{0.05}) & \vphantom{2} (\num{0.05})\\
-$HW_{i} \times Hispanic_{i}$ & \num{-0.16}*** & \num{-0.11}**\\
- & (\num{0.05}) & \vphantom{1} (\num{0.05})\\
-$HH_{i} \times Hispanic_{i}$ & \num{-0.11}** & \num{-0.06}\\
- & (\num{0.05}) & \vphantom{2} (\num{0.04})\\
-$WH_{i}$ & \num{0.03} & \num{0.01}\\
- & (\num{0.05}) & \vphantom{1} (\num{0.04})\\
-$HW_{i}$ & \num{0.01} & \num{0.04}\\
- & (\num{0.05}) & (\num{0.05})\\
-$HH_{i}$ & \num{-0.06} & \num{0.00}\\
- & (\num{0.05}) & (\num{0.04})\\
+$WH_{i}$ & \num{-0.09}*** & \num{-0.04}**\\
+ & (\num{0.02}) & (\num{0.02})\\
+$HW_{i}$ & \num{-0.14}*** & \num{-0.06}***\\
+ & (\num{0.02}) & (\num{0.01})\\
+$HH_{i}$ & \num{-0.16}*** & \num{-0.06}***\\
+ & (\num{0.01}) & (\num{0.01})\\
 \midrule
-$HW_{i} \times Hispanic_{i} - WH_{i} \times Hispanic_{i}$ & -0.03 & -0.04\\
- & (0.07) & (0.07)\\
+$HW_{i} - WH_{i}$ & -0.05** & -0.02\\
+ & (0.02) & (0.02)\\
 \midrule
@@
 Hours Worked & X & X\\
+Year & X & X\\
 Age & X & X\\
-Year FE & X & X\\
 Education &  & X\\
@@
 \begin{tablenotes}
-\item[1] \footnotesize{This table includes the estimation results of equation (\ref{eq:iden}).}
-\item[2] \footnotesize{The four groups stand for White Husband White Wife (WW), White Husband Hispanic Wife (WH), Hispanic Husband White (HW), and Hispanic Husband Hispanic Wife (HH).
-$Hispanic_{i}$ is a dummy variable equal to one if a person identifies as Hispanic and zero otherwise.}
+\item[1] \footnotesize{This table includes the estimation results of equation (\ref{eq:1a}).}
+\item[2] \footnotesize{The four groups stand for White Husband White Wife (WW), White Husband Hispanic Wife (WH), Hispanic Husband White (HW), and Hispanic Husband Hispanic Wife (HH).}
 \item[3] \footnotesize{The sample is restricted to men working full-time full-year and are waged and salaried workers.}
-\item[4] \footnotesize{Column one has the regression results when controlling for hours worked, age, and years of fixed effects. Column two has the results after controlling for education.}
+\item[4] \footnotesize{Column one has the regression results when controlling for hours worked, age, and fixed effects. Column two has the results after controlling for education.}
 \end{tablenotes}

### 2023-01-20 output/tables/tab07-regression.tex
push project to github

--- 2023-01-20 output/tables/tab05-regression.tex
@@
 
-\caption{Effect of Having Hispanic Last Name \label{tab:lastnamereg}}
+\caption{Hsiapnic–White Earnings Gap Using Hispanic Variable Only \label{tab:hispwhitegap}}
 \centering
@@
 \midrule
-$WH_{i}$ & \num{-0.09}*** & \num{-0.04}**\\
- & (\num{0.02}) & (\num{0.02})\\
-$HW_{i}$ & \num{-0.14}*** & \num{-0.06}***\\
- & (\num{0.02}) & (\num{0.01})\\
-$HH_{i}$ & \num{-0.16}*** & \num{-0.06}***\\
- & (\num{0.01}) & (\num{0.01})\\
+$Hispanic_{i}$ & \num{-0.17}*** & \num{-0.07}***\\
+ & (\num{0.00}) & (\num{0.00})\\
 \midrule
-$HW_{i} - WH_{i}$ & -0.05** & -0.02\\
- & (0.02) & (0.02)\\
-\midrule
-Observations & \num{129359} & \num{129359}\\
+Observations & \num{137977} & \num{137977}\\
 \textit{Controlling for:} &  & \\
 Hours Worked & X & X\\
-Year & X & X\\
 Age & X & X\\
+Year FE & X & X\\
 Education &  & X\\
@@
 \begin{tablenotes}
-\item[1] \footnotesize{This table includes the estimation results of equation (\ref{eq:1a}).}
-\item[2] \footnotesize{The four groups stand for White Husband White Wife (WW), White Husband Hispanic Wife (WH), Hispanic Husband White (HW), and Hispanic Husband Hispanic Wife (HH).}
-\item[3] \footnotesize{The sample is restricted to men working full-time full-year and are waged and salaried workers.}
+\item[1] \footnotesize{This table includes the estimation results of equation (\ref{eq:naive}).}
+\item[2] \footnotesize{$Hispanic_{i}$ is a dummy variable equal to one if a person identifies as Hispanic and zero otherwise.}
+\item[3] \footnotesize{The sample is restricted to men working full time full-year and are waged and salaried workers.}
 \item[4] \footnotesize{Column one has the regression results when controlling for hours worked, age, and fixed effects. Column two has the results after controlling for education.}

### 2023-01-20 programs/01-packages-wds.R

@@ -1,0 +1,126 @@
+# This a script to downlaod
+# packages and set working
+# directories
+
+# date: May 18th, 2022
+
+if (!require("pacman")) install.packages("pacman")
+pacman::p_load(tictoc, parallel, pbapply, future, 
+               future.apply, furrr, RhpcBLASctl, memoise, 
+               here, foreign, mfx, tidyverse, hrbrthemes, 
+               estimatr, ivreg, fixest, sandwich, lmtest, 
+               margins, vtable, broom, modelsummary, 
+               stargazer, fastDummies, recipes, dummy, 
+               gplots, haven, huxtable, kableExtra, 
+               gmodels, survey, gtsummary, data.table, 
+               tidyfast, dtplyr, microbenchmark, ggpubr, 
+               tibble, viridis, wesanderson, censReg, 
+               rstatix, srvyr, formatR, sysfonts, 
+               showtextdb, showtext, thematic, 
+               sampleSelection, textme, paletteer, 
+               wesanderson, patchwork, RStata, car,
+               #textme, lodown,
+               BiocManager, Polychrome, effects,
+               maps, sf, multcomp, cdlTools,
+               finalfit, ggtext, glue, scales, 
+               gganimate, ggrepel, MetBrewer, fs,
+               marginaleffects, gghighlight, ggview,
+               camcorder, rnaturalearth, rnaturalearthdata,
+               latex2exp, janitor)
+options("RStata.StataPath" = "/Applications/Stata/StataSE.app/Contents/MacOS/stata-se")
+options("RStata.StataVersion" = 17)
+            
+# devtools::install_github("thomasp85/patchwork")
+# devtools::install_github("ajdamico/lodown")
+## My preferred ggplot2 plotting theme (optional)
+## https://github.com/hrbrmstr/hrbrthemes
+# scale_fill_ipsum()
+# scale_color_ipsum()
+font_add_google("Fira Sans", "firasans")
+font_add_google("Fira Code", "firasans")
+
+showtext_auto()
+
+showtext_opts(dpi = 300)
+camcorder::gg_record(
+  dir = figures_wd,
+  device = "png",
+  width = 10,
+  height = 10 + 9 / 16, #16:9 ratio
+  units = 'cm',
+  dpi = 300
+)
+theme_customs <- function() {
+  theme_minimal(base_family = "IBM Plex Sans Condensed") +
+    theme(panel.grid.minor = element_blank(),
[…]

### 2023-01-20 programs/07-table-one-parent-types.R

@@ -1,0 +1,112 @@
+# Table of observation number
+# of different types of parents
+
+# table one
+
+# date: January 19th, 2023
+
+DATA <- read_dta(file.path(datasets,"BySexAnalysisData.dta"))
+
+# all observations
+
+ALL <- DATA %>%
+  summarise(n())
+# calculate the number 
+# of white fathers
+# white mothers
+
+WW <- DATA %>%
+  filter(Type == "White-White") %>%
+  summarise(n())
+# prettyNum(WW, big.mark = ",")
+WW_perc <- paste0("(",round(WW/ALL, digits = 2), ")")
+
+# calculate the number 
+# of hispanic fathers
+# white mothers
+
+HW <- DATA %>%
+  filter(Type == "Hispanic-White") %>%
+  summarise(n())
+HW_perc <- paste0("(",round(HW/ALL, digits = 2), ")")
+
+# calculate the number 
+# of white fathers
+# hispanic mothers
+
+WH <- DATA %>%
+  filter(Type == "White-Hispanic") %>%
+  summarise(n())
+WH_perc <- paste0("(",round(WH/ALL, digits = 2), ")")
+
+# calculate the number 
+# of hispanic fathers
+# hispanic mothers
+
+HH <- DATA %>%
+  filter(Type == "Hispanic-Hispanic") %>%
+  summarise(n())
+HH_perc <- paste0("(",round(HH/ALL, digits = 2), ")")
+
+# table
+
+first_col <- c(" ",
+               "Observations"
+)
+
+second_col <- c("\\specialcell{White Father \\\\ White Mother}",
+               paste0("\\specialcell{", 
+                      prettyNum(WW, big.mark = ","), 
+                      "\\\\",
+                      WW_perc,
+                      "}"
+                      )
+)
+
+third_col <- c("\\specialcell{White Father \\\\ Hispanic Mother}",
+                paste0("\\specialcell{", 
+                       prettyNum(WH, big.mark = ","), 
+                       "\\\\",
+                       WH_perc,
+                       "}"
+                )
+)
+
+fourth_col <- c("\\specialcell{Hispanic Father \\\\ White Mother}",
+                paste0("\\specialcell{", 
+                       prettyNum(HW, big.mark = ","), 
+                       "\\\\",
+                       HW_perc,
+                       "}"
+                )
+)
+
+fifth_col <- c("\\specialcell{Hispanic Father \\\\ Hispanic Mother}",
[…]

### 2023-01-20 programs/10-table-two-couple-types.R

@@ -1,0 +1,113 @@
+# Table of observation number
+# of different types of parents
+
+# table two
+
+# date: January 19th, 2023
+
+DATA <- read_dta(file.path(datasets,"ParentDataFull.dta"))
+
+# all observations
+
+ALL <- DATA %>%
+  summarise(n())
+# calculate the number 
+# of white Husbands
+# white Wifes
+
+WW <- DATA %>%
+  filter(CoupleType == "White-White") %>%
+  summarise(n())
+# prettyNum(WW, big.mark = ",")
+WW_perc <- paste0("(",round(WW/ALL, digits = 2), ")")
+
+# calculate the number 
+# of hispanic Husbands
+# white Wifes
+
+HW <- DATA %>%
+  filter(CoupleType == "Hispanic-White") %>%
+  summarise(n())
+HW_perc <- paste0("(",round(HW/ALL, digits = 2), ")")
+
+# calculate the number 
+# of white Husbands
+# hispanic Wifes
+
+WH <- DATA %>%
+  filter(CoupleType == "White-Hispanic") %>%
+  summarise(n())
+WH_perc <- paste0("(",round(WH/ALL, digits = 2), ")")
+
+# calculate the number 
+# of hispanic Husbands
+# hispanic Wifes
+
+HH <- DATA %>%
+  filter(CoupleType == "Hispanic-Hispanic") %>%
+  summarise(n())
+HH_perc <- paste0("(",round(HH/ALL, digits = 2), ")")
+
+# table
+
+first_col <- c(" ",
+               "Observations"
+)
+
+second_col <- c("\\specialcell{White Husband \\\\ White Wife}",
+                paste0("\\specialcell{", 
+                       prettyNum(WW, big.mark = ","), 
+                       "\\\\",
+                       WW_perc,
+                       "}"
+                )
+)
+
+third_col <- c("\\specialcell{White Husband \\\\ Hispanic Wife}",
+               paste0("\\specialcell{", 
+                      prettyNum(WH, big.mark = ","), 
+                      "\\\\",
+                      WH_perc,
+                      "}"
+               )
+)
+
+fourth_col <- c("\\specialcell{Hispanic Husband \\\\ White Wife}",
+                paste0("\\specialcell{", 
+                       prettyNum(HW, big.mark = ","), 
+                       "\\\\",
+                       HW_perc,
+                       "}"
+                )
+)
+
[…]

### 2023-01-20 programs/12-table-four-summary-stats.R

@@ -1,0 +1,780 @@
+# summary stats
+
+# table four
+# only hispanics
+
+# date: January 19th, 2023
+
+IndividualData <- read_dta(file.path(datasets,"BySexAnalysisData.dta")) |> 
+  mutate(Education = as.numeric(Education)) |> 
+  filter(Hispanic_ID == 1)
+ParentData     <- read_dta(file.path(datasets,"ParentDataFull.dta")) |> 
+  mutate(Tot_ed = HusbandEducation + WifeEducation)
+
+row1 <- c("Variables", 
+          "\\specialcell{White Father \\\\ White Mother \\\\ (WW) \\\\ (i)}",
+          "\\specialcell{White Father \\\\ Hispanic Mother \\\\ (WH) \\\\ (ii)}",
+          "\\specialcell{Hispanic Father \\\\ White Mother \\\\ (HW) \\\\ (iii)}",
+          "\\specialcell{Hispanic Father \\\\ Hispanic Mother \\\\ (HH) \\\\ (iv)}",
+          "\\specialcell{HH - WW \\\\ (v)}",
+          "\\specialcell{HW - WH \\\\ (vi)}"
+)
+dim(row1) <- c(1,7)
+row2 <- c("Panel A: Parent’s", " ", " ", " ", " ", " ", " ")
+dim(row2) <- c(1,7)
+
+# parent's panel
+
+## Husband's Education
+HusbandEducation <- ParentData |> 
+  group_by(CoupleType) |> 
+  summarise(HusEducation   = weighted.mean(HusbandEducation, na.rm = T, hhwt),
+            HusEducationSD = weighted.sd(HusbandEducation, na.rm = T, hhwt))
+
+model <- lm(HusbandEducation ~ 0 + WW + WM + MW + MM, 
+            data = ParentData, 
+            weights = hhwt)
+test1 <- tidy(glht(model, linfct = c("MM - WW = 0")))
+pvalues1 <- test1$std.error
+differences1 <- test1$estimate
+
+test2 <- tidy(glht(model, linfct = c("MW - WM = 0")))
+pvalues2 <- test2$std.error
+differences2 <- test2$estimate
+
+row3 <- c(
+  "Husband’seducation (Total Years)",
+  paste0("\\specialcell{", 
+         round(HusbandEducation[4,2], digits = 2), 
+         "\\\\",
+         "(",
+         round(HusbandEducation[4,3], digits = 2),
+         ")}"
+  ),
[…]

### 2023-01-20 programs/13-table-five-regression.R

@@ -1,0 +1,133 @@
+# summary stats
+
+# table Five
+# Regression
+
+# date: January 19th, 2023
+
+IndividualData <- read_dta(file.path(datasets,"BySexAnalysisData.dta")) |> 
+  mutate(Education = as.numeric(Education)) |> 
+  filter((WW == 1 & Hispanic_ID == 0) | WH == 1 | HW == 1 | HH == 1)
+
+ParentDummies <- c("WH", "HW", "HH", "uhrsworkly")
+              
+# By generation
+reg1 <- list(
+  "\\specialcell{(1) \\\\ Log annual earnings}" = feols(lninctot_1999 ~ .[ParentDummies] | year + age, se = "iid",
+                                                            data = IndividualData |> 
+                                                          filter(sex == 1 & FTFY == 1 & Self_employed == 0)),
+  "\\specialcell{(2) \\\\  Log annual earnings}" = feols(lninctot_1999 ~ .[ParentDummies] | year + age + educ, se = "iid",
+                                                         data = IndividualData |> 
+                                                           filter(sex == 1 & FTFY == 1 & Self_employed == 0))
+  
+)
+
+# calculate means to add
+# as a row
+
+
+test1 <- tidy(glht(reg1[[1]], linfct = c("HW - WH = 0")))
+differences1 <- case_when(abs(test1$estimate/test1$std.error)   >= 2.58  ~ paste0(sprintf("%2.5s",round_tidy(test1$estimate, digits = 2)), "***"),
+          abs(test1$estimate/test1$std.error)   <  2.58 & 
+            abs(test1$estimate/test1$std.error) >= 1.96  ~ paste0(sprintf("%2.5s",round_tidy(test1$estimate, digits = 2)), "**"),
+          abs(test1$estimate/test1$std.error)   <  1.96  &
+            abs(test1$estimate/test1$std.error) >= 1.645 ~ paste0(sprintf("%2.5s",round_tidy(test1$estimate, digits = 2)), "*"),
+          abs(test1$estimate/test1$std.error)   < 1.645  ~ round_tidy(test1$estimate, digits = 2)
+          )
+
+test2 <- tidy(glht(reg1[[2]], linfct = c("HW - WH = 0")))
+differences2 <- case_when(abs(test2$estimate/test2$std.error)   >= 2.58  ~ paste0(sprintf("%2.5s",round_tidy(test2$estimate, digits = 2)), "***"),
+                          abs(test2$estimate/test2$std.error)   <  2.58 & 
[…]

### 2023-01-20 programs/14-table-six-regression.R

@@ -1,0 +1,135 @@
+# summary stats
+
+# table Five
+# Regression
+
+# date: January 19th, 2023
+
+IndividualData <- read_dta(file.path(datasets,"BySexAnalysisData.dta")) |> 
+  mutate(Education = as.numeric(Education)) |> 
+  filter((WW == 1 & Hispanic_ID == 0) | WH == 1 | HW == 1 | HH == 1)
+
+ParentDummies <- c("WH:Hispanic_ID", "HW:Hispanic_ID", "HH:Hispanic_ID","WH", "HW", "HH", "uhrsworkly")
+
+# By generation
+reg1 <- list(
+  "\\specialcell{(1) \\\\ Log annual earnings}" = feols(lninctot_1999 ~ .[ParentDummies] | year + age, se = "iid",
+                                                        data = IndividualData |> 
+                                                          filter(sex == 1 & FTFY == 1 & Self_employed == 0)),
+  "\\specialcell{(2) \\\\  Log annual earnings}" = feols(lninctot_1999 ~ .[ParentDummies] | year + age + educ, se = "iid",
+                                                         data = IndividualData |> 
+                                                           filter(sex == 1 & FTFY == 1 & Self_employed == 0))
+  
+)
+
+# calculate means to add
+# as a row
+
+
+test1 <- tidy(glht(reg1[[1]], linfct = c("Hispanic_ID:HW - WH:Hispanic_ID = 0")))
+differences1 <- case_when(abs(test1$estimate/test1$std.error)   >= 2.58  ~ paste0(sprintf("%2.5s",round_tidy(test1$estimate, digits = 2)), "***"),
+                          abs(test1$estimate/test1$std.error)   <  2.58 & 
+                            abs(test1$estimate/test1$std.error) >= 1.96  ~ paste0(sprintf("%2.5s",round_tidy(test1$estimate, digits = 2)), "**"),
+                          abs(test1$estimate/test1$std.error)   <  1.96  &
+                            abs(test1$estimate/test1$std.error) >= 1.645 ~ paste0(sprintf("%2.5s",round_tidy(test1$estimate, digits = 2)), "*"),
+                          abs(test1$estimate/test1$std.error)   < 1.645  ~ round_tidy(test1$estimate, digits = 2)
+)
+
+test2 <- tidy(glht(reg1[[2]], linfct = c("Hispanic_ID:HW - WH:Hispanic_ID = 0")))
[…]

### 2023-01-20 my_paper/FinalDraft.tex
edits

@@
 & \beta_{4} WH_{i} \cdot Hispanic_{i} + \beta_{5} HW_{i}\cdot Hispanic_{i} +  \beta_{6} HH_{i} \cdot Hispanic_{i}+ \nonumber \\
-&X_{it}^{\prime} + \gamma_{t}+\phi_{it} \nonumber
+&X_{it} \pi + \gamma_{t}+\phi_{it} \nonumber
 \end{align}

### 2023-04-11 programs/10-table-two-couple-types.R
go back

@@ -111,3 +111,16 @@ knitr::kable(Table_cols, "latex", align = "lcccc",
   add_header_above(c(" " = 1, "Couples\' Type" = 4)) |> 
   save_kable(file.path(tables_wd,"tab02-observations-by-couple.tex")) %>% 
   save_kable(file.path(thesis_tabs,"tab02-observations-by-couple.tex"))
+
+# presentation table
+
+knitr::kable(Table_cols, "latex", align = "lcccc",
+             booktabs = T,
+             escape = F,
+             caption = "Couples' Type \\label{tab:mat2}") %>%
+  column_spec(1, bold = T) %>%
+  kable_classic(full_width = F) %>%
+  kable_styling(bootstrap_options = c("hover", "condensed"), 
+                latex_options = c("scale_down", "hold_position")) |>  
+  add_header_above(c(" " = 1, "Couples\' Type" = 4)) |> 
+  save_kable(file.path(dissertation_wd,"tables/tab02-observations-by-couple.tex"))

### 2023-04-11 programs/95-make-all.R

@@ -22,7 +22,7 @@ add_health_data_dir <- "/Users/hhadah/Dropbox/Research/My Research Data and Idea
 # pres_tabs <- paste0(mdir,"/presentations/jmp-presentation/tables")
 # pres_plots <- paste0(mdir,"/presentations/jmp-presentation/figures")
 
-
+dissertation_wd <- paste0("/Users/hhadah/Documents/GiT/my_thesis/presentation/dissertation")
 ### run do files and scripts
 
 # main scripts

### 2023-04-11 my_paper/tables/tab03-summary-stats.tex
edit tables for presentation

@@
 
-\caption{Children's outcome using parent's place of birth \label{tab:c&p1}}
+\caption{Summary statistics of outcomes using parent's place of birth \label{tab:c&p1}}
 \centering
@@
 \hspace{1em}Wife’seducation (Total Years) & \specialcell{12.74\\(2.12)} & \specialcell{11.03\\(3.92)} & \specialcell{11.54\\(3.12)} & \specialcell{8.6\\(4.13)} & \specialcell{-4.13\\(0.02)} & \specialcell{0.51\\(0.04)}\\
-\hspace{1em}Total Household seducation (Total Years) & \specialcell{25.78\\(4.08)} & \specialcell{23.35\\(6.51)} & \specialcell{22.19\\(6.69)} & \specialcell{17.54\\(7.83)} & \specialcell{-4.13\\(0.02)} & \specialcell{0.51\\(0.04)}\\
+\hspace{1em}Total Household seducation (Total Years) & \specialcell{25.78\\(4.08)} & \specialcell{23.35\\(6.51)} & \specialcell{22.19\\(6.69)} & \specialcell{17.54\\(7.83)} & \specialcell{-8.25\\(0.03)} & \specialcell{-1.16\\(0.07)}\\
 \textbf{Panel B: Education} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
@@
 \textbf{Panel C: Employment and Earnings} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
-\hspace{1em}Men’s Employment Rate & \specialcell{0.95\\(0.22)} & \specialcell{0.94\\(0.23)} & \specialcell{0.92\\(0.26)} & \specialcell{0.93\\(0.26)} & \specialcell{-0.02\\(0)} & \specialcell{-0.02\\(0)}\\
-\hspace{1em}Women’s Employment Rate & \specialcell{0.96\\(0.2)} & \specialcell{0.95\\(0.22)} & \specialcell{0.93\\(0.25)} & \specialcell{0.94\\(0.24)} & \specialcell{-0.02\\(0)} & \specialcell{-0.02\\(0)}\\
+\hspace{1em}Men’s Employment Rate & \specialcell{0.95\\(0.22)} & \specialcell{0.94\\(0.23)} & \specialcell{0.92\\(0.26)} & \specialcell{0.93\\(0.26)} & \specialcell{-0.02\\(0.00)} & \specialcell{-0.02\\(0.00)}\\
+\hspace{1em}Women’s Employment Rate & \specialcell{0.96\\(0.2)} & \specialcell{0.95\\(0.22)} & \specialcell{0.93\\(0.25)} & \specialcell{0.94\\(0.24)} & \specialcell{-0.02\\(0.00)} & \specialcell{-0.02\\(0.00)}\\
 \addlinespace

### 2023-04-11 my_paper/tables/tab04-summary-stats.tex
edit tables for presentation

--- 2023-01-20 my_paper/tables/tab03-summary-stats.tex
@@
 
-\caption{Children's outcome using parent's place of birth \label{tab:c&p1}}
+\caption{Summary statistics of outcomes using parent's place of birth only for those that self-identify as Hispanic \label{tab:c&p2}}
 \centering
@@
 \hspace{1em}Wife’seducation (Total Years) & \specialcell{12.74\\(2.12)} & \specialcell{11.03\\(3.92)} & \specialcell{11.54\\(3.12)} & \specialcell{8.6\\(4.13)} & \specialcell{-4.13\\(0.02)} & \specialcell{0.51\\(0.04)}\\
-\hspace{1em}Total Household seducation (Total Years) & \specialcell{25.78\\(4.08)} & \specialcell{23.35\\(6.51)} & \specialcell{22.19\\(6.69)} & \specialcell{17.54\\(7.83)} & \specialcell{-4.13\\(0.02)} & \specialcell{0.51\\(0.04)}\\
+\hspace{1em}Total Household seducation (Total Years) & \specialcell{25.78\\(4.08)} & \specialcell{23.35\\(6.51)} & \specialcell{22.19\\(6.69)} & \specialcell{17.54\\(7.83)} & \specialcell{-8.25\\(0.03)} & \specialcell{-1.16\\(0.07)}\\
 \textbf{Panel B: Education} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
 \addlinespace
-\hspace{1em}Men’s education (Total Years) & \specialcell{13.91\\(2.39)} & \specialcell{13.58\\(2.35)} & \specialcell{13.21\\(2.32)} & \specialcell{12.91\\(2.26)} & \specialcell{-1.00\\(0.01)} & \specialcell{-0.36\\(0.03)}\\
-\hspace{1em}Women’s education (Total Years) & \specialcell{14.29\\(2.41)} & \specialcell{13.87\\(2.47)} & \specialcell{13.42\\(2.35)} & \specialcell{13.27\\(2.37)} & \specialcell{-1.01\\(0.01)} & \specialcell{-0.46\\(0.03)}\\
+\hspace{1em}Men’s education (Total Years) & \specialcell{12.97\\(2.15)} & \specialcell{13.45\\(2.37)} & \specialcell{13.13\\(2.27)} & \specialcell{12.89\\(2.25)} & \specialcell{-0.08\\(0.01)} & \specialcell{-0.32\\(0.03)}\\
+\hspace{1em}Women’s education (Total Years) & \specialcell{13.23\\(2.25)} & \specialcell{13.75\\(2.41)} & \specialcell{13.32\\(2.34)} & \specialcell{13.26\\(2.37)} & \specialcell{0.03\\(0.01)} & \specialcell{-0.43\\(0.03)}\\
 \textbf{Panel C: Employment and Earnings} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
-\hspace{1em}Men’s Employment Rate & \specialcell{0.95\\(0.22)} & \specialcell{0.94\\(0.23)} & \specialcell{0.92\\(0.26)} & \specialcell{0.93\\(0.26)} & \specialcell{-0.02\\(0)} & \specialcell{-0.02\\(0)}\\
-\hspace{1em}Women’s Employment Rate & \specialcell{0.96\\(0.2)} & \specialcell{0.95\\(0.22)} & \specialcell{0.93\\(0.25)} & \specialcell{0.94\\(0.24)} & \specialcell{-0.02\\(0)} & \specialcell{-0.02\\(0)}\\
+\hspace{1em}Men’s Employment Rate & \specialcell{0.93\\(0.26)} & \specialcell{0.94\\(0.23)} & \specialcell{0.92\\(0.27)} & \specialcell{0.93\\(0.26)} & \specialcell{0.00\\(0.00)} & \specialcell{-0.02\\(0.00)}\\
+\hspace{1em}Women’s Employment Rate & \specialcell{0.94\\(0.25)} & \specialcell{0.94\\(0.23)} & \specialcell{0.93\\(0.26)} & \specialcell{0.94\\(0.24)} & \specialcell{0.00\\(0.00)} & \specialcell{-0.02\\(0.00)}\\
 \addlinespace
-\hspace{1em}Men’s Log Hourly Earnings & \specialcell{2.48\\(0.45)} & \specialcell{2.41\\(0.45)} & \specialcell{2.4\\(0.44)} & \specialcell{2.41\\(0.43)} & \specialcell{-0.07\\(0.01)} & \specialcell{-0.00\\(0.02)}\\
-\hspace{1em}Women’s Hourly Earnings & \specialcell{2.33\\(0.49)} & \specialcell{2.33\\(0.46)} & \specialcell{2.27\\(0.45)} & \specialcell{2.3\\(0.41)} & \specialcell{-0.02\\(0.01)} & \specialcell{-0.06\\(0.02)}\\
-\hspace{1em}Men’s Log Annual Earnings & \specialcell{10.25\\(1.01)} & \specialcell{10.08\\(1.05)} & \specialcell{10.04\\(1.06)} & \specialcell{10.01\\(1.04)} & \specialcell{-0.25\\(0.01)} & \specialcell{-0.04\\(0.03)}\\
-\hspace{1em}Women’s Hourly Earnings & \specialcell{9.46\\(1.78)} & \specialcell{9.54\\(1.64)} & \specialcell{9.47\\(1.57)} & \specialcell{9.53\\(1.52)} & \specialcell{0.07\\(0.02)} & \specialcell{-0.07\\(0.06)}\\
-\textbf{Panel D: Hispanic Identity} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
-\addlinespace
-\hspace{1em}Men & \specialcell{0.05} & \specialcell{0.81} & \specialcell{0.88} & \specialcell{0.97} &  & \\
-\hspace{1em}Women & \specialcell{0.06} & \specialcell{0.85} & \specialcell{0.87} & \specialcell{0.97} &  & \\
+\hspace{1em}Men’s Log Hourly Earnings & \specialcell{2.4\\(0.44)} & \specialcell{2.41\\(0.45)} & \specialcell{2.4\\(0.44)} & \specialcell{2.41\\(0.43)} & \specialcell{0.01\\(0.01)} & \specialcell{-0.01\\(0.02)}\\
+\hspace{1em}Women’s Hourly Earnings & \specialcell{2.26\\(0.43)} & \specialcell{2.32\\(0.45)} & \specialcell{2.27\\(0.45)} & \specialcell{2.3\\(0.41)} & \specialcell{0.04\\(0.01)} & \specialcell{-0.05\\(0.02)}\\
+\hspace{1em}Men’s Log Annual Earnings & \specialcell{10.02\\(1.02)} & \specialcell{10.06\\(1.06)} & \specialcell{10.03\\(0.99)} & \specialcell{10\\(1.04)} & \specialcell{-0.02\\(0.01)} & \specialcell{-0.03\\(0.04)}\\
+\hspace{1em}Women’s Hourly Earnings & \specialcell{9.44\\(1.59)} & \specialcell{9.55\\(1.59)} & \specialcell{9.47\\(1.55)} & \specialcell{9.52\\(1.52)} & \specialcell{0.08\\(0.02)} & \specialcell{-0.08\\(0.05)}\\
 \bottomrule

### 2023-04-11 programs/12-table-four-summary-stats.R

@@ -159,7 +159,7 @@ TotalEducation <- ParentData |>
   summarise(HHducation   = weighted.mean(Tot_ed, na.rm = T, hhwt),
             HHducationSD = weighted.sd(Tot_ed, na.rm = T, hhwt))
 
-model <- lm(WifeEducation ~ 0 + WW + WM + MW + MM, 
+model <- lm(Tot_ed ~ 0 + WW + WM + MW + MM, 
             data = ParentData, 
             weights = hhwt)
 test1 <- tidy(glht(model, linfct = c("MM - WW = 0")))
@@ -406,17 +406,17 @@ row10 <- c(
          ")}"
   ),
   paste0("\\specialcell{", 
-         round(differences1[[1]], digits = 2), 
+         sprintf("%.2f",round(differences1[[1]], digits = 2)), 
          "\\\\",
          "(",
-         round(pvalues1[[1]], digits = 2),
+         sprintf("%.2f",round(pvalues1[[1]], digits = 2)),
          ")}"
   ),
   paste0("\\specialcell{", 
          round(differences2[[1]], digits = 2), 
          "\\\\",
          "(",
-         round(pvalues2[[1]], digits = 2),
+         sprintf("%.2f",round(pvalues2[[1]], digits = 2)),
          ")}"
   )
 )
@@ -469,17 +469,17 @@ row11 <- c(
          ")}"
   ),
   paste0("\\specialcell{", 
-         round(differences1[[1]], digits = 2), 
+         sprintf("%.2f",round(differences1[[1]], digits = 2)), 
          "\\\\",
          "(",
-         round(pvalues1[[1]], digits = 2),
+         sprintf("%.2f",round(pvalues1[[1]], digits = 2)),
          ")}"
   ),
   paste0("\\specialcell{", 
          round(differences2[[1]], digits = 2), 
          "\\\\",
          "(",
-         round(pvalues2[[1]], digits = 2),
+         sprintf("%.2f",round(pvalues2[[1]], digits = 2)),
          ")}"
   )
 )
@@ -760,7 +760,7 @@ Table_rows <-  Table_rows |>
 knitr::kable(Table_rows, "latex", align = "lcccccc",
              booktabs = T,
              escape = F,
-             caption = "Children's outcome using parent's place of birth Only for those that Identify as Hispanic \\label{tab:c&p2}") %>%
[…]

### 2023-04-11 programs/14-table-six-regression.R

@@ -133,3 +133,22 @@ regression_tab %>%
 
 regression_tab %>%
   save_kable(file.path(thesis_tabs,"tab06-regression.tex"))
+
+# dissertaion table
+regression_tab <- modelsummary(reg1, fmt = 2,
+                               output = "latex",
+                               coef_map = cm,
+                               add_rows = all_row,
+                               gof_map = gm,
+                               escape = F,
+                               #gof_omit = 'DF|Deviance|R2|AIC|BIC|Log.Lik.|F|Std.Errors',
+                               stars= c('***' = 0.01, '**' = 0.05, '*' = 0.1)) %>%
+  kable_styling(bootstrap_options = c("hover", "condensed"), 
+                latex_options = c("scale_down", "hold_position")
+  ) |> 
+  row_spec(14, hline_after = T)
+
+
+regression_tab %>%
+  save_kable(file.path(dissertation_wd,"tables/tab06-regression.tex"))
+

### 2023-04-12 my_paper/Hadah-last-names.tex
edit link
add sig levels to sum tables and hilight defense columns

--- 2023-01-20 my_paper/FinalDraft.tex
@@
 
-%\begin{center}
-%\href{https://hussainhadah.com/worksinprogress/hispanicdiscrimnation/}{\textcolor{green!15!black!30!blue}{\footnotesize{\textsc{Click Here to Get the Most Updated Version}}}}
-%\end{center}
+\begin{center}
+\href{https://hhadah.github.io/hispanic-last-names/my_paper/Hadah-last-names.pdf}{\textcolor{green!15!black!30!blue}{\footnotesize{\textsc{Click Here to Get the Most Updated Version}}}}
+\end{center}
 
@@
 Hispanics constitute a large and growing portion of the population in the United States.
-As the number of Hispanics keeps increasing in the labor force, identifying whether they face ethnic discrimination grows in importance.
-Consequently, questions on assimilation and mobility grow in importance \citep{chettyUnitedStatesStill2014, chettyEffectsExposureBetter2016,chettyFadingAmericanDream2017,abramitzkyImmigrantsAssimilateMore2020a, abramitzkyNationImmigrantsAssimilation2014,abramitzkyCulturalAssimilationAge2016,chettyWhereLandOpportunity2014}, how identity could shift public opinions toward trade \citep{grossmanIdentityPoliticsTrade2021}, and how racial and gender attitudes affect the racial and gender earnings gaps \citep{charlesPrejudiceWagesEmpirical2008,charlesEffectsSexismAmerican2018}.
+As the number of Hispanics keeps increasing in the labor force, determining whether ethnic discrimination affects their labor market outcomes becomes increasingly crucial \citep{chettyUnitedStatesStill2014, chettyEffectsExposureBetter2016,chettyFadingAmericanDream2017,abramitzkyImmigrantsAssimilateMore2020a, abramitzkyNationImmigrantsAssimilation2014,abramitzkyCulturalAssimilationAge2016,chettyWhereLandOpportunity2014}, how identity could shift public opinions toward trade \citep{grossmanIdentityPoliticsTrade2021}, and how racial and gender attitudes affect the racial and gender earnings gaps \citep{charlesPrejudiceWagesEmpirical2008,charlesEffectsSexismAmerican2018}.
 Thus, it is important to understand whether a person's ethnicity affects their labor market outcomes.
+Assimilation and mobility are crucial because they reflect how well Hispanics can integrate into society and move up the socioeconomic ladder.
 
@@
 What is the effect of identifying as Hispanic on earnings?
-Moreover, I aim to show that comparing Hispanic Whites to non-Hispanic Whites might create an artificially higher earnings gap since the two groups differ on many observable characteristics.
-Others have attempted to compare how native-born White Hispanics fare to non-Hispanic Whites and foreign-born Hispanics.
+Moreover, I aim to show that comparing Hispanic Whites to non-Hispanic Whites might create an artificially higher earnings gap since the two groups differ on many observable characteristics.\footnote{By identifying as Hispanic, I refer to individuals who self-report their ethnicity as Hispanic on surveys or other data collection instruments.} \footnote{Observable characteristics refer to factors that can be measured and quantified, such as education level, work experience, and immigration status.} Others have attempted to compare how native-born White Hispanics fare to non-Hispanic Whites and foreign-born Hispanics.
 In \citet{antman2020ethnic,antmanEthnicAttritionObserved2016,antmanEthnicAttritionObserved2016a,antmanEthnicAttritionAssimilation2020}, the authors compare the health and educational outcomes of Hispanic Whites to non-Hispanic Whites and native-born Hispanics to foreign-born Hispanics.
@@
 Therefore, understanding discrimination against Hispanics and their children is an important task.
-Discrimination against Hispanics has a couple of negative consequences.
+Discrimination against Hispanics has a couple of negative consequences, for example discrimination against Hispanics can lead to reduced job opportunities, lower wages, and hindered assimilation.
 First, if Hispanics face a substantial labor market penalty for their ethnic identity, it could hinder assimilation and integration.
@@
 Identifying discrimination is difficult because of factors that affect labor market outcomes that are unobservable to economists---like innate ability.
-Also, I should be able to disentangle skills differentials between Hispanics and non-Hispanic Whites.
-To causally identify ethnic discrimination against Hispanics and its effects on labor market outcomes, a researcher should compare two identical people that only differ in their last name.
+Also, I should be able to disentangle skills differentials between Hispanics and non-Hispanic Whites.\footnote{Skills differentials refer to differences in education, work experience, and other factors that can affect a person's job performance and wages.} To causally identify ethnic discrimination against Hispanics and its effects on labor market outcomes, a researcher should compare two identical people that only differ in their last name.
 Since this approach is unfeasible, researchers took different approaches.
@@
 
-I will use a method developed by \citet{rubinstein2014pride} in this research.
-I will compare children from intermarriages.
+This study utilizes a method developed by \citet{rubinstein2014pride} in this research.
+I will compare children from inter-ethnic marriages.
 More precisely, I will reach children of Hispanic fathers and White mothers (henceforth HW) to children of White fathers and Hispanic mothers (henceforth  WH).
@@
 Children of HW and WH marriages have more similar observable characteristics than children of homogeneous marriages--- i.e., White fathers-White mothers and Hispanic fathers-Hispanic mothers.
-Moreover, children from a Hispanic father and White mother household will have a Hispanic last name from their fathers, allowing me to test for ethnic signals on annual log earnings.
+Moreover, children from a Hispanic father and White mother household will have a Hispanic last name from their fathers, enabling the investigation of how ethnic signals, such as having a Hispanic last name, affect annual log earnings.
 
@@
 
-This is not the first attempt to use names as signals for race and ethnicity \citep{fryer2004causes, rubinstein2014pride, bertrand2004emily}.
+Previous studies have also used names as a proxy for race and ethnicity \citep{fryer2004causes, rubinstein2014pride, bertrand2004emily}.
 \citet{fryer2004causes} point out that names can be a predictor of a person's race.
@@
 They did, however, find that having a Black name, after controlling for the home environment at birth, does not affect their labor market outcomes.
-\citet{rubinstein2014pride} compared the children of mixed marriages between Sephardic and Ashkenazis in Israel.
-They found that workers with Sephardic last names earn substantially less than those with an Ashkenazi last name.
+\citet{rubinstein2014pride} compared the children of mixed marriages between Sephardic and Ashkenazis in Israel.\footnote{Ashkenazi and Sephardic are two distinct Jewish ethnic groups.} They found that workers with Sephardic last names earn substantially less than those with an Ashkenazi last name.
 \citet{bertrand2004emily} conducted an audit study by sending employers identical resumes that differ in the ethnic and racial signal of a name (Black sounding name versus a White sounding one).
@@
 Moreover, since the CPS does not provide data on parents' characteristics, essential to determine the family background, I use the census to construct synthetic parents.
-The census offers a more significant sample of potential parents.
+The census offers a larger sample of potential parents.
 Similar to the CPS, the 1970 to 1990 censuses ask about the person's place of birth and the individual's race and ethnicity.
 I employ this information to construct "synthetic parents" using a method developed by \citet{rubinstein2014pride}.
+I construct the synthetic parents by linking husbands and wives in the census data to each other.
+I assume that parents have children between the ages of 25 and 40, I then link these synthetic parents using the year they were surveyed to the year of birth of the "children" in the CPS sample.
 
-I construct the synthetic parents by linking husbands and wives in the census data to each other.
-I assume that parents have children between the ages of 25 and 40.
-I then link these synthetic parents using the year they were surveyed to the year of birth of the "children" in the CPS sample.
-
-\subsection{Children of the four parents' types}
+\subsection{Children of the four parental types}
 I will use the CPS for my primary analysis of the effect of having a Hispanic last name on earnings.
@@
 Also, WH and HW are comparable, unlike WW and HH children.
-These two groups are important to my analysis.
+These two groups are important to the analysis.
 Children of HW are going to be the treatment group.
@@
 At the same time, the WH children have White fathers and will take his non-Hispanic, or White, last name.
-A first glance at the summary statistics of these two groups, table \ref{tab:c&p1} columns (ii) and (iii) can show that they are different from the children of non-intermarriages.
+A first glance at the summary statistics of these two groups, table \ref{tab:c&p1} columns (ii) and (iii) can show that they are different from the children of non-inter-ethnic marriages.
 
@@
 
-Moreover, in table \ref{tab:c&p2}, I present the summary statistics for the four types but limit the sample to those that only identify as Hispanic\footnote{Limiting the sample to those that identify as Hispanic only affects WH, HW, and HH.
+Moreover, in table \ref{tab:c&p2}, I present the summary statistics for the four types and limit the sample to those that only identify as Hispanic\footnote{Limiting the sample to those that identify as Hispanic only affects WH, HW, and HH.
 Members of the WW group are not affected by this restriction.}.
@@
 The one difference is that those who identify as Hispanic do slightly worse.
+
  %In columns five and six of table \ref{tab:c1}, I present the gaps between HH and WW children and HW and WH children, respectively.
@@
 It can also be caused by differences in innate abilities, skills, and parental investments during infancy.
-To avoid these threats to identification, I compare children of intermarriages, HW and WH.
+To avoid these threats to identification, I compare children of inter-ethnic marriages, HW and WH.
 WH and HW children are equally Hispanic physically but provide employers/labor market with different signals.\footnote{WH and HW children are both half White, half Hispanic.} WH children will have a non-Hispanic last name, while HW children will have a Hispanic last name.
@@
 
-\subsection{Estimating the effect of having a Hispanic last name}
+\section{Estimating the effect of having a Hispanic last name}
 
@@
 I find that the earnings gaps between WH and WW are 8.8\%, 14.1\% between HW and WW, and 16.1\% between HH and WW.
-Moreover, the difference in earnings between HW and WH children, which captures the effect of having a Hispanic last name, equals a statistically significant 5.3 percentage points.
+Moreover, the difference in earnings between HW and WH children, that captures the effect of having a Hispanic last name, equals a statistically significant 5.3 percentage points.
 Meaning by comparing inter-ethnic children, a person with a Hispanic last name earns 5.3 percentage points less than someone with a White last name.
 However, more than half of the earnings gaps could be explained by educational differences between WW children and WH, HW, and HH.
-When I control for education, the gap between the three groups and WW is cut between 50-64.0\%.
+When I control for education, the gap between the three groups and WW is cut between by 50\% to 64.0\%.
 Furthermore, the last name effect decreases to a statistically insignificant 1.6 percentage points earnings gap.
@@
 After controlling for education, an HW man earns 5.9\% less than a WW man, and an HH man earns 5.8\% less.
-The effect of having a Hispanic last name on earnings decreased to 1.6 percentage points and became statistically insignificant after controlling for education.
+The effect of having a Hispanic last name on earnings decreased from 5.3 to 1.6 percentage points and became statistically insignificant after controlling for education.
 
-Therefore, Hispanics face big earnings gaps when compared to Whites.
-By comparing the children of intermarriages, I could study the impact of having a Hispanic last name on earnings.
+Therefore, Hispanics face big earnings gaps compared to non-Hispanic Whites.
+By comparing the children of inter-ethnic marriages, I could study the impact of having a Hispanic last name on earnings.
 Having a Hispanic last name carries a cost in the labor market.
@@
 Comparing the coefficients of the interaction terms to those of the $WH_{i}$, $HW_{i}$, and $HH_{i}$ shows that members of the three groups that do not identify as Hispanic earn about the same as WW.
-Moreover, the difference between $HW_{i} \cdot Hispanic_{i}$  and $WH_{i} \cdot Hispanic_{i}$ captures the effect of having a Hispanic last name and identifying as Hispanic compared to having a White last name and identifying as Hispanic.
+Moreover, the difference between $HW_{i} \cdot Hispanic_{i}$  and $WH_{i} \cdot Hispanic_{i}$ captures the effect of having a Hispanic last name and identifying as Hispanic compared to individuals who identify as Hispanic but have a White last name.
 A person with a Hispanic last name that identifies as Hispanic earns 3 percentage points less than a person that identifies as Hispanic but has a White last name.
-The difference, however, is statistically insignificant.
+The difference between the two interaction terms is statistically insignificant.
 
@@
 Furthermore, the difference between $HW_{i} \cdot Hispanic_{i}$  and $WH_{i} \cdot Hispanic_{i}$ increases to 4.3 percentage points but remains statistically insignificant.
-\subsection{Discussion}
+
+\section{Discussion}
 
@@
 After controlling for education, the gap decreases by more than half and becomes 6\%, which is also similar to the gap between HH and WW after controlling for education.
-Since differences between Whites and Hispanics, on the observable and unobservables are large, estimating the gap HW-WH could provide us with a bound on how much of the discrimination against Hispanics is taste-based.
+Since the differences between Whites and Hispanics, both observable and unobservable, are large, estimating the gap HW-WH could provide us with a bound on how much of the discrimination against Hispanics is taste-based.
 
@@
 The upper bound of the confidence interval is equivalent to 4.5 percentage points.
-Therefore, I can put an upper bound on the portion of the Hispanic-White earnings gap that could be explained by taste based discrimination.
+Therefore, I can put a potential upper bound on the portion of the Hispanic-White earnings gap that could be explained by taste based discrimination.
 In this case, the upper bound is equal to 4.5 percentage points.
@@
 As the Hispanic population keeps growing in the United States, studying discrimination against this group is also important.
-In this paper, I investigate discrimination against Hispanics in the labor market.
-More specifically, I explore the effect of having a Hispanic last name and identifying as Hispanic on annual log earnings.
+In this paper, I examine discrimination against Hispanics in the labor market.
+More specifically, I examine the impact of Hispanic last names and Hispanic identification on annual log earnings.
 

### 2023-04-12 my_paper/tables/tab03-summary-stats.tex
add sig levels to sum tables and hilight defense columns

@@
 \textbf{Panel A: Parent’s} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
-\hspace{1em}Husband’seducation (Total Years) & \specialcell{13.05\\(2.44)} & \specialcell{12.32\\(3.33)} & \specialcell{10.65\\(4.39)} & \specialcell{8.93\\(4.41)} & \specialcell{-4.11\\(0.02)} & \specialcell{-1.67\\(0.04)}\\
-\hspace{1em}Wife’seducation (Total Years) & \specialcell{12.74\\(2.12)} & \specialcell{11.03\\(3.92)} & \specialcell{11.54\\(3.12)} & \specialcell{8.6\\(4.13)} & \specialcell{-4.13\\(0.02)} & \specialcell{0.51\\(0.04)}\\
-\hspace{1em}Total Household seducation (Total Years) & \specialcell{25.78\\(4.08)} & \specialcell{23.35\\(6.51)} & \specialcell{22.19\\(6.69)} & \specialcell{17.54\\(7.83)} & \specialcell{-8.25\\(0.03)} & \specialcell{-1.16\\(0.07)}\\
+\hspace{1em}Husband’seducation (Total Years) & \specialcell{13.05\\(2.44)} & \specialcell{12.32\\(3.33)} & \specialcell{10.65\\(4.39)} & \specialcell{8.93\\(4.41)} & \specialcell{-4.11**\\(0.02)} & \specialcell{-1.67**\\(0.04)}\\
+\hspace{1em}Wife’seducation (Total Years) & \specialcell{12.74\\(2.12)} & \specialcell{11.03\\(3.92)} & \specialcell{11.54\\(3.12)} & \specialcell{8.6\\(4.13)} & \specialcell{-4.13**\\(0.02)} & \specialcell{0.51**\\(0.04)}\\
+\hspace{1em}Total Household seducation (Total Years) & \specialcell{25.78\\(4.08)} & \specialcell{23.35\\(6.51)} & \specialcell{22.19\\(6.69)} & \specialcell{17.54\\(7.83)} & \specialcell{-8.25**\\(0.03)} & \specialcell{-1.16*\\(0.07)}\\
 \textbf{Panel B: Education} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
 \addlinespace
-\hspace{1em}Men’s education (Total Years) & \specialcell{13.91\\(2.39)} & \specialcell{13.58\\(2.35)} & \specialcell{13.21\\(2.32)} & \specialcell{12.91\\(2.26)} & \specialcell{-1.00\\(0.01)} & \specialcell{-0.36\\(0.03)}\\
-\hspace{1em}Women’s education (Total Years) & \specialcell{14.29\\(2.41)} & \specialcell{13.87\\(2.47)} & \specialcell{13.42\\(2.35)} & \specialcell{13.27\\(2.37)} & \specialcell{-1.01\\(0.01)} & \specialcell{-0.46\\(0.03)}\\
+\hspace{1em}Men’s education (Total Years) & \specialcell{13.91\\(2.39)} & \specialcell{13.58\\(2.35)} & \specialcell{13.21\\(2.32)} & \specialcell{12.91\\(2.26)} & \specialcell{-1.00***\\(0.01)} & \specialcell{-0.36**\\(0.03)}\\
+\hspace{1em}Women’s education (Total Years) & \specialcell{14.29\\(2.41)} & \specialcell{13.87\\(2.47)} & \specialcell{13.42\\(2.35)} & \specialcell{13.27\\(2.37)} & \specialcell{-1.01***\\(0.01)} & \specialcell{-0.46**\\(0.03)}\\
 \textbf{Panel C: Employment and Earnings} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
-\hspace{1em}Men’s Employment Rate & \specialcell{0.95\\(0.22)} & \specialcell{0.94\\(0.23)} & \specialcell{0.92\\(0.26)} & \specialcell{0.93\\(0.26)} & \specialcell{-0.02\\(0.00)} & \specialcell{-0.02\\(0.00)}\\
-\hspace{1em}Women’s Employment Rate & \specialcell{0.96\\(0.2)} & \specialcell{0.95\\(0.22)} & \specialcell{0.93\\(0.25)} & \specialcell{0.94\\(0.24)} & \specialcell{-0.02\\(0.00)} & \specialcell{-0.02\\(0.00)}\\
+\hspace{1em}Men’s Employment Rate & \specialcell{0.95\\(0.22)} & \specialcell{0.94\\(0.23)} & \specialcell{0.92\\(0.26)} & \specialcell{0.93\\(0.26)} & \specialcell{-0.02***\\(0.00)} & \specialcell{-0.02***\\(0.00)}\\
+\hspace{1em}Women’s Employment Rate & \specialcell{0.96\\(0.2)} & \specialcell{0.95\\(0.22)} & \specialcell{0.93\\(0.25)} & \specialcell{0.94\\(0.24)} & \specialcell{-0.02***\\(0.00)} & \specialcell{-0.02***\\(0.00)}\\
 \addlinespace
-\hspace{1em}Men’s Log Hourly Earnings & \specialcell{2.48\\(0.45)} & \specialcell{2.41\\(0.45)} & \specialcell{2.4\\(0.44)} & \specialcell{2.41\\(0.43)} & \specialcell{-0.07\\(0.01)} & \specialcell{-0.00\\(0.02)}\\
-\hspace{1em}Women’s Hourly Earnings & \specialcell{2.33\\(0.49)} & \specialcell{2.33\\(0.46)} & \specialcell{2.27\\(0.45)} & \specialcell{2.3\\(0.41)} & \specialcell{-0.02\\(0.01)} & \specialcell{-0.06\\(0.02)}\\
-\hspace{1em}Men’s Log Annual Earnings & \specialcell{10.25\\(1.01)} & \specialcell{10.08\\(1.05)} & \specialcell{10.04\\(1.06)} & \specialcell{10.01\\(1.04)} & \specialcell{-0.25\\(0.01)} & \specialcell{-0.04\\(0.03)}\\
-\hspace{1em}Women’s Hourly Earnings & \specialcell{9.46\\(1.78)} & \specialcell{9.54\\(1.64)} & \specialcell{9.47\\(1.57)} & \specialcell{9.53\\(1.52)} & \specialcell{0.07\\(0.02)} & \specialcell{-0.07\\(0.06)}\\
+\hspace{1em}Men’s Log Hourly Earnings & \specialcell{2.48\\(0.45)} & \specialcell{2.41\\(0.45)} & \specialcell{2.4\\(0.44)} & \specialcell{2.41\\(0.43)} & \specialcell{-0.07***\\(0.01)} & \specialcell{-0.00**\\(0.02)}\\
+\hspace{1em}Women’s Hourly Earnings & \specialcell{2.33\\(0.49)} & \specialcell{2.33\\(0.46)} & \specialcell{2.27\\(0.45)} & \specialcell{2.3\\(0.41)} & \specialcell{-0.02***\\(0.01)} & \specialcell{-0.06**\\(0.02)}\\
+\hspace{1em}Men’s Log Annual Earnings & \specialcell{10.25\\(1.01)} & \specialcell{10.08\\(1.05)} & \specialcell{10.04\\(1.06)} & \specialcell{10.01\\(1.04)} & \specialcell{-0.25**\\(0.01)} & \specialcell{-0.04**\\(0.03)}\\
+\hspace{1em}Women’s Hourly Earnings & \specialcell{9.46\\(1.78)} & \specialcell{9.54\\(1.64)} & \specialcell{9.47\\(1.57)} & \specialcell{9.53\\(1.52)} & \specialcell{0.07**\\(0.02)} & \specialcell{-0.07*\\(0.06)}\\
 \textbf{Panel D: Hispanic Identity} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\

### 2023-04-12 my_paper/tables/tab04-summary-stats.tex
add sig levels to sum tables and hilight defense columns

@@
 \textbf{Panel A: Parent’s} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
-\hspace{1em}Husband’seducation (Total Years) & \specialcell{13.05\\(2.44)} & \specialcell{12.32\\(3.33)} & \specialcell{10.65\\(4.39)} & \specialcell{8.93\\(4.41)} & \specialcell{-4.11\\(0.02)} & \specialcell{-1.67\\(0.04)}\\
-\hspace{1em}Wife’seducation (Total Years) & \specialcell{12.74\\(2.12)} & \specialcell{11.03\\(3.92)} & \specialcell{11.54\\(3.12)} & \specialcell{8.6\\(4.13)} & \specialcell{-4.13\\(0.02)} & \specialcell{0.51\\(0.04)}\\
-\hspace{1em}Total Household seducation (Total Years) & \specialcell{25.78\\(4.08)} & \specialcell{23.35\\(6.51)} & \specialcell{22.19\\(6.69)} & \specialcell{17.54\\(7.83)} & \specialcell{-8.25\\(0.03)} & \specialcell{-1.16\\(0.07)}\\
+\hspace{1em}Husband’seducation (Total Years) & \specialcell{13.05\\(2.44)} & \specialcell{12.32\\(3.33)} & \specialcell{10.65\\(4.39)} & \specialcell{8.93\\(4.41)} & \specialcell{-4.11**\\(0.02)} & \specialcell{-1.67**\\(0.04)}\\
+\hspace{1em}Wife’seducation (Total Years) & \specialcell{12.74\\(2.12)} & \specialcell{11.03\\(3.92)} & \specialcell{11.54\\(3.12)} & \specialcell{8.6\\(4.13)} & \specialcell{-4.13**\\(0.02)} & \specialcell{0.51**\\(0.04)}\\
+\hspace{1em}Total Household seducation (Total Years) & \specialcell{25.78\\(4.08)} & \specialcell{23.35\\(6.51)} & \specialcell{22.19\\(6.69)} & \specialcell{17.54\\(7.83)} & \specialcell{-8.25**\\(0.03)} & \specialcell{-1.16*\\(0.07)}\\
 \textbf{Panel B: Education} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
 \addlinespace
-\hspace{1em}Men’s education (Total Years) & \specialcell{12.97\\(2.15)} & \specialcell{13.45\\(2.37)} & \specialcell{13.13\\(2.27)} & \specialcell{12.89\\(2.25)} & \specialcell{-0.08\\(0.01)} & \specialcell{-0.32\\(0.03)}\\
-\hspace{1em}Women’s education (Total Years) & \specialcell{13.23\\(2.25)} & \specialcell{13.75\\(2.41)} & \specialcell{13.32\\(2.34)} & \specialcell{13.26\\(2.37)} & \specialcell{0.03\\(0.01)} & \specialcell{-0.43\\(0.03)}\\
+\hspace{1em}Men’s education (Total Years) & \specialcell{12.97\\(2.15)} & \specialcell{13.45\\(2.37)} & \specialcell{13.13\\(2.27)} & \specialcell{12.89\\(2.25)} & \specialcell{-0.08**\\(0.01)} & \specialcell{-0.32**\\(0.03)}\\
+\hspace{1em}Women’s education (Total Years) & \specialcell{13.23\\(2.25)} & \specialcell{13.75\\(2.41)} & \specialcell{13.32\\(2.34)} & \specialcell{13.26\\(2.37)} & \specialcell{0.03**\\(0.01)} & \specialcell{-0.43**\\(0.03)}\\
 \textbf{Panel C: Employment and Earnings} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
-\hspace{1em}Men’s Employment Rate & \specialcell{0.93\\(0.26)} & \specialcell{0.94\\(0.23)} & \specialcell{0.92\\(0.27)} & \specialcell{0.93\\(0.26)} & \specialcell{0.00\\(0.00)} & \specialcell{-0.02\\(0.00)}\\
-\hspace{1em}Women’s Employment Rate & \specialcell{0.94\\(0.25)} & \specialcell{0.94\\(0.23)} & \specialcell{0.93\\(0.26)} & \specialcell{0.94\\(0.24)} & \specialcell{0.00\\(0.00)} & \specialcell{-0.02\\(0.00)}\\
+\hspace{1em}Men’s Employment Rate & \specialcell{0.93\\(0.26)} & \specialcell{0.94\\(0.23)} & \specialcell{0.92\\(0.27)} & \specialcell{0.93\\(0.26)} & \specialcell{0.00***\\(0.00)} & \specialcell{-0.02***\\(0.00)}\\
+\hspace{1em}Women’s Employment Rate & \specialcell{0.94\\(0.25)} & \specialcell{0.94\\(0.23)} & \specialcell{0.93\\(0.26)} & \specialcell{0.94\\(0.24)} & \specialcell{0.00***\\(0.00)} & \specialcell{-0.02***\\(0.00)}\\
 \addlinespace
-\hspace{1em}Men’s Log Hourly Earnings & \specialcell{2.4\\(0.44)} & \specialcell{2.41\\(0.45)} & \specialcell{2.4\\(0.44)} & \specialcell{2.41\\(0.43)} & \specialcell{0.01\\(0.01)} & \specialcell{-0.01\\(0.02)}\\
-\hspace{1em}Women’s Hourly Earnings & \specialcell{2.26\\(0.43)} & \specialcell{2.32\\(0.45)} & \specialcell{2.27\\(0.45)} & \specialcell{2.3\\(0.41)} & \specialcell{0.04\\(0.01)} & \specialcell{-0.05\\(0.02)}\\
-\hspace{1em}Men’s Log Annual Earnings & \specialcell{10.02\\(1.02)} & \specialcell{10.06\\(1.06)} & \specialcell{10.03\\(0.99)} & \specialcell{10\\(1.04)} & \specialcell{-0.02\\(0.01)} & \specialcell{-0.03\\(0.04)}\\
-\hspace{1em}Women’s Hourly Earnings & \specialcell{9.44\\(1.59)} & \specialcell{9.55\\(1.59)} & \specialcell{9.47\\(1.55)} & \specialcell{9.52\\(1.52)} & \specialcell{0.08\\(0.02)} & \specialcell{-0.08\\(0.05)}\\
+\hspace{1em}Men’s Log Hourly Earnings & \specialcell{2.4\\(0.44)} & \specialcell{2.41\\(0.45)} & \specialcell{2.4\\(0.44)} & \specialcell{2.41\\(0.43)} & \specialcell{0.01***\\(0.01)} & \specialcell{-0.01**\\(0.02)}\\
+\hspace{1em}Women’s Hourly Earnings & \specialcell{2.26\\(0.43)} & \specialcell{2.32\\(0.45)} & \specialcell{2.27\\(0.45)} & \specialcell{2.3\\(0.41)} & \specialcell{0.04***\\(0.01)} & \specialcell{-0.05**\\(0.02)}\\
+\hspace{1em}Men’s Log Annual Earnings & \specialcell{10.02\\(1.02)} & \specialcell{10.06\\(1.06)} & \specialcell{10.03\\(0.99)} & \specialcell{10\\(1.04)} & \specialcell{-0.02**\\(0.01)} & \specialcell{-0.03**\\(0.04)}\\
+\hspace{1em}Women’s Hourly Earnings & \specialcell{9.44\\(1.59)} & \specialcell{9.55\\(1.59)} & \specialcell{9.47\\(1.55)} & \specialcell{9.52\\(1.52)} & \specialcell{0.08**\\(0.02)} & \specialcell{-0.08*\\(0.05)}\\
 \bottomrule

### 2023-04-12 programs/12-table-four-summary-stats.R

@@ -74,6 +74,7 @@ row3 <- c(
   ),
   paste0("\\specialcell{", 
          round(differences1[[1]], digits = 2), 
+         symnum(pvalues1[[1]], corr = FALSE, na = FALSE, cutpoints = c(0, 0.01, 0.05, 0.1), symbols = c("***", "**", "*")),
          "\\\\",
          "(",
          round(pvalues1[[1]], digits = 2),
@@ -81,6 +82,7 @@ row3 <- c(
   ),
   paste0("\\specialcell{", 
          round(differences2[[1]], digits = 2), 
+         symnum(pvalues2[[1]], corr = FALSE, na = FALSE, cutpoints = c(0, 0.01, 0.05, 0.1), symbols = c("***", "**", "*")),
          "\\\\",
          "(",
          round(pvalues2[[1]], digits = 2),
@@ -138,6 +140,7 @@ row4 <- c(
   ),
   paste0("\\specialcell{", 
          round(differences1[[1]], digits = 2), 
+         symnum(pvalues1[[1]], corr = FALSE, na = FALSE, cutpoints = c(0, 0.01, 0.05, 0.1), symbols = c("***", "**", "*")),
          "\\\\",
          "(",
          round(pvalues1[[1]], digits = 2),
@@ -145,6 +148,7 @@ row4 <- c(
   ),
   paste0("\\specialcell{", 
          round(differences2[[1]], digits = 2), 
+         symnum(pvalues2[[1]], corr = FALSE, na = FALSE, cutpoints = c(0, 0.01, 0.05, 0.1), symbols = c("***", "**", "*")),
          "\\\\",
          "(",
          round(pvalues2[[1]], digits = 2),
@@ -202,6 +206,7 @@ row5 <- c(
   ),
   paste0("\\specialcell{", 
          round(differences1[[1]], digits = 2), 
+         symnum(pvalues1[[1]], corr = FALSE, na = FALSE, cutpoints = c(0, 0.01, 0.05, 0.1), symbols = c("***", "**", "*")),
          "\\\\",
          "(",
          round(pvalues1[[1]], digits = 2),
@@ -209,6 +214,7 @@ row5 <- c(
   ),
   paste0("\\specialcell{", 
          round(differences2[[1]], digits = 2), 
+         symnum(pvalues2[[1]], corr = FALSE, na = FALSE, cutpoints = c(0, 0.01, 0.05, 0.1), symbols = c("***", "**", "*")),
[…]

### 2023-04-12 programs/13-table-five-regression.R

@@ -148,3 +148,41 @@ regression_tab <- modelsummary(reg1, fmt = 2,
 
 regression_tab %>%
   save_kable(file.path(dissertation_wd,"tables/tab05-regression.tex"))
+
+# hilight column 1
+regression_tab <- modelsummary(reg1, fmt = 2,
+                               output = "latex",
+                               coef_map = cm,
+                               add_rows = all_row,
+                               gof_map = gm,
+                               escape = F,
+                               #gof_omit = 'DF|Deviance|R2|AIC|BIC|Log.Lik.|F|Std.Errors',
+                               stars= c('***' = 0.01, '**' = 0.05, '*' = 0.1)) %>%
+  kable_styling(bootstrap_options = c("hover", "condensed"), 
+                latex_options = c("scale_down", "hold_position")
+  ) |> 
+  row_spec(8, hline_after = T) |> 
+  column_spec(2, background = "#ffff30")
+
+
+regression_tab %>%
+  save_kable(file.path(dissertation_wd,"tables/tab05a-regression.tex"))
+
+# hilight column 2
+regression_tab <- modelsummary(reg1, fmt = 2,
+                               output = "latex",
+                               coef_map = cm,
+                               add_rows = all_row,
+                               gof_map = gm,
+                               escape = F,
+                               #gof_omit = 'DF|Deviance|R2|AIC|BIC|Log.Lik.|F|Std.Errors',
+                               stars= c('***' = 0.01, '**' = 0.05, '*' = 0.1)) %>%
+  kable_styling(bootstrap_options = c("hover", "condensed"), 
+                latex_options = c("scale_down", "hold_position")
+  ) |> 
+  row_spec(8, hline_after = T) |> 
+  column_spec(3, background = "#ffff30")
+
+
+regression_tab %>%
+  save_kable(file.path(dissertation_wd,"tables/tab05b-regression.tex"))

### 2023-04-12 programs/14-table-six-regression.R

@@ -54,7 +54,7 @@ row_diff <- rbind(differences_row, pvalue_row)
 colnames(row_diff)<-LETTERS[1:3]
 
 row_diff <- as.data.frame(row_diff)
-attr(row_diff, 'position') <- c(13:14)
+attr(row_diff, 'position') <- c(7:8)
 
 controling_for <-  c("\\textit{Controlling for:}", " ", "")
 dim(controling_for) <- c(1,3)
@@ -66,14 +66,14 @@ controling_for <- as.data.frame(controling_for)
 hoursworked <- as.data.frame(hoursworked)
 
 all_row <- rbind(differences_row, pvalue_row, controling_for, hoursworked)
-attr(all_row, 'position') <- c(13:14, 16:17)
+attr(all_row, 'position') <- c(7:8, 10:11)
 
 cm <- c("WH:Hispanic_ID" = "$WH_{i} \\times Hispanic_{i}$",
         "Hispanic_ID:HW" = "$HW_{i} \\times Hispanic_{i}$",
-        "Hispanic_ID:HH" = "$HH_{i} \\times Hispanic_{i}$",
-        "WH" = "$WH_{i}$",
-        "HW" = "$HW_{i}$",
-        "HH" = "$HH_{i}$"
+        "Hispanic_ID:HH" = "$HH_{i} \\times Hispanic_{i}$"#,
+        # "WH" = "$WH_{i}$",
+        # "HW" = "$HW_{i}$",
+        # "HH" = "$HH_{i}$"
 ) 
 gm <- tibble::tribble(
   ~raw,        ~clean,          ~fmt,
@@ -125,7 +125,7 @@ regression_tab <- modelsummary(reg1, fmt = 2,
   footnote_as_chunk = F, title_format = c("italic"),
   escape = F, threeparttable = T
   ) |> 
-  row_spec(14, hline_after = T)
+  row_spec(8, hline_after = T)
 
 
 regression_tab %>%
@@ -146,9 +146,48 @@ regression_tab <- modelsummary(reg1, fmt = 2,
   kable_styling(bootstrap_options = c("hover", "condensed"), 
                 latex_options = c("scale_down", "hold_position")
   ) |> 
-  row_spec(14, hline_after = T)
+  row_spec(8, hline_after = T)
 
 
 regression_tab %>%
   save_kable(file.path(dissertation_wd,"tables/tab06-regression.tex"))
+
+# highlight collumn 2
+regression_tab <- modelsummary(reg1, fmt = 2,
+                               output = "latex",
[…]

### 2023-04-12 my_paper/Hadah-last-names.tex
some edits

@@
 Finally, I find that men identifying as Hispanic earn significantly less than those who do not, even when I control ancestry and education.
+Inter-ethnic children with a White father, and thus having a non-Hispanic last name, earn 14\% less than children of endogamous White parents.
+Inter-ethnic children with a Hispanic father, and thus having a Hispanic last name, earn 20\% less than children of endogamous White parents.
+Endogamous children with Hispanic parents earn 24\% less than children of endogamous White parents.
+A big part of the gaps could be explained by education.
+Inter-ethnic children with a White father and identify as Hispanic, and thus having a non-Hispanic last name, earn 14\% less than children of endogamous White parents.
+Inter-ethnic children with a Hispanic father and identify as Hispanic, and thus having a Hispanic last name, earn 18\% less than children of endogamous White parents.
+Endogamous children with Hispanic parents and identify as Hispanic earn 10\% less than children of endogamous White parents.
 
@@
 Overall, HH children do worse than WW children.
-HH has one less year of education compared to WW, 2.1 percentage points less likely to be employed, HH men earn 16.4\% less than WW men, and HH women make 12.3\% less than WW women.
+HH has one less year of education compared to WW, 2.1 percentage points less likely to be employed, HH men earn 18\% less than WW men, and HH women make 12.3\% less than WW women.
 Therefore, these two groups, HH and WW, are different in many observable.
@@
 
-Let $Y_{it}$ be the annual log earnings of person $i$ at time $t$.
-$WH_{i}$, $HW_{i}$ and $HH_{i}$ are indicator variables for the type of parents person $i$ has.
-$X_{it}$ is a vector of controls, $\gamma_{t}$ are year fixed effects and $\phi_{it}$ represents the error term.
+Let $Y_{ist}$ be the annual log earnings of person $i$ at time $t$.
+$WH_{ist}$, $HW_{ist}$ and $HH_{ist}$ are indicator variables for the type of parents person $i$ has.
+$X_{ist}$ is a vector of controls, $\gamma_{t}$ are year fixed effects, $\lambda_{s}$ are state fixed effects, and $\phi_{ist}$ represents the error term.
 The equation for this strategy is written as follows:
 \begin{equation} \label{eq:1a}
-Y_{it} = \beta_{1} WH_{i} +\beta_{2} HW_{i} + \beta_{3} HH_{i} + X_{it} \pi + \gamma_{t}+\phi_{it}
+Y_{ist} = \beta_{1} WH_{ist} +\beta_{2} HW_{ist} + \beta_{3} HH_{ist} + X_{ist} \pi + \gamma_{t} + \lambda_s + \phi_{ist}
 \end{equation}
@@
 
-Let $Y_{it}$ be the annual log earnings of person $i$ at time $t$.
-$WH_{i}$, $HW_{i}$ and $WW_{i}$ are indicator variables for the type of parents person $i$ has.
-$X_{it}$ is a vector of controls, $\gamma_{t}$ are year fixed effects and $\phi_{it}$ represents the error term.
-The addition to the previous model is the $Hispanic_{i}$ indicator variable.
-$Hispanic_{i}$ is equal to one if a person identifies as Hispanic and zero otherwise.
-I interact this term with the $WH_{i}$, $HW_{i}$ and $HH_{i}$.
+Let $Y_{ist}$ be the annual log earnings of person $i$ in state $s$ at time $t$.
+$WH_{ist}$, $HW_{ist}$ and $WW_{ist}$ are indicator variables for the type of parents person $i$ has.
+$X_{ist}$ is a vector of controls, $\gamma_{t}$ are year fixed effects, $\lambda_s$ are state fixed effects, and $\phi_{ist}$ represents the error term.
+The addition to the previous model is the $Hispanic_{ist}$ indicator variable.
+$Hispanic_{ist}$ is equal to one if a person identifies as Hispanic and zero otherwise.
+I interact this term with the $WH_{ist}$, $HW_{ist}$ and $HH_{ist}$.
 The regression is presented in the following equation.
@@
 \begin{align} \label{eq:iden}
-Y_{it} &= \beta_{1} HW_{i} +  \beta_{2} WH_{i} + \beta_{3} HH_{i} +\\
-& \beta_{4} WH_{i} \cdot Hispanic_{i} + \beta_{5} HW_{i}\cdot Hispanic_{i} +  \beta_{6} HH_{i} \cdot Hispanic_{i}+ \nonumber \\
-&X_{it} \pi + \gamma_{t}+\phi_{it} \nonumber
+Y_{ist} &= \beta_{1} HW_{ist} +  \beta_{2} WH_{ist} + \beta_{3} HH_{ist} +\\
+& \beta_{4} WH_{ist} \cdot Hispanic_{ist} + \beta_{5} HW_{ist}\cdot Hispanic_{ist} +  \beta_{6} HH_{ist} \cdot Hispanic_{ist}+ \nonumber \\
+&X_{ist} \pi + \gamma_{t} + \lambda_s + \phi_{ist} \nonumber
 \end{align}
@@
 The coefficients of interest from equation \ref{eq:iden} are $\beta_{4}$, $\beta_{5}$ and $\beta_{6}$.
-The interaction terms capture the earnings gaps between members of $WH_{i}$, $HW_{i}$, and $WW_{i}$ that identify as Hispanic and White WW children.
+The interaction terms capture the earnings gaps between members of $WH_{ist}$, $HW_{ist}$, and $WW_{ist}$ that identify as Hispanic and White WW children.
 
@@
 
-I find that the earnings gaps between WH and WW are 8.8\%, 14.1\% between HW and WW, and 16.1\% between HH and WW.
-Moreover, the difference in earnings between HW and WH children, that captures the effect of having a Hispanic last name, equals a statistically significant 5.3 percentage points.
-Meaning by comparing inter-ethnic children, a person with a Hispanic last name earns 5.3 percentage points less than someone with a White last name.
+I find that the earnings gaps between WH and WW are 14\%, 18\% between HW and WW, and 24\% between HH and WW.
+Moreover, the difference in earnings between HW and WH children, that captures the effect of having a Hispanic last name, equals a statistically significant 6 percentage points.
+Meaning by comparing inter-ethnic children, a person with a Hispanic last name earns 6 percentage points less than someone with a White last name.
 However, more than half of the earnings gaps could be explained by educational differences between WW children and WH, HW, and HH.
 When I control for education, the gap between the three groups and WW is cut between by 50\% to 64.0\%.
-Furthermore, the last name effect decreases to a statistically insignificant 1.6 percentage points earnings gap.
+Furthermore, the last name effect decreases to a statistically insignificant 2 percentage points earnings gap.
 
 I also find a significant earnings gap between those that identify as Hispanic and WW children.
-The gaps between WH, HW, and HH that identify as Hispanic and WW are, respectively, 13.4\%, 16.4\%, and 10.6\%.
+The gaps between WH, HW, and HH that identify as Hispanic and WW are, respectively, 14\%, 18\%, and 10\%.
 The difference between WH and HW that identify as Hispanic---captures the last name effect of those that identify as Hispanic--- is equal to 3 percentage points and is statistically insignificant.
 Controlling for education explains a big part of the gap.
-The gaps between WH, HW, and HH that identify as Hispanic and WW after controlling for education are 6.4\%, 10.8\%, and 6.2\%.
+The gaps between WH, HW, and HH that identify as Hispanic and WW after controlling for education are 8\%, 12\%, and 6\%.
 The WH--WW and HH--WW gaps for those identifying as Hispanic become statistically insignificant after controlling for education.
-Finally, the difference between WH and HW that identify as Hispanic that identifies as Hispanic remains statistically insignificant but increases to 4.3 percentage points.
+Finally, the difference between WH and HW that identify as Hispanic that identifies as Hispanic remains statistically insignificant but increases to 4 percentage points.
 
@@
 Overall, there is big Hispanic--White earnings gaps.
-Children of White father-Hispanic mothers (WH) earn 8.8\% less than children of White fathers-White mothers (WW) families.
-The gap between children of Hispanic father-White mothers (HW) and WW children is 14.1\%.
-The gap between children of Hispanic father-Hispanic mothers (HH) and WW children is 16.1\%.
+Children of White father-Hispanic mothers (WH) earn 14\% less than children of White fathers-White mothers (WW) families.
+The gap between children of Hispanic father-White mothers (HW) and WW children is 20\%.
+The gap between children of Hispanic father-Hispanic mothers (HH) and WW children is 24\%.
 Therefore, Hispanic children, inter-ethnic or children of homogeneous HH marriages, face a substantial earnings gap.
@@
 A WH child will have a White last name since their father is White, while an HW child will have a Hispanic last name like their Hispanic father.
-When I take the difference between HW and WH, I find a statistically significant gap of 5.3 percentage points, which means that an inter-ethnic person with a Hispanic last name earns 5.3 percentage points less than a similar inter-ethnic person but with a White last name.
+When I take the difference between HW and WH, I find a statistically significant gap of 6 percentage points, which means that an inter-ethnic person with a Hispanic last name earns 6 percentage points less than a similar inter-ethnic person but with a White last name.
 
@@
 The sample includes full-time, full-year, and waged and salaried men, and I control for hours worked and age and include year fixed effects (FE).
-The gap between WH and WW is 4.4\%, a reduction of 50\% from the baseline after controlling for education.
-The earning gaps between HW and HH, and WW were respectively reduced by 58\% and 64\%.
-After controlling for education, an HW man earns 5.9\% less than a WW man, and an HH man earns 5.8\% less.
-The effect of having a Hispanic last name on earnings decreased from 5.3 to 1.6 percentage points and became statistically insignificant after controlling for education.
+The gap between WH and WW is 9\%, a reduction of 35\% from the baseline after controlling for education.
+The earning gaps between HW and HH, and WW were respectively reduced by 45\% and 50\%.
+After controlling for education, an HW man earns 11\% less than a WW man, and an HH man earns 12\% less.
+The effect of having a Hispanic last name on earnings decreased from 6 to 2 percentage points and became statistically insignificant after controlling for education.
 
@@
 Having a Hispanic last name carries a cost in the labor market.
-A person with a Hispanic last name earns 5.3 percentage points less than someone with a White name.
+A person with a Hispanic last name earns 6 percentage points less than someone with a White name.
 However, educational differences between Hispanics and Whites could explain more than 50\% of the earning gaps and all of the Hispanic last name costs.
@@
 I find that those that identify as Hispanic face a big earnings gap compared to those that do not.
-A WH man that identifies as Hispanic earns 13.4\% less than a WW man.
-An HW man who identifies as Hispanic makes 16.4\% less than a WW man, and an HH man who is determined as Hispanic earns 10.6\% less than a WW man.
-Comparing the coefficients of the interaction terms to those of the $WH_{i}$, $HW_{i}$, and $HH_{i}$ shows that members of the three groups that do not identify as Hispanic earn about the same as WW.
-Moreover, the difference between $HW_{i} \cdot Hispanic_{i}$  and $WH_{i} \cdot Hispanic_{i}$ captures the effect of having a Hispanic last name and identifying as Hispanic compared to individuals who identify as Hispanic but have a White last name.
+A WH man that identifies as Hispanic earns 14\% less than a WW man.
+An HW man who identifies as Hispanic makes 18\% less than a WW man, and an HH man who is determined as Hispanic earns 10\% less than a WW man.
+Comparing the coefficients of the interaction terms to those of the $WH_{ist}$, $HW_{ist}$, and $HH_{ist}$ shows that members of the three groups that do not identify as Hispanic earn about the same as WW.
+Moreover, the difference between $HW_{ist} \cdot Hispanic_{ist}$  and $WH_{ist} \cdot Hispanic_{ist}$ captures the effect of having a Hispanic last name and identifying as Hispanic compared to individuals who identify as Hispanic but have a White last name.
 A person with a Hispanic last name that identifies as Hispanic earns 3 percentage points less than a person that identifies as Hispanic but has a White last name.
@@
 Education explains all the identity costs for members of WH and HH.
-Also, controlling for education decreases the earnings gap for identifying as Hispanic and being a member of HW by 34.1\%.
-An HW member that identifies as Hispanic earns 10.8\% less than a WW man.
-Furthermore, the difference between $HW_{i} \cdot Hispanic_{i}$  and $WH_{i} \cdot Hispanic_{i}$ increases to 4.3 percentage points but remains statistically insignificant.
+Also, controlling for education decreases the earnings gap for identifying as Hispanic and being a member of HW by 33\%.
+An HW member that identifies as Hispanic earns 12\% less than a WW man.
+Furthermore, the difference between $HW_{ist} \cdot Hispanic_{ist}$  and $WH_{ist} \cdot Hispanic_{ist}$ increases to 4 percentage points but remains statistically insignificant.
 
@@
  \begin{equation} \label{eq:naive}
- Y_{it} = \beta_{1} Hispanic_{i} + X_{it}^{\prime} + \gamma_{t}+\phi_{it}
+ Y_{ist} = \beta_{1} Hispanic_{ist} + X_{ist}^{\prime} + \gamma_{t} + \lambda_s + \phi_{ist}
  \end{equation}
 
-Where $Y_{it}$ in equation \ref{eq:naive} is the annual log earnings of person $i$ at time $t$.
-$Hispanic_{i}$ is a dummy variable equal to one if person $i$ is Hispanic and zero otherwise.
-$X_{it}$ is a vector of controls, $\gamma_{t}$ is time fixed effects and $\phi_{it}$ is the error term.
+Where $Y_{ist}$ in equation \ref{eq:naive} is the annual log earnings of person $i$ at time $t$.
+$Hispanic_{ist}$ is a dummy variable equal to one if person $i$ is Hispanic and zero otherwise.
+$X_{ist}$ is a vector of controls, $\gamma_{t}$ is time fixed effects and $\phi_{ist}$ is the error term.
 
@@
 I compare the children of inter-ethnic marriages to study the labor market impact of having a Hispanic last name.
-The earning gap between WH and WW is 8.8\%, 14.1\% between HW and WW, and 16.1\% between HH and WW.
-When I compare the earnings of HW and WH children, which captures the effect of having a Hispanic last name, HW children earn 5.3\% less than WH children.
-Thus, by comparing inter-ethnic children, a person with a Hispanic last name makes 5.3\% less than someone with a White last name.
+The earning gap between WH and WW is 14\%, 20\% between HW and WW, and 24\% between HH and WW.
+When I compare the earnings of HW and WH children, which captures the effect of having a Hispanic last name, HW children earn  6\% less than WH children.
+Thus, by comparing inter-ethnic children, a person with a Hispanic last name makes 6\% less than someone with a White last name.
 Controlling for education, however, explains more than half of the earnings gaps between WW children and WH, HW, and HH.
 When I control for education, the gap between the three groups and WW is cut between 50-64.0\%.
-Furthermore, the last name effect decreases to a statistically insignificant 1.6\% earnings gap.
+Furthermore, the last name effect decreases to a statistically insignificant 2\% earnings gap.
 
 I also find that men who identify as Hispanic earn significantly less than those who do not, even when I control ancestry and education.
-The gaps between WH, HW, and HH that identify as Hispanic and WW are, respectively, 13.4\%, 16.4\%, and 10.6\%.
+The gaps between WH, HW, and HH that identify as Hispanic and WW are, respectively, 14\%, 18\%, and 10\%.
 Controlling for education explains a big part of the gap.
-The gaps between WH, HW, and HH that identify as Hispanic and WW after controlling for education are 6.4\%, 10.8\%, and 6.2\%.
+The gaps between WH, HW, and HH that identify as Hispanic and WW after controlling for education are 8\%, 12\%, and 6\%.
 
@@
 
-% $WH_{i}$ &  & -0.088{*}{*}{*} &  & -0.044{*}{*}{*} & \tabularnewline
+% $WH_{ist}$ &  & -0.088{*}{*}{*} &  & -0.044{*}{*}{*} & \tabularnewline
 %  &  & (0.018) &  & (0.017) & \tabularnewline
 %  &  &  &  &  & \tabularnewline
-% $HW_{i}$ &  & -0.141{*}{*}{*} &  & -0.059{*}{*}{*} & \tabularnewline
+% $HW_{ist}$ &  & -0.141{*}{*}{*} &  & -0.059{*}{*}{*} & \tabularnewline
 %  &  & (0.015) &  & (0.014) & \tabularnewline
 %  &  &  &  &  & \tabularnewline
-% $HH_{i}$ &  & -0.161{*}{*}{*} &  & -0.058{*}{*}{*} & \tabularnewline
+% $HH_{ist}$ &  & -0.161{*}{*}{*} &  & -0.058{*}{*}{*} & \tabularnewline
 %  &  & (0.008) &  & (0.007) & \tabularnewline
@@
 % \midrule
-% $HW_{i}$ - $WH_{i}$ &  & -0.053{*}{*} &  & -0.016 & \tabularnewline
+% $HW_{ist}$ - $WH_{ist}$ &  & -0.053{*}{*} &  & -0.016 & \tabularnewline
 %  &  & (0.024) &  & (0.022) & \tabularnewline
@@
 %  &  &  &  &  & \tabularnewline
-% $WH_{i}\times Hispanic_{i}$ &  & -0.134{*}{*} &  & -0.064 & \tabularnewline
+% $WH_{ist}\times Hispanic_{ist}$ &  & -0.134{*}{*} &  & -0.064 & \tabularnewline
 %  &  & (0.053) &  & (0.049) & \tabularnewline
-% $HW_{i}\times Hispanic_{i}$ &  & -0.164{*}{*}{*} &  & -0.108{*}{*} & \tabularnewline
+% $HW_{ist}\times Hispanic_{ist}$ &  & -0.164{*}{*}{*} &  & -0.108{*}{*} & \tabularnewline
 %  &  & (0.052) &  & (0.048) & \tabularnewline
-% $HH_{i}\times Hispanic_{i}$ &  & -0.106{*}{*} &  & -0.062 & \tabularnewline
+% $HH_{ist}\times Hispanic_{ist}$ &  & -0.106{*}{*} &  & -0.062 & \tabularnewline
 %  &  & (0.049) &  & (0.045) & \tabularnewline
-% $WH_{i}$ &  & 0.027 &  & 0.011 & \tabularnewline
+% $WH_{ist}$ &  & 0.027 &  & 0.011 & \tabularnewline
 %  &  & (0.048) &  & (0.045) & \tabularnewline
-% $HW_{i}$ &  & 0.007 &  & 0.038 & \tabularnewline
+% $HW_{ist}$ &  & 0.007 &  & 0.038 & \tabularnewline
 %  &  & (0.050) &  & (0.046) & \tabularnewline
-% $HH_{i}$ &  & -0.057 &  & 0.003 & \tabularnewline
+% $HH_{ist}$ &  & -0.057 &  & 0.003 & \tabularnewline
 %  &  & (0.057) &  & (0.044) & \tabularnewline
 % \hline 
-% $HW_{i}\times Hispanic_{i}$ - $WH_{i}\times Hispanic_{i}$ &  & -0.030 &  & -0.043 & \tabularnewline
+% $HW_{ist}\times Hispanic_{ist}$ - $WH_{ist}\times Hispanic_{ist}$ &  & -0.030 &  & -0.043 & \tabularnewline
 %  &  & (0.074) &  & (0.068) & \tabularnewline
@@
 %    \item [\textit{Notes.}] The four groups stand for White Husband White Wife (WW), White Husband Hispanic Wife (WH), Hispanic Husband White (HW), and Hispanic Husband Hispanic Wife (HH).
-$Hispanic_{i}$ is a dummy variable equal to one if a person identifies as Hispanic and zero otherwise.
+$Hispanic_{ist}$ is a dummy variable equal to one if a person identifies as Hispanic and zero otherwise.
 %    \item [\textit{Notes.}] The sample is restricted to men working full-time full-year and are waged and salaried workers.
@@
 % \hline 
-% $Hispanic_{i}$ &  & -0.156{*}{*}{*} &  & -0.060{*}{*}{*} & \tabularnewline
+% $Hispanic_{ist}$ &  & -0.156{*}{*}{*} &  & -0.060{*}{*}{*} & \tabularnewline
 %  &  & (0.006) &  & (0.006) & \tabularnewline
@@
 %    \item [\textit{Notes.}] This table includes the estimation results of equation \ref{eq:naive}.
-%    \item [\textit{Notes.}] $Hispanic_{i}$ is a dummy variable equal to one if a person identifies as Hispanic and zero otherwise.
+%    \item [\textit{Notes.}] $Hispanic_{ist}$ is a dummy variable equal to one if a person identifies as Hispanic and zero otherwise.
 %    \item [\textit{Notes.}] The sample is restricted to men working full time full-year and are waged and salaried workers.

### 2023-04-12 programs/13-table-five-regression.R

@@ -11,10 +11,10 @@ ParentDummies <- c("WH", "HW", "HH", "uhrsworkly")
               
 # By generation
 reg1 <- list(
-  "\\specialcell{(1) \\\\ Log annual earnings}" = feols(lninctot_1999 ~ .[ParentDummies] | year + age, se = "iid",
+  "\\specialcell{(1) \\\\ Log annual earnings}" = feols(lninctot_1999 ~ .[ParentDummies] | year + age + statefip, vcov = ~statefip,
                                                             data = IndividualData |> 
                                                           filter(sex == 1 & FTFY == 1 & Self_employed == 0)),
-  "\\specialcell{(2) \\\\  Log annual earnings}" = feols(lninctot_1999 ~ .[ParentDummies] | year + age + educ, se = "iid",
+  "\\specialcell{(2) \\\\  Log annual earnings}" = feols(lninctot_1999 ~ .[ParentDummies] | year + age + educ + statefip, vcov = ~statefip,
                                                          data = IndividualData |> 
                                                            filter(sex == 1 & FTFY == 1 & Self_employed == 0))
   
@@ -65,7 +65,7 @@ hoursworked <- as.data.frame(hoursworked)
 
 all_row <- rbind(differences_row, pvalue_row, controling_for, hoursworked)
 
-attr(all_row, 'position') <- c(7:8, 10:11)
+attr(all_row, 'position') <- c(7:8, 12:13)
 
 cm <- c("WH" = "$WH_{i}$",
         "HW" = "$HW_{i}$",
@@ -74,7 +74,8 @@ cm <- c("WH" = "$WH_{i}$",
 gm <- tibble::tribble(
   ~raw,        ~clean,          ~fmt,
   "nobs",      "Observations",             0,
-  "FE: year", "Year", 0,
+  "FE: statefip", "State FE", 0,
+  "FE: year", "Year FE", 0,
   "FE: age", "Age", 0,
   "FE: educ", "Education", 0,
   "std.error.type", "Standard Errors", 0,
@@ -95,7 +96,8 @@ modelsummary(reg1, fmt = 2,
   footnote(number = c("\\\\footnotesize{This table includes the estimation results of equation (\\\\ref{eq:1a}).}",
                       "\\\\footnotesize{The four groups stand for White Husband White Wife (WW), White Husband Hispanic Wife (WH), Hispanic Husband White (HW), and Hispanic Husband Hispanic Wife (HH).}",
[…]

### 2023-04-12 programs/14-table-six-regression.R

@@ -13,10 +13,10 @@ ParentDummies <- c("WH:Hispanic_ID", "HW:Hispanic_ID", "HH:Hispanic_ID","WH", "H
 
 # By generation
 reg1 <- list(
-  "\\specialcell{(1) \\\\ Log annual earnings}" = feols(lninctot_1999 ~ .[ParentDummies] | year + age, se = "iid",
+  "\\specialcell{(1) \\\\ Log annual earnings}" = feols(lninctot_1999 ~ .[ParentDummies] | year + age + statefip, vcov = ~statefip,
                                                         data = IndividualData |> 
                                                           filter(sex == 1 & FTFY == 1 & Self_employed == 0)),
-  "\\specialcell{(2) \\\\  Log annual earnings}" = feols(lninctot_1999 ~ .[ParentDummies] | year + age + educ, se = "iid",
+  "\\specialcell{(2) \\\\  Log annual earnings}" = feols(lninctot_1999 ~ .[ParentDummies] | year + age + educ + statefip, vcov = ~statefip,
                                                          data = IndividualData |> 
                                                            filter(sex == 1 & FTFY == 1 & Self_employed == 0))
   
@@ -66,7 +66,7 @@ controling_for <- as.data.frame(controling_for)
 hoursworked <- as.data.frame(hoursworked)
 
 all_row <- rbind(differences_row, pvalue_row, controling_for, hoursworked)
-attr(all_row, 'position') <- c(7:8, 10:11)
+attr(all_row, 'position') <- c(7:8, 12:13)
 
 cm <- c("WH:Hispanic_ID" = "$WH_{i} \\times Hispanic_{i}$",
         "Hispanic_ID:HW" = "$HW_{i} \\times Hispanic_{i}$",
@@ -78,8 +78,9 @@ cm <- c("WH:Hispanic_ID" = "$WH_{i} \\times Hispanic_{i}$",
 gm <- tibble::tribble(
   ~raw,        ~clean,          ~fmt,
   "nobs",      "Observations",             0,
-  "FE: age", "Age", 0,
+  "FE: statefip", "State FE", 0,
   "FE: year", "Year FE", 0,
+  "FE: age", "Age", 0,
   "FE: educ", "Education", 0,
   "std.error.type", "Standard Errors", 0,
   #"r.squared", "R squared", 3
@@ -120,7 +121,8 @@ regression_tab <- modelsummary(reg1, fmt = 2,
[…]

### 2023-04-13 my_paper/Hadah-last-names.tex
edits
...

@@
 
-Let $Y_{ist}$ be the annual log earnings of person $i$ at time $t$.
+Let $Y_{ist}$ be the annual log earnings of person $i$ in state $s$ at time $t$.
 $WH_{ist}$, $HW_{ist}$ and $HH_{ist}$ are indicator variables for the type of parents person $i$ has.
@@
 From equation \ref{eq:naive}, $\beta_{1}$ estimates the earnings gap between Hispanics and Whites.
-When calculating the equation, before controlling for education, I find an earnings gap that is equal to 15.6\%.
-This gap is close to the gap between HH and WW.
-After controlling for education, the gap decreases by more than half and becomes 6\%, which is also similar to the gap between HH and WW after controlling for education.
+When calculating the equation, before controlling for education, I find an earnings gap that is equal to 17\%.
+After controlling for education, the gap decreases by more than half and becomes 7\%.
 Since the differences between Whites and Hispanics, both observable and unobservable, are large, estimating the gap HW-WH could provide us with a bound on how much of the discrimination against Hispanics is taste-based.
 
-The point estimate of the difference between HW and WH from equation \ref{eq:1a} is equal to 1.3 percentage points.
-The upper bound of the confidence interval is equivalent to 4.5 percentage points.
+The point estimate of the difference between HW and WH from equation \ref{eq:1a} is equal to 2 percentage points.
+The lower bound of the confidence interval is equal to 6 percentage points.
 Therefore, I can put a potential upper bound on the portion of the Hispanic-White earnings gap that could be explained by taste based discrimination.
-In this case, the upper bound is equal to 4.5 percentage points.
+In this case, the upper bound is equal to 6 percentage points.
 

### 2023-04-13 programs/15-table-seven-regression.R

@@ -92,3 +92,21 @@ regression_tab %>%
 
 regression_tab %>%
   save_kable(file.path(thesis_tabs,"tab07-regression.tex"))
+
+# dissertation table
+regression_tab <- modelsummary(reg1, fmt = 2,
+                               output = "latex",
+                               coef_map = cm,
+                               add_rows = all_row,
+                               gof_map = gm,
+                               escape = F,
+                               #gof_omit = 'DF|Deviance|R2|AIC|BIC|Log.Lik.|F|Std.Errors',
+                               stars= c('***' = 0.01, '**' = 0.05, '*' = 0.1)) %>%
+  kable_styling(bootstrap_options = c("hover", "condensed"), 
+                latex_options = c("scale_down", "hold_position")
+  )
+
+
+regression_tab %>%
+  save_kable(file.path(dissertation_wd,"tables/tab07-naive-regression.tex"))
+

### 2023-04-13 my_paper/Hadah-last-names.tex
...

@@
 From equation \ref{eq:naive}, $\beta_{1}$ estimates the earnings gap between Hispanics and Whites.
-When calculating the equation, before controlling for education, I find an earnings gap that is equal to 17\%.
-After controlling for education, the gap decreases by more than half and becomes 7\%.
+When calculating the equation, before controlling for education, I find an earnings gap that is equal to 22\%.
+After controlling for education, the gap decreases by more than half and becomes 11\%.
+Both of the estimates is similiar to the earnings gap between HH and WW.
 Since the differences between Whites and Hispanics, both observable and unobservable, are large, estimating the gap HW-WH could provide us with a bound on how much of the discrimination against Hispanics is taste-based.

### 2023-04-16 my_paper/Hadah-last-names.tex
some edits

--- 2023-04-13 my_paper/Hadah-last-names.tex
@@
 
-There is a plethora of evidence of ethnic and racial discrimination in the labor market \citep{bayer2018divergent, charles2008prejudice, card1992school, fryer2004causes, rubinstein2014pride, bertrand2004emily, juhn1991accounting}.
+A large literature documents substantial earnings gaps across race and ethnicity \citep{bayer2018divergent, charles2008prejudice, card1992school, fryer2004causes, rubinstein2014pride, bertrand2004emily, juhn1991accounting}.
 Hispanics constitute a large and growing portion of the population in the United States.
-As the number of Hispanics keeps increasing in the labor force, determining whether ethnic discrimination affects their labor market outcomes becomes increasingly crucial \citep{chettyUnitedStatesStill2014, chettyEffectsExposureBetter2016,chettyFadingAmericanDream2017,abramitzkyImmigrantsAssimilateMore2020a, abramitzkyNationImmigrantsAssimilation2014,abramitzkyCulturalAssimilationAge2016,chettyWhereLandOpportunity2014}, how identity could shift public opinions toward trade \citep{grossmanIdentityPoliticsTrade2021}, and how racial and gender attitudes affect the racial and gender earnings gaps \citep{charlesPrejudiceWagesEmpirical2008,charlesEffectsSexismAmerican2018}.
+As the number of Hispanics increases, determining whether ethnic discrimination affects their labor market outcomes becomes increasingly crucial \citep{chettyUnitedStatesStill2014, chettyEffectsExposureBetter2016,chettyFadingAmericanDream2017,abramitzkyImmigrantsAssimilateMore2020a, abramitzkyNationImmigrantsAssimilation2014,abramitzkyCulturalAssimilationAge2016,chettyWhereLandOpportunity2014}.
 Thus, it is important to understand whether a person's ethnicity affects their labor market outcomes.
@@
 The US population is growing in diversity.
-The proportion of non-Whites has increased by more than 10 percentage points from 1995 to 2019 (see figure \ref{fig:1}), and the number of Hispanics has grown by 9 percentage points over the same period (see figure \ref{fig:2}).
-Native-born White Hispanic men earn 21\% less than White men \citep{duncan2018identifying}.
-However, a substantial portion of the earnings gap is due to educational differences between Hispanics and Whites \citep{duncan2006hispanics, duncan2018socioeconomic}.
-Therefore, understanding discrimination against Hispanics and their children is an important task.
-Discrimination against Hispanics has a couple of negative consequences, for example discrimination against Hispanics can lead to reduced job opportunities, lower wages, and hindered assimilation.
-First, if Hispanics face a substantial labor market penalty for their ethnic identity, it could hinder assimilation and integration.
-Second, a significant labor market penalty contributes to the growing wage gap between Whites and other minority groups.
+The proportion of non-Whites has increased by more than 10 percentage points from 13 percent in 1995 to 23 percent in 2019.
+The number of Hispanics has grown by 9 percentage points from 9 percent in 1995 to 16 percent in 2019.\footnote{The portion of non-Whites and Hispanics is calculated using the Current Population Survey (CPS).} Native-born White Hispanic men earn 21\% less than White men, although a substantial portion of the earnings gap is due to educational differences between Hispanics and Whites \citep{duncan2006hispanics, duncan2018identifying, duncan2018socioeconomic}.
+Some of the earnings differences may also be due to discrimination which will have negative consequences.
+For example, discrimination against Hispanics can lead to reduced job opportunities, lower wages, and hinder assimilation.
 In this paper, I examine the role of having a Hispanic last name and identifying as Hispanic on labor market outcomes.
 
-Identifying discrimination is difficult because of factors that affect labor market outcomes that are unobservable to economists---like innate ability.
-Also, I should be able to disentangle skills differentials between Hispanics and non-Hispanic Whites.\footnote{Skills differentials refer to differences in education, work experience, and other factors that can affect a person's job performance and wages.} To causally identify ethnic discrimination against Hispanics and its effects on labor market outcomes, a researcher should compare two identical people that only differ in their last name.
-Since this approach is unfeasible, researchers took different approaches.
-\citet{bertrand2004emily} and \citet{fryer2004causes} conducted audit studies to examine the effect of Black sounding names on employment discrimination.
+Identifying discrimination is difficult because of factors that affect labor market outcomes that are unobservable to economists---such as unobserved skills.
+one strategy used by researchers is audit or resume studies.
+\citet{bertrand2004emily} conducted an audit study where identical resumes were sent to employers with White and Black Sounding names.
 This approach, however, has its drawbacks.
@@
 
-This study utilizes a method developed by \citet{rubinstein2014pride} in this research.
-I will compare children from inter-ethnic marriages.
-More precisely, I will reach children of Hispanic fathers and White mothers (henceforth HW) to children of White fathers and Hispanic mothers (henceforth  WH).
+This study utilizes a method developed by \citet{rubinstein2014pride}.
+I compare children from inter-ethnic marriages.
+More precisely, I compare children of Hispanic fathers and White mothers (henceforth HW) to children of White fathers and Hispanic mothers (henceforth  WH).
 This approach stems from the fact that marriages are not random.
@@
 \citep{averettBetterWorseRelationship2008, averettEconomicRealityBeauty1996, beckerTheoryMarriagePart1973, beckerTheoryMarriagePart1974, beckerTreatiseFamily1993, browningCollectiveUnitaryModels2006, chiapporiFatterAttractionAnthropometric2012}.
-Children of HW and WH marriages have more similar observable characteristics than children of homogeneous marriages--- i.e., White fathers-White mothers and Hispanic fathers-Hispanic mothers.
+Children of HW and WH marriages have more similar observable characteristics than children of endogamous/homogonous marriages--- i.e., White fathers-White mothers and Hispanic fathers-Hispanic mothers.
 Moreover, children from a Hispanic father and White mother household will have a Hispanic last name from their fathers, enabling the investigation of how ethnic signals, such as having a Hispanic last name, affect annual log earnings.
 
-Using the Current Population Survey (CPS) from 1994 to 2019 and the 1970-1990 US censuses, I will estimate the effect of having a Hispanic last name on the labor market outcomes.
-In other words, I will approximate the variation in labor market perception of ethnic signals, in this case having a Hispanic last name, on annual log earnings.
+Using the Current Population Survey (CPS) from 1994 to 2019 and the 1970-1990 US censuses, I estimate the effect of having a Hispanic last name on labor market outcomes.
+In other words, I approximate the variation in labor market perception of ethnic signals, in this case having a Hispanic last name, on annual log earnings.
 I also study the effect of identifying as Hispanic on labor market outcomes.
 
-To retrieve the effect of having a Hispanic last name, I compare the children of inter-ethnic marriages between Whites and Hispanics.
-I find that while men born to White father-Hispanic mothers earn less than men born to Hispanic father-White mothers, this gap is entirely explained by educational differences.
+I find that while men born to Hispanic father-White mothers earn less than men born to White father-Hispanic mothers, this gap is entirely explained by educational differences.
 Thus, I do not find a significant effect of having a Hispanic last name.
-Finally, I find that men identifying as Hispanic earn significantly less than those who do not, even when I control ancestry and education.
-Inter-ethnic children with a White father, and thus having a non-Hispanic last name, earn 14\% less than children of endogamous White parents.
+Finally, I find that men identifying as Hispanic earn significantly less than those who do not, even when I control for ancestry and education.
+
+% Inter-ethnic children with a White father, and thus having a non-Hispanic last name, earn 14\% less than children of endogamous White parents.
 Inter-ethnic children with a Hispanic father, and thus having a Hispanic last name, earn 20\% less than children of endogamous White parents.
@@
 
-A long list of economic papers investigated taste-based discrimination, statistical discrimination, and Black-White and gender earnings gaps that date back to the seminal work of \citet{becker2010economics}.
-\citet{bayer2018divergent} investigated the Black-White wage gap's narrowing and then divergence.
-\citet{card1992school} argued that the Black-White wage gap narrowed from 1960 to 1980 due to improvement in the quality of Black schools, and \citet{juhn1991accounting} showed that skills play a big part in the Black-White wage gap.
-Moreover,  \citet{charles2008prejudice} provided evidence that 25\% of the Black-White wage gap is due to prejudice.
-\citet{black1995discrimination} introduced a search model showing that with a taste for discrimination, minorities earn less than the majority group even when hired by an unprejudiced firm.
-In comparison, \citet{altonji2001employer} found that firms do not statistically discriminate based on race and ethnicity.
-Little has been done, however, about discrimination against Hispanics, which offers an opportunity to exploit the observable racial similarities between Hispanics and Whites.
-
 Previous studies have also used names as a proxy for race and ethnicity \citep{fryer2004causes, rubinstein2014pride, bertrand2004emily}.
@@
 They did, however, find that having a Black name, after controlling for the home environment at birth, does not affect their labor market outcomes.
-\citet{rubinstein2014pride} compared the children of mixed marriages between Sephardic and Ashkenazis in Israel.\footnote{Ashkenazi and Sephardic are two distinct Jewish ethnic groups.} They found that workers with Sephardic last names earn substantially less than those with an Ashkenazi last name.
+\citet{rubinstein2014pride} compared the children of mixed marriages between Sephardic and Ashkenazi Jews in Israel.\footnote{Ashkenazi and Sephardic are two distinct Jewish ethnic groups.} They found that workers with Sephardic last names earn substantially less than those with Ashkenazi last names.
 \citet{bertrand2004emily} conducted an audit study by sending employers identical resumes that differ in the ethnic and racial signal of a name (Black sounding name versus a White sounding one).
@@
 The rest of the paper is organized as follows.
-Section \ref{sec:data} provides an overview of the data and summary statistics of the sample I am using.
-Section \ref{sec:emp_model} introduces a formal empirical model.
-I go over the results in section \ref{sec:results}.
-Finally, I offer a brief conclusion in section \ref{sec:con}.
+Section \ref{sec:data} provides an overview of the data and summary statistics of the sample.
+Section \ref{sec:emp_model} introduces the empirical framework.
+I report the results in section \ref{sec:results}.
+Finally, I offer a brief conclusion in section \ref{sec:con1}.
 
 \section{Data}\label{sec:data}
+\subsection{Data}\label{sec:data1}
 
-For this paper, I will use two data sets.
-The Integrated Public Use Microdata Series (IPUMS) Current Population Survey (CPS) Annual Social and Economic (ASEC) \citep{cps2019} and the 1970 to 1990 US censuses \citep{acs2019}.
+I use two datasets:  the Integrated Public Use Microdata Series (IPUMS) Current Population Survey (CPS) Annual Social and Economic (ASEC) \citep{cps2019} and the 1970 to 1990 US censuses \citep{acs2019}.
 
-I use the CPS' rich data set to study the effect of having a Hispanic last name on a person's labor market outcomes.
-I take advantage of the fact that the CPS asks for parents' place of birth, ethnicity, and race.
+I use the CPS data set to study the effect of having a Hispanic last name on a person's labor market outcomes.
+I take advantage of the fact that the CPS asks parents' place of birth, ethnicity, and race.
 The data spans the period between 1994--- the earliest sample to ask about a parent's place of birth--- to 2019.
@@
 I construct the synthetic parents by linking husbands and wives in the census data to each other.
-I assume that parents have children between the ages of 25 and 40, I then link these synthetic parents using the year they were surveyed to the year of birth of the "children" in the CPS sample.
+I assume that parents have children between the ages of 25 and 40.
+I then link these synthetic parents using the year they were surveyed to the year of birth of the "children" in the CPS sample.
 
 \subsection{Children of the four parental types}
-I will use the CPS for my primary analysis of the effect of having a Hispanic last name on earnings.
+
+I use the CPS for my primary analysis of the effect of having a Hispanic last name on earnings.
 I restrict my sample to Whites, United States-born citizens aged 25 to 40 and born between 1970 to 1990.
@@
 \end{enumerate}
-The distribution of the four types of children is presented in figure \ref{fig:dist}.
-The majority of the sample, 95.14\%, is WW children.
+
+The distribution of the four types of children is presented in figure \ref{fig:dist} and table \ref{tab:mat1}.
+The majority of the sample (95.14\%) is WW children.
 The second biggest group is HH, which constitutes 3.36\% of the sample.
@@
 The summary statistics for the children of the four types of marriages are presented in table \ref{tab:c&p1}.
-Children of WW marriages (column i) do better on every measure while children of HH parents (column iv) do worse than other children on every measure.
+Children of WW marriages (column 1) do better on every measure while children of HH parents (column 4) do worse than other children on every measure.
 
 On the one hand, WW children are more educated, with approximately 14 years of education for men and women.
-man and woman WW children have an employment rate of 95\%.
-They have log hourly earnings of 2.557 for men and 2.435 for women.
-Also, their annual log earnings are 10.529 for men and 10.297 for women.
+WW children have an employment rate of 95\% for men and 96\% for women.
+They have log hourly earnings of 2.48 for men and 2.33 for women.
+Also, their annual log earnings are 10.25 for men and 9.46 for women.
 
-On the other hand, children of HH marriages have an average of 12.898 and 13.204 years of education for men and women, respectively.
-HH men and women have an employment rate of 93\%.
-Their annual log earnings are 10.365 and 10.175 for men and women, respectively.
+On the other hand, children of HH marriages have an average of 12.1 and 13.27 years of education for men and women, respectively.
+HH men and women have an employment rate of 93\% and 94\%.
+Their annual log earnings are 10.01 and 9.53 for men and women, respectively.
 
@@
 Children of HW are going to be the treatment group.
-HW's children have a Hispanic father, who will carry his Hispanic last name.
-At the same time, the WH children have White fathers and will take his non-Hispanic, or White, last name.
-A first glance at the summary statistics of these two groups, table \ref{tab:c&p1} columns (ii) and (iii) can show that they are different from the children of non-inter-ethnic marriages.
+HW children have a Hispanic father with a Hispanic last names.
+At the same time, the WH children have White fathers with a non-Hispanic, or White, last names.
+Summary statistics of these two groups, table \ref{tab:c&p1} columns (2) and (3) can show that they are different from the children of non-inter-ethnic marriages.
 
-WH children have an average of 13.366 and 13.741 years of education for men and women.
-The employment rate for men is 94.5\% and for women 95.0\%.
-HW man and woman children obtained 13.072 and 13.270 years of schooling.
-Men's employment rate is 92.7\% and women's 93.1\%.
+WH children have an average of 13.58 and 13.87 years of education for men and women.
+The employment rate for men is 94\% and for women 95.0\%.
+HW men and women obtained 13.21 and 13.42 years of schooling.
+Men's employment rate is 92\% and women's 93\%.
 
-In panel D of table \ref{tab:c&p1}, I present the rates at which members of each group self-identify as Hispanic \footnote{Hispanic self-identification is observed when an individual notes that they are Hispanic on the CPS questionnaire}.
-The vast majority of HH children, 97\%, identify as Hispanic.
+In panel C of table \ref{tab:c&p1}, I present the rates at which members of each group self-identify as Hispanic.
+\footnote{Hispanic self-identification is observed when an individual notes that they are Hispanic on the CPS questionnaire} The vast majority of HH children, 97\%, identify as Hispanic.
 There is considerable attrition in Hispanic identification of inter-ethnic children.
-Among WH children, 82.1\% of men and 85.0\% of women identified as Hispanic.
-Among HW children, 87.9\% of men and 85.9\% of women identified as Hispanic.
+Among WH children, 81\% of men and 85\% of women identify as Hispanic.
+Among HW children, 88\% of men and 87\% of women identify as Hispanic.
 These results align with the findings of a substantial body of literature that noted this attrition among children of immigrants, especially those coming from inter-ethnic marriages \citep{duncan2017complexity, duncan2018identifying, duncan2020new, antman2020ethnic}.
-Finally, a small portion of WW children, 6\%, identified as Hispanic.
-These 6\% of WW children are most likely third generation--- or more--- immigrants.
-For my analysis, I dropped the WW observations identified as Hispanic to compare with non-Hispanic Whites.
+Finally, a small portion of WW children, 5\% of men and 6\% of women, identified as Hispanic.
+These 6\% of WW children are most likely third generation--- or higher--- children of immigrants.
+For my analysis, I dropped the WW observations identify as Hispanic to compare with non-Hispanic Whites.
 
-In columns (v) and (vi), I calculate the differences in means between HH-WW (column v) and HW-WH (column vi).
+In columns (5) and (6), I calculate the differences in means between HH-WW (column 5) and HW-WH (column 6).
 Overall, HH children do worse than WW children.
-HH has one less year of education compared to WW, 2.1 percentage points less likely to be employed, HH men earn 18\% less than WW men, and HH women make 12.3\% less than WW women.
-Therefore, these two groups, HH and WW, are different in many observable.
+HH has one year less of education compared to WW, and are 2 percentage points less likely to be employed.
+HH men earn 7\% less than WW men, and HH women make 2\% less than WW women.
+Therefore, these two groups, HH and WW, are different on many observable qualities.
 
 The differentials between WH and HW are not as severe as the ones between HH and WW.
-HW children are still worse off than WH but are way better than HH.
-HW men attain 0.294 fewer years of education than WH men.
-HW women achieve 0.471 fewer years of schooling than WH women.
-HW children are 1.8-1.9 percentage points less likely to be employed.
-HW men earn the same annual earnings as WH men, and HW women earn 10.3\% less.
-These differences show that HW and WH children are different from children of homogeneous marriages and, thus, are more comparable.
+HW children are still worse off than WH but are much better than HH.
+HW men attain 0.36 fewer years of education than WH men.
+HW women achieve 0.46 fewer years of schooling than WH women.
+HW children are 2 percentage points less likely to be employed.
+HW men earn 4\% less annual earnings as WH men, and HW women earn 7\% less.
+These differences show that HW and WH children are different from children of endogamous/homogonous marriages and, thus, are more comparable.
 
-Moreover, in table \ref{tab:c&p2}, I present the summary statistics for the four types and limit the sample to those that only identify as Hispanic\footnote{Limiting the sample to those that identify as Hispanic only affects WH, HW, and HH.
-Members of the WW group are not affected by this restriction.}.
-The same pattern discussed above between HH--WW and HW--WH persists.
+In table \ref{tab:c&p2}, I present the summary statistics for the four types and limit the sample to those that only identify as Hispanic.\footnote{Limiting the sample to those who identify as Hispanic only affects WH, HW, and HH.
+Members of the WW group are not affected by this restriction.} The same pattern discussed above between HH--WW and HW--WH persists.
 The one difference is that those who identify as Hispanic do slightly worse.
@@
 
-Using the 1970 and 1990 censuses, I constructed a panel of synthetic parents.
+Using the 1970 and 1990 censuses, I constructed a data set of synthetic parents.
 The sample includes married White men and women.
@@
 Using the information provided in the census, I can link husbands and wives with each other.
-I assume that parents have children between the ages of 25 and 45.
+I assume that parents have children between the ages of 25 and 40.
 Therefore, my sample consists of married White men and women that were born in the 1930 to 1965 cohorts\footnote{The construction of "synthetic parents" follows the method used by \citet{rubinstein2014pride}.}.
@@
 I show the distribution of the four types of couples in table \ref{tab:mat1}.
-White husbands and White wives (WW) make up the majority of couples in the sample, 98.18\% (681,206 couples).
-Hispanic husbands and wives (HH) are the second largest group making up 1.04\% (7,193 couples)of all couples.
-White husbands and Hispanic wives (WH) couples are 0.40\% (2,750 couples) of the sample while Hispanic husbands and White wives (HW) are 0.39\% (2,710 couples).
+White husbands and White wives (WW) make up the majority of couples in the sample, 97\% (1,286,731 couples).
+Hispanic husbands and wives (HH) are the second largest group making up 2\% (20,911 couples)of all couples.
+White husbands and Hispanic wives (WH) couples are less than 1\% (7,178 couples) of the sample and Hispanic husbands and White wives (HW) are also less than 1\% (7,606 couples).
 
 I present the summary statistics of the four types of synthetic parents in table \ref{tab:c&p1}'s parent's panel.
-WW couples have higher education, 11.9 years for husbands and 11.8 for wives.
-As a household, WW couples have 25.194 years of schooling.
-Husbands in HH marriages have 8.5 years of education, while women have 8.2 years of schooling.
-As a household, HH couples have 17.539 years of education.
+WW couples have higher education, 13.05 years for husbands and 12.74 for wives.
+As a household, WW couples have 25.78 years of schooling.
+Husbands in HH marriages have 8.93 years of education, while women have 8.6 years of schooling.
+As a household, HH couples have 17.54 years of education.
 
-WH husbands have 10.3 years of education, while wives have an average of 9.6 years, the lowest of any group.
-WH household attained a total of 22.308 years of schooling in total.
-HW husbands and wives obtained 9.7 and 9.8 years of education, respectively.
-HW household attained a total of 21.473 years of schooling in total.
+WH husbands have 12.32 years of education, while wives have an average of 11.03 years.
+WH household attained a total of 23.35 years of schooling in total.
+HW husbands and wives obtained 10.65 and 11.54 years of education, respectively.
+HW household attained a total of 22.19 years of schooling in total.
 
-In columns (v) and (vi) of table \ref{tab:c&p1}, I calculated the differences in means between HH-WW households and HW-WH households.
-An HH Hispanic husband attained 3.779 fewer years of education than a WW husband.
-While an HH wife acquired 3.876 fewer years of schooling than a White WW wife.
-The total educational difference between HH and WW households equals 7.655 years of education.
+In columns 5 and 6 of table \ref{tab:c&p1}, I report the differences in means between HH-WW households and HW-WH households.
+An HH Hispanic husband attained 4.11 fewer years of education than a WW husband.
+While an HH wife acquired 4.13 fewer years of schooling than a White WW wife.
+The total educational difference between HH and WW households equals 8.25 years of education.
 The educational differences between HW and WH are less severe than those between HH and WW.
-A Hispanic husband in intermarriage attained 1.560 fewer years of schooling than a White husband in intermarriage.
-A White wife in intermarriage attained 0.726 years of education, more than a Hispanic wife in intermarriage.
-The difference in total years of education between HW and WH households is 0.835, which is an 89\% decrease from the HH--WW gap.
+A Hispanic husband in intermarriage attained 1.67 fewer years of schooling than a White husband in intermarriage.
+A White wife in intermarriage attained 0.51 years of education, more than a Hispanic wife in intermarriage.
+The difference in total years of education between HW and WH households is 1.16, which is an 85\% decrease from the HH--WW gap.
 These differences show that couples that intermarry are better and different than those that do not.
@@
 
-\section{Empirical Strategy}\label{sec:emp_model}
-In this section, I will present two empirical strategies.
-The first empirical strategy will estimate the effect of having a Hispanic last name on annual log earnings.
-The second empirical strategy will estimate the effect of identifying as Hispanic on log yearly earnings.
+\subsection{Empirical Strategy}\label{sec:emp_model}
+
+In this section, I present two empirical strategies.
+The first empirical strategy estimates the effect of having a Hispanic last name on annual log earnings.
+The second empirical strategy will estimate the effect of identifying as Hispanic on log annual earnings.
 
 The difference in means between Hispanics and non-Hispanic Whites could result from discrimination.
-It can also be caused by differences in innate abilities, skills, and parental investments during infancy.
-To avoid these threats to identification, I compare children of inter-ethnic marriages, HW and WH.
-WH and HW children are equally Hispanic physically but provide employers/labor market with different signals.\footnote{WH and HW children are both half White, half Hispanic.} WH children will have a non-Hispanic last name, while HW children will have a Hispanic last name.
+It can also be caused by differences in innate abilities, skills, and parental investments.
+While controlling for observable skill measures, I compare children of inter-ethnic marriages, HW and WH.
+WH and HW children are more similar in characteristics but provide employers, and the labor market, with different signals.\footnote{WH and HW children are both half White, half Hispanic.} WH children will have a non-Hispanic last name, while HW children will have a Hispanic last name.
 This is a method developed by \citet{rubinstein2014pride}.
 
-\section{Estimating the effect of having a Hispanic last name}
+\subsubsection{Estimating the effect of having a Hispanic last name}
 
+In this section I restrict the sample to WH and HW groups.
 Let $Y_{ist}$ be the annual log earnings of person $i$ in state $s$ at time $t$.
-$WH_{ist}$, $HW_{ist}$ and $HH_{ist}$ are indicator variables for the type of parents person $i$ has.
-$X_{ist}$ is a vector of controls, $\gamma_{t}$ are year fixed effects, $\lambda_{s}$ are state fixed effects, and $\phi_{ist}$ represents the error term.
+$HW_{ist}$ is an indicator variables for the type of parents person $i$ has.
+$X_{ist}$ is a vector of controls that includes age and numbers of hours worked, $\gamma_{t}$ are year fixed effects, $\lambda_{s}$ are state fixed effects, and $\phi_{ist}$ represents the error term.
 The equation for this strategy is written as follows:
 \begin{equation} \label{eq:1a}
-Y_{ist} = \beta_{1} WH_{ist} +\beta_{2} HW_{ist} + \beta_{3} HH_{ist} + X_{ist} \pi + \gamma_{t} + \lambda_s + \phi_{ist}
+Y_{ist} = \beta_{1} HW_{ist} + X_{ist} \pi + \gamma_{t} + \lambda_s + \phi_{ist}
 \end{equation}
 
-$\beta_{1}, \beta_{2}$ and $\beta_{3}$ are the coefficients of interest in this specification.
-They estimate the earnings gaps between the three groups and children of White-White parents.
-The difference between $\beta_{1}$ and $\beta_{2}$ captures the effect of having a Hispanic last name.
-This comes from the fact that the WH and HW children are similar.
-On the one hand, WH children will have a White father and carry his White last name.
-On the other hand, HW children will have a Hispanic father and carry his Hispanic last name.
-Therefore, \textit{ceteris paribus}, the difference will capture the effect of having a Hispanic last name.
+$\beta_{1}$ is the coefficient of interest in this specification.
+$\beta_{1}$ estimates the earnings gaps between children of inter-ethnic marriages who have a Spanish-sounding last name versus a White last name.
 
-\subsection{Estimating the effect of identifying as Hispanic}
+\subsubsection{Estimating the effect of identifying as Hispanic}
 
 In this section, I will present an estimation strategy that would allow me to capture the effect of identifying as a Hispanic.
-The empiricism in the section follows the theoretical identity model developed by \citet{akerlof2000economics}.
 
-Let $Y_{ist}$ be the annual log earnings of person $i$ in state $s$ at time $t$.
-$WH_{ist}$, $HW_{ist}$ and $WW_{ist}$ are indicator variables for the type of parents person $i$ has.
-$X_{ist}$ is a vector of controls, $\gamma_{t}$ are year fixed effects, $\lambda_s$ are state fixed effects, and $\phi_{ist}$ represents the error term.
-The addition to the previous model is the $Hispanic_{ist}$ indicator variable.
-$Hispanic_{ist}$ is equal to one if a person identifies as Hispanic and zero otherwise.
-I interact this term with the $WH_{ist}$, $HW_{ist}$ and $HH_{ist}$.
-The regression is presented in the following equation.
+\begin{equation} \label{eq:iden}
+Y_{ist} = \beta_{1} HW_{ist} + \beta_{2} Hispanic_i + \beta_{3} HW_{ist} \cdot Hispanic_i X_{ist} \pi + \gamma_{t} + \lambda_s + \phi_{ist}
+\end{equation}
 
-\begin{align} \label{eq:iden}
-Y_{ist} &= \beta_{1} HW_{ist} +  \beta_{2} WH_{ist} + \beta_{3} HH_{ist} +\\
-& \beta_{4} WH_{ist} \cdot Hispanic_{ist} + \beta_{5} HW_{ist}\cdot Hispanic_{ist} +  \beta_{6} HH_{ist} \cdot Hispanic_{ist}+ \nonumber \\
-&X_{ist} \pi + \gamma_{t} + \lambda_s + \phi_{ist} \nonumber
-\end{align}
+The coefficients of interest from equation \ref{eq:iden} are $\beta_{1}$, $\beta_{2}$ and $\beta_{3}$. The coefficients $\beta_{2}$ and $\beta_{3}$ capture the effects of Spanish-sounding last names.
 
-The coefficients of interest from equation \ref{eq:iden} are $\beta_{4}$, $\beta_{5}$ and $\beta_{6}$.
-The interaction terms capture the earnings gaps between members of $WH_{ist}$, $HW_{ist}$, and $WW_{ist}$ that identify as Hispanic and White WW children.
-
-\section{Results and Discussion}\label{sec:results}
+\subsection{Results}\label{sec:results1}
 
@@
 
-I find that the earnings gaps between WH and WW are 14\%, 18\% between HW and WW, and 24\% between HH and WW.
-Moreover, the difference in earnings between HW and WH children, that captures the effect of having a Hispanic last name, equals a statistically significant 6 percentage points.
-Meaning by comparing inter-ethnic children, a person with a Hispanic last name earns 6 percentage points less than someone with a White last name.
-However, more than half of the earnings gaps could be explained by educational differences between WW children and WH, HW, and HH.
-When I control for education, the gap between the three groups and WW is cut between by 50\% to 64.0\%.
-Furthermore, the last name effect decreases to a statistically insignificant 2 percentage points earnings gap.
-
-I also find a significant earnings gap between those that identify as Hispanic and WW children.
-The gaps between WH, HW, and HH that identify as Hispanic and WW are, respectively, 14\%, 18\%, and 10\%.
-The difference between WH and HW that identify as Hispanic---captures the last name effect of those that identify as Hispanic--- is equal to 3 percentage points and is statistically insignificant.
-Controlling for education explains a big part of the gap.
-The gaps between WH, HW, and HH that identify as Hispanic and WW after controlling for education are 8\%, 12\%, and 6\%.
-The WH--WW and HH--WW gaps for those identifying as Hispanic become statistically insignificant after controlling for education.
-Finally, the difference between WH and HW that identify as Hispanic that identifies as Hispanic remains statistically insignificant but increases to 4 percentage points.
+I find that an inter-ethnic person with a Spanish-sounding last names earns less than an inter-ethnic who has a White last name.
+A person with a Spanish-sounding last name earns 5 percentage points less than a person with a White last name.
+In other words, by comparing inter-ethnic children, a person with a Hispanic last name earns 5 percentage points less than someone with a White last name.
+However, more than half of the earnings gaps could be explained by educational differences.
+When I control for education, the last name effect decreases to a statistically insignificant 2 percentage points earnings gap.
+I also find a significant earnings gap between those that identify as Hispanic.
 
@@
 
-\section{Discussion}
+\subsection{Conclusion}\label{sec:con1}
 
- The results presented in tables \ref{tab:lastnamereg} and \ref{tab:identreg} have an implication that affects how Hispanic-White earnings gaps should be calculated.
-First, A big portion of the earnings gaps could be explained by education.
-However, more unobservable characteristics, like innate ability, could affect earnings outcomes.
-Therefore, comparing Hispanics to Whites could yield inflated gaps since the two groups are different.
-The reduced form regression of this approach could be written as:
-
- \begin{equation} \label{eq:naive}
- Y_{ist} = \beta_{1} Hispanic_{ist} + X_{ist}^{\prime} + \gamma_{t} + \lambda_s + \phi_{ist}
- \end{equation}
-
-Where $Y_{ist}$ in equation \ref{eq:naive} is the annual log earnings of person $i$ at time $t$.
-$Hispanic_{ist}$ is a dummy variable equal to one if person $i$ is Hispanic and zero otherwise.
-$X_{ist}$ is a vector of controls, $\gamma_{t}$ is time fixed effects and $\phi_{ist}$ is the error term.
-
-From equation \ref{eq:naive}, $\beta_{1}$ estimates the earnings gap between Hispanics and Whites.
-When calculating the equation, before controlling for education, I find an earnings gap that is equal to 17\%.
-After controlling for education, the gap decreases by more than half and becomes 7\%.
-Since the differences between Whites and Hispanics, both observable and unobservable, are large, estimating the gap HW-WH could provide us with a bound on how much of the discrimination against Hispanics is taste-based.
-
-The point estimate of the difference between HW and WH from equation \ref{eq:1a} is equal to 2 percentage points.
-The lower bound of the confidence interval is equal to 6 percentage points.
-Therefore, I can put a potential upper bound on the portion of the Hispanic-White earnings gap that could be explained by taste based discrimination.
-In this case, the upper bound is equal to 6 percentage points.
-
-\section{Conclusion}\label{sec:con}
-
-As the Hispanic population keeps growing in the United States, studying discrimination against this group is also important.
+As the Hispanic population grows in the United States, studying discrimination against this group becomes increasingly important.
 In this paper, I examine discrimination against Hispanics in the labor market.
@@
 Controlling for education, however, explains more than half of the earnings gaps between WW children and WH, HW, and HH.
-When I control for education, the gap between the three groups and WW is cut between 50-64.0\%.
-Furthermore, the last name effect decreases to a statistically insignificant 2\% earnings gap.
+When I control for education, the last name effect decreases to a statistically insignificant 2\% earnings gap.
 
-I also find that men who identify as Hispanic earn significantly less than those who do not, even when I control ancestry and education.
+I also find that men who identify as Hispanic earn significantly less than those who do not, even when I control for ancestry and education.
 The gaps between WH, HW, and HH that identify as Hispanic and WW are, respectively, 14\%, 18\%, and 10\%.
@@
 \section{Figures}\label{appendix:figures}
-
-\begin{center}
-\begin{figure}[H]
-\caption{proportion of non-Whites in the United States from 1995 to 2019}
-\includegraphics[width=\textwidth]{GraphNonWhites.png} 
-\label{fig:1}
-\end{figure}
-\end{center}
-
-\newpage
-
-\begin{center}
-\begin{figure}
-\caption{Hispanics as a percentage of the total population in \\
-the United States from 1995 to 2019.}
-\includegraphics[width=\textwidth]{HispanicUSA.png} \label{fig:2}
-\end{figure}
-\end{center}
-
-\newpage
 
@@
 
-\newpage
-
-\input{tables/tab07-regression}
-
 % \begin{table}[H]

### 2023-04-16 my_paper/tables/tab04-summary-stats.tex
some edits

@@
-\begin{table}[!h]
+\begin{table}[H]
 
@@
 \cmidrule(l{3pt}r{3pt}){2-5} \cmidrule(l{3pt}r{3pt}){6-7}
-Variables & \specialcell{White Father \\ White Mother \\ (WW) \\ (i)} & \specialcell{White Father \\ Hispanic Mother \\ (WH) \\ (ii)} & \specialcell{Hispanic Father \\ White Mother \\ (HW) \\ (iii)} & \specialcell{Hispanic Father \\ Hispanic Mother \\ (HH) \\ (iv)} & \specialcell{HH - WW \\ (v)} & \specialcell{HW - WH \\ (vi)}\\
+Variables & \specialcell{White Father \\ White Mother \\ (WW) \\ (1)} & \specialcell{White Father \\ Hispanic Mother \\ (WH) \\ (2)} & \specialcell{Hispanic Father \\ White Mother \\ (HW) \\ (3)} & \specialcell{Hispanic Father \\ Hispanic Mother \\ (HH) \\ (4)} & \specialcell{HH - WW \\ (5)} & \specialcell{HW - WH \\ (6)}\\
 \midrule
-\textbf{Panel A: Parent’s} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
+\textbf{Panel A: Children's Sunthetic Parent} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
 \hspace{1em}Husband’seducation (Total Years) & \specialcell{13.05\\(2.44)} & \specialcell{12.32\\(3.33)} & \specialcell{10.65\\(4.39)} & \specialcell{8.93\\(4.41)} & \specialcell{-4.11**\\(0.02)} & \specialcell{-1.67**\\(0.04)}\\
@@
 \hspace{1em}Women’s education (Total Years) & \specialcell{13.23\\(2.25)} & \specialcell{13.75\\(2.41)} & \specialcell{13.32\\(2.34)} & \specialcell{13.26\\(2.37)} & \specialcell{0.03**\\(0.01)} & \specialcell{-0.43**\\(0.03)}\\
-\textbf{Panel C: Employment and Earnings} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
+\textbf{Panel C: Children's Employment and Earnings} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
 \hspace{1em}Men’s Employment Rate & \specialcell{0.93\\(0.26)} & \specialcell{0.94\\(0.23)} & \specialcell{0.92\\(0.27)} & \specialcell{0.93\\(0.26)} & \specialcell{0.00***\\(0.00)} & \specialcell{-0.02***\\(0.00)}\\
@@
 \hspace{1em}Men’s Log Hourly Earnings & \specialcell{2.4\\(0.44)} & \specialcell{2.41\\(0.45)} & \specialcell{2.4\\(0.44)} & \specialcell{2.41\\(0.43)} & \specialcell{0.01***\\(0.01)} & \specialcell{-0.01**\\(0.02)}\\
-\hspace{1em}Women’s Hourly Earnings & \specialcell{2.26\\(0.43)} & \specialcell{2.32\\(0.45)} & \specialcell{2.27\\(0.45)} & \specialcell{2.3\\(0.41)} & \specialcell{0.04***\\(0.01)} & \specialcell{-0.05**\\(0.02)}\\
+\hspace{1em}Women’s Log Hourly Earnings & \specialcell{2.26\\(0.43)} & \specialcell{2.32\\(0.45)} & \specialcell{2.27\\(0.45)} & \specialcell{2.3\\(0.41)} & \specialcell{0.04***\\(0.01)} & \specialcell{-0.05**\\(0.02)}\\
 \hspace{1em}Men’s Log Annual Earnings & \specialcell{10.02\\(1.02)} & \specialcell{10.06\\(1.06)} & \specialcell{10.03\\(0.99)} & \specialcell{10\\(1.04)} & \specialcell{-0.02**\\(0.01)} & \specialcell{-0.03**\\(0.04)}\\
-\hspace{1em}Women’s Hourly Earnings & \specialcell{9.44\\(1.59)} & \specialcell{9.55\\(1.59)} & \specialcell{9.47\\(1.55)} & \specialcell{9.52\\(1.52)} & \specialcell{0.08**\\(0.02)} & \specialcell{-0.08*\\(0.05)}\\
+\hspace{1em}Women’s Log Annual Earnings & \specialcell{9.44\\(1.59)} & \specialcell{9.55\\(1.59)} & \specialcell{9.47\\(1.55)} & \specialcell{9.52\\(1.52)} & \specialcell{0.08**\\(0.02)} & \specialcell{-0.08*\\(0.05)}\\
 \bottomrule
@@
 \begin{tablenotes}
-\item[1] The data is restricted to native-born United States citizens between 1994 and 2019 who are also White and between the ages of 25 and 40.
+\item[1] Source: The 1970-1990 Census for synthetic parents, and 1994-2019 Current Population Surveys (CPS) for children's outcomes
+\item[2] The data is restricted to native-born United States citizens between 1994 and 2019 who are also White and between the ages of 25 and 40, and who identify as Hispanic.
 I identify the ethnicity of a person's parents through the parent's place of birth.
@@
 A parent is White if they were born in the United States.
-\item[2] In each column, I present the average statistics of the different types of people based on the ethnicities of their parents.
-In column one, I show the summary statistics of children of White fathers and White mothers.
-In column two, I present the summary statistics of children of White fathers and Hispanic mothers.
-In column three, I show the summary statistics of children of Hispanic fathers and White mothers.
-In column four, I present the summary statistics of children of Hispanic fathers and mothers.
-\item[3] Columns five and six have data on the HH--WW gaps (column five) and the HW--WH gaps (column six).
 \end{tablenotes}

### 2023-04-16 my_paper/tables/tab05-regression.tex
some edits

--- 2023-01-20 output/tables/tab05-regression.tex
@@
 \midrule
-$WH_{i}$ & \num{-0.09}*** & \num{-0.04}**\\
+$HW_{i}$ & \num{-0.05}** & \num{-0.02}\\
  & (\num{0.02}) & (\num{0.02})\\
-$HW_{i}$ & \num{-0.14}*** & \num{-0.06}***\\
- & (\num{0.02}) & (\num{0.01})\\
-$HH_{i}$ & \num{-0.16}*** & \num{-0.06}***\\
- & (\num{0.01}) & (\num{0.01})\\
 \midrule
-$HW_{i} - WH_{i}$ & -0.05** & -0.02\\
- & (0.02) & (0.02)\\
-\midrule
-Observations & \num{129359} & \num{129359}\\
+State FE & X & X\\
+Year FE & X & X\\
 \textit{Controlling for:} &  & \\
 Hours Worked & X & X\\
-Year & X & X\\
 Age & X & X\\
 Education &  & X\\
+Observations & \num{2423} & \num{2423}\\
 \bottomrule
@@
 \item[1] \footnotesize{This table includes the estimation results of equation (\ref{eq:1a}).}
-\item[2] \footnotesize{The four groups stand for White Husband White Wife (WW), White Husband Hispanic Wife (WH), Hispanic Husband White (HW), and Hispanic Husband Hispanic Wife (HH).}
+\item[2] \footnotesize{HW is an indicator variable that is equal to 1 if a person is the child of a Hispanic-father and White-mother.}
 \item[3] \footnotesize{The sample is restricted to men working full-time full-year and are waged and salaried workers.}
 \item[4] \footnotesize{Column one has the regression results when controlling for hours worked, age, and fixed effects. Column two has the results after controlling for education.}
+\item[5] \footnotesize{Standard errors are clustered on the state level.}
 \end{tablenotes}

### 2023-04-16 output/tables/tab01-observations-by-parents.tex
some edits

\begin{table}[H]

\caption{Number of Children by Parental Type \label{tab:mat1}}
\centering
\resizebox{\linewidth}{!}{
\begin{threeparttable}
\begin{tabular}[t]{>{}lcccc}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{4}{c}{Perental Type} \\
\cmidrule(l{3pt}r{3pt}){2-5}
  & \specialcell{White Father \\ White Mother} & \specialcell{White Father \\ Hispanic Mother} & \specialcell{Hispanic Father \\ White Mother} & \specialcell{Hispanic Father \\ Hispanic Mother}\\
\midrule
\textbf{\specialcell{Observations\\Share}} & \specialcell{3,828,024\\0.95} & \specialcell{24,306\\0.01} & \specialcell{35,841\\0.01} & \specialcell{135,300\\0.03}\\
\bottomrule
\end{tabular}
\begin{tablenotes}
\item[1] Source: Current Population Surveys (CPS) 1994-2019
\item[2] The sample includes Whites, who are married, and are between the ages 25 and 40. Ethnicity of a person's parents are identified by the parent's place of birth. A parent is Hispanic if she/he was born in a Spanish-speaking country. A parent is White she/he was born in the United States.
\end{tablenotes}
\end{threeparttable}}
\end{table}

### 2023-04-16 output/tables/tab02-observations-by-couple.tex
some edits

--- 2023-04-16 output/tables/tab01-observations-by-parents.tex
@@
 
-\caption{Number of Children by Parental Type \label{tab:mat1}}
+\caption{Couples' Type \label{tab:mat2}}
 \centering
@@
 \toprule
-\multicolumn{1}{c}{ } & \multicolumn{4}{c}{Perental Type} \\
+\multicolumn{1}{c}{ } & \multicolumn{4}{c}{Couples Type} \\
 \cmidrule(l{3pt}r{3pt}){2-5}
-  & \specialcell{White Father \\ White Mother} & \specialcell{White Father \\ Hispanic Mother} & \specialcell{Hispanic Father \\ White Mother} & \specialcell{Hispanic Father \\ Hispanic Mother}\\
+  & \specialcell{White Husband \\ White Wife} & \specialcell{White Husband \\ Hispanic Wife} & \specialcell{Hispanic Husband \\ White Wife} & \specialcell{Hispanic Husband \\ Hispanic Wife}\\
 \midrule
-\textbf{\specialcell{Observations\\Share}} & \specialcell{3,828,024\\0.95} & \specialcell{24,306\\0.01} & \specialcell{35,841\\0.01} & \specialcell{135,300\\0.03}\\
+\textbf{\specialcell{Observations\\Share}} & \specialcell{1,286,731\\0.97} & \specialcell{7,178\\0.01} & \specialcell{7,606\\0.01} & \specialcell{20,911\\0.02}\\
 \bottomrule
@@
 \begin{tablenotes}
-\item[1] Source: Current Population Surveys (CPS) 1994-2019
+\item[1] Source: 1970-1990 Census
 \item[2] The sample includes Whites, who are married, and are between the ages 25 and 40.

### 2023-04-16 output/tables/tab03a-summary-stats.tex
some edits

\begin{table}[H]

\caption{Summary statistics of synthetic parents by couple type \label{tab:synth}}
\centering
\resizebox{\linewidth}{!}{
\begin{threeparttable}
\begin{tabular}[t]{lcccccc}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{4}{c}{Father's and Mother's Ethnicities} & \multicolumn{2}{c}{Differences} \\
\cmidrule(l{3pt}r{3pt}){2-5} \cmidrule(l{3pt}r{3pt}){6-7}
Variables & \specialcell{White \\ White \\ (WW) \\ (1)} & \specialcell{White \\ Hispanic \\ (WH) \\ (2)} & \specialcell{Hispanic \\ White \\ (HW) \\ (3)} & \specialcell{Hispanic \\ Hispanic \\ (HH) \\ (4)} & \specialcell{HH - WW \\ (5)} & \specialcell{HW - WH \\ (6)}\\
\midrule
Husband's education (Total Years) & \specialcell{13.05\\(2.44)} & \specialcell{12.32\\(3.33)} & \specialcell{10.65\\(4.39)} & \specialcell{8.93\\(4.41)} & \specialcell{-4.11**\\(0.02)} & \specialcell{-1.67**\\(0.04)}\\
Wife's education (Total Years) & \specialcell{12.74\\(2.12)} & \specialcell{11.03\\(3.92)} & \specialcell{11.54\\(3.12)} & \specialcell{8.6\\(4.13)} & \specialcell{-4.13**\\(0.02)} & \specialcell{0.51**\\(0.04)}\\
Total Household seducation (Total Years) & \specialcell{25.78\\(4.08)} & \specialcell{23.35\\(6.51)} & \specialcell{22.19\\(6.69)} & \specialcell{17.54\\(7.83)} & \specialcell{-8.25**\\(0.03)} & \specialcell{-1.16*\\(0.07)}\\
Log Total Family Income & \specialcell{10.79\\(0.57)} & \specialcell{10.65\\(0.66)} & \specialcell{10.55\\(0.74)} & \specialcell{10.38\\(0.75)} & \specialcell{-0.41***\\(0)} & \specialcell{-0.1***\\(0.01)}\\
\bottomrule
\end{tabular}
\begin{tablenotes}
\item[1] Source: The 1970-1990 Census for synthetic parents, and 1994-2019 Current Population Surveys (CPS) for children's outcomes
\item[2] The data is restricted to native-born United States citizens between 1994 and 2019 who are also White and between the ages of 25 and 40. I identify the ethnicity of a person's parents through the parent's place of birth. A parent is Hispanic if they were born in a Spanish-speaking country. A parent is White if they were born in the United States.
\end{tablenotes}
\end{threeparttable}}
\end{table}

### 2023-04-16 output/tables/tab06-regression.tex
some edits

--- 2023-01-20 my_paper/tables/tab06-regression.tex
@@
 \midrule
-$WH_{i} \times Hispanic_{i}$ & \num{-0.13}** & \num{-0.06}\\
- & (\num{0.05}) & \vphantom{2} (\num{0.05})\\
-$HW_{i} \times Hispanic_{i}$ & \num{-0.16}*** & \num{-0.11}**\\
- & (\num{0.05}) & \vphantom{1} (\num{0.05})\\
-$HH_{i} \times Hispanic_{i}$ & \num{-0.11}** & \num{-0.06}\\
- & (\num{0.05}) & \vphantom{2} (\num{0.04})\\
-$WH_{i}$ & \num{0.03} & \num{0.01}\\
- & (\num{0.05}) & \vphantom{1} (\num{0.04})\\
-$HW_{i}$ & \num{0.01} & \num{0.04}\\
+$HW_{i} \times Hispanic_{i}$ & \num{-0.02} & \num{-0.02}\\
+ & (\num{0.08}) & (\num{0.07})\\
+$Hispanic_{i}$ & \num{-0.11}** & \num{-0.06}\\
  & (\num{0.05}) & (\num{0.05})\\
-$HH_{i}$ & \num{-0.06} & \num{0.00}\\
- & (\num{0.05}) & (\num{0.04})\\
+$HW_{i}$ & \num{-0.03} & \num{0.00}\\
+ & (\num{0.08}) & (\num{0.06})\\
 \midrule
-$HW_{i} \times Hispanic_{i} - WH_{i} \times Hispanic_{i}$ & -0.03 & -0.04\\
- & (0.07) & (0.07)\\
-\midrule
-Observations & \num{129359} & \num{129359}\\
+State FE & X & X\\
+Year FE & X & X\\
 \textit{Controlling for:} &  & \\
@@
 Age & X & X\\
-Year FE & X & X\\
 Education &  & X\\
+Observations & \num{2423} & \num{2423}\\
 \bottomrule
@@
 \item[1] \footnotesize{This table includes the estimation results of equation (\ref{eq:iden}).}
-\item[2] \footnotesize{The four groups stand for White Husband White Wife (WW), White Husband Hispanic Wife (WH), Hispanic Husband White (HW), and Hispanic Husband Hispanic Wife (HH).
-$Hispanic_{i}$ is a dummy variable equal to one if a person identifies as Hispanic and zero otherwise.}
+\item[2] \footnotesize{HW is an indicator variable that is equal to 1 if a person is the child of a Hispanic-father and White-mother, and Hispanic is an indicator variable that is equal to one if a person self-identifies as Hispanic.}
 \item[3] \footnotesize{The sample is restricted to men working full-time full-year and are waged and salaried workers.}
-\item[4] \footnotesize{Column one has the regression results when controlling for hours worked, age, and years of fixed effects. Column two has the results after controlling for education.}
+\item[4] \footnotesize{Column one has the regression results when controlling for hours worked, age, and fixed effects. Column two has the results after controlling for education.}
 \end{tablenotes}

### 2023-04-16 programs/07-table-one-parent-types.R

@@ -19,7 +19,7 @@ WW <- DATA %>%
   filter(Type == "White-White") %>%
   summarise(n())
 # prettyNum(WW, big.mark = ",")
-WW_perc <- paste0("(",round(WW/ALL, digits = 2), ")")
+WW_perc <- paste0(round(WW/ALL, digits = 2))
 
 # calculate the number 
 # of hispanic fathers
@@ -28,7 +28,7 @@ WW_perc <- paste0("(",round(WW/ALL, digits = 2), ")")
 HW <- DATA %>%
   filter(Type == "Hispanic-White") %>%
   summarise(n())
-HW_perc <- paste0("(",round(HW/ALL, digits = 2), ")")
+HW_perc <- paste0(round(HW/ALL, digits = 2))
 
 # calculate the number 
 # of white fathers
@@ -37,7 +37,7 @@ HW_perc <- paste0("(",round(HW/ALL, digits = 2), ")")
 WH <- DATA %>%
   filter(Type == "White-Hispanic") %>%
   summarise(n())
-WH_perc <- paste0("(",round(WH/ALL, digits = 2), ")")
+WH_perc <- paste0(round(WH/ALL, digits = 2))
 
 # calculate the number 
 # of hispanic fathers
@@ -46,13 +46,17 @@ WH_perc <- paste0("(",round(WH/ALL, digits = 2), ")")
 HH <- DATA %>%
   filter(Type == "Hispanic-Hispanic") %>%
   summarise(n())
-HH_perc <- paste0("(",round(HH/ALL, digits = 2), ")")
+HH_perc <- paste0(round(HH/ALL, digits = 2))
 
 # table
 
 first_col <- c(" ",
-               "Observations"
-)
+                paste0("\\specialcell{", 
+                "Observations", 
+                      "\\\\",
+                "Share",
+                      "}"
+))
 
 second_col <- c("\\specialcell{White Father \\\\ White Mother}",
                paste0("\\specialcell{", 
@@ -98,31 +102,35 @@ Table_cols <-  Table_cols |>
 knitr::kable(Table_cols, "latex", align = "lcccc",
              booktabs = T,
              escape = F,
-             caption = "Number of Children from the Different Types of Parents \\label{tab:mat1}") %>%
+             caption = "Number of Children by Parental Type \\label{tab:mat1}") %>%
   column_spec(1, bold = T) %>%
   kable_classic(full_width = F) %>%
[…]

### 2023-04-16 programs/11-table-three-summary-stats.R

@@ -12,15 +12,15 @@ ParentData     <- read_dta(file.path(datasets,"ParentDataFull.dta")) |>
   mutate(Tot_ed = HusbandEducation + WifeEducation)
 
 row1 <- c("Variables", 
-          "\\specialcell{White Father \\\\ White Mother \\\\ (WW) \\\\ (i)}",
-          "\\specialcell{White Father \\\\ Hispanic Mother \\\\ (WH) \\\\ (ii)}",
-          "\\specialcell{Hispanic Father \\\\ White Mother \\\\ (HW) \\\\ (iii)}",
-          "\\specialcell{Hispanic Father \\\\ Hispanic Mother \\\\ (HH) \\\\ (iv)}",
-          "\\specialcell{HH - WW \\\\ (v)}",
-          "\\specialcell{HW - WH \\\\ (vi)}"
+          "\\specialcell{White \\\\ White \\\\ (WW) \\\\ (1)}",
+          "\\specialcell{White \\\\ Hispanic \\\\ (WH) \\\\ (2)}",
+          "\\specialcell{Hispanic \\\\ White \\\\ (HW) \\\\ (3)}",
+          "\\specialcell{Hispanic \\\\ Hispanic \\\\ (HH) \\\\ (4)}",
+          "\\specialcell{HH - WW \\\\ (5)}",
+          "\\specialcell{HW - WH \\\\ (6)}"
           )
 dim(row1) <- c(1,7)
-row2 <- c("Panel A: Parent’s", " ", " ", " ", " ", " ", " ")
+row2 <- c("Panel A: Synthetic Parents", " ", " ", " ", " ", " ", " ")
 dim(row2) <- c(1,7)
 
 # parent's panel
@@ -43,7 +43,7 @@ pvalues2 <- test2$std.error
 differences2 <- test2$estimate
 
 row3 <- c(
-  "Husband’seducation (Total Years)",
+  "Husband's education (Total Years)",
   paste0("\\specialcell{", 
          round(HusbandEducation[4,2], digits = 2), 
          "\\\\",
@@ -109,7 +109,7 @@ pvalues2 <- test2$std.error
 differences2 <- test2$estimate
 
 row4 <- c(
-  "Wife’seducation (Total Years)",
+  "Wife's education (Total Years)",
   paste0("\\specialcell{", 
          round(WifeEducation[4,2], digits = 2), 
          "\\\\",
[…]

### 2023-04-16 programs/11a-table-three-summary-stats.R

@@ -1,0 +1,316 @@
+# summary stats
+
+# table three
+
+# date: January 19th, 2023
+
+ParentData     <- read_dta(file.path(datasets,"ParentDataFull.dta")) |> 
+  mutate(Tot_ed = HusbandEducation + WifeEducation)
+
+row1 <- c("Variables", 
+          "\\specialcell{White \\\\ White \\\\ (WW) \\\\ (1)}",
+          "\\specialcell{White \\\\ Hispanic \\\\ (WH) \\\\ (2)}",
+          "\\specialcell{Hispanic \\\\ White \\\\ (HW) \\\\ (3)}",
+          "\\specialcell{Hispanic \\\\ Hispanic \\\\ (HH) \\\\ (4)}",
+          "\\specialcell{HH - WW \\\\ (5)}",
+          "\\specialcell{HW - WH \\\\ (6)}"
+          )
+dim(row1) <- c(1,7)
+
+# parent's panel
+
+## Husband's Education
+HusbandEducation <- ParentData |> 
+  group_by(CoupleType) |> 
+  summarise(HusEducation   = weighted.mean(HusbandEducation, na.rm = T, hhwt),
+            HusEducationSD = weighted.sd(HusbandEducation, na.rm = T, hhwt))
+
+model <- lm(HusbandEducation ~ 0 + WW + WM + MW + MM, 
+    data = ParentData, 
+    weights = hhwt)
+test1 <- tidy(glht(model, linfct = c("MM - WW = 0")))
+pvalues1 <- test1$std.error
+differences1 <- test1$estimate
+
+test2 <- tidy(glht(model, linfct = c("MW - WM = 0")))
+pvalues2 <- test2$std.error
+differences2 <- test2$estimate
+
+row3 <- c(
+  "Husband's education (Total Years)",
+  paste0("\\specialcell{", 
+         round(HusbandEducation[4,2], digits = 2), 
+         "\\\\",
+         "(",
+         round(HusbandEducation[4,3], digits = 2),
+         ")}"
+  ),
+  paste0("\\specialcell{", 
+         round(HusbandEducation[3,2], digits = 2), 
+         "\\\\",
+         "(",
+         round(HusbandEducation[3,3], digits = 2),
+         ")}"
+  ),
+  paste0("\\specialcell{", 
+         round(HusbandEducation[2,2], digits = 2), 
+         "\\\\",
+         "(",
+         round(HusbandEducation[2,3], digits = 2),
+         ")}"
+  ),
+  paste0("\\specialcell{", 
[…]

### 2023-04-16 programs/12-table-four-summary-stats.R

@@ -12,15 +12,16 @@ ParentData     <- read_dta(file.path(datasets,"ParentDataFull.dta")) |>
   mutate(Tot_ed = HusbandEducation + WifeEducation)
 
 row1 <- c("Variables", 
-          "\\specialcell{White Father \\\\ White Mother \\\\ (WW) \\\\ (i)}",
-          "\\specialcell{White Father \\\\ Hispanic Mother \\\\ (WH) \\\\ (ii)}",
-          "\\specialcell{Hispanic Father \\\\ White Mother \\\\ (HW) \\\\ (iii)}",
-          "\\specialcell{Hispanic Father \\\\ Hispanic Mother \\\\ (HH) \\\\ (iv)}",
-          "\\specialcell{HH - WW \\\\ (v)}",
-          "\\specialcell{HW - WH \\\\ (vi)}"
-)
+          "\\specialcell{White Father \\\\ White Mother \\\\ (WW) \\\\ (1)}",
+          "\\specialcell{White Father \\\\ Hispanic Mother \\\\ (WH) \\\\ (2)}",
+          "\\specialcell{Hispanic Father \\\\ White Mother \\\\ (HW) \\\\ (3)}",
+          "\\specialcell{Hispanic Father \\\\ Hispanic Mother \\\\ (HH) \\\\ (4)}",
+          "\\specialcell{HH - WW \\\\ (5)}",
+          "\\specialcell{HW - WH \\\\ (6)}"
+          )
+
 dim(row1) <- c(1,7)
-row2 <- c("Panel A: Parent’s", " ", " ", " ", " ", " ", " ")
+row2 <- c("Panel A: Children's Sunthetic Parent", " ", " ", " ", " ", " ", " ")
 dim(row2) <- c(1,7)
 
 # parent's panel
@@ -365,7 +366,7 @@ dim(row8) <- c(1,7)
 
 # Panel C
 # Employment and Earnings
-row9 <- c("Panel C: Employment and Earnings ", " ", " ", " ", " ", " ", " ")
+row9 <- c("Panel C: Children's Employment and Earnings ", " ", " ", " ", " ", " ", " ")
 dim(row9) <- c(1,7)
 
 ## Man's employment
@@ -583,7 +584,7 @@ pvalues2 <- test2$std.error
 differences2 <- test2$estimate
 
 row13 <- c(
-  "Women’s Hourly Earnings",
+  "Women’s Log Hourly Earnings",
   paste0("\\specialcell{", 
[…]

### 2023-04-16 programs/13-table-five-regression.R

@@ -7,16 +7,16 @@ IndividualData <- read_dta(file.path(datasets,"BySexAnalysisData.dta")) |>
   mutate(Education = as.numeric(Education)) |> 
   filter((WW == 1 & Hispanic_ID == 0) | WH == 1 | HW == 1 | HH == 1)
 
-ParentDummies <- c("WH", "HW", "HH", "uhrsworkly")
+ParentDummies <- c("HW", "uhrsworkly")
               
 # By generation
 reg1 <- list(
   "\\specialcell{(1) \\\\ Log annual earnings}" = feols(lninctot_1999 ~ .[ParentDummies] | year + age + statefip, vcov = ~statefip,
                                                             data = IndividualData |> 
-                                                          filter(sex == 1 & FTFY == 1 & Self_employed == 0)),
+                                                          filter(sex == 1 & FTFY == 1 & Self_employed == 0 & (HW == 1 | WH == 1))),
   "\\specialcell{(2) \\\\  Log annual earnings}" = feols(lninctot_1999 ~ .[ParentDummies] | year + age + educ + statefip, vcov = ~statefip,
                                                          data = IndividualData |> 
-                                                           filter(sex == 1 & FTFY == 1 & Self_employed == 0))
+                                                           filter(sex == 1 & FTFY == 1 & Self_employed == 0 & (HW == 1 | WH == 1)))
   
 )
 
@@ -24,35 +24,35 @@ reg1 <- list(
 # as a row
 
 
-test1 <- tidy(glht(reg1[[1]], linfct = c("HW - WH = 0")))
-differences1 <- case_when(abs(test1$estimate/test1$std.error)   >= 2.58  ~ paste0(sprintf("%2.5s",round_tidy(test1$estimate, digits = 2)), "***"),
-          abs(test1$estimate/test1$std.error)   <  2.58 & 
-            abs(test1$estimate/test1$std.error) >= 1.96  ~ paste0(sprintf("%2.5s",round_tidy(test1$estimate, digits = 2)), "**"),
-          abs(test1$estimate/test1$std.error)   <  1.96  &
-            abs(test1$estimate/test1$std.error) >= 1.645 ~ paste0(sprintf("%2.5s",round_tidy(test1$estimate, digits = 2)), "*"),
-          abs(test1$estimate/test1$std.error)   < 1.645  ~ round_tidy(test1$estimate, digits = 2)
-          )
[…]

### 2023-04-17 my_paper/Hadah-last-names.tex
table edits

@@
 \section{Data}\label{sec:data}
-\subsection{Data}\label{sec:data1}
 
@@
 
-I present the summary statistics of the four types of synthetic parents in table \ref{tab:c&p1}'s parent's panel.
+I present the summary statistics of the four types of synthetic parents in table \ref{tab:synth}'s parent's panel.
 WW couples have higher education, 13.05 years for husbands and 12.74 for wives.
@@
 
-In columns 5 and 6 of table \ref{tab:c&p1}, I report the differences in means between HH-WW households and HW-WH households.
+In columns 5 and 6 of table \ref{tab:synth}, I report the differences in means between HH-WW households and HW-WH households.
 An HH Hispanic husband attained 4.11 fewer years of education than a WW husband.
@@
 
-\subsection{Empirical Strategy}\label{sec:emp_model}
+\section{Empirical Strategy}\label{sec:emp_model}
 
@@
 
-\subsubsection{Estimating the effect of having a Hispanic last name}
+\subsection{Estimating the effect of having a Hispanic last name}
 
@@
 The equation for this strategy is written as follows:
+
 \begin{equation} \label{eq:1a}
@@
 
-\subsubsection{Estimating the effect of identifying as Hispanic}
+\subsection{Estimating the effect of identifying as Hispanic}
 
@@
 \begin{equation} \label{eq:iden}
-Y_{ist} = \beta_{1} HW_{ist} + \beta_{2} Hispanic_i + \beta_{3} HW_{ist} \cdot Hispanic_i X_{ist} \pi + \gamma_{t} + \lambda_s + \phi_{ist}
+Y_{ist} = \beta_{1} HW_{ist} + \beta_{2} Hispanic_i + \beta_{3} HW_{ist} \cdot Hispanic_i + X_{ist} \pi + \gamma_{t} + \lambda_s + \phi_{ist}
 \end{equation}
@@
 
-\subsection{Results}\label{sec:results1}
+\section{Results}\label{sec:results}
 
@@
 
-\subsection{The effect of having a Hispanic last name on labor market outcomes}
+\subsection{The Effect of Having a Hispanic Last Name and Hispanic Identity on Labor Market Outcomes}
 
@@
 The sample in this analysis includes full-time, full-year, and waged and salaried men, and I control for hours worked and age and include year fixed effects (FE).
+I also restrict the sample to children of HW and WH (omitted) parents.
 Column one in table \ref{tab:lastnamereg} is the results without controlling for education.
-Overall, there is big Hispanic--White earnings gaps.
-Children of White father-Hispanic mothers (WH) earn 14\% less than children of White fathers-White mothers (WW) families.
-The gap between children of Hispanic father-White mothers (HW) and WW children is 20\%.
-The gap between children of Hispanic father-Hispanic mothers (HH) and WW children is 24\%.
-Therefore, Hispanic children, inter-ethnic or children of homogeneous HH marriages, face a substantial earnings gap.
+Overall, an inter-ethnic with a Hispanic last name earns 6 percentage points less than an inter-ethnic with a White last name.
+This gap, however, could be entirely explained by educational differences.
+An inter-ethnic with a Hispanic last name earns 6 percentage points less than an inter-ethnic with a White last name, but the result is statistically insignificant.
 
-The difference between the estimates of WH and HW would capture the effect of having a Hispanic last name.
-A WH child will have a White last name since their father is White, while an HW child will have a Hispanic last name like their Hispanic father.
-When I take the difference between HW and WH, I find a statistically significant gap of 6 percentage points, which means that an inter-ethnic person with a Hispanic last name earns 6 percentage points less than a similar inter-ethnic person but with a White last name.
+I provide the results to the estimation of equation \ref{eq:iden} in table \ref{tab:identreg}.
+The sample in this analysis includes full-time, full-year, and waged and salaried men, and I control for hours worked and age and include year fixed effects (FE).
+I also restrict the sample to children of HW and WH (omitted) parents.
+Column one in table \ref{tab:identreg} is the results without controlling for education.
+Overall, an inter-ethnic that identifies as Hispanic earns 11 percentage points less than an inter-ethnic that does not.
+This gap, however, could be entirely explained by educational differences.
+An inter-ethnic that identifies as Hispanic earns 6 percentage points less than an inter-ethnic that does not, but the result is statistically insignificant.
 
-The literature on the earnings gaps shows that educational differences could explain a big part of these gaps \citep{duncan2006hispanics, duncan2017complexity, duncan2018identifying, duncan2020new}.
-Therefore, I run the equation \ref{eq:1a}, but this time I control for education.
-The regression results are in column two of table \ref{tab:lastnamereg}.
-The sample includes full-time, full-year, and waged and salaried men, and I control for hours worked and age and include year fixed effects (FE).
-The gap between WH and WW is 9\%, a reduction of 35\% from the baseline after controlling for education.
-The earning gaps between HW and HH, and WW were respectively reduced by 45\% and 50\%.
-After controlling for education, an HW man earns 11\% less than a WW man, and an HH man earns 12\% less.
-The effect of having a Hispanic last name on earnings decreased from 6 to 2 percentage points and became statistically insignificant after controlling for education.
-
-Therefore, Hispanics face big earnings gaps compared to non-Hispanic Whites.
-By comparing the children of inter-ethnic marriages, I could study the impact of having a Hispanic last name on earnings.
-Having a Hispanic last name carries a cost in the labor market.
-A person with a Hispanic last name earns 6 percentage points less than someone with a White name.
-However, educational differences between Hispanics and Whites could explain more than 50\% of the earning gaps and all of the Hispanic last name costs.
-
-\subsection{The effect of identifying as Hispanic on labor market outcomes}
-
-In this section, I present the results of equation \ref{eq:iden} that estimates the effect of identifying as Hispanic on annual log earnings.
-I show the results in table \ref{tab:identreg}.
-The sample is restricted to full-time full-year (FTFY) salaried and waged men.
-I control for age and hours worked and include year fixed effects.
-
-I run the estimating equation without controlling for education in table \ref{tab:identreg} column 1.
-I find that those that identify as Hispanic face a big earnings gap compared to those that do not.
-A WH man that identifies as Hispanic earns 14\% less than a WW man.
-An HW man who identifies as Hispanic makes 18\% less than a WW man, and an HH man who is determined as Hispanic earns 10\% less than a WW man.
-Comparing the coefficients of the interaction terms to those of the $WH_{ist}$, $HW_{ist}$, and $HH_{ist}$ shows that members of the three groups that do not identify as Hispanic earn about the same as WW.
-Moreover, the difference between $HW_{ist} \cdot Hispanic_{ist}$  and $WH_{ist} \cdot Hispanic_{ist}$ captures the effect of having a Hispanic last name and identifying as Hispanic compared to individuals who identify as Hispanic but have a White last name.
-A person with a Hispanic last name that identifies as Hispanic earns 3 percentage points less than a person that identifies as Hispanic but has a White last name.
-The difference between the two interaction terms is statistically insignificant.
-
-In table  \ref{tab:identreg} column 2, I show the results of equation \ref{eq:iden} while controlling for education.
-Education explains all the identity costs for members of WH and HH.
-Also, controlling for education decreases the earnings gap for identifying as Hispanic and being a member of HW by 33\%.
-An HW member that identifies as Hispanic earns 12\% less than a WW man.
-Furthermore, the difference between $HW_{ist} \cdot Hispanic_{ist}$  and $WH_{ist} \cdot Hispanic_{ist}$ increases to 4 percentage points but remains statistically insignificant.
-
-\subsection{Conclusion}\label{sec:con1}
+\section{Conclusion}\label{sec:con1}
 
@@
 When I control for education, the last name effect decreases to a statistically insignificant 2\% earnings gap.
-
-I also find that men who identify as Hispanic earn significantly less than those who do not, even when I control for ancestry and education.
-The gaps between WH, HW, and HH that identify as Hispanic and WW are, respectively, 14\%, 18\%, and 10\%.
-Controlling for education explains a big part of the gap.
-The gaps between WH, HW, and HH that identify as Hispanic and WW after controlling for education are 8\%, 12\%, and 6\%.
 
@@
 
-\input{tables/tab03-summary-stats}
+\input{tables/tab03b-summary-stats}
+
+\newpage
+
+\input{tables/tab03a-summary-stats}
 

### 2023-04-17 my_paper/tables/tab03b-summary-stats.tex
table edits

\begin{landscape}
\begin{ThreePartTable}
\begin{TableNotes}
\item[1] Source: The 1970-1990 Census for synthetic parents, and 1994-2019 Current Population Surveys (CPS) for children's outcomes
\item[2] The data is restricted to native-born United States citizens between 1994 and 2019 who are also White and between the ages of 25 and 40. I identify the ethnicity of a person's parents through the parent's place of birth. A parent is Hispanic if they were born in a Spanish-speaking country. A parent is White if they were born in the United States.
\end{TableNotes}
\begin{longtable}[t]{>{\raggedright\arraybackslash}p{5cm}cccccc}
\caption{Summary statistics of children's outcomes using parent's place of birth \label{tab:c&p1}}\\
\toprule
\multicolumn{1}{c}{ } & \multicolumn{4}{c}{Father's and Mother's Ethnicities} & \multicolumn{2}{c}{Differences} \\
\cmidrule(l{3pt}r{3pt}){2-5} \cmidrule(l{3pt}r{3pt}){6-7}
Variables & \specialcell{White \\ White \\ (WW) \\ (1)} & \specialcell{White \\ Hispanic \\ (WH) \\ (2)} & \specialcell{Hispanic \\ White \\ (HW) \\ (3)} & \specialcell{Hispanic \\ Hispanic \\ (HH) \\ (4)} & \specialcell{HH - WW \\ (5)} & \specialcell{HW - WH \\ (6)}\\
\midrule
\endfirsthead
\caption[]{Summary statistics of children's outcomes using parent's place of birth  \textit{(continued)}}\\
\toprule
Variables & \specialcell{White \\ White \\ (WW) \\ (1)} & \specialcell{White \\ Hispanic \\ (WH) \\ (2)} & \specialcell{Hispanic \\ White \\ (HW) \\ (3)} & \specialcell{Hispanic \\ Hispanic \\ (HH) \\ (4)} & \specialcell{HH - WW \\ (5)} & \specialcell{HW - WH \\ (6)}\\
\midrule
\endhead
\midrule
\multicolumn{7}{r@{}}{\textit{(Continued on Next Page...)}}\
\endfoot
\bottomrule
\insertTableNotes
\endlastfoot
\textbf{Panel A: Children's Education} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
\hspace{1em}Men’s education (Total Years) & \specialcell{13.91\\(2.39)} & \specialcell{13.58\\(2.35)} & \specialcell{13.21\\(2.32)} & \specialcell{12.91\\(2.26)} & \specialcell{-1.00***\\(0.01)} & \specialcell{-0.36**\\(0.03)}\\
\hspace{1em}Women’s education (Total Years) & \specialcell{14.29\\(2.41)} & \specialcell{13.87\\(2.47)} & \specialcell{13.42\\(2.35)} & \specialcell{13.27\\(2.37)} & \specialcell{-1.01***\\(0.01)} & \specialcell{-0.46**\\(0.03)}\\
\textbf{Panel B: Children's Employment and Earnings} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
\hspace{1em}Men’s Unemployment Rate & \specialcell{0.05\\(0.78)} & \specialcell{0.06\\(0.77)} & \specialcell{0.08\\(0.74)} & \specialcell{0.07\\(0.74)} & \specialcell{0.02***\\(0.00)} & \specialcell{0.02***\\(0.00)}\\
\addlinespace
\hspace{1em}Women’s Unemployment Rate & \specialcell{0.04\\(0.8)} & \specialcell{0.05\\(0.22)} & \specialcell{0.07\\(0.75)} & \specialcell{0.06\\(0.76)} & \specialcell{0.02***\\(0.00)} & \specialcell{0.02***\\(0.00)}\\
\hspace{1em}Men’s Log Hourly Earnings & \specialcell{2.48\\(0.45)} & \specialcell{2.41\\(0.45)} & \specialcell{2.4\\(0.44)} & \specialcell{2.41\\(0.43)} & \specialcell{-0.07***\\(0.01)} & \specialcell{-0.00**\\(0.02)}\\
\hspace{1em}Women’s Log Hourly Earnings & \specialcell{2.33\\(0.49)} & \specialcell{2.33\\(0.46)} & \specialcell{2.27\\(0.45)} & \specialcell{2.3\\(0.41)} & \specialcell{-0.02***\\(0.01)} & \specialcell{-0.06**\\(0.02)}\\
\hspace{1em}Men’s Log Annual Earnings & \specialcell{10.25\\(1.01)} & \specialcell{10.08\\(1.05)} & \specialcell{10.04\\(1.06)} & \specialcell{10.01\\(1.04)} & \specialcell{-0.25**\\(0.01)} & \specialcell{-0.04**\\(0.03)}\\
\hspace{1em}Women’s Log Annual Earnings & \specialcell{9.46\\(1.78)} & \specialcell{9.54\\(1.64)} & \specialcell{9.47\\(1.57)} & \specialcell{9.53\\(1.52)} & \specialcell{0.07**\\(0.02)} & \specialcell{-0.07*\\(0.06)}\\
\addlinespace
\textbf{Panel C: Children's Hispanic Identity} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
\hspace{1em}Men & \specialcell{0.05} & \specialcell{0.81} & \specialcell{0.88} & \specialcell{0.97} &  & \\
\hspace{1em}Women & \specialcell{0.06} & \specialcell{0.85} & \specialcell{0.87} & \specialcell{0.97} &  & \\*
\end{longtable}
\end{ThreePartTable}
\end{landscape}

### 2023-04-17 output/tables/tab04-summary-stats.tex
push project to github
table edits

--- 2023-01-20 my_paper/tables/tab03-summary-stats.tex
@@
-\begin{table}[!h]
-
-\caption{Children's outcome using parent's place of birth \label{tab:c&p1}}
-\centering
-\resizebox{\linewidth}{!}{
-\begin{threeparttable}
-\begin{tabular}[t]{lcccccc}
+\begin{landscape}
+\begin{ThreePartTable}
+\begin{TableNotes}
+\item[1] Source: The 1970-1990 Census for synthetic parents, and 1994-2019 Current Population Surveys (CPS) for children's outcomes
+\item[2] The data is restricted to native-born United States citizens between 1994 and 2019 who are also White and between the ages of 25 and 40.
+I identify the ethnicity of a person's parents through the parent's place of birth.
+A parent is Hispanic if they were born in a Spanish-speaking country.
+A parent is White if they were born in the United States.
+\end{TableNotes}
+\begin{longtable}[t]{>{\raggedright\arraybackslash}p{5cm}cccccc}
+\caption{Summary statistics of outcomes using parent's place of birth only for those that self-identify as Hispanic \label{tab:c&p2}}\\
 \toprule
@@
 \cmidrule(l{3pt}r{3pt}){2-5} \cmidrule(l{3pt}r{3pt}){6-7}
-Variables & \specialcell{White Father \\ White Mother \\ (WW) \\ (i)} & \specialcell{White Father \\ Hispanic Mother \\ (WH) \\ (ii)} & \specialcell{Hispanic Father \\ White Mother \\ (HW) \\ (iii)} & \specialcell{Hispanic Father \\ Hispanic Mother \\ (HH) \\ (iv)} & \specialcell{HH - WW \\ (v)} & \specialcell{HW - WH \\ (vi)}\\
+Variables & \specialcell{White \\ White \\ (WW) \\ (1)} & \specialcell{White \\ Hispanic \\ (WH) \\ (2)} & \specialcell{Hispanic \\ White \\ (HW) \\ (3)} & \specialcell{Hispanic \\ Hispanic \\ (HH) \\ (4)} & \specialcell{HH - WW \\ (5)} & \specialcell{HW - WH \\ (6)}\\
 \midrule
-\textbf{Panel A: Parent’s} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
-\hspace{1em}Husband’seducation (Total Years) & \specialcell{13.05\\(2.44)} & \specialcell{12.32\\(3.33)} & \specialcell{10.65\\(4.39)} & \specialcell{8.93\\(4.41)} & \specialcell{-4.11\\(0.02)} & \specialcell{-1.67\\(0.04)}\\
-\hspace{1em}Wife’seducation (Total Years) & \specialcell{12.74\\(2.12)} & \specialcell{11.03\\(3.92)} & \specialcell{11.54\\(3.12)} & \specialcell{8.6\\(4.13)} & \specialcell{-4.13\\(0.02)} & \specialcell{0.51\\(0.04)}\\
-\hspace{1em}Total Household seducation (Total Years) & \specialcell{25.78\\(4.08)} & \specialcell{23.35\\(6.51)} & \specialcell{22.19\\(6.69)} & \specialcell{17.54\\(7.83)} & \specialcell{-4.13\\(0.02)} & \specialcell{0.51\\(0.04)}\\
-\textbf{Panel B: Education} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
+\endfirsthead
+\caption[]{Summary statistics of outcomes using parent's place of birth only for those that self-identify as Hispanic  \textit{(continued)}}\\
+\toprule
+Variables & \specialcell{White \\ White \\ (WW) \\ (1)} & \specialcell{White \\ Hispanic \\ (WH) \\ (2)} & \specialcell{Hispanic \\ White \\ (HW) \\ (3)} & \specialcell{Hispanic \\ Hispanic \\ (HH) \\ (4)} & \specialcell{HH - WW \\ (5)} & \specialcell{HW - WH \\ (6)}\\
+\midrule
+\endhead
+\midrule
+\multicolumn{7}{r@{}}{\textit{(Continued on Next Page...)}}\
+\endfoot
+\bottomrule
+\insertTableNotes
+\endlastfoot
+\textbf{Panel A: Children's Education} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
+\hspace{1em}Men’s education (Total Years) & \specialcell{12.97\\(2.15)} & \specialcell{13.45\\(2.37)} & \specialcell{13.13\\(2.27)} & \specialcell{12.89\\(2.25)} & \specialcell{-0.08**\\(0.01)} & \specialcell{-0.32**\\(0.03)}\\
+\hspace{1em}Women’s education (Total Years) & \specialcell{13.23\\(2.25)} & \specialcell{13.75\\(2.41)} & \specialcell{13.32\\(2.34)} & \specialcell{13.26\\(2.37)} & \specialcell{0.03**\\(0.01)} & \specialcell{-0.43**\\(0.03)}\\
+\textbf{Panel B: Children's Employment and Earnings} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
+\hspace{1em}Men’s Unemployment Rate & \specialcell{0.07\\(0.74)} & \specialcell{0.06\\(0.77)} & \specialcell{0.08\\(0.73)} & \specialcell{0.07\\(0.74)} & \specialcell{-0.00***\\(0.00)} & \specialcell{0.02***\\(0.00)}\\
 \addlinespace
-\hspace{1em}Men’s education (Total Years) & \specialcell{13.91\\(2.39)} & \specialcell{13.58\\(2.35)} & \specialcell{13.21\\(2.32)} & \specialcell{12.91\\(2.26)} & \specialcell{-1.00\\(0.01)} & \specialcell{-0.36\\(0.03)}\\
-\hspace{1em}Women’s education (Total Years) & \specialcell{14.29\\(2.41)} & \specialcell{13.87\\(2.47)} & \specialcell{13.42\\(2.35)} & \specialcell{13.27\\(2.37)} & \specialcell{-1.01\\(0.01)} & \specialcell{-0.46\\(0.03)}\\
-\textbf{Panel C: Employment and Earnings} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
-\hspace{1em}Men’s Employment Rate & \specialcell{0.95\\(0.22)} & \specialcell{0.94\\(0.23)} & \specialcell{0.92\\(0.26)} & \specialcell{0.93\\(0.26)} & \specialcell{-0.02\\(0)} & \specialcell{-0.02\\(0)}\\
-\hspace{1em}Women’s Employment Rate & \specialcell{0.96\\(0.2)} & \specialcell{0.95\\(0.22)} & \specialcell{0.93\\(0.25)} & \specialcell{0.94\\(0.24)} & \specialcell{-0.02\\(0)} & \specialcell{-0.02\\(0)}\\
-\addlinespace
-\hspace{1em}Men’s Log Hourly Earnings & \specialcell{2.48\\(0.45)} & \specialcell{2.41\\(0.45)} & \specialcell{2.4\\(0.44)} & \specialcell{2.41\\(0.43)} & \specialcell{-0.07\\(0.01)} & \specialcell{-0.00\\(0.02)}\\
-\hspace{1em}Women’s Hourly Earnings & \specialcell{2.33\\(0.49)} & \specialcell{2.33\\(0.46)} & \specialcell{2.27\\(0.45)} & \specialcell{2.3\\(0.41)} & \specialcell{-0.02\\(0.01)} & \specialcell{-0.06\\(0.02)}\\
-\hspace{1em}Men’s Log Annual Earnings & \specialcell{10.25\\(1.01)} & \specialcell{10.08\\(1.05)} & \specialcell{10.04\\(1.06)} & \specialcell{10.01\\(1.04)} & \specialcell{-0.25\\(0.01)} & \specialcell{-0.04\\(0.03)}\\
-\hspace{1em}Women’s Hourly Earnings & \specialcell{9.46\\(1.78)} & \specialcell{9.54\\(1.64)} & \specialcell{9.47\\(1.57)} & \specialcell{9.53\\(1.52)} & \specialcell{0.07\\(0.02)} & \specialcell{-0.07\\(0.06)}\\
-\textbf{Panel D: Hispanic Identity} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
-\addlinespace
-\hspace{1em}Men & \specialcell{0.05} & \specialcell{0.81} & \specialcell{0.88} & \specialcell{0.97} &  & \\
-\hspace{1em}Women & \specialcell{0.06} & \specialcell{0.85} & \specialcell{0.87} & \specialcell{0.97} &  & \\
-\bottomrule
-\end{tabular}
-\begin{tablenotes}
-\item[1] The data is restricted to native-born United States citizens between 1994 and 2019 who are also White and between the ages of 25 and 40.
-I identify the ethnicity of a person's parents through the parent's place of birth.
-A parent is Hispanic if they were born in a Spanish-speaking country.
-A parent is White if they were born in the United States.
-\item[2] In each column, I present the average statistics of the different types of people based on the ethnicities of their parents.
-In column one, I show the summary statistics of children of White fathers and White mothers.
-In column two, I present the summary statistics of children of White fathers and Hispanic mothers.
-In column three, I show the summary statistics of children of Hispanic fathers and White mothers.
-In column four, I present the summary statistics of children of Hispanic fathers and mothers.
-\item[3] Columns five and six have data on the HH--WW gaps (column five) and the HW--WH gaps (column six).
-\end{tablenotes}
-\end{threeparttable}}
-\end{table}
+\hspace{1em}Women’s Unemployment Rate & \specialcell{0.06\\(0.75)} & \specialcell{0.06\\(0.23)} & \specialcell{0.07\\(0.74)} & \specialcell{0.06\\(0.76)} & \specialcell{-0.00***\\(0.00)} & \specialcell{0.02***\\(0.00)}\\
+\hspace{1em}Men’s Log Hourly Earnings & \specialcell{2.4\\(0.44)} & \specialcell{2.41\\(0.45)} & \specialcell{2.4\\(0.44)} & \specialcell{2.41\\(0.43)} & \specialcell{0.01***\\(0.01)} & \specialcell{-0.01**\\(0.02)}\\
+\hspace{1em}Women’s Log Hourly Earnings & \specialcell{2.26\\(0.43)} & \specialcell{2.32\\(0.45)} & \specialcell{2.27\\(0.45)} & \specialcell{2.3\\(0.41)} & \specialcell{0.04***\\(0.01)} & \specialcell{-0.05**\\(0.02)}\\
+\hspace{1em}Men’s Log Annual Earnings & \specialcell{10.02\\(1.02)} & \specialcell{10.06\\(1.06)} & \specialcell{10.03\\(0.99)} & \specialcell{10\\(1.04)} & \specialcell{-0.02**\\(0.01)} & \specialcell{-0.03**\\(0.04)}\\
+\hspace{1em}Women’s Log Annual Earnings & \specialcell{9.44\\(1.59)} & \specialcell{9.55\\(1.59)} & \specialcell{9.47\\(1.55)} & \specialcell{9.52\\(1.52)} & \specialcell{0.08**\\(0.02)} & \specialcell{-0.08*\\(0.05)}\\*
+\end{longtable}
+\end{ThreePartTable}
+\end{landscape}

### 2023-04-17 output/tables/tab05-regression.tex
some edits
table edits

@@
 \begin{threeparttable}
-\begin{tabular}[t]{lcc}
+\begin{tabular}[t]{lcccc}
 \toprule
-  & \specialcell{(1) \\ Log annual earnings} & \specialcell{(2) \\  Log annual earnings}\\
+  & \specialcell{(1) \\ Log annual \\ earnings} & \specialcell{(2) \\ Log annual \\ earnings} & \specialcell{(3) \\ Log annual \\ earnings} & \specialcell{(4) \\  Log annual \\ earnings}\\
 \midrule
-$WH_{i}$ & \num{-0.09}*** & \num{-0.04}**\\
- & (\num{0.02}) & (\num{0.02})\\
-$HW_{i}$ & \num{-0.14}*** & \num{-0.06}***\\
- & (\num{0.02}) & (\num{0.01})\\
-$HH_{i}$ & \num{-0.16}*** & \num{-0.06}***\\
- & (\num{0.01}) & (\num{0.01})\\
+$HW_{i}$ & \num{-0.06}** & \num{-0.07}** & \num{-0.05}** & \num{-0.02}\\
+ & (\num{0.03}) & (\num{0.03}) & (\num{0.02}) & (\num{0.02})\\
+Constant & \num{10.40}*** & \num{9.49}*** &  & \\
+ & (\num{0.04}) & (\num{0.07}) &  & \\
 \midrule
-$HW_{i} - WH_{i}$ & -0.05** & -0.02\\
- & (0.02) & (0.02)\\
-\midrule
-Observations & \num{129359} & \num{129359}\\
-\textit{Controlling for:} &  & \\
-Hours Worked & X & X\\
-Year & X & X\\
-Age & X & X\\
-Education &  & X\\
+\textit{Controlling for:} &  &  &  & \\
+Hours Worked &  & X & X & X\\
+State FE &  &  & X & X\\
+Year FE &  &  & X & X\\
+Age &  &  & X & X\\
+Education &  &  &  & X\\
+Observations & \num{2423} & \num{2423} & \num{2423} & \num{2423}\\
 \bottomrule
-\multicolumn{3}{l}{\rule{0pt}{1em}* p $<$ 0.1, ** p $<$ 0.05, *** p $<$ 0.01}\\
+\multicolumn{5}{l}{\rule{0pt}{1em}* p $<$ 0.1, ** p $<$ 0.05, *** p $<$ 0.01}\\
 \end{tabular}

### 2023-04-17 output/tables/tab06-regression.tex
table edits

@@
 \begin{threeparttable}
-\begin{tabular}[t]{lcc}
+\begin{tabular}[t]{lcccc}
 \toprule
-  & \specialcell{(1) \\ Log annual earnings} & \specialcell{(2) \\  Log annual earnings}\\
+  & \specialcell{(1) \\ Log annual \\ earnings} & \specialcell{(2) \\ Log annual \\ earnings} & \specialcell{(3) \\ Log annual \\ earnings} & \specialcell{(4) \\  Log annual \\ earnings}\\
 \midrule
-$HW_{i} \times Hispanic_{i}$ & \num{-0.02} & \num{-0.02}\\
- & (\num{0.08}) & (\num{0.07})\\
-$Hispanic_{i}$ & \num{-0.11}** & \num{-0.06}\\
- & (\num{0.05}) & (\num{0.05})\\
-$HW_{i}$ & \num{-0.03} & \num{0.00}\\
- & (\num{0.08}) & (\num{0.06})\\
+$HW_{i} \times Hispanic_{i}$ & \num{-0.01} & \num{0.04} & \num{-0.02} & \num{-0.02}\\
+ & (\num{0.09}) & (\num{0.09}) & (\num{0.08}) & (\num{0.07})\\
+$Hispanic_{i}$ & \num{-0.16}** & \num{-0.15}** & \num{-0.11}** & \num{-0.06}\\
+ & (\num{0.08}) & (\num{0.07}) & (\num{0.05}) & (\num{0.05})\\
+$HW_{i}$ & \num{-0.05} & \num{-0.09} & \num{-0.03} & \num{0.00}\\
+ & (\num{0.09}) & (\num{0.08}) & (\num{0.08}) & (\num{0.06})\\
+Constant & \num{10.54}*** & \num{9.63}*** &  & \\
+ & (\num{0.08}) & (\num{0.11}) &  & \\
 \midrule
-State FE & X & X\\
-Year FE & X & X\\
-\textit{Controlling for:} &  & \\
-Hours Worked & X & X\\
-Age & X & X\\
-Education &  & X\\
-Observations & \num{2423} & \num{2423}\\
+$HW_{i} \times Hispanic_{i} + Hispanic_{i}$ & -0.17** & -0.11 & -0.13* & -0.08\\
+ & (0.08) & (0.07) & (0.07) & (0.06)\\
+\midrule
+State FE &  &  & X & X\\
+Year FE &  &  & X & X\\
+\textit{Controlling for:} &  &  &  & \\
+Hours Worked &  & X & X & X\\
+Age &  &  & X & X\\
+Education &  &  &  & X\\
+Observations & \num{2423} & \num{2423} & \num{2423} & \num{2423}\\
 \bottomrule
-\multicolumn{3}{l}{\rule{0pt}{1em}* p $<$ 0.1, ** p $<$ 0.05, *** p $<$ 0.01}\\
+\multicolumn{5}{l}{\rule{0pt}{1em}* p $<$ 0.1, ** p $<$ 0.05, *** p $<$ 0.01}\\
 \end{tabular}

### 2023-04-17 programs/11a-table-three-summary-stats.R

@@ -310,6 +310,7 @@ knitr::kable(Table_rows, "latex", align = "lcccccc",
   ) |> 
   add_header_above(c(" " = 1, "Father's and Mother's Ethnicities" = 4,
                      "Differences" = 2)) |> 
+  column_spec(1, width = "5cm") |> 
   save_kable(file.path(tables_wd,"tab03a-summary-stats.tex")) |> 
   save_kable(file.path(thesis_tabs,"tab03a-summary-stats.tex"))  |> 
   save_kable(file.path("/Users/hhadah/Documents/GiT/my_thesis/tables","tab03a-summary-stats.tex"))

### 2023-04-17 programs/11b-table-three-summary-stats.R

@@ -1,232 +1,29 @@
 # summary stats
 
-# table four
-# only hispanics
+# table three
+
+library(Hmisc)
 
 # date: January 19th, 2023
 
 IndividualData <- read_dta(file.path(datasets,"BySexAnalysisData.dta")) |> 
-  mutate(Education = as.numeric(Education)) |> 
-  filter(Hispanic_ID == 1)
+  mutate(Education = as.numeric(Education))
 ParentData     <- read_dta(file.path(datasets,"ParentDataFull.dta")) |> 
   mutate(Tot_ed = HusbandEducation + WifeEducation)
 
 row1 <- c("Variables", 
-          "\\specialcell{White Father \\\\ White Mother \\\\ (WW) \\\\ (1)}",
-          "\\specialcell{White Father \\\\ Hispanic Mother \\\\ (WH) \\\\ (2)}",
-          "\\specialcell{Hispanic Father \\\\ White Mother \\\\ (HW) \\\\ (3)}",
-          "\\specialcell{Hispanic Father \\\\ Hispanic Mother \\\\ (HH) \\\\ (4)}",
+          "\\specialcell{White \\\\ White \\\\ (WW) \\\\ (1)}",
+          "\\specialcell{White \\\\ Hispanic \\\\ (WH) \\\\ (2)}",
+          "\\specialcell{Hispanic \\\\ White \\\\ (HW) \\\\ (3)}",
+          "\\specialcell{Hispanic \\\\ Hispanic \\\\ (HH) \\\\ (4)}",
           "\\specialcell{HH - WW \\\\ (5)}",
           "\\specialcell{HW - WH \\\\ (6)}"
           )
-
 dim(row1) <- c(1,7)
-row2 <- c("Panel A: Children's Sunthetic Parent", " ", " ", " ", " ", " ", " ")
-dim(row2) <- c(1,7)
 
-# parent's panel
-
-## Husband's Education
-HusbandEducation <- ParentData |> 
-  group_by(CoupleType) |> 
-  summarise(HusEducation   = weighted.mean(HusbandEducation, na.rm = T, hhwt),
-            HusEducationSD = weighted.sd(HusbandEducation, na.rm = T, hhwt))
-
-model <- lm(HusbandEducation ~ 0 + WW + WM + MW + MM, 
-            data = ParentData, 
-            weights = hhwt)
-test1 <- tidy(glht(model, linfct = c("MM - WW = 0")))
[…]

### 2023-04-17 programs/13-table-five-regression.R

@@ -11,10 +11,16 @@ ParentDummies <- c("HW", "uhrsworkly")
               
 # By generation
 reg1 <- list(
-  "\\specialcell{(1) \\\\ Log annual earnings}" = feols(lninctot_1999 ~ .[ParentDummies] | year + age + statefip, vcov = ~statefip,
+  "\\specialcell{(1) \\\\ Log annual \\\\ earnings}" = feols(lninctot_1999 ~ HW, vcov = ~statefip,
                                                             data = IndividualData |> 
                                                           filter(sex == 1 & FTFY == 1 & Self_employed == 0 & (HW == 1 | WH == 1))),
-  "\\specialcell{(2) \\\\  Log annual earnings}" = feols(lninctot_1999 ~ .[ParentDummies] | year + age + educ + statefip, vcov = ~statefip,
+  "\\specialcell{(2) \\\\ Log annual \\\\ earnings}" = feols(lninctot_1999 ~ .[ParentDummies], vcov = ~statefip,
+                                                            data = IndividualData |> 
+                                                          filter(sex == 1 & FTFY == 1 & Self_employed == 0 & (HW == 1 | WH == 1))),
+  "\\specialcell{(3) \\\\ Log annual \\\\ earnings}" = feols(lninctot_1999 ~ .[ParentDummies] | year + age + statefip, vcov = ~statefip,
+                                                            data = IndividualData |> 
+                                                          filter(sex == 1 & FTFY == 1 & Self_employed == 0 & (HW == 1 | WH == 1))),
+  "\\specialcell{(4) \\\\  Log annual \\\\ earnings}" = feols(lninctot_1999 ~ .[ParentDummies] | year + age + educ + statefip, vcov = ~statefip,
                                                          data = IndividualData |> 
                                                            filter(sex == 1 & FTFY == 1 & Self_employed == 0 & (HW == 1 | WH == 1)))
   
@@ -54,11 +60,11 @@ reg1 <- list(
 # row_diff <- as.data.frame(row_diff)
 # attr(row_diff, 'position') <- c(7:8)
 
-controling_for <-  c("\\textit{Controlling for:}", " ", "")
-dim(controling_for) <- c(1,3)
+controling_for <-  c("\\textit{Controlling for:}", " ", "", " ", "")
[…]

### 2023-04-17 programs/14-table-six-regression.R

@@ -20,10 +20,16 @@ ParentDummies <- c(#"WH:Hispanic_ID",
 
 # By generation
 reg1 <- list(
-  "\\specialcell{(1) \\\\ Log annual earnings}" = feols(lninctot_1999 ~ .[ParentDummies] | year + age + statefip, vcov = ~statefip,
+  "\\specialcell{(1) \\\\ Log annual \\\\ earnings}" = feols(lninctot_1999 ~ HW*Hispanic_ID, vcov = ~statefip,
                                                         data = IndividualData |> 
                                                           filter(sex == 1 & FTFY == 1 & Self_employed == 0 & (HW == 1 | WH == 1))),
-  "\\specialcell{(2) \\\\  Log annual earnings}" = feols(lninctot_1999 ~ .[ParentDummies] | year + age + educ + statefip, vcov = ~statefip,
+  "\\specialcell{(2) \\\\ Log annual \\\\ earnings}" = feols(lninctot_1999 ~ .[ParentDummies] , vcov = ~statefip,
+                                                        data = IndividualData |> 
+                                                          filter(sex == 1 & FTFY == 1 & Self_employed == 0 & (HW == 1 | WH == 1))),
+  "\\specialcell{(3) \\\\ Log annual \\\\ earnings}" = feols(lninctot_1999 ~ .[ParentDummies] | year + age + statefip, vcov = ~statefip,
+                                                        data = IndividualData |> 
+                                                          filter(sex == 1 & FTFY == 1 & Self_employed == 0 & (HW == 1 | WH == 1))),
+  "\\specialcell{(4) \\\\  Log annual \\\\ earnings}" = feols(lninctot_1999 ~ .[ParentDummies] | year + age + educ + statefip, vcov = ~statefip,
                                                          data = IndividualData |> 
                                                            filter(sex == 1 & FTFY == 1 & Self_employed == 0 & (HW == 1 | WH == 1) ))
   
@@ -33,57 +39,77 @@ reg1 <- list(
 # as a row
 
 
-# test1 <- tidy(glht(reg1[[1]], linfct = c("Hispanic_ID:HW - WH:Hispanic_ID = 0")))
[…]

### 2023-04-18 my_paper/tables/tab05-regression.tex
edits

@@
 \begin{threeparttable}
-\begin{tabular}[t]{lcc}
+\begin{tabular}[t]{lcccc}
 \toprule
-  & \specialcell{(1) \\ Log annual earnings} & \specialcell{(2) \\  Log annual earnings}\\
+  & \specialcell{(1) \\ Log annual \\ earnings} & \specialcell{(2) \\ Log annual \\ earnings} & \specialcell{(3) \\ Log annual \\ earnings} & \specialcell{(4) \\  Log annual \\ earnings}\\
 \midrule
-$HW_{i}$ & \num{-0.05}** & \num{-0.02}\\
- & (\num{0.02}) & (\num{0.02})\\
+$HW_{i}$ & \num{-0.06}** & \num{-0.07}** & \num{-0.05}** & \num{-0.02}\\
+ & (\num{0.03}) & (\num{0.03}) & (\num{0.02}) & (\num{0.02})\\
+Constant & \num{10.40}*** & \num{9.49}*** &  & \\
+ & (\num{0.04}) & (\num{0.07}) &  & \\
 \midrule
-State FE & X & X\\
-Year FE & X & X\\
-\textit{Controlling for:} &  & \\
-Hours Worked & X & X\\
-Age & X & X\\
-Education &  & X\\
-Observations & \num{2423} & \num{2423}\\
+\textit{Controlling for:} &  &  &  & \\
+Hours Worked &  & X & X & X\\
+State FE &  &  & X & X\\
+Year FE &  &  & X & X\\
+Age FE &  &  & X & X\\
+Education FE &  &  &  & X\\
+Observations & \num{2423} & \num{2423} & \num{2423} & \num{2423}\\
 \bottomrule
-\multicolumn{3}{l}{\rule{0pt}{1em}* p $<$ 0.1, ** p $<$ 0.05, *** p $<$ 0.01}\\
+\multicolumn{5}{l}{\rule{0pt}{1em}* p $<$ 0.1, ** p $<$ 0.05, *** p $<$ 0.01}\\
 \end{tabular}
 \begin{tablenotes}
-\item[1] \footnotesize{This table includes the estimation results of equation (\ref{eq:1a}).}
-\item[2] \footnotesize{HW is an indicator variable that is equal to 1 if a person is the child of a Hispanic-father and White-mother.}
-\item[3] \footnotesize{The sample is restricted to men working full-time full-year and are waged and salaried workers.}
-\item[4] \footnotesize{Column one has the regression results when controlling for hours worked, age, and fixed effects. Column two has the results after controlling for education.}
-\item[5] \footnotesize{Standard errors are clustered on the state level.}
+\item[1] {\setstretch{1.0}\footnotesize{This table includes the estimation results of equation (\ref{eq:1a}).}}
+\item[2] {\setstretch{1.0}\footnotesize{HW is an indicator variable that is equal to 1 if a person is the child of a Hispanic-father and White-mother.}}
+\item[3] {\setstretch{1.0}\footnotesize{The sample is restricted to men working full-time full-year and are waged and salaried workers.}}
+\item[4] {\setstretch{1.0}\footnotesize{Column one has the regression results when controlling for hours worked, age, and fixed effects. Column two has the results after controlling for education.}}
+\item[5] {\setstretch{1.0}\footnotesize{Standard errors are clustered on the state level.}}
 \end{tablenotes}

### 2023-04-18 output/tables/tab06-regression.tex
edits

@@
 \centering
-\resizebox{\linewidth}{!}{
 \begin{threeparttable}
@@
 \midrule
+\textit{Controlling for:} &  &  &  & \\
+Hours Worked &  & X & X & X\\
 State FE &  &  & X & X\\
 Year FE &  &  & X & X\\
-\textit{Controlling for:} &  &  &  & \\
-Hours Worked &  & X & X & X\\
-Age &  &  & X & X\\
-Education &  &  &  & X\\
+Age FE &  &  & X & X\\
+Education FE &  &  &  & X\\
 Observations & \num{2423} & \num{2423} & \num{2423} & \num{2423}\\
@@
 \end{tablenotes}
-\end{threeparttable}}
+\end{threeparttable}
 \end{table}

### 2023-04-18 programs/13-table-five-regression.R

@@ -87,8 +87,8 @@ gm <- tibble::tribble(
   ~raw,        ~clean,          ~fmt,
   "FE: statefip", "State FE", 0,
   "FE: year", "Year FE", 0,
-  "FE: age", "Age", 0,
-  "FE: educ", "Education", 0,
+  "FE: age", "Age FE", 0,
+  "FE: educ", "Education FE", 0,
   "std.error.type", "Standard Errors", 0,
   "nobs",      "Observations",             0,
   #"r.squared", "R squared", 3
@@ -124,14 +124,14 @@ regression_tab <- modelsummary(reg1, fmt = 2,
                                #gof_omit = 'DF|Deviance|R2|AIC|BIC|Log.Lik.|F|Std.Errors',
                                stars= c('***' = 0.01, '**' = 0.05, '*' = 0.1),
                                title = "Effect of Having Hispanic Last Name \\label{tab:lastnamereg}") %>%
-  kable_styling(bootstrap_options = c("hover", "condensed"), 
-                latex_options = c("scale_down", "HOLD_position")
+  kable_styling(
+                latex_options = c("scale_down", "HOLD_position", "footnotesize", "\\setlength{\\footnotesep}{1\\baselineskip}")
   ) %>%
-  footnote(number = c("\\\\footnotesize{This table includes the estimation results of equation (\\\\ref{eq:1a}).}",
-                      "\\\\footnotesize{HW is an indicator variable that is equal to 1 if a person is the child of a Hispanic-father and White-mother.}",
-                      "\\\\footnotesize{The sample is restricted to men working full-time full-year and are waged and salaried workers.}",
-                      "\\\\footnotesize{Column one has the regression results when controlling for hours worked, age, and fixed effects. Column two has the results after controlling for education.}",
-                      "\\\\footnotesize{Standard errors are clustered on the state level.}"
+  footnote(number = c("{\\\\setstretch{1.0}\\\\footnotesize{This table includes the estimation results of equation (\\\\ref{eq:1a}).}}",
+                      "{\\\\setstretch{1.0}\\\\footnotesize{HW is an indicator variable that is equal to 1 if a person is the child of a Hispanic-father and White-mother.}}",
+                      "{\\\\setstretch{1.0}\\\\footnotesize{The sample is restricted to men working full-time full-year and are waged and salaried workers.}}",
[…]

### 2023-04-20 my_paper/tables/tab06-regression.tex
add sig levels to sum tables and hilight defense columns
edit tables for typos

@@
-\begin{table}[!h]
+\begin{table}[H]
 
-\caption{Effect of Having Hispanic Last Name \label{tab:identreg}}
+\caption{Effect of Having Hispanic Last Name by Hispanic Identity\label{tab:identreg}}
 \centering
@@
 \begin{threeparttable}
-\begin{tabular}[t]{lcc}
+\begin{tabular}[t]{lcccc}
 \toprule
-  & \specialcell{(1) \\ Log annual earnings} & \specialcell{(2) \\  Log annual earnings}\\
+  & \specialcell{(1) \\ Log annual \\ earnings} & \specialcell{(2) \\ Log annual \\ earnings} & \specialcell{(3) \\ Log annual \\ earnings} & \specialcell{(4) \\  Log annual \\ earnings}\\
 \midrule
-$WH_{i} \times Hispanic_{i}$ & \num{-0.13}** & \num{-0.06}\\
- & (\num{0.05}) & \vphantom{2} (\num{0.05})\\
-$HW_{i} \times Hispanic_{i}$ & \num{-0.16}*** & \num{-0.11}**\\
- & (\num{0.05}) & \vphantom{1} (\num{0.05})\\
-$HH_{i} \times Hispanic_{i}$ & \num{-0.11}** & \num{-0.06}\\
- & (\num{0.05}) & \vphantom{2} (\num{0.04})\\
-$WH_{i}$ & \num{0.03} & \num{0.01}\\
- & (\num{0.05}) & \vphantom{1} (\num{0.04})\\
-$HW_{i}$ & \num{0.01} & \num{0.04}\\
- & (\num{0.05}) & (\num{0.05})\\
-$HH_{i}$ & \num{-0.06} & \num{0.00}\\
- & (\num{0.05}) & (\num{0.04})\\
+$HW_{ist} \times Hispanic_{ist}$ & \num{-0.01} & \num{0.04} & \num{-0.02} & \num{-0.02}\\
+ & (\num{0.09}) & (\num{0.09}) & (\num{0.08}) & (\num{0.07})\\
+$Hispanic_{ist}$ & \num{-0.16}** & \num{-0.15}** & \num{-0.11}** & \num{-0.06}\\
+ & (\num{0.08}) & (\num{0.07}) & (\num{0.05}) & (\num{0.05})\\
+$HW_{ist}$ & \num{-0.05} & \num{-0.09} & \num{-0.03} & \num{0.00}\\
+ & (\num{0.09}) & (\num{0.08}) & (\num{0.08}) & (\num{0.06})\\
+Constant & \num{10.54}*** & \num{9.63}*** &  & \\
+ & (\num{0.08}) & (\num{0.11}) &  & \\
 \midrule
-$HW_{i} \times Hispanic_{i} - WH_{i} \times Hispanic_{i}$ & -0.03 & -0.04\\
- & (0.07) & (0.07)\\
+$HW_{i} \times Hispanic_{ist} + HW_{ist}$ & -0.06** & -0.05** & -0.05** & -0.02\\
+ & (0.03) & (0.02) & (0.02) & (0.02)\\
 \midrule
-Observations & \num{129359} & \num{129359}\\
-\textit{Controlling for:} &  & \\
-Hours Worked & X & X\\
-Age & X & X\\
-Year FE & X & X\\
-Education &  & X\\
+\textit{Controlling for:} &  &  &  & \\
+Hours Worked &  & X & X & X\\
+State FE &  &  & X & X\\
+Year FE &  &  & X & X\\
+Age FE &  &  & X & X\\
+Education FE &  &  &  & X\\
+Observations & \num{2423} & \num{2423} & \num{2423} & \num{2423}\\
 \bottomrule
-\multicolumn{3}{l}{\rule{0pt}{1em}* p $<$ 0.1, ** p $<$ 0.05, *** p $<$ 0.01}\\
+\multicolumn{5}{l}{\rule{0pt}{1em}* p $<$ 0.1, ** p $<$ 0.05, *** p $<$ 0.01}\\
 \end{tabular}
@@
 \item[1] \footnotesize{This table includes the estimation results of equation (\ref{eq:iden}).}
-\item[2] \footnotesize{The four groups stand for White Husband White Wife (WW), White Husband Hispanic Wife (WH), Hispanic Husband White (HW), and Hispanic Husband Hispanic Wife (HH).
-$Hispanic_{i}$ is a dummy variable equal to one if a person identifies as Hispanic and zero otherwise.}
+\item[2] \footnotesize{HW is an indicator variable that is equal to 1 if a person is the child of a Hispanic-father and White-mother, and Hispanic is an indicator variable that is equal to one if a person self-identifies as Hispanic.}
 \item[3] \footnotesize{The sample is restricted to men working full-time full-year and are waged and salaried workers.}
-\item[4] \footnotesize{Column one has the regression results when controlling for hours worked, age, and years of fixed effects. Column two has the results after controlling for education.}
+\item[4] \footnotesize{Column one has the regression results when controlling for hours worked, age, education, year and state fixed effects.
+Column two has the results after controlling for education.}
 \end{tablenotes}

### 2023-04-20 output/tables/tab05-regression.tex
edit tables for typos

@@
 \midrule
-$HW_{i}$ & \num{-0.06}** & \num{-0.07}** & \num{-0.05}** & \num{-0.02}\\
+$HW_{ist}$ & \num{-0.06}** & \num{-0.07}** & \num{-0.05}** & \num{-0.02}\\
  & (\num{0.03}) & (\num{0.03}) & (\num{0.02}) & (\num{0.02})\\
@@
 Year FE &  &  & X & X\\
-Age &  &  & X & X\\
-Education &  &  &  & X\\
+Age FE &  &  & X & X\\
+Education FE &  &  &  & X\\
 Observations & \num{2423} & \num{2423} & \num{2423} & \num{2423}\\
@@
 \begin{tablenotes}
-\item[1] \footnotesize{This table includes the estimation results of equation (\ref{eq:1a}).}
-\item[2] \footnotesize{HW is an indicator variable that is equal to 1 if a person is the child of a Hispanic-father and White-mother.}
-\item[3] \footnotesize{The sample is restricted to men working full-time full-year and are waged and salaried workers.}
-\item[4] \footnotesize{Column one has the regression results when controlling for hours worked, age, and fixed effects. Column two has the results after controlling for education.}
-\item[5] \footnotesize{Standard errors are clustered on the state level.}
+\item[1] {\setstretch{1.0}\footnotesize{This table includes the estimation results of equation (\ref{eq:1a}).}}
+\item[2] {\setstretch{1.0}\footnotesize{HW is an indicator variable that is equal to 1 if a person is the child of a Hispanic-father and White-mother.}}
+\item[3] {\setstretch{1.0}\footnotesize{The sample is restricted to men working full-time full-year and are wage and salary workers.}}
+\item[4] {\setstretch{1.0}\footnotesize{Column one has the regression results when controlling for hours worked, age, education, year and state fixed effects.
+Column two has the results after controlling for education.}}
+\item[5] {\setstretch{1.0}\footnotesize{Standard errors are clustered on the state level.}}
 \end{tablenotes}
-\end{threeparttable}}
+\end{threeparttable}
 \end{table}

### 2023-04-20 programs/07-table-one-parent-types.R

@@ -108,7 +108,7 @@ knitr::kable(Table_cols, "latex", align = "lcccc",
   kable_styling(bootstrap_options = c("hover", "condensed"), 
                 latex_options = c("scale_down", "HOLD_position")) |> 
   footnote(number = c("Source: Current Population Surveys (CPS) 1994-2019",
-                      "The sample includes Whites, who are married, and are between the ages 25 and 40. Ethnicity of a person's parents are identified by the parent's place of birth. A parent is Hispanic if she/he was born in a Spanish-speaking country. A parent is White she/he was born in the United States."),
+                      "The sample includes Whites, who are married, and are between the ages 25 and 40. Ethnicity of a person's parents are identified by the parent's place of birth. A parent is Hispanic if she/he was born in a Spanish-speaking country. A parent is White if she/he was born in the United States."),
            footnote_as_chunk = F, title_format = c("italic"),
            escape = F, threeparttable = T
   ) |> 

### 2023-04-20 programs/10-table-two-couple-types.R

@@ -109,7 +109,7 @@ knitr::kable(Table_cols, "latex", align = "lcccc",
   kable_styling(bootstrap_options = c("hover", "condensed"), 
                 latex_options = c("scale_down", "HOLD_position")) |> 
   footnote(number = c("Source: 1970-1990 Census",
-                      "The sample includes Whites, who are married, and are between the ages 25 and 40. Ethnicity of a person's parents are identified by the parent's place of birth. A parent is Hispanic if she/he was born in a Spanish-speaking country. A parent is White she/he was born in the United States."),
+                      "The sample includes Whites, who are married, and are between the ages 25 and 40. Ethnicity of a person's parents are identified by the parent's place of birth. A parent is Hispanic if she/he was born in a Spanish-speaking country. A parent is White if she/he was born in the United States."),
            footnote_as_chunk = F, title_format = c("italic"),
            escape = F, threeparttable = T
   ) |> 

### 2023-04-20 programs/11a-table-three-summary-stats.R

@@ -304,7 +304,7 @@ knitr::kable(Table_rows, "latex", align = "lcccccc",
               #   "repeat_header"
                 )) |> 
   footnote(number = c("Source: The 1970-1990 Census for synthetic parents, and 1994-2019 Current Population Surveys (CPS) for children's outcomes",
-                      "The data is restricted to native-born United States citizens between 1994 and 2019 who are also White and between the ages of 25 and 40. I identify the ethnicity of a person's parents through the parent's place of birth. A parent is Hispanic if they were born in a Spanish-speaking country. A parent is White if they were born in the United States."),
+                      "The data is restricted to native-born United States citizens who are also White and between the ages of 25 and 40. I identify the ethnicity of a person's parents through the parent's place of birth. A parent is Hispanic if they were born in a Spanish-speaking country. A parent is White if they were born in the United States."),
            footnote_as_chunk = F, title_format = c("italic"),
            escape = F, threeparttable = T
   ) |> 

### 2023-04-20 programs/14-table-six-regression.R

@@ -39,7 +39,7 @@ reg1 <- list(
 # as a row
 
 
-test1 <- tidy(glht(reg1[[1]], linfct = c("HW:Hispanic_ID + Hispanic_ID = 0")))
+test1 <- tidy(glht(reg1[[1]], linfct = c("HW:Hispanic_ID + HW = 0")))
 differences1 <- case_when(abs(test1$estimate/test1$std.error)   >= 2.58  ~ paste0(sprintf("%2.5s",round_tidy(test1$estimate, digits = 2)), "***"),
                           abs(test1$estimate/test1$std.error)   <  2.58 & 
                             abs(test1$estimate/test1$std.error) >= 1.96  ~ paste0(sprintf("%2.5s",round_tidy(test1$estimate, digits = 2)), "**"),
@@ -48,7 +48,7 @@ differences1 <- case_when(abs(test1$estimate/test1$std.error)   >= 2.58  ~ paste
                           abs(test1$estimate/test1$std.error)   < 1.645  ~ round_tidy(test1$estimate, digits = 2)
 )
 
-test2 <- tidy(glht(reg1[[2]], linfct = c("HW:Hispanic_ID + Hispanic_ID = 0")))
+test2 <- tidy(glht(reg1[[2]], linfct = c("HW:Hispanic_ID + HW = 0")))
 differences2 <- case_when(abs(test2$estimate/test2$std.error)   >= 2.58  ~ paste0(sprintf("%2.5s",round_tidy(test2$estimate, digits = 2)), "***"),
                           abs(test2$estimate/test2$std.error)   <  2.58 & 
                             abs(test2$estimate/test2$std.error) >= 1.96  ~ paste0(sprintf("%2.5s",round_tidy(test2$estimate, digits = 2)), "**"),
@@ -57,7 +57,7 @@ differences2 <- case_when(abs(test2$estimate/test2$std.error)   >= 2.58  ~ paste
                           abs(test2$estimate/test2$std.error)   < 1.645  ~ round_tidy(test2$estimate, digits = 2)
 )
 
-test3 <- tidy(glht(reg1[[3]], linfct = c("HW:Hispanic_ID + Hispanic_ID = 0")))
+test3 <- tidy(glht(reg1[[3]], linfct = c("HW:Hispanic_ID + HW = 0")))
 differences3 <- case_when(abs(test3$estimate/test3$std.error)   >= 2.58  ~ paste0(sprintf("%2.5s",round_tidy(test3$estimate, digits = 2)), "***"),
[…]

### 2023-04-29 my_paper/Hadah-last-names.tex
edits
edit paper to match thesis

@@
 % *****************************************************************
+
+\usepackage{lscape}
+\usepackage{threeparttablex}
+\usepackage{longtable}
 
@@
 \begin{abstract}
-In this paper, I compare the children of inter-ethnic marriages to study the impact of having a Hispanic last name.
-While men born to Hispanic father-White mothers earn less than those born to White father-Hispanic mothers, the gap could be explained entirely by educational differences.
+In this paper, I study discrimination against Hispanics in the labor market.
+I compare the children of inter-ethnic marriages to study the impact of having a Hispanic last name.
+While males born to Hispanic father-White mothers earn less than those born to White father-Hispanic mothers, the gap could be completely explained by educational differences.
 I also study the effect of identifying as Hispanic on earnings.
-I find that men that identify as Hispanic earn significantly less than those that do not, even after controlling for educational differences.
+I find that men who identify as Hispanic earn significantly less than those who do not, even after controlling for educational differences.
 \end{abstract}
@@
 They also find that native-born Hispanics are more likely than their foreign-born counterparts to report poor health.
-\citet{davilaChangesRelativeEarnings2008}, the authors documents many gaps in labor market outcomes between Hispanics and Whites.
+\citet{davilaChangesRelativeEarnings2008} documents many gaps in labor market outcomes between Hispanics and Whites.
 They attribute a big part of this gap to differences in education, experience, immigration status, and regional differences.
@@
 Identifying discrimination is difficult because of factors that affect labor market outcomes that are unobservable to economists---such as unobserved skills.
-one strategy used by researchers is audit or resume studies.
-\citet{bertrand2004emily} conducted an audit study where identical resumes were sent to employers with White and Black Sounding names.
+One strategy used by researchers is audit or resume studies.
+\citet{bertrand2004emily} conducted an audit study where identical resumes were sent to employers with White and Black-sounding names.
 This approach, however, has its drawbacks.
@@
 \citep{averettBetterWorseRelationship2008, averettEconomicRealityBeauty1996, beckerTheoryMarriagePart1973, beckerTheoryMarriagePart1974, beckerTreatiseFamily1993, browningCollectiveUnitaryModels2006, chiapporiFatterAttractionAnthropometric2012}.
-Children of HW and WH marriages have more similar observable characteristics than children of endogamous/homogonous marriages--- i.e., White fathers-White mothers and Hispanic fathers-Hispanic mothers.
+Children of HW and WH marriages have more similar observable characteristics than children of endogamous/homogamous marriages--- i.e., White fathers-White mothers and Hispanic fathers-Hispanic mothers.
 Moreover, children from a Hispanic father and White mother household will have a Hispanic last name from their fathers, enabling the investigation of how ethnic signals, such as having a Hispanic last name, affect annual log earnings.
 
-Using the Current Population Survey (CPS) from 1994 to 2019 and the 1970-1990 US censuses, I estimate the effect of having a Hispanic last name on labor market outcomes.
+Using the Current Population Survey (CPS) from 1994 to 2019 and the 1970 to 1990 US censuses, I estimate the effect of having a Hispanic last name on labor market outcomes.
 In other words, I approximate the variation in labor market perception of ethnic signals, in this case having a Hispanic last name, on annual log earnings.
@@
 On the one hand, WW children are more educated, with approximately 14 years of education for men and women.
-WW children have an employment rate of 95\% for men and 96\% for women.
+WW children have an unemployment rate of 5\% for men and 4\% for women.
 They have log hourly earnings of 2.48 for men and 2.33 for women.
@@
 On the other hand, children of HH marriages have an average of 12.1 and 13.27 years of education for men and women, respectively.
-HH men and women have an employment rate of 93\% and 94\%.
+HH men and women have an unemployment rate of 7\% and 6\%.
 Their annual log earnings are 10.01 and 9.53 for men and women, respectively.
@@
 WH children have an average of 13.58 and 13.87 years of education for men and women.
-The employment rate for men is 94\% and for women 95.0\%.
+The unemployment rate for men is 6\% and for women 5\%.
 HW men and women obtained 13.21 and 13.42 years of schooling.
-Men's employment rate is 92\% and women's 93\%.
+Men's unemployment rate is 8\% and women's 7\%.
 
@@
 These 6\% of WW children are most likely third generation--- or higher--- children of immigrants.
-For my analysis, I dropped the WW observations identify as Hispanic to compare with non-Hispanic Whites.
+For my analysis, I dropped the WW observations who identify as Hispanic when I compare with non-Hispanic Whites.
 
@@
 HW men earn 4\% less annual earnings as WH men, and HW women earn 7\% less.
-These differences show that HW and WH children are different from children of endogamous/homogonous marriages and, thus, are more comparable.
+These differences show that HW and WH children are different from children of endogamous/homogamous marriages and, thus, are more comparable.
 
@@
 
- %In columns five and six of table \ref{tab:c1}, I present the gaps between HH and WW children and HW and WH children, respectively.
-The gap between HH and WW children (column five) shows a consistent story, HH children are worse in education and work more.
-HH children have more kids.
-While the gap between HW and WH does not have a clear-cut direction.
-Even though WH children obtain more years of education and have a higher employment rate, HW and WH men have the same fertility rates.
-HW women have a higher fertility rate than WH women.
-WH and HW women have the same employment rate.
-
 \subsection{Synthetic parents}
 
-Using the 1970 and 1990 censuses, I constructed a data set of synthetic parents.
+Using the 1970 to 1990 censuses, I constructed a data set of synthetic parents.
 The sample includes married White men and women.
@@
 
-I show the distribution of the four types of couples in table \ref{tab:mat1}.
+I show the distribution of the four types of couples in table \ref{tab:mat2}.
 White husbands and White wives (WW) make up the majority of couples in the sample, 97\% (1,286,731 couples).
@@
 
-I present the summary statistics of the four types of synthetic parents in table \ref{tab:synth}'s parent's panel.
+I present the summary statistics of the four types of synthetic parents in table \ref{tab:synth}.
 WW couples have higher education, 13.05 years for husbands and 12.74 for wives.
@@
 The first empirical strategy estimates the effect of having a Hispanic last name on annual log earnings.
-The second empirical strategy will estimate the effect of identifying as Hispanic on log annual earnings.
+The second empirical strategy estimates the effect of identifying as Hispanic on log annual earnings.
 
@@
 The equation for this strategy is written as follows:
-
 \begin{equation} \label{eq:1a}
@@
 
-The coefficients of interest from equation \ref{eq:iden} are $\beta_{1}$, $\beta_{2}$ and $\beta_{3}$. The coefficients $\beta_{2}$ and $\beta_{3}$ capture the effects of Spanish-sounding last names.
+The coefficients of interest from equation \ref{eq:iden} are $\beta_{1}$, $\beta_{2}$ and $\beta_{3}$. The coefficients $\beta_{1}$ and $\beta_{3}$ capture the effects of Spanish-sounding last names.
 
@@
 The results are shown in tables \ref{tab:lastnamereg} and \ref{tab:identreg}.
-All estimations are done on a sample of White, native-born men between the ages of 25 and 40 employed for full-time full-year (FTFY) and are waged and salaried workers.
+All estimations are done on a sample of White, native-born men between the ages of 25 and 40 employed for full-time full-year (FTFY) and are wage and salary workers.
 
@@
 I provide the results to the estimation of equation \ref{eq:1a} in table \ref{tab:lastnamereg}.
-The sample in this analysis includes full-time, full-year, and waged and salaried men, and I control for hours worked and age and include year fixed effects (FE).
+The sample in this analysis includes full-time, full-year, and wage and salary men, and I control for hours worked and age and include year fixed effects (FE).
 I also restrict the sample to children of HW and WH (omitted) parents.
@@
 I provide the results to the estimation of equation \ref{eq:iden} in table \ref{tab:identreg}.
-The sample in this analysis includes full-time, full-year, and waged and salaried men, and I control for hours worked and age and include year fixed effects (FE).
-I also restrict the sample to children of HW and WH (omitted) parents.
+The sample in this analysis includes full-time, full-year, and wage and salary men, and I control for hours worked and age and include year fixed effects (FE).
+I also restrict the sample to children of HW and WH parents.
+The omitted group is children of WH parents that do not identify as Hispanic.
 Column one in table \ref{tab:identreg} is the results without controlling for education.
@@
 This gap, however, could be entirely explained by educational differences.
-An inter-ethnic that identifies as Hispanic earns 6 percentage points less than an inter-ethnic that does not, but the result is statistically insignificant.
+An inter-ethnic that identifies as Hispanic---i.e.
+$HW_{ist} + HW_{ist} \cdot Hispanic_i$---earns 6 percentage points less than an inter-ethnic that does not, but the result is statistically insignificant.
 
@@
 I compare the children of inter-ethnic marriages to study the labor market impact of having a Hispanic last name.
-The earning gap between WH and WW is 14\%, 20\% between HW and WW, and 24\% between HH and WW.
-When I compare the earnings of HW and WH children, which captures the effect of having a Hispanic last name, HW children earn  6\% less than WH children.
-Thus, by comparing inter-ethnic children, a person with a Hispanic last name makes 6\% less than someone with a White last name.
-Controlling for education, however, explains more than half of the earnings gaps between WW children and WH, HW, and HH.
-When I control for education, the last name effect decreases to a statistically insignificant 2\% earnings gap.
+When I compare the earnings of HW and WH children, which captures the effect of having a Hispanic last name, HW children earn  6 percentage points less than WH children.
+Thus, by comparing inter-ethnic children, a person with a Hispanic last name makes 6 percentage points less than someone with a White last name.
+When I control for education, the last name effect decreases to a statistically insignificant 2 percentage points earnings gap.
+Moreover, I estimate the effect of identifying as Hispanic and having a Hispanic last name.
+I find that an inter-ethnic child with a Hispanic last name and identifies as Hispanic earns 5 percentage points less that an inter-ethnic child with native sounding last name and identifies as Hispanic.
+The earnings gap, however, could be entirely explained by educational differences.
 

### 2023-05-09 my_paper/Hadah-last-names.tex
finish implementing the code that would merge synthetic parents (ACS) to children (CPS) addressing Yona's comments

@@
 \begin{center}
-\begin{figure}
+\begin{figure}[h!]
 \caption{Inter-racial marriages as a percent of all marriages \\
@@
 \begin{center}
-\begin{figure}
+\begin{figure}[h!]
 \caption{Distribution of the four types of children.}
@@
 \newpage
-\begin{center}
-\begin{figure}
-\caption{Chart explaining which synthetic parents and children \\
-and when we observe them.}
+
+\begin{figure}[h!]
+\centering
+\caption{Chart explaining which synthetic parents and children and when we observe them.}
 \begin{tikzpicture}[node distance =2cm]
-\node (start) [startstop] {Synthetic Parents: Observed in 1970 to 1990. $25 \leq age \leq 40$};
+\node (start) [startstop] {Synthetic Parents: Observed in 1960 to 2000. $25 \leq age \leq 40$, White, with kids};
 
-\node (pro1) [process, below of = start, yshift = -2cm] {Children: Observed between 1994 and 2019. Born between 1970 to 1990 and $25 \leq age \leq 40$};
+\node (pro1) [process, below of = start, yshift = -2cm] {Children: Observed between 1994 and 2019. Born between 1954 to 1999 and $25 \leq age \leq 40$};
 
@@
 \end{figure}
-\end{center}
 

### 2023-05-09 my_paper/tables/tab04-summary-stats.tex
finish implementing the code that would merge synthetic parents (ACS) to children (CPS) addressing Yona's comments

@@
-\begin{table}[H]
-
-\caption{Summary statistics of outcomes using parent's place of birth only for those that self-identify as Hispanic \label{tab:c&p2}}
-\centering
-\resizebox{\linewidth}{!}{
-\begin{threeparttable}
-\begin{tabular}[t]{lcccccc}
+\begin{landscape}
+\begin{ThreePartTable}
+\begin{TableNotes}
+\item[1] Source: The 1950-2000 Census for synthetic parents, and 1994-2019 Current Population Surveys (CPS) for children's outcomes
+\item[2] The data is restricted to native-born United States citizens between 1994 and 2019 who are also White and between the ages of 25 and 40.
+I identify the ethnicity of a person's parents through the parent's place of birth.
+A parent is Hispanic if they were born in a Spanish-speaking country.
+A parent is White if they were born in the United States.
+\end{TableNotes}
+\begin{longtable}[t]{>{\raggedright\arraybackslash}p{5cm}cccccc}
+\caption{Summary statistics of outcomes using parent's place of birth only for those that self-identify as Hispanic \label{tab:c&p2}}\\
 \toprule
@@
 \cmidrule(l{3pt}r{3pt}){2-5} \cmidrule(l{3pt}r{3pt}){6-7}
-Variables & \specialcell{White Father \\ White Mother \\ (WW) \\ (1)} & \specialcell{White Father \\ Hispanic Mother \\ (WH) \\ (2)} & \specialcell{Hispanic Father \\ White Mother \\ (HW) \\ (3)} & \specialcell{Hispanic Father \\ Hispanic Mother \\ (HH) \\ (4)} & \specialcell{HH - WW \\ (5)} & \specialcell{HW - WH \\ (6)}\\
+Variables & \specialcell{White \\ White \\ (WW) \\ (1)} & \specialcell{White \\ Hispanic \\ (WH) \\ (2)} & \specialcell{Hispanic \\ White \\ (HW) \\ (3)} & \specialcell{Hispanic \\ Hispanic \\ (HH) \\ (4)} & \specialcell{HH - WW \\ (5)} & \specialcell{HW - WH \\ (6)}\\
 \midrule
-\textbf{Panel A: Children's Sunthetic Parent} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
-\hspace{1em}Husband’seducation (Total Years) & \specialcell{13.05\\(2.44)} & \specialcell{12.32\\(3.33)} & \specialcell{10.65\\(4.39)} & \specialcell{8.93\\(4.41)} & \specialcell{-4.11**\\(0.02)} & \specialcell{-1.67**\\(0.04)}\\
-\hspace{1em}Wife’seducation (Total Years) & \specialcell{12.74\\(2.12)} & \specialcell{11.03\\(3.92)} & \specialcell{11.54\\(3.12)} & \specialcell{8.6\\(4.13)} & \specialcell{-4.13**\\(0.02)} & \specialcell{0.51**\\(0.04)}\\
-\hspace{1em}Total Household seducation (Total Years) & \specialcell{25.78\\(4.08)} & \specialcell{23.35\\(6.51)} & \specialcell{22.19\\(6.69)} & \specialcell{17.54\\(7.83)} & \specialcell{-8.25**\\(0.03)} & \specialcell{-1.16*\\(0.07)}\\
-\textbf{Panel B: Education} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
+\endfirsthead
+\caption[]{Summary statistics of outcomes using parent's place of birth only for those that self-identify as Hispanic  \textit{(continued)}}\\
+\toprule
+Variables & \specialcell{White \\ White \\ (WW) \\ (1)} & \specialcell{White \\ Hispanic \\ (WH) \\ (2)} & \specialcell{Hispanic \\ White \\ (HW) \\ (3)} & \specialcell{Hispanic \\ Hispanic \\ (HH) \\ (4)} & \specialcell{HH - WW \\ (5)} & \specialcell{HW - WH \\ (6)}\\
+\midrule
+\endhead
+\midrule
+\multicolumn{7}{r@{}}{\textit{(Continued on Next Page...)}}\
+\endfoot
+\bottomrule
+\insertTableNotes
+\endlastfoot
+\textbf{Panel A: Children's Education} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
+\hspace{1em}Men’s education (Total Years) & \specialcell{13.82\\(2.42)} & \specialcell{13.4\\(2.38)} & \specialcell{13.08\\(2.27)} & \specialcell{12.87\\(2.3)} & \specialcell{-0.94***\\(0.01)} & \specialcell{-0.32**\\(0.03)}\\
+\hspace{1em}Women’s education (Total Years) & \specialcell{14.06\\(2.37)} & \specialcell{13.62\\(2.39)} & \specialcell{13.26\\(2.35)} & \specialcell{13.22\\(2.39)} & \specialcell{-0.84***\\(0.01)} & \specialcell{-0.36**\\(0.02)}\\
+\textbf{Panel B: Children's Employment and Earnings} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
+\hspace{1em}Men’s Unemployment Rate & \specialcell{0.04\\(0.8)} & \specialcell{0.06\\(0.76)} & \specialcell{0.07\\(0.74)} & \specialcell{0.07\\(0.75)} & \specialcell{0.02***\\(0.00)} & \specialcell{0.01***\\(0.00)}\\
 \addlinespace
-\hspace{1em}Men’s education (Total Years) & \specialcell{12.97\\(2.15)} & \specialcell{13.45\\(2.37)} & \specialcell{13.13\\(2.27)} & \specialcell{12.89\\(2.25)} & \specialcell{-0.08**\\(0.01)} & \specialcell{-0.32**\\(0.03)}\\
-\hspace{1em}Women’s education (Total Years) & \specialcell{13.23\\(2.25)} & \specialcell{13.75\\(2.41)} & \specialcell{13.32\\(2.34)} & \specialcell{13.26\\(2.37)} & \specialcell{0.03**\\(0.01)} & \specialcell{-0.43**\\(0.03)}\\
-\textbf{Panel C: Children's Employment and Earnings} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
-\hspace{1em}Men’s Employment Rate & \specialcell{0.93\\(0.26)} & \specialcell{0.94\\(0.23)} & \specialcell{0.92\\(0.27)} & \specialcell{0.93\\(0.26)} & \specialcell{0.00***\\(0.00)} & \specialcell{-0.02***\\(0.00)}\\
-\hspace{1em}Women’s Employment Rate & \specialcell{0.94\\(0.25)} & \specialcell{0.94\\(0.23)} & \specialcell{0.93\\(0.26)} & \specialcell{0.94\\(0.24)} & \specialcell{0.00***\\(0.00)} & \specialcell{-0.02***\\(0.00)}\\
-\addlinespace
-\hspace{1em}Men’s Log Hourly Earnings & \specialcell{2.4\\(0.44)} & \specialcell{2.41\\(0.45)} & \specialcell{2.4\\(0.44)} & \specialcell{2.41\\(0.43)} & \specialcell{0.01***\\(0.01)} & \specialcell{-0.01**\\(0.02)}\\
-\hspace{1em}Women’s Log Hourly Earnings & \specialcell{2.26\\(0.43)} & \specialcell{2.32\\(0.45)} & \specialcell{2.27\\(0.45)} & \specialcell{2.3\\(0.41)} & \specialcell{0.04***\\(0.01)} & \specialcell{-0.05**\\(0.02)}\\
-\hspace{1em}Men’s Log Annual Earnings & \specialcell{10.02\\(1.02)} & \specialcell{10.06\\(1.06)} & \specialcell{10.03\\(0.99)} & \specialcell{10\\(1.04)} & \specialcell{-0.02**\\(0.01)} & \specialcell{-0.03**\\(0.04)}\\
-\hspace{1em}Women’s Log Annual Earnings & \specialcell{9.44\\(1.59)} & \specialcell{9.55\\(1.59)} & \specialcell{9.47\\(1.55)} & \specialcell{9.52\\(1.52)} & \specialcell{0.08**\\(0.02)} & \specialcell{-0.08*\\(0.05)}\\
-\bottomrule
-\end{tabular}
-\begin{tablenotes}
-\item[1] Source: The 1970-1990 Census for synthetic parents, and 1994-2019 Current Population Surveys (CPS) for children's outcomes
-\item[2] The data is restricted to native-born United States citizens between 1994 and 2019 who are also White and between the ages of 25 and 40, and who identify as Hispanic.
-I identify the ethnicity of a person's parents through the parent's place of birth.
-A parent is Hispanic if they were born in a Spanish-speaking country.
-A parent is White if they were born in the United States.
-\end{tablenotes}
-\end{threeparttable}}
-\end{table}
+\hspace{1em}Women’s Unemployment Rate & \specialcell{0.04\\(0.81)} & \specialcell{0.06\\(0.24)} & \specialcell{0.07\\(0.75)} & \specialcell{0.06\\(0.76)} & \specialcell{0.02***\\(0.00)} & \specialcell{0.01***\\(0.00)}\\
+\hspace{1em}Men’s Log Hourly Earnings & \specialcell{2.51\\(0.45)} & \specialcell{2.43\\(0.46)} & \specialcell{2.42\\(0.44)} & \specialcell{2.42\\(0.43)} & \specialcell{-0.09***\\(0)} & \specialcell{-0.00**\\(0.02)}\\
+\hspace{1em}Women’s Log Hourly Earnings & \specialcell{2.32\\(0.49)} & \specialcell{2.31\\(0.45)} & \specialcell{2.28\\(0.45)} & \specialcell{2.31\\(0.42)} & \specialcell{-0.02***\\(0)} & \specialcell{-0.03**\\(0.02)}\\
+\hspace{1em}Men’s Log Annual Earnings & \specialcell{10.29\\(1.01)} & \specialcell{10.09\\(1.05)} & \specialcell{10.06\\(0.97)} & \specialcell{10.01\\(1.04)} & \specialcell{-0.28***\\(0.01)} & \specialcell{-0.03**\\(0.03)}\\
+\hspace{1em}Women’s Log Annual Earnings & \specialcell{9.43\\(1.77)} & \specialcell{9.52\\(1.59)} & \specialcell{9.47\\(1.56)} & \specialcell{9.53\\(1.49)} & \specialcell{0.1**\\(0.02)} & \specialcell{-0.05*\\(0.05)}\\*
+\end{longtable}
+\end{ThreePartTable}
+\end{landscape}

### 2023-05-09 output/tables/tab03a-summary-stats.tex
table edits
finish implementing the code that would merge synthetic parents (ACS) to children (CPS) addressing Yona's comments

@@
 \begin{threeparttable}
-\begin{tabular}[t]{lcccccc}
+\begin{tabular}[t]{>{\raggedright\arraybackslash}p{5cm}cccccc}
 \toprule
@@
 \midrule
-Husband's education (Total Years) & \specialcell{13.05\\(2.44)} & \specialcell{12.32\\(3.33)} & \specialcell{10.65\\(4.39)} & \specialcell{8.93\\(4.41)} & \specialcell{-4.11**\\(0.02)} & \specialcell{-1.67**\\(0.04)}\\
-Wife's education (Total Years) & \specialcell{12.74\\(2.12)} & \specialcell{11.03\\(3.92)} & \specialcell{11.54\\(3.12)} & \specialcell{8.6\\(4.13)} & \specialcell{-4.13**\\(0.02)} & \specialcell{0.51**\\(0.04)}\\
-Total Household seducation (Total Years) & \specialcell{25.78\\(4.08)} & \specialcell{23.35\\(6.51)} & \specialcell{22.19\\(6.69)} & \specialcell{17.54\\(7.83)} & \specialcell{-8.25**\\(0.03)} & \specialcell{-1.16*\\(0.07)}\\
-Log Total Family Income & \specialcell{10.79\\(0.57)} & \specialcell{10.65\\(0.66)} & \specialcell{10.55\\(0.74)} & \specialcell{10.38\\(0.75)} & \specialcell{-0.41***\\(0)} & \specialcell{-0.1***\\(0.01)}\\
+Husband's education (Total Years) & \specialcell{12.75\\(0.61)} & \specialcell{11.77\\(1.87)} & \specialcell{10.25\\(2.44)} & \specialcell{8.64\\(2.01)} & \specialcell{-4.11***\\(0.00)} & \specialcell{-1.52***\\(0.00)}\\
+Wife's education (Total Years) & \specialcell{12.47\\(0.56)} & \specialcell{10.40\\(2.23)} & \specialcell{11.11\\(1.75)} & \specialcell{8.49\\(1.92)} & \specialcell{-3.98***\\(0.00)} & \specialcell{0.70***\\(0.00)}\\
+Total Household seducation (Total Years) & \specialcell{25.22\\(1.17)} & \specialcell{22.17\\(4.05)} & \specialcell{21.36\\(4.10)} & \specialcell{17.13\\(3.88)} & \specialcell{-8.09***\\(0.00)} & \specialcell{-0.81***\\(0.01)}\\
+Log Total Family Income & \specialcell{10.72\\(0.09)} & \specialcell{10.55\\(0.26)} & \specialcell{10.46\\(0.27)} & \specialcell{10.27\\(0.21)} & \specialcell{-0.45***\\(0.00)} & \specialcell{-0.10***\\(0.00)}\\
+Fertility & \specialcell{3.77\\(0.40)} & \specialcell{3.98\\(0.71)} & \specialcell{4.15\\(0.79)} & \specialcell{4.23\\(0.64)} & \specialcell{0.46***\\(0.00)} & \specialcell{0.17***\\(0.00)}\\
 \bottomrule
@@
 \begin{tablenotes}
-\item[1] Source: The 1970-1990 Census for synthetic parents, and 1994-2019 Current Population Surveys (CPS) for children's outcomes
-\item[2] The data is restricted to native-born United States citizens between 1994 and 2019 who are also White and between the ages of 25 and 40.
+\item[1] Source: The 1950-2000 Census for synthetic parents, and 1994-2019 Current Population Surveys (CPS) for children's outcomes
+\item[2] The data is restricted to native-born United States citizens who are also White, between the ages of 25 and 40, and have kids.
 I identify the ethnicity of a person's parents through the parent's place of birth.

### 2023-05-09 output/tables/tab03b-summary-stats.tex
finish implementing the code that would merge synthetic parents (ACS) to children (CPS) addressing Yona's comments

--- 2023-04-17 my_paper/tables/tab03b-summary-stats.tex
@@
 \begin{TableNotes}
-\item[1] Source: The 1970-1990 Census for synthetic parents, and 1994-2019 Current Population Surveys (CPS) for children's outcomes
-\item[2] The data is restricted to native-born United States citizens between 1994 and 2019 who are also White and between the ages of 25 and 40.
+\item[1] Source: The 1950-2000 Census for synthetic parents, and 1994-2019 Current Population Surveys (CPS) for children's outcomes
+\item[2] The data is restricted to native-born United States citizens who are also White and between the ages of 25 and 40.
 I identify the ethnicity of a person's parents through the parent's place of birth.
@@
 \textbf{Panel A: Children's Education} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
-\hspace{1em}Men’s education (Total Years) & \specialcell{13.91\\(2.39)} & \specialcell{13.58\\(2.35)} & \specialcell{13.21\\(2.32)} & \specialcell{12.91\\(2.26)} & \specialcell{-1.00***\\(0.01)} & \specialcell{-0.36**\\(0.03)}\\
-\hspace{1em}Women’s education (Total Years) & \specialcell{14.29\\(2.41)} & \specialcell{13.87\\(2.47)} & \specialcell{13.42\\(2.35)} & \specialcell{13.27\\(2.37)} & \specialcell{-1.01***\\(0.01)} & \specialcell{-0.46**\\(0.03)}\\
+\hspace{1em}Men’s education (Total Years) & \specialcell{13.82\\(2.42)} & \specialcell{13.58\\(2.40)} & \specialcell{13.22\\(2.34)} & \specialcell{12.90\\(2.31)} & \specialcell{-0.92***\\(0.01)} & \specialcell{-0.36**\\(0.02)}\\
+\hspace{1em}Women’s education (Total Years) & \specialcell{14.06\\(2.37)} & \specialcell{13.79\\(2.44)} & \specialcell{13.42\\(2.38)} & \specialcell{13.24\\(2.39)} & \specialcell{-0.82***\\(0.01)} & \specialcell{-0.37**\\(0.02)}\\
 \textbf{Panel B: Children's Employment and Earnings} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
-\hspace{1em}Men’s Unemployment Rate & \specialcell{0.05\\(0.78)} & \specialcell{0.06\\(0.77)} & \specialcell{0.08\\(0.74)} & \specialcell{0.07\\(0.74)} & \specialcell{0.02***\\(0.00)} & \specialcell{0.02***\\(0.00)}\\
+\hspace{1em}Men’s Unemployment Rate & \specialcell{0.04\\(0.80)} & \specialcell{0.05\\(0.77)} & \specialcell{0.07\\(0.75)} & \specialcell{0.07\\(0.75)} & \specialcell{0.02***\\(0.00)} & \specialcell{0.01***\\(0.00)}\\
 \addlinespace
-\hspace{1em}Women’s Unemployment Rate & \specialcell{0.04\\(0.8)} & \specialcell{0.05\\(0.22)} & \specialcell{0.07\\(0.75)} & \specialcell{0.06\\(0.76)} & \specialcell{0.02***\\(0.00)} & \specialcell{0.02***\\(0.00)}\\
-\hspace{1em}Men’s Log Hourly Earnings & \specialcell{2.48\\(0.45)} & \specialcell{2.41\\(0.45)} & \specialcell{2.4\\(0.44)} & \specialcell{2.41\\(0.43)} & \specialcell{-0.07***\\(0.01)} & \specialcell{-0.00**\\(0.02)}\\
-\hspace{1em}Women’s Log Hourly Earnings & \specialcell{2.33\\(0.49)} & \specialcell{2.33\\(0.46)} & \specialcell{2.27\\(0.45)} & \specialcell{2.3\\(0.41)} & \specialcell{-0.02***\\(0.01)} & \specialcell{-0.06**\\(0.02)}\\
-\hspace{1em}Men’s Log Annual Earnings & \specialcell{10.25\\(1.01)} & \specialcell{10.08\\(1.05)} & \specialcell{10.04\\(1.06)} & \specialcell{10.01\\(1.04)} & \specialcell{-0.25**\\(0.01)} & \specialcell{-0.04**\\(0.03)}\\
-\hspace{1em}Women’s Log Annual Earnings & \specialcell{9.46\\(1.78)} & \specialcell{9.54\\(1.64)} & \specialcell{9.47\\(1.57)} & \specialcell{9.53\\(1.52)} & \specialcell{0.07**\\(0.02)} & \specialcell{-0.07*\\(0.06)}\\
+\hspace{1em}Women’s Unemployment Rate & \specialcell{0.04\\(0.81)} & \specialcell{0.05\\(0.22)} & \specialcell{0.06\\(0.76)} & \specialcell{0.06\\(0.76)} & \specialcell{0.02***\\(0.00)} & \specialcell{0.01***\\(0.00)}\\
+\hspace{1em}Men’s Log Hourly Earnings & \specialcell{2.51\\(0.45)} & \specialcell{2.44\\(0.47)} & \specialcell{2.43\\(0.45)} & \specialcell{2.42\\(0.43)} & \specialcell{-0.09***\\(0.00)} & \specialcell{-0.01**\\(0.01)}\\
+\hspace{1em}Women’s Log Hourly Earnings & \specialcell{2.32\\(0.49)} & \specialcell{2.32\\(0.46)} & \specialcell{2.29\\(0.46)} & \specialcell{2.31\\(0.42)} & \specialcell{-0.02***\\(0.00)} & \specialcell{-0.03**\\(0.01)}\\
+\hspace{1em}Men’s Log Annual Earnings & \specialcell{10.29\\(1.01)} & \specialcell{10.12\\(1.05)} & \specialcell{10.08\\(1.01)} & \specialcell{10.01\\(1.04)} & \specialcell{-0.28***\\(0.01)} & \specialcell{-0.04**\\(0.03)}\\
+\hspace{1em}Women’s Log Annual Earnings & \specialcell{9.43\\(1.77)} & \specialcell{9.51\\(1.63)} & \specialcell{9.46\\(1.59)} & \specialcell{9.53\\(1.50)} & \specialcell{0.10**\\(0.02)} & \specialcell{-0.05**\\(0.05)}\\
 \addlinespace
 \textbf{Panel C: Children's Hispanic Identity} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
-\hspace{1em}Men & \specialcell{0.05} & \specialcell{0.81} & \specialcell{0.88} & \specialcell{0.97} &  & \\
-\hspace{1em}Women & \specialcell{0.06} & \specialcell{0.85} & \specialcell{0.87} & \specialcell{0.97} &  & \\*
+\hspace{1em}Men & \specialcell{0.04} & \specialcell{0.74} & \specialcell{0.83} & \specialcell{0.96} &  & \\
+\hspace{1em}Women & \specialcell{0.05} & \specialcell{0.78} & \specialcell{0.82} & \specialcell{0.97} &  & \\*
 \end{longtable}

### 2023-05-09 output/tables/tab06-regression.tex
finish implementing the code that would merge synthetic parents (ACS) to children (CPS) addressing Yona's comments

--- 2023-04-16 output/tables/tab06-regression.tex
@@
 
-\caption{Effect of Having Hispanic Last Name \label{tab:identreg}}
+\caption{Effect of Having Hispanic Last Name by Hispanic Identity\label{tab:identreg}}
 \centering
@@
 \begin{threeparttable}
-\begin{tabular}[t]{lcc}
+\begin{tabular}[t]{lccccc}
 \toprule
-  & \specialcell{(1) \\ Log annual earnings} & \specialcell{(2) \\  Log annual earnings}\\
+  & \specialcell{(1) \\ Log annual \\ earnings} & \specialcell{(2) \\ Log annual \\ earnings} & \specialcell{(3) \\ Log annual \\ earnings} & \specialcell{(4) \\  Log annual \\ earnings} & \specialcell{(5) \\  Log annual \\ earnings}\\
 \midrule
-$HW_{i} \times Hispanic_{i}$ & \num{-0.02} & \num{-0.02}\\
- & (\num{0.08}) & (\num{0.07})\\
-$Hispanic_{i}$ & \num{-0.11}** & \num{-0.06}\\
- & (\num{0.05}) & (\num{0.05})\\
-$HW_{i}$ & \num{-0.03} & \num{0.00}\\
- & (\num{0.08}) & (\num{0.06})\\
+$HW_{ist} \times Hispanic_{ist}$ & \num{-0.09} & \num{-0.06} & \num{-0.05} & \num{-0.04} & \num{-0.04}\\
+ & (\num{0.07}) & (\num{0.07}) & (\num{0.06}) & (\num{0.06}) & (\num{0.06})\\
+$Hispanic_{ist}$ & \num{-0.13}** & \num{-0.11} & \num{-0.09}** & \num{-0.03} & \num{-0.02}\\
+ & (\num{0.07}) & (\num{0.07}) & (\num{0.04}) & (\num{0.05}) & (\num{0.04})\\
+$HW_{ist}$ & \num{0.03} & \num{0.01} & \num{0.00} & \num{0.02} & \num{0.02}\\
+ & (\num{0.06}) & (\num{0.06}) & (\num{0.06}) & (\num{0.06}) & (\num{0.05})\\
+Constant & \num{10.53}*** & \num{9.57}*** &  &  & \\
+ & (\num{0.05}) & (\num{0.08}) &  &  & \\
 \midrule
-State FE & X & X\\
-Year FE & X & X\\
-\textit{Controlling for:} &  & \\
-Hours Worked & X & X\\
-Age & X & X\\
-Education &  & X\\
-Observations & \num{2423} & \num{2423}\\
+$HW_{ist} \times Hispanic_{ist} + HW_{ist}$ & -0.06*** & -0.05** & -0.05** & -0.02 & -0.01\\
+ & (0.02) & (0.02) & (0.02) & (0.02) & (0.02)\\
+\midrule
+\textit{Controlling for:} &  &  &  &  & \\
+Hours Worked &  & X & X & X & X\\
+State FE &  &  & X & X & X\\
+Year FE &  &  & X & X & X\\
+Age FE &  &  & X & X & X\\
+Education FE &  &  &  & X & X\\
+Parental Background &  &  &  &  & X\\
+Observations & \num{3581} & \num{3581} & \num{3581} & \num{3581} & \num{3581}\\
 \bottomrule
-\multicolumn{3}{l}{\rule{0pt}{1em}* p $<$ 0.1, ** p $<$ 0.05, *** p $<$ 0.01}\\
+\multicolumn{6}{l}{\rule{0pt}{1em}* p $<$ 0.1, ** p $<$ 0.05, *** p $<$ 0.01}\\
 \end{tabular}
@@
 \item[3] \footnotesize{The sample is restricted to men working full-time full-year and are waged and salaried workers.}
-\item[4] \footnotesize{Column one has the regression results when controlling for hours worked, age, and fixed effects. Column two has the results after controlling for education.}
+\item[4] \footnotesize{Column one has the regression results when controlling for hours worked, age, education, year and state fixed effects.
+Column two has the results after controlling for education.}
+\item[5] {\setstretch{1.0}\footnotesize{Standard errors are clustered on the state level.}}
 \end{tablenotes}

### 2023-05-09 programs/01-packages-wds.R

@@ -26,7 +26,7 @@ pacman::p_load(tictoc, parallel, pbapply, future,
                gganimate, ggrepel, MetBrewer, fs,
                marginaleffects, gghighlight, #ggview,
                camcorder, rnaturalearth, rnaturalearthdata,
-               latex2exp, janitor)
+               latex2exp, janitor, extrafont)
 options("RStata.StataPath" = "/Applications/Stata/StataSE.app/Contents/MacOS/stata-se")
 options("RStata.StataVersion" = 17)
             
@@ -38,9 +38,11 @@ options("RStata.StataVersion" = 17)
 # scale_color_ipsum()
 font_add_google("Fira Sans", "firasans")
 font_add_google("Fira Code", "firasans")
+font_add_google("Quattrocento", "lmmodern")
 
 showtext_auto()
-
+loadfonts(device = "all")
+# font_import(pattern = "lmroman*") 
 showtext_opts(dpi = 300)
 camcorder::gg_record(
   dir = figures_wd,
@@ -51,7 +53,7 @@ camcorder::gg_record(
   dpi = 300
 )
 theme_customs <- function() {
-  theme_minimal(base_family = "IBM Plex Sans Condensed") +
+  theme_minimal(base_family = "LM Roman 10") +
     theme(panel.grid.minor = element_blank(),
           plot.background = element_rect(fill = "white", color = NA),
           plot.title = element_text(face = "bold"),
@@ -63,7 +65,7 @@ theme_customs <- function() {
 }
 
 theme_customs_map <- function() {
-  theme_minimal(base_family = "IBM Plex Sans Condensed") +
+  theme_minimal(base_family = "LM Roman 10") +
     theme(panel.grid.major = element_blank(), 
           panel.grid.minor = element_blank(),
           panel.background = element_blank(), 
@@ -79,11 +81,11 @@ theme_customs_map <- function() {
 }
 # Make labels use IBM Plex Sans by default
 update_geom_defaults("label", 
-                     list(family = "IBM Plex Sans Condensed"))
+                     list(family = "LM Roman 10"))
 update_geom_defaults(ggtext::GeomRichText, 
-                     list(family = "IBM Plex Sans Condensed"))
+                     list(family = "LM Roman 10"))
 update_geom_defaults("label_repel", 
-                     list(family = "IBM Plex Sans Condensed"))
+                     list(family = "LM Roman 10"))
 
 # Use the Johnson color palette
 clrs <- met.brewer("Johnson")

### 2023-05-09 programs/12-table-two-couple-types.R

@@ -122,8 +122,8 @@ knitr::kable(Table_cols, "latex", align = "lcccc",
   kable_classic(full_width = F) %>%
   kable_styling(bootstrap_options = c("hover", "condensed"), 
                 latex_options = c("scale_down", "HOLD_position")) |> 
-  footnote(number = c("Source: 1970-1990 Census",
-                      "The sample includes Whites, who are married, and are between the ages 25 and 40. Ethnicity of a person's parents are identified by the parent's place of birth. A parent is Hispanic if she/he was born in a Spanish-speaking country. A parent is White if she/he was born in the United States."),
+  footnote(number = c("Source: 1950-2000 Census",
+                      "The sample includes Whites, who are married, with kids, and are between the ages 25 and 40. Ethnicity of a person's parents are identified by the parent's place of birth. A parent is Hispanic if she/he was born in a Spanish-speaking country. A parent is White if she/he was born in the United States."),
            footnote_as_chunk = F, title_format = c("italic"),
            escape = F, threeparttable = T
   ) |> 

### 2023-05-09 programs/14-table-three-summary-stats.R

@@ -4,7 +4,7 @@
 
 # date: January 19th, 2023
 
-ParentData     <- read_dta(file.path(datasets,"ParentDataFull.dta")) |> 
+ParentData     <- read_csv(file.path(datasets,"CPS_synth.csv")) |> 
   mutate(Tot_ed = HusbandEducation + WifeEducation)
 
 row1 <- c("Variables", 
@@ -22,64 +22,63 @@ dim(row1) <- c(1,7)
 ## Husband's Education
 HusbandEducation <- ParentData |> 
   group_by(CoupleType) |> 
-  summarise(HusEducation   = weighted.mean(HusbandEducation, na.rm = T, hhwt),
-            HusEducationSD = weighted.sd(HusbandEducation, na.rm = T, hhwt))
+  summarise(HusEducation   = mean(HusbandEducation, na.rm = T),
+            HusEducationSD = sd(HusbandEducation, na.rm = T))
 
-model <- lm(HusbandEducation ~ 0 + WW + WM + MW + MM, 
-    data = ParentData, 
-    weights = hhwt)
-test1 <- tidy(glht(model, linfct = c("MM - WW = 0")))
+model <- lm(HusbandEducation ~ 0 + WW + WH + HW + HH, 
+    data = ParentData)
+test1 <- tidy(glht(model, linfct = c("HH - WW = 0")))
 pvalues1 <- test1$std.error
 differences1 <- test1$estimate
 
-test2 <- tidy(glht(model, linfct = c("MW - WM = 0")))
+test2 <- tidy(glht(model, linfct = c("HW - WH = 0")))
 pvalues2 <- test2$std.error
 differences2 <- test2$estimate
 
 row3 <- c(
   "Husband's education (Total Years)",
   paste0("\\specialcell{", 
-         round(HusbandEducation[4,2], digits = 2), 
+         sprintf("%0.2f",round(HusbandEducation[4,2], digits = 2)),  
          "\\\\",
          "(",
-         round(HusbandEducation[4,3], digits = 2),
+         sprintf("%0.2f",round(HusbandEducation[4,3], digits = 2)), 
          ")}"
   ),
   paste0("\\specialcell{", 
-         round(HusbandEducation[3,2], digits = 2), 
+         sprintf("%0.2f",round(HusbandEducation[3,2], digits = 2)),  
          "\\\\",
          "(",
-         round(HusbandEducation[3,3], digits = 2),
+         sprintf("%0.2f",round(HusbandEducation[3,3], digits = 2)), 
[…]

### 2023-05-09 programs/16-table-four-summary-stats.R

@@ -1,232 +1,29 @@
 # summary stats
 
-# table four
-# only hispanics
+# table three
+
+# library(Hmisc)
 
 # date: January 19th, 2023
 
-IndividualData <- read_dta(file.path(datasets,"BySexAnalysisData.dta")) |> 
+IndividualData <- read_csv(file.path(datasets,"CPS_synth.csv")) |> 
   mutate(Education = as.numeric(Education)) |> 
-  filter(Hispanic_ID == 1)
-ParentData     <- read_dta(file.path(datasets,"ParentDataFull.dta")) |> 
-  mutate(Tot_ed = HusbandEducation + WifeEducation)
+  filter(WW == 1 | (WH == 1 & Hispanic_ID == 1) | (HW == 1 & Hispanic_ID == 1) | (HH == 1 & Hispanic_ID == 1))
 
+CrossTable(IndividualData$Type, IndividualData$Hispanic_ID)
 row1 <- c("Variables", 
-          "\\specialcell{White Father \\\\ White Mother \\\\ (WW) \\\\ (1)}",
-          "\\specialcell{White Father \\\\ Hispanic Mother \\\\ (WH) \\\\ (2)}",
-          "\\specialcell{Hispanic Father \\\\ White Mother \\\\ (HW) \\\\ (3)}",
-          "\\specialcell{Hispanic Father \\\\ Hispanic Mother \\\\ (HH) \\\\ (4)}",
+          "\\specialcell{White \\\\ White \\\\ (WW) \\\\ (1)}",
+          "\\specialcell{White \\\\ Hispanic \\\\ (WH) \\\\ (2)}",
+          "\\specialcell{Hispanic \\\\ White \\\\ (HW) \\\\ (3)}",
+          "\\specialcell{Hispanic \\\\ Hispanic \\\\ (HH) \\\\ (4)}",
           "\\specialcell{HH - WW \\\\ (5)}",
           "\\specialcell{HW - WH \\\\ (6)}"
           )
-
 dim(row1) <- c(1,7)
-row2 <- c("Panel A: Children's Sunthetic Parent", " ", " ", " ", " ", " ", " ")
-dim(row2) <- c(1,7)
 
-# parent's panel
-
-## Husband's Education
-HusbandEducation <- ParentData |> 
-  group_by(CoupleType) |> 
-  summarise(HusEducation   = weighted.mean(HusbandEducation, na.rm = T, hhwt),
[…]

### 2023-05-09 programs/17-table-five-regression.R

@@ -3,12 +3,31 @@
 
 # date: January 19th, 2023
 
-IndividualData <- read_dta(file.path(datasets,"BySexAnalysisData.dta")) |> 
+# open data, filter to only have the relevant data
+# and impute empty parental charachtersitcs with mean
+IndividualData <- read_csv(file.path(datasets,"CPS_synth.csv")) |> 
   mutate(Education = as.numeric(Education)) |> 
-  filter((WW == 1 & Hispanic_ID == 0) | WH == 1 | HW == 1 | HH == 1)
-
+  filter((WW == 1 & Hispanic_ID == 0) | WH == 1 | HW == 1 | HH == 1) |> 
+  mutate(HusbandEducation = case_when(!is.na(HusbandEducation) ~ HusbandEducation,
+                                       is.na(HusbandEducation) ~ mean(HusbandEducation, na.rm = TRUE)),
+         WifeEducation = case_when(!is.na(WifeEducation) ~ WifeEducation,
+                                    is.na(WifeEducation) ~ mean(WifeEducation, na.rm = TRUE)),
+         Husband_ftotval = case_when(!is.na(Husband_ftotval) ~ Husband_ftotval,
+                                       is.na(Husband_ftotval) ~ mean(Husband_ftotval, na.rm = TRUE)),
+                                )
+# create a total education variable
+# and a income quintile variable
+# and education quintile variable
+IndividualData  <- IndividualData  |> 
+  mutate(TotalHHEducation = HusbandEducation + WifeEducation,
+         IncomeQuint      = cut(Husband_ftotval, breaks = quantile(Husband_ftotval, probs = c(0, 0.2, 0.4, 0.6, 0.8, 1), na.rm = TRUE), include.lowest = TRUE)
+          )
+IndividualData  <- IndividualData  |> 
+  mutate(EducationQuint      = cut(TotalHHEducation, breaks = quantile(TotalHHEducation, probs = c(0, 0.2, 0.4, 0.6, 0.8, 1), na.rm = TRUE), include.lowest = TRUE)
+          )
 ParentDummies <- c("HW", "uhrsworkly")
-              
+
+ParentControls  <- c("IncomeQuint", "EducationQuint")
 # By generation
 reg1 <- list(
   "\\specialcell{(1) \\\\ Log annual \\\\ earnings}" = feols(lninctot_1999 ~ HW, vcov = ~statefip,
@@ -22,9 +41,20 @@ reg1 <- list(
[…]

### 2023-05-09 programs/17-table-four-summary-stats.R

@@ -1,602 +1,0 @@
-# summary stats
-
-# table three
-
-library(Hmisc)
-
-# date: January 19th, 2023
-
-IndividualData <- read_dta(file.path(datasets,"BySexAnalysisData.dta")) |> 
-  mutate(Education = as.numeric(Education)) |> 
-  filter(Hispanic_ID == 1)
-
-row1 <- c("Variables", 
-          "\\specialcell{White \\\\ White \\\\ (WW) \\\\ (1)}",
-          "\\specialcell{White \\\\ Hispanic \\\\ (WH) \\\\ (2)}",
-          "\\specialcell{Hispanic \\\\ White \\\\ (HW) \\\\ (3)}",
-          "\\specialcell{Hispanic \\\\ Hispanic \\\\ (HH) \\\\ (4)}",
-          "\\specialcell{HH - WW \\\\ (5)}",
-          "\\specialcell{HW - WH \\\\ (6)}"
-          )
-dim(row1) <- c(1,7)
-
-# Panel A
-# Age and Education
-row6 <- c("Panel A: Children's Education ", " ", " ", " ", " ", " ", " ")
-dim(row6) <- c(1,7)
-
-## Man's Education
-ManEducation <- IndividualData |> 
-  filter(sex == 1) |> 
-  group_by(Type) |> 
-  summarise(EducationMean   = mean(Education, na.rm = T),
-            EducationSD = sd(Education, na.rm = T))
-
-model <- lm(Education ~ 0 + WW + WH + HW + HH, 
-             data = IndividualData |> filter(sex == 1))
-test1 <- tidy(glht(model, linfct = c("HH - WW = 0")))
-pvalues1 <- test1$std.error
-differences1 <- test1$estimate
-
-test2 <- tidy(glht(model, linfct = c("HW - WH = 0")))
-pvalues2 <- test2$std.error
-differences2 <- test2$estimate
-
-row7 <- c(
-  "Men’s education (Total Years)",
-  paste0("\\specialcell{", 
-         round(ManEducation[4,2], digits = 2), 
-         "\\\\",
-         "(",
-         round(ManEducation[4,3], digits = 2),
-         ")}"
-  ),
-  paste0("\\specialcell{", 
-         round(ManEducation[3,2], digits = 2), 
-         "\\\\",
-         "(",
-         round(ManEducation[3,3], digits = 2),
-         ")}"
-  ),
-  paste0("\\specialcell{", 
[…]

### 2023-05-09 programs/18-table-six-regression.R

@@ -5,9 +5,29 @@
 
 # date: January 19th, 2023
 
-IndividualData <- read_dta(file.path(datasets,"BySexAnalysisData.dta")) |> 
+# open data, filter to only have the relevant data
+# and impute empty parental charachtersitcs with mean
+
+IndividualData <- read_csv(file.path(datasets,"CPS_synth.csv")) |> 
   mutate(Education = as.numeric(Education)) |> 
-  filter((WW == 1 & Hispanic_ID == 0) | WH == 1 | HW == 1 | HH == 1)
+  filter((WW == 1 & Hispanic_ID == 0) | WH == 1 | HW == 1 | HH == 1)|> 
+  mutate(HusbandEducation = case_when(!is.na(HusbandEducation) ~ HusbandEducation,
+                                       is.na(HusbandEducation) ~ mean(HusbandEducation, na.rm = TRUE)),
+         WifeEducation = case_when(!is.na(WifeEducation) ~ WifeEducation,
+                                    is.na(WifeEducation) ~ mean(WifeEducation, na.rm = TRUE)),
+          Husband_ftotval = case_when(!is.na(Husband_ftotval) ~ Husband_ftotval,
+                                       is.na(Husband_ftotval) ~ mean(Husband_ftotval, na.rm = TRUE)))
+
+# create a total education variable
+# and a income quintile variable
+# and education quintile variable
+IndividualData  <- IndividualData  |> 
+  mutate(TotalHHEducation = HusbandEducation + WifeEducation,
+         IncomeQuint      = cut(Husband_ftotval, breaks = quantile(Husband_ftotval, probs = c(0, 0.2, 0.4, 0.6, 0.8, 1), na.rm = TRUE), include.lowest = TRUE)
+          )
+IndividualData  <- IndividualData  |> 
+  mutate(EducationQuint      = cut(TotalHHEducation, breaks = quantile(TotalHHEducation, probs = c(0, 0.2, 0.4, 0.6, 0.8, 1), na.rm = TRUE), include.lowest = TRUE)
+          )
 
 ParentDummies <- c(#"WH:Hispanic_ID", 
                    "HW:Hispanic_ID", 
@@ -17,6 +37,7 @@ ParentDummies <- c(#"WH:Hispanic_ID",
                    "HW", 
                    #"HH", 
                    "uhrsworkly")
+ParentControls  <- c("EducationQuint", "IncomeQuint")
 
 # By generation
 reg1 <- list(
@@ -30,6 +51,9 @@ reg1 <- list(
[…]

### 2023-05-09 my_paper/Hadah-last-names.tex
change latex doc and pdf to incorporate new changes

@@
 
-Using the Current Population Survey (CPS) from 1994 to 2019 and the 1970 to 1990 US censuses, I estimate the effect of having a Hispanic last name on labor market outcomes.
+Using the Current Population Survey (CPS) from 1994 to 2019 and the 1960 to 2000 US censuses, I estimate the effect of having a Hispanic last name on labor market outcomes.
 In other words, I approximate the variation in labor market perception of ethnic signals, in this case having a Hispanic last name, on annual log earnings.
@@
 
-I use two datasets:  the Integrated Public Use Microdata Series (IPUMS) Current Population Survey (CPS) Annual Social and Economic (ASEC) \citep{cps2019} and the 1970 to 1990 US censuses \citep{acs2019}.
+I use two datasets:  the Integrated Public Use Microdata Series (IPUMS) Current Population Survey (CPS) Annual Social and Economic (ASEC) \citep{cps2019} and the 1960 to 2000 US censuses \citep{acs2019}.
 
@@
 The census offers a larger sample of potential parents.
-Similar to the CPS, the 1970 to 1990 censuses ask about the person's place of birth and the individual's race and ethnicity.
+Similar to the CPS, the 1960 to 2000 censuses ask about the person's place of birth and the individual's race and ethnicity.
 I employ this information to construct "synthetic parents" using a method developed by \citet{rubinstein2014pride}.
@@
 I assume that parents have children between the ages of 25 and 40.
-I then link these synthetic parents using the year they were surveyed to the year of birth of the "children" in the CPS sample.
+I then link these synthetic parents using the birth year of child, and place of birth of parents to the year of birth of the "children" and the palce of birth of parents in the CPS sample.
 
@@
 I use the CPS for my primary analysis of the effect of having a Hispanic last name on earnings.
-I restrict my sample to Whites, United States-born citizens aged 25 to 40 and born between 1970 to 1990.
+I restrict my sample to Whites, United States-born citizens aged 25 to 40 and born between 1960 to 2000.
 Taking advantage of data on parents' place of birth, I divide the sample into four groups depending on their parent's ethnicity.
@@
 The distribution of the four types of children is presented in figure \ref{fig:dist} and table \ref{tab:mat1}.
-The majority of the sample (95.14\%) is WW children.
-The second biggest group is HH, which constitutes 3.36\% of the sample.
-Inter-ethnic children, WH and HW, make up 1.78\% of the sample with 60,147 observations.
-Even though WH and HW are only 1.78\% of the sample, I have plenty of observations to carry out an analysis.
+The majority of the sample (95.96\%) is WW children.
+The second biggest group is HH, which constitutes 2.69\% of the sample.
+Inter-ethnic children, WH and HW, make up 1.35\% of the sample with 90,325 observations.
+Even though WH and HW are only 1.35\% of the sample, I have plenty of observations to carry out an analysis.
 The summary statistics for the children of the four types of marriages are presented in table \ref{tab:c&p1}.
@@
 On the one hand, WW children are more educated, with approximately 14 years of education for men and women.
-WW children have an unemployment rate of 5\% for men and 4\% for women.
-They have log hourly earnings of 2.48 for men and 2.33 for women.
-Also, their annual log earnings are 10.25 for men and 9.46 for women.
+WW children have an unemployment rate of 4\% for men and women.
+They have log hourly earnings of 2.51 for men and 2.32 for women.
+Also, their annual log earnings are 10.29 for men and 9.43 for women.
 
-On the other hand, children of HH marriages have an average of 12.1 and 13.27 years of education for men and women, respectively.
+On the other hand, children of HH marriages have an average of 12.9 and 13.24 years of education for men and women, respectively.
 HH men and women have an unemployment rate of 7\% and 6\%.
@@
 
-WH children have an average of 13.58 and 13.87 years of education for men and women.
-The unemployment rate for men is 6\% and for women 5\%.
-HW men and women obtained 13.21 and 13.42 years of schooling.
-Men's unemployment rate is 8\% and women's 7\%.
+WH children have an average of 13.58 and 13.79 years of education for men and women.
+The unemployment rate for men and women is 5\%.
+HW men and women obtained 13.22 and 13.42 years of schooling.
+Men's unemployment rate is 7\% and women's 6\%.
 
 In panel C of table \ref{tab:c&p1}, I present the rates at which members of each group self-identify as Hispanic.
-\footnote{Hispanic self-identification is observed when an individual notes that they are Hispanic on the CPS questionnaire} The vast majority of HH children, 97\%, identify as Hispanic.
+\footnote{Hispanic self-identification is observed when an individual notes that they are Hispanic on the CPS questionnaire} The vast majority of HH children, 96\% for men and 97\% for women, identify as Hispanic.
 There is considerable attrition in Hispanic identification of inter-ethnic children.
-Among WH children, 81\% of men and 85\% of women identify as Hispanic.
-Among HW children, 88\% of men and 87\% of women identify as Hispanic.
+Among WH children, 74\% of men and 78\% of women identify as Hispanic.
+Among HW children, 83\% of men and 82\% of women identify as Hispanic.
 These results align with the findings of a substantial body of literature that noted this attrition among children of immigrants, especially those coming from inter-ethnic marriages \citep{duncan2017complexity, duncan2018identifying, duncan2020new, antman2020ethnic}.
-Finally, a small portion of WW children, 5\% of men and 6\% of women, identified as Hispanic.
-These 6\% of WW children are most likely third generation--- or higher--- children of immigrants.
+Finally, a small portion of WW children, 4\% of men and 5\% of women, identified as Hispanic.
+These 5\% of WW children are most likely third generation--- or higher--- children of immigrants.
 For my analysis, I dropped the WW observations who identify as Hispanic when I compare with non-Hispanic Whites.
@@
 Overall, HH children do worse than WW children.
-HH has one year less of education compared to WW, and are 2 percentage points less likely to be employed.
-HH men earn 7\% less than WW men, and HH women make 2\% less than WW women.
+HH has one year less of education compared to WW, and are 2 percentage points more likely to be unemployed.
+HH men earn 9\% less than WW men, and HH women make 2\% less than WW women.
 Therefore, these two groups, HH and WW, are different on many observable qualities.
@@
 HW men attain 0.36 fewer years of education than WH men.
-HW women achieve 0.46 fewer years of schooling than WH women.
-HW children are 2 percentage points less likely to be employed.
-HW men earn 4\% less annual earnings as WH men, and HW women earn 7\% less.
+HW women achieve 0.37 fewer years of schooling than WH women.
+HW children are 1 percentage points more likely to be unemployed.
+HW men earn 1\% less annual earnings as WH men, and HW women earn 3\% less.
 These differences show that HW and WH children are different from children of endogamous/homogamous marriages and, thus, are more comparable.
@@
 
-Using the 1970 to 1990 censuses, I constructed a data set of synthetic parents.
+Using the 1960 to 2000 censuses, I constructed a data set of synthetic parents.
 The sample includes married White men and women.
@@
 I assume that parents have children between the ages of 25 and 40.
-Therefore, my sample consists of married White men and women that were born in the 1930 to 1965 cohorts\footnote{The construction of "synthetic parents" follows the method used by \citet{rubinstein2014pride}.}.
+Therefore, my sample consists of married White men and women with children that were born in the 1920 to 1975 cohorts\footnote{The construction of "synthetic parents" follows the method used by \citet{rubinstein2014pride}.}.
 
 I show the distribution of the four types of couples in table \ref{tab:mat2}.
-White husbands and White wives (WW) make up the majority of couples in the sample, 97\% (1,286,731 couples).
-Hispanic husbands and wives (HH) are the second largest group making up 2\% (20,911 couples)of all couples.
-White husbands and Hispanic wives (WH) couples are less than 1\% (7,178 couples) of the sample and Hispanic husbands and White wives (HW) are also less than 1\% (7,606 couples).
+White husbands and White wives (WW) make up the majority of couples in the sample, 96\% (5,141,737 couples).
+Hispanic husbands and wives (HH) are the second largest group making up 2\% (119,749 couples) of all couples.
+White husbands and Hispanic wives (WH) couples are less 1\% (33,097 couples) of the sample and Hispanic husbands and White wives (HW) are also 1\% (37,847 couples).
 
 I present the summary statistics of the four types of synthetic parents in table \ref{tab:synth}.
-WW couples have higher education, 13.05 years for husbands and 12.74 for wives.
-As a household, WW couples have 25.78 years of schooling.
-Husbands in HH marriages have 8.93 years of education, while women have 8.6 years of schooling.
-As a household, HH couples have 17.54 years of education.
+WW couples have higher education, 12.75 years for husbands and 12.47 for wives.
+As a household, WW couples have 25.22 years of schooling.
+Additionally, the log total family income for a WW household is 10.72 and have 3.77 children.
+Husbands in HH marriages have 8.64 years of education, while women have 8.49 years of schooling.
+As a household, HH couples have 17.13 years of education, log total family income of 10.27, and have 4.23 children.
 
-WH husbands have 12.32 years of education, while wives have an average of 11.03 years.
-WH household attained a total of 23.35 years of schooling in total.
-HW husbands and wives obtained 10.65 and 11.54 years of education, respectively.
-HW household attained a total of 22.19 years of schooling in total.
+WH husbands have 11.77 years of education, while wives have an average of 10.40 years.
+WH household attained a total of 22.17 years of schooling in total, a log total family income of 10.55, and 3.98 children.
+HW husbands and wives obtained 10.25 and 11.11 years of education, respectively.
+HW household attained a total of 21.36 years of schooling in total, a log total family income of 10.46, and 4.15 children.
 
@@
 An HH Hispanic husband attained 4.11 fewer years of education than a WW husband.
-While an HH wife acquired 4.13 fewer years of schooling than a White WW wife.
-The total educational difference between HH and WW households equals 8.25 years of education.
+While an HH wife acquired 3.98 fewer years of schooling than a White WW wife.
+The total educational difference between HH and WW households equals 8.09 years of education.
 The educational differences between HW and WH are less severe than those between HH and WW.
-A Hispanic husband in intermarriage attained 1.67 fewer years of schooling than a White husband in intermarriage.
-A White wife in intermarriage attained 0.51 years of education, more than a Hispanic wife in intermarriage.
-The difference in total years of education between HW and WH households is 1.16, which is an 85\% decrease from the HH--WW gap.
+A Hispanic husband in intermarriage attained 1.52 fewer years of schooling than a White husband in intermarriage.
+A White wife in intermarriage attained 0.70 years of education, more than a Hispanic wife in intermarriage.
+The difference in total years of education between HW and WH households is 0.81, which is an 90\% decrease from the HH--WW gap.
+The total family income of an HH household is 45\% lower than a WW household, while the total family income of an HW household is 10\% lower than a WH household.
 These differences show that couples that intermarry are better and different than those that do not.
@@
 Column one in table \ref{tab:identreg} is the results without controlling for education.
-Overall, an inter-ethnic that identifies as Hispanic earns 11 percentage points less than an inter-ethnic that does not.
+Overall, an inter-ethnic that identifies as Hispanic earns 13 percentage points less than an inter-ethnic that does not.
 This gap, however, could be entirely explained by educational differences.
 An inter-ethnic that identifies as Hispanic---i.e.
-$HW_{ist} + HW_{ist} \cdot Hispanic_i$---earns 6 percentage points less than an inter-ethnic that does not, but the result is statistically insignificant.
+$HW_{ist} + HW_{ist} \cdot Hispanic_i$---earns 6 percentage points less than an inter-ethnic that does not, but the result could be entirely explained by educational differences.
 
@@
 
-% \begin{table}[H]
-%   \centering
-%   \renewcommand*\TPTnoteLabel[1]{\parbox[b]{3em}{\hfill#1\,}}
-%   \begin{adjustbox}{max width=\textwidth}
-%   \begin{threeparttable}
-% \caption{\textsc{Number of children from the different types of parents.}}\label{tab:mat1}
-% \footnotesize
-% \begin{tabular}{cccccccccccccc}
-% \hline 
-%  &  &  & \multicolumn{10}{c}{Perent's Type} & \tabularnewline
-% \cline{4-13} 
-%  &  &  & White Father &  &  & White Father &  &  & Hispanic father &  &  & Hispanic father & \tabularnewline
-%  &  &  & White Mother &  &  & Hispanic Mother &  &  & White Mother &  &  & Hispanic Mother & \tabularnewline
-% \hline 
-%  &  &  &  &  &  &  &  &  &  &  &  &  & \tabularnewline
-%  & Number &  & 3,828,024 &  &  & 24,306 &  &  & 35,841 &  &  & 135,300 & \tabularnewline
-%  & (\%) &  & (95.14\%) &  &  & (0.89\%) &  &  & (0.89\%) &  &  & (3.36\%) & \tabularnewline
-%  &  &  &  &  &  &  &  &  &  &  &  &  & \tabularnewline
-% \hline 
-% \end{tabular}
-%  \begin{tablenotes}
-%    \item [\textit{Notes.}] The data is restricted to people interviewed between 1994 and 2019, also White, married, and between the ages of 18 and 65.
-I identify the ethnicity of a person's parents through the parent's place of birth.
-A parent is Hispanic if both her parents were born in a Spanish-speaking country.
-A parent is White if born parents were born in the United States.
-%   \end{tablenotes}
-%   \end{threeparttable}
-%   \end{adjustbox}
-% \end{table}
-
 \newpage
@@
 \input{tables/tab02-observations-by-couple}
-
-% \begin{table}[H]
-% \begin{adjustbox}{max width=\textwidth}
-%   \centering
-%   \renewcommand*\TPTnoteLabel[1]{\parbox[b]{3em}{\hfill#1\,}}
-%   \begin{threeparttable}
-% \caption{\textsc{Parents' types}}\label{tab:mat2}
-% \footnotesize
-% \begin{tabular}{cccccccccccccc}
-% \hline 
-%  &  &  & \multicolumn{10}{c}{Couples' Type} & \tabularnewline
-% \cline{4-13} 
-%  &  &  & White Husband &  &  & White Husband &  &  & Hispanic Husband &  &  & Hispanic Husband & \tabularnewline
-%  &  &  & White Wife &  &  & Hispanic Wife &  &  & White Wife &  &  & Hispanic Wife & \tabularnewline
-% \hline 
-%  &  &  &  &  &  &  &  &  &  &  &  &  & \tabularnewline
-%  & Number &  & 1,198,446 &  &  & 3,819 &  &  & 3,413 &  &  & 6,920 & \tabularnewline
-%  & (\%) &  & (98.83\%) &  &  & (0.31\%) &  &  & (0.28\%) &  &  & (0.57\%) & \tabularnewline
-%  &  &  &  &  &  &  &  &  &  &  &  &  & \tabularnewline
-% \hline 
-% \end{tabular}
-
-%   \begin{tablenotes}
-%     \item [\textit{Notes.}] The data is restricted to people interviewed in 1970 and 1960 and also White and married.
-I identify the ethnicity of a person through their place of birth.
-A parent is Hispanic if they were born in a Spanish-speaking country.
-A parent is White if they were born in the United States.
-%     \item [\textit{Notes.}] The table includes information on the proportion of the four types of synthetic parents that I have constructed.
-%   \end{tablenotes}
-%   \end{threeparttable}
-% \end{adjustbox}
-% \end{table}
 
@@
 
-% \newpage
-
-% \begin{table}[H]
-%   \centering
-%   \renewcommand*\TPTnoteLabel[1]{\parbox[b]{3em}{\hfill#1\,}}
-%   \begin{adjustbox}{max width=\textwidth}
-%   \begin{threeparttable}
-% \caption{\textsc{Children's outcome using parent's place of birth}}\label{tab:c&p1}
-% \footnotesize
-% \begin{tabular}{llccccccccr@{\extracolsep{0pt}.}l}
-% \toprule 
-%  &  & \multicolumn{4}{c}{Father's and Mother's Ethnicities} &  & \multicolumn{5}{c}{Differences}\tabularnewline
-% \cmidrule{3-6} \cmidrule{8-12} 
-%  &  &  &  &  &  &  &  &  &  & \multicolumn{2}{c}{}\tabularnewline
-% Variables &  & White Father & White Father & Hispanic Father & Hispanic Father &  & HH - WW &  & HW-WH & \multicolumn{2}{c}{}\tabularnewline
-%  &  & White Mother & Hispanic Mother & White Mother & Hispanic Mother &  &  &  &  & \multicolumn{2}{c}{}\tabularnewline
-%  &  & (WW) & (WH) & (HW) & (HH) &  &  &  &  & \multicolumn{2}{c}{}\tabularnewline
-%  &  & (i) & (ii) & (iii) & (iv) &  & (v) &  & (vi) & \multicolumn{2}{c}{}\tabularnewline
-% \midrule
-% \textbf{Parent's Panel:} &  &  &  &  &  &  &  &  &  & \multicolumn{2}{c}{}\tabularnewline
-% Husband's education (total years) &  & 12.727 & 11.847 & 10.286 & 8.949 &  & -3.779{*}{*}{*} &  & -1.560{*}{*}{*} & \multicolumn{2}{c}{}\tabularnewline
-%  &  & (2.616) & (3.658) & (4.602) & (4.422) &  & (0.018) &  & (0.041) & \multicolumn{2}{c}{}\tabularnewline
-% Wife's education (total years) &  & 12.467 & 10.462 & 11.188 & 8.590 &  & -3.876{*}{*}{*} &  & 0.726{*}{*}{*} & \multicolumn{2}{c}{}\tabularnewline
-%  &  & (2.182) & (4.139) & (3.394) & (4.120) &  & (0.015) &  & (0.035) & \multicolumn{2}{c}{}\tabularnewline
-% Total Household education (years) &  & 25.194 & 22.308 & 21.473 & 17.539 &  & -7.655{*}{*}{*} &  & -0.835{*}{*}{*} & \multicolumn{2}{c}{}\tabularnewline
-%  &  & (4.300) & (7.034) & (7.202) & (7.836) &  & (0.029) &  & (0.068) & \multicolumn{2}{c}{}\tabularnewline
-% \textbf{Panel A:} &  &  &  &  &  &  &  &  &  & \multicolumn{2}{c}{}\tabularnewline
-% \textbf{age and education} &  &  &  &  &  &  &  &  &  & \multicolumn{2}{c}{}\tabularnewline
-% man's Education &  & 13.861 & 13.366 & 13.072 & 12.898 &  & -0.962{*}{*}{*} &  & -0.294{*}{*}{*} & \multicolumn{2}{c}{}\tabularnewline
-%  &  & (2.421) & (2.317) & (2.270) & (2.356) &  &  &  &  & \multicolumn{2}{c}{}\tabularnewline
-% woman's Education &  & 14.200 & 13.741 & 13.270 & 13.204 &  & -0.996{*}{*}{*} &  & -0.471{*}{*}{*} & \multicolumn{2}{c}{}\tabularnewline
-%  &  & (2.424) & (2.467) & (2.342) & (2.453) &  &  &  &  & \multicolumn{2}{c}{}\tabularnewline
-% \textbf{Panel B: Employment \& Earnings} &  &  &  &  &  &  &  &  &  & \multicolumn{2}{c}{}\tabularnewline
-% men' employment rate &  & 0.952 & 0.945 & 0.927 & 0.931 &  & -0.021{*}{*}{*} &  & -0.018{*}{*}{*} & \multicolumn{2}{c}{}\tabularnewline
-%  &  & (0.215) & (0.227) & (0.260) & (0.254) &  &  &  &  & \multicolumn{2}{c}{}\tabularnewline
-% women' employment rate &  & 0.956 & 0.950 & 0.931 & 0.936 &  & -0.020{*}{*}{*} &  & -0.019{*}{*}{*} & \multicolumn{2}{c}{}\tabularnewline
-%  &  & (0.205) & (0.218) & (0.253) & (0.244) &  &  &  &  & \multicolumn{2}{c}{}\tabularnewline
-% men' log hourly earnings &  & 2.557 & 2.524 & 2.468 & 2.505 &  & -0.041{*}{*}{*} &  & -0.056{*}{*}{*} & \multicolumn{2}{c}{}\tabularnewline
-%  &  & (0.399) & (0.421) & (0.423) & (0.401) &  &  &  &  & \multicolumn{2}{c}{}\tabularnewline
-% women' log hourly earnings &  & 2.435 & 2.444 & 2.385 & 2.389 &  & -0.016{*}{*} &  & -0.061{*}{*}{*} & \multicolumn{2}{c}{}\tabularnewline
-%  &  & (0.415) & (0.426) & (0.375) & (0.376) &  &  &  &  & \multicolumn{2}{c}{}\tabularnewline
-% men' log annual earnings &  & 10.529 & 10.435 & 10.386 & 10.365 &  & -0.164{*}{*}{*} &  & -0.049 & \multicolumn{2}{c}{}\tabularnewline
-%  &  & (0.703) & (0.637) & (0.872) & (0.622) &  &  &  &  & \multicolumn{2}{c}{}\tabularnewline
-% women' log annual earnings &  & 10.297 & 10.262 & 10.159 & 10.175 &  & -0.123{*}{*}{*} &  & -0.103{*}{*}{*} & \multicolumn{2}{c}{}\tabularnewline
-%  &  & (0.633) & (0.613) & (0.541) & (0.650) &  &  &  &  & \multicolumn{2}{c}{}\tabularnewline
-% \textbf{Panel D: Identity} &  &  &  &  &  &  &  &  &  & \multicolumn{2}{c}{}\tabularnewline
-% Proportion of men &  & 0.060 & 0.821 & 0.879 & 0.969 &  &  &  &  & \multicolumn{2}{c}{}\tabularnewline
-% identifying as Hispanic &  &  &  &  &  &  &  &  &  & \multicolumn{2}{c}{}\tabularnewline
-% Proportion of women &  & 0.059 & 0.850 & 0.859 & 0.970 &  &  &  &  & \multicolumn{2}{c}{}\tabularnewline
-% identifying as Hispanic &  &  &  &  &  &  &  &  &  & \multicolumn{2}{c}{}\tabularnewline
-%  &  &  &  &  &  &  &  &  &  & \multicolumn{2}{c}{}\tabularnewline
-% \bottomrule
-% \end{tabular}
-%   \begin{tablenotes}
-%    \item [\textit{Notes.}] The data is restricted to native-born United States citizens between 1994 and 2019 who are also White and between the ages of 25 and 40.
-I identify the ethnicity of a person's parents through the parent's place of birth.
-A parent is Hispanic if they were born in a Spanish-speaking country.
-A parent is White if they were born in the United States.
-%    \item [\textit{Notes.}] In each column, I present the average statistics of the different types of people based on the ethnicities of their parents.
-In column one, I show the summary statistics of children of White fathers and White mothers.
-In column two, I present the summary statistics of children of White fathers and Hispanic mothers.
-In column three, I show the summary statistics of children of Hispanic fathers and White mothers.
-In column four, I present the summary statistics of children of Hispanic fathers and mothers.
-%    \item [\textit{Notes.}] Columns five and six have data on the HH--WW gaps (column five) and the HW--WH gaps (column six).
-
-%   \end{tablenotes}
-%   \end{threeparttable}
-%   \end{adjustbox}
-% \end{table}
-
 \newpage
@@
 \input{tables/tab04-summary-stats}
-
-% \begin{table}[H]
-%   \centering
-%   \renewcommand*\TPTnoteLabel[1]{\parbox[b]{3em}{\hfill#1\,}}
-%   \begin{adjustbox}{max width=\textwidth}
-%   \begin{threeparttable}
-% \caption{\textsc{Children's outcomes using parent's place of birth only for those that identify as Hispanic}}\label{tab:c&p2}
-% \footnotesize
-% \begin{tabular}{llccccccccr@{\extracolsep{0pt}.}l}
-% \hline 
-%  &  & \multicolumn{4}{c}{Father's and Mother's Ethnicities} &  & \multicolumn{5}{c}{Differences}\tabularnewline
-% \cline{3-6} \cline{8-12} 
-%  &  &  &  &  &  &  &  &  &  & \multicolumn{2}{c}{}\tabularnewline
-% Variables &  & White Father & White Father & Hispanic Father & Hispanic Father &  & HH - WW &  & HW-WH & \multicolumn{2}{c}{}\tabularnewline
-%  &  & White Mother & Hispanic Mother & White Mother & Hispanic Mother &  &  &  &  & \multicolumn{2}{c}{}\tabularnewline
-%  &  & (WW) & (WH) & (HW) & (HH) &  &  &  &  & \multicolumn{2}{c}{}\tabularnewline
-%  &  & (i) & (ii) & (iii) & (iv) &  & (v) &  & (vi) & \multicolumn{2}{c}{}\tabularnewline
-% \hline 
-% \textbf{Parent's Panel:} &  &  &  &  &  &  &  &  &  & \multicolumn{2}{c}{}\tabularnewline
-% Husband's education (total years) &  & 11.856 & 10.211 & 9.665 & 8.427 &  & -3.429{*}{*}{*} &  & -0.546{*}{*}{*} & \multicolumn{2}{c}{}\tabularnewline
-%  &  & (2.922) & (3.876) & (4.095) & (4.228) &  & (0.035) &  & (0.079) & \multicolumn{2}{c}{}\tabularnewline
-% Wife's education (total years) &  & 11.793 & 9.556 & 9.791 & 8.137 &  & -3.656{*}{*}{*} &  & 0.235{*}{*}{*} & \multicolumn{2}{c}{}\tabularnewline
-%  &  & (2.329) & (3.751) & (3.524) & (3.899) &  & (0.028) &  & (0.063) & \multicolumn{2}{c}{}\tabularnewline
-% Total Household education (years) &  & 23.650 & 19.767 & 19.456 & 16.564 &  & -7.085{*}{*}{*} &  & -0.311{*}{*}{*} & \multicolumn{2}{c}{}\tabularnewline
-%  &  & (4.716) & (6.843) & (6.896) & (7.338) &  & (0.057) &  & (0.128) & \multicolumn{2}{c}{}\tabularnewline
-% \textbf{Panel A:} &  &  &  &  &  &  &  &  &  & \multicolumn{2}{c}{}\tabularnewline
-% \textbf{age and education} &  &  &  &  &  &  &  &  &  & \multicolumn{2}{c}{}\tabularnewline
-% man's Education &  & 13.861 & 13.237 & 13.008 & 12.874 &  & -0.986{*}{*}{*} &  & -0.230{*}{*}{*} & \multicolumn{2}{c}{}\tabularnewline
-%  &  & (2.421) & (2.313) & (2.217) & (2.347) &  &  &  &  & \multicolumn{2}{c}{}\tabularnewline
-% woman's Education &  & 14.200 & 13.603 & 13.133 & 13.182 &  & -1.018{*}{*}{*} &  & -0.470{*}{*}{*} & \multicolumn{2}{c}{}\tabularnewline
-%  &  & (2.424) & (2.391) & (2.325) & (2.453) &  &  &  &  & \multicolumn{2}{c}{}\tabularnewline
-% \textbf{Panel D: Employment \& Earnings} &  &  &  &  &  &  &  &  &  & \multicolumn{2}{c}{}\tabularnewline
-% men' employment rate &  & 0.952 & 0.943 & 0.925 & 0.930 &  & -0.021{*}{*}{*} &  & -0.018{*}{*}{*} & \multicolumn{2}{c}{}\tabularnewline
-%  &  & (0.215) & (0.232) & (0.264) & (0.254) &  &  &  &  & \multicolumn{2}{c}{}\tabularnewline
-% women' employment rate &  & 0.956 & 0.948 & 0.928 & 0.936 &  & -0.020{*}{*}{*} &  & -0.020{*}{*}{*} & \multicolumn{2}{c}{}\tabularnewline
-%  &  & (0.205) & (0.222) & (0.258) & (0.245) &  &  &  &  & \multicolumn{2}{c}{}\tabularnewline
-% men' log hourly earnings &  & 2.501 & 2.470 & 2.408 & 2.460 &  & -0.041{*}{*}{*} &  & -0.062{*}{*}{*} & \multicolumn{2}{c}{}\tabularnewline
-%  &  & (0.454) & (0.458) & (0.463) & (0.434) &  &  &  &  & \multicolumn{2}{c}{}\tabularnewline
-% women' log hourly earnings &  & 2.339 & 2.358 & 2.288 & 2.322 &  & -0.017{*}{*} &  & -0.071{*}{*}{*} & \multicolumn{2}{c}{}\tabularnewline
-%  &  & (0.494) & (0.499) & (0.464) & (0.423) &  &  &  &  & \multicolumn{2}{c}{}\tabularnewline
-% men' log annual earnings &  & 10.529 & 10.414 & 10.394 & 10.364 &  & -0.165{*}{*}{*} &  & -0.020 & \multicolumn{2}{c}{}\tabularnewline
-%  &  & (0.703) & (0.610) & (0.704) & (0.625) &  &  &  &  & \multicolumn{2}{c}{}\tabularnewline
-% women' log annual earnings &  & 10.297 & 10.240 & 10.136 & 10.172 &  & -0.126{*}{*}{*} &  & -0.104{*}{*}{*} & \multicolumn{2}{c}{}\tabularnewline
-%  &  & (0.634) & (0.635) & (0.537) & (0.652) &  &  &  &  & \multicolumn{2}{c}{}\tabularnewline
-%  &  &  &  &  &  &  &  &  &  & \multicolumn{2}{c}{}\tabularnewline
-% \hline 
-% \end{tabular}
-%   \begin{tablenotes}
-%    \item [\textit{Notes.}] The data is restricted to native-born United States citizens between 1994 and 2019 who are also White and between the ages of 25 and 40.
-I identify the ethnicity of a person's parents through the parent's place of birth.
-A parent is Hispanic if they were born in a Spanish-speaking country.
-A parent is White if they were born in the United States.
-%    \item [\textit{Notes.}] In each column, I present the average statistics of those that identify as Hispanic by the different types of people based on the ethnicities of their parents.
-In column one, I present the summary statistics of children of White fathers and White mothers.
-In column two, I show the summary statistics of children of White fathers and Hispanic mothers identifying as Hispanic.
-In column three, I present the summary statistics of children of Hispanic fathers and White mothers that identify as Hispanic.
-In column four, I show the summary statistics of children of Hispanic fathers and Hispanic mothers that identify as Hispanic.
-%    \item [\textit{Notes.}] Columns five and six have data on the HH--WW gaps (column five) and the HW--WH gaps (column six).
-
-%   \end{tablenotes}
-%   \end{threeparttable}
-%   \end{adjustbox}
-% \end{table}
 
@@
 
-% \begin{table}[H]
-%   \centering
-%   \renewcommand*\TPTnoteLabel[1]{\parbox[b]{3em}{\hfill#1\,}}
-%     \begin{adjustbox}{max width=\textwidth}
-%   \begin{threeparttable}
-% \caption{\textsc{Effect of having a Hispanic last name}}\label{tab:lastnamereg}
-% \footnotesize
-
-% \begin{tabular}{llcccc}
-% \toprule 
-%  &  &  &  &  & \tabularnewline
-% Variables &  & Log annual earnings &  & Log annual earnings & \tabularnewline
-% \midrule
-
-% $WH_{ist}$ &  & -0.088{*}{*}{*} &  & -0.044{*}{*}{*} & \tabularnewline
-%  &  & (0.018) &  & (0.017) & \tabularnewline
-%  &  &  &  &  & \tabularnewline
-% $HW_{ist}$ &  & -0.141{*}{*}{*} &  & -0.059{*}{*}{*} & \tabularnewline
-%  &  & (0.015) &  & (0.014) & \tabularnewline
-%  &  &  &  &  & \tabularnewline
-% $HH_{ist}$ &  & -0.161{*}{*}{*} &  & -0.058{*}{*}{*} & \tabularnewline
-%  &  & (0.008) &  & (0.007) & \tabularnewline
-%  &  &  &  &  & \tabularnewline
-% \midrule
-% $HW_{ist}$ - $WH_{ist}$ &  & -0.053{*}{*} &  & -0.016 & \tabularnewline
-%  &  & (0.024) &  & (0.022) & \tabularnewline
-%  &  &  &  &  & \tabularnewline
-% \midrule
-% \textit{Controlling for:} &  &  &  &  & \tabularnewline
-% Education &  & No &  & Yes & \tabularnewline
-% Hours worked &  & Yes &  & Yes & \tabularnewline
-% Age &  & Yes &  & Yes & \tabularnewline
-% Year FE &  & Yes &  & Yes & \tabularnewline
-% \bottomrule
-% \end{tabular}
-
-%   \begin{tablenotes}
-%    \item [\textit{Notes.}] This table includes the estimation results of equation \ref{eq:1a}.
-%    \item [\textit{Notes.}] The four groups stand for White Husband White Wife (WW), White Husband Hispanic Wife (WH), Hispanic Husband White (HW), and Hispanic Husband Hispanic Wife (HH).
-%    \item [\textit{Notes.}] The sample is restricted to men working full-time full-year and are waged and salaried workers.
-%     \item [\textit{Notes.}] Column one has the regression results when controlling for hours worked, age, and fixed effects. Column two has the results after controlling for education.
-%   \end{tablenotes}
-%   \end{threeparttable}
-%   \end{adjustbox}
-% \end{table}
-
 \newpage
@@
 
-% \begin{table}[H]
-%   \centering
-%   \renewcommand*\TPTnoteLabel[1]{\parbox[b]{3em}{\hfill#1\,}}
-%     \begin{adjustbox}{max width=\textwidth}
-%   \begin{threeparttable}
-% \caption{\textsc{Last name effect}}\label{tab:identreg}
-% \footnotesize
-
-% \begin{tabular}{lccccc}
-% \hline 
-%  &  & Log annual earnings &  & Log annual earnings & \tabularnewline
-% Variables &  &  &  &  & \tabularnewline
-% \hline 
-%  &  &  &  &  & \tabularnewline
-% $WH_{ist}\times Hispanic_{ist}$ &  & -0.134{*}{*} &  & -0.064 & \tabularnewline
-%  &  & (0.053) &  & (0.049) & \tabularnewline
-% $HW_{ist}\times Hispanic_{ist}$ &  & -0.164{*}{*}{*} &  & -0.108{*}{*} & \tabularnewline
-%  &  & (0.052) &  & (0.048) & \tabularnewline
-% $HH_{ist}\times Hispanic_{ist}$ &  & -0.106{*}{*} &  & -0.062 & \tabularnewline
-%  &  & (0.049) &  & (0.045) & \tabularnewline
-% $WH_{ist}$ &  & 0.027 &  & 0.011 & \tabularnewline
-%  &  & (0.048) &  & (0.045) & \tabularnewline
-% $HW_{ist}$ &  & 0.007 &  & 0.038 & \tabularnewline
-%  &  & (0.050) &  & (0.046) & \tabularnewline
-% $HH_{ist}$ &  & -0.057 &  & 0.003 & \tabularnewline
-%  &  & (0.057) &  & (0.044) & \tabularnewline
-% \hline 
-% $HW_{ist}\times Hispanic_{ist}$ - $WH_{ist}\times Hispanic_{ist}$ &  & -0.030 &  & -0.043 & \tabularnewline
-%  &  & (0.074) &  & (0.068) & \tabularnewline
-% \hline 
-% \textit{Controlling for:} &  &  &  &  & \tabularnewline
-% Education &  & No &  & Yes & \tabularnewline
-% Hours worked &  & Yes &  & Yes & \tabularnewline
-% Age &  & Yes &  & Yes & \tabularnewline
-% Year FE &  & Yes &  & Yes & \tabularnewline
-% \hline 
-% \end{tabular}
-
-%   \begin{tablenotes}
-%    \item [\textit{Notes.}] This table includes the estimation results of equation \ref{eq:iden}.
-%    \item [\textit{Notes.}] The four groups stand for White Husband White Wife (WW), White Husband Hispanic Wife (WH), Hispanic Husband White (HW), and Hispanic Husband Hispanic Wife (HH).
-$Hispanic_{ist}$ is a dummy variable equal to one if a person identifies as Hispanic and zero otherwise.
-%    \item [\textit{Notes.}] The sample is restricted to men working full-time full-year and are waged and salaried workers.
-%     \item [\textit{Notes.}] Column one has the regression results when controlling for hours worked, age, and years of fixed effects. Column two has the results after controlling for education.
-%   \end{tablenotes}
-%   \end{threeparttable}
-%   \end{adjustbox}
-% \end{table}
-
-% \begin{table}[H]
-%   \centering
-%   \renewcommand*\TPTnoteLabel[1]{\parbox[b]{3em}{\hfill#1\,}}
-%     \begin{adjustbox}{max width=\textwidth}
-%   \begin{threeparttable}
-% \caption{\textsc{Hsiapnic--White earnings gap using Hispanic variable only}}\label{tab:hispwhitegap}
-% \footnotesize
-
-% \begin{tabular}{lccccc}
-% \hline 
-%  &  &  &  &  & \tabularnewline
-% Variables &  & Log annual earnings &  & Log annual earnings & \tabularnewline
-% \hline 
-% $Hispanic_{ist}$ &  & -0.156{*}{*}{*} &  & -0.060{*}{*}{*} & \tabularnewline
-%  &  & (0.006) &  & (0.006) & \tabularnewline
-% \textit{Controlling for:} &  &  &  &  & \tabularnewline
-% Education &  & No &  & Yes & \tabularnewline
-% Hours worked &  & Yes &  & Yes & \tabularnewline
-% Age &  & Yes &  & Yes & \tabularnewline
-% Year FE &  & Yes &  & Yes & \tabularnewline
-% \hline 
-% \end{tabular}
-
-%   \begin{tablenotes}
-%    \item [\textit{Notes.}] This table includes the estimation results of equation \ref{eq:naive}.
-%    \item [\textit{Notes.}] $Hispanic_{ist}$ is a dummy variable equal to one if a person identifies as Hispanic and zero otherwise.
-%    \item [\textit{Notes.}] The sample is restricted to men working full time full-year and are waged and salaried workers.
-%     \item [\textit{Notes.}] Column one has the regression results when controlling for hours worked, age, and fixed effects. Column two has the results after controlling for education.
-%   \end{tablenotes}
-%   \end{threeparttable}
-%   \end{adjustbox}
-% \end{table}
-
 \end{appendices}
 
-%\begin{table}[!htbp]
-%\caption{T-tests across  groups by several variables} \label{tab:table8}
-%\resizebox{0.95\textwidth}{!}{
-%	  \estauto{sumstat_parents2.tex}{8}{c}
-%	  }
-%\small{Note:  *** p$<$0.01, ** p$<$0.05, * p$<$0.10.}  
-%\end{table}
-
 \end{document}

### 2023-05-09 programs/12-table-two-couple-types.R

@@ -108,7 +108,7 @@ knitr::kable(Table_cols, "html", align = "lcccc",
   kable_classic(full_width = F) %>%
   kable_styling(bootstrap_options = c("hover", "condensed"), 
                 latex_options = c("scale_down", "HOLD_position")) |> 
-  footnote(number = c("Source: 1970-1990 Census",
+  footnote(number = c("Source: 1960-1990 Census",
                       "The sample includes Whites, who are married, and are between the ages 25 and 40. Ethnicity of a person's parents are identified by the parent's place of birth. A parent is Hispanic if she/he was born in a Spanish-speaking country. A parent is White if she/he was born in the United States."),
            footnote_as_chunk = F, title_format = c("italic"),
            escape = F, threeparttable = T
@@ -122,7 +122,7 @@ knitr::kable(Table_cols, "latex", align = "lcccc",
   kable_classic(full_width = F) %>%
   kable_styling(bootstrap_options = c("hover", "condensed"), 
                 latex_options = c("scale_down", "HOLD_position")) |> 
-  footnote(number = c("Source: 1950-2000 Census",
+  footnote(number = c("Source: 1960-2000 Census",
                       "The sample includes Whites, who are married, with kids, and are between the ages 25 and 40. Ethnicity of a person's parents are identified by the parent's place of birth. A parent is Hispanic if she/he was born in a Spanish-speaking country. A parent is White if she/he was born in the United States."),
            footnote_as_chunk = F, title_format = c("italic"),
            escape = F, threeparttable = T

### 2023-05-09 programs/14-table-three-summary-stats.R

@@ -166,7 +166,7 @@ pvalues2 <- test2$std.error
 differences2 <- test2$estimate
 
 row5 <- c(
-  "Total Household seducation (Total Years)",
+  "Total Household education (Total Years)",
   paste0("\\specialcell{", 
          sprintf("%0.2f",round(TotalEducation[4,2], digits = 2)),  
          "\\\\",
@@ -362,7 +362,7 @@ knitr::kable(Table_rows, "html", align = "lcccccc",
                 "HOLD_position"
               #   "repeat_header"
                 )) |> 
-  footnote(number = c("Source: The 1950-2000 Census for synthetic parents, and 1994-2019 Current Population Surveys (CPS) for children's outcomes",
+  footnote(number = c("Source: The 1960-2000 Census for synthetic parents, and 1994-2019 Current Population Surveys (CPS) for children's outcomes",
                       "The data is restricted to native-born United States citizens who are also White, between the ages of 25 and 40, and have kids. I identify the ethnicity of a person's parents through the parent's place of birth. A parent is Hispanic if they were born in a Spanish-speaking country. A parent is White if they were born in the United States."),
            footnote_as_chunk = F, title_format = c("italic"),
            escape = F, threeparttable = T
@@ -382,7 +382,7 @@ knitr::kable(Table_rows, "latex", align = "lcccccc",
                 "HOLD_position"
               #   "repeat_header"
                 )) |> 
-  footnote(number = c("Source: The 1950-2000 Census for synthetic parents, and 1994-2019 Current Population Surveys (CPS) for children's outcomes",
+  footnote(number = c("Source: The 1960-2000 Census for synthetic parents, and 1994-2019 Current Population Surveys (CPS) for children's outcomes",
                       "The data is restricted to native-born United States citizens who are also White, between the ages of 25 and 40, and have kids. I identify the ethnicity of a person's parents through the parent's place of birth. A parent is Hispanic if they were born in a Spanish-speaking country. A parent is White if they were born in the United States."),
            footnote_as_chunk = F, title_format = c("italic"),
            escape = F, threeparttable = T

### 2023-05-09 programs/15-table-three-summary-stats.R

@@ -650,7 +650,7 @@ knitr::kable(Table_rows, "latex", align = "lcccccc",
                 "repeat_header"
                 ),
                 repeat_header_continued = "\\textit{(Continued on Next Page...)}") |> 
-  footnote(number = c("Source: The 1950-2000 Census for synthetic parents, and 1994-2019 Current Population Surveys (CPS) for children's outcomes",
+  footnote(number = c("Source: The 1960-2000 Census for synthetic parents, and 1994-2019 Current Population Surveys (CPS) for children's outcomes",
                       "The data is restricted to native-born United States citizens who are also White and between the ages of 25 and 40. I identify the ethnicity of a person's parents through the parent's place of birth. A parent is Hispanic if they were born in a Spanish-speaking country. A parent is White if they were born in the United States."),
            footnote_as_chunk = F, title_format = c("italic"),
            escape = F, threeparttable = T
@@ -680,7 +680,7 @@ knitr::kable(Table_rows, "html", align = "lcccccc",
                 "repeat_header"
                 ),
                 repeat_header_continued = "\\textit{(Continued on Next Page...)}") |> 
-  footnote(number = c("Source: The 1950-2000 Census for synthetic parents, and 1994-2019 Current Population Surveys (CPS) for children's outcomes",
+  footnote(number = c("Source: The 1960-2000 Census for synthetic parents, and 1994-2019 Current Population Surveys (CPS) for children's outcomes",
                       "The data is restricted to native-born United States citizens who are also White and between the ages of 25 and 40. I identify the ethnicity of a person's parents through the parent's place of birth. A parent is Hispanic if they were born in a Spanish-speaking country. A parent is White if they were born in the United States."),
            footnote_as_chunk = F, title_format = c("italic"),
            escape = F, threeparttable = T

### 2023-05-11 my_paper/Hadah-last-names.tex
some edits

@@
 I compare the children of inter-ethnic marriages to study the impact of having a Hispanic last name.
-While males born to Hispanic father-White mothers earn less than those born to White father-Hispanic mothers, the gap could be completely explained by educational differences.
+While males born to Hispanic father-White mothers earn 6 percentage points less than those born to White father-Hispanic mothers, the gap could be completely explained by educational differences.
 I also study the effect of identifying as Hispanic on earnings.
-I find that men who identify as Hispanic earn significantly less than those who do not, even after controlling for educational differences.
+I find that men with a Spanish-sounding last name who identify as Hispanic earn significantly less than those who do not identify as Hispanic, but the gap could also be explained by educational differences.
 \end{abstract}
@@
 \pagebreak
-
-%\section{Questions}
-%
-%In this paper, I will answer the following questions:
-%
-%\begin{enumerate}
-%\item Do Hispanics face discrimination in the labor market due to an ethnic signal (last name)?
-%
-%\item Does the marriage market penalize or reward Hispanics for their ethnicity? 
-%
-%\item How do Hispanic men and women fare in inter-ethnic marriages? 
-%
-%\item Is there assortative matching in inter-ethnic marriages?
-%
-%\item Is there a difference in identity cost for those that identify as Hispanics and those that I induce are Hispanic through their parents' place of birth?
-%
-%\item What are the minority groups that faced a marriage market penalty? (Historical analysis)
-%
-%\item How did the marriage market penalty change over time for the different minority groups?
-%
-%\end{enumerate}
 \section{Introduction}\label{sec:intro}

### 2023-05-11 programs/18-table-six-regression.R

@@ -39,6 +39,10 @@ ParentDummies <- c(#"WH:Hispanic_ID",
                    "uhrsworkly")
 ParentControls  <- c("EducationQuint", "IncomeQuint")
 
+IndividualData |> 
+  filter(sex == 1 & FTFY == 1 & Self_employed == 0 & (HW == 1 | WH == 1)) |> 
+  group_by(Hispanic_ID, HW, WH) |>
+  summarise(mean(lninctot_1999, na.rm = TRUE))
 # By generation
 reg1 <- list(
   "\\specialcell{(1) \\\\ Log annual \\\\ earnings}" = feols(lninctot_1999 ~ HW*Hispanic_ID, vcov = ~statefip,

### 2023-05-11 my_paper/Hadah-last-names.tex
edit latex doc

@@
 
-The coefficients of interest from equation \ref{eq:iden} are $\beta_{1}$, $\beta_{2}$ and $\beta_{3}$. The coefficients $\beta_{1}$ and $\beta_{3}$ capture the effects of Spanish-sounding last names.
+The coefficients of interest from equation \ref{eq:iden} are $\beta_{1}$, $\beta_{2}$ and $\beta_{3}$.
+The coefficients $\beta_{1}$ and $\beta_{3}$ capture the effects of Spanish-sounding last names for $HW_{ist}$ that identify as Hispanic \textit{versus} those who do not.
 
@@
 Column one in table \ref{tab:identreg} is the results without controlling for education.
-Overall, an inter-ethnic that identifies as Hispanic earns 13 percentage points less than an inter-ethnic that does not.
-This gap, however, could be entirely explained by educational differences.
-An inter-ethnic that identifies as Hispanic---i.e.
+An inter-ethnic $HW_{ist}$ that identifies as Hispanic \textit{versus} $HW_{ist}$ who do not---i.e.
 $HW_{ist} + HW_{ist} \cdot Hispanic_i$---earns 6 percentage points less than an inter-ethnic that does not, but the result could be entirely explained by educational differences.

### 2023-07-12 programs/13a-table-three-summary-stats-mex.R
create sum tab for mex and nonmex hispanics

@@ -11,12 +11,6 @@ ParentData     <- read_csv(file.path(datasets,"CPS_synth.csv")) |>
 ParentData <- ParentData |> 
        filter(BirthPlaceMom %in% c("Mexico", "USA") & BirthPlaceDad %in% c("Mexico", "USA"))
 
-# Only Non-Mexicans
-`%notin%` <- Negate(`%in%`)
-ParentData <- ParentData |> 
-       filter(BirthPlaceMom %notin% c("Mexico") & BirthPlaceDad %notin% c("Mexico"))
-
-
 row1 <- c("Variables", 
           "\\specialcell{White \\\\ White \\\\ (WW) \\\\ (1)}",
           "\\specialcell{White \\\\ Hispanic \\\\ (WH) \\\\ (2)}",
@@ -365,7 +359,7 @@ knitr::kable(Table_rows, "html", align = "lcccccc",
              booktabs = T,
              escape = F,
        #       longtable = T, 
-             caption = "Summary statistics of synthetic parents by couple type \\label{tab:synth}") |>
+             caption = "Summary statistics of synthetic parents by couple type (Mexican Hispanics) \\label{tab:synthmex}") |>
   kable_classic(full_width = F) |>
   kable_styling(bootstrap_options = c("hover", "condensed"), 
                 latex_options = c("scale_down", 
@@ -385,7 +379,7 @@ knitr::kable(Table_rows, "latex", align = "lcccccc",
              booktabs = T,
              escape = F,
        #       longtable = T, 
-             caption = "Summary statistics of synthetic parents by couple type \\label{tab:synth}") |>
+             caption = "Summary statistics of synthetic parents by couple type (Mexican Hispanics) \\label{tab:synthmex}") |>
   kable_classic(full_width = F) |>
   kable_styling(bootstrap_options = c("hover", "condensed"), 
                 latex_options = c("scale_down", 
@@ -393,14 +387,14 @@ knitr::kable(Table_rows, "latex", align = "lcccccc",
               #   "repeat_header"
                 )) |> 
   footnote(number = c("Source: The 1960-2000 Census for synthetic parents, and 1994-2019 Current Population Surveys (CPS) for children's outcomes",
[…]

### 2023-07-12 programs/13b-table-three-summary-stats-nonmex.R

@@ -7,10 +7,6 @@
 ParentData     <- read_csv(file.path(datasets,"CPS_synth.csv")) |> 
   mutate(Tot_ed = HusbandEducation + WifeEducation)
 
-# Only Mexicans and Native Born
-ParentData <- ParentData |> 
-       filter(BirthPlaceMom %in% c("Mexico", "USA") & BirthPlaceDad %in% c("Mexico", "USA"))
-
 # Only Non-Mexicans
 `%notin%` <- Negate(`%in%`)
 ParentData <- ParentData |> 
@@ -365,7 +361,7 @@ knitr::kable(Table_rows, "html", align = "lcccccc",
              booktabs = T,
              escape = F,
        #       longtable = T, 
-             caption = "Summary statistics of synthetic parents by couple type \\label{tab:synth}") |>
+             caption = "Summary statistics of synthetic parents by couple type (Non-Mexican Hispanics) \\label{tab:syntnonmex}") |>
   kable_classic(full_width = F) |>
   kable_styling(bootstrap_options = c("hover", "condensed"), 
                 latex_options = c("scale_down", 
@@ -385,7 +381,7 @@ knitr::kable(Table_rows, "latex", align = "lcccccc",
              booktabs = T,
              escape = F,
        #       longtable = T, 
-             caption = "Summary statistics of synthetic parents by couple type \\label{tab:synth}") |>
+             caption = "Summary statistics of synthetic parents by couple type (Non-Mexican Hispanics) \\label{tab:syntnonmex}") |>
   kable_classic(full_width = F) |>
   kable_styling(bootstrap_options = c("hover", "condensed"), 
                 latex_options = c("scale_down", 
@@ -393,14 +389,14 @@ knitr::kable(Table_rows, "latex", align = "lcccccc",
               #   "repeat_header"
                 )) |> 
   footnote(number = c("Source: The 1960-2000 Census for synthetic parents, and 1994-2019 Current Population Surveys (CPS) for children's outcomes",
-                      "The data is restricted to native-born United States citizens who are also White, between the ages of 25 and 40, and have kids. I identify the ethnicity of a person's parents through the parent's place of birth. A parent is Hispanic if they were born in a Spanish-speaking country. A parent is White if they were born in the United States."),
[…]

### 2023-07-14 my_paper/Hadah-last-names.tex
edit pdf
create sum tab for mex and nonmex hispanics
edit

@@
 \selectfont{\Large{\MakeUppercase{The Impact of Hispanic Last Names and Identity on Labor Market Outcomes}}}}}}
-	\author{{\fontfamily{pbk}\selectfont\large{\textsc{Hussain Hadah}}\thanks{Department of Economics, University of Houston, 206A McElhinney Hall, 3623 Cullen Boulevard, Houston, TX 77204-5019, United States (e-mail: \email{hhadah@uh.edu}).}}}
+	\author{{\fontfamily{pbk}\selectfont\large{\textsc{Hussain Hadah}}\thanks{Department of Economics, Tulane University, 6823 St.
+Charles Ave., New Orleans, LA 70118, United States (e-mail: \email{hhadah@tulane.edu}).}}}
 	\date{\fontfamily{pbk}\selectfont\normalsize{\textsc{\today}}}
@@
 I find that men with a Spanish-sounding last name who identify as Hispanic earn significantly less than those who do not identify as Hispanic, but the gap could also be explained by educational differences.
+
+\keywords{Economics of Minorities, Race, and Immigrants; Discrimination and Prejudice.}
+\JEL{J71; J64}
+
 \end{abstract}
@@
 $\beta_{1}$ is the coefficient of interest in this specification.
-$\beta_{1}$ estimates the earnings gaps between children of inter-ethnic marriages who have a Spanish-sounding last name versus a White last name.
+$\beta_{1}$ represents the earnings gaps between children of inter-ethnic marriages who have a Spanish-sounding last name versus a White last name.
 
@@
 
+\input{tables/tab03a-summary-stats-mex}
+
+\newpage
+
+\input{tables/tab03a-summary-stats-nonmex}
+
+\newpage
+
 \input{tables/tab04-summary-stats}

### 2023-07-22 manuscripts/Hussain-Hadah/Hussain-Hadah.tex
add elsavier manuscript

--- 2023-07-14 my_paper/Hadah-last-names.tex
@@
-\documentclass{wptemp}
-\usepackage{adjustbox}
-\usepackage{dcolumn}    % aligning decimals
-    \newcolumntype{d}[1]{D{.}{.}{#1}}
-% *****************************************************************
-% Estout LaTeX wrapper
-% *****************************************************************
+%%
+%% Copyright 2019-2021 Elsevier Ltd
+%%
+%% This file is part of the 'CAS Bundle'.
+%% --------------------------------------
+%%
+%% It may be distributed under the conditions of the LaTeX Project Public
+%% License, either version 1.2 of this license or (at your option) any
+%% later version.  The latest version of this license is in
+%%    http://www.latex-project.org/lppl.txt
+%% and version 1.2 or later is part of all distributions of LaTeX
+%% version 1999/12/01 or later.
+%%
+%% The list of all files belonging to the 'CAS Bundle' is
+%% given in the file `manifest.txt'.
+%%
+%% Template article for cas-sc documentclass for
+%% single column output.
+
+\documentclass[a4paper,fleqn]{cas-sc}
+
+% If the frontmatter runs over more than one page
+% use the longmktitle option.
+
+%\documentclass[a4paper,fleqn,longmktitle]{cas-sc}
+
+%\usepackage[numbers]{natbib}
+\usepackage[authoryear]{natbib}
+%\usepackage[authoryear,longnamesfirst]{natbib}
+
+%%%Author macros
+\def\tsc#1{\csdef{#1}{\textsc{\lowercase{#1}}\xspace}}
+\tsc{WGM}
+\tsc{QE}
+%%%
+
+% Uncomment and use as if needed
+%\newtheorem{theorem}{Theorem}
+%\newtheorem{lemma}[theorem]{Lemma}
+%\newdefinition{rmk}{Remark}
+%\newproof{pf}{Proof}
+%\newproof{pot}{Proof of Theorem \ref{thm}}
 
@@
 \usepackage{longtable}
-
-%%Original code developed by Jörg Weber: see
-%% https://www.jwe.cc/2012/03/stata-latex-tables-estout/
-%% and
-%% https://www.jwe.cc/blog/
-
-\let\estinput=\input % define a new input command so that we can still flatten the document
-
-\newcommand{\estwide}[3]{
-		\vspace{.75ex}{
-			%\textsymbols% Note the added command here
-			\begin{tabular*}
-			{\textwidth}{@{\hskip\tabcolsep\extracolsep\fill}l*{#2}{#3}}
-			\toprule
-			\estinput{#1}
-			\bottomrule
-			\addlinespace[.75ex]
-			\end{tabular*}
-			}
-		}
-
-\newcommand{\estauto}[3]{
-		\vspace{.75ex}{
-			%\textsymbols% Note the added command here
-			\begin{tabular}{l*{#2}{#3}}
-			\toprule
-			\estinput{#1}
-			\bottomrule
-			\addlinespace[.75ex]
-			\end{tabular}
-			}
-		}
-
 % Allow line breaks with \\ in specialcells
@@
 
-\newcommand{\sym}[1]{\rlap{#1}}% Thanks David Carlisle
+\usepackage{dcolumn}    % aligning decimals
+    \newcolumntype{d}[1]{D{.}{.}{#1}}
 
-%%%%%%%%%%% End of wrapper %%%%%%%%%%%%%%%%%%%%%
-
+\usepackage{siunitx}
 %%%%%%%%%%% TiKz %%%%%%%%%%%%%%%%%%%%%
@@
 \tikzstyle{arrow} = [thick,->,>=stealth]
-
 \begin{document}
 
-\linespread{1}{\title{{\fontfamily{ptm}
-\selectfont{\Large{\MakeUppercase{The Impact of Hispanic Last Names and Identity on Labor Market Outcomes}}}}}}
-	\author{{\fontfamily{pbk}\selectfont\large{\textsc{Hussain Hadah}}\thanks{Department of Economics, Tulane University, 6823 St.
-Charles Ave., New Orleans, LA 70118, United States (e-mail: \email{hhadah@tulane.edu}).}}}
-	\date{\fontfamily{pbk}\selectfont\normalsize{\textsc{\today}}}
-	\maketitle
+\let\WriteBookmarks\relax
+\def\floatpagepagefraction{1}
+\def\textpagefraction{.001}
 
-\begin{center}
-\href{https://hhadah.github.io/hispanic-last-names/my_paper/Hadah-last-names.pdf}{\textcolor{green!15!black!30!blue}{\footnotesize{\textsc{Click Here to Get the Most Updated Version}}}}
-\end{center}
+% Short title
+\shorttitle{The impact of Hispanic last names and identity on labor market outcomes}
 
-\pagenumbering{arabic}
+% Short author
+\shortauthors{H. Hadah}
+
+% Main title of the paper
+\title [mode = title]{The impact of Hispanic last names and identity on labor market outcomes}
+
+\author[]{Hussain Hadah}[orcid=https://orcid.org/0000-0002-8705-6386]
+\ead{hhadah@tulane.edu}
+
+\affiliation[]{organization={Department of Economics, Tulane University},
+            addressline={6823 St. Charles Ave.},
+            city={New Orleans},
+%          citysep={}, % Uncomment if no comma needed between city and postcode
+            postcode={70118},
+            state={LA},
+            country={United States}}
 
@@
 I find that men with a Spanish-sounding last name who identify as Hispanic earn significantly less than those who do not identify as Hispanic, but the gap could also be explained by educational differences.
-
-\keywords{Economics of Minorities, Race, and Immigrants; Discrimination and Prejudice.}
-\JEL{J71; J64}
-
 \end{abstract}
 
-\pagebreak
+\begin{keywords}
+ Economics of Minorities, Race, and Immigrants\sep Discrimination and Prejudice\\
+ \JEL{J71\sep J64}
+\end{keywords}
+
+\maketitle
+
 \section{Introduction}\label{sec:intro}
@@
 
+\begin{figure}[t]
+\centering
+\includegraphics[scale=0.25]{Fig/interracialovertime.png}
+\caption{Inter-racial marriages as a percent of all marriagesin the United States from 1995 to 2020.}
+\label{fig:3}
+\end{figure}
+
 This study utilizes a method developed by \citet{rubinstein2014pride}.
@@
 
-% Inter-ethnic children with a White father, and thus having a non-Hispanic last name, earn 14\% less than children of endogamous White parents.
-Inter-ethnic children with a Hispanic father, and thus having a Hispanic last name, earn 20\% less than children of endogamous White parents.
-Endogamous children with Hispanic parents earn 24\% less than children of endogamous White parents.
-A big part of the gaps could be explained by education.
-Inter-ethnic children with a White father and identify as Hispanic, and thus having a non-Hispanic last name, earn 14\% less than children of endogamous White parents.
-Inter-ethnic children with a Hispanic father and identify as Hispanic, and thus having a Hispanic last name, earn 18\% less than children of endogamous White parents.
-Endogamous children with Hispanic parents and identify as Hispanic earn 10\% less than children of endogamous White parents.
-
 Previous studies have also used names as a proxy for race and ethnicity \citep{fryer2004causes, rubinstein2014pride, bertrand2004emily}.
@@
 Section \ref{sec:emp_model} introduces the empirical framework.
-I report the results in section \ref{sec:results}.
-Finally, I offer a brief conclusion in section \ref{sec:con1}.
+I report the results in Section \ref{sec:results}.
+Finally, I offer a brief conclusion in Section \ref{sec:con1}.
 
@@
 
-The distribution of the four types of children is presented in figure \ref{fig:dist} and table \ref{tab:mat1}.
+\begin{figure}[t]
+\centering
+\includegraphics[scale=0.4]{Fig/PirChart2.png}
+\caption{Distribution of the four types of children.}
+\label{fig:dist}
+\end{figure}
+
+\begin{figure}[t]
+\centering
+\begin{tikzpicture}[node distance =2cm]
+\node (start) [startstop] {Synthetic Parents: Observed in 1960 to 2000. $25 \leq age \leq 40$, White, with kids};
+
+\node (pro1) [process, below of = start, yshift = -2cm] {Children: Observed between 1994 and 2019. Born between 1954 to 1999 and $25 \leq age \leq 40$};
+
+\draw [arrow] (start) -- (pro1);
+\end{tikzpicture}
+\caption{Chart explaining which synthetic parents and children and when we observe them.}
+\label{flowchart1}
+\end{figure}
+
+The distribution of the four types of children is presented in Fig.
+\ref{fig:dist} and Table \ref{tab:mat1}.
 The majority of the sample (95.96\%) is WW children.
@@
 Even though WH and HW are only 1.35\% of the sample, I have plenty of observations to carry out an analysis.
-The summary statistics for the children of the four types of marriages are presented in table \ref{tab:c&p1}.
+The summary statistics for the children of the four types of marriages are presented in Table \ref{tab:c&p1}.
 Children of WW marriages (column 1) do better on every measure while children of HH parents (column 4) do worse than other children on every measure.
+
+%table1
+\input{tables/tab01-observations-by-parents}
+%table2
+\input{tables/tab02-observations-by-couple}
 
@@
 At the same time, the WH children have White fathers with a non-Hispanic, or White, last names.
-Summary statistics of these two groups, table \ref{tab:c&p1} columns (2) and (3) can show that they are different from the children of non-inter-ethnic marriages.
+Summary statistics of these two groups, Table \ref{tab:c&p1} columns (2) and (3) can show that they are different from the children of non-inter-ethnic marriages.
+
+%table3
+\input{tables/tab03b-summary-stats}
 
@@
 
-In panel C of table \ref{tab:c&p1}, I present the rates at which members of each group self-identify as Hispanic.
+In panel C of Table \ref{tab:c&p1}, I present the rates at which members of each group self-identify as Hispanic.
 \footnote{Hispanic self-identification is observed when an individual notes that they are Hispanic on the CPS questionnaire} The vast majority of HH children, 96\% for men and 97\% for women, identify as Hispanic.
@@
 
-In table \ref{tab:c&p2}, I present the summary statistics for the four types and limit the sample to those that only identify as Hispanic.\footnote{Limiting the sample to those who identify as Hispanic only affects WH, HW, and HH.
+In Table \ref{tab:c&p2}, I present the summary statistics for the four types and limit the sample to those that only identify as Hispanic.\footnote{Limiting the sample to those who identify as Hispanic only affects WH, HW, and HH.
 Members of the WW group are not affected by this restriction.} The same pattern discussed above between HH--WW and HW--WH persists.
 The one difference is that those who identify as Hispanic do slightly worse.
+
+%table4
+\input{tables/tab03a-summary-stats}
 
@@
 
-I show the distribution of the four types of couples in table \ref{tab:mat2}.
+%table5
+\input{tables/tab04-summary-stats}
+
+I show the distribution of the four types of couples in Table \ref{tab:mat2}.
 White husbands and White wives (WW) make up the majority of couples in the sample, 96\% (5,141,737 couples).
@@
 
-I present the summary statistics of the four types of synthetic parents in table \ref{tab:synth}.
+I present the summary statistics of the four types of synthetic parents in Table \ref{tab:synth}.
 WW couples have higher education, 12.75 years for husbands and 12.47 for wives.
@@
 
-In columns 5 and 6 of table \ref{tab:synth}, I report the differences in means between HH-WW households and HW-WH households.
+In columns 5 and 6 of Table \ref{tab:synth}, I report the differences in means between HH-WW households and HW-WH households.
 An HH Hispanic husband attained 4.11 fewer years of education than a WW husband.
@@
 These differences show that couples that intermarry are better and different than those that do not.
-
-%In table \ref{tab:p1} columns five and six, I provide the  HH-WW and HW-WH gaps.
-WW couples are more educated and earn more than HH couples.
-Also, WH husbands are more educated and earn more than HW husbands.
-In comparison, WH wives are more educated than HW wives.
-Even though White husbands in intermarriages are more educated and earn more than husbands in HW marriages, they are more comparable to each other than the non-intermarriage couples.
 
@@
 
+%table6
+\input{tables/tab05-regression}
 \subsection{Estimating the effect of having a Hispanic last name}
@@
 
-The coefficients of interest from equation \ref{eq:iden} are $\beta_{1}$, $\beta_{2}$ and $\beta_{3}$.
+The coefficients of interest from Eq.
+\eqref{eq:iden} are $\beta_{1}$, $\beta_{2}$ and $\beta_{3}$.
 The coefficients $\beta_{1}$ and $\beta_{3}$ capture the effects of Spanish-sounding last names for $HW_{ist}$ that identify as Hispanic \textit{versus} those who do not.
-
+%table7
+\input{tables/tab06-regression}
 \section{Results}\label{sec:results}
 
-In this section, I present the results from estimating the two specifications in equations \ref{eq:1a} and \ref{eq:iden}.
-The results are shown in tables \ref{tab:lastnamereg} and \ref{tab:identreg}.
+In this section, I present the results from estimating the two specifications in Eqs.
+\eqref{eq:1a} and \eqref{eq:iden}.
+The results are shown in Tables \ref{tab:lastnamereg} and \ref{tab:identreg}.
 All estimations are done on a sample of White, native-born men between the ages of 25 and 40 employed for full-time full-year (FTFY) and are wage and salary workers.
@@
 
-I provide the results to the estimation of equation \ref{eq:1a} in table \ref{tab:lastnamereg}.
+I provide the results to the estimation of Eq.
+\eqref{eq:1a} in Table \ref{tab:lastnamereg}.
 The sample in this analysis includes full-time, full-year, and wage and salary men, and I control for hours worked and age and include year fixed effects (FE).
 I also restrict the sample to children of HW and WH (omitted) parents.
-Column one in table \ref{tab:lastnamereg} is the results without controlling for education.
+Column one in Table \ref{tab:lastnamereg} is the results without controlling for education.
 Overall, an inter-ethnic with a Hispanic last name earns 6 percentage points less than an inter-ethnic with a White last name.
@@
 
-I provide the results to the estimation of equation \ref{eq:iden} in table \ref{tab:identreg}.
+I provide the results to the estimation of Eq.
+\eqref{eq:iden} in Table \ref{tab:identreg}.
 The sample in this analysis includes full-time, full-year, and wage and salary men, and I control for hours worked and age and include year fixed effects (FE).
@@
 The omitted group is children of WH parents that do not identify as Hispanic.
-Column one in table \ref{tab:identreg} is the results without controlling for education.
+Column one in Table \ref{tab:identreg} is the results without controlling for education.
 An inter-ethnic $HW_{ist}$ that identifies as Hispanic \textit{versus} $HW_{ist}$ who do not---i.e.
@@
 
-% \nocite{*}
-\pagebreak
-\bibliography{HispanicDiscriminationMarriageMarket}
-\newpage
-\pagebreak
-\begin{appendices}
-
-\section{Figures}\label{appendix:figures}
-
-\begin{center}
-\begin{figure}[h!]
-\caption{Inter-racial marriages as a percent of all marriages \\
- in the United States from 1995 to 2020.}
-\includegraphics[width=\textwidth]{interracialovertime.png} 
-\label{fig:3}
-\end{figure}
-\end{center}
-
-\newpage
-
-\begin{center}
-\begin{figure}[h!]
-\caption{Distribution of the four types of children.}
-\includegraphics[width=\textwidth]{PirChart2.png} 
-\label{fig:dist}
-\end{figure}
-\end{center}
-
-\newpage
-
-\begin{figure}[h!]
-\centering
-\caption{Chart explaining which synthetic parents and children and when we observe them.}
-\begin{tikzpicture}[node distance =2cm]
-\node (start) [startstop] {Synthetic Parents: Observed in 1960 to 2000. $25 \leq age \leq 40$, White, with kids};
-
-\node (pro1) [process, below of = start, yshift = -2cm] {Children: Observed between 1994 and 2019. Born between 1954 to 1999 and $25 \leq age \leq 40$};
-
-\draw [arrow] (start) -- (pro1);
-\end{tikzpicture}
-\label{flowchart1}
-\end{figure}
-
-\newpage
-
-\section{tables}\label{appendix:tables}
-
-\input{tables/tab01-observations-by-parents}
-
-\newpage
-
-\input{tables/tab02-observations-by-couple}
-
-\newpage
-
-\input{tables/tab03b-summary-stats}
-
-\newpage
-
-\input{tables/tab03a-summary-stats}
-
-\newpage
-
-\input{tables/tab03a-summary-stats-mex}
-
-\newpage
-
-\input{tables/tab03a-summary-stats-nonmex}
-
-\newpage
-
-\input{tables/tab04-summary-stats}
-
-\newpage
-
-\input{tables/tab05-regression}
-
-\newpage
-
-\input{tables/tab06-regression}
-
-\end{appendices}
+\bibliographystyle{cas-model2-names}
+\bibliography{bibliography}
 

### 2023-07-22 manuscripts/Hussain-Hadah/tables/tab02-observations-by-couple.tex
add elsavier manuscript

--- 2023-04-16 output/tables/tab02-observations-by-couple.tex
@@
-\begin{table}[H]
-
+\begin{table}[t]
 \caption{Couples' Type \label{tab:mat2}}
-\centering
-\resizebox{\linewidth}{!}{
+%\resizebox{\linewidth}{!}{
 \begin{threeparttable}
@@
 \midrule
-\textbf{\specialcell{Observations\\Share}} & \specialcell{1,286,731\\0.97} & \specialcell{7,178\\0.01} & \specialcell{7,606\\0.01} & \specialcell{20,911\\0.02}\\
+\textbf{\specialcell{Observations\\Share}} & \specialcell{5,141,737\\0.96} & \specialcell{33,097\\0.01} & \specialcell{37,847\\0.01} & \specialcell{119,749\\0.02}\\
 \bottomrule
@@
 \begin{tablenotes}
-\item[1] Source: 1970-1990 Census
-\item[2] The sample includes Whites, who are married, and are between the ages 25 and 40.
+\item[1] Source: 1960-2000 Census
+\item[2] The sample includes Whites, who are married, with kids, and are between the ages 25 and 40.
 Ethnicity of a person's parents are identified by the parent's place of birth.
 A parent is Hispanic if she/he was born in a Spanish-speaking country.
-A parent is White she/he was born in the United States.
+A parent is White if she/he was born in the United States.
 \end{tablenotes}
-\end{threeparttable}}
+\end{threeparttable}
+%}
 \end{table}

### 2023-07-22 manuscripts/Hussain-Hadah/tables/tab03a-summary-stats-mex.tex
add elsavier manuscript

--- 2023-05-09 output/tables/tab03a-summary-stats.tex
@@
 
-\caption{Summary statistics of synthetic parents by couple type \label{tab:synth}}
+\caption{Summary statistics of synthetic parents by couple type (Mexican Hispanics) \label{tab:synthmex}}
 \centering
@@
 \midrule
-Husband's education (Total Years) & \specialcell{12.75\\(0.61)} & \specialcell{11.77\\(1.87)} & \specialcell{10.25\\(2.44)} & \specialcell{8.64\\(2.01)} & \specialcell{-4.11***\\(0.00)} & \specialcell{-1.52***\\(0.00)}\\
-Wife's education (Total Years) & \specialcell{12.47\\(0.56)} & \specialcell{10.40\\(2.23)} & \specialcell{11.11\\(1.75)} & \specialcell{8.49\\(1.92)} & \specialcell{-3.98***\\(0.00)} & \specialcell{0.70***\\(0.00)}\\
-Total Household seducation (Total Years) & \specialcell{25.22\\(1.17)} & \specialcell{22.17\\(4.05)} & \specialcell{21.36\\(4.10)} & \specialcell{17.13\\(3.88)} & \specialcell{-8.09***\\(0.00)} & \specialcell{-0.81***\\(0.01)}\\
-Log Total Family Income & \specialcell{10.72\\(0.09)} & \specialcell{10.55\\(0.26)} & \specialcell{10.46\\(0.27)} & \specialcell{10.27\\(0.21)} & \specialcell{-0.45***\\(0.00)} & \specialcell{-0.10***\\(0.00)}\\
-Fertility & \specialcell{3.77\\(0.40)} & \specialcell{3.98\\(0.71)} & \specialcell{4.15\\(0.79)} & \specialcell{4.23\\(0.64)} & \specialcell{0.46***\\(0.00)} & \specialcell{0.17***\\(0.00)}\\
+Husband's education (Total Years) & \specialcell{12.75\\(0.61)} & \specialcell{10.80\\(1.41)} & \specialcell{9.14\\(1.52)} & \specialcell{7.85\\(1.24)} & \specialcell{-4.90***\\(0.00)} & \specialcell{-1.65***\\(0.01)}\\
+Wife's education (Total Years) & \specialcell{12.47\\(0.56)} & \specialcell{9.27\\(1.71)} & \specialcell{10.46\\(1.32)} & \specialcell{7.80\\(1.36)} & \specialcell{-4.67***\\(0.00)} & \specialcell{1.19***\\(0.00)}\\
+Total Household education (Total Years) & \specialcell{25.22\\(1.17)} & \specialcell{20.06\\(3.09)} & \specialcell{19.60\\(2.80)} & \specialcell{15.65\\(2.60)} & \specialcell{-9.57***\\(0.00)} & \specialcell{-0.47**\\(0.01)}\\
+Log Total Family Income & \specialcell{10.72\\(0.09)} & \specialcell{10.42\\(0.14)} & \specialcell{10.33\\(0.12)} & \specialcell{10.20\\(0.07)} & \specialcell{-0.52***\\(0.00)} & \specialcell{-0.09***\\(0.00)}\\
+Fertility & \specialcell{3.77\\(0.40)} & \specialcell{4.26\\(0.64)} & \specialcell{4.33\\(0.64)} & \specialcell{4.46\\(0.49)} & \specialcell{0.69***\\(0.00)} & \specialcell{0.07***\\(0.00)}\\
 \bottomrule
@@
 \begin{tablenotes}
-\item[1] Source: The 1950-2000 Census for synthetic parents, and 1994-2019 Current Population Surveys (CPS) for children's outcomes
+\item[1] Source: The 1960-2000 Census for synthetic parents, and 1994-2019 Current Population Surveys (CPS) for children's outcomes
 \item[2] The data is restricted to native-born United States citizens who are also White, between the ages of 25 and 40, and have kids.
 I identify the ethnicity of a person's parents through the parent's place of birth.
-A parent is Hispanic if they were born in a Spanish-speaking country.
+A parent is Hispanic if they were born in a Mexico.
 A parent is White if they were born in the United States.

### 2023-07-22 manuscripts/Hussain-Hadah/tables/tab03a-summary-stats-nonmex.tex
add elsavier manuscript

--- 2023-07-22 manuscripts/Hussain-Hadah/tables/tab03a-summary-stats-mex.tex
@@
 
-\caption{Summary statistics of synthetic parents by couple type (Mexican Hispanics) \label{tab:synthmex}}
+\caption{Summary statistics of synthetic parents by couple type (Non-Mexican Hispanics) \label{tab:syntnonmex}}
 \centering
@@
 \midrule
-Husband's education (Total Years) & \specialcell{12.75\\(0.61)} & \specialcell{10.80\\(1.41)} & \specialcell{9.14\\(1.52)} & \specialcell{7.85\\(1.24)} & \specialcell{-4.90***\\(0.00)} & \specialcell{-1.65***\\(0.01)}\\
-Wife's education (Total Years) & \specialcell{12.47\\(0.56)} & \specialcell{9.27\\(1.71)} & \specialcell{10.46\\(1.32)} & \specialcell{7.80\\(1.36)} & \specialcell{-4.67***\\(0.00)} & \specialcell{1.19***\\(0.00)}\\
-Total Household education (Total Years) & \specialcell{25.22\\(1.17)} & \specialcell{20.06\\(3.09)} & \specialcell{19.60\\(2.80)} & \specialcell{15.65\\(2.60)} & \specialcell{-9.57***\\(0.00)} & \specialcell{-0.47**\\(0.01)}\\
-Log Total Family Income & \specialcell{10.72\\(0.09)} & \specialcell{10.42\\(0.14)} & \specialcell{10.33\\(0.12)} & \specialcell{10.20\\(0.07)} & \specialcell{-0.52***\\(0.00)} & \specialcell{-0.09***\\(0.00)}\\
-Fertility & \specialcell{3.77\\(0.40)} & \specialcell{4.26\\(0.64)} & \specialcell{4.33\\(0.64)} & \specialcell{4.46\\(0.49)} & \specialcell{0.69***\\(0.00)} & \specialcell{0.07***\\(0.00)}\\
+Husband's education (Total Years) & \specialcell{12.75\\(0.61)} & \specialcell{13.35\\(1.14)} & \specialcell{12.63\\(1.81)} & \specialcell{10.41\\(1.88)} & \specialcell{-2.34***\\(0.00)} & \specialcell{-0.72***\\(0.01)}\\
+Wife's education (Total Years) & \specialcell{12.47\\(0.56)} & \specialcell{12.46\\(1.22)} & \specialcell{12.46\\(1.49)} & \specialcell{10.18\\(1.74)} & \specialcell{-2.29***\\(0.00)} & \specialcell{0.00***\\(0.01)}\\
+Total Household education (Total Years) & \specialcell{25.22\\(1.17)} & \specialcell{25.80\\(2.18)} & \specialcell{25.08\\(3.12)} & \specialcell{20.59\\(3.50)} & \specialcell{-4.63***\\(0.00)} & \specialcell{-0.72**\\(0.01)}\\
+Log Total Family Income & \specialcell{10.72\\(0.09)} & \specialcell{10.76\\(0.24)} & \specialcell{10.69\\(0.28)} & \specialcell{10.36\\(0.27)} & \specialcell{-0.36***\\(0.00)} & \specialcell{-0.07***\\(0.00)}\\
+Fertility & \specialcell{3.77\\(0.40)} & \specialcell{3.52\\(0.49)} & \specialcell{3.71\\(0.80)} & \specialcell{3.85\\(0.61)} & \specialcell{0.09***\\(0.00)} & \specialcell{0.19***\\(0.00)}\\
 \bottomrule
@@
 I identify the ethnicity of a person's parents through the parent's place of birth.
-A parent is Hispanic if they were born in a Mexico.
+A parent is Hispanic if they were born in a Spanish-speaking country other than Mexico.
 A parent is White if they were born in the United States.

### 2023-07-22 manuscripts/Hussain-Hadah/tables/tab03a-summary-stats.tex
add elsavier manuscript

--- 2023-05-09 output/tables/tab03a-summary-stats.tex
@@
-\begin{table}[H]
-
+\begin{table}[t]
 \caption{Summary statistics of synthetic parents by couple type \label{tab:synth}}
-\centering
-\resizebox{\linewidth}{!}{
+%\resizebox{\linewidth}{!}{
 \begin{threeparttable}
@@
 Wife's education (Total Years) & \specialcell{12.47\\(0.56)} & \specialcell{10.40\\(2.23)} & \specialcell{11.11\\(1.75)} & \specialcell{8.49\\(1.92)} & \specialcell{-3.98***\\(0.00)} & \specialcell{0.70***\\(0.00)}\\
-Total Household seducation (Total Years) & \specialcell{25.22\\(1.17)} & \specialcell{22.17\\(4.05)} & \specialcell{21.36\\(4.10)} & \specialcell{17.13\\(3.88)} & \specialcell{-8.09***\\(0.00)} & \specialcell{-0.81***\\(0.01)}\\
+Total Household education (Total Years) & \specialcell{25.22\\(1.17)} & \specialcell{22.17\\(4.05)} & \specialcell{21.36\\(4.10)} & \specialcell{17.13\\(3.88)} & \specialcell{-8.09***\\(0.00)} & \specialcell{-0.81***\\(0.01)}\\
 Log Total Family Income & \specialcell{10.72\\(0.09)} & \specialcell{10.55\\(0.26)} & \specialcell{10.46\\(0.27)} & \specialcell{10.27\\(0.21)} & \specialcell{-0.45***\\(0.00)} & \specialcell{-0.10***\\(0.00)}\\
@@
 \begin{tablenotes}
-\item[1] Source: The 1950-2000 Census for synthetic parents, and 1994-2019 Current Population Surveys (CPS) for children's outcomes
+\item[1] Source: The 1960-2000 Census for synthetic parents, and 1994-2019 Current Population Surveys (CPS) for children's outcomes
 \item[2] The data is restricted to native-born United States citizens who are also White, between the ages of 25 and 40, and have kids.
@@
 \end{tablenotes}
-\end{threeparttable}}
+\end{threeparttable}
+%}
 \end{table}

### 2023-07-22 manuscripts/Hussain-Hadah/tables/tab03b-summary-stats.tex
add elsavier manuscript

--- 2023-05-09 output/tables/tab03b-summary-stats.tex
@@
-\begin{landscape}
+%\begin{landscape}
 \begin{ThreePartTable}
-\begin{TableNotes}
-\item[1] Source: The 1950-2000 Census for synthetic parents, and 1994-2019 Current Population Surveys (CPS) for children's outcomes
-\item[2] The data is restricted to native-born United States citizens who are also White and between the ages of 25 and 40.
-I identify the ethnicity of a person's parents through the parent's place of birth.
-A parent is Hispanic if they were born in a Spanish-speaking country.
-A parent is White if they were born in the United States.
-\end{TableNotes}
 \begin{longtable}[t]{>{\raggedright\arraybackslash}p{5cm}cccccc}
@@
 \bottomrule
-\insertTableNotes
+\multicolumn{7}{l}{\parbox{15cm}{$^1$Source: The 1960-2000 Census for synthetic parents, and 1994-2019 Current Population Surveys (CPS) for children's outcomes}}\\
+\multicolumn{7}{l}{\parbox{15cm}{$^2$The data is restricted to native-born United States citizens who are also White and between the ages of 25 and 40.
+I identify the ethnicity of a person's parents through the parent's place of birth.
+A parent is Hispanic if they were born in a Spanish-speaking country.
+A parent is White if they were born in the United States.}}\\
 \endlastfoot
 \textbf{Panel A: Children's Education} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
-\hspace{1em}Men’s education (Total Years) & \specialcell{13.82\\(2.42)} & \specialcell{13.58\\(2.40)} & \specialcell{13.22\\(2.34)} & \specialcell{12.90\\(2.31)} & \specialcell{-0.92***\\(0.01)} & \specialcell{-0.36**\\(0.02)}\\
-\hspace{1em}Women’s education (Total Years) & \specialcell{14.06\\(2.37)} & \specialcell{13.79\\(2.44)} & \specialcell{13.42\\(2.38)} & \specialcell{13.24\\(2.39)} & \specialcell{-0.82***\\(0.01)} & \specialcell{-0.37**\\(0.02)}\\
+\hspace{1em}Men's education (Total Years) & \specialcell{13.82\\(2.42)} & \specialcell{13.58\\(2.40)} & \specialcell{13.22\\(2.34)} & \specialcell{12.90\\(2.31)} & \specialcell{-0.92***\\(0.01)} & \specialcell{-0.36**\\(0.02)}\\
+\hspace{1em}Women's education (Total Years) & \specialcell{14.06\\(2.37)} & \specialcell{13.79\\(2.44)} & \specialcell{13.42\\(2.38)} & \specialcell{13.24\\(2.39)} & \specialcell{-0.82***\\(0.01)} & \specialcell{-0.37**\\(0.02)}\\
 \textbf{Panel B: Children's Employment and Earnings} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
-\hspace{1em}Men’s Unemployment Rate & \specialcell{0.04\\(0.80)} & \specialcell{0.05\\(0.77)} & \specialcell{0.07\\(0.75)} & \specialcell{0.07\\(0.75)} & \specialcell{0.02***\\(0.00)} & \specialcell{0.01***\\(0.00)}\\
+\hspace{1em}Men's Unemployment Rate & \specialcell{0.04\\(0.80)} & \specialcell{0.05\\(0.77)} & \specialcell{0.07\\(0.75)} & \specialcell{0.07\\(0.75)} & \specialcell{0.02***\\(0.00)} & \specialcell{0.01***\\(0.00)}\\
 \addlinespace
-\hspace{1em}Women’s Unemployment Rate & \specialcell{0.04\\(0.81)} & \specialcell{0.05\\(0.22)} & \specialcell{0.06\\(0.76)} & \specialcell{0.06\\(0.76)} & \specialcell{0.02***\\(0.00)} & \specialcell{0.01***\\(0.00)}\\
-\hspace{1em}Men’s Log Hourly Earnings & \specialcell{2.51\\(0.45)} & \specialcell{2.44\\(0.47)} & \specialcell{2.43\\(0.45)} & \specialcell{2.42\\(0.43)} & \specialcell{-0.09***\\(0.00)} & \specialcell{-0.01**\\(0.01)}\\
-\hspace{1em}Women’s Log Hourly Earnings & \specialcell{2.32\\(0.49)} & \specialcell{2.32\\(0.46)} & \specialcell{2.29\\(0.46)} & \specialcell{2.31\\(0.42)} & \specialcell{-0.02***\\(0.00)} & \specialcell{-0.03**\\(0.01)}\\
-\hspace{1em}Men’s Log Annual Earnings & \specialcell{10.29\\(1.01)} & \specialcell{10.12\\(1.05)} & \specialcell{10.08\\(1.01)} & \specialcell{10.01\\(1.04)} & \specialcell{-0.28***\\(0.01)} & \specialcell{-0.04**\\(0.03)}\\
-\hspace{1em}Women’s Log Annual Earnings & \specialcell{9.43\\(1.77)} & \specialcell{9.51\\(1.63)} & \specialcell{9.46\\(1.59)} & \specialcell{9.53\\(1.50)} & \specialcell{0.10**\\(0.02)} & \specialcell{-0.05**\\(0.05)}\\
+\hspace{1em}Women's Unemployment Rate & \specialcell{0.04\\(0.81)} & \specialcell{0.05\\(0.22)} & \specialcell{0.06\\(0.76)} & \specialcell{0.06\\(0.76)} & \specialcell{0.02***\\(0.00)} & \specialcell{0.01***\\(0.00)}\\
+\hspace{1em}Men's Log Hourly Earnings & \specialcell{2.51\\(0.45)} & \specialcell{2.44\\(0.47)} & \specialcell{2.43\\(0.45)} & \specialcell{2.42\\(0.43)} & \specialcell{-0.09***\\(0.00)} & \specialcell{-0.01**\\(0.01)}\\
+\hspace{1em}Women's Log Hourly Earnings & \specialcell{2.32\\(0.49)} & \specialcell{2.32\\(0.46)} & \specialcell{2.29\\(0.46)} & \specialcell{2.31\\(0.42)} & \specialcell{-0.02***\\(0.00)} & \specialcell{-0.03**\\(0.01)}\\
+\hspace{1em}Men's Log Annual Earnings & \specialcell{10.29\\(1.01)} & \specialcell{10.12\\(1.05)} & \specialcell{10.08\\(1.01)} & \specialcell{10.01\\(1.04)} & \specialcell{-0.28***\\(0.01)} & \specialcell{-0.04**\\(0.03)}\\
+\hspace{1em}Women's Log Annual Earnings & \specialcell{9.43\\(1.77)} & \specialcell{9.51\\(1.63)} & \specialcell{9.46\\(1.59)} & \specialcell{9.53\\(1.50)} & \specialcell{0.10**\\(0.02)} & \specialcell{-0.05**\\(0.05)}\\
 \addlinespace
@@
 \end{ThreePartTable}
-\end{landscape}
+%\end{landscape}

### 2023-07-22 manuscripts/Hussain-Hadah/tables/tab04-summary-stats.tex
add elsavier manuscript

--- 2023-05-09 my_paper/tables/tab04-summary-stats.tex
@@
-\begin{landscape}
+%\begin{landscape}
 \begin{ThreePartTable}
-\begin{TableNotes}
-\item[1] Source: The 1950-2000 Census for synthetic parents, and 1994-2019 Current Population Surveys (CPS) for children's outcomes
-\item[2] The data is restricted to native-born United States citizens between 1994 and 2019 who are also White and between the ages of 25 and 40.
-I identify the ethnicity of a person's parents through the parent's place of birth.
-A parent is Hispanic if they were born in a Spanish-speaking country.
-A parent is White if they were born in the United States.
-\end{TableNotes}
 \begin{longtable}[t]{>{\raggedright\arraybackslash}p{5cm}cccccc}
@@
 \bottomrule
-\insertTableNotes
+\multicolumn{7}{l}{\parbox{15cm}{$^1$ Source: The 1960-2000 Census for synthetic parents, and 1994-2019 Current Population Surveys (CPS) for children's outcomes}}\\
+\multicolumn{7}{l}{\parbox{15cm}{$^2$ The data is restricted to native-born United States citizens between 1994 and 2019 who are also White and between the ages of 25 and 40.
+I identify the ethnicity of a person's parents through the parent's place of birth.
+A parent is Hispanic if they were born in a Spanish-speaking country.
+A parent is White if they were born in the United States.}}\\
 \endlastfoot
 \textbf{Panel A: Children's Education} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
-\hspace{1em}Men’s education (Total Years) & \specialcell{13.82\\(2.42)} & \specialcell{13.4\\(2.38)} & \specialcell{13.08\\(2.27)} & \specialcell{12.87\\(2.3)} & \specialcell{-0.94***\\(0.01)} & \specialcell{-0.32**\\(0.03)}\\
-\hspace{1em}Women’s education (Total Years) & \specialcell{14.06\\(2.37)} & \specialcell{13.62\\(2.39)} & \specialcell{13.26\\(2.35)} & \specialcell{13.22\\(2.39)} & \specialcell{-0.84***\\(0.01)} & \specialcell{-0.36**\\(0.02)}\\
+\hspace{1em}Men's education (Total Years) & \specialcell{13.82\\(2.42)} & \specialcell{13.4\\(2.38)} & \specialcell{13.08\\(2.27)} & \specialcell{12.87\\(2.3)} & \specialcell{-0.94***\\(0.01)} & \specialcell{-0.32**\\(0.03)}\\
+\hspace{1em}Women's education (Total Years) & \specialcell{14.06\\(2.37)} & \specialcell{13.62\\(2.39)} & \specialcell{13.26\\(2.35)} & \specialcell{13.22\\(2.39)} & \specialcell{-0.84***\\(0.01)} & \specialcell{-0.36**\\(0.02)}\\
 \textbf{Panel B: Children's Employment and Earnings} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
@@
 \addlinespace
-\hspace{1em}Women’s Unemployment Rate & \specialcell{0.04\\(0.81)} & \specialcell{0.06\\(0.24)} & \specialcell{0.07\\(0.75)} & \specialcell{0.06\\(0.76)} & \specialcell{0.02***\\(0.00)} & \specialcell{0.01***\\(0.00)}\\
-\hspace{1em}Men’s Log Hourly Earnings & \specialcell{2.51\\(0.45)} & \specialcell{2.43\\(0.46)} & \specialcell{2.42\\(0.44)} & \specialcell{2.42\\(0.43)} & \specialcell{-0.09***\\(0)} & \specialcell{-0.00**\\(0.02)}\\
-\hspace{1em}Women’s Log Hourly Earnings & \specialcell{2.32\\(0.49)} & \specialcell{2.31\\(0.45)} & \specialcell{2.28\\(0.45)} & \specialcell{2.31\\(0.42)} & \specialcell{-0.02***\\(0)} & \specialcell{-0.03**\\(0.02)}\\
-\hspace{1em}Men’s Log Annual Earnings & \specialcell{10.29\\(1.01)} & \specialcell{10.09\\(1.05)} & \specialcell{10.06\\(0.97)} & \specialcell{10.01\\(1.04)} & \specialcell{-0.28***\\(0.01)} & \specialcell{-0.03**\\(0.03)}\\
-\hspace{1em}Women’s Log Annual Earnings & \specialcell{9.43\\(1.77)} & \specialcell{9.52\\(1.59)} & \specialcell{9.47\\(1.56)} & \specialcell{9.53\\(1.49)} & \specialcell{0.1**\\(0.02)} & \specialcell{-0.05*\\(0.05)}\\*
+\hspace{1em}Women's Unemployment Rate & \specialcell{0.04\\(0.81)} & \specialcell{0.06\\(0.24)} & \specialcell{0.07\\(0.75)} & \specialcell{0.06\\(0.76)} & \specialcell{0.02***\\(0.00)} & \specialcell{0.01***\\(0.00)}\\
+\hspace{1em}Men's Log Hourly Earnings & \specialcell{2.51\\(0.45)} & \specialcell{2.43\\(0.46)} & \specialcell{2.42\\(0.44)} & \specialcell{2.42\\(0.43)} & \specialcell{-0.09***\\(0)} & \specialcell{-0.00**\\(0.02)}\\
+\hspace{1em}Women's Log Hourly Earnings & \specialcell{2.32\\(0.49)} & \specialcell{2.31\\(0.45)} & \specialcell{2.28\\(0.45)} & \specialcell{2.31\\(0.42)} & \specialcell{-0.02***\\(0)} & \specialcell{-0.03**\\(0.02)}\\
+\hspace{1em}Men's Log Annual Earnings & \specialcell{10.29\\(1.01)} & \specialcell{10.09\\(1.05)} & \specialcell{10.06\\(0.97)} & \specialcell{10.01\\(1.04)} & \specialcell{-0.28***\\(0.01)} & \specialcell{-0.03**\\(0.03)}\\
+\hspace{1em}Women's Log Annual Earnings & \specialcell{9.43\\(1.77)} & \specialcell{9.52\\(1.59)} & \specialcell{9.47\\(1.56)} & \specialcell{9.53\\(1.49)} & \specialcell{0.1**\\(0.02)} & \specialcell{-0.05*\\(0.05)}\\*
 \end{longtable}
 \end{ThreePartTable}
-\end{landscape}
+%\end{landscape}

### 2023-07-22 manuscripts/Hussain-Hadah/tables/tab05-regression.tex
add elsavier manuscript

--- 2023-05-09 output/tables/tab06-regression.tex
@@
-\begin{table}[H]
-
-\caption{Effect of Having Hispanic Last Name by Hispanic Identity\label{tab:identreg}}
-\centering
-\resizebox{\linewidth}{!}{
+\begin{table}[t]
+\caption{Effect of Having Hispanic Last Name \label{tab:lastnamereg}}
 \begin{threeparttable}
@@
 \midrule
-$HW_{ist} \times Hispanic_{ist}$ & \num{-0.09} & \num{-0.06} & \num{-0.05} & \num{-0.04} & \num{-0.04}\\
- & (\num{0.07}) & (\num{0.07}) & (\num{0.06}) & (\num{0.06}) & (\num{0.06})\\
-$Hispanic_{ist}$ & \num{-0.13}** & \num{-0.11} & \num{-0.09}** & \num{-0.03} & \num{-0.02}\\
- & (\num{0.07}) & (\num{0.07}) & (\num{0.04}) & (\num{0.05}) & (\num{0.04})\\
-$HW_{ist}$ & \num{0.03} & \num{0.01} & \num{0.00} & \num{0.02} & \num{0.02}\\
- & (\num{0.06}) & (\num{0.06}) & (\num{0.06}) & (\num{0.06}) & (\num{0.05})\\
-Constant & \num{10.53}*** & \num{9.57}*** &  &  & \\
- & (\num{0.05}) & (\num{0.08}) &  &  & \\
-\midrule
-$HW_{ist} \times Hispanic_{ist} + HW_{ist}$ & -0.06*** & -0.05** & -0.05** & -0.02 & -0.01\\
- & (0.02) & (0.02) & (0.02) & (0.02) & (0.02)\\
+$HW_{ist}$ & \num{-0.06}*** & \num{-0.05}*** & \num{-0.05}** & \num{-0.02} & \num{-0.01}\\
+ & (\num{0.02}) & (\num{0.02}) & (\num{0.02}) & (\num{0.02}) & (\num{0.02})\\
+Constant & \num{10.42}*** & \num{9.46}*** &  &  & \\
+ & (\num{0.04}) & (\num{0.07}) &  &  & \\
 \midrule
@@
 \begin{tablenotes}
-\item[1] \footnotesize{This table includes the estimation results of equation (\ref{eq:iden}).}
-\item[2] \footnotesize{HW is an indicator variable that is equal to 1 if a person is the child of a Hispanic-father and White-mother, and Hispanic is an indicator variable that is equal to one if a person self-identifies as Hispanic.}
-\item[3] \footnotesize{The sample is restricted to men working full-time full-year and are waged and salaried workers.}
-\item[4] \footnotesize{Column one has the regression results when controlling for hours worked, age, education, year and state fixed effects.
-Column two has the results after controlling for education.}
-\item[5] {\setstretch{1.0}\footnotesize{Standard errors are clustered on the state level.}}
+\item[1] {\footnotesize{This table includes the estimation results of equation (\ref{eq:1a}).}}
+\item[2] {\footnotesize{HW is an indicator variable that is equal to 1 if a person is the child of a Hispanic-father and White-mother.}}
+\item[3] {\footnotesize{The sample is restricted to men working full-time full-year and are wage and salary workers.}}
+\item[4] {\footnotesize{Column one has the regression results when controlling for hours worked, age, education, year and state fixed effects.
+Column two has the results after controlling for education.}}
+\item[5] {\footnotesize{Standard errors are clustered on the state level.}}
 \end{tablenotes}
-\end{threeparttable}}
+\end{threeparttable}
 \end{table}

### 2023-07-28 manuscripts/Hussain-Hadah/Hadah-Hispanic-Last-Names.tex
manuscript to JEBO

--- 2023-07-22 manuscripts/Hussain-Hadah/Hussain-Hadah.tex
@@
 \documentclass[a4paper,fleqn]{cas-sc}
+
+% \usepackage{mathptmx} % To set the font to Times New Roman
+% \let\cdot\relax
+\usepackage{amssymb}
+
+\usepackage{graphicx}
+\usepackage{caption}
+
+% Define a custom command to set the font for all tables
+\newcommand{\tablefont}{\fontfamily{ptm}\selectfont}
 
@@
 % Short title
-\shorttitle{The impact of Hispanic last names and identity on labor market outcomes}
+\shorttitle{The Impact of Hispanic Last Names and Identity on Labor Market Outcomes}
 
 % Short author
-\shortauthors{H. Hadah}
+\shortauthors{Hadah}
 
 % Main title of the paper
-\title [mode = title]{The impact of Hispanic last names and identity on labor market outcomes}
+\title [mode = title]{The Impact of Hispanic Last Names and Identity on Labor Market Outcomes}
 
@@
 I compare the children of inter-ethnic marriages to study the impact of having a Hispanic last name.
-While males born to Hispanic father-White mothers earn 6 percentage points less than those born to White father-Hispanic mothers, the gap could be completely explained by educational differences.
+While males born to Hispanic father-White mothers earn 5 percentage points less than those born to White father-Hispanic mothers, the gap could be completely explained by educational differences.
 I also study the effect of identifying as Hispanic on earnings.
@@
  Economics of Minorities, Race, and Immigrants\sep Discrimination and Prejudice\\
- \JEL{J71\sep J64}
+ \JEL{J15, J64, J71}
 \end{keywords}
@@
 \maketitle
-
 \section{Introduction}\label{sec:intro}
@@
 A large literature documents substantial earnings gaps across race and ethnicity \citep{bayer2018divergent, charles2008prejudice, card1992school, fryer2004causes, rubinstein2014pride, bertrand2004emily, juhn1991accounting}.
-Hispanics constitute a large and growing portion of the population in the United States.
+Hispanics constitute a large and growing portion of the population in the United States, see figure \ref{fig:3}.
 As the number of Hispanics increases, determining whether ethnic discrimination affects their labor market outcomes becomes increasingly crucial \citep{chettyUnitedStatesStill2014, chettyEffectsExposureBetter2016,chettyFadingAmericanDream2017,abramitzkyImmigrantsAssimilateMore2020a, abramitzkyNationImmigrantsAssimilation2014,abramitzkyCulturalAssimilationAge2016,chettyWhereLandOpportunity2014}.
@@
 \centering
-\includegraphics[scale=0.25]{Fig/interracialovertime.png}
-\caption{Inter-racial marriages as a percent of all marriagesin the United States from 1995 to 2020.}
+\includegraphics[scale=0.50]{Fig/Hispanics_all.png}
+\captionsetup{width=0.85\linewidth} % Adjust the width to your preference
+\caption{Hispanics as a percent of the population in the United States from 1994 to 2021.}
 \label{fig:3}
+
+\begin{minipage}{\linewidth}
+\footnotesize
+\textit{Source:} The data source is the Current Population Survey (CPS) from 1994 to 2021.\par
+\textit{Note:} The graph shows the percent of the population in the US that self-identifies as Hispanic.
+\end{minipage}
 \end{figure}
@@
 \centering
-\includegraphics[scale=0.4]{Fig/PirChart2.png}
+\includegraphics[scale=0.5]{Fig/PirChart2.png}
 \caption{Distribution of the four types of children.}
 \label{fig:dist}
+\begin{minipage}{\linewidth}
+\footnotesize
+\textit{Source:} The data source is the Current Population Survey (CPS) from 1994 to 2021.\par
+\end{minipage}
 \end{figure}
@@
 
-The distribution of the four types of children is presented in Fig.
-\ref{fig:dist} and Table \ref{tab:mat1}.
+The distribution of the four types of children is presented in figure \ref{fig:dist} and table \ref{tab:mat1}.
 The majority of the sample (95.96\%) is WW children.
@@
 Even though WH and HW are only 1.35\% of the sample, I have plenty of observations to carry out an analysis.
-The summary statistics for the children of the four types of marriages are presented in Table \ref{tab:c&p1}.
+The summary statistics for the children of the four types of marriages are presented in table \ref{tab:c&p1}.
 Children of WW marriages (column 1) do better on every measure while children of HH parents (column 4) do worse than other children on every measure.
@@
 At the same time, the WH children have White fathers with a non-Hispanic, or White, last names.
-Summary statistics of these two groups, Table \ref{tab:c&p1} columns (2) and (3) can show that they are different from the children of non-inter-ethnic marriages.
+Summary statistics of these two groups, table \ref{tab:c&p1} columns (2) and (3) can show that they are different from the children of non-inter-ethnic marriages.
 
@@
 
-In panel C of Table \ref{tab:c&p1}, I present the rates at which members of each group self-identify as Hispanic.
+In panel C of table \ref{tab:c&p1}, I present the rates at which members of each group self-identify as Hispanic.
 \footnote{Hispanic self-identification is observed when an individual notes that they are Hispanic on the CPS questionnaire} The vast majority of HH children, 96\% for men and 97\% for women, identify as Hispanic.
@@
 
-In Table \ref{tab:c&p2}, I present the summary statistics for the four types and limit the sample to those that only identify as Hispanic.\footnote{Limiting the sample to those who identify as Hispanic only affects WH, HW, and HH.
+In table \ref{tab:c&p2}, I present the summary statistics for the four types and limit the sample to those that only identify as Hispanic.\footnote{Limiting the sample to those who identify as Hispanic only affects WH, HW, and HH.
 Members of the WW group are not affected by this restriction.} The same pattern discussed above between HH--WW and HW--WH persists.
@@
 I assume that parents have children between the ages of 25 and 40.
-Therefore, my sample consists of married White men and women with children that were born in the 1920 to 1975 cohorts\footnote{The construction of "synthetic parents" follows the method used by \citet{rubinstein2014pride}.}.
+Therefore, my sample consists of married White men and women with children that were born in the 1920 to 1975 cohorts, see figure \ref{flowchart1} for a flowchart of the linking process.\footnote{The construction of "synthetic parents" follows the method used by \citet{rubinstein2014pride}.}
 
@@
 
-I show the distribution of the four types of couples in Table \ref{tab:mat2}.
+I show the distribution of the four types of couples in table \ref{tab:mat2}.
 White husbands and White wives (WW) make up the majority of couples in the sample, 96\% (5,141,737 couples).
@@
 
-I present the summary statistics of the four types of synthetic parents in Table \ref{tab:synth}.
+I present the summary statistics of the four types of synthetic parents in table \ref{tab:synth}.
 WW couples have higher education, 12.75 years for husbands and 12.47 for wives.
@@
 
-In columns 5 and 6 of Table \ref{tab:synth}, I report the differences in means between HH-WW households and HW-WH households.
+In columns 5 and 6 of table \ref{tab:synth}, I report the differences in means between HH-WW households and HW-WH households.
 An HH Hispanic husband attained 4.11 fewer years of education than a WW husband.
@@
 I provide the results to the estimation of Eq.
-\eqref{eq:1a} in Table \ref{tab:lastnamereg}.
+\eqref{eq:1a} in table \ref{tab:lastnamereg}.
 The sample in this analysis includes full-time, full-year, and wage and salary men, and I control for hours worked and age and include year fixed effects (FE).
 I also restrict the sample to children of HW and WH (omitted) parents.
-Column one in Table \ref{tab:lastnamereg} is the results without controlling for education.
-Overall, an inter-ethnic with a Hispanic last name earns 6 percentage points less than an inter-ethnic with a White last name.
+Column one in table \ref{tab:lastnamereg} is the results without controlling for education.
+Overall, an inter-ethnic with a Hispanic last name earns 5 percentage points less than an inter-ethnic with a White last name.
 This gap, however, could be entirely explained by educational differences.
-An inter-ethnic with a Hispanic last name earns 6 percentage points less than an inter-ethnic with a White last name, but the result is statistically insignificant.
+An inter-ethnic with a Hispanic last name earns 2 percentage points less than an inter-ethnic with a White last name, but the result is statistically insignificant.
 
 I provide the results to the estimation of Eq.
-\eqref{eq:iden} in Table \ref{tab:identreg}.
+\eqref{eq:iden} in table \ref{tab:identreg}.
 The sample in this analysis includes full-time, full-year, and wage and salary men, and I control for hours worked and age and include year fixed effects (FE).
@@
 The omitted group is children of WH parents that do not identify as Hispanic.
-Column one in Table \ref{tab:identreg} is the results without controlling for education.
+Column one in table \ref{tab:identreg} is the results without controlling for education.
 An inter-ethnic $HW_{ist}$ that identifies as Hispanic \textit{versus} $HW_{ist}$ who do not---i.e.
-$HW_{ist} + HW_{ist} \cdot Hispanic_i$---earns 6 percentage points less than an inter-ethnic that does not, but the result could be entirely explained by educational differences.
+$HW_{ist} + HW_{ist} \cdot Hispanic_i$---earns 5 percentage points less than an inter-ethnic that does not, but the result could be entirely explained by educational differences.
 
@@
 I compare the children of inter-ethnic marriages to study the labor market impact of having a Hispanic last name.
-When I compare the earnings of HW and WH children, which captures the effect of having a Hispanic last name, HW children earn  6 percentage points less than WH children.
-Thus, by comparing inter-ethnic children, a person with a Hispanic last name makes 6 percentage points less than someone with a White last name.
+When I compare the earnings of HW and WH children, which captures the effect of having a Hispanic last name, HW children earn  5 percentage points less than WH children.
+Thus, by comparing inter-ethnic children, a person with a Hispanic last name makes 5 percentage points less than someone with a White last name.
 When I control for education, the last name effect decreases to a statistically insignificant 2 percentage points earnings gap.
@@
 
+While the earnings gap between children of inter-ethnic parents with and without a Hispanic last name disappears when controlling for education, it does not necessarily indicate the absence of discrimination.
+Education itself can be influenced by bias, potentially resulting in divergent outcomes.
+Consequently, further research is essential to comprehensively understand the earnings gaps between Hispanics and Whites.
+
+\section*{Declaration of Competing Interest}
+The author declares that he has no relevant or material financial interests that relate to the research described in this paper.
+
+\section*{Data availability}
+
+Data is available \href{https://github.com/hhadah/hispanic-last-names/tree/main/data/datasets}{here}.
+
 \bibliographystyle{cas-model2-names}

### 2023-07-28 manuscripts/Hussain-Hadah/title-page.tex
manuscript to JEBO

--- 2023-07-22 manuscripts/Hussain-Hadah/Hussain-Hadah.tex
@@
             city={New Orleans},
-%          citysep={}, % Uncomment if no comma needed between city and postcode
             postcode={70118},
@@
 Economics of Minorities, Race, and Immigrants\sep Discrimination and Prejudice\\
- \JEL{J71\sep J64}
+\JEL{J15, J64, J71}
 \end{keywords}
@@
 
-\section{Introduction}\label{sec:intro}
-
-A large literature documents substantial earnings gaps across race and ethnicity \citep{bayer2018divergent, charles2008prejudice, card1992school, fryer2004causes, rubinstein2014pride, bertrand2004emily, juhn1991accounting}.
-Hispanics constitute a large and growing portion of the population in the United States.
-As the number of Hispanics increases, determining whether ethnic discrimination affects their labor market outcomes becomes increasingly crucial \citep{chettyUnitedStatesStill2014, chettyEffectsExposureBetter2016,chettyFadingAmericanDream2017,abramitzkyImmigrantsAssimilateMore2020a, abramitzkyNationImmigrantsAssimilation2014,abramitzkyCulturalAssimilationAge2016,chettyWhereLandOpportunity2014}.
-Thus, it is important to understand whether a person's ethnicity affects their labor market outcomes.
-Assimilation and mobility are crucial because they reflect how well Hispanics can integrate into society and move up the socioeconomic ladder.
-
-In this paper, I answer the following questions.
-Does having a Hispanic last name affect labor market outcomes?
-What is the effect of identifying as Hispanic on earnings?
-Moreover, I aim to show that comparing Hispanic Whites to non-Hispanic Whites might create an artificially higher earnings gap since the two groups differ on many observable characteristics.\footnote{By identifying as Hispanic, I refer to individuals who self-report their ethnicity as Hispanic on surveys or other data collection instruments.} \footnote{Observable characteristics refer to factors that can be measured and quantified, such as education level, work experience, and immigration status.} Others have attempted to compare how native-born White Hispanics fare to non-Hispanic Whites and foreign-born Hispanics.
-In \citet{antman2020ethnic,antmanEthnicAttritionObserved2016,antmanEthnicAttritionObserved2016a,antmanEthnicAttritionAssimilation2020}, the authors compare the health and educational outcomes of Hispanic Whites to non-Hispanic Whites and native-born Hispanics to foreign-born Hispanics.
-They find gaps in education and health between Hispanics and Whites.
-They also find that native-born Hispanics are more likely than their foreign-born counterparts to report poor health.
-\citet{davilaChangesRelativeEarnings2008} documents many gaps in labor market outcomes between Hispanics and Whites.
-They attribute a big part of this gap to differences in education, experience, immigration status, and regional differences.
-
-The US population is growing in diversity.
-The proportion of non-Whites has increased by more than 10 percentage points from 13 percent in 1995 to 23 percent in 2019.
-The number of Hispanics has grown by 9 percentage points from 9 percent in 1995 to 16 percent in 2019.\footnote{The portion of non-Whites and Hispanics is calculated using the Current Population Survey (CPS).} Native-born White Hispanic men earn 21\% less than White men, although a substantial portion of the earnings gap is due to educational differences between Hispanics and Whites \citep{duncan2006hispanics, duncan2018identifying, duncan2018socioeconomic}.
-Some of the earnings differences may also be due to discrimination which will have negative consequences.
-For example, discrimination against Hispanics can lead to reduced job opportunities, lower wages, and hinder assimilation.
-In this paper, I examine the role of having a Hispanic last name and identifying as Hispanic on labor market outcomes.
-
-Identifying discrimination is difficult because of factors that affect labor market outcomes that are unobservable to economists---such as unobserved skills.
-One strategy used by researchers is audit or resume studies.
-\citet{bertrand2004emily} conducted an audit study where identical resumes were sent to employers with White and Black-sounding names.
-This approach, however, has its drawbacks.
-Audit studies only observe callbacks, not wages.
-
-\begin{figure}[t]
-\centering
-\includegraphics[scale=0.25]{Fig/interracialovertime.png}
-\caption{Inter-racial marriages as a percent of all marriagesin the United States from 1995 to 2020.}
-\label{fig:3}
-\end{figure}
-
-This study utilizes a method developed by \citet{rubinstein2014pride}.
-I compare children from inter-ethnic marriages.
-More precisely, I compare children of Hispanic fathers and White mothers (henceforth HW) to children of White fathers and Hispanic mothers (henceforth  WH).
-This approach stems from the fact that marriages are not random.
-Couples match on several observable characteristics like income, schooling, socio-economic background, etc.
-\citep{averettBetterWorseRelationship2008, averettEconomicRealityBeauty1996, beckerTheoryMarriagePart1973, beckerTheoryMarriagePart1974, beckerTreatiseFamily1993, browningCollectiveUnitaryModels2006, chiapporiFatterAttractionAnthropometric2012}.
-Children of HW and WH marriages have more similar observable characteristics than children of endogamous/homogamous marriages--- i.e., White fathers-White mothers and Hispanic fathers-Hispanic mothers.
-Moreover, children from a Hispanic father and White mother household will have a Hispanic last name from their fathers, enabling the investigation of how ethnic signals, such as having a Hispanic last name, affect annual log earnings.
-
-Using the Current Population Survey (CPS) from 1994 to 2019 and the 1960 to 2000 US censuses, I estimate the effect of having a Hispanic last name on labor market outcomes.
-In other words, I approximate the variation in labor market perception of ethnic signals, in this case having a Hispanic last name, on annual log earnings.
-I also study the effect of identifying as Hispanic on labor market outcomes.
-
-I find that while men born to Hispanic father-White mothers earn less than men born to White father-Hispanic mothers, this gap is entirely explained by educational differences.
-Thus, I do not find a significant effect of having a Hispanic last name.
-Finally, I find that men identifying as Hispanic earn significantly less than those who do not, even when I control for ancestry and education.
-
-Previous studies have also used names as a proxy for race and ethnicity \citep{fryer2004causes, rubinstein2014pride, bertrand2004emily}.
-\citet{fryer2004causes} point out that names can be a predictor of a person's race.
-Specifically, they provide a rising pattern among Blacks having different names than Whites.
-They did, however, find that having a Black name, after controlling for the home environment at birth, does not affect their labor market outcomes.
-\citet{rubinstein2014pride} compared the children of mixed marriages between Sephardic and Ashkenazi Jews in Israel.\footnote{Ashkenazi and Sephardic are two distinct Jewish ethnic groups.} They found that workers with Sephardic last names earn substantially less than those with Ashkenazi last names.
-\citet{bertrand2004emily} conducted an audit study by sending employers identical resumes that differ in the ethnic and racial signal of a name (Black sounding name versus a White sounding one).
-They found that resumes with Black-sounding names received substantially fewer callbacks than their White counterparts.
-
-The rest of the paper is organized as follows.
-Section \ref{sec:data} provides an overview of the data and summary statistics of the sample.
-Section \ref{sec:emp_model} introduces the empirical framework.
-I report the results in Section \ref{sec:results}.
-Finally, I offer a brief conclusion in Section \ref{sec:con1}.
-
-\section{Data}\label{sec:data}
-
-I use two datasets:  the Integrated Public Use Microdata Series (IPUMS) Current Population Survey (CPS) Annual Social and Economic (ASEC) \citep{cps2019} and the 1960 to 2000 US censuses \citep{acs2019}.
-
-I use the CPS data set to study the effect of having a Hispanic last name on a person's labor market outcomes.
-I take advantage of the fact that the CPS asks parents' place of birth, ethnicity, and race.
-The data spans the period between 1994--- the earliest sample to ask about a parent's place of birth--- to 2019.
-Moreover, since the CPS does not provide data on parents' characteristics, essential to determine the family background, I use the census to construct synthetic parents.
-The census offers a larger sample of potential parents.
-Similar to the CPS, the 1960 to 2000 censuses ask about the person's place of birth and the individual's race and ethnicity.
-I employ this information to construct "synthetic parents" using a method developed by \citet{rubinstein2014pride}.
-I construct the synthetic parents by linking husbands and wives in the census data to each other.
-I assume that parents have children between the ages of 25 and 40.
-I then link these synthetic parents using the birth year of child, and place of birth of parents to the year of birth of the "children" and the palce of birth of parents in the CPS sample.
-
-\subsection{Children of the four parental types}
-
-I use the CPS for my primary analysis of the effect of having a Hispanic last name on earnings.
-I restrict my sample to Whites, United States-born citizens aged 25 to 40 and born between 1960 to 2000.
-Taking advantage of data on parents' place of birth, I divide the sample into four groups depending on their parent's ethnicity.
-Mothers or fathers are Hispanic if they were born in a Spanish-speaking country and Puerto Rico, and White if they were born in the United States.\footnote{The list of Spanish Speaking countries include: Argentina, Bolivia, Chile, Colombia, Costa Rica, Cuba, Dominican Republic, Ecuador, El Salvador, Guatemala, Equatorial Guinea, Honduras, Mexico, Nicaragua, Panama, Paraguay, Peru, Spain, Uruguay, Venezuela.} Therefore, an observation can be the product of four types of parents:
-\begin{enumerate}
-\item White father and White mother (hereafter WW)
-\item White father and Hispanic mother (hereafter WH)
-\item Hispanic father and White mother (hereafter HW)
-\item Hispanic father and Hispanic mother (hereafter HH).
-\end{enumerate}
-
-\begin{figure}[t]
-\centering
-\includegraphics[scale=0.4]{Fig/PirChart2.png}
-\caption{Distribution of the four types of children.}
-\label{fig:dist}
-\end{figure}
-
-\begin{figure}[t]
-\centering
-\begin{tikzpicture}[node distance =2cm]
-\node (start) [startstop] {Synthetic Parents: Observed in 1960 to 2000. $25 \leq age \leq 40$, White, with kids};
-
-\node (pro1) [process, below of = start, yshift = -2cm] {Children: Observed between 1994 and 2019. Born between 1954 to 1999 and $25 \leq age \leq 40$};
-
-\draw [arrow] (start) -- (pro1);
-\end{tikzpicture}
-\caption{Chart explaining which synthetic parents and children and when we observe them.}
-\label{flowchart1}
-\end{figure}
-
-The distribution of the four types of children is presented in Fig.
-\ref{fig:dist} and Table \ref{tab:mat1}.
-The majority of the sample (95.96\%) is WW children.
-The second biggest group is HH, which constitutes 2.69\% of the sample.
-Inter-ethnic children, WH and HW, make up 1.35\% of the sample with 90,325 observations.
-Even though WH and HW are only 1.35\% of the sample, I have plenty of observations to carry out an analysis.
-The summary statistics for the children of the four types of marriages are presented in Table \ref{tab:c&p1}.
-Children of WW marriages (column 1) do better on every measure while children of HH parents (column 4) do worse than other children on every measure.
-
-%table1
-\input{tables/tab01-observations-by-parents}
-%table2
-\input{tables/tab02-observations-by-couple}
-
-On the one hand, WW children are more educated, with approximately 14 years of education for men and women.
-WW children have an unemployment rate of 4\% for men and women.
-They have log hourly earnings of 2.51 for men and 2.32 for women.
-Also, their annual log earnings are 10.29 for men and 9.43 for women.
-
-On the other hand, children of HH marriages have an average of 12.9 and 13.24 years of education for men and women, respectively.
-HH men and women have an unemployment rate of 7\% and 6\%.
-Their annual log earnings are 10.01 and 9.53 for men and women, respectively.
-
-Inter-ethnic children, WH and HW, have lower educational attainment and income than WW children, but they are better than HH children.
-Also, WH and HW are comparable, unlike WW and HH children.
-These two groups are important to the analysis.
-Children of HW are going to be the treatment group.
-HW children have a Hispanic father with a Hispanic last names.
-At the same time, the WH children have White fathers with a non-Hispanic, or White, last names.
-Summary statistics of these two groups, Table \ref{tab:c&p1} columns (2) and (3) can show that they are different from the children of non-inter-ethnic marriages.
-
-%table3
-\input{tables/tab03b-summary-stats}
-
-WH children have an average of 13.58 and 13.79 years of education for men and women.
-The unemployment rate for men and women is 5\%.
-HW men and women obtained 13.22 and 13.42 years of schooling.
-Men's unemployment rate is 7\% and women's 6\%.
-
-In panel C of Table \ref{tab:c&p1}, I present the rates at which members of each group self-identify as Hispanic.
-\footnote{Hispanic self-identification is observed when an individual notes that they are Hispanic on the CPS questionnaire} The vast majority of HH children, 96\% for men and 97\% for women, identify as Hispanic.
-There is considerable attrition in Hispanic identification of inter-ethnic children.
-Among WH children, 74\% of men and 78\% of women identify as Hispanic.
-Among HW children, 83\% of men and 82\% of women identify as Hispanic.
-These results align with the findings of a substantial body of literature that noted this attrition among children of immigrants, especially those coming from inter-ethnic marriages \citep{duncan2017complexity, duncan2018identifying, duncan2020new, antman2020ethnic}.
-Finally, a small portion of WW children, 4\% of men and 5\% of women, identified as Hispanic.
-These 5\% of WW children are most likely third generation--- or higher--- children of immigrants.
-For my analysis, I dropped the WW observations who identify as Hispanic when I compare with non-Hispanic Whites.
-
-In columns (5) and (6), I calculate the differences in means between HH-WW (column 5) and HW-WH (column 6).
-Overall, HH children do worse than WW children.
-HH has one year less of education compared to WW, and are 2 percentage points more likely to be unemployed.
-HH men earn 9\% less than WW men, and HH women make 2\% less than WW women.
-Therefore, these two groups, HH and WW, are different on many observable qualities.
-
-The differentials between WH and HW are not as severe as the ones between HH and WW.
-HW children are still worse off than WH but are much better than HH.
-HW men attain 0.36 fewer years of education than WH men.
-HW women achieve 0.37 fewer years of schooling than WH women.
-HW children are 1 percentage points more likely to be unemployed.
-HW men earn 1\% less annual earnings as WH men, and HW women earn 3\% less.
-These differences show that HW and WH children are different from children of endogamous/homogamous marriages and, thus, are more comparable.
-
-In Table \ref{tab:c&p2}, I present the summary statistics for the four types and limit the sample to those that only identify as Hispanic.\footnote{Limiting the sample to those who identify as Hispanic only affects WH, HW, and HH.
-Members of the WW group are not affected by this restriction.} The same pattern discussed above between HH--WW and HW--WH persists.
-The one difference is that those who identify as Hispanic do slightly worse.
-
-%table4
-\input{tables/tab03a-summary-stats}
-
-\subsection{Synthetic parents}
-
-Using the 1960 to 2000 censuses, I constructed a data set of synthetic parents.
-The sample includes married White men and women.
-Even though the census asks a person whether they are Hispanic or not, I took advantage of the questions on place of birth to create a proxy for ethnicity.
-I consider a Hispanic person as White and born in a Spanish-speaking country.
-Consequently, Whites in the sample are people who are White and native-born.
-Using the information provided in the census, I can link husbands and wives with each other.
-I assume that parents have children between the ages of 25 and 40.
-Therefore, my sample consists of married White men and women with children that were born in the 1920 to 1975 cohorts\footnote{The construction of "synthetic parents" follows the method used by \citet{rubinstein2014pride}.}.
-
-%table5
-\input{tables/tab04-summary-stats}
-
-I show the distribution of the four types of couples in Table \ref{tab:mat2}.
-White husbands and White wives (WW) make up the majority of couples in the sample, 96\% (5,141,737 couples).
-Hispanic husbands and wives (HH) are the second largest group making up 2\% (119,749 couples) of all couples.
-White husbands and Hispanic wives (WH) couples are less 1\% (33,097 couples) of the sample and Hispanic husbands and White wives (HW) are also 1\% (37,847 couples).
-
-I present the summary statistics of the four types of synthetic parents in Table \ref{tab:synth}.
-WW couples have higher education, 12.75 years for husbands and 12.47 for wives.
-As a household, WW couples have 25.22 years of schooling.
-Additionally, the log total family income for a WW household is 10.72 and have 3.77 children.
-Husbands in HH marriages have 8.64 years of education, while women have 8.49 years of schooling.
-As a household, HH couples have 17.13 years of education, log total family income of 10.27, and have 4.23 children.
-
-WH husbands have 11.77 years of education, while wives have an average of 10.40 years.
-WH household attained a total of 22.17 years of schooling in total, a log total family income of 10.55, and 3.98 children.
-HW husbands and wives obtained 10.25 and 11.11 years of education, respectively.
-HW household attained a total of 21.36 years of schooling in total, a log total family income of 10.46, and 4.15 children.
-
-In columns 5 and 6 of Table \ref{tab:synth}, I report the differences in means between HH-WW households and HW-WH households.
-An HH Hispanic husband attained 4.11 fewer years of education than a WW husband.
-While an HH wife acquired 3.98 fewer years of schooling than a White WW wife.
-The total educational difference between HH and WW households equals 8.09 years of education.
-The educational differences between HW and WH are less severe than those between HH and WW.
-A Hispanic husband in intermarriage attained 1.52 fewer years of schooling than a White husband in intermarriage.
-A White wife in intermarriage attained 0.70 years of education, more than a Hispanic wife in intermarriage.
-The difference in total years of education between HW and WH households is 0.81, which is an 90\% decrease from the HH--WW gap.
-The total family income of an HH household is 45\% lower than a WW household, while the total family income of an HW household is 10\% lower than a WH household.
-These differences show that couples that intermarry are better and different than those that do not.
-
-\section{Empirical Strategy}\label{sec:emp_model}
-
-In this section, I present two empirical strategies.
-The first empirical strategy estimates the effect of having a Hispanic last name on annual log earnings.
-The second empirical strategy estimates the effect of identifying as Hispanic on log annual earnings.
-
-The difference in means between Hispanics and non-Hispanic Whites could result from discrimination.
-It can also be caused by differences in innate abilities, skills, and parental investments.
-While controlling for observable skill measures, I compare children of inter-ethnic marriages, HW and WH.
-WH and HW children are more similar in characteristics but provide employers, and the labor market, with different signals.\footnote{WH and HW children are both half White, half Hispanic.} WH children will have a non-Hispanic last name, while HW children will have a Hispanic last name.
-This is a method developed by \citet{rubinstein2014pride}.
-
-%table6
-\input{tables/tab05-regression}
-\subsection{Estimating the effect of having a Hispanic last name}
-
-In this section I restrict the sample to WH and HW groups.
-Let $Y_{ist}$ be the annual log earnings of person $i$ in state $s$ at time $t$.
-$HW_{ist}$ is an indicator variables for the type of parents person $i$ has.
-$X_{ist}$ is a vector of controls that includes age and numbers of hours worked, $\gamma_{t}$ are year fixed effects, $\lambda_{s}$ are state fixed effects, and $\phi_{ist}$ represents the error term.
-The equation for this strategy is written as follows:
-\begin{equation} \label{eq:1a}
-Y_{ist} = \beta_{1} HW_{ist} + X_{ist} \pi + \gamma_{t} + \lambda_s + \phi_{ist}
-\end{equation}
-
-$\beta_{1}$ is the coefficient of interest in this specification.
-$\beta_{1}$ represents the earnings gaps between children of inter-ethnic marriages who have a Spanish-sounding last name versus a White last name.
-
-\subsection{Estimating the effect of identifying as Hispanic}
-
-In this section, I will present an estimation strategy that would allow me to capture the effect of identifying as a Hispanic.
-
-\begin{equation} \label{eq:iden}
-Y_{ist} = \beta_{1} HW_{ist} + \beta_{2} Hispanic_i + \beta_{3} HW_{ist} \cdot Hispanic_i + X_{ist} \pi + \gamma_{t} + \lambda_s + \phi_{ist}
-\end{equation}
-
-The coefficients of interest from Eq.
-\eqref{eq:iden} are $\beta_{1}$, $\beta_{2}$ and $\beta_{3}$.
-The coefficients $\beta_{1}$ and $\beta_{3}$ capture the effects of Spanish-sounding last names for $HW_{ist}$ that identify as Hispanic \textit{versus} those who do not.
-%table7
-\input{tables/tab06-regression}
-\section{Results}\label{sec:results}
-
-In this section, I present the results from estimating the two specifications in Eqs.
-\eqref{eq:1a} and \eqref{eq:iden}.
-The results are shown in Tables \ref{tab:lastnamereg} and \ref{tab:identreg}.
-All estimations are done on a sample of White, native-born men between the ages of 25 and 40 employed for full-time full-year (FTFY) and are wage and salary workers.
-
-I find that an inter-ethnic person with a Spanish-sounding last names earns less than an inter-ethnic who has a White last name.
-A person with a Spanish-sounding last name earns 5 percentage points less than a person with a White last name.
-In other words, by comparing inter-ethnic children, a person with a Hispanic last name earns 5 percentage points less than someone with a White last name.
-However, more than half of the earnings gaps could be explained by educational differences.
-When I control for education, the last name effect decreases to a statistically insignificant 2 percentage points earnings gap.
-I also find a significant earnings gap between those that identify as Hispanic.
-
-\subsection{The Effect of Having a Hispanic Last Name and Hispanic Identity on Labor Market Outcomes}
-
-I provide the results to the estimation of Eq.
-\eqref{eq:1a} in Table \ref{tab:lastnamereg}.
-The sample in this analysis includes full-time, full-year, and wage and salary men, and I control for hours worked and age and include year fixed effects (FE).
-I also restrict the sample to children of HW and WH (omitted) parents.
-Column one in Table \ref{tab:lastnamereg} is the results without controlling for education.
-Overall, an inter-ethnic with a Hispanic last name earns 6 percentage points less than an inter-ethnic with a White last name.
-This gap, however, could be entirely explained by educational differences.
-An inter-ethnic with a Hispanic last name earns 6 percentage points less than an inter-ethnic with a White last name, but the result is statistically insignificant.
-
-I provide the results to the estimation of Eq.
-\eqref{eq:iden} in Table \ref{tab:identreg}.
-The sample in this analysis includes full-time, full-year, and wage and salary men, and I control for hours worked and age and include year fixed effects (FE).
-I also restrict the sample to children of HW and WH parents.
-The omitted group is children of WH parents that do not identify as Hispanic.
-Column one in Table \ref{tab:identreg} is the results without controlling for education.
-An inter-ethnic $HW_{ist}$ that identifies as Hispanic \textit{versus} $HW_{ist}$ who do not---i.e.
-$HW_{ist} + HW_{ist} \cdot Hispanic_i$---earns 6 percentage points less than an inter-ethnic that does not, but the result could be entirely explained by educational differences.
-
-\section{Conclusion}\label{sec:con1}
-
-As the Hispanic population grows in the United States, studying discrimination against this group becomes increasingly important.
-In this paper, I examine discrimination against Hispanics in the labor market.
-More specifically, I examine the impact of Hispanic last names and Hispanic identification on annual log earnings.
-
-I compare the children of inter-ethnic marriages to study the labor market impact of having a Hispanic last name.
-When I compare the earnings of HW and WH children, which captures the effect of having a Hispanic last name, HW children earn  6 percentage points less than WH children.
-Thus, by comparing inter-ethnic children, a person with a Hispanic last name makes 6 percentage points less than someone with a White last name.
-When I control for education, the last name effect decreases to a statistically insignificant 2 percentage points earnings gap.
-Moreover, I estimate the effect of identifying as Hispanic and having a Hispanic last name.
-I find that an inter-ethnic child with a Hispanic last name and identifies as Hispanic earns 5 percentage points less that an inter-ethnic child with native sounding last name and identifies as Hispanic.
-The earnings gap, however, could be entirely explained by educational differences.
-
-\bibliographystyle{cas-model2-names}
-\bibliography{bibliography}
-
 \end{document}

### 2023-07-28 my_paper/Hadah-last-names.tex
manuscript to JEBO

@@
 I find that an inter-ethnic child with a Hispanic last name and identifies as Hispanic earns 5 percentage points less that an inter-ethnic child with native sounding last name and identifies as Hispanic.
-The earnings gap, however, could be entirely explained by educational differences.
+
+While the earnings gap between children of inter-ethnic parents with and without a Hispanic last name disappears when controlling for education, it does not necessarily indicate the absence of discrimination.
+Education itself can be influenced by bias, potentially resulting in divergent outcomes.
+Consequently, further research is essential to comprehensively understand the earnings gaps between Hispanics and Whites.
 

### 2023-08-14 manuscripts/Hussain-Hadah/Hadah-Hispanic-Last-Names.tex
edit JMP.tex

@@
 \documentclass[a4paper,fleqn]{cas-sc}
+\usepackage{setspace} % Load the setspace package for line spacing
+
+% Set line spacing to double spaced
+\doublespacing
 
@@
 \begin{abstract}
+\singlespacing % Set single spacing for the abstract
 In this paper, I study discrimination against Hispanics in the labor market.
@@
 \begin{keywords}
+\singlespacing % Set single spacing for keywords
+
  Economics of Minorities, Race, and Immigrants\sep Discrimination and Prejudice\\

### 2023-08-14 my_paper/Hadah-last-names.tex
edit JMP.tex

@@
 
-\begin{center}
-\begin{figure}[h!]
-\caption{Inter-racial marriages as a percent of all marriages \\
- in the United States from 1995 to 2020.}
-\includegraphics[width=\textwidth]{interracialovertime.png} 
-\label{fig:3}
-\end{figure}
-\end{center}
+% \begin{center}
+% \begin{figure}[h!]
+% \caption{Inter-racial marriages as a percent of all marriages \\
+%  in the United States from 1995 to 2020.}
+% \includegraphics[width=\textwidth]{figures/Hispanics_all.png} 
+% \label{fig:3}
+% \end{figure}
+% \end{center}
 
-\newpage
+% \newpage
 
@@
 
-\newpage
+% \newpage
 
-\begin{figure}[h!]
-\centering
-\caption{Chart explaining which synthetic parents and children and when we observe them.}
-\begin{tikzpicture}[node distance =2cm]
-\node (start) [startstop] {Synthetic Parents: Observed in 1960 to 2000. $25 \leq age \leq 40$, White, with kids};
+% \begin{figure}[h!]
+% \centering
+% \caption{Chart explaining which synthetic parents and children and when we observe them.}
+% \begin{tikzpicture}[node distance =2cm]
+% \node (start) [startstop] {Synthetic Parents: Observed in 1960 to 2000. $25 \leq age \leq 40$, White, with kids};
 
-\node (pro1) [process, below of = start, yshift = -2cm] {Children: Observed between 1994 and 2019. Born between 1954 to 1999 and $25 \leq age \leq 40$};
+% \node (pro1) [process, below of = start, yshift = -2cm] {Children: Observed between 1994 and 2019. Born between 1954 to 1999 and $25 \leq age \leq 40$};
 
-\draw [arrow] (start) -- (pro1);
-\end{tikzpicture}
-\label{flowchart1}
-\end{figure}
+% \draw [arrow] (start) -- (pro1);
+% \end{tikzpicture}
+% \label{flowchart1}
+% \end{figure}
 

### 2023-08-14 my_paper/Hadah-last-names.tex
submit paper to JODE

@@
 %% https://www.jwe.cc/blog/
+
+% Redefine figure and table counters to use Arabic numerals
+\renewcommand{\thefigure}{\arabic{figure}}
+\renewcommand{\thetable}{\arabic{table}}
 
@@
 
-\linespread{1}{\title{{\fontfamily{ptm}
-\selectfont{\Large{\MakeUppercase{The Impact of Hispanic Last Names and Identity on Labor Market Outcomes}}}}}}
+\linespread{1}{\title{{\fontfamily{ptm}\selectfont{\Large{\MakeUppercase{The Impact of Hispanic Last Names and Identity on Labor Market Outcomes}}}}
+\thanks{I thank Professors Willa Friedman, Chinhui Juhn, Vikram Maheshri, and Yona Rubinstein for their support and advice.
+I also thank Aimee Chin, Steven Craig, German Cubas, Elaine Liu, Fan Wang, and the participants of the Applied Microeconomics Workshop at the University of Houston.}
+}}
 	\author{{\fontfamily{pbk}\selectfont\large{\textsc{Hussain Hadah}}\thanks{Department of Economics, Tulane University, 6823 St.
-Charles Ave., New Orleans, LA 70118, United States (e-mail: \email{hhadah@tulane.edu}).}}}
+Charles Ave., New Orleans, LA 70118, United States, (phone: +1-602-393-8077, e-mail: \email{hhadah@tulane.edu}).
+Competing interests: The author declares none.}}}
 	\date{\fontfamily{pbk}\selectfont\normalsize{\textsc{\today}}}
@@
 
-\newpage
+% \newpage
 
-\input{tables/tab03a-summary-stats-mex}
+% \input{tables/tab03a-summary-stats-mex}
 
-\newpage
+% \newpage
 
-\input{tables/tab03a-summary-stats-nonmex}
+% \input{tables/tab03a-summary-stats-nonmex}
 

### 2023-08-22 my_paper/Hadah-last-names-draft.tex
add new latex template

--- 2023-08-14 my_paper/Hadah-last-names.tex
@@
-\documentclass{wptemp}
-\usepackage{adjustbox}
-\usepackage{dcolumn}    % aligning decimals
-    \newcolumntype{d}[1]{D{.}{.}{#1}}
-% *****************************************************************
-% Estout LaTeX wrapper
-% *****************************************************************
+\documentclass[12pt,english]{article}
 
-\usepackage{lscape}
-\usepackage{threeparttablex}
-\usepackage{longtable}
+\input{fragments/preamble_packages.tex}
+\input{fragments/preamble_layout.tex}
+\input{fragments/preamble_citation.tex}
+\bibliography{bib/bibliography}
 
-%%Original code developed by Jörg Weber: see
-%% https://www.jwe.cc/2012/03/stata-latex-tables-estout/
-%% and
-%% https://www.jwe.cc/blog/
-
-\let\estinput=\input % define a new input command so that we can still flatten the document
-
-\newcommand{\estwide}[3]{
-		\vspace{.75ex}{
-			%\textsymbols% Note the added command here
-			\begin{tabular*}
-			{\textwidth}{@{\hskip\tabcolsep\extracolsep\fill}l*{#2}{#3}}
-			\toprule
-			\estinput{#1}
-			\bottomrule
-			\addlinespace[.75ex]
-			\end{tabular*}
-			}
-		}
-
-\newcommand{\estauto}[3]{
-		\vspace{.75ex}{
-			%\textsymbols% Note the added command here
-			\begin{tabular}{l*{#2}{#3}}
-			\toprule
-			\estinput{#1}
-			\bottomrule
-			\addlinespace[.75ex]
-			\end{tabular}
-			}
-		}
-
-% Allow line breaks with \\ in specialcells
-\newcommand{\specialcell}[2][c]{%
-    \begin{tabular}[#1]{@{}c@{}}#2\end{tabular}
-}
-
-\newcommand{\sym}[1]{\rlap{#1}}% Thanks David Carlisle
-
-%%%%%%%%%%% End of wrapper %%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%% TiKz %%%%%%%%%%%%%%%%%%%%%
-\usepackage{tikz}
-\usetikzlibrary{shapes.geometric, arrows}
-
-\tikzstyle{startstop} = [rectangle, rounded corners, minimum width=3cm, minimum height=1cm,text centered, draw=black, fill=red!30]
-
-\tikzstyle{io} = [trapezium, trapezium left angle=70, trapezium right angle=110, minimum width=3cm, minimum height=1cm, text centered, text width =5cm, draw=black, fill=blue!30]
-
-\tikzstyle{process} = [rectangle, minimum width=3cm, minimum height=1cm, text centered, text width =5cm, draw=black, fill=orange!30]
-
-\tikzstyle{decision} = [diamond, minimum width=3cm, minimum height=1cm, text centered, draw=black, fill=green!30]
-
-\tikzstyle{arrow} = [thick,->,>=stealth]
+\makeatother
 
@@
 
-\linespread{1}{\title{{\fontfamily{ptm}
-\selectfont{\Large{\MakeUppercase{The Impact of Hispanic Last Names and Identity on Labor Market Outcomes}}}}}}
-	\author{{\fontfamily{pbk}\selectfont\large{\textsc{Hussain Hadah}}\thanks{Department of Economics, Tulane University, 6823 St.
-Charles Ave., New Orleans, LA 70118, United States (e-mail: \email{hhadah@tulane.edu}).}}}
-	\date{\fontfamily{pbk}\selectfont\normalsize{\textsc{\today}}}
+\title{The Impact of Hispanic Last Names and Identity on Labor Market Outcomes}
+
+\author{\href{https://hussainhadah.com/}{Hussain Hadah} \thanks{Hussain Hadah: Department of Economics, Tulane University, 6823 St.
+Charles Ave., New Orleans, LA 70118, United States (e-mail: \href{mailto:hhadah@tulane.edu}{hhadah@tulane.edu}).
+I thank Professors Willa Friedman, Chinhui Juhn, Vikram Maheshri, and Yona Rubinstein for their support and advice.
+I also thank Aimee Chin, Steven Craig, German Cubas, Elaine Liu, Fan Wang, and the participants of the Applied Microeconomics Workshop at the University of Houston.}}
+
 \maketitle
@@
 \begin{center}
-\href{https://hhadah.github.io/hispanic-last-names/my_paper/Hadah-last-names.pdf}{\textcolor{green!15!black!30!blue}{\footnotesize{\textsc{Click Here to Get the Most Updated Version}}}}
+\href{https://hhadah.github.io/hispanic-last-names/my_paper/Hadah-last-names-draft.pdf}{\textcolor{green!15!black!30!blue}{\footnotesize{\textsc{Click Here to Get the Most Updated Version}}}}
 \end{center}
 
-\pagenumbering{arabic}
-
 \begin{abstract}
-In this paper, I study discrimination against Hispanics in the labor market.
+\singlespacing In this paper, I study discrimination against Hispanics in the labor market.
 I compare the children of inter-ethnic marriages to study the impact of having a Hispanic last name.
@@
 
-\keywords{Economics of Minorities, Race, and Immigrants; Discrimination and Prejudice.}
-\JEL{J71; J64}
+\noindent\textbf{JEL Classification}: J71; J64
+
+\noindent\textbf{Keywords}: Economics of Minorities, Race, and Immigrants; Discrimination and Prejudice
 
@@
 
-\pagebreak
+\vfill
+\pagebreak{}
+
 \section{Introduction}\label{sec:intro}
 
-A large literature documents substantial earnings gaps across race and ethnicity \citep{bayer2018divergent, charles2008prejudice, card1992school, fryer2004causes, rubinstein2014pride, bertrand2004emily, juhn1991accounting}.
+A large literature documents substantial earnings gaps across race and ethnicity \autocite{bayer2018divergent, charles2008prejudice, card1992school, fryer2004causes, rubinstein2014pride, bertrand2004emily, juhn1991accounting}.
 Hispanics constitute a large and growing portion of the population in the United States.
-As the number of Hispanics increases, determining whether ethnic discrimination affects their labor market outcomes becomes increasingly crucial \citep{chettyUnitedStatesStill2014, chettyEffectsExposureBetter2016,chettyFadingAmericanDream2017,abramitzkyImmigrantsAssimilateMore2020a, abramitzkyNationImmigrantsAssimilation2014,abramitzkyCulturalAssimilationAge2016,chettyWhereLandOpportunity2014}.
+As the number of Hispanics increases, determining whether ethnic discrimination affects their labor market outcomes becomes increasingly crucial \autocite{chettyUnitedStatesStill2014, chettyEffectsExposureBetter2016,chettyFadingAmericanDream2017,abramitzkyImmigrantsAssimilateMore2020a, abramitzkyNationImmigrantsAssimilation2014,abramitzkyCulturalAssimilationAge2016,chettyWhereLandOpportunity2014}.
 Thus, it is important to understand whether a person's ethnicity affects their labor market outcomes.
@@
 Moreover, I aim to show that comparing Hispanic Whites to non-Hispanic Whites might create an artificially higher earnings gap since the two groups differ on many observable characteristics.\footnote{By identifying as Hispanic, I refer to individuals who self-report their ethnicity as Hispanic on surveys or other data collection instruments.} \footnote{Observable characteristics refer to factors that can be measured and quantified, such as education level, work experience, and immigration status.} Others have attempted to compare how native-born White Hispanics fare to non-Hispanic Whites and foreign-born Hispanics.
-In \citet{antman2020ethnic,antmanEthnicAttritionObserved2016,antmanEthnicAttritionObserved2016a,antmanEthnicAttritionAssimilation2020}, the authors compare the health and educational outcomes of Hispanic Whites to non-Hispanic Whites and native-born Hispanics to foreign-born Hispanics.
+In \textcite{antman2020ethnic,antmanEthnicAttritionObserved2016,antmanEthnicAttritionObserved2016a,antmanEthnicAttritionAssimilation2020}, the authors compare the health and educational outcomes of Hispanic Whites to non-Hispanic Whites and native-born Hispanics to foreign-born Hispanics.
 They find gaps in education and health between Hispanics and Whites.
 They also find that native-born Hispanics are more likely than their foreign-born counterparts to report poor health.
-\citet{davilaChangesRelativeEarnings2008} documents many gaps in labor market outcomes between Hispanics and Whites.
+\textcite{davilaChangesRelativeEarnings2008} documents many gaps in labor market outcomes between Hispanics and Whites.
 They attribute a big part of this gap to differences in education, experience, immigration status, and regional differences.
@@
 The proportion of non-Whites has increased by more than 10 percentage points from 13 percent in 1995 to 23 percent in 2019.
-The number of Hispanics has grown by 9 percentage points from 9 percent in 1995 to 16 percent in 2019.\footnote{The portion of non-Whites and Hispanics is calculated using the Current Population Survey (CPS).} Native-born White Hispanic men earn 21\% less than White men, although a substantial portion of the earnings gap is due to educational differences between Hispanics and Whites \citep{duncan2006hispanics, duncan2018identifying, duncan2018socioeconomic}.
+The number of Hispanics has grown by 9 percentage points from 9 percent in 1995 to 16 percent in 2019.\footnote{The portion of non-Whites and Hispanics is calculated using the Current Population Survey (CPS).} Native-born White Hispanic men earn 21\% less than White men, although a substantial portion of the earnings gap is due to educational differences between Hispanics and Whites \autocite{duncan2006hispanics, duncan2018identifying, duncan2018socioeconomic}.
 Some of the earnings differences may also be due to discrimination which will have negative consequences.
@@
 One strategy used by researchers is audit or resume studies.
-\citet{bertrand2004emily} conducted an audit study where identical resumes were sent to employers with White and Black-sounding names.
+\textcite{bertrand2004emily} conducted an audit study where identical resumes were sent to employers with White and Black-sounding names.
 This approach, however, has its drawbacks.
@@
 
-This study utilizes a method developed by \citet{rubinstein2014pride}.
+This study utilizes a method developed by \textcite{rubinstein2014pride}.
 I compare children from inter-ethnic marriages.
@@
 Couples match on several observable characteristics like income, schooling, socio-economic background, etc.
-\citep{averettBetterWorseRelationship2008, averettEconomicRealityBeauty1996, beckerTheoryMarriagePart1973, beckerTheoryMarriagePart1974, beckerTreatiseFamily1993, browningCollectiveUnitaryModels2006, chiapporiFatterAttractionAnthropometric2012}.
+\autocite{averettBetterWorseRelationship2008, averettEconomicRealityBeauty1996, beckerTheoryMarriagePart1973, beckerTheoryMarriagePart1974, beckerTreatiseFamily1993, browningCollectiveUnitaryModels2006, chiapporiFatterAttractionAnthropometric2012}.
 Children of HW and WH marriages have more similar observable characteristics than children of endogamous/homogamous marriages--- i.e., White fathers-White mothers and Hispanic fathers-Hispanic mothers.
@@
 
-Previous studies have also used names as a proxy for race and ethnicity \citep{fryer2004causes, rubinstein2014pride, bertrand2004emily}.
-\citet{fryer2004causes} point out that names can be a predictor of a person's race.
+Previous studies have also used names as a proxy for race and ethnicity \autocite{fryer2004causes, rubinstein2014pride, bertrand2004emily}.
+\textcite{fryer2004causes} point out that names can be a predictor of a person's race.
 Specifically, they provide a rising pattern among Blacks having different names than Whites.
 They did, however, find that having a Black name, after controlling for the home environment at birth, does not affect their labor market outcomes.
-\citet{rubinstein2014pride} compared the children of mixed marriages between Sephardic and Ashkenazi Jews in Israel.\footnote{Ashkenazi and Sephardic are two distinct Jewish ethnic groups.} They found that workers with Sephardic last names earn substantially less than those with Ashkenazi last names.
-\citet{bertrand2004emily} conducted an audit study by sending employers identical resumes that differ in the ethnic and racial signal of a name (Black sounding name versus a White sounding one).
+\textcite{rubinstein2014pride} compared the children of mixed marriages between Sephardic and Ashkenazi Jews in Israel.\footnote{Ashkenazi and Sephardic are two distinct Jewish ethnic groups.} They found that workers with Sephardic last names earn substantially less than those with Ashkenazi last names.
+\textcite{bertrand2004emily} conducted an audit study by sending employers identical resumes that differ in the ethnic and racial signal of a name (Black sounding name versus a White sounding one).
 They found that resumes with Black-sounding names received substantially fewer callbacks than their White counterparts.
@@
 
-I use two datasets:  the Integrated Public Use Microdata Series (IPUMS) Current Population Survey (CPS) Annual Social and Economic (ASEC) \citep{cps2019} and the 1960 to 2000 US censuses \citep{acs2019}.
+I use two datasets:  the Integrated Public Use Microdata Series (IPUMS) Current Population Survey (CPS) Annual Social and Economic (ASEC) \autocite{cps2019} and the 1960 to 2000 US censuses \autocite{acs2019}.
 
@@
 Similar to the CPS, the 1960 to 2000 censuses ask about the person's place of birth and the individual's race and ethnicity.
-I employ this information to construct "synthetic parents" using a method developed by \citet{rubinstein2014pride}.
+I employ this information to construct "synthetic parents" using a method developed by \textcite{rubinstein2014pride}.
 I construct the synthetic parents by linking husbands and wives in the census data to each other.
@@
 Among HW children, 83\% of men and 82\% of women identify as Hispanic.
-These results align with the findings of a substantial body of literature that noted this attrition among children of immigrants, especially those coming from inter-ethnic marriages \citep{duncan2017complexity, duncan2018identifying, duncan2020new, antman2020ethnic}.
+These results align with the findings of a substantial body of literature that noted this attrition among children of immigrants, especially those coming from inter-ethnic marriages \autocite{duncan2017complexity, duncan2018identifying, duncan2020new, antman2020ethnic}.
 Finally, a small portion of WW children, 4\% of men and 5\% of women, identified as Hispanic.
@@
 I assume that parents have children between the ages of 25 and 40.
-Therefore, my sample consists of married White men and women with children that were born in the 1920 to 1975 cohorts\footnote{The construction of "synthetic parents" follows the method used by \citet{rubinstein2014pride}.}.
+Therefore, my sample consists of married White men and women with children that were born in the 1920 to 1975 cohorts\footnote{The construction of "synthetic parents" follows the method used by \textcite{rubinstein2014pride}.}.
 
@@
 WH and HW children are more similar in characteristics but provide employers, and the labor market, with different signals.\footnote{WH and HW children are both half White, half Hispanic.} WH children will have a non-Hispanic last name, while HW children will have a Hispanic last name.
-This is a method developed by \citet{rubinstein2014pride}.
+This is a method developed by \textcite{rubinstein2014pride}.
 
@@
 
-% \nocite{*}
 \pagebreak
-\bibliography{HispanicDiscriminationMarriageMarket}
-\newpage
+\begingroup
+\setstretch{1.0}
+%\setstretch{1.1}
+\setlength\bibitemsep{0pt}
+\printbibliography
+\endgroup
 \pagebreak
+
 \begin{appendices}
-
-\section{Figures}\label{appendix:figures}
+\section{Figures}\label{appendix:figs}
 
 % \begin{center}
-% \begin{figure}[h!]
-% \caption{Inter-racial marriages as a percent of all marriages \\
-%  in the United States from 1995 to 2020.}
-% \includegraphics[width=\textwidth]{figures/Hispanics_all.png} 
-% \label{fig:3}
+% \begin{figure}[H]
+% \caption{Histogram Before Targetting}
+% \includegraphics[width=\textwidth]{img/hist_dep_original.jpg} 
+% \label{fig:hist}
 % \end{figure}
@@
 
-% \newpage
+% \pagebreak
 
@@
 
-% \newpage
-
-% \begin{figure}[h!]
-% \centering
-% \caption{Chart explaining which synthetic parents and children and when we observe them.}
-% \begin{tikzpicture}[node distance =2cm]
-% \node (start) [startstop] {Synthetic Parents: Observed in 1960 to 2000. $25 \leq age \leq 40$, White, with kids};
-
-% \node (pro1) [process, below of = start, yshift = -2cm] {Children: Observed between 1994 and 2019. Born between 1954 to 1999 and $25 \leq age \leq 40$};
-
-% \draw [arrow] (start) -- (pro1);
-% \end{tikzpicture}
-% \label{flowchart1}
-% \end{figure}
-
 \newpage
 
-\section{tables}\label{appendix:tables}
+\section{Tables}\label{appendix:tabs}
 

### 2023-11-30 my_paper/tables/tab05-regression 2.tex
work on some of the ref comments

--- 2023-07-22 manuscripts/Hussain-Hadah/tables/tab05-regression.tex
@@
-\begin{table}[t]
+\begin{table}[H]
+
 \caption{Effect of Having Hispanic Last Name \label{tab:lastnamereg}}
+\centering
 \begin{threeparttable}
@@
 \begin{tablenotes}
-\item[1] {\footnotesize{This table includes the estimation results of equation (\ref{eq:1a}).}}
-\item[2] {\footnotesize{HW is an indicator variable that is equal to 1 if a person is the child of a Hispanic-father and White-mother.}}
-\item[3] {\footnotesize{The sample is restricted to men working full-time full-year and are wage and salary workers.}}
-\item[4] {\footnotesize{Column one has the regression results when controlling for hours worked, age, education, year and state fixed effects.
+\item[1] {\setstretch{1.0}\footnotesize{This table includes the estimation results of equation (\ref{eq:1a}).}}
+\item[2] {\setstretch{1.0}\footnotesize{HW is an indicator variable that is equal to 1 if a person is the child of a Hispanic-father and White-mother.}}
+\item[3] {\setstretch{1.0}\footnotesize{The sample is restricted to men working full-time full-year and are wage and salary workers.}}
+\item[4] {\setstretch{1.0}\footnotesize{Column one has the regression results when controlling for hours worked, age, education, year and state fixed effects.
 Column two has the results after controlling for education.}}
-\item[5] {\footnotesize{Standard errors are clustered on the state level.}}
+\item[5] {\setstretch{1.0}\footnotesize{Standard errors are clustered on the state level.}}
 \end{tablenotes}

### 2023-11-30 my_paper/tables/tab08-regression-weekearn-mex.tex
work on some of the ref comments

--- 2023-04-20 output/tables/tab05-regression.tex
@@
 
-\caption{Effect of Having Hispanic Last Name \label{tab:lastnamereg}}
+\caption{Effect of Having Hispanic Last Name: Hispanics with Mexican Ancestry  \label{tab:lastnamereg-weekearm-mex}}
 \centering
@@
 \toprule
-  & \specialcell{(1) \\ Log annual \\ earnings} & \specialcell{(2) \\ Log annual \\ earnings} & \specialcell{(3) \\ Log annual \\ earnings} & \specialcell{(4) \\  Log annual \\ earnings}\\
+  & \specialcell{(1) \\ Log weekly \\ earnings} & \specialcell{(2) \\ Log weekly \\ earnings} & \specialcell{(3) \\  Log weekly \\ earnings} & \specialcell{(4) \\  Log weekly \\ earnings}\\
 \midrule
-$HW_{ist}$ & \num{-0.06}** & \num{-0.07}** & \num{-0.05}** & \num{-0.02}\\
- & (\num{0.03}) & (\num{0.03}) & (\num{0.02}) & (\num{0.02})\\
-Constant & \num{10.40}*** & \num{9.49}*** &  & \\
- & (\num{0.04}) & (\num{0.07}) &  & \\
+$HW_{ist}$ & \num{-0.01} & \num{-0.03}* & \num{0.00} & \num{-0.01}\\
+ & (\num{0.02}) & (\num{0.02}) & (\num{0.01}) & (\num{0.01})\\
+Constant & \num{6.53}*** &  &  & \\
+ & (\num{0.04}) &  &  & \\
 \midrule
 \textit{Controlling for:} &  &  &  & \\
-Hours Worked &  & X & X & X\\
-State FE &  &  & X & X\\
-Year FE &  &  & X & X\\
-Age FE &  &  & X & X\\
-Education FE &  &  &  & X\\
-Observations & \num{2423} & \num{2423} & \num{2423} & \num{2423}\\
+State FE &  & X & X & X\\
+Year FE &  & X & X & X\\
+Age FE &  & X & X & X\\
+Education FE &  &  & X & X\\
+Parental Background &  &  &  & X\\
+Observations & \num{4535} & \num{4535} & \num{4535} & \num{4535}\\
 \bottomrule
@@
 \begin{tablenotes}
-\item[1] {\setstretch{1.0}\footnotesize{This table includes the estimation results of equation (\ref{eq:1a}).}}
+\item[1] {\setstretch{1.0}\footnotesize{This table includes the estimation results of equation (\ref{eq:1a}) where the dependent variable is log weekly earnings.}}
 \item[2] {\setstretch{1.0}\footnotesize{HW is an indicator variable that is equal to 1 if a person is the child of a Hispanic-father and White-mother.}}
-\item[3] {\setstretch{1.0}\footnotesize{The sample is restricted to men working full-time full-year and are wage and salary workers.}}
+\item[3] {\setstretch{1.0}\footnotesize{The sample is restricted to men working full-time and are wage and salary workers.}}
 \item[4] {\setstretch{1.0}\footnotesize{Column one has the regression results when controlling for hours worked, age, education, year and state fixed effects.

### 2023-11-30 output/tables/tab09-regression-weekearn-nonmex.tex
work on some of the ref comments

--- 2023-11-30 my_paper/tables/tab08-regression-weekearn-mex.tex
@@
 
-\caption{Effect of Having Hispanic Last Name: Hispanics with Mexican Ancestry  \label{tab:lastnamereg-weekearm-mex}}
+\caption{Effect of Having Hispanic Last Name: Hispanics with Non-Mexican Ancestry  \label{tab:lastnamereg-weekearm-nonmex}}
 \centering
@@
 \midrule
-$HW_{ist}$ & \num{-0.01} & \num{-0.03}* & \num{0.00} & \num{-0.01}\\
- & (\num{0.02}) & (\num{0.02}) & (\num{0.01}) & (\num{0.01})\\
-Constant & \num{6.53}*** &  &  & \\
- & (\num{0.04}) &  &  & \\
+$HW_{ist}$ & \num{-0.04} & \num{-0.04} & \num{-0.01} & \num{-0.01}\\
+ & (\num{0.03}) & (\num{0.02}) & (\num{0.02}) & (\num{0.03})\\
+Constant & \num{6.71}*** &  &  & \\
+ & (\num{0.03}) &  &  & \\
 \midrule
@@
 Parental Background &  &  &  & X\\
-Observations & \num{4535} & \num{4535} & \num{4535} & \num{4535}\\
+Observations & \num{2859} & \num{2859} & \num{2859} & \num{2859}\\
 \bottomrule

### 2023-11-30 programs/01-packages-wds.R

@@ -27,7 +27,7 @@ pacman::p_load(tictoc, parallel, pbapply, future,
                marginaleffects, gghighlight, #ggview,
                camcorder, rnaturalearth, rnaturalearthdata,
                latex2exp, janitor, extrafont)
-options("RStata.StataPath" = "/Applications/Stata/StataSE.app/Contents/MacOS/stata-se")
+options("RStata.StataPath" = "/Applications/Stata/StataMP.app/Contents/MacOS/stata-mp")
 options("RStata.StataVersion" = 17)
             
 # devtools::install_github("thomasp85/patchwork")

### 2023-11-30 programs/13c-table-three-summary-stats-spanish.R

@@ -8,22 +8,16 @@ ParentData     <- read_csv(file.path(datasets,"CPS_synth.csv")) |>
   mutate(Tot_ed = HusbandEducation + WifeEducation)
 
 # Only Mexicans and Native Born
-# ParentData <- ParentData |> 
-#        filter(BirthPlaceMom %in% c("Mexico", "USA") & BirthPlaceDad %in% c("Mexico", "USA"))
-
-# Only Non-Mexicans
-# `%notin%` <- Negate(`%in%`)
-# ParentData <- ParentData |> 
-#        filter(BirthPlaceMom %notin% c("Mexico") & BirthPlaceDad %notin% c("Mexico"))
-
+ParentData <- ParentData |> 
+       filter(BirthPlaceMom %in% c("Spain", "USA") & BirthPlaceDad %in% c("Spain", "USA"))
 
 row1 <- c("Variables", 
           "\\specialcell{White \\\\ White \\\\ (WW) \\\\ (1)}",
-          "\\specialcell{White \\\\ Hispanic \\\\ (WH) \\\\ (2)}",
-          "\\specialcell{Hispanic \\\\ White \\\\ (HW) \\\\ (3)}",
-          "\\specialcell{Hispanic \\\\ Hispanic \\\\ (HH) \\\\ (4)}",
-          "\\specialcell{HH - WW \\\\ (5)}",
-          "\\specialcell{HW - WH \\\\ (6)}"
+          "\\specialcell{White \\\\ Spanish \\\\ (WS) \\\\ (2)}",
+          "\\specialcell{Spanish \\\\ White \\\\ (SW) \\\\ (3)}",
+          "\\specialcell{Spanish \\\\ Spanish \\\\ (SS) \\\\ (4)}",
+          "\\specialcell{SS - WW \\\\ (5)}",
+          "\\specialcell{SW - WS \\\\ (6)}"
           )
 dim(row1) <- c(1,7)
 
@@ -365,7 +359,7 @@ knitr::kable(Table_rows, "html", align = "lcccccc",
              booktabs = T,
              escape = F,
        #       longtable = T, 
-             caption = "Summary statistics of synthetic parents by couple type \\label{tab:synth}") |>
+             caption = "Summary statistics of synthetic parents by couple type (Spanish) \\label{tab:synthspain}") |>
   kable_classic(full_width = F) |>
   kable_styling(bootstrap_options = c("hover", "condensed"), 
                 latex_options = c("scale_down", 
[…]

### 2023-12-06 my_paper/Hadah-last-names-draft.tex
switch places of keywords and JEL
work on some of the ref comments
addressing comments from referee

@@
 
-\noindent\textbf{JEL Classification}: J71; J64
+\noindent\textbf{Keywords}: Economics of Minorities, Race, and Immigrants; Discrimination and Prejudice
 
-\noindent\textbf{Keywords}: Economics of Minorities, Race, and Immigrants; Discrimination and Prejudice
+\noindent\textbf{JEL Classification}: J71; J64; J15
 
@@
 Column one in table \ref{tab:lastnamereg} is the results without controlling for education.
-Overall, an inter-ethnic with a Hispanic last name earns 6 percentage points less than an inter-ethnic with a White last name.
+Overall, an inter-ethnic with a Hispanic last name earns 5 percentage points less than an inter-ethnic with a White last name.
 This gap, however, could be entirely explained by educational differences.
-An inter-ethnic with a Hispanic last name earns 6 percentage points less than an inter-ethnic with a White last name, but the result is statistically insignificant.
+An inter-ethnic with a Hispanic last name earns 1 percentage points less than an inter-ethnic with a White last name, but the result is statistically insignificant.
+
+Since Hispanics are a very heterogeneous group, I conduct a heterogeneity analysis on different samples of Hispanics.
+To increase the sample size of my analysis I estimate equation \ref{eq:1a} using weekly earnings as a dependent variable.
+Weekly earnings are available in all of the monthly CPS surveys and not just the March Supplement.
+I present the results to these estimations in tables \ref{tab:lastnamereg-weekearm}-\ref{tab:lastnamereg-weekearm-cub}.
+
+First, in Table \ref{tab:lastnamereg-weekearm}, I present the results to the estimation of equation \ref{eq:1a} using weekly earnings as a dependent variable on the full sample.
+Similar to the analysis before, I find that a person with a Hispanic last name earns 4 p.p.
+less that a person without a Hispanic last name.
+This gap could be explained by Educational differences.
+Second, in Table \ref{tab:lastnamereg-weekearm-mex}, I present the results to the estimation of equation \ref{eq:1a} using weekly earnings as a dependent variable on a sample of Mexican Hispanics.
+I find that a Mexican with a Hispanic sounding last name earns 3 p.p.
+less than a Mexican with a native sounding last name.
+The gap becomes an imprecise zero after controlling for education.
+Third, in Table \ref{tab:lastnamereg-weekearm-nonmex}, I present the results to the estimation of equation \ref{eq:1a} using weekly earnings as a dependent variable on a sample of non-Mexican Hispanics.
+The gaps between non-Mexicans with Hispanic sounding last names and those with native sounding last names were all negative and statistically insignificant.
+I find that a non-Mexican with a Hispanic sounding last name earns 4 p.p.
+less than a Mexican with a native sounding last name.
+The gap becomes 1 p.p.
+after controlling for education.
+Finally, in Table \ref{tab:lastnamereg-weekearm-cub}, I present the results to the estimation of equation \ref{eq:1a} using weekly earnings as a dependent variable on a sample of Cubans.
+I find that a Cuban with Hispanic sounding last name earns 3 p.p.
+less than a Cuban with a native sounding last name but the gap was statistically insignificant.
 
@@
 
-% \newpage
+\newpage
 
-% \input{tables/tab03a-summary-stats-mex}
+\input{tables/tab03a-summary-stats-mex}
 
-% \newpage
+\newpage
 
-% \input{tables/tab03a-summary-stats-nonmex}
+\input{tables/tab03a-summary-stats-nonmex}
 
@@
 
+\input{tables/tab07-regression-weekearn}
+
+\newpage
+
+\input{tables/tab08-regression-weekearn-mex}
+
+\newpage
+
+\input{tables/tab09-regression-weekearn-nonmex}
+
+\newpage
+
+\input{tables/tab10-regression-weekearn-cuba}
+
+\newpage
+
 \input{tables/tab06-regression}

### 2023-12-06 my_paper/tables/tab07-regression-weekearn.tex
addressing comments from referee

--- 2023-11-30 my_paper/tables/tab08-regression-weekearn-mex.tex
@@
 
-\caption{Effect of Having Hispanic Last Name: Hispanics with Mexican Ancestry  \label{tab:lastnamereg-weekearm-mex}}
+\caption{Effect of Having Hispanic Last Name \label{tab:lastnamereg-weekearm}}
 \centering
@@
 \midrule
-$HW_{ist}$ & \num{-0.01} & \num{-0.03}* & \num{0.00} & \num{-0.01}\\
- & (\num{0.02}) & (\num{0.02}) & (\num{0.01}) & (\num{0.01})\\
-Constant & \num{6.53}*** &  &  & \\
- & (\num{0.04}) &  &  & \\
+$HW_{ist}$ & \num{-0.03}** & \num{-0.04}*** & \num{-0.01} & \num{-0.01}\\
+ & (\num{0.01}) & (\num{0.01}) & (\num{0.01}) & (\num{0.01})\\
+Constant & \num{6.60}*** &  &  & \\
+ & (\num{0.03}) &  &  & \\
 \midrule
@@
 Parental Background &  &  &  & X\\
-Observations & \num{4535} & \num{4535} & \num{4535} & \num{4535}\\
+Observations & \num{6592} & \num{6592} & \num{6592} & \num{6592}\\
 \bottomrule

### 2023-12-06 my_paper/tables/tab10-regression-weekearn-cuba.tex
addressing comments from referee

--- 2023-12-06 my_paper/tables/tab07-regression-weekearn.tex
@@
 
-\caption{Effect of Having Hispanic Last Name \label{tab:lastnamereg-weekearm}}
+\caption{Effect of Having Hispanic Last Name: Hispanics with Cuban Ancestry  \label{tab:lastnamereg-weekearm-cub}}
 \centering
@@
 \midrule
-$HW_{ist}$ & \num{-0.03}** & \num{-0.04}*** & \num{-0.01} & \num{-0.01}\\
- & (\num{0.01}) & (\num{0.01}) & (\num{0.01}) & (\num{0.01})\\
-Constant & \num{6.60}*** &  &  & \\
- & (\num{0.03}) &  &  & \\
+$HW_{ist}$ & \num{-0.08}** & \num{-0.03} & \num{-0.01} & \num{-0.02}\\
+ & (\num{0.04}) & (\num{0.04}) & (\num{0.04}) & (\num{0.05})\\
+Constant & \num{6.69}*** &  &  & \\
+ & (\num{0.05}) &  &  & \\
 \midrule
@@
 Parental Background &  &  &  & X\\
-Observations & \num{6592} & \num{6592} & \num{6592} & \num{6592}\\
+Observations & \num{1182} & \num{1182} & \num{1182} & \num{1182}\\
 \bottomrule

### 2023-12-06 programs/04-ACS-prep.r

@@ -1,0 +1,111 @@
+# prep ACS data to be merged with 
+# CPS data
+
+# table two
+
+# date: May 1st, 2023
+
+DATA <- read_dta(file.path(datasets,"ParentDataFull.dta")) |> 
+        mutate(BirthPlaceMom = case_when(bpld_mom  <= 12092  ~ "USA",
+           								 bpld_mom  == 11000  ~ "Puerto Rico",
+           								 bpld_mom  == 30005  ~ "Argentina",
+           								 bpld_mom  == 30010  ~ "Bolivia",
+		   								 bpld_mom  == 30015  ~ "Brazil",
+           								 bpld_mom  == 30020  ~ "Chile",
+           								 bpld_mom  == 30025  ~ "Colombia",
+           								 bpld_mom  == 26010  ~ "Dominican Republic",
+           								 bpld_mom  == 30030  ~ "Ecuador",
+           								 bpld_mom  == 21020  ~ "Costa Rica",
+           								 bpld_mom  == 21030  ~ "El Salvador",
+           								 bpld_mom  == 21040  ~ "Guatemala",
+           								 bpld_mom  == 21050  ~ "Honduras",
+           								 bpld_mom  == 25000  ~ "Cuba",
+           								 bpld_mom  == 20000  ~ "Mexico",
+           								 bpld_mom  == 21060  ~ "Nicaragua",
+           								 bpld_mom  == 21070  ~ "Panama",
+           								 bpld_mom  == 30050  ~ "Peru",
+           								 bpld_mom  == 30060  ~ "Uruguay",
+           								 bpld_mom  == 30065  ~ "Venezuela",
+           								 bpld_mom  == 43800  ~ "Spain",
+           								 bpld_mom  == 29900  ~ "Americas",
+		   								 bpld_mom  == 21000  ~ "Cental America",
+           								 bpld_mom  == 30000  ~ "South America"),
+               BirthPlaceDad = case_when(bpld_pop  <= 12092  ~ "USA",
+           								 bpld_pop  == 11000  ~ "Puerto Rico",
+           								 bpld_pop  == 30005  ~ "Argentina",
+           								 bpld_pop  == 30010  ~ "Bolivia",
+		   								 bpld_pop  == 30015  ~ "Brazil",
+           								 bpld_pop  == 30020  ~ "Chile",
+           								 bpld_pop  == 30025  ~ "Colombia",
[…]

### 2023-12-06 programs/11-table-one-parent-types.R

@@ -1,0 +1,175 @@
+# Table of observation number
+# of different types of parents
+
+# table one
+
+# date: January 19th, 2023
+
+DATA <- read_csv(file.path(datasets,"CPS_synth.csv"))
+
+# all observations
+
+ALL <- DATA %>%
+  summarise(n())
+# calculate the number 
+# of white fathers
+# white mothers
+
+WW <- DATA %>%
+  filter(Type == "White-White") %>%
+  summarise(n())
+# prettyNum(WW, big.mark = ",")
+WW_perc <- paste0(round(WW/ALL, digits = 2))
+
+# calculate the number 
+# of hispanic fathers
+# white mothers
+
+HW <- DATA %>%
+  filter(Type == "Hispanic-White") %>%
+  summarise(n())
+HW_perc <- paste0(round(HW/ALL, digits = 2))
+
+# calculate the number 
+# of white fathers
+# hispanic mothers
+
+WH <- DATA %>%
+  filter(Type == "White-Hispanic") %>%
+  summarise(n())
+WH_perc <- paste0(round(WH/ALL, digits = 2))
+
+# calculate the number 
+# of hispanic fathers
+# hispanic mothers
+
+HH <- DATA %>%
+  filter(Type == "Hispanic-Hispanic") %>%
+  summarise(n())
+HH_perc <- paste0(round(HH/ALL, digits = 2))
+
+# table
+
+first_col <- c(" ",
+                paste0("\\specialcell{", 
+                "Observations", 
+                      "\\\\",
+                "Share",
+                      "}"
+))
+
+second_col <- c("\\specialcell{White Father \\\\ White Mother}",
+               paste0("\\specialcell{", 
+                      prettyNum(WW, big.mark = ","), 
+                      "\\\\",
+                      WW_perc,
+                      "}"
+                      )
+)
+
+third_col <- c("\\specialcell{White Father \\\\ Hispanic Mother}",
+                paste0("\\specialcell{", 
+                       prettyNum(WH, big.mark = ","), 
+                       "\\\\",
+                       WH_perc,
+                       "}"
+                )
+)
+
+fourth_col <- c("\\specialcell{Hispanic Father \\\\ White Mother}",
+                paste0("\\specialcell{", 
+                       prettyNum(HW, big.mark = ","), 
+                       "\\\\",
+                       HW_perc,
+                       "}"
+                )
+)
+
+fifth_col <- c("\\specialcell{Hispanic Father \\\\ Hispanic Mother}",
+                paste0("\\specialcell{", 
[…]

### 2023-12-06 programs/13-table-three-summary-stats.R

@@ -1,0 +1,406 @@
+# summary stats
+
+# table three
+
+# date: January 19th, 2023
+
+ParentData     <- read_csv(file.path(datasets,"CPS_synth.csv")) |> 
+  mutate(Tot_ed = HusbandEducation + WifeEducation)
+
+# Only Mexicans and Native Born
+# ParentData <- ParentData |> 
+#        filter(BirthPlaceMom %in% c("Mexico", "USA") & BirthPlaceDad %in% c("Mexico", "USA"))
+
+# Only Non-Mexicans
+# `%notin%` <- Negate(`%in%`)
+# ParentData <- ParentData |> 
+#        filter(BirthPlaceMom %notin% c("Mexico") & BirthPlaceDad %notin% c("Mexico"))
+
+
+row1 <- c("Variables", 
+          "\\specialcell{White \\\\ White \\\\ (WW) \\\\ (1)}",
+          "\\specialcell{White \\\\ Hispanic \\\\ (WH) \\\\ (2)}",
+          "\\specialcell{Hispanic \\\\ White \\\\ (HW) \\\\ (3)}",
+          "\\specialcell{Hispanic \\\\ Hispanic \\\\ (HH) \\\\ (4)}",
+          "\\specialcell{HH - WW \\\\ (5)}",
+          "\\specialcell{HW - WH \\\\ (6)}"
+          )
+dim(row1) <- c(1,7)
+
+# parent's panel
+
+## Husband's Education
+HusbandEducation <- ParentData |> 
+  group_by(CoupleType) |> 
+  summarise(HusEducation   = mean(HusbandEducation, na.rm = T),
+            HusEducationSD = sd(HusbandEducation, na.rm = T))
+
+model <- lm(HusbandEducation ~ 0 + WW + WH + HW + HH, 
+    data = ParentData)
+test1 <- tidy(glht(model, linfct = c("HH - WW = 0")))
+pvalues1 <- test1$std.error
+differences1 <- test1$estimate
+
+test2 <- tidy(glht(model, linfct = c("HW - WH = 0")))
+pvalues2 <- test2$std.error
+differences2 <- test2$estimate
+
+row3 <- c(
+  "Husband's education (Total Years)",
+  paste0("\\specialcell{", 
+         sprintf("%0.2f",round(HusbandEducation[4,2], digits = 2)),  
+         "\\\\",
+         "(",
+         sprintf("%0.2f",round(HusbandEducation[4,3], digits = 2)), 
+         ")}"
+  ),
[…]

### 2023-12-06 programs/13a-table-three-summary-stats-mex.R

@@ -1,0 +1,400 @@
+# summary stats
+
+# table three
+
+# date: January 19th, 2023
+
+ParentData     <- read_csv(file.path(datasets,"CPS_synth.csv")) |> 
+  mutate(Tot_ed = HusbandEducation + WifeEducation)
+
+# Only Mexicans and Native Born
+ParentData <- ParentData |> 
+       filter(BirthPlaceMom %in% c("Mexico", "USA") & BirthPlaceDad %in% c("Mexico", "USA"))
+
+row1 <- c("Variables", 
+          "\\specialcell{White \\\\ White \\\\ (WW) \\\\ (1)}",
+          "\\specialcell{White \\\\ Hispanic \\\\ (WH) \\\\ (2)}",
+          "\\specialcell{Hispanic \\\\ White \\\\ (HW) \\\\ (3)}",
+          "\\specialcell{Hispanic \\\\ Hispanic \\\\ (HH) \\\\ (4)}",
+          "\\specialcell{HH - WW \\\\ (5)}",
+          "\\specialcell{HW - WH \\\\ (6)}"
+          )
+dim(row1) <- c(1,7)
+
+# parent's panel
+
+## Husband's Education
+HusbandEducation <- ParentData |> 
+  group_by(CoupleType) |> 
+  summarise(HusEducation   = mean(HusbandEducation, na.rm = T),
+            HusEducationSD = sd(HusbandEducation, na.rm = T))
+
+model <- lm(HusbandEducation ~ 0 + WW + WH + HW + HH, 
+    data = ParentData)
+test1 <- tidy(glht(model, linfct = c("HH - WW = 0")))
+pvalues1 <- test1$std.error
+differences1 <- test1$estimate
+
+test2 <- tidy(glht(model, linfct = c("HW - WH = 0")))
+pvalues2 <- test2$std.error
+differences2 <- test2$estimate
+
+row3 <- c(
+  "Husband's education (Total Years)",
+  paste0("\\specialcell{", 
+         sprintf("%0.2f",round(HusbandEducation[4,2], digits = 2)),  
+         "\\\\",
+         "(",
+         sprintf("%0.2f",round(HusbandEducation[4,3], digits = 2)), 
+         ")}"
+  ),
+  paste0("\\specialcell{", 
+         sprintf("%0.2f",round(HusbandEducation[3,2], digits = 2)),  
+         "\\\\",
+         "(",
+         sprintf("%0.2f",round(HusbandEducation[3,3], digits = 2)), 
[…]

### 2023-12-06 programs/13b-table-three-summary-stats-nonmex.R

@@ -1,0 +1,402 @@
+# summary stats
+
+# table three
+
+# date: January 19th, 2023
+
+ParentData     <- read_csv(file.path(datasets,"CPS_synth.csv")) |> 
+  mutate(Tot_ed = HusbandEducation + WifeEducation)
+
+# Only Non-Mexicans
+`%notin%` <- Negate(`%in%`)
+ParentData <- ParentData |> 
+       filter(BirthPlaceMom %notin% c("Mexico") & BirthPlaceDad %notin% c("Mexico"))
+
+
+row1 <- c("Variables", 
+          "\\specialcell{White \\\\ White \\\\ (WW) \\\\ (1)}",
+          "\\specialcell{White \\\\ Hispanic \\\\ (WH) \\\\ (2)}",
+          "\\specialcell{Hispanic \\\\ White \\\\ (HW) \\\\ (3)}",
+          "\\specialcell{Hispanic \\\\ Hispanic \\\\ (HH) \\\\ (4)}",
+          "\\specialcell{HH - WW \\\\ (5)}",
+          "\\specialcell{HW - WH \\\\ (6)}"
+          )
+dim(row1) <- c(1,7)
+
+# parent's panel
+
+## Husband's Education
+HusbandEducation <- ParentData |> 
+  group_by(CoupleType) |> 
+  summarise(HusEducation   = mean(HusbandEducation, na.rm = T),
+            HusEducationSD = sd(HusbandEducation, na.rm = T))
+
+model <- lm(HusbandEducation ~ 0 + WW + WH + HW + HH, 
+    data = ParentData)
+test1 <- tidy(glht(model, linfct = c("HH - WW = 0")))
+pvalues1 <- test1$std.error
+differences1 <- test1$estimate
+
+test2 <- tidy(glht(model, linfct = c("HW - WH = 0")))
+pvalues2 <- test2$std.error
+differences2 <- test2$estimate
+
+row3 <- c(
+  "Husband's education (Total Years)",
+  paste0("\\specialcell{", 
+         sprintf("%0.2f",round(HusbandEducation[4,2], digits = 2)),  
+         "\\\\",
+         "(",
+         sprintf("%0.2f",round(HusbandEducation[4,3], digits = 2)), 
+         ")}"
+  ),
+  paste0("\\specialcell{", 
+         sprintf("%0.2f",round(HusbandEducation[3,2], digits = 2)),  
+         "\\\\",
+         "(",
[…]

### 2023-12-06 programs/13c-table-three-summary-stats-spanish.R

@@ -1,0 +1,400 @@
+# summary stats
+
+# table three
+
+# date: January 19th, 2023
+
+ParentData     <- read_csv(file.path(datasets,"CPS_synth.csv")) |> 
+  mutate(Tot_ed = HusbandEducation + WifeEducation)
+
+# Only Mexicans and Native Born
+ParentData <- ParentData |> 
+       filter(BirthPlaceMom %in% c("Spain", "USA") & BirthPlaceDad %in% c("Spain", "USA"))
+
+row1 <- c("Variables", 
+          "\\specialcell{White \\\\ White \\\\ (WW) \\\\ (1)}",
+          "\\specialcell{White \\\\ Spanish \\\\ (WS) \\\\ (2)}",
+          "\\specialcell{Spanish \\\\ White \\\\ (SW) \\\\ (3)}",
+          "\\specialcell{Spanish \\\\ Spanish \\\\ (SS) \\\\ (4)}",
+          "\\specialcell{SS - WW \\\\ (5)}",
+          "\\specialcell{SW - WS \\\\ (6)}"
+          )
+dim(row1) <- c(1,7)
+
+# parent's panel
+
+## Husband's Education
+HusbandEducation <- ParentData |> 
+  group_by(CoupleType) |> 
+  summarise(HusEducation   = mean(HusbandEducation, na.rm = T),
+            HusEducationSD = sd(HusbandEducation, na.rm = T))
+
+model <- lm(HusbandEducation ~ 0 + WW + WH + HW + HH, 
+    data = ParentData)
+test1 <- tidy(glht(model, linfct = c("HH - WW = 0")))
+pvalues1 <- test1$std.error
+differences1 <- test1$estimate
+
+test2 <- tidy(glht(model, linfct = c("HW - WH = 0")))
+pvalues2 <- test2$std.error
+differences2 <- test2$estimate
+
+row3 <- c(
+  "Husband's education (Total Years)",
+  paste0("\\specialcell{", 
+         sprintf("%0.2f",round(HusbandEducation[4,2], digits = 2)),  
+         "\\\\",
+         "(",
+         sprintf("%0.2f",round(HusbandEducation[4,3], digits = 2)), 
+         ")}"
+  ),
+  paste0("\\specialcell{", 
+         sprintf("%0.2f",round(HusbandEducation[3,2], digits = 2)),  
+         "\\\\",
+         "(",
+         sprintf("%0.2f",round(HusbandEducation[3,3], digits = 2)), 
[…]

### 2023-12-06 programs/14-table-three-summary-stats.R

@@ -1,0 +1,396 @@
+# summary stats
+
+# table three
+
+# date: January 19th, 2023
+
+ParentData     <- read_csv(file.path(datasets,"CPS_synth.csv")) |> 
+  mutate(Tot_ed = HusbandEducation + WifeEducation)
+
+row1 <- c("Variables", 
+          "\\specialcell{White \\\\ White \\\\ (WW) \\\\ (1)}",
+          "\\specialcell{White \\\\ Hispanic \\\\ (WH) \\\\ (2)}",
+          "\\specialcell{Hispanic \\\\ White \\\\ (HW) \\\\ (3)}",
+          "\\specialcell{Hispanic \\\\ Hispanic \\\\ (HH) \\\\ (4)}",
+          "\\specialcell{HH - WW \\\\ (5)}",
+          "\\specialcell{HW - WH \\\\ (6)}"
+          )
+dim(row1) <- c(1,7)
+
+# parent's panel
+
+## Husband's Education
+HusbandEducation <- ParentData |> 
+  group_by(CoupleType) |> 
+  summarise(HusEducation   = mean(HusbandEducation, na.rm = T),
+            HusEducationSD = sd(HusbandEducation, na.rm = T))
+
+model <- lm(HusbandEducation ~ 0 + WW + WH + HW + HH, 
+    data = ParentData)
+test1 <- tidy(glht(model, linfct = c("HH - WW = 0")))
+pvalues1 <- test1$std.error
+differences1 <- test1$estimate
+
+test2 <- tidy(glht(model, linfct = c("HW - WH = 0")))
+pvalues2 <- test2$std.error
+differences2 <- test2$estimate
+
+row3 <- c(
+  "Husband's education (Total Years)",
+  paste0("\\specialcell{", 
+         sprintf("%0.2f",round(HusbandEducation[4,2], digits = 2)),  
+         "\\\\",
+         "(",
+         sprintf("%0.2f",round(HusbandEducation[4,3], digits = 2)), 
+         ")}"
+  ),
+  paste0("\\specialcell{", 
+         sprintf("%0.2f",round(HusbandEducation[3,2], digits = 2)),  
+         "\\\\",
+         "(",
+         sprintf("%0.2f",round(HusbandEducation[3,3], digits = 2)), 
+         ")}"
+  ),
+  paste0("\\specialcell{", 
+         sprintf("%0.2f",round(HusbandEducation[2,2], digits = 2)),  
+         "\\\\",
+         "(",
[…]

### 2023-12-06 programs/15-table-four-summary-stats.R

@@ -1,0 +1,603 @@
+# summary stats
+
+# table three
+
+# library(Hmisc)
+
+# date: January 19th, 2023
+
+IndividualData <- read_csv(file.path(datasets,"CPS_synth.csv")) |> 
+  mutate(Education = as.numeric(Education)) |> 
+  filter(WW == 1 | (WH == 1 & Hispanic_ID == 1) | (HW == 1 & Hispanic_ID == 1) | (HH == 1 & Hispanic_ID == 1))
+
+CrossTable(IndividualData$Type, IndividualData$Hispanic_ID)
+row1 <- c("Variables", 
+          "\\specialcell{White \\\\ White \\\\ (WW) \\\\ (1)}",
+          "\\specialcell{White \\\\ Hispanic \\\\ (WH) \\\\ (2)}",
+          "\\specialcell{Hispanic \\\\ White \\\\ (HW) \\\\ (3)}",
+          "\\specialcell{Hispanic \\\\ Hispanic \\\\ (HH) \\\\ (4)}",
+          "\\specialcell{HH - WW \\\\ (5)}",
+          "\\specialcell{HW - WH \\\\ (6)}"
+          )
+dim(row1) <- c(1,7)
+
+# Panel A
+# Age and Education
+row6 <- c("Panel A: Children's Education ", " ", " ", " ", " ", " ", " ")
+dim(row6) <- c(1,7)
+
+## Man's Education
+ManEducation <- IndividualData |> 
+  filter(sex == 1) |> 
+  group_by(Type) |> 
+  summarise(EducationMean   = mean(Education, na.rm = T),
+            EducationSD = sd(Education, na.rm = T))
+
+model <- lm(Education ~ 0 + WW + WH + HW + HH, 
+             data = IndividualData |> filter(sex == 1))
+test1 <- tidy(glht(model, linfct = c("HH - WW = 0")))
+pvalues1 <- test1$std.error
+differences1 <- test1$estimate
+
+test2 <- tidy(glht(model, linfct = c("HW - WH = 0")))
+pvalues2 <- test2$std.error
+differences2 <- test2$estimate
+
+row7 <- c(
+  "Men’s education (Total Years)",
+  paste0("\\specialcell{", 
+         round(ManEducation[4,2], digits = 2), 
+         "\\\\",
+         "(",
+         round(ManEducation[4,3], digits = 2),
+         ")}"
+  ),
+  paste0("\\specialcell{", 
+         round(ManEducation[3,2], digits = 2), 
+         "\\\\",
+         "(",
[…]

### 2023-12-06 programs/16-table-five-regression.R

@@ -1,0 +1,184 @@
+# table Five
+# Regression
+
+# date: January 19th, 2023
+
+# open data, filter to only have the relevant data
+# and impute empty parental charachtersitcs with mean
+IndividualData <- read_csv(file.path(datasets,"CPS_synth.csv")) |> 
+  mutate(Education = as.numeric(Education)) |> 
+  filter((WW == 1 & Hispanic_ID == 0) | WH == 1 | HW == 1 | HH == 1) |> 
+  mutate(HusbandEducation = case_when(!is.na(HusbandEducation) ~ HusbandEducation,
+                                       is.na(HusbandEducation) ~ mean(HusbandEducation, na.rm = TRUE)),
+         WifeEducation = case_when(!is.na(WifeEducation) ~ WifeEducation,
+                                    is.na(WifeEducation) ~ mean(WifeEducation, na.rm = TRUE)),
+         Husband_ftotval = case_when(!is.na(Husband_ftotval) ~ Husband_ftotval,
+                                       is.na(Husband_ftotval) ~ mean(Husband_ftotval, na.rm = TRUE)),
+                                )
+# create a total education variable
+# and a income quintile variable
+# and education quintile variable
+IndividualData  <- IndividualData  |> 
+  mutate(TotalHHEducation = HusbandEducation + WifeEducation,
+         IncomeQuint      = cut(Husband_ftotval, breaks = quantile(Husband_ftotval, probs = c(0, 0.2, 0.4, 0.6, 0.8, 1), na.rm = TRUE), include.lowest = TRUE)
+          )
+IndividualData  <- IndividualData  |> 
+  mutate(EducationQuint      = cut(TotalHHEducation, breaks = quantile(TotalHHEducation, probs = c(0, 0.2, 0.4, 0.6, 0.8, 1), na.rm = TRUE), include.lowest = TRUE)
+          )
+ParentDummies <- c("HW", "uhrsworkly")
+
+ParentControls  <- c("IncomeQuint", "EducationQuint")
+# By generation
+reg1 <- list(
+  "\\specialcell{(1) \\\\ Log annual \\\\ earnings}" = feols(lninctot_1999 ~ HW, vcov = ~statefip,
+                                                            data = IndividualData |> 
+                                                          filter(sex == 1 & FTFY == 1 & Self_employed_ASEC == 0 & (HW == 1 | WH == 1))),
[…]

### 2023-12-06 programs/17-table-six-regression.R

@@ -1,0 +1,222 @@
+# summary stats
+
+# table Five
+# Regression
+
+# date: January 19th, 2023
+
+# open data, filter to only have the relevant data
+# and impute empty parental charachtersitcs with mean
+
+IndividualData <- read_csv(file.path(datasets,"CPS_synth.csv")) |> 
+  mutate(Education = as.numeric(Education)) |> 
+  filter((WW == 1 & Hispanic_ID == 0) | WH == 1 | HW == 1 | HH == 1)|> 
+  mutate(HusbandEducation = case_when(!is.na(HusbandEducation) ~ HusbandEducation,
+                                       is.na(HusbandEducation) ~ mean(HusbandEducation, na.rm = TRUE)),
+         WifeEducation = case_when(!is.na(WifeEducation) ~ WifeEducation,
+                                    is.na(WifeEducation) ~ mean(WifeEducation, na.rm = TRUE)),
+          Husband_ftotval = case_when(!is.na(Husband_ftotval) ~ Husband_ftotval,
+                                       is.na(Husband_ftotval) ~ mean(Husband_ftotval, na.rm = TRUE)))
+
+# create a total education variable
+# and a income quintile variable
+# and education quintile variable
+IndividualData  <- IndividualData  |> 
+  mutate(TotalHHEducation = HusbandEducation + WifeEducation,
+         IncomeQuint      = cut(Husband_ftotval, breaks = quantile(Husband_ftotval, probs = c(0, 0.2, 0.4, 0.6, 0.8, 1), na.rm = TRUE), include.lowest = TRUE)
+          )
+IndividualData  <- IndividualData  |> 
+  mutate(EducationQuint      = cut(TotalHHEducation, breaks = quantile(TotalHHEducation, probs = c(0, 0.2, 0.4, 0.6, 0.8, 1), na.rm = TRUE), include.lowest = TRUE)
+          )
+
+ParentDummies <- c(#"WH:Hispanic_ID", 
+                   "HW:Hispanic_ID", 
+                   "Hispanic_ID",
+                   #"HH:Hispanic_ID",
+                   #"WH", 
+                   "HW", 
+                   #"HH", 
+                   "uhrsworkly")
+ParentControls  <- c("EducationQuint", "IncomeQuint")
+
+IndividualData |> 
+  filter(sex == 1 & FTFY == 1 & Self_employed == 0 & (HW == 1 | WH == 1)) |> 
[…]

### 2023-12-06 programs/18-table-seven-regression.R

@@ -1,0 +1,112 @@
+# summary stats
+
+# table Five
+# Regression
+
+# date: January 19th, 2023
+
+IndividualData <- read_dta(file.path(datasets,"BySexAnalysisData.dta")) |> 
+  mutate(Education = as.numeric(Education))
+
+# By generation
+reg1 <- list(
+  "\\specialcell{(1) \\\\ Log annual earnings}" = feols(lninctot_1999 ~ Hispanic_ID+  uhrsworkly | year + age+ statefip, vcov = ~statefip,
+                                                        data = IndividualData |> 
+                                                          filter(sex == 1 & FTFY == 1 & Self_employed == 0)),
+  "\\specialcell{(2) \\\\  Log annual earnings}" = feols(lninctot_1999 ~ Hispanic_ID+  uhrsworkly | year + age + educ+ statefip, vcov = ~statefip,
+                                                         data = IndividualData |> 
+                                                           filter(sex == 1 & FTFY == 1 & Self_employed == 0))
+  
+)
+
+
+controling_for <-  c("\\textit{Controlling for:}", " ", "")
+dim(controling_for) <- c(1,3)
+
+hoursworked <-  c("Hours Worked", "X", "X")
+dim(hoursworked) <- c(1,3)
+
+controling_for <- as.data.frame(controling_for)
+hoursworked <- as.data.frame(hoursworked)
+
+all_row <- rbind(controling_for, hoursworked)
+attr(all_row, 'position') <- c(4:5)
+
+cm <- c("Hispanic_ID" = "$Hispanic_{i}$"
+) 
+gm <- tibble::tribble(
+  ~raw,        ~clean,          ~fmt,
+  "nobs",      "Observations",             0,
+  "FE: year", "Year FE", 0,
+  "FE: statefip", "State FE", 0,
+  "FE: age", "Age", 0,
+  "FE: educ", "Education", 0,
+  "std.error.type", "Standard Errors", 0,
+  #"r.squared", "R squared", 3
+)
+
+modelsummary(reg1, fmt = 2,  
+             coef_map = cm,
+             add_rows = all_row,
+             gof_map = gm,
+             escape = F,
+             #gof_omit = 'DF|Deviance|R2|AIC|BIC|Log.Lik.|F|Std.Errors',
+             stars= c('***' = 0.01, '**' = 0.05, '*' = 0.1),
[…]

### 2023-12-06 programs/95-make-all.R

@@ -1,0 +1,55 @@
+#######################################################################
+# Master script
+#######################################################################
+
+## Clear out existing environment
+gc()
+rm(list = ls()) 
+
+mdir <- "~/Dropbox/Research/My Research Data and Ideas/hispanic-last-names"
+datasets <- paste0(mdir,"/data/datasets")
+raw <- paste0(mdir,"/data/raw")
+
+mdir_git     <- "~/Documents/GiT/hispanic-last-names"
+datasets_git <- paste0(mdir_git,"/data/datasets") 
+raw_git      <- paste0(mdir_git,"/data/raw") 
+tables_wd    <- paste0(mdir_git,"/output/tables") 
+figures_wd   <- paste0(mdir_git,"/output/figures") 
+programs     <- paste0(mdir_git,"/programs") 
+thesis_tabs  <- paste0(mdir_git,"/my_paper/tables") 
+thesis_plots <- paste0(mdir_git,"/my_paper/figure") 
+add_health_data_dir <- "/Users/hhadah/Dropbox/Research/My Research Data and Ideas/Identification_Paper/Data/Datasets"
+ACS_data  <- "/Users/hhadah/Documents/GiT/jmp-decomposition-replication/data/raw"
+
+# pres_tabs <- paste0(mdir,"/presentations/jmp-presentation/tables")
+# pres_plots <- paste0(mdir,"/presentations/jmp-presentation/figures")
+
+# dissertation_wd <- paste0("/Users/hhadah/Documents/GiT/my_thesis/presentation/dissertation")
+### run do files and scripts
+
+# main scripts
+source(file.path(programs,"01-packages-wds.r")) # set up package
+stata(file.path(programs,"02-ACS-Synt-Parents-DoFile.do")) # run ACS do file
+stata(file.path(programs,"03-CleaningDataDoFile.do")) # clean data
+source(file.path(programs,"04-ACS-prep.r")) # prep data
+stata(file.path(programs,"05-CPS-Do-File.do")) # CPS Do File
+stata(file.path(programs,"06-CPS-DataCleanUpScript.do")) # CPS data cleaning
+# stata(file.path(programs,"07-BySexAnalysis.do")) # by sex analysis
+# stata(file.path(programs,"08-CoupleAnalysis.do")) # couple analysis
+# stata(file.path(programs,"09-Regressions.do")) # regressions
+source(file.path(programs,"10-merge-cps-acs.r")) # merge cps and acs
+source(file.path(programs,"11-table-one-parent-types.R")) # table one
[…]

### 2023-12-11 my_paper/Hadah-last-names-draft.tex
edits to abstract and intro

@@
 \begin{abstract}
-\singlespacing In this paper, I study discrimination against Hispanics in the labor market.
-I compare the children of inter-ethnic marriages to study the impact of having a Hispanic last name.
-While males born to Hispanic father-White mothers earn 6 percentage points less than those born to White father-Hispanic mothers, the gap could be completely explained by educational differences.
-I also study the effect of identifying as Hispanic on earnings.
-I find that men with a Spanish-sounding last name who identify as Hispanic earn significantly less than those who do not identify as Hispanic, but the gap could also be explained by educational differences.
+\singlespacing Do people with ethnically Hispanic names face discrimination in the labor market?
+In this study, I analyze the causal impact of Hispanic-sounding surnames on wages, using the unique context of inter-ethnic marriages and the distinct characteristics of White and Hispanic surnames.
+We compare the wages of individuals with one White and one Hispanic parent, observing that those with Hispanic surnames often encounter wage disparities.
+Our findings show a significant wage difference favoring individuals with White surnames.
+Although males born to Hispanic fathers and White mothers earn 5 percentage points less than those born to White fathers and Hispanic mothers, this gap can be completely attributed to differences in education levels.
+However, this should not be interpreted as an absence of discrimination; rather, it potentially reflects discrimination in human capital accumulation.
+Additionally, I explore the impact of self-identifying as Hispanic on earnings.
+Men with a Spanish-sounding last name who identify as Hispanic earn significantly less than those who do not, but again, this gap can be largely explained by educational disparities.
+This suggests that discrimination may extend beyond the labor market, influencing educational opportunities and outcomes as well.
 
@@
 
-Identifying discrimination is difficult because of factors that affect labor market outcomes that are unobservable to economists---such as unobserved skills.
+Identifying discrimination is difficult because of factors that affect labor market outcomes that are unobservable to economists---such as unobserved skills and separating it from prejudice and stereotypes.
 One strategy used by researchers is audit or resume studies.
@@
 More precisely, I compare children of Hispanic fathers and White mothers (henceforth HW) to children of White fathers and Hispanic mothers (henceforth  WH).
-This approach stems from the fact that marriages are not random.
+This approach stems from the fact that there is a strong selection among many characteristics, and thus, marriages are not random.
 Couples match on several observable characteristics like income, schooling, socio-economic background, etc.

### 2023-12-24 my_paper/Hadah-last-names-draft.tex
Finish writing a new version

@@
 Men with a Spanish-sounding last name who identify as Hispanic earn significantly less than those who do not, but again, this gap can be largely explained by educational disparities.
-This suggests that discrimination may extend beyond the labor market, influencing educational opportunities and outcomes as well.
+This suggests that discrimination may extend beyond the labor market, influencing educational opportunities and outcomes too.
 
@@
 
-Using the Current Population Survey (CPS) from 1994 to 2019 and the 1960 to 2000 US censuses, I estimate the effect of having a Hispanic last name on labor market outcomes.
-In other words, I approximate the variation in labor market perception of ethnic signals, in this case having a Hispanic last name, on annual log earnings.
-I also study the effect of identifying as Hispanic on labor market outcomes.
-
-I find that while men born to Hispanic father-White mothers earn less than men born to White father-Hispanic mothers, this gap is entirely explained by educational differences.
-Thus, I do not find a significant effect of having a Hispanic last name.
-Finally, I find that men identifying as Hispanic earn significantly less than those who do not, even when I control for ancestry and education.
-
-% Inter-ethnic children with a White father, and thus having a non-Hispanic last name, earn 14\% less than children of endogamous White parents.
-Inter-ethnic children with a Hispanic father, and thus having a Hispanic last name, earn 20\% less than children of endogamous White parents.
-Endogamous children with Hispanic parents earn 24\% less than children of endogamous White parents.
-A big part of the gaps could be explained by education.
-Inter-ethnic children with a White father and identify as Hispanic, and thus having a non-Hispanic last name, earn 14\% less than children of endogamous White parents.
-Inter-ethnic children with a Hispanic father and identify as Hispanic, and thus having a Hispanic last name, earn 18\% less than children of endogamous White parents.
-Endogamous children with Hispanic parents and identify as Hispanic earn 10\% less than children of endogamous White parents.
+The main identifying assumption of my empirical strategy depends on the assumption that people born to Hispanic fathers and White mothers are similar to their White fathers Hispanic mothers peer in all observable aspects and characteristics that are important in the labor market.
+Consequently, the only difference between the two groups is the variation in which group is more likely to have a Hispanic sounding last name.
+Children from mixed ethnic backgrounds may appear physically similar to those from single ethnicity backgrounds, but the influence of family dynamics and upbringing, crucial in developing skills and personal characteristics, varies with the pattern of mixed ethnic marriages.
+Particularly in a society where Hispanic ancestry is perceived negatively, it raises the question of what kind of White women would choose Hispanic men.
+Furthermore, even if such unions were formed randomly, children from White-Hispanic homes might benefit from more favorable family conditions than those from Hispanic-White homes, considering Whites generally have stronger socio-economic backgrounds than Hispanics.
+These factors introduce doubts regarding whether children from Hispanic-White families receive comparable familial support and influences, either genetically or environmentally, as those from White-Hispanic families.
 
@@
 They found that resumes with Black-sounding names received substantially fewer callbacks than their White counterparts.
+
+Using the Current Population Survey (CPS) from 1994 to 2019 and the 1960 to 2000 US censuses, I estimate the effect of having a Hispanic last name on labor market outcomes.
+In other words, I approximate the variation in labor market perception of ethnic signals, in this case having a Hispanic last name, on annual log earnings.
+I also study the effect of identifying as Hispanic on labor market outcomes.
+
+I find that while men born to Hispanic father-White mothers earn less than men born to White father-Hispanic mothers, this gap is entirely explained by educational differences.
+I find that inter-ethnic children with Hispanic sounding last name earn 5 percentage points less than those without a Hispanic last name.
+The gap becomes statistically insignificant after controlling for education.
+Thus, I do not find a significant effect of having a Hispanic last name.
+Finally, I find that men identifying as Hispanic earn significantly less than those who do not, even when I control for ancestry and education.
 
@@
 I assume that parents have children between the ages of 25 and 40.
-I then link these synthetic parents using the birth year of child, and place of birth of parents to the year of birth of the "children" and the palce of birth of parents in the CPS sample.
+I then link these synthetic parents using the birth year of child, and parents' places of birth to the year of birth of the "children" and the parents' places of birth in the CPS sample.
 
@@
 
-The distribution of the four types of children is presented in figure \ref{fig:dist} and table \ref{tab:mat1}.
-The majority of the sample (95.96\%) is WW children.
-The second biggest group is HH, which constitutes 2.69\% of the sample.
+The distribution of the four types of children is presented Table \ref{tab:mat1}.
+The majority of the sample (96\%) is WW children.
+The second biggest group is HH, which constitutes 3\% of the sample.
 Inter-ethnic children, WH and HW, make up 1.35\% of the sample with 90,325 observations.
 Even though WH and HW are only 1.35\% of the sample, I have plenty of observations to carry out an analysis.
-The summary statistics for the children of the four types of marriages are presented in table \ref{tab:c&p1}.
+The summary statistics for the children of the four types of marriages are presented in Table \ref{tab:c&p1}.
 Children of WW marriages (column 1) do better on every measure while children of HH parents (column 4) do worse than other children on every measure.
+Children of WH (column 2) and HW (column 3) marriages fall in between WW and HH children.
+Moreover, the summary statistics for those that also identify as Hispanic is reported in in Table \ref{tab:c&p2}.
 
-On the one hand, WW children are more educated, with approximately 14 years of education for men and women.
+% On the one hand, WW children are more educated, with approximately 14 years of education for men and women.
 WW children have an unemployment rate of 4\% for men and women.
@@
 
-On the other hand, children of HH marriages have an average of 12.9 and 13.24 years of education for men and women, respectively.
+% On the other hand, children of HH marriages have an average of 12.9 and 13.24 years of education for men and women, respectively.
 HH men and women have an unemployment rate of 7\% and 6\%.
@@
 
-Inter-ethnic children, WH and HW, have lower educational attainment and income than WW children, but they are better than HH children.
+% Inter-ethnic children, WH and HW, have lower educational attainment and income than WW children, but they are better than HH children.
 Also, WH and HW are comparable, unlike WW and HH children.
@@
 At the same time, the WH children have White fathers with a non-Hispanic, or White, last names.
-Summary statistics of these two groups, table \ref{tab:c&p1} columns (2) and (3) can show that they are different from the children of non-inter-ethnic marriages.
+Summary statistics of these two groups, Table \ref{tab:c&p1} columns (2) and (3) can show that they are different from the children of non-inter-ethnic marriages.
 
-WH children have an average of 13.58 and 13.79 years of education for men and women.
+% WH children have an average of 13.58 and 13.79 years of education for men and women.
 The unemployment rate for men and women is 5\%.
@@
 
-In panel C of table \ref{tab:c&p1}, I present the rates at which members of each group self-identify as Hispanic.
+% In panel C of Table \ref{tab:c&p1}, I present the rates at which members of each group self-identify as Hispanic.
 \footnote{Hispanic self-identification is observed when an individual notes that they are Hispanic on the CPS questionnaire} The vast majority of HH children, 96\% for men and 97\% for women, identify as Hispanic.
@@
 
-In columns (5) and (6), I calculate the differences in means between HH-WW (column 5) and HW-WH (column 6).
+% In columns (5) and (6), I calculate the differences in means between HH-WW (column 5) and HW-WH (column 6).
 Overall, HH children do worse than WW children.
@@
 
-The differentials between WH and HW are not as severe as the ones between HH and WW.
+% The differentials between WH and HW are not as severe as the ones between HH and WW.
 HW children are still worse off than WH but are much better than HH.
@@
 
-In table \ref{tab:c&p2}, I present the summary statistics for the four types and limit the sample to those that only identify as Hispanic.\footnote{Limiting the sample to those who identify as Hispanic only affects WH, HW, and HH.
+% In Table \ref{tab:c&p2}, I present the summary statistics for the four types and limit the sample to those that only identify as Hispanic.\footnote{Limiting the sample to those who identify as Hispanic only affects WH, HW, and HH.
 Members of the WW group are not affected by this restriction.} The same pattern discussed above between HH--WW and HW--WH persists.
@@
 
-I show the distribution of the four types of couples in table \ref{tab:mat2}.
+I show the distribution of the four types of couples in Table \ref{tab:mat2}.
 White husbands and White wives (WW) make up the majority of couples in the sample, 96\% (5,141,737 couples).
@@
 White husbands and Hispanic wives (WH) couples are less 1\% (33,097 couples) of the sample and Hispanic husbands and White wives (HW) are also 1\% (37,847 couples).
-
-I present the summary statistics of the four types of synthetic parents in table \ref{tab:synth}.
-WW couples have higher education, 12.75 years for husbands and 12.47 for wives.
-As a household, WW couples have 25.22 years of schooling.
-Additionally, the log total family income for a WW household is 10.72 and have 3.77 children.
-Husbands in HH marriages have 8.64 years of education, while women have 8.49 years of schooling.
-As a household, HH couples have 17.13 years of education, log total family income of 10.27, and have 4.23 children.
-
-WH husbands have 11.77 years of education, while wives have an average of 10.40 years.
-WH household attained a total of 22.17 years of schooling in total, a log total family income of 10.55, and 3.98 children.
-HW husbands and wives obtained 10.25 and 11.11 years of education, respectively.
-HW household attained a total of 21.36 years of schooling in total, a log total family income of 10.46, and 4.15 children.
-
-In columns 5 and 6 of table \ref{tab:synth}, I report the differences in means between HH-WW households and HW-WH households.
-An HH Hispanic husband attained 4.11 fewer years of education than a WW husband.
-While an HH wife acquired 3.98 fewer years of schooling than a White WW wife.
-The total educational difference between HH and WW households equals 8.09 years of education.
-The educational differences between HW and WH are less severe than those between HH and WW.
-A Hispanic husband in intermarriage attained 1.52 fewer years of schooling than a White husband in intermarriage.
-A White wife in intermarriage attained 0.70 years of education, more than a Hispanic wife in intermarriage.
-The difference in total years of education between HW and WH households is 0.81, which is an 90\% decrease from the HH--WW gap.
-The total family income of an HH household is 45\% lower than a WW household, while the total family income of an HW household is 10\% lower than a WH household.
-These differences show that couples that intermarry are better and different than those that do not.
-
-%In table \ref{tab:p1} columns five and six, I provide the  HH-WW and HW-WH gaps.
-WW couples are more educated and earn more than HH couples.
-Also, WH husbands are more educated and earn more than HW husbands.
-In comparison, WH wives are more educated than HW wives.
-Even though White husbands in intermarriages are more educated and earn more than husbands in HW marriages, they are more comparable to each other than the non-intermarriage couples.
+I present the summary statistics of the parents in Table \ref{tab:synth}.
 
@@
 Let $Y_{ist}$ be the annual log earnings of person $i$ in state $s$ at time $t$.
-$HW_{ist}$ is an indicator variables for the type of parents person $i$ has.
+$HW_{ist}$ is an indicator variables for the type of parents of person $i$ has.
 $X_{ist}$ is a vector of controls that includes age and numbers of hours worked, $\gamma_{t}$ are year fixed effects, $\lambda_{s}$ are state fixed effects, and $\phi_{ist}$ represents the error term.
@@
 
+\subsection{Threats to Identification}
+
+The central assumption underpinning my estimation strategy is based on the hypothesis that individuals born to a Hispanic father and White mother exhibit comparable characteristics to their peers of White-Hispanic descent, especially in areas vital to the labor market.
+This assumption includes similarities in educational background, skill sets, and work experiences that are significant determinants in employment opportunities, salary levels, and career advancement.
+
+Two primary reasons underscore that this assumption.
+First, there is a significant selection in the marriage market.
+Since belonging to the Hispanic out-group comes with a negative penalty in society, one might wonder who are the White women that would be willing to give their kids a Hispanic last-name that would potentially negatively influence their futures.
+Second, fathers and mothers influence human capital accumulation differently \autocite{kimball2009risk,magruder2010intergenerational}.
+If marriage was random, then WH households might be a better environment for children that HA.
+Using the CPS and the US Census, I will evaluate these empirical concerns.
+
+The previous two threats to identification could be addressed by the selection of partners among inter-ethnic couples.
+The average White person exhibits better observable characteristics than the Average Hispanic.
+Consequently, if marriages were random, the White husband in an WH marriage would have better traits than wife, and the White wife in the HW marriage would have the stronger traits.
+The random matching coupled with the fact that parents are not perfect substitutes in the inter-generational transmission of human capital would render the identification invalid since the parental background will affect labor market outcomes.
+
+Marriages, however, are not random.
+There is large body of empirical and theoretical work that indicates that marriages exhibits strong selection, or what is referred to as assortative matching, on traits \autocite{averettBetterWorseRelationship2008, averettEconomicRealityBeauty1996, beckerTheoryMarriagePart1973, beckerTheoryMarriagePart1974, beckerTreatiseFamily1993, browningCollectiveUnitaryModels2006, chiapporiFatterAttractionAnthropometric2012}.
+Also, \textcite{duncanIntermarriageIntergenerationalTransmission2011} shows a similar pattern of assortative matching among Mexicans in the United States.
+These papers suggest that inter-ethnic WH and HW marriages will, on average, consist of partners sharing similar characteristics.
+Consequently, the literature would predict that both WH and HW parents will be similar in all aspects that are relevant to a child's labor market outcomes.
+I will go over the empirical evidence from the data to show that this is actually the case.
+
+\section{From the Data: The Differences Between HW and WH Couples}\label{sec:hw-wh-couples-data}
+
+In this section, I explore the empirical data to affirm the validity of my empirical strategy, as presented in Table \ref{tab:synth}, which details the educational and economic profiles of parents from four different ethnic groupings—White White (WW), White Hispanic (WH), Hispanic White (HW), and Hispanic Hispanic (HH).
+This table meticulously lays out the average outcomes and discrepancies for each group, highlighting the impact of inter-ethnic marriages on children's prospects.
+These results show that there is selection in marriage, and that the differences between children born to Hispanic fathers and White mother (HW) and children born to White fathers and Hispanic mother (WH) are less severe that those between the other two groups.
+Consequently, comparing WH and HW children to each other to analyze discrimination against Hispanics in the labor market.
+
+Using the ``synthetic'' parents sample that I constructed using the Census data, I examine the family background of the different type of children.
+WW couples have higher education, 12.58 years for husbands and 12.36 for wives.
+As a household, WW couples have 24.95 years of schooling.
+Husbands in HH marriages have 8.64 years of education, while women have 8.49 years of schooling.
+As a household, HH couples have 17.13 years of education.
+As predicted, the inter-ethnic couples exhibit matching in which partners marry people similar to them.
+WH husbands have 11.82 years of education, while wives have an average of 10.71 years.
+WH household attained a total of 22.68 years of schooling in total.
+HW husbands have 10.33 years of education, while wives have an average of 11.01 years.
+HW household attained a total of 21.50 years of schooling in total.
+Both Hispanic men and women in inter-ethnic couples marry white spouses that are, on average, more educated the the average HH couple.
+More importantly for human capital accumulation, the mother in the HW marriages---the children that have a Spanish sounding last-name---are more educated than their WH peers.
+Since mothers could be more important in the a child's human capital accumulation and education, this would suggest that a child with a Hispanic last-name should complete more years of education.
+
+The data reveals that WW couples have the highest total household education, amounting to 24.95 years, significantly surpassing the 17.69 years of HH couples.
+This stark contrast underlines the substantial educational divide between these groups.
+Inter-ethnic couples, WH and HW, display intermediate educational achievements, with WH households totaling 22.68 years of education and HW households slightly behind at 21.50 years.
+Notably, HW husbands are less educated (10.33 years) compared to WH husbands (11.82 years), yet HW wives surpass their WH counterparts with 11.01 years of education, suggesting a balance in educational attainment within these marriages.
+This factor is particularly pertinent for children with Hispanic last names who might derive greater benefits from their mother's higher education.
+
+In terms of labor market performance, WW households boast the highest log total family income at 10.75, while HH households fall at the lower end with 10.42.
+Among inter-ethnic couples, WH households have a slightly higher total income of 10.65 compared to HW households at 10.60.
+Notably, the difference in husbands' log hourly earnings between HW and WH is marginal, indicating HW men earn 4\% less than their WH counterparts.
+Furthermore, HW women actually surpass WH women in earnings, with HW wives earning 1.75 and WH wives earning 1.73 in log hourly earnings, respectively.
+This reversal in the typical earning pattern not only speaks to the closing economic disparities between these groups but also implies potentially greater economic contributions from HW women to their families, which could be advantageous for children and to the developments of traits that are important in the labor market.
+
+The table also sheds light on fertility trends, noting that WW couples have less children than HH couples, reflecting broader socio-economic and cultural patterns.
+The difference in fertility between HW and WH couples is positive but significantly lower than the difference in fertility between HH and WW couples.
+
+The evidence presented in this section supports a robust empirical strategy, revealing significant selection in marriage and distinctive educational and income patterns among different types of families.
+Inter-ethnic couples exhibit comparable levels of education and earnings.
+This suggests that the children of HW families are likely positioned for better educational outcomes, critical in understanding discrimination in the labor market.
+Notably, the disparities between HW and WH families are considerably less pronounced than those between other groups, emphasizing the importance of comparing these children directly.
+Such comparisons shed light on the nuanced dynamics of ethnicity, education, and economic outcomes in inter-ethnic marriages.
+\footnote{I present in Tables \ref{tab:synthmex} and \ref{tab:syntnonmex} the summary statistics which detail the educational and economic profiles of parents from four different ethnic groupings—White White (WW), White Hispanic (WH), Hispanic White (HW), and Hispanic Hispanic (HH) on sub-samples of Hispanics of Mexican and non-Mexican ancestries.
+I find similar results that describe a selection into inter-ethnic marriages among the two groups.}
+
+Despite higher levels of education and income among Hispanic White (HW) mothers compared to White Hispanic (WH) mothers, suggesting an expectation of greater educational attainment for HW children, data reveals a discrepancy.
+Specifically, HW children complete, on average, 0.4 fewer years of education compared to their WH peers (refer to Table \ref{tab:c&p1}).
+This gap may suggest potential discrimination or barriers in educational access for HW children.
+
 \section{Results}\label{sec:results}
 
-In this section, I present the results from estimating the two specifications in equations \ref{eq:1a} and \ref{eq:iden}.
-The results are shown in tables \ref{tab:lastnamereg} and \ref{tab:identreg}.
-All estimations are done on a sample of White, native-born men between the ages of 25 and 40 employed for full-time full-year (FTFY) and are wage and salary workers.
+In this section, I present the results from estimating the two specifications presented in equations \ref{eq:1a} and \ref{eq:iden}.
+I estimate the mean wages of White, U.S.-born, Hispanic men aged 25-40 who are employed full-time and full-year (FTFY) as waged and salaried workers.
+The results are shown in Tables \ref{tab:lastnamereg} and \ref{tab:identreg}.
 
@@
 However, more than half of the earnings gaps could be explained by educational differences.
-When I control for education, the last name effect decreases to a statistically insignificant 2 percentage points earnings gap.
+When I control for education, the last name effect decreases to a statistically insignificant 1 percentage points earnings gap.
 I also find a significant earnings gap between those that identify as Hispanic.
@@
 
-I provide the results to the estimation of equation \ref{eq:1a} in table \ref{tab:lastnamereg}.
-The sample in this analysis includes full-time, full-year, and wage and salary men, and I control for hours worked and age and include year fixed effects (FE).
+I provide the results to the estimation of equation \ref{eq:1a} in Table \ref{tab:lastnamereg}.
+I estimate the mean wages of White, U.S.-born, Hispanic men aged 25-40 who are employed full-time and full-year (FTFY) as waged and salaried workers.
 I also restrict the sample to children of HW and WH (omitted) parents.
-Column one in table \ref{tab:lastnamereg} is the results without controlling for education.
-Overall, an inter-ethnic with a Hispanic last name earns 5 percentage points less than an inter-ethnic with a White last name.
+Column one in Table \ref{tab:lastnamereg} is the average crude earnings gap in log annual earnings between HW workers and their WH peers.
+In the next 4 columns, I introduce the results with controls for hours worked, state FE, year FE, age FE, Education FE, and parental background.
+
+Overall, the crude gap between HW and WH workers is equal to 5 percentage points (Table \ref{tab:lastnamereg} column 1).
+An inter-ethnic with a Hispanic last name earns 5 percentage points less than an inter-ethnic with a White last name.
+Even after controlling for hours worked, and including state, year, and age FEs in the estimation, the gap stays at 5 percentage points.
 This gap, however, could be entirely explained by educational differences.
@@
 less than a Mexican with a native sounding last name.
-The gap becomes an imprecise zero after controlling for education.
+The gap becomes an imprecise zero after controlling for education and parental background.
 Third, in Table \ref{tab:lastnamereg-weekearm-nonmex}, I present the results to the estimation of equation \ref{eq:1a} using weekly earnings as a dependent variable on a sample of non-Mexican Hispanics.
-The gaps between non-Mexicans with Hispanic sounding last names and those with native sounding last names were all negative and statistically insignificant.
-I find that a non-Mexican with a Hispanic sounding last name earns 4 p.p.
+The gap between non-Mexicans with Hispanic sounding last names and those with native sounding last names was similarly explained by educational differences.
+I find that a non-Mexican with a Hispanic sounding last name earns 3 p.p.
 less than a Mexican with a native sounding last name.
-The gap becomes 1 p.p.
-after controlling for education.
+The gap becomes an imprecise zero after controlling for education and parental background.
 Finally, in Table \ref{tab:lastnamereg-weekearm-cub}, I present the results to the estimation of equation \ref{eq:1a} using weekly earnings as a dependent variable on a sample of Cubans.
-I find that a Cuban with Hispanic sounding last name earns 3 p.p.
+I find that a Cuban with Hispanic sounding last name earns 4 p.p.
 less than a Cuban with a native sounding last name but the gap was statistically insignificant.
 
-I provide the results to the estimation of equation \ref{eq:iden} in table \ref{tab:identreg}.
+I provide the results to the estimation of equation \ref{eq:iden} in Table \ref{tab:identreg}.
 The sample in this analysis includes full-time, full-year, and wage and salary men, and I control for hours worked and age and include year fixed effects (FE).
@@
 When I compare the earnings of HW and WH children, which captures the effect of having a Hispanic last name, HW children earn  6 percentage points less than WH children.
-Thus, by comparing inter-ethnic children, a person with a Hispanic last name makes 6 percentage points less than someone with a White last name.
+Thus, by comparing inter-ethnic children, a person with a Hispanic last name makes 5 percentage points less than someone with a White last name.
 When I control for education, the last name effect decreases to a statistically insignificant 2 percentage points earnings gap.
@@
 Education itself can be influenced by bias, potentially resulting in divergent outcomes.
-Consequently, further research is essential to comprehensively understand the earnings gaps between Hispanics and Whites.
+This is especially the case when parental characteristics indicate that people with a Hispanic-sounding last-names should in theory complete more years of education---since mothers of inter-ethnic children with a Hispanic last-names have more years of education and earn more.
+Consequently, further research is needed to comprehensively understand the earnings gaps between Hispanics and Whites.
 
@@
 \begin{appendices}
-\section{Figures}\label{appendix:figs}
+% \section{Figures}\label{appendix:figs}
 
@@
 
-\begin{center}
-\begin{figure}[h!]
-\caption{Distribution of the four types of children.}
-\includegraphics[width=\textwidth]{PirChart2.png} 
-\label{fig:dist}
-\end{figure}
-\end{center}
+% \begin{center}
+% \begin{figure}[h!]
+% \caption{Distribution of the four types of children.}
+% \includegraphics[width=\textwidth]{PirChart2.png} 
+% \label{fig:dist}
+% \end{figure}
+% \end{center}
 
-\newpage
+% \newpage
 
@@
 
-\input{tables/tab02-observations-by-couple}
+\input{tables/tab03b-summary-stats}
 
@@
 
-\input{tables/tab03b-summary-stats}
+\input{tables/tab04-summary-stats}
+
+\newpage
+
+\input{tables/tab02-observations-by-couple}
 
@@
 \input{tables/tab03a-summary-stats-nonmex}
-
-\newpage
-
-\input{tables/tab04-summary-stats}
 

### 2023-12-24 programs/04-ACS-prep.r

@@ -55,6 +55,7 @@ DATA <- read_dta(file.path(datasets,"ParentDataFull.dta")) |>
 		   								 bpld_pop  == 21000  ~ "Cental America",
            								 bpld_pop  == 30000  ~ "South America"))
 CrossTable(DATA$BirthPlaceMom)
+write_dta(DATA, file.path(datasets,"ParentDataFull.dta"))
 # create variables for parents' outcomes
 
 DATA  |> 
@@ -68,8 +69,16 @@ DATA  |>
               HusbandEducation      = weighted.mean(HusbandEducation, weights = Husband_wt, na.rm = TRUE),
               # Wife education
               WifeEducation         = weighted.mean(WifeEducation, weights = Wife_wt, na.rm = TRUE),
-              # Husband income
+              # family income
               Husband_ftotval       = weighted.mean(Husband_ftotval, weights = Husband_wt, na.rm = TRUE),
+			  # Husband hourly inctot
+              Husband_inctot        = weighted.mean(Hus_log_hourly_earnings_inctot, weights = Husband_wt, na.rm = TRUE),
+			  # Husband hourly incwage
+              Husband_incwage       = weighted.mean(Hus_log_hourly_earnings_incwage, weights = Husband_wt, na.rm = TRUE),
+			  # Wife hourly inctot
+              Wife_inctot           = weighted.mean(Wife_log_hourly_earnings_inctot, weights = Wife_wt, na.rm = TRUE),
+			  # Wife hourly incwage
+              Wife_incwage          = weighted.mean(Wife_log_hourly_earnings_incwage, weights = Wife_wt, na.rm = TRUE),
               # Fertility
               Fertility             = weighted.mean(Fertility, weights = Wife_wt, na.rm = TRUE)
               )
@@ -88,24 +97,43 @@ ParentData  <-  DATA  |>
               WifeEducation         = weighted.mean(WifeEducation, weights = Wife_wt, na.rm = TRUE),
               # Husband income
               Husband_ftotval       = weighted.mean(Husband_ftotval, weights = Husband_wt, na.rm = TRUE),
+			  # Husband hourly inctot
+              Husband_inctot        = weighted.mean(Hus_log_hourly_earnings_inctot, weights = Husband_wt, na.rm = TRUE),
+			  # Husband hourly incwage
+              Husband_incwage       = weighted.mean(Hus_log_hourly_earnings_incwage, weights = Husband_wt, na.rm = TRUE),
+			  # Wife hourly inctot
[…]

### 2023-12-24 programs/10-merge-cps-acs.r

@@ -24,7 +24,11 @@ SyntheticParents  <- read.csv(file.path(datasets_git,"ParentData.csv")) |>
                "LAWife",           
                "HusbandEducation", 
                "WifeEducation",    
-               "Husband_ftotval",  
+               "Husband_ftotval",
+			   "Husband_inctot",
+			   "Husband_incwage",
+			   "Wife_inctot",
+			   "Wife_incwage",  
                "Fertility")
 
 CPS  <- read_dta(file.path(datasets,"BySexAnalysisData.dta"))|> 
@@ -93,6 +97,52 @@ MergedData  <- left_join(CPS,
 summary(MergedData$Husband_ftotval)
 summary(MergedData$HusbandEducation)
 summary(MergedData$WifeEducation)
+CrossTable(MergedData$Type)
 
+MergedData  |> 
+	    group_by(Type) |>
+    	summarise(
+              # Husband is foreign born
+              LAHusband             = weighted.mean(LAHusband, weights = Husband_wt, na.rm = TRUE),
+              # Wife is foreign born
+              LAWife                = weighted.mean(LAWife, weights = Wife_wt, na.rm = TRUE),
+              # Husband education
+              HusbandEducation      = weighted.mean(HusbandEducation, weights = Husband_wt, na.rm = TRUE),
+              # Wife education
+              WifeEducation         = weighted.mean(WifeEducation, weights = Wife_wt, na.rm = TRUE),
+			  # Husband hourly inctot
+              Husband_ftotval        = weighted.mean(Husband_ftotval, weights = Husband_wt, na.rm = TRUE),
+			  # Husband hourlt incwage
+              Husband_incwage       = weighted.mean(Husband_incwage, weights = Husband_wt, na.rm = TRUE),
+			  # Wife hourly inctot
+              Wife_inctot           = weighted.mean(Wife_inctot, weights = Wife_wt, na.rm = TRUE),
+			  # Wife hourlt incwage
+              Wife_incwage          = weighted.mean(Wife_incwage, na.rm = TRUE),
+              # Fertility
+              Fertility             = weighted.mean(Fertility, weights = Wife_wt, na.rm = TRUE)
+              )
+
+DATA  |> 
+	    group_by(CoupleType) |>
+    	summarise(
+              # Husband is foreign born
+              LAHusband             = weighted.mean(LAHusband, weights = Husband_wt, na.rm = TRUE),
+              # Wife is foreign born
[…]

### 2023-12-24 programs/11-table-one-parent-types.R

@@ -99,6 +99,20 @@ Table_cols <- cbind(first_col, second_col,  third_col, fourth_col, fifth_col)
 Table_cols <-  Table_cols |> 
   row_to_names(row_number = 1)
 
+knitr::kable(Table_cols, "html", align = "lcccc",
+             booktabs = T,
+             escape = F,
+             caption = "Number of Children by Parental Type \\label{tab:mat1}") %>%
+  column_spec(1, bold = T) %>%
+  kable_classic(full_width = F) %>%
+  kable_styling(bootstrap_options = c("hover", "condensed"), 
+                latex_options = c("scale_down", "HOLD_position")) |> 
+  footnote(number = c("Source: Current Population Surveys (CPS) 1994-2019",
+                      "The sample includes Whites, who are married, and are between the ages 25 and 40. Ethnicity of a person's parents are identified by the parent's place of birth. A parent is Hispanic if she/he was born in a Spanish-speaking country. A parent is White if she/he was born in the United States."),
+           footnote_as_chunk = F, title_format = c("italic"),
+           escape = F, threeparttable = T
+  ) |> 
+  add_header_above(c(" " = 1, "Perental Type" = 4))
 
 knitr::kable(Table_cols, "latex", align = "lcccc",
              booktabs = T,
@@ -119,57 +133,41 @@ knitr::kable(Table_cols, "latex", align = "lcccc",
   save_kable(file.path("/Users/hhadah/Documents/GiT/my_thesis/tables","tab01-observations-by-parents.tex"))
 
 
-# presentation table
-knitr::kable(Table_cols, "latex", align = "lcccc",
-             booktabs = T,
-             escape = F,
-             caption = "Number of Children by Parental Type \\label{tab:mat1}") %>%
-  column_spec(1, bold = T) %>%
-  kable_classic(full_width = F) %>%
-  kable_styling(bootstrap_options = c("hover", "condensed"), 
-                latex_options = c("scale_down", "hold_position")) |> 
-  footnote(number = c("Source: Current Population Surveys (CPS) 1994-2019",
-                      "The sample includes Whites, who are married, and are between the ages 18 and 65. Ethnicity of a person's parents are identified by the parent's place of birth. A parent is Hispanic if she/he was born in a Spanish-speaking country. A parent is White she/he was born in the United States."),
[…]

### 2023-12-24 programs/12-table-two-couple-types.R

@@ -129,19 +129,4 @@ knitr::kable(Table_cols, "latex", align = "lcccc",
   ) |> 
   add_header_above(c(" " = 1, "Couples Type" = 4)) |> 
   save_kable(file.path(tables_wd,"tab02-observations-by-couple.tex")) %>% 
-  save_kable(file.path(thesis_tabs,"tab02-observations-by-couple.tex")) |> 
-    save_kable(file.path("/Users/hhadah/Documents/GiT/my_thesis/tables","tab02-observations-by-couple.tex"))
-
-
-# presentation table
-
-knitr::kable(Table_cols, "latex", align = "lcccc",
-             booktabs = T,
-             escape = F,
-             caption = "Couples' Type \\label{tab:mat2}") %>%
-  column_spec(1, bold = T) %>%
-  kable_classic(full_width = F) %>%
-  kable_styling(bootstrap_options = c("hover", "condensed"), 
-                latex_options = c("scale_down", "hold_position")) |>  
-  add_header_above(c(" " = 1, "Couples Type" = 4)) |> 
-  save_kable(file.path(dissertation_wd,"tables/tab02-observations-by-couple.tex"))
+  save_kable(file.path(thesis_tabs,"tab02-observations-by-couple.tex"))

### 2023-12-24 programs/13-table-three-summary-stats.R

@@ -4,7 +4,7 @@
 
 # date: January 19th, 2023
 
-ParentData     <- read_csv(file.path(datasets,"CPS_synth.csv")) |> 
+ParentData     <- read_dta(file.path(datasets,"ParentDataFull.dta")) |> 
   mutate(Tot_ed = HusbandEducation + WifeEducation)
 
 # Only Mexicans and Native Born
@@ -45,7 +45,7 @@ test2 <- tidy(glht(model, linfct = c("HW - WH = 0")))
 pvalues2 <- test2$std.error
 differences2 <- test2$estimate
 
-row3 <- c(
+row2 <- c(
   "Husband's education (Total Years)",
   paste0("\\specialcell{", 
          sprintf("%0.2f",round(HusbandEducation[4,2], digits = 2)),  
@@ -92,7 +92,7 @@ row3 <- c(
          ")}"
   )
 )
-dim(row3) <- c(1,7)
+dim(row2) <- c(1,7)
 
 ## Wife's Education
 WifeEducation <- ParentData |> 
@@ -110,7 +110,7 @@ test2 <- tidy(glht(model, linfct = c("HW - WH = 0")))
 pvalues2 <- test2$std.error
 differences2 <- test2$estimate
 
-row4 <- c(
+row3 <- c(
   "Wife's education (Total Years)",
   paste0("\\specialcell{", 
          sprintf("%0.2f",round(WifeEducation[4,2], digits = 2)),  
@@ -157,7 +157,7 @@ row4 <- c(
          ")}"
   )
 )
-dim(row4) <- c(1,7)
+dim(row3) <- c(1,7)
 
 ## Total Education
 TotalEducation <- ParentData |> 
@@ -175,7 +175,7 @@ test2 <- tidy(glht(model, linfct = c("HW - WH = 0")))
 pvalues2 <- test2$std.error
 differences2 <- test2$estimate
 
-row5 <- c(
+row4 <- c(
   "Total Household education (Total Years)",
   paste0("\\specialcell{", 
          sprintf("%0.2f",round(TotalEducation[4,2], digits = 2)),  
@@ -222,7 +222,7 @@ row5 <- c(
          ")}"
   )
 )
-dim(row5) <- c(1,7)
+dim(row4) <- c(1,7)
 
 ## Total Family Income
 TotalFamInc <- ParentData |> 
@@ -240,7 +240,7 @@ test2 <- tidy(glht(model, linfct = c("HW - WH = 0")))
[…]

### 2023-12-24 programs/13a-table-three-summary-stats-mex.R

@@ -4,7 +4,7 @@
 
 # date: January 19th, 2023
 
-ParentData     <- read_csv(file.path(datasets,"CPS_synth.csv")) |> 
+ParentData     <- read_dta(file.path(datasets,"ParentDataFull.dta")) |> 
   mutate(Tot_ed = HusbandEducation + WifeEducation)
 
 # Only Mexicans and Native Born
@@ -39,7 +39,7 @@ test2 <- tidy(glht(model, linfct = c("HW - WH = 0")))
 pvalues2 <- test2$std.error
 differences2 <- test2$estimate
 
-row3 <- c(
+row2 <- c(
   "Husband's education (Total Years)",
   paste0("\\specialcell{", 
          sprintf("%0.2f",round(HusbandEducation[4,2], digits = 2)),  
@@ -86,7 +86,7 @@ row3 <- c(
          ")}"
   )
 )
-dim(row3) <- c(1,7)
+dim(row2) <- c(1,7)
 
 ## Wife's Education
 WifeEducation <- ParentData |> 
@@ -104,7 +104,7 @@ test2 <- tidy(glht(model, linfct = c("HW - WH = 0")))
 pvalues2 <- test2$std.error
 differences2 <- test2$estimate
 
-row4 <- c(
+row3 <- c(
   "Wife's education (Total Years)",
   paste0("\\specialcell{", 
          sprintf("%0.2f",round(WifeEducation[4,2], digits = 2)),  
@@ -151,7 +151,7 @@ row4 <- c(
          ")}"
   )
 )
-dim(row4) <- c(1,7)
+dim(row3) <- c(1,7)
 
 ## Total Education
 TotalEducation <- ParentData |> 
@@ -169,7 +169,7 @@ test2 <- tidy(glht(model, linfct = c("HW - WH = 0")))
 pvalues2 <- test2$std.error
 differences2 <- test2$estimate
 
-row5 <- c(
+row4 <- c(
   "Total Household education (Total Years)",
   paste0("\\specialcell{", 
          sprintf("%0.2f",round(TotalEducation[4,2], digits = 2)),  
@@ -216,7 +216,7 @@ row5 <- c(
          ")}"
   )
 )
-dim(row5) <- c(1,7)
+dim(row4) <- c(1,7)
 
 ## Total Family Income
 TotalFamInc <- ParentData |> 
@@ -234,7 +234,7 @@ test2 <- tidy(glht(model, linfct = c("HW - WH = 0")))
[…]

### 2023-12-24 programs/13b-table-three-summary-stats-nonmex.R

@@ -4,7 +4,7 @@
 
 # date: January 19th, 2023
 
-ParentData     <- read_csv(file.path(datasets,"CPS_synth.csv")) |> 
+ParentData     <- read_dta(file.path(datasets,"ParentDataFull.dta")) |> 
   mutate(Tot_ed = HusbandEducation + WifeEducation)
 
 # Only Non-Mexicans
@@ -41,7 +41,7 @@ test2 <- tidy(glht(model, linfct = c("HW - WH = 0")))
 pvalues2 <- test2$std.error
 differences2 <- test2$estimate
 
-row3 <- c(
+row2 <- c(
   "Husband's education (Total Years)",
   paste0("\\specialcell{", 
          sprintf("%0.2f",round(HusbandEducation[4,2], digits = 2)),  
@@ -88,7 +88,7 @@ row3 <- c(
          ")}"
   )
 )
-dim(row3) <- c(1,7)
+dim(row2) <- c(1,7)
 
 ## Wife's Education
 WifeEducation <- ParentData |> 
@@ -106,7 +106,7 @@ test2 <- tidy(glht(model, linfct = c("HW - WH = 0")))
 pvalues2 <- test2$std.error
 differences2 <- test2$estimate
 
-row4 <- c(
+row3 <- c(
   "Wife's education (Total Years)",
   paste0("\\specialcell{", 
          sprintf("%0.2f",round(WifeEducation[4,2], digits = 2)),  
@@ -153,7 +153,7 @@ row4 <- c(
          ")}"
   )
 )
-dim(row4) <- c(1,7)
+dim(row3) <- c(1,7)
 
 ## Total Education
 TotalEducation <- ParentData |> 
@@ -171,7 +171,7 @@ test2 <- tidy(glht(model, linfct = c("HW - WH = 0")))
 pvalues2 <- test2$std.error
 differences2 <- test2$estimate
 
-row5 <- c(
+row4 <- c(
   "Total Household education (Total Years)",
   paste0("\\specialcell{", 
          sprintf("%0.2f",round(TotalEducation[4,2], digits = 2)),  
@@ -218,7 +218,7 @@ row5 <- c(
          ")}"
   )
 )
-dim(row5) <- c(1,7)
+dim(row4) <- c(1,7)
 
 ## Total Family Income
 TotalFamInc <- ParentData |> 
@@ -236,7 +236,7 @@ test2 <- tidy(glht(model, linfct = c("HW - WH = 0")))
[…]

### 2023-12-24 programs/14-table-three-summary-stats.R

@@ -2,11 +2,15 @@
 
 # table three
 
+# library(Hmisc)
+
 # date: January 19th, 2023
 
-ParentData     <- read_csv(file.path(datasets,"CPS_synth.csv")) |> 
-  mutate(Tot_ed = HusbandEducation + WifeEducation)
+IndividualData <- read_csv(file.path(datasets,"CPS_synth.csv")) |> 
+  mutate(Education = as.numeric(Education)) |> 
+  filter(WW == 1 | (WH == 1 ) | (HW == 1 ) | (HH == 1 ))
 
+CrossTable(IndividualData$Type, IndividualData$Hispanic_ID)
 row1 <- c("Variables", 
           "\\specialcell{White \\\\ White \\\\ (WW) \\\\ (1)}",
           "\\specialcell{White \\\\ Hispanic \\\\ (WH) \\\\ (2)}",
@@ -17,276 +21,20 @@ row1 <- c("Variables",
           )
 dim(row1) <- c(1,7)
 
-# parent's panel
-
-## Husband's Education
-HusbandEducation <- ParentData |> 
-  group_by(CoupleType) |> 
-  summarise(HusEducation   = mean(HusbandEducation, na.rm = T),
-            HusEducationSD = sd(HusbandEducation, na.rm = T))
-
-model <- lm(HusbandEducation ~ 0 + WW + WH + HW + HH, 
-    data = ParentData)
-test1 <- tidy(glht(model, linfct = c("HH - WW = 0")))
-pvalues1 <- test1$std.error
-differences1 <- test1$estimate
-
-test2 <- tidy(glht(model, linfct = c("HW - WH = 0")))
-pvalues2 <- test2$std.error
-differences2 <- test2$estimate
-
-row3 <- c(
-  "Husband's education (Total Years)",
-  paste0("\\specialcell{", 
-         sprintf("%0.2f",round(HusbandEducation[4,2], digits = 2)),  
-         "\\\\",
-         "(",
-         sprintf("%0.2f",round(HusbandEducation[4,3], digits = 2)), 
-         ")}"
-  ),
-  paste0("\\specialcell{", 
-         sprintf("%0.2f",round(HusbandEducation[3,2], digits = 2)),  
-         "\\\\",
-         "(",
-         sprintf("%0.2f",round(HusbandEducation[3,3], digits = 2)), 
-         ")}"
-  ),
-  paste0("\\specialcell{", 
-         sprintf("%0.2f",round(HusbandEducation[2,2], digits = 2)),  
[…]

### 2023-12-24 programs/16-table-five-regression.R

@@ -179,6 +179,3 @@ regression_tab %>%
 
 regression_tab %>%
   save_kable(file.path(thesis_tabs,"tab05-regression.tex"))
-regression_tab %>%
-  save_kable(file.path("/Users/hhadah/Documents/GiT/my_thesis/tables","tab05-regression.tex"))
-

### 2023-12-24 programs/17-table-six-regression.R

@@ -217,6 +217,3 @@ regression_tab %>%
 
 regression_tab %>%
   save_kable(file.path(thesis_tabs,"tab06-regression.tex"))
-regression_tab %>%
-  save_kable(file.path("/Users/hhadah/Documents/GiT/my_thesis/tables","tab06-regression.tex"))
-

### 2023-12-24 programs/19-table-eight-regression.R

@@ -15,7 +15,7 @@ ParentDummies <- c("WH", "HW", "HH")
 reg1 <- list(
   "\\specialcell{(1) \\\\ Education}" = feols(Education ~ .[ParentDummies] | year + age, se = "iid",
                                                         data = IndividualData))
-  )
+
 
 # calculate means to add
 # as a row

### 2023-12-25 manuscripts/economic-inquiry/fragment/preamble_package.tex
Add manuscript submission

--- 2023-07-28 my_paper/Hadah-last-names.tex
@@
-\documentclass{wptemp}
+%%%%%%%%%%%%%%%%%%%%%%%%%
+%% SEC: PACKAGEs MAIN  %%
+%%%%%%%%%%%%%%%%%%%%%%%%%
+\usepackage{mathpazo}
+\usepackage{eulervm}
+\usepackage{amsmath}
+\usepackage{amssymb}
+\usepackage[utf8]{inputenc}
+\usepackage[T1]{fontenc}
+\usepackage{color}
+% \usepackage[monochrome]{xcolor}
+\usepackage{babel}
+\usepackage{csquotes}
+\usepackage{setspace}
+\usepackage{graphicx}
+\usepackage{booktabs,dcolumn}
+\usepackage{array}
+\usepackage{multirow}
+\usepackage[font=singlespacing, skip=3pt]{caption}
+\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
+%\usepackage[usenames,dvipsnames,svgnames,table,monochrome]{xcolor} % test black and white color results, 2018-11-21 08:48
+\usepackage{float}
+\usepackage[export]{adjustbox}[2011/08/13]
+\usepackage{enumitem}
+% \usepackage{subfig}
+%\usepackage[nolists, tablesfirst, nomarkers]{endfloat}
+% \usepackage[unicode=true,pdfusetitle,
+%  bookmarks=true,bookmarksnumbered=false,bookmarksopen=false,
+%  breaklinks=false,pdfborder={0 0 1},backref=false,colorlinks=false]
+%  {hyperref}
+\usepackage[colorlinks=true, linkcolor=blue, citecolor=blue, plainpages=false, pdfpagelabels=true, urlcolor=blue]{hyperref}
+\usepackage{geometry}
+\usepackage{ragged2e}
+\usepackage{appendix}
+\usepackage{subcaption} % For subfigures
+
+\usepackage{makecell}
 \usepackage{adjustbox}
-\usepackage{dcolumn}    % aligning decimals
-    \newcolumntype{d}[1]{D{.}{.}{#1}}
+\usepackage{siunitx}
+\usepackage{indentfirst}
+\newcolumntype{d}{S[input-symbols = ()]}
+\usepackage{lscape}
+\usepackage{longtable}
+\newcommand{\source}[1]{\caption*{Source: {#1}} }
+\usepackage{threeparttable} % For tables with notes
+\usepackage{threeparttablex} % For extending threeparttable functionality
+
+% Redefine the footnoterule command
+\usepackage[hang,flushmargin]{footmisc}
+\renewcommand{\footnoterule}{%
+    \kern-3pt % Adjust the vertical position of the line
+    \hrule width 0.2\columnwidth height 0.4pt % Width and height of the line
+    \kern2.6pt % Adjust the spacing between the line and the footnotes
+}
+
+\renewcommand{\thetable}{\arabic{table}}
+\renewcommand{\thefigure}{\arabic{figure}}
+
 % *****************************************************************
@@
 
-\usepackage{lscape}
-\usepackage{threeparttablex}
-\usepackage{longtable}
-
 %%Original code developed by Jörg Weber: see
 %% https://www.jwe.cc/2012/03/stata-latex-tables-estout/
-%% and
+%% andwho
 %% https://www.jwe.cc/blog/
@@
 
-% Allow line breaks with \\ in specialcells
+% Allow line breaks with \ in specialcells
 \newcommand{\specialcell}[2][c]{%
@@
 
-\newcommand{\sym}[1]{\rlap{#1}}% Thanks David Carlisle
+% \newcommand{\sym}[1]{\rlap{#1}}% Thanks David Carlisle
 
@@
 \tikzstyle{arrow} = [thick,->,>=stealth]
-
-\begin{document}
-
-\linespread{1}{\title{{\fontfamily{ptm}
-\selectfont{\Large{\MakeUppercase{The Impact of Hispanic Last Names and Identity on Labor Market Outcomes}}}}}}
-	\author{{\fontfamily{pbk}\selectfont\large{\textsc{Hussain Hadah}}\thanks{Department of Economics, Tulane University, 6823 St.
-Charles Ave., New Orleans, LA 70118, United States (e-mail: \email{hhadah@tulane.edu}).}}}
-	\date{\fontfamily{pbk}\selectfont\normalsize{\textsc{\today}}}
-	\maketitle
-
-\begin{center}
-\href{https://hhadah.github.io/hispanic-last-names/my_paper/Hadah-last-names.pdf}{\textcolor{green!15!black!30!blue}{\footnotesize{\textsc{Click Here to Get the Most Updated Version}}}}
-\end{center}
-
-\pagenumbering{arabic}
-
-\begin{abstract} 
-In this paper, I study discrimination against Hispanics in the labor market.
-I compare the children of inter-ethnic marriages to study the impact of having a Hispanic last name.
-While males born to Hispanic father-White mothers earn 6 percentage points less than those born to White father-Hispanic mothers, the gap could be completely explained by educational differences.
-I also study the effect of identifying as Hispanic on earnings.
-I find that men with a Spanish-sounding last name who identify as Hispanic earn significantly less than those who do not identify as Hispanic, but the gap could also be explained by educational differences.
-
-\keywords{Economics of Minorities, Race, and Immigrants; Discrimination and Prejudice.}
-\JEL{J71; J64}
-
-\end{abstract}
-
-\pagebreak
-\section{Introduction}\label{sec:intro}
-
-A large literature documents substantial earnings gaps across race and ethnicity \citep{bayer2018divergent, charles2008prejudice, card1992school, fryer2004causes, rubinstein2014pride, bertrand2004emily, juhn1991accounting}.
-Hispanics constitute a large and growing portion of the population in the United States.
-As the number of Hispanics increases, determining whether ethnic discrimination affects their labor market outcomes becomes increasingly crucial \citep{chettyUnitedStatesStill2014, chettyEffectsExposureBetter2016,chettyFadingAmericanDream2017,abramitzkyImmigrantsAssimilateMore2020a, abramitzkyNationImmigrantsAssimilation2014,abramitzkyCulturalAssimilationAge2016,chettyWhereLandOpportunity2014}.
-Thus, it is important to understand whether a person's ethnicity affects their labor market outcomes.
-Assimilation and mobility are crucial because they reflect how well Hispanics can integrate into society and move up the socioeconomic ladder.
-
-In this paper, I answer the following questions.
-Does having a Hispanic last name affect labor market outcomes?
-What is the effect of identifying as Hispanic on earnings?
-Moreover, I aim to show that comparing Hispanic Whites to non-Hispanic Whites might create an artificially higher earnings gap since the two groups differ on many observable characteristics.\footnote{By identifying as Hispanic, I refer to individuals who self-report their ethnicity as Hispanic on surveys or other data collection instruments.} \footnote{Observable characteristics refer to factors that can be measured and quantified, such as education level, work experience, and immigration status.} Others have attempted to compare how native-born White Hispanics fare to non-Hispanic Whites and foreign-born Hispanics.
-In \citet{antman2020ethnic,antmanEthnicAttritionObserved2016,antmanEthnicAttritionObserved2016a,antmanEthnicAttritionAssimilation2020}, the authors compare the health and educational outcomes of Hispanic Whites to non-Hispanic Whites and native-born Hispanics to foreign-born Hispanics.
-They find gaps in education and health between Hispanics and Whites.
-They also find that native-born Hispanics are more likely than their foreign-born counterparts to report poor health.
-\citet{davilaChangesRelativeEarnings2008} documents many gaps in labor market outcomes between Hispanics and Whites.
-They attribute a big part of this gap to differences in education, experience, immigration status, and regional differences.
-
-The US population is growing in diversity.
-The proportion of non-Whites has increased by more than 10 percentage points from 13 percent in 1995 to 23 percent in 2019.
-The number of Hispanics has grown by 9 percentage points from 9 percent in 1995 to 16 percent in 2019.\footnote{The portion of non-Whites and Hispanics is calculated using the Current Population Survey (CPS).} Native-born White Hispanic men earn 21\% less than White men, although a substantial portion of the earnings gap is due to educational differences between Hispanics and Whites \citep{duncan2006hispanics, duncan2018identifying, duncan2018socioeconomic}.
-Some of the earnings differences may also be due to discrimination which will have negative consequences.
-For example, discrimination against Hispanics can lead to reduced job opportunities, lower wages, and hinder assimilation.
-In this paper, I examine the role of having a Hispanic last name and identifying as Hispanic on labor market outcomes.
-
-Identifying discrimination is difficult because of factors that affect labor market outcomes that are unobservable to economists---such as unobserved skills.
-One strategy used by researchers is audit or resume studies.
-\citet{bertrand2004emily} conducted an audit study where identical resumes were sent to employers with White and Black-sounding names.
-This approach, however, has its drawbacks.
-Audit studies only observe callbacks, not wages.
-
-This study utilizes a method developed by \citet{rubinstein2014pride}.
-I compare children from inter-ethnic marriages.
-More precisely, I compare children of Hispanic fathers and White mothers (henceforth HW) to children of White fathers and Hispanic mothers (henceforth  WH).
-This approach stems from the fact that marriages are not random.
-Couples match on several observable characteristics like income, schooling, socio-economic background, etc.
-\citep{averettBetterWorseRelationship2008, averettEconomicRealityBeauty1996, beckerTheoryMarriagePart1973, beckerTheoryMarriagePart1974, beckerTreatiseFamily1993, browningCollectiveUnitaryModels2006, chiapporiFatterAttractionAnthropometric2012}.
-Children of HW and WH marriages have more similar observable characteristics than children of endogamous/homogamous marriages--- i.e., White fathers-White mothers and Hispanic fathers-Hispanic mothers.
-Moreover, children from a Hispanic father and White mother household will have a Hispanic last name from their fathers, enabling the investigation of how ethnic signals, such as having a Hispanic last name, affect annual log earnings.
-
-Using the Current Population Survey (CPS) from 1994 to 2019 and the 1960 to 2000 US censuses, I estimate the effect of having a Hispanic last name on labor market outcomes.
-In other words, I approximate the variation in labor market perception of ethnic signals, in this case having a Hispanic last name, on annual log earnings.
-I also study the effect of identifying as Hispanic on labor market outcomes.
-
-I find that while men born to Hispanic father-White mothers earn less than men born to White father-Hispanic mothers, this gap is entirely explained by educational differences.
-Thus, I do not find a significant effect of having a Hispanic last name.
-Finally, I find that men identifying as Hispanic earn significantly less than those who do not, even when I control for ancestry and education.
-
-% Inter-ethnic children with a White father, and thus having a non-Hispanic last name, earn 14\% less than children of endogamous White parents.
-Inter-ethnic children with a Hispanic father, and thus having a Hispanic last name, earn 20\% less than children of endogamous White parents.
-Endogamous children with Hispanic parents earn 24\% less than children of endogamous White parents.
-A big part of the gaps could be explained by education.
-Inter-ethnic children with a White father and identify as Hispanic, and thus having a non-Hispanic last name, earn 14\% less than children of endogamous White parents.
-Inter-ethnic children with a Hispanic father and identify as Hispanic, and thus having a Hispanic last name, earn 18\% less than children of endogamous White parents.
-Endogamous children with Hispanic parents and identify as Hispanic earn 10\% less than children of endogamous White parents.
-
-Previous studies have also used names as a proxy for race and ethnicity \citep{fryer2004causes, rubinstein2014pride, bertrand2004emily}.
-\citet{fryer2004causes} point out that names can be a predictor of a person's race.
-Specifically, they provide a rising pattern among Blacks having different names than Whites.
-They did, however, find that having a Black name, after controlling for the home environment at birth, does not affect their labor market outcomes.
-\citet{rubinstein2014pride} compared the children of mixed marriages between Sephardic and Ashkenazi Jews in Israel.\footnote{Ashkenazi and Sephardic are two distinct Jewish ethnic groups.} They found that workers with Sephardic last names earn substantially less than those with Ashkenazi last names.
-\citet{bertrand2004emily} conducted an audit study by sending employers identical resumes that differ in the ethnic and racial signal of a name (Black sounding name versus a White sounding one).
-They found that resumes with Black-sounding names received substantially fewer callbacks than their White counterparts.
-
-The rest of the paper is organized as follows.
-Section \ref{sec:data} provides an overview of the data and summary statistics of the sample.
-Section \ref{sec:emp_model} introduces the empirical framework.
-I report the results in section \ref{sec:results}.
-Finally, I offer a brief conclusion in section \ref{sec:con1}.
-
-\section{Data}\label{sec:data}
-
-I use two datasets:  the Integrated Public Use Microdata Series (IPUMS) Current Population Survey (CPS) Annual Social and Economic (ASEC) \citep{cps2019} and the 1960 to 2000 US censuses \citep{acs2019}.
-
-I use the CPS data set to study the effect of having a Hispanic last name on a person's labor market outcomes.
-I take advantage of the fact that the CPS asks parents' place of birth, ethnicity, and race.
-The data spans the period between 1994--- the earliest sample to ask about a parent's place of birth--- to 2019.
-Moreover, since the CPS does not provide data on parents' characteristics, essential to determine the family background, I use the census to construct synthetic parents.
-The census offers a larger sample of potential parents.
-Similar to the CPS, the 1960 to 2000 censuses ask about the person's place of birth and the individual's race and ethnicity.
-I employ this information to construct "synthetic parents" using a method developed by \citet{rubinstein2014pride}.
-I construct the synthetic parents by linking husbands and wives in the census data to each other.
-I assume that parents have children between the ages of 25 and 40.
-I then link these synthetic parents using the birth year of child, and place of birth of parents to the year of birth of the "children" and the palce of birth of parents in the CPS sample.
-
-\subsection{Children of the four parental types}
-
-I use the CPS for my primary analysis of the effect of having a Hispanic last name on earnings.
-I restrict my sample to Whites, United States-born citizens aged 25 to 40 and born between 1960 to 2000.
-Taking advantage of data on parents' place of birth, I divide the sample into four groups depending on their parent's ethnicity.
-Mothers or fathers are Hispanic if they were born in a Spanish-speaking country and Puerto Rico, and White if they were born in the United States.\footnote{The list of Spanish Speaking countries include: Argentina, Bolivia, Chile, Colombia, Costa Rica, Cuba, Dominican Republic, Ecuador, El Salvador, Guatemala, Equatorial Guinea, Honduras, Mexico, Nicaragua, Panama, Paraguay, Peru, Spain, Uruguay, Venezuela.} Therefore, an observation can be the product of four types of parents:
-\begin{enumerate}
-\item White father and White mother (hereafter WW) 
-\item White father and Hispanic mother (hereafter WH)
-\item Hispanic father and White mother (hereafter HW)
-\item Hispanic father and Hispanic mother (hereafter HH).
-\end{enumerate}
-
-The distribution of the four types of children is presented in figure \ref{fig:dist} and table \ref{tab:mat1}.
-The majority of the sample (95.96\%) is WW children.
-The second biggest group is HH, which constitutes 2.69\% of the sample.
-Inter-ethnic children, WH and HW, make up 1.35\% of the sample with 90,325 observations.
-Even though WH and HW are only 1.35\% of the sample, I have plenty of observations to carry out an analysis.
-The summary statistics for the children of the four types of marriages are presented in table \ref{tab:c&p1}.
-Children of WW marriages (column 1) do better on every measure while children of HH parents (column 4) do worse than other children on every measure.
-
-On the one hand, WW children are more educated, with approximately 14 years of education for men and women.
-WW children have an unemployment rate of 4\% for men and women.
-They have log hourly earnings of 2.51 for men and 2.32 for women.
-Also, their annual log earnings are 10.29 for men and 9.43 for women.
-
-On the other hand, children of HH marriages have an average of 12.9 and 13.24 years of education for men and women, respectively.
-HH men and women have an unemployment rate of 7\% and 6\%.
-Their annual log earnings are 10.01 and 9.53 for men and women, respectively.
-
-Inter-ethnic children, WH and HW, have lower educational attainment and income than WW children, but they are better than HH children.
-Also, WH and HW are comparable, unlike WW and HH children.
-These two groups are important to the analysis.
-Children of HW are going to be the treatment group.
-HW children have a Hispanic father with a Hispanic last names.
-At the same time, the WH children have White fathers with a non-Hispanic, or White, last names.
-Summary statistics of these two groups, table \ref{tab:c&p1} columns (2) and (3) can show that they are different from the children of non-inter-ethnic marriages.
-
-WH children have an average of 13.58 and 13.79 years of education for men and women.
-The unemployment rate for men and women is 5\%.
-HW men and women obtained 13.22 and 13.42 years of schooling.
-Men's unemployment rate is 7\% and women's 6\%.
-
-In panel C of table \ref{tab:c&p1}, I present the rates at which members of each group self-identify as Hispanic.
-\footnote{Hispanic self-identification is observed when an individual notes that they are Hispanic on the CPS questionnaire} The vast majority of HH children, 96\% for men and 97\% for women, identify as Hispanic.
-There is considerable attrition in Hispanic identification of inter-ethnic children.
-Among WH children, 74\% of men and 78\% of women identify as Hispanic.
-Among HW children, 83\% of men and 82\% of women identify as Hispanic.
-These results align with the findings of a substantial body of literature that noted this attrition among children of immigrants, especially those coming from inter-ethnic marriages \citep{duncan2017complexity, duncan2018identifying, duncan2020new, antman2020ethnic}.
-Finally, a small portion of WW children, 4\% of men and 5\% of women, identified as Hispanic.
-These 5\% of WW children are most likely third generation--- or higher--- children of immigrants.
-For my analysis, I dropped the WW observations who identify as Hispanic when I compare with non-Hispanic Whites.
-
-In columns (5) and (6), I calculate the differences in means between HH-WW (column 5) and HW-WH (column 6).
-Overall, HH children do worse than WW children.
-HH has one year less of education compared to WW, and are 2 percentage points more likely to be unemployed.
-HH men earn 9\% less than WW men, and HH women make 2\% less than WW women.
-Therefore, these two groups, HH and WW, are different on many observable qualities.
-
-The differentials between WH and HW are not as severe as the ones between HH and WW.
-HW children are still worse off than WH but are much better than HH.
-HW men attain 0.36 fewer years of education than WH men.
-HW women achieve 0.37 fewer years of schooling than WH women.
-HW children are 1 percentage points more likely to be unemployed.
-HW men earn 1\% less annual earnings as WH men, and HW women earn 3\% less.
-These differences show that HW and WH children are different from children of endogamous/homogamous marriages and, thus, are more comparable.
-
-In table \ref{tab:c&p2}, I present the summary statistics for the four types and limit the sample to those that only identify as Hispanic.\footnote{Limiting the sample to those who identify as Hispanic only affects WH, HW, and HH.
-Members of the WW group are not affected by this restriction.} The same pattern discussed above between HH--WW and HW--WH persists.
-The one difference is that those who identify as Hispanic do slightly worse.
-
-\subsection{Synthetic parents}
-
-Using the 1960 to 2000 censuses, I constructed a data set of synthetic parents.
-The sample includes married White men and women.
-Even though the census asks a person whether they are Hispanic or not, I took advantage of the questions on place of birth to create a proxy for ethnicity.
-I consider a Hispanic person as White and born in a Spanish-speaking country.
-Consequently, Whites in the sample are people who are White and native-born.
-Using the information provided in the census, I can link husbands and wives with each other.
-I assume that parents have children between the ages of 25 and 40.
-Therefore, my sample consists of married White men and women with children that were born in the 1920 to 1975 cohorts\footnote{The construction of "synthetic parents" follows the method used by \citet{rubinstein2014pride}.}.
-
-I show the distribution of the four types of couples in table \ref{tab:mat2}.
-White husbands and White wives (WW) make up the majority of couples in the sample, 96\% (5,141,737 couples).
-Hispanic husbands and wives (HH) are the second largest group making up 2\% (119,749 couples) of all couples.
-White husbands and Hispanic wives (WH) couples are less 1\% (33,097 couples) of the sample and Hispanic husbands and White wives (HW) are also 1\% (37,847 couples).
-
-I present the summary statistics of the four types of synthetic parents in table \ref{tab:synth}.
-WW couples have higher education, 12.75 years for husbands and 12.47 for wives.
-As a household, WW couples have 25.22 years of schooling.
-Additionally, the log total family income for a WW household is 10.72 and have 3.77 children.
-Husbands in HH marriages have 8.64 years of education, while women have 8.49 years of schooling.
-As a household, HH couples have 17.13 years of education, log total family income of 10.27, and have 4.23 children.
-
-WH husbands have 11.77 years of education, while wives have an average of 10.40 years.
-WH household attained a total of 22.17 years of schooling in total, a log total family income of 10.55, and 3.98 children.
-HW husbands and wives obtained 10.25 and 11.11 years of education, respectively.
-HW household attained a total of 21.36 years of schooling in total, a log total family income of 10.46, and 4.15 children.
-
-In columns 5 and 6 of table \ref{tab:synth}, I report the differences in means between HH-WW households and HW-WH households.
-An HH Hispanic husband attained 4.11 fewer years of education than a WW husband.
-While an HH wife acquired 3.98 fewer years of schooling than a White WW wife.
-The total educational difference between HH and WW households equals 8.09 years of education.
-The educational differences between HW and WH are less severe than those between HH and WW.
-A Hispanic husband in intermarriage attained 1.52 fewer years of schooling than a White husband in intermarriage.
-A White wife in intermarriage attained 0.70 years of education, more than a Hispanic wife in intermarriage.
-The difference in total years of education between HW and WH households is 0.81, which is an 90\% decrease from the HH--WW gap.
-The total family income of an HH household is 45\% lower than a WW household, while the total family income of an HW household is 10\% lower than a WH household.
-These differences show that couples that intermarry are better and different than those that do not.
-
-%In table \ref{tab:p1} columns five and six, I provide the  HH-WW and HW-WH gaps.
-WW couples are more educated and earn more than HH couples.
-Also, WH husbands are more educated and earn more than HW husbands.
-In comparison, WH wives are more educated than HW wives.
-Even though White husbands in intermarriages are more educated and earn more than husbands in HW marriages, they are more comparable to each other than the non-intermarriage couples.
-
-\section{Empirical Strategy}\label{sec:emp_model}
-
-In this section, I present two empirical strategies.
-The first empirical strategy estimates the effect of having a Hispanic last name on annual log earnings.
-The second empirical strategy estimates the effect of identifying as Hispanic on log annual earnings.
-
-The difference in means between Hispanics and non-Hispanic Whites could result from discrimination.
-It can also be caused by differences in innate abilities, skills, and parental investments.
-While controlling for observable skill measures, I compare children of inter-ethnic marriages, HW and WH.
-WH and HW children are more similar in characteristics but provide employers, and the labor market, with different signals.\footnote{WH and HW children are both half White, half Hispanic.} WH children will have a non-Hispanic last name, while HW children will have a Hispanic last name.
-This is a method developed by \citet{rubinstein2014pride}.
-
-\subsection{Estimating the effect of having a Hispanic last name}
-
-In this section I restrict the sample to WH and HW groups.
-Let $Y_{ist}$ be the annual log earnings of person $i$ in state $s$ at time $t$.
-$HW_{ist}$ is an indicator variables for the type of parents person $i$ has.
-$X_{ist}$ is a vector of controls that includes age and numbers of hours worked, $\gamma_{t}$ are year fixed effects, $\lambda_{s}$ are state fixed effects, and $\phi_{ist}$ represents the error term.
-The equation for this strategy is written as follows:
-\begin{equation} \label{eq:1a}
-Y_{ist} = \beta_{1} HW_{ist} + X_{ist} \pi + \gamma_{t} + \lambda_s + \phi_{ist}
-\end{equation}
-
-$\beta_{1}$ is the coefficient of interest in this specification.
-$\beta_{1}$ represents the earnings gaps between children of inter-ethnic marriages who have a Spanish-sounding last name versus a White last name.
-
-\subsection{Estimating the effect of identifying as Hispanic}
-
-In this section, I will present an estimation strategy that would allow me to capture the effect of identifying as a Hispanic.
-
-\begin{equation} \label{eq:iden}
-Y_{ist} = \beta_{1} HW_{ist} + \beta_{2} Hispanic_i + \beta_{3} HW_{ist} \cdot Hispanic_i + X_{ist} \pi + \gamma_{t} + \lambda_s + \phi_{ist}
-\end{equation}
-
-The coefficients of interest from equation \ref{eq:iden} are $\beta_{1}$, $\beta_{2}$ and $\beta_{3}$.
-The coefficients $\beta_{1}$ and $\beta_{3}$ capture the effects of Spanish-sounding last names for $HW_{ist}$ that identify as Hispanic \textit{versus} those who do not.
-
-\section{Results}\label{sec:results}
-
-In this section, I present the results from estimating the two specifications in equations \ref{eq:1a} and \ref{eq:iden}.
-The results are shown in tables \ref{tab:lastnamereg} and \ref{tab:identreg}.
-All estimations are done on a sample of White, native-born men between the ages of 25 and 40 employed for full-time full-year (FTFY) and are wage and salary workers.
-
-I find that an inter-ethnic person with a Spanish-sounding last names earns less than an inter-ethnic who has a White last name.
-A person with a Spanish-sounding last name earns 5 percentage points less than a person with a White last name.
-In other words, by comparing inter-ethnic children, a person with a Hispanic last name earns 5 percentage points less than someone with a White last name.
-However, more than half of the earnings gaps could be explained by educational differences.
-When I control for education, the last name effect decreases to a statistically insignificant 2 percentage points earnings gap.
-I also find a significant earnings gap between those that identify as Hispanic.
-
-\subsection{The Effect of Having a Hispanic Last Name and Hispanic Identity on Labor Market Outcomes}
-
-I provide the results to the estimation of equation \ref{eq:1a} in table \ref{tab:lastnamereg}.
-The sample in this analysis includes full-time, full-year, and wage and salary men, and I control for hours worked and age and include year fixed effects (FE).
-I also restrict the sample to children of HW and WH (omitted) parents.
-Column one in table \ref{tab:lastnamereg} is the results without controlling for education.
-Overall, an inter-ethnic with a Hispanic last name earns 6 percentage points less than an inter-ethnic with a White last name.
-This gap, however, could be entirely explained by educational differences.
-An inter-ethnic with a Hispanic last name earns 6 percentage points less than an inter-ethnic with a White last name, but the result is statistically insignificant.
-
-I provide the results to the estimation of equation \ref{eq:iden} in table \ref{tab:identreg}.
-The sample in this analysis includes full-time, full-year, and wage and salary men, and I control for hours worked and age and include year fixed effects (FE).
-I also restrict the sample to children of HW and WH parents.
-The omitted group is children of WH parents that do not identify as Hispanic.
-Column one in table \ref{tab:identreg} is the results without controlling for education.
-An inter-ethnic $HW_{ist}$ that identifies as Hispanic \textit{versus} $HW_{ist}$ who do not---i.e.
-$HW_{ist} + HW_{ist} \cdot Hispanic_i$---earns 6 percentage points less than an inter-ethnic that does not, but the result could be entirely explained by educational differences.
-
-\section{Conclusion}\label{sec:con1}
-
-As the Hispanic population grows in the United States, studying discrimination against this group becomes increasingly important.
-In this paper, I examine discrimination against Hispanics in the labor market.
-More specifically, I examine the impact of Hispanic last names and Hispanic identification on annual log earnings.
-
-I compare the children of inter-ethnic marriages to study the labor market impact of having a Hispanic last name.
-When I compare the earnings of HW and WH children, which captures the effect of having a Hispanic last name, HW children earn  6 percentage points less than WH children.
-Thus, by comparing inter-ethnic children, a person with a Hispanic last name makes 6 percentage points less than someone with a White last name.
-When I control for education, the last name effect decreases to a statistically insignificant 2 percentage points earnings gap.
-Moreover, I estimate the effect of identifying as Hispanic and having a Hispanic last name.
-I find that an inter-ethnic child with a Hispanic last name and identifies as Hispanic earns 5 percentage points less that an inter-ethnic child with native sounding last name and identifies as Hispanic.
-
-While the earnings gap between children of inter-ethnic parents with and without a Hispanic last name disappears when controlling for education, it does not necessarily indicate the absence of discrimination.
-Education itself can be influenced by bias, potentially resulting in divergent outcomes.
-Consequently, further research is essential to comprehensively understand the earnings gaps between Hispanics and Whites.
-
-% \nocite{*}
-\pagebreak
-\bibliography{HispanicDiscriminationMarriageMarket}
-\newpage
-\pagebreak
-\begin{appendices}
-
-\section{Figures}\label{appendix:figures}
-
-\begin{center}
-\begin{figure}[h!]
-\caption{Inter-racial marriages as a percent of all marriages \\
- in the United States from 1995 to 2020.}
-\includegraphics[width=\textwidth]{interracialovertime.png} 
-\label{fig:3}
-\end{figure}
-\end{center}
-
-\newpage
-
-\begin{center}
-\begin{figure}[h!]
-\caption{Distribution of the four types of children.}
-\includegraphics[width=\textwidth]{PirChart2.png} 
-\label{fig:dist}
-\end{figure}
-\end{center}
-
-\newpage
-
-\begin{figure}[h!]
-\centering
-\caption{Chart explaining which synthetic parents and children and when we observe them.}
-\begin{tikzpicture}[node distance =2cm]
-\node (start) [startstop] {Synthetic Parents: Observed in 1960 to 2000. $25 \leq age \leq 40$, White, with kids};
-
-\node (pro1) [process, below of = start, yshift = -2cm] {Children: Observed between 1994 and 2019. Born between 1954 to 1999 and $25 \leq age \leq 40$};
-
-\draw [arrow] (start) -- (pro1);
-\end{tikzpicture}
-\label{flowchart1}
-\end{figure}
-
-\newpage
-
-\section{tables}\label{appendix:tables}
-
-\input{tables/tab01-observations-by-parents}
-
-\newpage
-
-\input{tables/tab02-observations-by-couple}
-
-\newpage
-
-\input{tables/tab03b-summary-stats}
-
-\newpage
-
-\input{tables/tab03a-summary-stats}
-
-\newpage
-
-\input{tables/tab03a-summary-stats-mex}
-
-\newpage
-
-\input{tables/tab03a-summary-stats-nonmex}
-
-\newpage
-
-\input{tables/tab04-summary-stats}
-
-\newpage
-
-\input{tables/tab05-regression}
-
-\newpage
-
-\input{tables/tab06-regression}
-
-\end{appendices}
-
-\end{document}

### 2023-12-25 manuscripts/economic-inquiry/manuscript_title_etc.tex
Add manuscript submission

--- 2023-12-11 my_paper/Hadah-last-names-draft.tex
@@
-\documentclass[12pt,english]{article}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% 1. Define Keywords, JEL
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\newcommand{\PAPERKEYWORDS}{\textbf{Keywords}: Economics of Minorities, Race, and Immigrants; Discrimination and Prejudice}
+\newcommand{\PAPERJEL}{\textbf{JEL}: J71; J64; J15}
 
-\input{fragments/preamble_packages.tex}
-\input{fragments/preamble_layout.tex}
-\input{fragments/preamble_citation.tex}
-\bibliography{bib/bibliography}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% 2. Define Title
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\newcommand{\PAPERTITLE}{\href{https://hhadah.github.io/hispanic-last-names/my_paper/Hadah-last-names-draft.pdf}{The Impact of Hispanic Last Names and Identity on Labor Market Outcomes}}
 
-\makeatother
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% 3. Define Authors contact information
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\newcommand{\AUTHORHADAH}{Hussain Hadah}
+\newcommand{\AUTHORHADAHURL}{https://orcid.org/0000-0002-8705-6386}
+\newcommand{\AUTHORHADAHINFO}{\href{\AUTHORHADAHURL}{\AUTHORHADAH}: Department of Economics, Tulane University, 6823 St.
+Charles Ave., New Orleans, LA 70118, United States (e-mail: \href{mailto:hhadah@tulane.edu}{hhadah@tulane.edu}, phone: +1-602-393-8077)}
 
-\begin{document}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% 4. Define Thanks
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\newcommand{\ACKNOWLEDGMENTS}{
+I thank Professors Willa Friedman, Chinhui Juhn, Vikram Maheshri, and Yona Rubinstein for their support and advice.
+I also thank Aimee Chin, Steven Craig, German Cubas, Elaine Liu, Fan HADAH, and the participants of the Applied Microeconomics Workshop at the University of Houston, and European Society for Population Economics (ESPE) for helpful feedback.}
 
-\title{The Impact of Hispanic Last Names and Identity on Labor Market Outcomes}
-
-\author{\href{https://hussainhadah.com/}{Hussain Hadah} \thanks{Hussain Hadah: Department of Economics, Tulane University, 6823 St.
-Charles Ave., New Orleans, LA 70118, United States (e-mail: \href{mailto:hhadah@tulane.edu}{hhadah@tulane.edu}).
-I thank Professors Willa Friedman, Chinhui Juhn, Vikram Maheshri, and Yona Rubinstein for their support and advice.
-I also thank Aimee Chin, Steven Craig, German Cubas, Elaine Liu, Fan Wang, and the participants of the Applied Microeconomics Workshop at the University of Houston.}}
-
-\maketitle
-
-\begin{center}
-\href{https://hhadah.github.io/hispanic-last-names/my_paper/Hadah-last-names-draft.pdf}{\textcolor{green!15!black!30!blue}{\footnotesize{\textsc{Click Here to Get the Most Updated Version}}}}
-\end{center}
-
-\begin{abstract}
-\singlespacing Do people with ethnically Hispanic names face discrimination in the labor market?
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% 5. Define Abstract
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\newcommand{\PAPERABSTRACT}{
+Do people with ethnically Hispanic names face discrimination in the labor market?
 In this study, I analyze the causal impact of Hispanic-sounding surnames on wages, using the unique context of inter-ethnic marriages and the distinct characteristics of White and Hispanic surnames.
@@
 Men with a Spanish-sounding last name who identify as Hispanic earn significantly less than those who do not, but again, this gap can be largely explained by educational disparities.
-This suggests that discrimination may extend beyond the labor market, influencing educational opportunities and outcomes as well.
+This suggests that discrimination may extend beyond the labor market, influencing educational opportunities and outcomes too.
+\PAPERJEL}
 
-\noindent\textbf{Keywords}: Economics of Minorities, Race, and Immigrants; Discrimination and Prejudice
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% 6. Define citation or availability of latest draft
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\newcommand{\PAPERDOIURL}{https://doi.org/10.1086/711654}
+\newcommand{\PAPERINFO}{
 
-\noindent\textbf{JEL Classification}: J71; J64; J15
-
-\end{abstract}
-
-\vfill
-\pagebreak{}
-
-\section{Introduction}\label{sec:intro}
-
-A large literature documents substantial earnings gaps across race and ethnicity \autocite{bayer2018divergent, charles2008prejudice, card1992school, fryer2004causes, rubinstein2014pride, bertrand2004emily, juhn1991accounting}.
-Hispanics constitute a large and growing portion of the population in the United States.
-As the number of Hispanics increases, determining whether ethnic discrimination affects their labor market outcomes becomes increasingly crucial \autocite{chettyUnitedStatesStill2014, chettyEffectsExposureBetter2016,chettyFadingAmericanDream2017,abramitzkyImmigrantsAssimilateMore2020a, abramitzkyNationImmigrantsAssimilation2014,abramitzkyCulturalAssimilationAge2016,chettyWhereLandOpportunity2014}.
-Thus, it is important to understand whether a person's ethnicity affects their labor market outcomes.
-Assimilation and mobility are crucial because they reflect how well Hispanics can integrate into society and move up the socioeconomic ladder.
-
-In this paper, I answer the following questions.
-Does having a Hispanic last name affect labor market outcomes?
-What is the effect of identifying as Hispanic on earnings?
-Moreover, I aim to show that comparing Hispanic Whites to non-Hispanic Whites might create an artificially higher earnings gap since the two groups differ on many observable characteristics.\footnote{By identifying as Hispanic, I refer to individuals who self-report their ethnicity as Hispanic on surveys or other data collection instruments.} \footnote{Observable characteristics refer to factors that can be measured and quantified, such as education level, work experience, and immigration status.} Others have attempted to compare how native-born White Hispanics fare to non-Hispanic Whites and foreign-born Hispanics.
-In \textcite{antman2020ethnic,antmanEthnicAttritionObserved2016,antmanEthnicAttritionObserved2016a,antmanEthnicAttritionAssimilation2020}, the authors compare the health and educational outcomes of Hispanic Whites to non-Hispanic Whites and native-born Hispanics to foreign-born Hispanics.
-They find gaps in education and health between Hispanics and Whites.
-They also find that native-born Hispanics are more likely than their foreign-born counterparts to report poor health.
-\textcite{davilaChangesRelativeEarnings2008} documents many gaps in labor market outcomes between Hispanics and Whites.
-They attribute a big part of this gap to differences in education, experience, immigration status, and regional differences.
-
-The US population is growing in diversity.
-The proportion of non-Whites has increased by more than 10 percentage points from 13 percent in 1995 to 23 percent in 2019.
-The number of Hispanics has grown by 9 percentage points from 9 percent in 1995 to 16 percent in 2019.\footnote{The portion of non-Whites and Hispanics is calculated using the Current Population Survey (CPS).} Native-born White Hispanic men earn 21\% less than White men, although a substantial portion of the earnings gap is due to educational differences between Hispanics and Whites \autocite{duncan2006hispanics, duncan2018identifying, duncan2018socioeconomic}.
-Some of the earnings differences may also be due to discrimination which will have negative consequences.
-For example, discrimination against Hispanics can lead to reduced job opportunities, lower wages, and hinder assimilation.
-In this paper, I examine the role of having a Hispanic last name and identifying as Hispanic on labor market outcomes.
-
-Identifying discrimination is difficult because of factors that affect labor market outcomes that are unobservable to economists---such as unobserved skills and separating it from prejudice and stereotypes.
-One strategy used by researchers is audit or resume studies.
-\textcite{bertrand2004emily} conducted an audit study where identical resumes were sent to employers with White and Black-sounding names.
-This approach, however, has its drawbacks.
-Audit studies only observe callbacks, not wages.
-
-This study utilizes a method developed by \textcite{rubinstein2014pride}.
-I compare children from inter-ethnic marriages.
-More precisely, I compare children of Hispanic fathers and White mothers (henceforth HW) to children of White fathers and Hispanic mothers (henceforth  WH).
-This approach stems from the fact that there is a strong selection among many characteristics, and thus, marriages are not random.
-Couples match on several observable characteristics like income, schooling, socio-economic background, etc.
-\autocite{averettBetterWorseRelationship2008, averettEconomicRealityBeauty1996, beckerTheoryMarriagePart1973, beckerTheoryMarriagePart1974, beckerTreatiseFamily1993, browningCollectiveUnitaryModels2006, chiapporiFatterAttractionAnthropometric2012}.
-Children of HW and WH marriages have more similar observable characteristics than children of endogamous/homogamous marriages--- i.e., White fathers-White mothers and Hispanic fathers-Hispanic mothers.
-Moreover, children from a Hispanic father and White mother household will have a Hispanic last name from their fathers, enabling the investigation of how ethnic signals, such as having a Hispanic last name, affect annual log earnings.
-
-Using the Current Population Survey (CPS) from 1994 to 2019 and the 1960 to 2000 US censuses, I estimate the effect of having a Hispanic last name on labor market outcomes.
-In other words, I approximate the variation in labor market perception of ethnic signals, in this case having a Hispanic last name, on annual log earnings.
-I also study the effect of identifying as Hispanic on labor market outcomes.
-
-I find that while men born to Hispanic father-White mothers earn less than men born to White father-Hispanic mothers, this gap is entirely explained by educational differences.
-Thus, I do not find a significant effect of having a Hispanic last name.
-Finally, I find that men identifying as Hispanic earn significantly less than those who do not, even when I control for ancestry and education.
-
-% Inter-ethnic children with a White father, and thus having a non-Hispanic last name, earn 14\% less than children of endogamous White parents.
-Inter-ethnic children with a Hispanic father, and thus having a Hispanic last name, earn 20\% less than children of endogamous White parents.
-Endogamous children with Hispanic parents earn 24\% less than children of endogamous White parents.
-A big part of the gaps could be explained by education.
-Inter-ethnic children with a White father and identify as Hispanic, and thus having a non-Hispanic last name, earn 14\% less than children of endogamous White parents.
-Inter-ethnic children with a Hispanic father and identify as Hispanic, and thus having a Hispanic last name, earn 18\% less than children of endogamous White parents.
-Endogamous children with Hispanic parents and identify as Hispanic earn 10\% less than children of endogamous White parents.
-
-Previous studies have also used names as a proxy for race and ethnicity \autocite{fryer2004causes, rubinstein2014pride, bertrand2004emily}.
-\textcite{fryer2004causes} point out that names can be a predictor of a person's race.
-Specifically, they provide a rising pattern among Blacks having different names than Whites.
-They did, however, find that having a Black name, after controlling for the home environment at birth, does not affect their labor market outcomes.
-\textcite{rubinstein2014pride} compared the children of mixed marriages between Sephardic and Ashkenazi Jews in Israel.\footnote{Ashkenazi and Sephardic are two distinct Jewish ethnic groups.} They found that workers with Sephardic last names earn substantially less than those with Ashkenazi last names.
-\textcite{bertrand2004emily} conducted an audit study by sending employers identical resumes that differ in the ethnic and racial signal of a name (Black sounding name versus a White sounding one).
-They found that resumes with Black-sounding names received substantially fewer callbacks than their White counterparts.
-
-The rest of the paper is organized as follows.
-Section \ref{sec:data} provides an overview of the data and summary statistics of the sample.
-Section \ref{sec:emp_model} introduces the empirical framework.
-I report the results in section \ref{sec:results}.
-Finally, I offer a brief conclusion in section \ref{sec:con1}.
-
-\section{Data}\label{sec:data}
-
-I use two datasets:  the Integrated Public Use Microdata Series (IPUMS) Current Population Survey (CPS) Annual Social and Economic (ASEC) \autocite{cps2019} and the 1960 to 2000 US censuses \autocite{acs2019}.
-
-I use the CPS data set to study the effect of having a Hispanic last name on a person's labor market outcomes.
-I take advantage of the fact that the CPS asks parents' place of birth, ethnicity, and race.
-The data spans the period between 1994--- the earliest sample to ask about a parent's place of birth--- to 2019.
-Moreover, since the CPS does not provide data on parents' characteristics, essential to determine the family background, I use the census to construct synthetic parents.
-The census offers a larger sample of potential parents.
-Similar to the CPS, the 1960 to 2000 censuses ask about the person's place of birth and the individual's race and ethnicity.
-I employ this information to construct "synthetic parents" using a method developed by \textcite{rubinstein2014pride}.
-I construct the synthetic parents by linking husbands and wives in the census data to each other.
-I assume that parents have children between the ages of 25 and 40.
-I then link these synthetic parents using the birth year of child, and place of birth of parents to the year of birth of the "children" and the palce of birth of parents in the CPS sample.
-
-\subsection{Children of the four parental types}
-
-I use the CPS for my primary analysis of the effect of having a Hispanic last name on earnings.
-I restrict my sample to Whites, United States-born citizens aged 25 to 40 and born between 1960 to 2000.
-Taking advantage of data on parents' place of birth, I divide the sample into four groups depending on their parent's ethnicity.
-Mothers or fathers are Hispanic if they were born in a Spanish-speaking country and Puerto Rico, and White if they were born in the United States.\footnote{The list of Spanish Speaking countries include: Argentina, Bolivia, Chile, Colombia, Costa Rica, Cuba, Dominican Republic, Ecuador, El Salvador, Guatemala, Equatorial Guinea, Honduras, Mexico, Nicaragua, Panama, Paraguay, Peru, Spain, Uruguay, Venezuela.} Therefore, an observation can be the product of four types of parents:
-\begin{enumerate}
-\item White father and White mother (hereafter WW) 
-\item White father and Hispanic mother (hereafter WH)
-\item Hispanic father and White mother (hereafter HW)
-\item Hispanic father and Hispanic mother (hereafter HH).
-\end{enumerate}
-
-The distribution of the four types of children is presented in figure \ref{fig:dist} and table \ref{tab:mat1}.
-The majority of the sample (95.96\%) is WW children.
-The second biggest group is HH, which constitutes 2.69\% of the sample.
-Inter-ethnic children, WH and HW, make up 1.35\% of the sample with 90,325 observations.
-Even though WH and HW are only 1.35\% of the sample, I have plenty of observations to carry out an analysis.
-The summary statistics for the children of the four types of marriages are presented in table \ref{tab:c&p1}.
-Children of WW marriages (column 1) do better on every measure while children of HH parents (column 4) do worse than other children on every measure.
-
-On the one hand, WW children are more educated, with approximately 14 years of education for men and women.
-WW children have an unemployment rate of 4\% for men and women.
-They have log hourly earnings of 2.51 for men and 2.32 for women.
-Also, their annual log earnings are 10.29 for men and 9.43 for women.
-
-On the other hand, children of HH marriages have an average of 12.9 and 13.24 years of education for men and women, respectively.
-HH men and women have an unemployment rate of 7\% and 6\%.
-Their annual log earnings are 10.01 and 9.53 for men and women, respectively.
-
-Inter-ethnic children, WH and HW, have lower educational attainment and income than WW children, but they are better than HH children.
-Also, WH and HW are comparable, unlike WW and HH children.
-These two groups are important to the analysis.
-Children of HW are going to be the treatment group.
-HW children have a Hispanic father with a Hispanic last names.
-At the same time, the WH children have White fathers with a non-Hispanic, or White, last names.
-Summary statistics of these two groups, table \ref{tab:c&p1} columns (2) and (3) can show that they are different from the children of non-inter-ethnic marriages.
-
-WH children have an average of 13.58 and 13.79 years of education for men and women.
-The unemployment rate for men and women is 5\%.
-HW men and women obtained 13.22 and 13.42 years of schooling.
-Men's unemployment rate is 7\% and women's 6\%.
-
-In panel C of table \ref{tab:c&p1}, I present the rates at which members of each group self-identify as Hispanic.
-\footnote{Hispanic self-identification is observed when an individual notes that they are Hispanic on the CPS questionnaire} The vast majority of HH children, 96\% for men and 97\% for women, identify as Hispanic.
-There is considerable attrition in Hispanic identification of inter-ethnic children.
-Among WH children, 74\% of men and 78\% of women identify as Hispanic.
-Among HW children, 83\% of men and 82\% of women identify as Hispanic.
-These results align with the findings of a substantial body of literature that noted this attrition among children of immigrants, especially those coming from inter-ethnic marriages \autocite{duncan2017complexity, duncan2018identifying, duncan2020new, antman2020ethnic}.
-Finally, a small portion of WW children, 4\% of men and 5\% of women, identified as Hispanic.
-These 5\% of WW children are most likely third generation--- or higher--- children of immigrants.
-For my analysis, I dropped the WW observations who identify as Hispanic when I compare with non-Hispanic Whites.
-
-In columns (5) and (6), I calculate the differences in means between HH-WW (column 5) and HW-WH (column 6).
-Overall, HH children do worse than WW children.
-HH has one year less of education compared to WW, and are 2 percentage points more likely to be unemployed.
-HH men earn 9\% less than WW men, and HH women make 2\% less than WW women.
-Therefore, these two groups, HH and WW, are different on many observable qualities.
-
-The differentials between WH and HW are not as severe as the ones between HH and WW.
-HW children are still worse off than WH but are much better than HH.
-HW men attain 0.36 fewer years of education than WH men.
-HW women achieve 0.37 fewer years of schooling than WH women.
-HW children are 1 percentage points more likely to be unemployed.
-HW men earn 1\% less annual earnings as WH men, and HW women earn 3\% less.
-These differences show that HW and WH children are different from children of endogamous/homogamous marriages and, thus, are more comparable.
-
-In table \ref{tab:c&p2}, I present the summary statistics for the four types and limit the sample to those that only identify as Hispanic.\footnote{Limiting the sample to those who identify as Hispanic only affects WH, HW, and HH.
-Members of the WW group are not affected by this restriction.} The same pattern discussed above between HH--WW and HW--WH persists.
-The one difference is that those who identify as Hispanic do slightly worse.
-
-\subsection{Synthetic parents}
-
-Using the 1960 to 2000 censuses, I constructed a data set of synthetic parents.
-The sample includes married White men and women.
-Even though the census asks a person whether they are Hispanic or not, I took advantage of the questions on place of birth to create a proxy for ethnicity.
-I consider a Hispanic person as White and born in a Spanish-speaking country.
-Consequently, Whites in the sample are people who are White and native-born.
-Using the information provided in the census, I can link husbands and wives with each other.
-I assume that parents have children between the ages of 25 and 40.
-Therefore, my sample consists of married White men and women with children that were born in the 1920 to 1975 cohorts\footnote{The construction of "synthetic parents" follows the method used by \textcite{rubinstein2014pride}.}.
-
-I show the distribution of the four types of couples in table \ref{tab:mat2}.
-White husbands and White wives (WW) make up the majority of couples in the sample, 96\% (5,141,737 couples).
-Hispanic husbands and wives (HH) are the second largest group making up 2\% (119,749 couples) of all couples.
-White husbands and Hispanic wives (WH) couples are less 1\% (33,097 couples) of the sample and Hispanic husbands and White wives (HW) are also 1\% (37,847 couples).
-
-I present the summary statistics of the four types of synthetic parents in table \ref{tab:synth}.
-WW couples have higher education, 12.75 years for husbands and 12.47 for wives.
-As a household, WW couples have 25.22 years of schooling.
-Additionally, the log total family income for a WW household is 10.72 and have 3.77 children.
-Husbands in HH marriages have 8.64 years of education, while women have 8.49 years of schooling.
-As a household, HH couples have 17.13 years of education, log total family income of 10.27, and have 4.23 children.
-
-WH husbands have 11.77 years of education, while wives have an average of 10.40 years.
-WH household attained a total of 22.17 years of schooling in total, a log total family income of 10.55, and 3.98 children.
-HW husbands and wives obtained 10.25 and 11.11 years of education, respectively.
-HW household attained a total of 21.36 years of schooling in total, a log total family income of 10.46, and 4.15 children.
-
-In columns 5 and 6 of table \ref{tab:synth}, I report the differences in means between HH-WW households and HW-WH households.
-An HH Hispanic husband attained 4.11 fewer years of education than a WW husband.
-While an HH wife acquired 3.98 fewer years of schooling than a White WW wife.
-The total educational difference between HH and WW households equals 8.09 years of education.
-The educational differences between HW and WH are less severe than those between HH and WW.
-A Hispanic husband in intermarriage attained 1.52 fewer years of schooling than a White husband in intermarriage.
-A White wife in intermarriage attained 0.70 years of education, more than a Hispanic wife in intermarriage.
-The difference in total years of education between HW and WH households is 0.81, which is an 90\% decrease from the HH--WW gap.
-The total family income of an HH household is 45\% lower than a WW household, while the total family income of an HW household is 10\% lower than a WH household.
-These differences show that couples that intermarry are better and different than those that do not.
-
-%In table \ref{tab:p1} columns five and six, I provide the  HH-WW and HW-WH gaps.
-WW couples are more educated and earn more than HH couples.
-Also, WH husbands are more educated and earn more than HW husbands.
-In comparison, WH wives are more educated than HW wives.
-Even though White husbands in intermarriages are more educated and earn more than husbands in HW marriages, they are more comparable to each other than the non-intermarriage couples.
-
-\section{Empirical Strategy}\label{sec:emp_model}
-
-In this section, I present two empirical strategies.
-The first empirical strategy estimates the effect of having a Hispanic last name on annual log earnings.
-The second empirical strategy estimates the effect of identifying as Hispanic on log annual earnings.
-
-The difference in means between Hispanics and non-Hispanic Whites could result from discrimination.
-It can also be caused by differences in innate abilities, skills, and parental investments.
-While controlling for observable skill measures, I compare children of inter-ethnic marriages, HW and WH.
-WH and HW children are more similar in characteristics but provide employers, and the labor market, with different signals.\footnote{WH and HW children are both half White, half Hispanic.} WH children will have a non-Hispanic last name, while HW children will have a Hispanic last name.
-This is a method developed by \textcite{rubinstein2014pride}.
-
-\subsection{Estimating the effect of having a Hispanic last name}
-
-In this section I restrict the sample to WH and HW groups.
-Let $Y_{ist}$ be the annual log earnings of person $i$ in state $s$ at time $t$.
-$HW_{ist}$ is an indicator variables for the type of parents person $i$ has.
-$X_{ist}$ is a vector of controls that includes age and numbers of hours worked, $\gamma_{t}$ are year fixed effects, $\lambda_{s}$ are state fixed effects, and $\phi_{ist}$ represents the error term.
-The equation for this strategy is written as follows:
-\begin{equation} \label{eq:1a}
-Y_{ist} = \beta_{1} HW_{ist} + X_{ist} \pi + \gamma_{t} + \lambda_s + \phi_{ist}
-\end{equation}
-
-$\beta_{1}$ is the coefficient of interest in this specification.
-$\beta_{1}$ represents the earnings gaps between children of inter-ethnic marriages who have a Spanish-sounding last name versus a White last name.
-
-\subsection{Estimating the effect of identifying as Hispanic}
-
-In this section, I will present an estimation strategy that would allow me to capture the effect of identifying as a Hispanic.
-
-\begin{equation} \label{eq:iden}
-Y_{ist} = \beta_{1} HW_{ist} + \beta_{2} Hispanic_i + \beta_{3} HW_{ist} \cdot Hispanic_i + X_{ist} \pi + \gamma_{t} + \lambda_s + \phi_{ist}
-\end{equation}
-
-The coefficients of interest from equation \ref{eq:iden} are $\beta_{1}$, $\beta_{2}$ and $\beta_{3}$.
-The coefficients $\beta_{1}$ and $\beta_{3}$ capture the effects of Spanish-sounding last names for $HW_{ist}$ that identify as Hispanic \textit{versus} those who do not.
-
-\section{Results}\label{sec:results}
-
-In this section, I present the results from estimating the two specifications in equations \ref{eq:1a} and \ref{eq:iden}.
-The results are shown in tables \ref{tab:lastnamereg} and \ref{tab:identreg}.
-All estimations are done on a sample of White, native-born men between the ages of 25 and 40 employed for full-time full-year (FTFY) and are wage and salary workers.
-
-I find that an inter-ethnic person with a Spanish-sounding last names earns less than an inter-ethnic who has a White last name.
-A person with a Spanish-sounding last name earns 5 percentage points less than a person with a White last name.
-In other words, by comparing inter-ethnic children, a person with a Hispanic last name earns 5 percentage points less than someone with a White last name.
-However, more than half of the earnings gaps could be explained by educational differences.
-When I control for education, the last name effect decreases to a statistically insignificant 2 percentage points earnings gap.
-I also find a significant earnings gap between those that identify as Hispanic.
-
-\subsection{The Effect of Having a Hispanic Last Name and Hispanic Identity on Labor Market Outcomes}
-
-I provide the results to the estimation of equation \ref{eq:1a} in table \ref{tab:lastnamereg}.
-The sample in this analysis includes full-time, full-year, and wage and salary men, and I control for hours worked and age and include year fixed effects (FE).
-I also restrict the sample to children of HW and WH (omitted) parents.
-Column one in table \ref{tab:lastnamereg} is the results without controlling for education.
-Overall, an inter-ethnic with a Hispanic last name earns 5 percentage points less than an inter-ethnic with a White last name.
-This gap, however, could be entirely explained by educational differences.
-An inter-ethnic with a Hispanic last name earns 1 percentage points less than an inter-ethnic with a White last name, but the result is statistically insignificant.
-
-Since Hispanics are a very heterogeneous group, I conduct a heterogeneity analysis on different samples of Hispanics.
-To increase the sample size of my analysis I estimate equation \ref{eq:1a} using weekly earnings as a dependent variable.
-Weekly earnings are available in all of the monthly CPS surveys and not just the March Supplement.
-I present the results to these estimations in tables \ref{tab:lastnamereg-weekearm}-\ref{tab:lastnamereg-weekearm-cub}.
-
-First, in Table \ref{tab:lastnamereg-weekearm}, I present the results to the estimation of equation \ref{eq:1a} using weekly earnings as a dependent variable on the full sample.
-Similar to the analysis before, I find that a person with a Hispanic last name earns 4 p.p.
-less that a person without a Hispanic last name.
-This gap could be explained by Educational differences.
-Second, in Table \ref{tab:lastnamereg-weekearm-mex}, I present the results to the estimation of equation \ref{eq:1a} using weekly earnings as a dependent variable on a sample of Mexican Hispanics.
-I find that a Mexican with a Hispanic sounding last name earns 3 p.p.
-less than a Mexican with a native sounding last name.
-The gap becomes an imprecise zero after controlling for education.
-Third, in Table \ref{tab:lastnamereg-weekearm-nonmex}, I present the results to the estimation of equation \ref{eq:1a} using weekly earnings as a dependent variable on a sample of non-Mexican Hispanics.
-The gaps between non-Mexicans with Hispanic sounding last names and those with native sounding last names were all negative and statistically insignificant.
-I find that a non-Mexican with a Hispanic sounding last name earns 4 p.p.
-less than a Mexican with a native sounding last name.
-The gap becomes 1 p.p.
-after controlling for education.
-Finally, in Table \ref{tab:lastnamereg-weekearm-cub}, I present the results to the estimation of equation \ref{eq:1a} using weekly earnings as a dependent variable on a sample of Cubans.
-I find that a Cuban with Hispanic sounding last name earns 3 p.p.
-less than a Cuban with a native sounding last name but the gap was statistically insignificant.
-
-I provide the results to the estimation of equation \ref{eq:iden} in table \ref{tab:identreg}.
-The sample in this analysis includes full-time, full-year, and wage and salary men, and I control for hours worked and age and include year fixed effects (FE).
-I also restrict the sample to children of HW and WH parents.
-The omitted group is children of WH parents that do not identify as Hispanic.
-Column one in table \ref{tab:identreg} is the results without controlling for education.
-An inter-ethnic $HW_{ist}$ that identifies as Hispanic \textit{versus} $HW_{ist}$ who do not---i.e.
-$HW_{ist} + HW_{ist} \cdot Hispanic_i$---earns 6 percentage points less than an inter-ethnic that does not, but the result could be entirely explained by educational differences.
-
-\section{Conclusion}\label{sec:con1}
-
-As the Hispanic population grows in the United States, studying discrimination against this group becomes increasingly important.
-In this paper, I examine discrimination against Hispanics in the labor market.
-More specifically, I examine the impact of Hispanic last names and Hispanic identification on annual log earnings.
-
-I compare the children of inter-ethnic marriages to study the labor market impact of having a Hispanic last name.
-When I compare the earnings of HW and WH children, which captures the effect of having a Hispanic last name, HW children earn  6 percentage points less than WH children.
-Thus, by comparing inter-ethnic children, a person with a Hispanic last name makes 6 percentage points less than someone with a White last name.
-When I control for education, the last name effect decreases to a statistically insignificant 2 percentage points earnings gap.
-Moreover, I estimate the effect of identifying as Hispanic and having a Hispanic last name.
-I find that an inter-ethnic child with a Hispanic last name and identifies as Hispanic earns 5 percentage points less that an inter-ethnic child with native sounding last name and identifies as Hispanic.
-
-While the earnings gap between children of inter-ethnic parents with and without a Hispanic last name disappears when controlling for education, it does not necessarily indicate the absence of discrimination.
-Education itself can be influenced by bias, potentially resulting in divergent outcomes.
-Consequently, further research is essential to comprehensively understand the earnings gaps between Hispanics and Whites.
-
-\pagebreak
-\begingroup
-\setstretch{1.0}
-%\setstretch{1.1}
-\setlength\bibitemsep{0pt}
-\printbibliography
-\endgroup
-\pagebreak
-
-\begin{appendices}
-\section{Figures}\label{appendix:figs}
-
-% \begin{center}
-% \begin{figure}[H]
-% \caption{Histogram Before Targetting}
-% \includegraphics[width=\textwidth]{img/hist_dep_original.jpg} 
-% \label{fig:hist}
-% \end{figure}
-% \end{center}
-
-% \pagebreak
-
-\begin{center}
-\begin{figure}[h!]
-\caption{Distribution of the four types of children.}
-\includegraphics[width=\textwidth]{PirChart2.png} 
-\label{fig:dist}
-\end{figure}
-\end{center}
-
-\newpage
-
-\section{Tables}\label{appendix:tabs}
-
-\input{tables/tab01-observations-by-parents}
-
-\newpage
-
-\input{tables/tab02-observations-by-couple}
-
-\newpage
-
-\input{tables/tab03b-summary-stats}
-
-\newpage
-
-\input{tables/tab03a-summary-stats}
-
-\newpage
-
-\input{tables/tab03a-summary-stats-mex}
-
-\newpage
-
-\input{tables/tab03a-summary-stats-nonmex}
-
-\newpage
-
-\input{tables/tab04-summary-stats}
-
-\newpage
-
-\input{tables/tab05-regression}
-
-\newpage
-
-\input{tables/tab07-regression-weekearn}
-
-\newpage
-
-\input{tables/tab08-regression-weekearn-mex}
-
-\newpage
-
-\input{tables/tab09-regression-weekearn-nonmex}
-
-\newpage
-
-\input{tables/tab10-regression-weekearn-cuba}
-
-\newpage
-
-\input{tables/tab06-regression}
-
-\end{appendices}
-
-\end{document}
+ \url{\PAPERDOIURL}.
+}

### 2023-12-25 manuscripts/Hussain-Hadah/tables/tab03-summary-stats.tex
....

--- 2023-04-12 my_paper/tables/tab03-summary-stats.tex
@@
-\begin{table}[!h]
-
+\begin{table}[H]
+\tablefont
 \caption{Summary statistics of outcomes using parent's place of birth \label{tab:c&p1}}
@@
 \cmidrule(l{3pt}r{3pt}){2-5} \cmidrule(l{3pt}r{3pt}){6-7}
-Variables & \specialcell{White Father \\ White Mother \\ (WW) \\ (i)} & \specialcell{White Father \\ Hispanic Mother \\ (WH) \\ (ii)} & \specialcell{Hispanic Father \\ White Mother \\ (HW) \\ (iii)} & \specialcell{Hispanic Father \\ Hispanic Mother \\ (HH) \\ (iv)} & \specialcell{HH - WW \\ (v)} & \specialcell{HW - WH \\ (vi)}\\
+Variables & \specialcell{White Father \\ White Mother \\ (WW) \\ (1)} & \specialcell{White Father \\ Hispanic Mother \\ (WH) \\ (2)} & \specialcell{Hispanic Father \\ White Mother \\ (HW) \\ (3)} & \specialcell{Hispanic Father \\ Hispanic Mother \\ (HH) \\ (4)} & \specialcell{HH - WW \\ (5)} & \specialcell{HW - WH \\ (6)}\\
 \midrule
-\textbf{Panel A: Parent’s} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
-\hspace{1em}Husband’seducation (Total Years) & \specialcell{13.05\\(2.44)} & \specialcell{12.32\\(3.33)} & \specialcell{10.65\\(4.39)} & \specialcell{8.93\\(4.41)} & \specialcell{-4.11**\\(0.02)} & \specialcell{-1.67**\\(0.04)}\\
-\hspace{1em}Wife’seducation (Total Years) & \specialcell{12.74\\(2.12)} & \specialcell{11.03\\(3.92)} & \specialcell{11.54\\(3.12)} & \specialcell{8.6\\(4.13)} & \specialcell{-4.13**\\(0.02)} & \specialcell{0.51**\\(0.04)}\\
+\textbf{Panel A: Synthetic Parents} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
+\hspace{1em}Husband's education (Total Years) & \specialcell{13.05\\(2.44)} & \specialcell{12.32\\(3.33)} & \specialcell{10.65\\(4.39)} & \specialcell{8.93\\(4.41)} & \specialcell{-4.11**\\(0.02)} & \specialcell{-1.67**\\(0.04)}\\
+\hspace{1em}Wife's education (Total Years) & \specialcell{12.74\\(2.12)} & \specialcell{11.03\\(3.92)} & \specialcell{11.54\\(3.12)} & \specialcell{8.6\\(4.13)} & \specialcell{-4.13**\\(0.02)} & \specialcell{0.51**\\(0.04)}\\
 \hspace{1em}Total Household seducation (Total Years) & \specialcell{25.78\\(4.08)} & \specialcell{23.35\\(6.51)} & \specialcell{22.19\\(6.69)} & \specialcell{17.54\\(7.83)} & \specialcell{-8.25**\\(0.03)} & \specialcell{-1.16*\\(0.07)}\\
-\textbf{Panel B: Education} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
+\textbf{Panel B: Children's Education} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
 \addlinespace
@@
 \hspace{1em}Women’s education (Total Years) & \specialcell{14.29\\(2.41)} & \specialcell{13.87\\(2.47)} & \specialcell{13.42\\(2.35)} & \specialcell{13.27\\(2.37)} & \specialcell{-1.01***\\(0.01)} & \specialcell{-0.46**\\(0.03)}\\
-\textbf{Panel C: Employment and Earnings} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
+\textbf{Panel C: Children's Employment and Earnings} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
 \hspace{1em}Men’s Employment Rate & \specialcell{0.95\\(0.22)} & \specialcell{0.94\\(0.23)} & \specialcell{0.92\\(0.26)} & \specialcell{0.93\\(0.26)} & \specialcell{-0.02***\\(0.00)} & \specialcell{-0.02***\\(0.00)}\\
@@
 \hspace{1em}Men’s Log Hourly Earnings & \specialcell{2.48\\(0.45)} & \specialcell{2.41\\(0.45)} & \specialcell{2.4\\(0.44)} & \specialcell{2.41\\(0.43)} & \specialcell{-0.07***\\(0.01)} & \specialcell{-0.00**\\(0.02)}\\
-\hspace{1em}Women’s Hourly Earnings & \specialcell{2.33\\(0.49)} & \specialcell{2.33\\(0.46)} & \specialcell{2.27\\(0.45)} & \specialcell{2.3\\(0.41)} & \specialcell{-0.02***\\(0.01)} & \specialcell{-0.06**\\(0.02)}\\
+\hspace{1em}Women’s Log Hourly Earnings & \specialcell{2.33\\(0.49)} & \specialcell{2.33\\(0.46)} & \specialcell{2.27\\(0.45)} & \specialcell{2.3\\(0.41)} & \specialcell{-0.02***\\(0.01)} & \specialcell{-0.06**\\(0.02)}\\
 \hspace{1em}Men’s Log Annual Earnings & \specialcell{10.25\\(1.01)} & \specialcell{10.08\\(1.05)} & \specialcell{10.04\\(1.06)} & \specialcell{10.01\\(1.04)} & \specialcell{-0.25**\\(0.01)} & \specialcell{-0.04**\\(0.03)}\\
-\hspace{1em}Women’s Hourly Earnings & \specialcell{9.46\\(1.78)} & \specialcell{9.54\\(1.64)} & \specialcell{9.47\\(1.57)} & \specialcell{9.53\\(1.52)} & \specialcell{0.07**\\(0.02)} & \specialcell{-0.07*\\(0.06)}\\
-\textbf{Panel D: Hispanic Identity} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
+\hspace{1em}Women’s Log Annual Earnings & \specialcell{9.46\\(1.78)} & \specialcell{9.54\\(1.64)} & \specialcell{9.47\\(1.57)} & \specialcell{9.53\\(1.52)} & \specialcell{0.07**\\(0.02)} & \specialcell{-0.07*\\(0.06)}\\
+\textbf{Panel D: Children's Hispanic Identity} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
 \addlinespace
@@
 \begin{tablenotes}
-\item[1] The data is restricted to native-born United States citizens between 1994 and 2019 who are also White and between the ages of 25 and 40.
+\item[1] Source: The 1970-1990 Census for synthetic parents, and 1994-2019 Current Population Surveys (CPS) for children's outcomes
+\item[2] The data is restricted to native-born United States citizens between 1994 and 2019 who are also White and between the ages of 25 and 40.
 I identify the ethnicity of a person's parents through the parent's place of birth.

### 2023-12-25 manuscripts/economic-inquiry/manuscript_main.tex
....

--- 2023-12-24 my_paper/Hadah-last-names-draft.tex
@@
-\documentclass[12pt,english]{article}
-
-\input{fragments/preamble_packages.tex}
-\input{fragments/preamble_layout.tex}
-\input{fragments/preamble_citation.tex}
-\bibliography{bib/bibliography}
-
-\makeatother
-
-\begin{document}
-
-\title{The Impact of Hispanic Last Names and Identity on Labor Market Outcomes}
-
-\author{\href{https://hussainhadah.com/}{Hussain Hadah} \thanks{Hussain Hadah: Department of Economics, Tulane University, 6823 St.
-Charles Ave., New Orleans, LA 70118, United States (e-mail: \href{mailto:hhadah@tulane.edu}{hhadah@tulane.edu}).
-I thank Professors Willa Friedman, Chinhui Juhn, Vikram Maheshri, and Yona Rubinstein for their support and advice.
-I also thank Aimee Chin, Steven Craig, German Cubas, Elaine Liu, Fan Wang, and the participants of the Applied Microeconomics Workshop at the University of Houston.}}
-
-\maketitle
-
-\begin{center}
-\href{https://hhadah.github.io/hispanic-last-names/my_paper/Hadah-last-names-draft.pdf}{\textcolor{green!15!black!30!blue}{\footnotesize{\textsc{Click Here to Get the Most Updated Version}}}}
-\end{center}
-
-\begin{abstract}
-\singlespacing Do people with ethnically Hispanic names face discrimination in the labor market?
-In this study, I analyze the causal impact of Hispanic-sounding surnames on wages, using the unique context of inter-ethnic marriages and the distinct characteristics of White and Hispanic surnames.
-We compare the wages of individuals with one White and one Hispanic parent, observing that those with Hispanic surnames often encounter wage disparities.
-Our findings show a significant wage difference favoring individuals with White surnames.
-Although males born to Hispanic fathers and White mothers earn 5 percentage points less than those born to White fathers and Hispanic mothers, this gap can be completely attributed to differences in education levels.
-However, this should not be interpreted as an absence of discrimination; rather, it potentially reflects discrimination in human capital accumulation.
-Additionally, I explore the impact of self-identifying as Hispanic on earnings.
-Men with a Spanish-sounding last name who identify as Hispanic earn significantly less than those who do not, but again, this gap can be largely explained by educational disparities.
-This suggests that discrimination may extend beyond the labor market, influencing educational opportunities and outcomes too.
-
-\noindent\textbf{Keywords}: Economics of Minorities, Race, and Immigrants; Discrimination and Prejudice
-
-\noindent\textbf{JEL Classification}: J71; J64; J15
-
-\end{abstract}
-
-\vfill
-\pagebreak{}
-
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Main Text
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Introduction}\label{sec:intro}
@@
 Consequently, further research is needed to comprehensively understand the earnings gaps between Hispanics and Whites.
-
-\pagebreak
-\begingroup
-\setstretch{1.0}
-%\setstretch{1.1}
-\setlength\bibitemsep{0pt}
-\printbibliography
-\endgroup
-\pagebreak
-
-\begin{appendices}
-% \section{Figures}\label{appendix:figs}
-
-% \begin{center}
-% \begin{figure}[H]
-% \caption{Histogram Before Targetting}
-% \includegraphics[width=\textwidth]{img/hist_dep_original.jpg} 
-% \label{fig:hist}
-% \end{figure}
-% \end{center}
-
-% \pagebreak
-
-% \begin{center}
-% \begin{figure}[h!]
-% \caption{Distribution of the four types of children.}
-% \includegraphics[width=\textwidth]{PirChart2.png} 
-% \label{fig:dist}
-% \end{figure}
-% \end{center}
-
-% \newpage
-
-\section{Tables}\label{appendix:tabs}
-
-\input{tables/tab01-observations-by-parents}
-
-\newpage
-
-\input{tables/tab03b-summary-stats}
-
-\newpage
-
-\input{tables/tab04-summary-stats}
-
-\newpage
-
-\input{tables/tab02-observations-by-couple}
-
-\newpage
-
-\input{tables/tab03a-summary-stats}
-
-\newpage
-
-\input{tables/tab03a-summary-stats-mex}
-
-\newpage
-
-\input{tables/tab03a-summary-stats-nonmex}
-
-\newpage
-
-\input{tables/tab05-regression}
-
-\newpage
-
-\input{tables/tab07-regression-weekearn}
-
-\newpage
-
-\input{tables/tab08-regression-weekearn-mex}
-
-\newpage
-
-\input{tables/tab09-regression-weekearn-nonmex}
-
-\newpage
-
-\input{tables/tab10-regression-weekearn-cuba}
-
-\newpage
-
-\input{tables/tab06-regression}
-
-\end{appendices}
-
-\end{document}

### 2023-12-25 programs/04-ACS-prep.r

@@ -1,0 +1,139 @@
+# prep ACS data to be merged with 
+# CPS data
+
+# table two
+
+# date: May 1st, 2023
+
+DATA <- read_dta(file.path(datasets,"ParentDataFull.dta")) |> 
+        mutate(BirthPlaceMom = case_when(bpld_mom  <= 12092  ~ "USA",
+           								 bpld_mom  == 11000  ~ "Puerto Rico",
+           								 bpld_mom  == 30005  ~ "Argentina",
+           								 bpld_mom  == 30010  ~ "Bolivia",
+		   								 bpld_mom  == 30015  ~ "Brazil",
+           								 bpld_mom  == 30020  ~ "Chile",
+           								 bpld_mom  == 30025  ~ "Colombia",
+           								 bpld_mom  == 26010  ~ "Dominican Republic",
+           								 bpld_mom  == 30030  ~ "Ecuador",
+           								 bpld_mom  == 21020  ~ "Costa Rica",
+           								 bpld_mom  == 21030  ~ "El Salvador",
+           								 bpld_mom  == 21040  ~ "Guatemala",
+           								 bpld_mom  == 21050  ~ "Honduras",
+           								 bpld_mom  == 25000  ~ "Cuba",
+           								 bpld_mom  == 20000  ~ "Mexico",
+           								 bpld_mom  == 21060  ~ "Nicaragua",
+           								 bpld_mom  == 21070  ~ "Panama",
+           								 bpld_mom  == 30050  ~ "Peru",
+           								 bpld_mom  == 30060  ~ "Uruguay",
+           								 bpld_mom  == 30065  ~ "Venezuela",
+           								 bpld_mom  == 43800  ~ "Spain",
+           								 bpld_mom  == 29900  ~ "Americas",
+		   								 bpld_mom  == 21000  ~ "Cental America",
+           								 bpld_mom  == 30000  ~ "South America"),
+               BirthPlaceDad = case_when(bpld_pop  <= 12092  ~ "USA",
+           								 bpld_pop  == 11000  ~ "Puerto Rico",
+           								 bpld_pop  == 30005  ~ "Argentina",
+           								 bpld_pop  == 30010  ~ "Bolivia",
+		   								 bpld_pop  == 30015  ~ "Brazil",
+           								 bpld_pop  == 30020  ~ "Chile",
+           								 bpld_pop  == 30025  ~ "Colombia",
[…]

### 2023-12-25 programs/11-table-one-parent-types.R

@@ -1,0 +1,173 @@
+# Table of observation number
+# of different types of parents
+
+# table one
+
+# date: January 19th, 2023
+
+DATA <- read_csv(file.path(datasets,"CPS_synth.csv"))
+
+# all observations
+
+ALL <- DATA %>%
+  summarise(n())
+# calculate the number 
+# of white fathers
+# white mothers
+
+WW <- DATA %>%
+  filter(Type == "White-White") %>%
+  summarise(n())
+# prettyNum(WW, big.mark = ",")
+WW_perc <- paste0(round(WW/ALL, digits = 2))
+
+# calculate the number 
+# of hispanic fathers
+# white mothers
+
+HW <- DATA %>%
+  filter(Type == "Hispanic-White") %>%
+  summarise(n())
+HW_perc <- paste0(round(HW/ALL, digits = 2))
+
+# calculate the number 
+# of white fathers
+# hispanic mothers
+
+WH <- DATA %>%
+  filter(Type == "White-Hispanic") %>%
+  summarise(n())
+WH_perc <- paste0(round(WH/ALL, digits = 2))
+
+# calculate the number 
+# of hispanic fathers
+# hispanic mothers
+
+HH <- DATA %>%
+  filter(Type == "Hispanic-Hispanic") %>%
+  summarise(n())
+HH_perc <- paste0(round(HH/ALL, digits = 2))
+
+# table
+
+first_col <- c(" ",
+                paste0("\\specialcell{", 
+                "Observations", 
+                      "\\\\",
+                "Share",
+                      "}"
+))
+
+second_col <- c("\\specialcell{White Father \\\\ White Mother}",
+               paste0("\\specialcell{", 
+                      prettyNum(WW, big.mark = ","), 
+                      "\\\\",
+                      WW_perc,
+                      "}"
+                      )
+)
+
+third_col <- c("\\specialcell{White Father \\\\ Hispanic Mother}",
+                paste0("\\specialcell{", 
+                       prettyNum(WH, big.mark = ","), 
+                       "\\\\",
+                       WH_perc,
+                       "}"
+                )
+)
+
+fourth_col <- c("\\specialcell{Hispanic Father \\\\ White Mother}",
+                paste0("\\specialcell{", 
+                       prettyNum(HW, big.mark = ","), 
+                       "\\\\",
+                       HW_perc,
+                       "}"
+                )
+)
+
+fifth_col <- c("\\specialcell{Hispanic Father \\\\ Hispanic Mother}",
+                paste0("\\specialcell{", 
[…]

### 2023-12-25 programs/111-add-health-data.R

@@ -1,0 +1,466 @@
+# Data cleaning and wrangling
+
+# date: May 14th, 2022
+
+# Open data
+
+# AddHealth <- read_dta(file.path(workdir,"AddHealth_Cleaned.dta"))
+# 
+# write_csv(AddHealth, file.path(workdir,"AddHealth_Cleaned.csv"))
+
+# AddHealth <- read_dta(file.path(workdir,"AddHealth_Cleaned.dta"))
+# write_csv(AddHealth, file.path(workdir,"AddHealth_Cleaned.csv"))
+AddHealth <- read_csv(file.path(add_health_data_dir,"AddHealth_Cleaned.csv"))
+
+####################
+# Data cleaning
+####################
+
+# Demographics
+AddHealth <- AddHealth %>%
+  mutate(SCHL_Sex = ifelse(SCHL_Sex==2,1,0),
+         Black_w1 = ifelse(Black_w1==1 & Hispanic_w1 == 0 & White_w1 == 0 & Native_w1 == 0 & Asian_w1 == 0 & Other_w1 == 0,1,0),
+         Black_w3 = ifelse(Black_w3==1 & Hispanic_w3 == 0 & White_w3 == 0 & Native_w3 == 0 & Asian_w3 == 0,1,0),
+         Other_w1 = ifelse(Other_w1==1 & Hispanic_w1 == 0  & White_w1 == 0 & Native_w1 == 0 & Asian_w1 == 0 & Black_w1 == 0,1,0),
+         White_w1 = ifelse(White_w1==1 & Hispanic_w1 == 0   & Other_w1 == 0 & Native_w1 == 0 & Asian_w1 == 0 & Black_w1 == 0,1,0),
+         Native_w1 = ifelse(Native_w1==1 & Hispanic_w1 == 0    & Other_w1 == 0 & White_w1 == 0 & Asian_w1 == 0 & Black_w1 == 0,1,0),
+         Asian_w1 = ifelse(Asian_w1==1 & Hispanic_w1 == 0     & Other_w1 == 0 & White_w1 == 0 & Native_w1 == 0 & Black_w1 == 0,1,0),
+         White_w3 = ifelse(White_w3==1 & Hispanic_w3 == 0  & Black_w3 == 0 & Native_w3 == 0 & Asian_w3 == 0,1,0),
+         Race_w1 = case_when(White_w1 == 1 ~"White",
+                             Black_w1 == 1 ~"Black",
+                             Other_w1 == 1 ~"Other",
[…]

### 2023-12-25 programs/12-table-two-couple-types.R

@@ -1,0 +1,132 @@
+# Table of observation number
+# of different types of parents
+
+# table two
+
+# date: January 19th, 2023
+
+DATA <- read_dta(file.path(datasets,"ParentDataFull.dta"))
+
+# all observations
+
+ALL <- DATA %>%
+  summarise(n())
+# calculate the number 
+# of white Husbands
+# white Wifes
+
+WW <- DATA %>%
+  filter(CoupleType == "White-White") %>%
+  summarise(n())
+# prettyNum(WW, big.mark = ",")
+WW_perc <- paste0(round(WW/ALL, digits = 2))
+
+# calculate the number 
+# of hispanic Husbands
+# white Wifes
+
+HW <- DATA %>%
+  filter(CoupleType == "Hispanic-White") %>%
+  summarise(n())
+HW_perc <- paste0(round(HW/ALL, digits = 2))
+
+# calculate the number 
+# of white Husbands
+# hispanic Wifes
+
+WH <- DATA %>%
+  filter(CoupleType == "White-Hispanic") %>%
+  summarise(n())
+WH_perc <- paste0(round(WH/ALL, digits = 2))
+
+# calculate the number 
+# of hispanic Husbands
+# hispanic Wifes
+
+HH <- DATA %>%
+  filter(CoupleType == "Hispanic-Hispanic") %>%
+  summarise(n())
+HH_perc <- paste0(round(HH/ALL, digits = 2))
+
+# table
+
+first_col <- c(" ",
+              paste0("\\specialcell{", 
+              "Observations", 
+                      "\\\\",
+                      "Share",
+                      "}"
+               )
+)
+
+second_col <- c("\\specialcell{White Husband \\\\ White Wife}",
+                paste0("\\specialcell{", 
+                       prettyNum(WW, big.mark = ","), 
+                       "\\\\",
+                       WW_perc,
+                       "}"
+                )
+)
+
+third_col <- c("\\specialcell{White Husband \\\\ Hispanic Wife}",
+               paste0("\\specialcell{", 
+                      prettyNum(WH, big.mark = ","), 
+                      "\\\\",
+                      WH_perc,
+                      "}"
+               )
+)
+
+fourth_col <- c("\\specialcell{Hispanic Husband \\\\ White Wife}",
+                paste0("\\specialcell{", 
+                       prettyNum(HW, big.mark = ","), 
+                       "\\\\",
+                       HW_perc,
+                       "}"
[…]

### 2023-12-25 programs/13-table-three-summary-stats.R

@@ -1,0 +1,534 @@
+# summary stats
+
+# table three
+
+# date: January 19th, 2023
+
+ParentData     <- read_dta(file.path(datasets,"ParentDataFull.dta")) |> 
+  mutate(Tot_ed = HusbandEducation + WifeEducation)
+
+# Only Mexicans and Native Born
+# ParentData <- ParentData |> 
+#        filter(BirthPlaceMom %in% c("Mexico", "USA") & BirthPlaceDad %in% c("Mexico", "USA"))
+
+# Only Non-Mexicans
+# `%notin%` <- Negate(`%in%`)
+# ParentData <- ParentData |> 
+#        filter(BirthPlaceMom %notin% c("Mexico") & BirthPlaceDad %notin% c("Mexico"))
+
+
+row1 <- c("Variables", 
+          "\\specialcell{White \\\\ White \\\\ (WW) \\\\ (1)}",
+          "\\specialcell{White \\\\ Hispanic \\\\ (WH) \\\\ (2)}",
+          "\\specialcell{Hispanic \\\\ White \\\\ (HW) \\\\ (3)}",
+          "\\specialcell{Hispanic \\\\ Hispanic \\\\ (HH) \\\\ (4)}",
+          "\\specialcell{HH - WW \\\\ (5)}",
+          "\\specialcell{HW - WH \\\\ (6)}"
+          )
+dim(row1) <- c(1,7)
+
+# parent's panel
+
+## Husband's Education
+HusbandEducation <- ParentData |> 
+  group_by(CoupleType) |> 
+  summarise(HusEducation   = mean(HusbandEducation, na.rm = T),
+            HusEducationSD = sd(HusbandEducation, na.rm = T))
+
+model <- lm(HusbandEducation ~ 0 + WW + WH + HW + HH, 
+    data = ParentData)
+test1 <- tidy(glht(model, linfct = c("HH - WW = 0")))
+pvalues1 <- test1$std.error
+differences1 <- test1$estimate
+
+test2 <- tidy(glht(model, linfct = c("HW - WH = 0")))
+pvalues2 <- test2$std.error
+differences2 <- test2$estimate
+
+row2 <- c(
+  "Husband's education (Total Years)",
+  paste0("\\specialcell{", 
+         sprintf("%0.2f",round(HusbandEducation[4,2], digits = 2)),  
+         "\\\\",
+         "(",
+         sprintf("%0.2f",round(HusbandEducation[4,3], digits = 2)), 
+         ")}"
+  ),
[…]

### 2023-12-25 programs/13a-table-three-summary-stats-mex.R

@@ -1,0 +1,530 @@
+# summary stats
+
+# table three
+
+# date: January 19th, 2023
+
+ParentData     <- read_dta(file.path(datasets,"ParentDataFull.dta")) |> 
+  mutate(Tot_ed = HusbandEducation + WifeEducation)
+
+# Only Mexicans and Native Born
+ParentData <- ParentData |> 
+       filter(BirthPlaceMom %in% c("Mexico", "USA") & BirthPlaceDad %in% c("Mexico", "USA"))
+
+row1 <- c("Variables", 
+          "\\specialcell{White \\\\ White \\\\ (WW) \\\\ (1)}",
+          "\\specialcell{White \\\\ Hispanic \\\\ (WH) \\\\ (2)}",
+          "\\specialcell{Hispanic \\\\ White \\\\ (HW) \\\\ (3)}",
+          "\\specialcell{Hispanic \\\\ Hispanic \\\\ (HH) \\\\ (4)}",
+          "\\specialcell{HH - WW \\\\ (5)}",
+          "\\specialcell{HW - WH \\\\ (6)}"
+          )
+dim(row1) <- c(1,7)
+
+# parent's panel
+
+## Husband's Education
+HusbandEducation <- ParentData |> 
+  group_by(CoupleType) |> 
+  summarise(HusEducation   = mean(HusbandEducation, na.rm = T),
+            HusEducationSD = sd(HusbandEducation, na.rm = T))
+
+model <- lm(HusbandEducation ~ 0 + WW + WH + HW + HH, 
+    data = ParentData)
+test1 <- tidy(glht(model, linfct = c("HH - WW = 0")))
+pvalues1 <- test1$std.error
+differences1 <- test1$estimate
+
+test2 <- tidy(glht(model, linfct = c("HW - WH = 0")))
+pvalues2 <- test2$std.error
+differences2 <- test2$estimate
+
+row2 <- c(
+  "Husband's education (Total Years)",
+  paste0("\\specialcell{", 
+         sprintf("%0.2f",round(HusbandEducation[4,2], digits = 2)),  
+         "\\\\",
+         "(",
+         sprintf("%0.2f",round(HusbandEducation[4,3], digits = 2)), 
+         ")}"
+  ),
+  paste0("\\specialcell{", 
+         sprintf("%0.2f",round(HusbandEducation[3,2], digits = 2)),  
+         "\\\\",
+         "(",
[…]

### 2023-12-25 programs/13b-table-three-summary-stats-nonmex.R

@@ -1,0 +1,532 @@
+# summary stats
+
+# table three
+
+# date: January 19th, 2023
+
+ParentData     <- read_dta(file.path(datasets,"ParentDataFull.dta")) |> 
+  mutate(Tot_ed = HusbandEducation + WifeEducation)
+
+# Only Non-Mexicans
+`%notin%` <- Negate(`%in%`)
+ParentData <- ParentData |> 
+       filter(BirthPlaceMom %notin% c("Mexico") & BirthPlaceDad %notin% c("Mexico"))
+
+
+row1 <- c("Variables", 
+          "\\specialcell{White \\\\ White \\\\ (WW) \\\\ (1)}",
+          "\\specialcell{White \\\\ Hispanic \\\\ (WH) \\\\ (2)}",
+          "\\specialcell{Hispanic \\\\ White \\\\ (HW) \\\\ (3)}",
+          "\\specialcell{Hispanic \\\\ Hispanic \\\\ (HH) \\\\ (4)}",
+          "\\specialcell{HH - WW \\\\ (5)}",
+          "\\specialcell{HW - WH \\\\ (6)}"
+          )
+dim(row1) <- c(1,7)
+
+# parent's panel
+
+## Husband's Education
+HusbandEducation <- ParentData |> 
+  group_by(CoupleType) |> 
+  summarise(HusEducation   = mean(HusbandEducation, na.rm = T),
+            HusEducationSD = sd(HusbandEducation, na.rm = T))
+
+model <- lm(HusbandEducation ~ 0 + WW + WH + HW + HH, 
+    data = ParentData)
+test1 <- tidy(glht(model, linfct = c("HH - WW = 0")))
+pvalues1 <- test1$std.error
+differences1 <- test1$estimate
+
+test2 <- tidy(glht(model, linfct = c("HW - WH = 0")))
+pvalues2 <- test2$std.error
+differences2 <- test2$estimate
+
+row2 <- c(
+  "Husband's education (Total Years)",
+  paste0("\\specialcell{", 
+         sprintf("%0.2f",round(HusbandEducation[4,2], digits = 2)),  
+         "\\\\",
+         "(",
+         sprintf("%0.2f",round(HusbandEducation[4,3], digits = 2)), 
+         ")}"
+  ),
+  paste0("\\specialcell{", 
+         sprintf("%0.2f",round(HusbandEducation[3,2], digits = 2)),  
+         "\\\\",
+         "(",
[…]

### 2023-12-25 programs/13c-table-three-summary-stats-spanish.R

@@ -1,0 +1,400 @@
+# summary stats
+
+# table three
+
+# date: January 19th, 2023
+
+ParentData     <- read_dta(file.path(datasets,"ParentDataFull.dta")) |> 
+  mutate(Tot_ed = HusbandEducation + WifeEducation)
+
+# Only Mexicans and Native Born
+ParentData <- ParentData |> 
+       filter(BirthPlaceMom %in% c("Spain", "USA") & BirthPlaceDad %in% c("Spain", "USA"))
+
+row1 <- c("Variables", 
+          "\\specialcell{White \\\\ White \\\\ (WW) \\\\ (1)}",
+          "\\specialcell{White \\\\ Spanish \\\\ (WS) \\\\ (2)}",
+          "\\specialcell{Spanish \\\\ White \\\\ (SW) \\\\ (3)}",
+          "\\specialcell{Spanish \\\\ Spanish \\\\ (SS) \\\\ (4)}",
+          "\\specialcell{SS - WW \\\\ (5)}",
+          "\\specialcell{SW - WS \\\\ (6)}"
+          )
+dim(row1) <- c(1,7)
+
+# parent's panel
+
+## Husband's Education
+HusbandEducation <- ParentData |> 
+  group_by(CoupleType) |> 
+  summarise(HusEducation   = mean(HusbandEducation, na.rm = T),
+            HusEducationSD = sd(HusbandEducation, na.rm = T))
+
+model <- lm(HusbandEducation ~ 0 + WW + WH + HW + HH, 
+    data = ParentData)
+test1 <- tidy(glht(model, linfct = c("HH - WW = 0")))
+pvalues1 <- test1$std.error
+differences1 <- test1$estimate
+
+test2 <- tidy(glht(model, linfct = c("HW - WH = 0")))
+pvalues2 <- test2$std.error
+differences2 <- test2$estimate
+
+row3 <- c(
+  "Husband's education (Total Years)",
+  paste0("\\specialcell{", 
+         sprintf("%0.2f",round(HusbandEducation[4,2], digits = 2)),  
+         "\\\\",
+         "(",
+         sprintf("%0.2f",round(HusbandEducation[4,3], digits = 2)), 
+         ")}"
+  ),
+  paste0("\\specialcell{", 
+         sprintf("%0.2f",round(HusbandEducation[3,2], digits = 2)),  
+         "\\\\",
+         "(",
+         sprintf("%0.2f",round(HusbandEducation[3,3], digits = 2)), 
[…]

### 2023-12-25 programs/14-table-three-summary-stats.R

@@ -1,0 +1,717 @@
+# summary stats
+
+# table three
+
+# library(Hmisc)
+
+# date: January 19th, 2023
+
+IndividualData <- read_csv(file.path(datasets,"CPS_synth.csv")) |> 
+  mutate(Education = as.numeric(Education)) |> 
+  filter(WW == 1 | (WH == 1 ) | (HW == 1 ) | (HH == 1 ))
+
+CrossTable(IndividualData$Type, IndividualData$Hispanic_ID)
+row1 <- c("Variables", 
+          "\\specialcell{White \\\\ White \\\\ (WW) \\\\ (1)}",
+          "\\specialcell{White \\\\ Hispanic \\\\ (WH) \\\\ (2)}",
+          "\\specialcell{Hispanic \\\\ White \\\\ (HW) \\\\ (3)}",
+          "\\specialcell{Hispanic \\\\ Hispanic \\\\ (HH) \\\\ (4)}",
+          "\\specialcell{HH - WW \\\\ (5)}",
+          "\\specialcell{HW - WH \\\\ (6)}"
+          )
+dim(row1) <- c(1,7)
+
+# Panel A
+# Age and Education
+row6 <- c("Panel A: Children's Education ", " ", " ", " ", " ", " ", " ")
+dim(row6) <- c(1,7)
+
+## Man's Education
+ManEducation <- IndividualData |> 
+  filter(sex == 1) |> 
+  group_by(Type) |> 
+  summarise(EducationMean   = mean(Education, na.rm = T),
+            EducationSD = sd(Education, na.rm = T))
+
+model <- lm(Education ~ 0 + WW + WH + HW + HH, 
+             data = IndividualData |> filter(sex == 1))
+test1 <- tidy(glht(model, linfct = c("HH - WW = 0")))
+pvalues1 <- test1$std.error
+differences1 <- test1$estimate
+
+test2 <- tidy(glht(model, linfct = c("HW - WH = 0")))
+pvalues2 <- test2$std.error
+differences2 <- test2$estimate
+
+row7 <- c(
+  "Men’s education (Total Years)",
+  paste0("\\specialcell{", 
+         round(ManEducation[4,2], digits = 2), 
+         "\\\\",
+         "(",
+         round(ManEducation[4,3], digits = 2),
+         ")}"
+  ),
+  paste0("\\specialcell{", 
+         round(ManEducation[3,2], digits = 2), 
+         "\\\\",
+         "(",
+         round(ManEducation[3,3], digits = 2),
[…]

### 2023-12-25 programs/17-table-six-regression.R

@@ -1,0 +1,219 @@
+# summary stats
+
+# table Five
+# Regression
+
+# date: January 19th, 2023
+
+# open data, filter to only have the relevant data
+# and impute empty parental charachtersitcs with mean
+
+IndividualData <- read_csv(file.path(datasets,"CPS_synth.csv")) |> 
+  mutate(Education = as.numeric(Education)) |> 
+  filter((WW == 1 & Hispanic_ID == 0) | WH == 1 | HW == 1 | HH == 1)|> 
+  mutate(HusbandEducation = case_when(!is.na(HusbandEducation) ~ HusbandEducation,
+                                       is.na(HusbandEducation) ~ mean(HusbandEducation, na.rm = TRUE)),
+         WifeEducation = case_when(!is.na(WifeEducation) ~ WifeEducation,
+                                    is.na(WifeEducation) ~ mean(WifeEducation, na.rm = TRUE)),
+          Husband_ftotval = case_when(!is.na(Husband_ftotval) ~ Husband_ftotval,
+                                       is.na(Husband_ftotval) ~ mean(Husband_ftotval, na.rm = TRUE)))
+
+# create a total education variable
+# and a income quintile variable
+# and education quintile variable
+IndividualData  <- IndividualData  |> 
+  mutate(TotalHHEducation = HusbandEducation + WifeEducation,
+         IncomeQuint      = cut(Husband_ftotval, breaks = quantile(Husband_ftotval, probs = c(0, 0.2, 0.4, 0.6, 0.8, 1), na.rm = TRUE), include.lowest = TRUE)
+          )
+IndividualData  <- IndividualData  |> 
+  mutate(EducationQuint      = cut(TotalHHEducation, breaks = quantile(TotalHHEducation, probs = c(0, 0.2, 0.4, 0.6, 0.8, 1), na.rm = TRUE), include.lowest = TRUE)
+          )
+
+ParentDummies <- c(#"WH:Hispanic_ID", 
+                   "HW:Hispanic_ID", 
+                   "Hispanic_ID",
+                   #"HH:Hispanic_ID",
+                   #"WH", 
+                   "HW", 
+                   #"HH", 
+                   "uhrsworkly")
+ParentControls  <- c("EducationQuint", "IncomeQuint")
+
+IndividualData |> 
+  filter(sex == 1 & FTFY == 1 & Self_employed == 0 & (HW == 1 | WH == 1)) |> 
[…]

### 2023-12-25 programs/19-table-eight-regression.R

@@ -1,0 +1,118 @@
+# summary stats
+
+# table Five
+# Regression
+
+# date: January 19th, 2023
+
+IndividualData <- read_dta(file.path(datasets,"BySexAnalysisData.dta")) |> 
+  mutate(Education = as.numeric(Education)) |> 
+  filter((WW == 1 & Hispanic_ID == 0) | WH == 1 | HW == 1 | HH == 1)
+
+ParentDummies <- c("WH", "HW", "HH")
+
+# By generation
+reg1 <- list(
+  "\\specialcell{(1) \\\\ Education}" = feols(Education ~ .[ParentDummies] | year + age, se = "iid",
+                                                        data = IndividualData))
+
+
+# calculate means to add
+# as a row
+
+
+test1 <- tidy(glht(reg1[[1]], linfct = c("HW - WH = 0")))
+differences1 <- case_when(abs(test1$estimate/test1$std.error)   >= 2.58  ~ paste0(sprintf("%2.5s",round_tidy(test1$estimate, digits = 2)), "***"),
+                          abs(test1$estimate/test1$std.error)   <  2.58 & 
+                            abs(test1$estimate/test1$std.error) >= 1.96  ~ paste0(sprintf("%2.5s",round_tidy(test1$estimate, digits = 2)), "**"),
+                          abs(test1$estimate/test1$std.error)   <  1.96  &
+                            abs(test1$estimate/test1$std.error) >= 1.645 ~ paste0(sprintf("%2.5s",round_tidy(test1$estimate, digits = 2)), "*"),
+                          abs(test1$estimate/test1$std.error)   < 1.645  ~ round_tidy(test1$estimate, digits = 2)
+)
+
+
+
+differences_row <-  c('$HW_{i} - WH_{i}$', differences1)
+dim(differences_row) <- c(1,2)
+
+pvalue_row <-  c(' ', paste0("(", round(test1$std.error, digits = 2)))
+dim(pvalue_row) <- c(1,2)
+
+row_diff <- rbind(differences_row, pvalue_row)
+colnames(row_diff)<-LETTERS[1:2]
+
+row_diff <- as.data.frame(row_diff)
+attr(row_diff, 'position') <- c(7:8)
+
+controling_for <-  c("\\textit{Controlling for:}", " ")
+dim(controling_for) <- c(1,2)
+
+
+controling_for <- as.data.frame(controling_for)
+
[…]

### 2023-12-25 programs/22-table-nine-summary-stats-synth-grands.R

@@ -1,0 +1,920 @@
+# summary stats
+
+# table three
+
+library(Hmisc)
+
+# date: January 19th, 2023
+
+IndividualData <- read_dta(file.path(datasets,"BySexAnalysisData.dta")) |> 
+  mutate(Education = as.numeric(Education))
+ParentData     <- read_dta(file.path(datasets,"ParentDataFull.dta")) |> 
+  mutate(Tot_ed = HusbandEducation + WifeEducation)
+
+row1 <- c("Variables", 
+          "\\specialcell{White \\\\ White \\\\ (WW) \\\\ (1)}",
+          "\\specialcell{White \\\\ Hispanic \\\\ (WH) \\\\ (2)}",
+          "\\specialcell{Hispanic \\\\ White \\\\ (HW) \\\\ (3)}",
+          "\\specialcell{Hispanic \\\\ Hispanic \\\\ (HH) \\\\ (4)}",
+          "\\specialcell{HH - WW \\\\ (5)}",
+          "\\specialcell{HW - WH \\\\ (6)}"
+          )
+dim(row1) <- c(1,7)
+row2 <- c("Panel A: Synthetic Parents", " ", " ", " ", " ", " ", " ")
+dim(row2) <- c(1,7)
+
+# parent's panel
+
+## Husband's Education
+HusbandEducation <- ParentData |> 
+  group_by(CoupleType) |> 
+  summarise(HusEducation   = weighted.mean(HusbandEducation, na.rm = T, hhwt),
+            HusEducationSD = weighted.sd(HusbandEducation, na.rm = T, hhwt))
+
+model <- lm(HusbandEducation ~ 0 + WW + WM + MW + MM, 
+    data = ParentData, 
+    weights = hhwt)
+test1 <- tidy(glht(model, linfct = c("MM - WW = 0")))
+pvalues1 <- test1$std.error
+differences1 <- test1$estimate
+
+test2 <- tidy(glht(model, linfct = c("MW - WM = 0")))
+pvalues2 <- test2$std.error
+differences2 <- test2$estimate
+
+row3 <- c(
+  "Husband's education (Total Years)",
+  paste0("\\specialcell{", 
+         round(HusbandEducation[4,2], digits = 2), 
+         "\\\\",
+         "(",
+         round(HusbandEducation[4,3], digits = 2),
+         ")}"
+  ),
+  paste0("\\specialcell{", 
+         round(HusbandEducation[3,2], digits = 2), 
[…]

### 2023-12-25 programs/95-make-all.R

@@ -1,0 +1,55 @@
+#######################################################################
+# Master script
+#######################################################################
+
+## Clear out existing environment
+gc()
+rm(list = ls()) 
+
+mdir <- "~/Dropbox/Research/My Research Data and Ideas/hispanic-last-names"
+datasets <- paste0(mdir,"/data/datasets")
+raw <- paste0(mdir,"/data/raw")
+
+mdir_git     <- "~/Documents/GiT/hispanic-last-names"
+datasets_git <- paste0(mdir_git,"/data/datasets") 
+raw_git      <- paste0(mdir_git,"/data/raw") 
+tables_wd    <- paste0(mdir_git,"/output/tables") 
+figures_wd   <- paste0(mdir_git,"/output/figures") 
+programs     <- paste0(mdir_git,"/programs") 
+thesis_tabs  <- paste0(mdir_git,"/my_paper/tables") 
+thesis_plots <- paste0(mdir_git,"/my_paper/figure") 
+add_health_data_dir <- "/Users/hhadah/Dropbox/Research/My Research Data and Ideas/Identification_Paper/Data/Datasets"
+ACS_data  <- "/Users/hhadah/Documents/GiT/jmp-decomposition-replication/data/raw"
+
+# pres_tabs <- paste0(mdir,"/presentations/jmp-presentation/tables")
+# pres_plots <- paste0(mdir,"/presentations/jmp-presentation/figures")
+
+# dissertation_wd <- paste0("/Users/hhadah/Documents/GiT/my_thesis/presentation/dissertation")
+### run do files and scripts
+
+# main scripts
+source(file.path(programs,"01-packages-wds.r")) # set up package
+# stata(file.path(programs,"02-ACS-Synt-Parents-DoFile.do")) # run ACS do file
+# stata(file.path(programs,"03-CleaningDataDoFile.do")) # clean data
+source(file.path(programs,"04-ACS-prep.r")) # prep data
+# stata(file.path(programs,"05-CPS-Do-File.do")) # CPS Do File
+# stata(file.path(programs,"06-CPS-DataCleanUpScript.do")) # CPS data cleaning
+# stata(file.path(programs,"07-BySexAnalysis.do")) # by sex analysis
+# stata(file.path(programs,"08-CoupleAnalysis.do")) # couple analysis
+# stata(file.path(programs,"09-Regressions.do")) # regressions
+source(file.path(programs,"10-merge-cps-acs.r")) # merge cps and acs
+source(file.path(programs,"11-table-one-parent-types.R")) # table one
[…]

### 2023-12-28 my_paper/Hadah-last-names-draft.tex
....
edit abstract

@@
 In this study, I analyze the causal impact of Hispanic-sounding surnames on wages, using the unique context of inter-ethnic marriages and the distinct characteristics of White and Hispanic surnames.
-We compare the wages of individuals with one White and one Hispanic parent, observing that those with Hispanic surnames often encounter wage disparities.
-Our findings show a significant wage difference favoring individuals with White surnames.
+I compare the wages of individuals with one White and one Hispanic parent, observing that those with Hispanic surnames often encounter wage disparities.
+The findings in this paper show a significant wage difference favoring individuals with White surnames.
 Although males born to Hispanic fathers and White mothers earn 5 percentage points less than those born to White fathers and Hispanic mothers, this gap can be completely attributed to differences in education levels.
-However, this should not be interpreted as an absence of discrimination; rather, it potentially reflects discrimination in human capital accumulation.
+However, this should not be interpreted as an absence of discrimination; rather, it potentially reflects discrimination in access to education.
 Additionally, I explore the impact of self-identifying as Hispanic on earnings.
@@
 
+\end{abstract}
+
 \noindent\textbf{Keywords}: Economics of Minorities, Race, and Immigrants; Discrimination and Prejudice
@@
 \noindent\textbf{JEL Classification}: J71; J64; J15
-
-\end{abstract}
 

### 2024-01-15 manuscripts/economic-inquiry/manuscript_main.tex
submission to EI

@@
 
-The main identifying assumption of my empirical strategy depends on the assumption that people born to Hispanic fathers and White mothers are similar to their White fathers Hispanic mothers peer in all observable aspects and characteristics that are important in the labor market.
+The main identifying assumption of my empirical strategy depends on the assumption that people born to HW parents are similar to their WH peer in all observable aspects and characteristics that are important in the labor market.
 Consequently, the only difference between the two groups is the variation in which group is more likely to have a Hispanic sounding last name.
@@
 
-Using the Current Population Survey (CPS) from 1994 to 2019 and the 1960 to 2000 US censuses, I estimate the effect of having a Hispanic last name on labor market outcomes.
+Using the CPS from 1994 to 2019 and the 1960 to 2000 US censuses, I estimate the effect of having a Hispanic last name on labor market outcomes.
 In other words, I approximate the variation in labor market perception of ethnic signals, in this case having a Hispanic last name, on annual log earnings.
@@
 $HW_{ist}$ is an indicator variables for the type of parents of person $i$ has.
-$X_{ist}$ is a vector of controls that includes age and numbers of hours worked, $\gamma_{t}$ are year fixed effects, $\lambda_{s}$ are state fixed effects, and $\phi_{ist}$ represents the error term.
+$X_{ist}$ is a vector of controls that includes age and numbers of hours worked, $\gamma_{t}$ are year fixed effects (FE), $\lambda_{s}$ are state FE, and $\phi_{ist}$ represents the error term.
 The equation for this strategy is written as follows:
@@
 
-The central assumption underpinning my estimation strategy is based on the hypothesis that individuals born to a Hispanic father and White mother exhibit comparable characteristics to their peers of White-Hispanic descent, especially in areas vital to the labor market.
+The central assumption underpinning my estimation strategy is based on the hypothesis that individuals born to a HW person exhibits comparable characteristics to their peers of WH descent, especially in areas vital to the labor market.
 This assumption includes similarities in educational background, skill sets, and work experiences that are significant determinants in employment opportunities, salary levels, and career advancement.
@@
 This table meticulously lays out the average outcomes and discrepancies for each group, highlighting the impact of inter-ethnic marriages on children's prospects.
-These results show that there is selection in marriage, and that the differences between children born to Hispanic fathers and White mother (HW) and children born to White fathers and Hispanic mother (WH) are less severe that those between the other two groups.
+These results show that there is selection in marriage, and that the differences between children born to HW and children born to WH are less severe that those between the other two groups.
 Consequently, comparing WH and HW children to each other to analyze discrimination against Hispanics in the labor market.
@@
 I provide the results to the estimation of equation \ref{eq:1a} in Table \ref{tab:lastnamereg}.
-I estimate the mean wages of White, U.S.-born, Hispanic men aged 25-40 who are employed full-time and full-year (FTFY) as waged and salaried workers.
+I estimate the mean wages of White, U.S.-born, Hispanic men aged 25-40 who are employed FTFY as waged and salaried workers.
 I also restrict the sample to children of HW and WH (omitted) parents.
@@
 I provide the results to the estimation of equation \ref{eq:iden} in Table \ref{tab:identreg}.
-The sample in this analysis includes full-time, full-year, and wage and salary men, and I control for hours worked and age and include year fixed effects (FE).
+The sample in this analysis includes full-time, full-year, and wage and salary men, and I control for hours worked and age and include year FE.
 I also restrict the sample to children of HW and WH parents.

### 2024-01-15 manuscripts/economic-inquiry/manuscript_title_etc.tex
submission to EI

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\newcommand{\PAPERTITLE}{\href{https://hhadah.github.io/hispanic-last-names/my_paper/Hadah-last-names-draft.pdf}{The Impact of Hispanic Last Names and Identity on Labor Market Outcomes}}
+% \newcommand{\PAPERTITLE}{\href{https://hhadah.github.io/hispanic-last-names/my_paper/Hadah-last-names-draft.pdf}{The Impact of Hispanic Last Names and Identity on Labor Market Outcomes}}
+\newcommand{\PAPERTITLE}{The Impact of Hispanic Last Names and Identity on Labor Market Outcomes}
 
@@
 \newcommand{\PAPERABSTRACT}{
+% This study examines if Hispanic-sounding surnames lead to labor market discrimination.
+Analyzing wages within inter-ethnic marriages, it finds significant wage disparities disadvantaging individuals with Hispanic surnames compared to those with White surnames.
+Males with Hispanic fathers and White mothers earn less than their counterparts with White fathers and Hispanic mothers, primarily due to educational differences.
+However, this may indicate discrimination in education rather than an absence of labor market bias.
+Additionally, Hispanic-identifying men with Hispanic surnames face earnings gaps, largely explainable by education.
 Do people with ethnically Hispanic names face discrimination in the labor market?

### 2024-02-11 manuscripts/ilr-review/submit_journal-cover.tex
ILR Review submission latex doc

\documentclass[12pt,english]{article}

\input{preamble_package}
% Citation comes loading other packages, so that footnotes follow biblatex-chicago style
\input{preamble_citation}

\begin{document}

\thispagestyle{empty}
\begingroup
  \doublespacing
  \centering
  \LARGE Submission to ILR Review \\[0.25em]
  % \LARGE ACCEPTED MANUSCRIPT: MS2020XXX \\[1.0em]
\endgroup
I am including in this document:
\begin{enumerate}
    \item Title page information
    \begin{itemize}
        \item Title and Author
        \item Abstract
        \item Keywords
    \end{itemize}
    \item The manuscript
    \begin{itemize}
        \item Main manuscript text
        \item Main manuscript references
        \item Main manuscript figures
        \item Main manuscript tables
    \end{itemize}    
    \item The online appendix, which is separately paged.
\end{enumerate}

\section*{Acknowledgment}
I thank Professors Willa Friedman, Chinhui Juhn, Vikram Maheshri, and Yona Rubinstein for their support and advice. I also thank Aimee Chin, Steven Craig, German Cubas, Elaine Liu, Fan Wang, and the participants of the Applied Microeconomics Workshop at the University of Houston, and European Society for Population Economics (ESPE) for helpful feedback.

\section*{Data and Programs}

The data that support the findings of this study will be openly available to all researchers after the review process. For immediate information regarding the data and/or computer programs used for this study, please contact Hussain Hadah at \href{mailto:hhadah@tulane.edu}{hhadah@tulane.edu}.

\clearpage

\end{document}

### 2024-05-18 manuscripts/ilr-review copy/manuscript_main.tex
create new table

--- 2024-01-15 manuscripts/economic-inquiry/manuscript_main.tex
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Introduction}\label{sec:intro}
 
@@
 
-Using the CPS from 1994 to 2019 and the 1960 to 2000 US censuses, I estimate the effect of having a Hispanic last name on labor market outcomes.
+% Using the CPS from 1994 to 2019 and the 1960 to 2000 US censuses, I estimate the effect of having a Hispanic last name on labor market outcomes.
 In other words, I approximate the variation in labor market perception of ethnic signals, in this case having a Hispanic last name, on annual log earnings.
@@
 
-I find that while men born to Hispanic father-White mothers earn less than men born to White father-Hispanic mothers, this gap is entirely explained by educational differences.
+% I find that while men born to Hispanic father-White mothers earn less than men born to White father-Hispanic mothers, this gap is entirely explained by educational differences.
 I find that inter-ethnic children with Hispanic sounding last name earn 5 percentage points less than those without a Hispanic last name.
@@
 
-The rest of the paper is organized as follows.
-Section \ref{sec:data} provides an overview of the data and summary statistics of the sample.
-Section \ref{sec:emp_model} introduces the empirical framework.
-I report the results in section \ref{sec:results}.
-Finally, I offer a brief conclusion in section \ref{sec:con1}.
-
-\section{Data}\label{sec:data}
+\section*{Data}\label{sec:data}
 
@@
 
-\subsection{Children of the four parental types}
+\subsection*{Children of the four parental types}
 
@@
 
-% On the one hand, WW children are more educated, with approximately 14 years of education for men and women.
-WW children have an unemployment rate of 4\% for men and women.
-They have log hourly earnings of 2.51 for men and 2.32 for women.
-Also, their annual log earnings are 10.29 for men and 9.43 for women.
-
-% On the other hand, children of HH marriages have an average of 12.9 and 13.24 years of education for men and women, respectively.
-HH men and women have an unemployment rate of 7\% and 6\%.
-Their annual log earnings are 10.01 and 9.53 for men and women, respectively.
-
-% Inter-ethnic children, WH and HW, have lower educational attainment and income than WW children, but they are better than HH children.
-Also, WH and HW are comparable, unlike WW and HH children.
-These two groups are important to the analysis.
-Children of HW are going to be the treatment group.
-HW children have a Hispanic father with a Hispanic last names.
-At the same time, the WH children have White fathers with a non-Hispanic, or White, last names.
-Summary statistics of these two groups, Table \ref{tab:c&p1} columns (2) and (3) can show that they are different from the children of non-inter-ethnic marriages.
-
-% WH children have an average of 13.58 and 13.79 years of education for men and women.
-The unemployment rate for men and women is 5\%.
-HW men and women obtained 13.22 and 13.42 years of schooling.
-Men's unemployment rate is 7\% and women's 6\%.
-
-% In panel C of Table \ref{tab:c&p1}, I present the rates at which members of each group self-identify as Hispanic.
-\footnote{Hispanic self-identification is observed when an individual notes that they are Hispanic on the CPS questionnaire} The vast majority of HH children, 96\% for men and 97\% for women, identify as Hispanic.
-There is considerable attrition in Hispanic identification of inter-ethnic children.
-Among WH children, 74\% of men and 78\% of women identify as Hispanic.
-Among HW children, 83\% of men and 82\% of women identify as Hispanic.
-These results align with the findings of a substantial body of literature that noted this attrition among children of immigrants, especially those coming from inter-ethnic marriages \autocite{duncan2017complexity, duncan2018identifying, duncan2020new, antman2020ethnic}.
-Finally, a small portion of WW children, 4\% of men and 5\% of women, identified as Hispanic.
-These 5\% of WW children are most likely third generation--- or higher--- children of immigrants.
-For my analysis, I dropped the WW observations who identify as Hispanic when I compare with non-Hispanic Whites.
-
-% In columns (5) and (6), I calculate the differences in means between HH-WW (column 5) and HW-WH (column 6).
-Overall, HH children do worse than WW children.
-HH has one year less of education compared to WW, and are 2 percentage points more likely to be unemployed.
-HH men earn 9\% less than WW men, and HH women make 2\% less than WW women.
-Therefore, these two groups, HH and WW, are different on many observable qualities.
-
-% The differentials between WH and HW are not as severe as the ones between HH and WW.
-HW children are still worse off than WH but are much better than HH.
-HW men attain 0.36 fewer years of education than WH men.
-HW women achieve 0.37 fewer years of schooling than WH women.
-HW children are 1 percentage points more likely to be unemployed.
-HW men earn 1\% less annual earnings as WH men, and HW women earn 3\% less.
-These differences show that HW and WH children are different from children of endogamous/homogamous marriages and, thus, are more comparable.
-
-% In Table \ref{tab:c&p2}, I present the summary statistics for the four types and limit the sample to those that only identify as Hispanic.\footnote{Limiting the sample to those who identify as Hispanic only affects WH, HW, and HH.
-Members of the WW group are not affected by this restriction.} The same pattern discussed above between HH--WW and HW--WH persists.
-The one difference is that those who identify as Hispanic do slightly worse.
-
-\subsection{Synthetic parents}
+\subsection*{Synthetic parents}
 
@@
 
-\section{Empirical Strategy}\label{sec:emp_model}
+\section*{Empirical Strategy}\label{sec:emp_model}
 
@@
 
-\subsection{Estimating the effect of having a Hispanic last name}
+\subsection*{Estimating the effect of having a Hispanic last name}
 
@@
 
-\subsection{Estimating the effect of identifying as Hispanic}
+\subsection*{Estimating the effect of identifying as Hispanic}
 
@@
 
-\subsection{Threats to Identification}
+\subsection*{Threats to Identification}
 
@@
 
-\section{From the Data: The Differences Between HW and WH Couples}\label{sec:hw-wh-couples-data}
+\section*{From the Data: The Differences Between HW and WH Couples}\label{sec:hw-wh-couples-data}
 
@@
 
-\section{Results}\label{sec:results}
+\section*{Results}\label{sec:results}
 
@@
 
-\subsection{The Effect of Having a Hispanic Last Name and Hispanic Identity on Labor Market Outcomes}
+\subsection*{The Effect of Having a Hispanic Last Name and Hispanic Identity on Labor Market Outcomes}
 
@@
 
-\section{Conclusion}\label{sec:con1}
+\section*{Conclusion}\label{sec:con1}
 

### 2024-05-18 output/tables/tab11-measurement-error.tex
create new table

\begin{table}[H]
\centering\centering
\caption{Self-reported Hispanic Identity Among First-Generation Hispanic Immigrants and Native-Born \label{tab:mat2}}
\centering
\resizebox{\ifdim\width>\linewidth\linewidth\else\width\fi}{!}{
\begin{threeparttable}
\begin{tabular}[t]{>{}lcccc}
\toprule
  & \specialcell{Native Born Husband} & \specialcell{Spanish-Speaking \\ Place of Birth \\ Husband} & \specialcell{Native Born Wife} & \specialcell{Spanish-Speaking \\ Place of Birth \\ Wife}\\
\midrule
\textbf{\specialcell{Percent White\\Percent Hispanic}} & \specialcell{0.97\\0.03} & \specialcell{0.03\\0.97} & \specialcell{0.97\\0.03} & \specialcell{0.03\\0.97}\\
\bottomrule
\end{tabular}
\begin{tablenotes}
\item[1] Source: 1960-2000 Census
\item[2] The sample includes Whites, who are married, and are between the ages 25 and 40.
\end{tablenotes}
\end{threeparttable}}
\end{table}

### 2024-05-18 programs/24-hispanic-firstgen-parents.R

@@ -1,0 +1,178 @@
+# summary stats
+
+# table about first
+# generation immigrants
+# Native born
+# and Hisapnic identity
+
+# date: January 19th, 2023
+
+ParentData     <- read_dta(file.path(datasets,"ParentDataFull.dta")) |> 
+  mutate(Tot_ed = HusbandEducation + WifeEducation)
+
+CrossTable(ParentData$LAHusband, ParentData$Hispanic_Husband)
+CrossTable(ParentData$LAWife, ParentData$Hispanic_Wife)
+
+# Spanish-speaking country
+# place of birth husbands
+# in America
+
+LA_Husbands <- ParentData %>%
+    filter(LAHusband == 1) %>%
+    summarise(n())
+
+# Spanish-speaking country
+# place of birth husbands
+# in America
+
+LA_Hispanic_Husbands <- ParentData %>%
+    filter(LAHusband == 1 & Hispanic_Husband == 1) %>%
+    summarise(n())
+
+# Percent Spanish-speaking country
+# and Hispanic husbands
+
+LA_Hispanic_Husbands_perc <- as.numeric(paste0(round(LA_Hispanic_Husbands/LA_Husbands, digits = 2)))
+
+# Native born husbands
+# in America
+
+NA_Husbands <- ParentData %>%
+    filter(LAHusband == 0) %>%
+    summarise(n())
+
+# Native born and Hispanic
+# husbands
+
+NA_Hispanic_Husbands <- ParentData %>%
+    filter(LAHusband == 0 & Hispanic_Husband == 1) %>%
+    summarise(n())
+
+# Percent Native born and Hispanic
+# husbands
+
+NA_Hispanic_Husbands_perc <- as.numeric(paste0(round(NA_Hispanic_Husbands/NA_Husbands, digits = 2)))
+
+# Spanish-speaking country
+# place of birth wives
+# in America
+
+LA_Wifes <- ParentData %>%
+    filter(LAWife == 1) %>%
+    summarise(n())
+
+# Spanish-speaking country
+# place of birth wives
+# in America
+
+LA_Hispanic_Wifes <- ParentData %>%
+    filter(LAWife == 1 & Hispanic_Wife == 1) %>%
+    summarise(n())
+
+# Percent Spanish-speaking country
+# and Hispanic wives
+
[…]

### 2024-05-20 manuscripts/latest-revisions/manuscript_title_etc.tex
JPAM edits

--- 2023-12-25 manuscripts/economic-inquiry/manuscript_title_etc.tex
@@
 I thank Professors Willa Friedman, Chinhui Juhn, Vikram Maheshri, and Yona Rubinstein for their support and advice.
-I also thank Aimee Chin, Steven Craig, German Cubas, Elaine Liu, Fan HADAH, and the participants of the Applied Microeconomics Workshop at the University of Houston, and European Society for Population Economics (ESPE) for helpful feedback.}
+I also thank Aimee Chin, Steven Craig, German Cubas, Elaine Liu, Fan Wang, and the participants of the Applied Microeconomics Workshop at the University of Houston, and European Society for Population Economics (ESPE) for helpful feedback.}
 
@@
 \newcommand{\PAPERABSTRACT}{
-Do people with ethnically Hispanic names face discrimination in the labor market?
-In this study, I analyze the causal impact of Hispanic-sounding surnames on wages, using the unique context of inter-ethnic marriages and the distinct characteristics of White and Hispanic surnames.
-We compare the wages of individuals with one White and one Hispanic parent, observing that those with Hispanic surnames often encounter wage disparities.
-Our findings show a significant wage difference favoring individuals with White surnames.
-Although males born to Hispanic fathers and White mothers earn 5 percentage points less than those born to White fathers and Hispanic mothers, this gap can be completely attributed to differences in education levels.
-However, this should not be interpreted as an absence of discrimination; rather, it potentially reflects discrimination in human capital accumulation.
-Additionally, I explore the impact of self-identifying as Hispanic on earnings.
-Men with a Spanish-sounding last name who identify as Hispanic earn significantly less than those who do not, but again, this gap can be largely explained by educational disparities.
-This suggests that discrimination may extend beyond the labor market, influencing educational opportunities and outcomes too.
+Do individuals with ethnically Hispanic names face labor market discrimination?
+In this study, I analyze the impact of Hispanic-sounding surnames on wages, considering inter-ethnic children and the distinct traits of native and Hispanic surnames.
+I examine the earnings of individuals with one White and one Hispanic parent, finding that those with Hispanic surnames often experience wage disparities.
+My findings reveal a notable wage gap favoring individuals with White surnames.
+Although males born to Hispanic fathers and White mothers earn 5 percentage points less than those born to White fathers and Hispanic mothers, this difference can be attributed to educational variances.
+However, this shouldn't imply an absence of discrimination; it may reflect discrimination in human capital accumulation.
+Furthermore, I investigate the impact of self-identifying as Hispanic on earnings.
+Men with a Spanish-sounding last name who identify as Hispanic earn notably less than those who do not, largely due to educational differences.
 \PAPERJEL}

### 2024-05-20 manuscripts/latest-revisions/submit_journal-title-abstract.tex
JPAM edits

\documentclass[12pt,english]{article}

\input{preamble_package}
% Citation comes loading other packages, so that footnotes follow biblatex-chicago style
\input{preamble_citation}

\begin{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Part I. Title Page, Authors, Abstract, etc. 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Load in title page information, title, author information, abstract, etc are stored here
\input{manuscript_title_etc}
\pagenumbering{roman}
\setcounter{page}{1}

% Part I.a.1 Title
\section*{Title and Author}
\subsection*{Title}
\begin{itemize}[label={}, leftmargin=*]
    \item Title: \textbf{\PAPERTITLE}
\end{itemize}
% Part I.a.2 Author info
\subsection*{Author}
\begin{itemize}[label={}, leftmargin=*]
    \item List of authors:
    \begin{enumerate}
        \item \AUTHORHADAHINFO
    \end{enumerate}
    \item Corresponding author:
    \begin{itemize}
        \item \textbf{\AUTHORHADAHINFO}
    \end{itemize}
\end{itemize}
\clearpage

\subsection*{Data Availability Statement}
The data that support the findings of this study will be openly available to all researchers after the review process. For immediate information regarding the data and/or computer programs used for this study, please contact Hussain Hadah at hhadah@tulane.edu.

\subsection*{Funding Statement}

The author has no funding for the research, authorship, and/or publication of this article to report.

\subsection*{Conflict of Interest Disclosure}
The author declare that they have no conflicts of interest regarding the publication of this manuscript.

\clearpage

\end{document}

### 2024-05-20 my_paper/Hadah-last-names-draft.tex
edit abstract
JPAM edits

@@
 I compare the wages of individuals with one White and one Hispanic parent, observing that those with Hispanic surnames often encounter wage disparities.
-The findings in this paper show a significant wage difference favoring individuals with White surnames.
+The findings in this paper show a significant wage difference favoring individuals with White or native sounding surnames.
 Although males born to Hispanic fathers and White mothers earn 5 percentage points less than those born to White fathers and Hispanic mothers, this gap can be completely attributed to differences in education levels.
@@
 I use the CPS for my primary analysis of the effect of having a Hispanic last name on earnings.
-I restrict my sample to Whites, United States-born citizens aged 25 to 40 and born between 1960 to 2000.
+I restrict my sample to people that racially identified as White, United States-born citizens aged 25 to 40 and born between 1960 to 2000.
 Taking advantage of data on parents' place of birth, I divide the sample into four groups depending on their parent's ethnicity.
@@
 Moreover, the summary statistics for those that also identify as Hispanic is reported in in Table \ref{tab:c&p2}.
-
-% On the one hand, WW children are more educated, with approximately 14 years of education for men and women.
-WW children have an unemployment rate of 4\% for men and women.
-They have log hourly earnings of 2.51 for men and 2.32 for women.
-Also, their annual log earnings are 10.29 for men and 9.43 for women.
-
-% On the other hand, children of HH marriages have an average of 12.9 and 13.24 years of education for men and women, respectively.
-HH men and women have an unemployment rate of 7\% and 6\%.
-Their annual log earnings are 10.01 and 9.53 for men and women, respectively.
-
-% Inter-ethnic children, WH and HW, have lower educational attainment and income than WW children, but they are better than HH children.
-Also, WH and HW are comparable, unlike WW and HH children.
-These two groups are important to the analysis.
-Children of HW are going to be the treatment group.
-HW children have a Hispanic father with a Hispanic last names.
-At the same time, the WH children have White fathers with a non-Hispanic, or White, last names.
-Summary statistics of these two groups, Table \ref{tab:c&p1} columns (2) and (3) can show that they are different from the children of non-inter-ethnic marriages.
-
-% WH children have an average of 13.58 and 13.79 years of education for men and women.
-The unemployment rate for men and women is 5\%.
-HW men and women obtained 13.22 and 13.42 years of schooling.
-Men's unemployment rate is 7\% and women's 6\%.
-
-% In panel C of Table \ref{tab:c&p1}, I present the rates at which members of each group self-identify as Hispanic.
-\footnote{Hispanic self-identification is observed when an individual notes that they are Hispanic on the CPS questionnaire} The vast majority of HH children, 96\% for men and 97\% for women, identify as Hispanic.
-There is considerable attrition in Hispanic identification of inter-ethnic children.
-Among WH children, 74\% of men and 78\% of women identify as Hispanic.
-Among HW children, 83\% of men and 82\% of women identify as Hispanic.
-These results align with the findings of a substantial body of literature that noted this attrition among children of immigrants, especially those coming from inter-ethnic marriages \autocite{duncan2017complexity, duncan2018identifying, duncan2020new, antman2020ethnic}.
-Finally, a small portion of WW children, 4\% of men and 5\% of women, identified as Hispanic.
-These 5\% of WW children are most likely third generation--- or higher--- children of immigrants.
-For my analysis, I dropped the WW observations who identify as Hispanic when I compare with non-Hispanic Whites.
-
-% In columns (5) and (6), I calculate the differences in means between HH-WW (column 5) and HW-WH (column 6).
-Overall, HH children do worse than WW children.
-HH has one year less of education compared to WW, and are 2 percentage points more likely to be unemployed.
-HH men earn 9\% less than WW men, and HH women make 2\% less than WW women.
-Therefore, these two groups, HH and WW, are different on many observable qualities.
-
-% The differentials between WH and HW are not as severe as the ones between HH and WW.
-HW children are still worse off than WH but are much better than HH.
-HW men attain 0.36 fewer years of education than WH men.
-HW women achieve 0.37 fewer years of schooling than WH women.
-HW children are 1 percentage points more likely to be unemployed.
-HW men earn 1\% less annual earnings as WH men, and HW women earn 3\% less.
-These differences show that HW and WH children are different from children of endogamous/homogamous marriages and, thus, are more comparable.
-
-% In Table \ref{tab:c&p2}, I present the summary statistics for the four types and limit the sample to those that only identify as Hispanic.\footnote{Limiting the sample to those who identify as Hispanic only affects WH, HW, and HH.
-Members of the WW group are not affected by this restriction.} The same pattern discussed above between HH--WW and HW--WH persists.
-The one difference is that those who identify as Hispanic do slightly worse.
 
@@
 I will go over the empirical evidence from the data to show that this is actually the case.
+
+Another threat to identification could arise from measurement error in using the place of birth of parents in the CPS data as a proxy for a parent's ethnicity and last name.
+The CPS only asks about the place of birth of parents and not their ethnic or racial identity.
+Therefore, it is possible that a native-born father could be a second-generation or later immigrant from a Spanish-speaking country.
+However, this scenario is highly improbable.
+Most Hispanics from 1960 to 2000 were first-generation immigrants, and the number of second-generation or later was very small.
+This is supported by data from the 1960 to 2000 censuses (see Table \ref{tab:mat3}).
+Only 3\% of native-born Americans identified as Hispanic, indicating the small number of second-generation or later Hispanic immigrants in the sample.
+Therefore, the probability that an inter-ethnic child with a native-born father that is also a second generation+ Hispanic immigrant is very small.
 
@@
 
+\input{tables/tab11-measurement-error}
+
+\newpage
+
 \input{tables/tab03a-summary-stats-mex}

### 2024-05-20 programs/24-hispanic-firstgen-parents.R

@@ -164,7 +164,7 @@ knitr::kable(Table_cols, "html", align = "lcccc",
 knitr::kable(Table_cols, "latex", align = "lcccc",
              booktabs = T,
              escape = F,
-             caption = "Self-reported Hispanic Identity Among First-Generation Hispanic Immigrants and Native-Born \\label{tab:mat2}") %>%
+             caption = "Self-reported Hispanic Identity Among First-Generation Hispanic Immigrants and Native-Born \\label{tab:mat3}") %>%
   column_spec(1, bold = T) %>%
   kable_classic(full_width = F) %>%
   kable_styling(bootstrap_options = c("hover", "condensed"), 

### 2024-08-08 manuscripts/labour-v5/manuscript_main.tex
LABOUR Comments

--- 2024-05-18 manuscripts/ilr-review copy/manuscript_main.tex
@@
 
-The main identifying assumption of my empirical strategy depends on the assumption that people born to HW parents are similar to their WH peer in all observable aspects and characteristics that are important in the labor market.
+The main identifying assumption of my empirical strategy depends on the assumption that people born to HW parents are similar to their WH peer in all observable---and unobservable---aspects and characteristics that are important in the labor market.
 Consequently, the only difference between the two groups is the variation in which group is more likely to have a Hispanic sounding last name.
@@
 
-% Using the CPS from 1994 to 2019 and the 1960 to 2000 US censuses, I estimate the effect of having a Hispanic last name on labor market outcomes.
-In other words, I approximate the variation in labor market perception of ethnic signals, in this case having a Hispanic last name, on annual log earnings.
-I also study the effect of identifying as Hispanic on labor market outcomes.
+Moreover, a large literature of audit studies in education economics investigated the effect of racial and ethnic signal on access to education.
+\textcite{bergman2018education} found that students with Hispanic sounding names received 2 percentage points fewer responses from schools than students with White sounding names.
+\textcite{janssen2022guidance} found that guidance counselors restricted Asian students from advanced opportunities.
+\textcite{gaddis2024racial} found significant discrimination in interactions with principals against Hispanic and Chinese American families.
+Finally, \textcite{bourabain2023school} found more evidence of discrimination in access to education against students with underprivileged backgrounds in the Flemish education system.
+This paper contributes to this literature by providing evidence that discrimination in access to education could lead to lower earnings for Hispanics.
 
-% I find that while men born to Hispanic father-White mothers earn less than men born to White father-Hispanic mothers, this gap is entirely explained by educational differences.
-I find that inter-ethnic children with Hispanic sounding last name earn 5 percentage points less than those without a Hispanic last name.
-The gap becomes statistically insignificant after controlling for education.
-Thus, I do not find a significant effect of having a Hispanic last name.
-Finally, I find that men identifying as Hispanic earn significantly less than those who do not, even when I control for ancestry and education.
+The rest of the paper is organized as follows.
+In Section \ref{sec:data}, I describe the data used in this paper.
+In Section \ref{sec:emp_model}, I present the empirical strategy.
+In Section \ref{sec:results}, I present the results from the estimation of the two specifications.
+Finally, in Section \ref{sec:con1}, I conclude.
 
-\section*{Data}\label{sec:data}
+\section{Data}\label{sec:data}
 
@@
 
-\section*{Empirical Strategy}\label{sec:emp_model}
+\section{Empirical Strategy}\label{sec:emp_model}
 
@@
 
+Another threat to identification could arise from measurement error in using the place of birth of parents in the CPS data as a proxy for a parent's ethnicity and last name.
+The CPS only asks about the place of birth of parents and not their ethnic or racial identity.
+Therefore, it is possible that a native-born father could be a second-generation or later immigrant from a Spanish-speaking country.
+However, this scenario is highly improbable.
+Most Hispanics from 1960 to 2000 were first-generation immigrants, and the number of second-generation or later was very small.
+This is supported by data from the 1960 to 2000 censuses (see Table \ref{tab:mat3}).
+Only 3\% of native-born Americans identified as Hispanic, indicating the small number of second-generation or later Hispanic immigrants in the sample.
+Therefore, the probability that an inter-ethnic child with a native-born father that is also a second generation+ Hispanic immigrant is very small.
+
 \section*{From the Data: The Differences Between HW and WH Couples}\label{sec:hw-wh-couples-data}
@@
 More importantly for human capital accumulation, the mother in the HW marriages---the children that have a Spanish sounding last-name---are more educated than their WH peers.
-Since mothers could be more important in the a child's human capital accumulation and education, this would suggest that a child with a Hispanic last-name should complete more years of education.
+Since mothers could be more important in the a child's human capital accumulation and education, this would suggest that a child with a Hispanic last-name should complete more years of education \autocite{gould2020does}.
 
@@
 
-\section*{Results}\label{sec:results}
+\section{Results}\label{sec:results}
 
@@
 
-\section*{Conclusion}\label{sec:con1}
+There is a large literature in education economics that shows disparities in access to education.
+In an audit study of charter and traditional public schools, \textcite{bergman2018education,gaddis2024racial} find that students with Hispanic names---compared to students that are presumed White---are less likely to get a response to inquiries from schools.
+This could be an indication that people with Hispanic sounding names could end up lower access to education, including high value-added schools.
+Therefore, the earnings gap between people with Hispanic sounding last names and those with native sounding last names could be due to differences in access to education that are in turn due to discrimination.
+
+\section{Conclusion}\label{sec:con1}
 
@@
 While the earnings gap between children of inter-ethnic parents with and without a Hispanic last name disappears when controlling for education, it does not necessarily indicate the absence of discrimination.
-Education itself can be influenced by bias, potentially resulting in divergent outcomes.
+Education itself can be influenced by bias, potentially resulting in divergent outcomes \autocite{bergman2018education,gaddis2024racial}.
 This is especially the case when parental characteristics indicate that people with a Hispanic-sounding last-names should in theory complete more years of education---since mothers of inter-ethnic children with a Hispanic last-names have more years of education and earn more.

### 2024-08-08 manuscripts/labour-v5/manuscript_title_etc.tex
LABOUR Comments

--- 2024-05-20 manuscripts/latest-revisions/manuscript_title_etc.tex
@@
 \newcommand{\PAPERABSTRACT}{
-Do individuals with ethnically Hispanic names face labor market discrimination?
+% Do individuals with ethnically Hispanic names face labor market discrimination?
 In this study, I analyze the impact of Hispanic-sounding surnames on wages, considering inter-ethnic children and the distinct traits of native and Hispanic surnames.
@@
 Men with a Spanish-sounding last name who identify as Hispanic earn notably less than those who do not, largely due to educational differences.
+
+Do individuals with Hispanic names face labor market discrimination?
+This study analyzes the impact of Hispanic-sounding surnames on wages among inter-ethnic children with one White and one Hispanic parent.
+I find that individuals with Hispanic surnames often earn less, with a notable wage gap favoring those with White surnames.
+Males born to Hispanic fathers and White mothers earn 5 percentage points less than those born to White fathers and Hispanic mothers, largely due to educational differences.
+Additionally, men with Spanish-sounding last names who identify as Hispanic earn significantly less, again due to educational disparities.
 \PAPERJEL}

### 2024-08-11 manuscripts/labour-v5/manuscript_main.tex
new edits

@@
 In this section, I present two empirical strategies.
-The first empirical strategy estimates the effect of having a Hispanic last name on annual log earnings.
+The first empirical strategy estimates the effect of having a Hispanic last name on educational outcomes and earnings.
 The second empirical strategy estimates the effect of identifying as Hispanic on log annual earnings.
@@
 In this section I restrict the sample to WH and HW groups.
-Let $Y_{ist}$ be the annual log earnings of person $i$ in state $s$ at time $t$.
+Let $Y_{ist}$ be the outcome of interest for person $i$ in state $s$ at time $t$.
 $HW_{ist}$ is an indicator variables for the type of parents of person $i$ has.
-$X_{ist}$ is a vector of controls that includes age and numbers of hours worked, $\gamma_{t}$ are year fixed effects (FE), $\lambda_{s}$ are state FE, and $\phi_{ist}$ represents the error term.
+$X_{ist}$ is a vector of controls that includes age and numbers of hours worked, $\gamma_{st}$ are state-year fixed effects (FE), and $\phi_{ist}$ represents the error term.
 The equation for this strategy is written as follows:
 \begin{equation} \label{eq:1a}
-Y_{ist} = \beta_{1} HW_{ist} + X_{ist} \pi + \gamma_{t} + \lambda_s + \phi_{ist}
+Y_{ist} = \beta_{1} HW_{ist} + X_{ist} \pi + \gamma_{st} + \phi_{ist}
 \end{equation}
@@
 $\beta_{1}$ is the coefficient of interest in this specification.
-$\beta_{1}$ represents the earnings gaps between children of inter-ethnic marriages who have a Spanish-sounding last name versus a White last name.
+$\beta_{1}$ represents the gaps in outcomes between children of inter-ethnic marriages who have a Spanish-sounding last name versus a White last name.
 
@@
 \begin{equation} \label{eq:iden}
-Y_{ist} = \beta_{1} HW_{ist} + \beta_{2} Hispanic_i + \beta_{3} HW_{ist} \cdot Hispanic_i + X_{ist} \pi + \gamma_{t} + \lambda_s + \phi_{ist}
+Y_{ist} = \beta_{1} HW_{ist} + \beta_{2} Hispanic_i + \beta_{3} HW_{ist} \cdot Hispanic_i + X_{ist} \pi + \gamma_{st} + \phi_{ist}
 \end{equation}
@@
 In this section, I present the results from estimating the two specifications presented in equations \ref{eq:1a} and \ref{eq:iden}.
-I estimate the mean wages of White, U.S.-born, Hispanic men aged 25-40 who are employed full-time and full-year (FTFY) as waged and salaried workers.
+I estimate the mean educational outcomes of White, Native-born, Hispanics aged 25-40.
+I estimate the mean wages of White, Native-born, Hispanic men aged 25-40 who are employed full-time and full-year (FTFY) as waged and salaried workers.
 The results are shown in Tables \ref{tab:lastnamereg} and \ref{tab:identreg}.
 
-I find that an inter-ethnic person with a Spanish-sounding last names earns less than an inter-ethnic who has a White last name.
+First, I find that an inter-ethnic person with a Spanish-sounding last names receive less total years of education than an inter-ethnic who has a White last name.
+A person with a Spanish-sounding last name receives 0.2 years of education less than a person with a White last name, have the same high school dropout rate as a person with a White last name, and 2 percentage points less likely to have an associate degree.
+People with a Spanish-sounding last name are 3 percentage points less likely to have a bachelor's degree.
+
+Second, I find that an inter-ethnic person with a Spanish-sounding last names earns less than an inter-ethnic who has a White last name.
 A person with a Spanish-sounding last name earns 5 percentage points less than a person with a White last name.
@@
 I also find a significant earnings gap between those that identify as Hispanic.
+
+\subsection*{The Effect of Having a Hispanic Last Name on Educational Outcomes}
+
+I provide the results of the estimation of equation \ref{eq:1a} in Table \ref{tab:lastname-ed-reg}.
+I estimate the mean educational outcomes of White, U.S.-born Hispanics ages 25-40.
+I also restrict the sample to children of HW and WH parents.
+The omitted group is children of WH parents.
+Column one in Table \ref{tab:lastname-ed-reg} is the difference in total years of education between HW children and their WH peers.
+Column two is the difference in the probability of having been a high school dropout.
+Column three is the difference in the probability of having an associate degree.
+Column four is the difference in the probability of having a bachelor's degree.
+All regressions include controls for age, parental education and income, and state-year fixed effects.
+
+Overall, there is a significant gap in total years of education between HW and WH children.
+HW children receive 0.2 years of education less than WH children.
+The gap between HW and WH women is larger than the gap between HW and WH men.
+Women with a Hispanic last name receive 0.25 years of education less than WH women.
+The gap between HW and WH men is equal to 0.16 years.
+
+Even though there is a significant, albeit small, gap in total years of education between HW and WH children, there is no significant gap in the probability of being a high school dropout.
+The gap between HW and WH high school dropouts is equal to a statistically insignificant 1 percentage point.
+The same is true for HW women (2 percentage points) and men (0 percentage points).
+
+However, there is a notable difference when it comes to higher education outcomes.
+HW children are 2 percentage points less likely to earn an associate degree and 3 percentage points less likely to earn a bachelor's degree compared to their WH peers.
+These gaps are slightly larger for HW women, who are 3 percentage points less likely to earn an associate degree and 4 percentage points less likely to earn a bachelor's degree.
+For HW men, the gap is 2 percentage points for an associate degree and statistically insignificant for a bachelor's degree.
+
+These results suggest that while the overall educational attainment gap between HW and WH children is small in terms of years of education, the disparities become more pronounced when considering higher education milestones, particularly for HW women.
+However, given the fact that HWs have ,more educated mothers than WHs, we would expect them to have higher levels of educational attainment \autocite{kimball2009risk, gould2020does}.
+This could indicate potential barriers or discrimination in access to higher education for HW children, particularly.
 
@@
 
-There is a large literature in education economics that shows disparities in access to education.
-In an audit study of charter and traditional public schools, \textcite{bergman2018education,gaddis2024racial} find that students with Hispanic names---compared to students that are presumed White---are less likely to get a response to inquiries from schools.
-This could be an indication that people with Hispanic sounding names could end up lower access to education, including high value-added schools.
-Therefore, the earnings gap between people with Hispanic sounding last names and those with native sounding last names could be due to differences in access to education that are in turn due to discrimination.
-
 \section{Conclusion}\label{sec:con1}
@@
 
-I compare the children of inter-ethnic marriages to study the labor market impact of having a Hispanic last name.
+I compare the children of inter-ethnic marriages to study the labor and educational markets impact of having a Hispanic last name.
+I find that inter-ethnic people with Spanish-sounding last names receive 0.2 years of education less than inter-ethnic counterparts that have a native-sounding last name.
 When I compare the earnings of HW and WH children, which captures the effect of having a Hispanic last name, HW children earn  6 percentage points less than WH children.
@@
 
+There is a large literature in education economics that shows disparities in access to education, which this paper contributes to.
+In an audit study of charter and traditional public schools, \textcite{bergman2018education,gaddis2024racial} find that students with Hispanic names---compared to students that are presumed White---are less likely to get a response to inquiries from schools.
+This could be an indication that people with Hispanic sounding names could end up lower access to education, including high value-added schools.
+Therefore, the earnings gap between people with Hispanic sounding last names and those with native sounding last names could be due to differences in access to education that are in turn due to discrimination.
+This paper provides further evidence, using observational data, that discrimination in access to education could lead to lower earnings for Hispanics.
+
 While the earnings gap between children of inter-ethnic parents with and without a Hispanic last name disappears when controlling for education, it does not necessarily indicate the absence of discrimination.

### 2024-08-11 manuscripts/labour-v5/manuscript_title_etc.tex
new edits

@@
 % \newcommand{\PAPERTITLE}{\href{https://hhadah.github.io/hispanic-last-names/my_paper/Hadah-last-names-draft.pdf}{The Impact of Hispanic Last Names and Identity on Labor Market Outcomes}}
-\newcommand{\PAPERTITLE}{The Impact of Hispanic Last Names and Identity on Labor Market Outcomes}
+\newcommand{\PAPERTITLE}{The Impact of Hispanic Last Names and Identity on Educational and Labor Market Outcomes}
 
@@
 I find that individuals with Hispanic surnames often earn less, with a notable wage gap favoring those with White surnames.
+People born to Hispanic fathers and White mothers receive 0.3 years of education less than those born to White fathers and Hispanic mothers.
 Males born to Hispanic fathers and White mothers earn 5 percentage points less than those born to White fathers and Hispanic mothers, largely due to educational differences.

### 2024-08-11 manuscripts/labour-v5/tab04-summary-stats.tex
new edits

--- 2023-05-09 my_paper/tables/tab04-summary-stats.tex
@@
 \begin{TableNotes}
-\item[1] Source: The 1950-2000 Census for synthetic parents, and 1994-2019 Current Population Surveys (CPS) for children's outcomes
+\item[1] Source: The 1994-2019 Current Population Surveys (CPS) for children's outcomes
 \item[2] The data is restricted to native-born United States citizens between 1994 and 2019 who are also White and between the ages of 25 and 40.
+The summary statistics for White-Hispanic (WH), Hispanic-White (HW), and Hispanic-Hispanic (HH) are restricted to the sample of children that self-identify as Hispanic or Latino.
 I identify the ethnicity of a person's parents through the parent's place of birth.
@@
 \begin{longtable}[t]{>{\raggedright\arraybackslash}p{5cm}cccccc}
-\caption{Summary statistics of outcomes using parent's place of birth only for those that self-identify as Hispanic \label{tab:c&p2}}\\
+\caption{Summary Statistics of Outcomes Using Parent's Place of Birth Only for Those That Self-Identify as Hispanic \label{tab:c&p2}}\\
 \toprule
@@
 \endfirsthead
-\caption[]{Summary statistics of outcomes using parent's place of birth only for those that self-identify as Hispanic  \textit{(continued)}}\\
+\caption[]{Summary Statistics of Outcomes Using Parent's Place of Birth Only for Those That Self-Identify as Hispanic  \textit{(continued)}}\\
 \toprule
@@
 \midrule
-\multicolumn{7}{r@{}}{\textit{(Continued on Next Page...)}}\
+\multicolumn{7}{r@{}}{\textit{(Continued on Next Page...)}}\\
 \endfoot
@@
 \textbf{Panel A: Children's Education} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
-\hspace{1em}Men’s education (Total Years) & \specialcell{13.82\\(2.42)} & \specialcell{13.4\\(2.38)} & \specialcell{13.08\\(2.27)} & \specialcell{12.87\\(2.3)} & \specialcell{-0.94***\\(0.01)} & \specialcell{-0.32**\\(0.03)}\\
+\hspace{1em}Men’s education (Total Years) & \specialcell{13.82\\(2.42)} & \specialcell{13.4\\(2.38)} & \specialcell{13.07\\(2.27)} & \specialcell{12.87\\(2.3)} & \specialcell{-0.94***\\(0.01)} & \specialcell{-0.32**\\(0.03)}\\
 \hspace{1em}Women’s education (Total Years) & \specialcell{14.06\\(2.37)} & \specialcell{13.62\\(2.39)} & \specialcell{13.26\\(2.35)} & \specialcell{13.22\\(2.39)} & \specialcell{-0.84***\\(0.01)} & \specialcell{-0.36**\\(0.02)}\\
+\hspace{1em}Men’s High School Dropout Rate & \specialcell{0.35\\(0.48)} & \specialcell{0.45\\(0.5)} & \specialcell{0.44\\(0.5)} & \specialcell{0.44\\(0.5)} & \specialcell{0.09***\\(0)} & \specialcell{-0.01***\\(0.01)}\\
+\hspace{1em}Women’s High School Dropout Rate & \specialcell{0.36\\(0.48)} & \specialcell{0.45\\(0.5)} & \specialcell{0.46\\(0.5)} & \specialcell{0.46\\(0.5)} & \specialcell{0.1***\\(0)} & \specialcell{0.01***\\(0.01)}\\
+\addlinespace
+\hspace{1em}Men’s Associate Degree Rate & \specialcell{0.15\\(0.36)} & \specialcell{0.15\\(0.36)} & \specialcell{0.12\\(0.32)} & \specialcell{0.11\\(0.32)} & \specialcell{-0.03***\\(0)} & \specialcell{-0.03***\\(0)}\\
+\hspace{1em}Women’s Associate Degree Rate & \specialcell{0.19\\(0.39)} & \specialcell{0.17\\(0.37)} & \specialcell{0.14\\(0.35)} & \specialcell{0.15\\(0.36)} & \specialcell{-0.04***\\(0)} & \specialcell{-0.03***\\(0)}\\
+\hspace{1em}Men’s BA Degree Rate & \specialcell{0.32\\(0.47)} & \specialcell{0.23\\(0.42)} & \specialcell{0.19\\(0.39)} & \specialcell{0.17\\(0.38)} & \specialcell{-0.15***\\(0)} & \specialcell{-0.04***\\(0.01)}\\
+\hspace{1em}Women’s BA Degree Rate & \specialcell{0.36\\(0.48)} & \specialcell{0.28\\(0.45)} & \specialcell{0.22\\(0.41)} & \specialcell{0.22\\(0.41)} & \specialcell{-0.14***\\(0)} & \specialcell{-0.06***\\(0)}\\
+\hspace{1em}Men’s Professional Degree Rate & \specialcell{0.09\\(0.28)} & \specialcell{0.07\\(0.25)} & \specialcell{0.05\\(0.21)} & \specialcell{0.04\\(0.19)} & \specialcell{-0.05***\\(0)} & \specialcell{-0.02***\\(0)}\\
+\addlinespace
+\hspace{1em}Women’s Professional Degree Rate & \specialcell{0.1\\(0.3)} & \specialcell{0.08\\(0.26)} & \specialcell{0.07\\(0.25)} & \specialcell{0.06\\(0.23)} & \specialcell{-0.04***\\(0)} & \specialcell{-0.01***\\(0)}\\
 \textbf{Panel B: Children's Employment and Earnings} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
 \hspace{1em}Men’s Unemployment Rate & \specialcell{0.04\\(0.8)} & \specialcell{0.06\\(0.76)} & \specialcell{0.07\\(0.74)} & \specialcell{0.07\\(0.75)} & \specialcell{0.02***\\(0.00)} & \specialcell{0.01***\\(0.00)}\\
-\addlinespace
 \hspace{1em}Women’s Unemployment Rate & \specialcell{0.04\\(0.81)} & \specialcell{0.06\\(0.24)} & \specialcell{0.07\\(0.75)} & \specialcell{0.06\\(0.76)} & \specialcell{0.02***\\(0.00)} & \specialcell{0.01***\\(0.00)}\\
 \hspace{1em}Men’s Log Hourly Earnings & \specialcell{2.51\\(0.45)} & \specialcell{2.43\\(0.46)} & \specialcell{2.42\\(0.44)} & \specialcell{2.42\\(0.43)} & \specialcell{-0.09***\\(0)} & \specialcell{-0.00**\\(0.02)}\\
+\addlinespace
 \hspace{1em}Women’s Log Hourly Earnings & \specialcell{2.32\\(0.49)} & \specialcell{2.31\\(0.45)} & \specialcell{2.28\\(0.45)} & \specialcell{2.31\\(0.42)} & \specialcell{-0.02***\\(0)} & \specialcell{-0.03**\\(0.02)}\\
 \hspace{1em}Men’s Log Annual Earnings & \specialcell{10.29\\(1.01)} & \specialcell{10.09\\(1.05)} & \specialcell{10.06\\(0.97)} & \specialcell{10.01\\(1.04)} & \specialcell{-0.28***\\(0.01)} & \specialcell{-0.03**\\(0.03)}\\
-\hspace{1em}Women’s Log Annual Earnings & \specialcell{9.43\\(1.77)} & \specialcell{9.52\\(1.59)} & \specialcell{9.47\\(1.56)} & \specialcell{9.53\\(1.49)} & \specialcell{0.1**\\(0.02)} & \specialcell{-0.05*\\(0.05)}\\*
+\hspace{1em}Women’s Log Annual Earnings & \specialcell{10.13\\(0.78)} & \specialcell{10.04\\(0.79)} & \specialcell{10.02\\(0.72)} & \specialcell{10.01\\(0.73)} & \specialcell{-0.12***\\(0.01)} & \specialcell{-0.02**\\(0.03)}\\*
 \end{longtable}

### 2024-08-11 my_paper/tables/tab01-observations-by-parents.tex
new edits

@@
-\begin{table}[!h]
-
-\caption{Number of Children from the Different Types of Parents \label{tab:mat1}}
+\begin{table}[H]
+\centering\centering
+\caption{Number of Children by Parental Type \label{tab:mat1}}
 \centering
-\resizebox{\linewidth}{!}{
+\resizebox{\ifdim\width>\linewidth\linewidth\else\width\fi}{!}{
 \begin{threeparttable}
@@
 \toprule
-\multicolumn{1}{c}{ } & \multicolumn{4}{c}{Perent's Type} \\
+\multicolumn{1}{c}{ } & \multicolumn{4}{c}{Perental Type} \\
 \cmidrule(l{3pt}r{3pt}){2-5}
-Parents' Type & \specialcell{White Father \\ White Mother} & \specialcell{White Father \\ Hispanic Mother} & \specialcell{Hispanic Father \\ White Mother} & \specialcell{Hispanic Father \\ Hispanic Mother}\\
+  & \specialcell{White Father \\ White Mother} & \specialcell{White Father \\ Hispanic Mother} & \specialcell{Hispanic Father \\ White Mother} & \specialcell{Hispanic Father \\ Hispanic Mother}\\
 \midrule
-\textbf{Observations} & \specialcell{3,828,024\\(0.95)} & \specialcell{24,306\\(0.01)} & \specialcell{35,841\\(0.01)} & \specialcell{135,300\\(0.03)}\\
+\textbf{\specialcell{Observations\\Share}} & \specialcell{6,405,696\\0.96} & \specialcell{38,937\\0.01} & \specialcell{51,090\\0.01} & \specialcell{179,152\\0.03}\\
 \bottomrule
@@
 \begin{tablenotes}
-\item[1] The data is restricted to people interviewed between 1994 and 2019, also White, married, and between the ages of 18 and 65.
-I identify the ethnicity of a person's parents through the parent's place of birth.
-A parent is Hispanic if both her parents were born in a Spanish-speaking country.
-A parent is White if born parents were born in the United States.
+\item[1] Source: Current Population Surveys (CPS) 1994-2019
+\item[2] The sample includes Whites, who are married, and are between the ages 25 and 40.
+Ethnicity of a person's parents are identified by the parent's place of birth.
+A parent is Hispanic if she/he was born in a Spanish-speaking country.
+A parent is White if she/he was born in the United States.
 \end{tablenotes}

### 2024-08-11 my_paper/tables/tab04-summary-stats.tex
change latex doc and pdf to incorporate new changes
new edits

@@
 \textbf{Panel A: Children's Education} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
-\hspace{1em}Men’s education (Total Years) & \specialcell{13.82\\(2.42)} & \specialcell{13.4\\(2.38)} & \specialcell{13.08\\(2.27)} & \specialcell{12.87\\(2.3)} & \specialcell{-0.94***\\(0.01)} & \specialcell{-0.32**\\(0.03)}\\
+\hspace{1em}Men’s education (Total Years) & \specialcell{13.82\\(2.42)} & \specialcell{13.4\\(2.38)} & \specialcell{13.07\\(2.27)} & \specialcell{12.87\\(2.3)} & \specialcell{-0.94***\\(0.01)} & \specialcell{-0.32**\\(0.03)}\\
 \hspace{1em}Women’s education (Total Years) & \specialcell{14.06\\(2.37)} & \specialcell{13.62\\(2.39)} & \specialcell{13.26\\(2.35)} & \specialcell{13.22\\(2.39)} & \specialcell{-0.84***\\(0.01)} & \specialcell{-0.36**\\(0.02)}\\
+\hspace{1em}Men’s Highschool Dropout Rate & \specialcell{0.35\\(0.48)} & \specialcell{0.45\\(0.5)} & \specialcell{0.44\\(0.5)} & \specialcell{0.44\\(0.5)} & \specialcell{0.09***\\(0)} & \specialcell{-0.01***\\(0.01)}\\
+\hspace{1em}Women’s Highschool Dropout Rate & \specialcell{0.36\\(0.48)} & \specialcell{0.45\\(0.5)} & \specialcell{0.46\\(0.5)} & \specialcell{0.46\\(0.5)} & \specialcell{0.1***\\(0)} & \specialcell{0.01***\\(0.01)}\\
+\addlinespace
+\hspace{1em}Men’s Associate Degree Rate & \specialcell{0.15\\(0.36)} & \specialcell{0.15\\(0.36)} & \specialcell{0.12\\(0.32)} & \specialcell{0.11\\(0.32)} & \specialcell{-0.03***\\(0)} & \specialcell{-0.03***\\(0)}\\
+\hspace{1em}Women’s Associate Degree Rate & \specialcell{0.19\\(0.39)} & \specialcell{0.17\\(0.37)} & \specialcell{0.14\\(0.35)} & \specialcell{0.15\\(0.36)} & \specialcell{-0.04***\\(0)} & \specialcell{-0.03***\\(0)}\\
+\hspace{1em}Men’s BA Degree Rate & \specialcell{0.32\\(0.47)} & \specialcell{0.23\\(0.42)} & \specialcell{0.19\\(0.39)} & \specialcell{0.17\\(0.38)} & \specialcell{-0.15***\\(0)} & \specialcell{-0.04***\\(0.01)}\\
+\hspace{1em}Women’s BA Degree Rate & \specialcell{0.36\\(0.48)} & \specialcell{0.28\\(0.45)} & \specialcell{0.22\\(0.41)} & \specialcell{0.22\\(0.41)} & \specialcell{-0.14***\\(0)} & \specialcell{-0.06***\\(0)}\\
+\hspace{1em}Men’s Professional Degree Rate & \specialcell{0.09\\(0.28)} & \specialcell{0.07\\(0.25)} & \specialcell{0.05\\(0.21)} & \specialcell{0.04\\(0.19)} & \specialcell{-0.05***\\(0)} & \specialcell{-0.02***\\(0)}\\
+\addlinespace
+\hspace{1em}Women’s Professional Degree Rate & \specialcell{0.1\\(0.3)} & \specialcell{0.08\\(0.26)} & \specialcell{0.07\\(0.25)} & \specialcell{0.06\\(0.23)} & \specialcell{-0.04***\\(0)} & \specialcell{-0.01***\\(0)}\\
 \textbf{Panel B: Children's Employment and Earnings} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
 \hspace{1em}Men’s Unemployment Rate & \specialcell{0.04\\(0.8)} & \specialcell{0.06\\(0.76)} & \specialcell{0.07\\(0.74)} & \specialcell{0.07\\(0.75)} & \specialcell{0.02***\\(0.00)} & \specialcell{0.01***\\(0.00)}\\
-\addlinespace
 \hspace{1em}Women’s Unemployment Rate & \specialcell{0.04\\(0.81)} & \specialcell{0.06\\(0.24)} & \specialcell{0.07\\(0.75)} & \specialcell{0.06\\(0.76)} & \specialcell{0.02***\\(0.00)} & \specialcell{0.01***\\(0.00)}\\
 \hspace{1em}Men’s Log Hourly Earnings & \specialcell{2.51\\(0.45)} & \specialcell{2.43\\(0.46)} & \specialcell{2.42\\(0.44)} & \specialcell{2.42\\(0.43)} & \specialcell{-0.09***\\(0)} & \specialcell{-0.00**\\(0.02)}\\
+\addlinespace
 \hspace{1em}Women’s Log Hourly Earnings & \specialcell{2.32\\(0.49)} & \specialcell{2.31\\(0.45)} & \specialcell{2.28\\(0.45)} & \specialcell{2.31\\(0.42)} & \specialcell{-0.02***\\(0)} & \specialcell{-0.03**\\(0.02)}\\
 \hspace{1em}Men’s Log Annual Earnings & \specialcell{10.29\\(1.01)} & \specialcell{10.09\\(1.05)} & \specialcell{10.06\\(0.97)} & \specialcell{10.01\\(1.04)} & \specialcell{-0.28***\\(0.01)} & \specialcell{-0.03**\\(0.03)}\\
-\hspace{1em}Women’s Log Annual Earnings & \specialcell{9.43\\(1.77)} & \specialcell{9.52\\(1.59)} & \specialcell{9.47\\(1.56)} & \specialcell{9.53\\(1.49)} & \specialcell{0.1**\\(0.02)} & \specialcell{-0.05*\\(0.05)}\\*
+\hspace{1em}Women’s Log Annual Earnings & \specialcell{10.13\\(0.78)} & \specialcell{10.04\\(0.79)} & \specialcell{10.02\\(0.72)} & \specialcell{10.01\\(0.73)} & \specialcell{-0.12***\\(0.01)} & \specialcell{-0.02**\\(0.03)}\\*
 \end{longtable}

### 2024-08-11 output/tables/tab02-observations-by-couple.tex
new edits

@@
 \begin{table}[H]
-
+\centering\centering
 \caption{Couples' Type \label{tab:mat2}}
 \centering
-\resizebox{\linewidth}{!}{
+\resizebox{\ifdim\width>\linewidth\linewidth\else\width\fi}{!}{
 \begin{threeparttable}
@@
 \begin{tablenotes}
-\item[1] Source: 1970-1990 Census
-\item[2] The sample includes Whites, who are married, and are between the ages 25 and 40.
+\item[1] Source: 1960-2000 Census
+\item[2] The sample includes Whites, who are married, with kids, and are between the ages 25 and 40.
 Ethnicity of a person's parents are identified by the parent's place of birth.
 A parent is Hispanic if she/he was born in a Spanish-speaking country.
-A parent is White she/he was born in the United States.
+A parent is White if she/he was born in the United States.
 \end{tablenotes}

### 2024-08-11 output/tables/tab05-regression.tex
new edits

@@
 \begin{table}[H]
-
-\caption{Effect of Having Hispanic Last Name \label{tab:lastnamereg}}
+\centering\centering
+\caption{Effect of Having Hispanic Last Name (Log Annual Earnings) \label{tab:lastnamereg}}
 \centering
 \begin{threeparttable}
-\begin{tabular}[t]{lcccc}
+\begin{tabular}[t]{lccccc}
 \toprule
-  & \specialcell{(1) \\ Log annual \\ earnings} & \specialcell{(2) \\ Log annual \\ earnings} & \specialcell{(3) \\ Log annual \\ earnings} & \specialcell{(4) \\  Log annual \\ earnings}\\
+  & \specialcell{(1) \\ Log annual \\ earnings} & \specialcell{(2) \\ Log annual \\ earnings} & \specialcell{(3) \\ Log annual \\ earnings} & \specialcell{(4) \\  Log annual \\ earnings} & \specialcell{(5) \\  Log annual \\ earnings}\\
 \midrule
-$HW_{ist}$ & \num{-0.06}** & \num{-0.07}** & \num{-0.05}** & \num{-0.02}\\
- & (\num{0.03}) & (\num{0.03}) & (\num{0.02}) & (\num{0.02})\\
-Constant & \num{10.40}*** & \num{9.49}*** &  & \\
- & (\num{0.04}) & (\num{0.07}) &  & \\
+$HW_{ist}$ & \num{-0.05}*** & \num{-0.05}** & \num{-0.05}* & \num{-0.02} & \num{-0.01}\\
+ & (\num{0.02}) & (\num{0.02}) & (\num{0.02}) & (\num{0.02}) & (\num{0.02})\\
+Constant & \num{10.42}*** & \num{9.46}*** &  &  & \\
+ & (\num{0.04}) & (\num{0.07}) &  &  & \\
 \midrule
-\textit{Controlling for:} &  &  &  & \\
-Hours Worked &  & X & X & X\\
-State FE &  &  & X & X\\
-Year FE &  &  & X & X\\
-Age FE &  &  & X & X\\
-Education FE &  &  &  & X\\
-Observations & \num{2423} & \num{2423} & \num{2423} & \num{2423}\\
+\textit{Controlling for:} &  &  &  &  & \\
+Hours Worked &  & X & X & X & X\\
+Age &  &  & X & X & X\\
+State FE &  &  & X & X & X\\
+Year FE &  &  & X & X & X\\
+State-Year FE &  &  & X & X & X\\
+Education &  &  &  & X & X\\
+Parental Background &  &  &  &  & X\\
+Observations & \num{3621} & \num{3621} & \num{3621} & \num{3621} & \num{3621}\\
 \bottomrule
-\multicolumn{5}{l}{\rule{0pt}{1em}* p $<$ 0.1, ** p $<$ 0.05, *** p $<$ 0.01}\\
+\multicolumn{6}{l}{\rule{0pt}{1em}* p $<$ 0.1, ** p $<$ 0.05, *** p $<$ 0.01}\\
 \end{tabular}

### 2024-08-11 output/tables/tab09-regression-weekearn-nonmex.tex
new edits

@@
 \begin{table}[H]
-
+\centering\centering
 \caption{Effect of Having Hispanic Last Name: Hispanics with Non-Mexican Ancestry  \label{tab:lastnamereg-weekearm-nonmex}}
@@
 \midrule
-$HW_{ist}$ & \num{-0.04} & \num{-0.04} & \num{-0.01} & \num{-0.01}\\
- & (\num{0.03}) & (\num{0.02}) & (\num{0.02}) & (\num{0.03})\\
-Constant & \num{6.71}*** &  &  & \\
+$HW_{ist}$ & \num{-0.04} & \num{-0.07}** & \num{-0.03} & \num{-0.03}\\
+ & (\num{0.03}) & (\num{0.03}) & (\num{0.03}) & (\num{0.03})\\
+Constant & \num{6.70}*** &  &  & \\
  & (\num{0.03}) &  &  & \\
@@
 Year FE &  & X & X & X\\
-Age FE &  & X & X & X\\
-Education FE &  &  & X & X\\
+State-Year FE &  & X & X & X\\
+Age &  & X & X & X\\
+Education &  &  & X & X\\
 Parental Background &  &  &  & X\\
-Observations & \num{2859} & \num{2859} & \num{2859} & \num{2859}\\
+Observations & \num{2853} & \num{2853} & \num{2853} & \num{2853}\\
 \bottomrule

### 2024-08-11 output/tables/tab10-regression-weekearn-cuba.tex
new edits

--- 2023-12-06 my_paper/tables/tab10-regression-weekearn-cuba.tex
@@
 \begin{table}[H]
-
+\centering\centering
 \caption{Effect of Having Hispanic Last Name: Hispanics with Cuban Ancestry  \label{tab:lastnamereg-weekearm-cub}}
@@
 \midrule
-$HW_{ist}$ & \num{-0.08}** & \num{-0.03} & \num{-0.01} & \num{-0.02}\\
- & (\num{0.04}) & (\num{0.04}) & (\num{0.04}) & (\num{0.05})\\
-Constant & \num{6.69}*** &  &  & \\
+$HW_{ist}$ & \num{-0.08}** & \num{-0.06} & \num{-0.02} & \num{-0.03}\\
+ & (\num{0.04}) & (\num{0.06}) & (\num{0.05}) & (\num{0.05})\\
+Constant & \num{6.68}*** &  &  & \\
  & (\num{0.05}) &  &  & \\
@@
 Year FE &  & X & X & X\\
-Age FE &  & X & X & X\\
-Education FE &  &  & X & X\\
+State-Year FE &  & X & X & X\\
+Age &  & X & X & X\\
+Education &  &  & X & X\\
 Parental Background &  &  &  & X\\
-Observations & \num{1182} & \num{1182} & \num{1182} & \num{1182}\\
+Observations & \num{1179} & \num{1179} & \num{1179} & \num{1179}\\
 \bottomrule

### 2024-08-11 output/tables/tab12-regression.tex
new edits

--- 2024-08-11 output/tables/tab05-regression.tex
@@
 \centering\centering
-\caption{Effect of Having Hispanic Last Name (Log Annual Earnings) \label{tab:lastnamereg}}
+\caption{Effect of Having Hispanic Last Name \label{tab:lastnamereg}}
 \centering
@@
 Age &  &  & X & X & X\\
-State FE &  &  & X & X & X\\
-Year FE &  &  & X & X & X\\
-State-Year FE &  &  & X & X & X\\
 Education &  &  &  & X & X\\
+NA & NA & NA & NA & NA & \vphantom{2} NA\\
+NA & NA & NA & NA & NA & \vphantom{1} NA\\
+NA & NA & NA & NA & NA & NA\\
 Parental Background &  &  &  &  & X\\
@@
 \item[2] {\setstretch{1.0}\footnotesize{HW is an indicator variable that is equal to 1 if a person is the child of a Hispanic-father and White-mother.}}
-\item[3] {\setstretch{1.0}\footnotesize{The sample is restricted to men working full-time full-year and are wage and salary workers.}}
-\item[4] {\setstretch{1.0}\footnotesize{Column one has the regression results when controlling for hours worked, age, education, year and state fixed effects.
-Column two has the results after controlling for education.}}
-\item[5] {\setstretch{1.0}\footnotesize{Standard errors are clustered on the state level.}}
+\item[3] {\setstretch{1.0}\footnotesize{Standard errors are clustered on the state level.}}
 \end{tablenotes}

### 2024-08-11 output/tables/tab13-regression.tex
new edits

\begin{table}[H]
\centering\centering
\caption{Effect of Having Hispanic Last Name \label{tab:lastname-ed-reg}}
\centering
\begin{threeparttable}
\begin{tabular}[t]{lcccc}
\toprule
  & \specialcell{(1) \\ Years of \\ Education} & \specialcell{(2) \\ Highschool \\ Dropout} & \specialcell{(3) \\ Associate Degree} & \specialcell{(4) \\ Bachelor Degree}\\
\midrule
\addlinespace[0.5em]
\multicolumn{5}{l}{\textit{Panel A: Full Sample}}\\
\midrule \hspace{1em}$HW_{ist}$ & -0.20*** & 0.01 & -0.02*** & -0.03***\\
\hspace{1em} & (0.05) & (0.01) & (0.01) & (0.01)\\
\hspace{1em}Observations & 88377 & 90027 & 66927 & 90027\\
\addlinespace[0.5em]
\multicolumn{5}{l}{\textit{Panel B: Women}}\\
\midrule \hspace{1em}$HW_{ist}$ & -0.25*** & 0.02 & -0.03*** & -0.04***\\
\hspace{1em} & (0.06) & (0.01) & (0.01) & (0.01)\\
\hspace{1em}Observations & 46516 & 47302 & 34334 & 47302\\
\addlinespace[0.5em]
\multicolumn{5}{l}{\textit{Panel C: Men}}\\
\midrule \hspace{1em}$HW_{ist}$ & -0.16** & 0.00 & -0.02** & -0.02\\
\hspace{1em} & (0.07) & (0.01) & (0.01) & (0.01)\\
\hspace{1em}Full Sample's Mean & 13.48 & 0.43 & 0.15 & 0.26\\
\hspace{1em}Observations & 41861 & 42725 & 32593 & 42725\\
Women's Mean & 13.58 & 0.44 & 0.16 & 0.27\\
Men's Mean & 13.38 & 0.43 & 0.14 & 0.24\\
\bottomrule
\multicolumn{5}{l}{\rule{0pt}{1em}* p $<$ 0.1, ** p $<$ 0.05, *** p $<$ 0.01}\\
\end{tabular}
\begin{tablenotes}
\item[1] {\setstretch{1.0}\footnotesize{This table includes the estimation results of equation (\ref{eq:1a}). All regressions include state-year fixed effects.}}
\item[2] {\setstretch{1.0}\footnotesize{HW is an indicator variable that is equal to 1 if a person is the child of a Hispanic-father and White-mother.}}
\item[3] {\setstretch{1.0}\footnotesize{Standard errors are clustered on the state level.}}
\end{tablenotes}
\end{threeparttable}
\end{table}

### 2024-08-11 programs/11-table-one-parent-types.R

@@ -56,7 +56,8 @@ first_col <- c(" ",
                       "\\\\",
                 "Share",
                       "}"
-))
+                      )
+                      )
 
 second_col <- c("\\specialcell{White Father \\\\ White Mother}",
                paste0("\\specialcell{", 

### 2024-08-11 programs/15-table-four-summary-stats.R

@@ -157,9 +157,558 @@ row8 <- c(
 )
 dim(row8) <- c(1,7)
 
+## Man's High School dropout
+ManEducation <- IndividualData |> 
+  filter(sex == 1) |> 
+  group_by(Type) |> 
+  summarise(
+    hs_dropout_mean = mean(hs_dropout, na.rm = TRUE),
+    hs_dropout_sd   = sd(hs_dropout, na.rm = TRUE)
+  )
+
+model <- lm(hs_dropout ~ 0 + WW + WH + HW + HH, 
+            data = IndividualData |> filter(sex == 1))
+
+test1 <- tidy(glht(model, linfct = c("HH - WW = 0")))
+pvalues1 <- test1$std.error
+differences1 <- test1$estimate
+
+test2 <- tidy(glht(model, linfct = c("HW - WH = 0")))
+pvalues2 <- test2$std.error
+differences2 <- test2$estimate
+
+row16 <- c(
+  "Men’s High School Dropout Rate",
+  paste0("\\specialcell{", 
+         round(ManEducation[4,2], digits = 2), 
+         "\\\\",
+         "(",
+         round(ManEducation[4,3], digits = 2),
+         ")}"
+  ),
+  paste0("\\specialcell{", 
+         round(ManEducation[3,2], digits = 2), 
+         "\\\\",
+         "(",
+         round(ManEducation[3,3], digits = 2),
+         ")}"
+  ),
+  paste0("\\specialcell{", 
+         round(ManEducation[2,2], digits = 2), 
+         "\\\\",
+         "(",
+         round(ManEducation[2,3], digits = 2),
+         ")}"
+  ),
+  paste0("\\specialcell{", 
+         round(ManEducation[1,2], digits = 2), 
+         "\\\\",
+         "(",
+         round(ManEducation[1,3], digits = 2),
+         ")}"
+  ),
+  paste0("\\specialcell{", 
+         sprintf("%.2f", round(differences1, 2)), 
+         symnum(pvalues1[[1]], corr = FALSE, na = FALSE, cutpoints = c(0, 0.01, 0.05, 0.1), symbols = c("***", "**", "*")),
+         "\\\\",
+         "(",
+         round(pvalues1[[1]], digits = 2),
+         ")}"
+  ),
+  paste0("\\specialcell{", 
+         round(differences2[[1]], digits = 2), 
[…]

### 2024-08-11 programs/16-table-five-regression.R

@@ -36,13 +36,13 @@ reg1 <- list(
   "\\specialcell{(2) \\\\ Log annual \\\\ earnings}" = feols(lninctot_1999 ~ .[ParentDummies], vcov = ~statefip,
                                                             data = IndividualData |> 
                                                           filter(sex == 1 & FTFY == 1 & Self_employed_ASEC == 0 & (HW == 1 | WH == 1))),
-  "\\specialcell{(3) \\\\ Log annual \\\\ earnings}" = feols(lninctot_1999 ~ .[ParentDummies] | year + age + statefip, vcov = ~statefip,
+  "\\specialcell{(3) \\\\ Log annual \\\\ earnings}" = feols(lninctot_1999 ~ .[ParentDummies] + age | year*statefip, vcov = ~statefip,
                                                             data = IndividualData |> 
                                                           filter(sex == 1 & FTFY == 1 & Self_employed_ASEC == 0 & (HW == 1 | WH == 1))),
-  "\\specialcell{(4) \\\\  Log annual \\\\ earnings}" = feols(lninctot_1999 ~ .[ParentDummies] | year + age + educ + statefip, vcov = ~statefip,
+  "\\specialcell{(4) \\\\  Log annual \\\\ earnings}" = feols(lninctot_1999 ~ .[ParentDummies] + age + educ | year*statefip, vcov = ~statefip,
                                                          data = IndividualData |> 
                                                            filter(sex == 1 & FTFY == 1 & Self_employed_ASEC == 0 & (HW == 1 | WH == 1))),
-  "\\specialcell{(5) \\\\  Log annual \\\\ earnings}" = feols(lninctot_1999 ~ .[ParentDummies] + .[ParentControls] | year + age + educ + statefip, vcov = ~statefip,
+  "\\specialcell{(5) \\\\  Log annual \\\\ earnings}" = feols(lninctot_1999 ~ .[ParentDummies] + .[ParentControls] + age + educ | year*statefip, vcov = ~statefip,
                                                          data = IndividualData |> 
                                                            filter(sex == 1 & FTFY == 1 & Self_employed_ASEC == 0 & (HW == 1 | WH == 1)))
   
@@ -93,25 +93,33 @@ reg1 <- list(
[…]

### 2024-08-11 programs/17-table-six-regression.R

@@ -51,13 +51,13 @@ reg1 <- list(
   "\\specialcell{(2) \\\\ Log annual \\\\ earnings}" = feols(lninctot_1999 ~ .[ParentDummies] , vcov = ~statefip,
                                                         data = IndividualData |> 
                                                           filter(sex == 1 & FTFY == 1 & Self_employed == 0 & (HW == 1 | WH == 1))),
-  "\\specialcell{(3) \\\\ Log annual \\\\ earnings}" = feols(lninctot_1999 ~ .[ParentDummies] | year + age + statefip, vcov = ~statefip,
+  "\\specialcell{(3) \\\\ Log annual \\\\ earnings}" = feols(lninctot_1999 ~ .[ParentDummies] + age | year*statefip, vcov = ~statefip,
                                                         data = IndividualData |> 
                                                           filter(sex == 1 & FTFY == 1 & Self_employed == 0 & (HW == 1 | WH == 1))),
-  "\\specialcell{(4) \\\\  Log annual \\\\ earnings}" = feols(lninctot_1999 ~ .[ParentDummies] | year + age + educ + statefip, vcov = ~statefip,
+  "\\specialcell{(4) \\\\  Log annual \\\\ earnings}" = feols(lninctot_1999 ~ .[ParentDummies] + age + educ | year*statefip, vcov = ~statefip,
                                                          data = IndividualData |> 
                                                            filter(sex == 1 & FTFY == 1 & Self_employed == 0 & (HW == 1 | WH == 1) )),
-  "\\specialcell{(5) \\\\  Log annual \\\\ earnings}" = feols(lninctot_1999 ~ .[ParentDummies] + .[ParentControls]| year + age + educ + statefip, vcov = ~statefip,
+  "\\specialcell{(5) \\\\  Log annual \\\\ earnings}" = feols(lninctot_1999 ~ .[ParentDummies] + .[ParentControls] + age + educ | year*statefip, vcov = ~statefip,
                                                          data = IndividualData |> 
                                                            filter(sex == 1 & FTFY == 1 & Self_employed == 0 & (HW == 1 | WH == 1) ))
   
@@ -130,20 +130,30 @@ dim(controling_for) <- c(1,6)
[…]

### 2024-08-11 programs/18-table-seven-regression.R

@@ -10,10 +10,10 @@ IndividualData <- read_dta(file.path(datasets,"BySexAnalysisData.dta")) |>
 
 # By generation
 reg1 <- list(
-  "\\specialcell{(1) \\\\ Log annual earnings}" = feols(lninctot_1999 ~ Hispanic_ID+  uhrsworkly | year + age+ statefip, vcov = ~statefip,
+  "\\specialcell{(1) \\\\ Log annual earnings}" = feols(lninctot_1999 ~ Hispanic_ID+  uhrsworkly + age | year*statefip, vcov = ~statefip,
                                                         data = IndividualData |> 
                                                           filter(sex == 1 & FTFY == 1 & Self_employed == 0)),
-  "\\specialcell{(2) \\\\  Log annual earnings}" = feols(lninctot_1999 ~ Hispanic_ID+  uhrsworkly | year + age + educ+ statefip, vcov = ~statefip,
+  "\\specialcell{(2) \\\\  Log annual earnings}" = feols(lninctot_1999 ~ Hispanic_ID+  uhrsworkly + age + educ | year*statefip, vcov = ~statefip,
                                                          data = IndividualData |> 
                                                            filter(sex == 1 & FTFY == 1 & Self_employed == 0))
   
@@ -24,12 +24,16 @@ controling_for <-  c("\\textit{Controlling for:}", " ", "")
 dim(controling_for) <- c(1,3)
 
 hoursworked <-  c("Hours Worked", "X", "X")
+educ_cont   <-  c("Education", "", "X")
+
 dim(hoursworked) <- c(1,3)
+dim(educ_cont) <- c(1,3)
 
 controling_for <- as.data.frame(controling_for)
 hoursworked <- as.data.frame(hoursworked)
+educ_cont <- as.data.frame(educ_cont)
 
-all_row <- rbind(controling_for, hoursworked)
+all_row <- rbind(controling_for, hoursworked, educ_cont)
 attr(all_row, 'position') <- c(4:5)
 
 cm <- c("Hispanic_ID" = "$Hispanic_{i}$"
@@ -39,8 +43,9 @@ gm <- tibble::tribble(
   "nobs",      "Observations",             0,
   "FE: year", "Year FE", 0,
   "FE: statefip", "State FE", 0,
-  "FE: age", "Age", 0,
-  "FE: educ", "Education", 0,
+  "FE: year:statefip", "State-Year FE", 0,
[…]

### 2024-08-11 programs/23-outgoing-rotation-regression.R

@@ -35,13 +35,13 @@ reg1 <- list(
   "\\specialcell{(1) \\\\ Log weekly \\\\ earnings}" = feols(logearnweek ~ HW, vcov = ~statefip,
                                                             data = IndividualData |> 
                                                           filter(sex == 1 & FullTime == 1 & Self_employed == 0 & (HW == 1 | WH == 1))),
-  "\\specialcell{(2) \\\\ Log weekly \\\\ earnings}" = feols(logearnweek ~ .[ParentDummies] | year + age + statefip, vcov = ~statefip,
+  "\\specialcell{(2) \\\\ Log weekly \\\\ earnings}" = feols(logearnweek ~ .[ParentDummies] + age | year*statefip, vcov = ~statefip,
                                                             data = IndividualData |> 
                                                           filter(sex == 1 & FullTime == 1 & Self_employed == 0 & (HW == 1 | WH == 1))),
-  "\\specialcell{(3) \\\\  Log weekly \\\\ earnings}" = feols(logearnweek ~ .[ParentDummies] | year + age + educ + statefip, vcov = ~statefip,
+  "\\specialcell{(3) \\\\  Log weekly \\\\ earnings}" = feols(logearnweek ~ .[ParentDummies] + age + educ | year*statefip, vcov = ~statefip,
                                                          data = IndividualData |> 
                                                            filter(sex == 1 & FullTime == 1 & Self_employed == 0 & (HW == 1 | WH == 1))),
-  "\\specialcell{(4) \\\\  Log weekly \\\\ earnings}" = feols(logearnweek ~ .[ParentDummies] + .[ParentControls] | year + age + educ + statefip, vcov = ~statefip,
+  "\\specialcell{(4) \\\\  Log weekly \\\\ earnings}" = feols(logearnweek ~ .[ParentDummies] + .[ParentControls]  + age + educ | year*statefip, vcov = ~statefip,
                                                          data = IndividualData |> 
                                                            filter(sex == 1 & FullTime == 1 & Self_employed == 0 & (HW == 1 | WH == 1)))
   
@@ -51,21 +51,29 @@ reg1 <- list(
 controling_for <-  c("\\textit{Controlling for:}", " ", "", " ", "")
 dim(controling_for) <- c(1,5)
 
[…]

### 2024-08-11 programs/26-educ-regressions.R

@@ -1,0 +1,192 @@
+# table Five
+# Regression
+
+# date: January 19th, 2023
+
+# open data, filter to only have the relevant data
+# and impute empty parental charachtersitcs with mean
+IndividualData <- read_csv(file.path(datasets,"CPS_synth.csv")) |> 
+  mutate(Education = as.numeric(Education)) |> 
+  filter(HW == 1 | WH == 1) |> 
+  mutate(HusbandEducation = case_when(!is.na(HusbandEducation) ~ HusbandEducation,
+                                       is.na(HusbandEducation) ~ mean(HusbandEducation, na.rm = TRUE)),
+         WifeEducation = case_when(!is.na(WifeEducation) ~ WifeEducation,
+                                    is.na(WifeEducation) ~ mean(WifeEducation, na.rm = TRUE)),
+         Husband_ftotval = case_when(!is.na(Husband_ftotval) ~ Husband_ftotval,
+                                       is.na(Husband_ftotval) ~ mean(Husband_ftotval, na.rm = TRUE)),
+                                )
+# create a total education variable
+# and a income quintile variable
+# and education quintile variable
+IndividualData  <- IndividualData  |> 
+  mutate(TotalHHEducation = HusbandEducation + WifeEducation,
+         IncomeQuint      = cut(Husband_ftotval, breaks = quantile(Husband_ftotval, probs = c(0, 0.2, 0.4, 0.6, 0.8, 1), na.rm = TRUE), include.lowest = TRUE)
+          )
+IndividualData  <- IndividualData  |> 
+  mutate(EducationQuint      = cut(TotalHHEducation, breaks = quantile(TotalHHEducation, probs = c(0, 0.2, 0.4, 0.6, 0.8, 1), na.rm = TRUE), include.lowest = TRUE)
+          )
+ParentDummies <- c("HW", "uhrsworkly")
+
+ParentControls  <- c("IncomeQuint", "EducationQuint")
+# By generation
+feols(Education ~ HW  , vcov = ~statefip, data = IndividualData)
+feols(Education ~ HW  | year + statefip, vcov = ~statefip, data = IndividualData)
+feols(Education ~ HW  +  age +.[ParentControls]  | year + statefip, vcov = ~statefip, data = IndividualData)
+feols(Education ~ HW  +  age +.[ParentControls]  | year*statefip, vcov = ~statefip, data = IndividualData)
[…]

### 2024-08-28 manuscripts/02-ilr-review-v2/manuscript_title_etc.tex
....

--- 2024-05-20 manuscripts/latest-revisions/manuscript_title_etc.tex
@@
 I thank Professors Willa Friedman, Chinhui Juhn, Vikram Maheshri, and Yona Rubinstein for their support and advice.
-I also thank Aimee Chin, Steven Craig, German Cubas, Elaine Liu, Fan Wang, and the participants of the Applied Microeconomics Workshop at the University of Houston, and European Society for Population Economics (ESPE) for helpful feedback.}
+I also thank Aimee Chin, Steven Craig, German Cubas, Elaine Liu, Fan HADAH, and the participants of the Applied Microeconomics Workshop at the University of Houston, and European Society for Population Economics (ESPE) for helpful feedback.}
 

### 2024-08-28 manuscripts/05-Hussain-Hadah/Hadah-Hispanic-Last-Names.tex
....

--- 2023-08-14 manuscripts/Hussain-Hadah/Hadah-Hispanic-Last-Names.tex
@@
             % citysep={}, % Uncomment if no comma needed between city and postcode
-            postcode={70118},
-            state={LA},
+            % postcode={70118},
+            state={LA 70118},
             country={United States}}
@@
 \singlespacing % Set single spacing for the abstract
-In this paper, I study discrimination against Hispanics in the labor market.
-I compare the children of inter-ethnic marriages to study the impact of having a Hispanic last name.
-While males born to Hispanic father-White mothers earn 5 percentage points less than those born to White father-Hispanic mothers, the gap could be completely explained by educational differences.
-I also study the effect of identifying as Hispanic on earnings.
-I find that men with a Spanish-sounding last name who identify as Hispanic earn significantly less than those who do not identify as Hispanic, but the gap could also be explained by educational differences.
+Do individuals with Hispanic names face labor market discrimination?
+This study analyzes the impact of Hispanic-sounding surnames on wages among inter-ethnic children with one White and one Hispanic parent.
+I find that individuals with Hispanic surnames often earn less, with a notable wage gap favoring those with White surnames.
+People born to Hispanic fathers and White mothers receive 0.3 years of education less than those born to White fathers and Hispanic mothers.
+Males born to Hispanic fathers and White mothers earn 5 percentage points less than those born to White fathers and Hispanic mothers, largely due to educational differences.
+Additionally, men with Spanish-sounding last names who identify as Hispanic earn significantly less, again due to educational disparities.
 \end{abstract}
@@
 \maketitle
-\section{Introduction}\label{sec:intro}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Main Text
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\section{Introduction}
 
 A large literature documents substantial earnings gaps across race and ethnicity \citep{bayer2018divergent, charles2008prejudice, card1992school, fryer2004causes, rubinstein2014pride, bertrand2004emily, juhn1991accounting}.
-Hispanics constitute a large and growing portion of the population in the United States, see figure \ref{fig:3}.
+Hispanics constitute a large and growing portion of the population in the United States.
 As the number of Hispanics increases, determining whether ethnic discrimination affects their labor market outcomes becomes increasingly crucial \citep{chettyUnitedStatesStill2014, chettyEffectsExposureBetter2016,chettyFadingAmericanDream2017,abramitzkyImmigrantsAssimilateMore2020a, abramitzkyNationImmigrantsAssimilation2014,abramitzkyCulturalAssimilationAge2016,chettyWhereLandOpportunity2014}.
@@
 
-In this paper, I answer the following questions.
-Does having a Hispanic last name affect labor market outcomes?
-What is the effect of identifying as Hispanic on earnings?
-Moreover, I aim to show that comparing Hispanic Whites to non-Hispanic Whites might create an artificially higher earnings gap since the two groups differ on many observable characteristics.\footnote{By identifying as Hispanic, I refer to individuals who self-report their ethnicity as Hispanic on surveys or other data collection instruments.} \footnote{Observable characteristics refer to factors that can be measured and quantified, such as education level, work experience, and immigration status.} Others have attempted to compare how native-born White Hispanics fare to non-Hispanic Whites and foreign-born Hispanics.
+In this paper, I answer the following question.
+Does having a Hispanic last name affect education and labor market outcomes?
+Moreover, I aim to show that comparing Hispanic Whites to non-Hispanic Whites might create an artificially higher earnings gap since the two groups differ on many observable characteristics.
+\footnote{Observable characteristics refer to factors that can be measured and quantified, such as education level, work experience, and immigration status.} Others have attempted to compare how native-born White Hispanics fare to non-Hispanic Whites and foreign-born Hispanics.
 In \citet{antman2020ethnic,antmanEthnicAttritionObserved2016,antmanEthnicAttritionObserved2016a,antmanEthnicAttritionAssimilation2020}, the authors compare the health and educational outcomes of Hispanic Whites to non-Hispanic Whites and native-born Hispanics to foreign-born Hispanics.
@@
 For example, discrimination against Hispanics can lead to reduced job opportunities, lower wages, and hinder assimilation.
-In this paper, I examine the role of having a Hispanic last name and identifying as Hispanic on labor market outcomes.
+In this paper, I examine the role of having a Hispanic last name on labor market and educational outcomes outcomes.
 
-Identifying discrimination is difficult because of factors that affect labor market outcomes that are unobservable to economists---such as unobserved skills.
+Identifying discrimination is difficult because of factors that affect labor market outcomes that are unobservable to economists---such as unobserved skills and separating it from prejudice and stereotypes.
 One strategy used by researchers is audit or resume studies.
@@
 
-\begin{figure}[t]
-\centering
-\includegraphics[scale=0.50]{Fig/Hispanics_all.png}
-\captionsetup{width=0.85\linewidth} % Adjust the width to your preference
-\caption{Hispanics as a percent of the population in the United States from 1994 to 2021.}
-\label{fig:3}
-
-\begin{minipage}{\linewidth}
-\footnotesize
-\textit{Source:} The data source is the Current Population Survey (CPS) from 1994 to 2021.\par
-\textit{Note:} The graph shows the percent of the population in the US that self-identifies as Hispanic.
-\end{minipage}
-\end{figure}
-
 This study utilizes a method developed by \citet{rubinstein2014pride}.
@@
 
-Using the Current Population Survey (CPS) from 1994 to 2019 and the 1960 to 2000 US censuses, I estimate the effect of having a Hispanic last name on labor market outcomes.
-In other words, I approximate the variation in labor market perception of ethnic signals, in this case having a Hispanic last name, on annual log earnings.
-I also study the effect of identifying as Hispanic on labor market outcomes.
-
-I find that while men born to Hispanic father-White mothers earn less than men born to White father-Hispanic mothers, this gap is entirely explained by educational differences.
-Thus, I do not find a significant effect of having a Hispanic last name.
-Finally, I find that men identifying as Hispanic earn significantly less than those who do not, even when I control for ancestry and education.
-
-Previous studies have also used names as a proxy for race and ethnicity \citep{fryer2004causes, rubinstein2014pride, bertrand2004emily}.
-\citet{fryer2004causes} point out that names can be a predictor of a person's race.
-Specifically, they provide a rising pattern among Blacks having different names than Whites.
-They did, however, find that having a Black name, after controlling for the home environment at birth, does not affect their labor market outcomes.
-\citet{rubinstein2014pride} compared the children of mixed marriages between Sephardic and Ashkenazi Jews in Israel.\footnote{Ashkenazi and Sephardic are two distinct Jewish ethnic groups.} They found that workers with Sephardic last names earn substantially less than those with Ashkenazi last names.
-\citet{bertrand2004emily} conducted an audit study by sending employers identical resumes that differ in the ethnic and racial signal of a name (Black sounding name versus a White sounding one).
-They found that resumes with Black-sounding names received substantially fewer callbacks than their White counterparts.
-
-The rest of the paper is organized as follows.
-Section \ref{sec:data} provides an overview of the data and summary statistics of the sample.
-Section \ref{sec:emp_model} introduces the empirical framework.
-I report the results in Section \ref{sec:results}.
-Finally, I offer a brief conclusion in Section \ref{sec:con1}.
+The main identifying assumption of my empirical strategy depends on the assumption that people born to HW parents are similar to their WH peer in all observable---and unobservable---aspects and characteristics that are important in the labor market.
+Consequently, the only difference between the two groups is the variation in which group is more likely to have a Hispanic sounding last name.
+Children from mixed ethnic backgrounds may appear physically similar to those from single ethnicity backgrounds, but the influence of family dynamics and upbringing, crucial in developing skills and personal characteristics, varies with the pattern of mixed ethnic marriages.
+Particularly in a society where Hispanic ancestry is perceived negatively, it raises the question of what kind of White women would choose Hispanic men.
+Furthermore, even if such unions were formed randomly, children from White-Hispanic homes might benefit from more favorable family conditions than those from Hispanic-White homes, considering Whites generally have stronger socio-economic backgrounds than Hispanics.
+These factors introduce doubts regarding whether children from Hispanic-White families receive comparable familial support and influences, either genetically or environmentally, as those from White-Hispanic families.
 
@@
 
-I use two datasets:  the Integrated Public Use Microdata Series (IPUMS) Current Population Survey (CPS) Annual Social and Economic (ASEC) \citep{cps2019} and the 1960 to 2000 US censuses \citep{acs2019}.
+I use two datasets: the IPUMS Current Population Survey (CPS) Annual Social and Economic (ASEC) \citep{cps2019} and the 1960 to 2000 US censuses \citep{acs2019}.
 
-I use the CPS data set to study the effect of having a Hispanic last name on a person's labor market outcomes.
-I take advantage of the fact that the CPS asks parents' place of birth, ethnicity, and race.
-The data spans the period between 1994--- the earliest sample to ask about a parent's place of birth--- to 2019.
-Moreover, since the CPS does not provide data on parents' characteristics, essential to determine the family background, I use the census to construct synthetic parents.
-The census offers a larger sample of potential parents.
-Similar to the CPS, the 1960 to 2000 censuses ask about the person's place of birth and the individual's race and ethnicity.
-I employ this information to construct "synthetic parents" using a method developed by \citet{rubinstein2014pride}.
-I construct the synthetic parents by linking husbands and wives in the census data to each other.
-I assume that parents have children between the ages of 25 and 40.
-I then link these synthetic parents using the birth year of child, and place of birth of parents to the year of birth of the "children" and the palce of birth of parents in the CPS sample.
+The CPS data (1994-2019) is used to study the effect of Hispanic last names on labor market outcomes.
+The census data is employed to construct synthetic parents using the method developed by \citet{rubinstein2014pride}.
 
@@
 
-I use the CPS for my primary analysis of the effect of having a Hispanic last name on earnings.
-I restrict my sample to Whites, United States-born citizens aged 25 to 40 and born between 1960 to 2000.
-Taking advantage of data on parents' place of birth, I divide the sample into four groups depending on their parent's ethnicity.
-Mothers or fathers are Hispanic if they were born in a Spanish-speaking country and Puerto Rico, and White if they were born in the United States.\footnote{The list of Spanish Speaking countries include: Argentina, Bolivia, Chile, Colombia, Costa Rica, Cuba, Dominican Republic, Ecuador, El Salvador, Guatemala, Equatorial Guinea, Honduras, Mexico, Nicaragua, Panama, Paraguay, Peru, Spain, Uruguay, Venezuela.} Therefore, an observation can be the product of four types of parents:
-\begin{enumerate}
-\item White father and White mother (hereafter WW)
-\item White father and Hispanic mother (hereafter WH)
-\item Hispanic father and White mother (hereafter HW)
-\item Hispanic father and Hispanic mother (hereafter HH).
-\end{enumerate}
+The CPS sample is restricted to White, US-born citizens aged 25-40, born between 1960-2000.
+Parents are classified as Hispanic if born in a Spanish-speaking country or Puerto Rico, and White if US-born.
+Four parental types are identified: (1) White father and White mother (WW), (2) White father and Hispanic mother (WH), (3) Hispanic father and White mother (HW), and (4) Hispanic father and Hispanic mother (HH).
 
-\begin{figure}[t]
-\centering
-\includegraphics[scale=0.5]{Fig/PirChart2.png}
-\caption{Distribution of the four types of children.}
-\label{fig:dist}
-\begin{minipage}{\linewidth}
-\footnotesize
-\textit{Source:} The data source is the Current Population Survey (CPS) from 1994 to 2021.\par
-\end{minipage}
-\end{figure}
-
-\begin{figure}[t]
-\centering
-\begin{tikzpicture}[node distance =2cm]
-\node (start) [startstop] {Synthetic Parents: Observed in 1960 to 2000. $25 \leq age \leq 40$, White, with kids};
-
-\node (pro1) [process, below of = start, yshift = -2cm] {Children: Observed between 1994 and 2019. Born between 1954 to 1999 and $25 \leq age \leq 40$};
-
-\draw [arrow] (start) -- (pro1);
-\end{tikzpicture}
-\caption{Chart explaining which synthetic parents and children and when we observe them.}
-\label{flowchart1}
-\end{figure}
-
-The distribution of the four types of children is presented in figure \ref{fig:dist} and table \ref{tab:mat1}.
-The majority of the sample (95.96\%) is WW children.
-The second biggest group is HH, which constitutes 2.69\% of the sample.
-Inter-ethnic children, WH and HW, make up 1.35\% of the sample with 90,325 observations.
-Even though WH and HW are only 1.35\% of the sample, I have plenty of observations to carry out an analysis.
-The summary statistics for the children of the four types of marriages are presented in table \ref{tab:c&p1}.
-Children of WW marriages (column 1) do better on every measure while children of HH parents (column 4) do worse than other children on every measure.
-
-%table1
-\input{tables/tab01-observations-by-parents}
-%table2
-\input{tables/tab02-observations-by-couple}
-
-On the one hand, WW children are more educated, with approximately 14 years of education for men and women.
-WW children have an unemployment rate of 4\% for men and women.
-They have log hourly earnings of 2.51 for men and 2.32 for women.
-Also, their annual log earnings are 10.29 for men and 9.43 for women.
-
-On the other hand, children of HH marriages have an average of 12.9 and 13.24 years of education for men and women, respectively.
-HH men and women have an unemployment rate of 7\% and 6\%.
-Their annual log earnings are 10.01 and 9.53 for men and women, respectively.
-
-Inter-ethnic children, WH and HW, have lower educational attainment and income than WW children, but they are better than HH children.
-Also, WH and HW are comparable, unlike WW and HH children.
-These two groups are important to the analysis.
-Children of HW are going to be the treatment group.
-HW children have a Hispanic father with a Hispanic last names.
-At the same time, the WH children have White fathers with a non-Hispanic, or White, last names.
-Summary statistics of these two groups, table \ref{tab:c&p1} columns (2) and (3) can show that they are different from the children of non-inter-ethnic marriages.
-
-%table3
-\input{tables/tab03b-summary-stats}
-
-WH children have an average of 13.58 and 13.79 years of education for men and women.
-The unemployment rate for men and women is 5\%.
-HW men and women obtained 13.22 and 13.42 years of schooling.
-Men's unemployment rate is 7\% and women's 6\%.
-
-In panel C of table \ref{tab:c&p1}, I present the rates at which members of each group self-identify as Hispanic.
-\footnote{Hispanic self-identification is observed when an individual notes that they are Hispanic on the CPS questionnaire} The vast majority of HH children, 96\% for men and 97\% for women, identify as Hispanic.
-There is considerable attrition in Hispanic identification of inter-ethnic children.
-Among WH children, 74\% of men and 78\% of women identify as Hispanic.
-Among HW children, 83\% of men and 82\% of women identify as Hispanic.
-These results align with the findings of a substantial body of literature that noted this attrition among children of immigrants, especially those coming from inter-ethnic marriages \citep{duncan2017complexity, duncan2018identifying, duncan2020new, antman2020ethnic}.
-Finally, a small portion of WW children, 4\% of men and 5\% of women, identified as Hispanic.
-These 5\% of WW children are most likely third generation--- or higher--- children of immigrants.
-For my analysis, I dropped the WW observations who identify as Hispanic when I compare with non-Hispanic Whites.
-
-In columns (5) and (6), I calculate the differences in means between HH-WW (column 5) and HW-WH (column 6).
-Overall, HH children do worse than WW children.
-HH has one year less of education compared to WW, and are 2 percentage points more likely to be unemployed.
-HH men earn 9\% less than WW men, and HH women make 2\% less than WW women.
-Therefore, these two groups, HH and WW, are different on many observable qualities.
-
-The differentials between WH and HW are not as severe as the ones between HH and WW.
-HW children are still worse off than WH but are much better than HH.
-HW men attain 0.36 fewer years of education than WH men.
-HW women achieve 0.37 fewer years of schooling than WH women.
-HW children are 1 percentage points more likely to be unemployed.
-HW men earn 1\% less annual earnings as WH men, and HW women earn 3\% less.
-These differences show that HW and WH children are different from children of endogamous/homogamous marriages and, thus, are more comparable.
-
-In table \ref{tab:c&p2}, I present the summary statistics for the four types and limit the sample to those that only identify as Hispanic.\footnote{Limiting the sample to those who identify as Hispanic only affects WH, HW, and HH.
-Members of the WW group are not affected by this restriction.} The same pattern discussed above between HH--WW and HW--WH persists.
-The one difference is that those who identify as Hispanic do slightly worse.
-
-%table4
-\input{tables/tab03a-summary-stats}
+Table \ref{tab:combined-mat} shows the distribution of these types.
+WW children comprise 96\% of the sample, HH 3\%, and inter-ethnic children (WH and HW) 1.35\%.
+Summary statistics for each group are presented in Table \ref{tab:combined-sum}.
 
@@
 
-Using the 1960 to 2000 censuses, I constructed a data set of synthetic parents.
-The sample includes married White men and women.
-Even though the census asks a person whether they are Hispanic or not, I took advantage of the questions on place of birth to create a proxy for ethnicity.
-I consider a Hispanic person as White and born in a Spanish-speaking country.
-Consequently, Whites in the sample are people who are White and native-born.
-Using the information provided in the census, I can link husbands and wives with each other.
-I assume that parents have children between the ages of 25 and 40.
-Therefore, my sample consists of married White men and women with children that were born in the 1920 to 1975 cohorts, see figure \ref{flowchart1} for a flowchart of the linking process.\footnote{The construction of "synthetic parents" follows the method used by \citet{rubinstein2014pride}.}
+Using the 1960-2000 censuses, I constructed a dataset of synthetic parents, including married White men and women born between 1920-1975.
+Hispanic individuals are defined as White and born in a Spanish-speaking country, while Whites are native-born.
 
-%table5
-\input{tables/tab04-summary-stats}
-
-I show the distribution of the four types of couples in table \ref{tab:mat2}.
-White husbands and White wives (WW) make up the majority of couples in the sample, 96\% (5,141,737 couples).
-Hispanic husbands and wives (HH) are the second largest group making up 2\% (119,749 couples) of all couples.
-White husbands and Hispanic wives (WH) couples are less 1\% (33,097 couples) of the sample and Hispanic husbands and White wives (HW) are also 1\% (37,847 couples).
-
-I present the summary statistics of the four types of synthetic parents in table \ref{tab:synth}.
-WW couples have higher education, 12.75 years for husbands and 12.47 for wives.
-As a household, WW couples have 25.22 years of schooling.
-Additionally, the log total family income for a WW household is 10.72 and have 3.77 children.
-Husbands in HH marriages have 8.64 years of education, while women have 8.49 years of schooling.
-As a household, HH couples have 17.13 years of education, log total family income of 10.27, and have 4.23 children.
-
-WH husbands have 11.77 years of education, while wives have an average of 10.40 years.
-WH household attained a total of 22.17 years of schooling in total, a log total family income of 10.55, and 3.98 children.
-HW husbands and wives obtained 10.25 and 11.11 years of education, respectively.
-HW household attained a total of 21.36 years of schooling in total, a log total family income of 10.46, and 4.15 children.
-
-In columns 5 and 6 of table \ref{tab:synth}, I report the differences in means between HH-WW households and HW-WH households.
-An HH Hispanic husband attained 4.11 fewer years of education than a WW husband.
-While an HH wife acquired 3.98 fewer years of schooling than a White WW wife.
-The total educational difference between HH and WW households equals 8.09 years of education.
-The educational differences between HW and WH are less severe than those between HH and WW.
-A Hispanic husband in intermarriage attained 1.52 fewer years of schooling than a White husband in intermarriage.
-A White wife in intermarriage attained 0.70 years of education, more than a Hispanic wife in intermarriage.
-The difference in total years of education between HW and WH households is 0.81, which is an 90\% decrease from the HH--WW gap.
-The total family income of an HH household is 45\% lower than a WW household, while the total family income of an HW household is 10\% lower than a WH household.
-These differences show that couples that intermarry are better and different than those that do not.
+Table \ref{tab:combined-mat} shows the distribution of couple types: WW (96\%), HH (2\%), WH (<1\%), and HW (1\%). Summary statistics for the parents are presented in Table \ref{tab:combined-sum}.
 
@@
 
-In this section, I present two empirical strategies.
-The first empirical strategy estimates the effect of having a Hispanic last name on annual log earnings.
-The second empirical strategy estimates the effect of identifying as Hispanic on log annual earnings.
+In this section, I present two empirical strategies. The first empirical strategy estimates the effect of having a Hispanic last name on educational outcomes and earnings.
 
@@
 
-%table6
-\input{tables/tab05-regression}
 \subsection{Estimating the effect of having a Hispanic last name}
@@
 In this section I restrict the sample to WH and HW groups.
-Let $Y_{ist}$ be the annual log earnings of person $i$ in state $s$ at time $t$.
-$HW_{ist}$ is an indicator variables for the type of parents person $i$ has.
-$X_{ist}$ is a vector of controls that includes age and numbers of hours worked, $\gamma_{t}$ are year fixed effects, $\lambda_{s}$ are state fixed effects, and $\phi_{ist}$ represents the error term.
+Let $Y_{ist}$ be the outcome of interest for person $i$ in state $s$ at time $t$.
+$HW_{ist}$ is an indicator variables for the type of parents of person $i$ has.
+$X_{ist}$ is a vector of controls that includes age and numbers of hours worked, $\gamma_{st}$ are state-year fixed effects (FE), and $\phi_{ist}$ represents the error term.
 The equation for this strategy is written as follows:
 \begin{equation} \label{eq:1a}
-Y_{ist} = \beta_{1} HW_{ist} + X_{ist} \pi + \gamma_{t} + \lambda_s + \phi_{ist}
+Y_{ist} = \beta_{1} HW_{ist} + X_{ist} \pi + \gamma_{st} + \phi_{ist}
 \end{equation}
@@
 $\beta_{1}$ is the coefficient of interest in this specification.
-$\beta_{1}$ represents the earnings gaps between children of inter-ethnic marriages who have a Spanish-sounding last name versus a White last name.
+$\beta_{1}$ represents the gaps in outcomes between children of inter-ethnic marriages who have a Spanish-sounding last name versus a White last name.
 
-\subsection{Estimating the effect of identifying as Hispanic}
+\subsection{Threats to Identification}
 
-In this section, I will present an estimation strategy that would allow me to capture the effect of identifying as a Hispanic.
+The central assumption of my estimation strategy is that individuals born to HW parents have comparable characteristics to their WH peers, particularly in areas crucial to the labor market.
+This assumption faces two main challenges.
+First, Selection in the marriage market: Given the potential societal penalties associated with Hispanic identity, the characteristics of White women who marry Hispanic men may be unique.
+Second, Differential parental influence: Fathers and mothers may impact human capital accumulation differently \citep{kimball2009risk,magruder2010intergenerational}.
 
-\begin{equation} \label{eq:iden}
-Y_{ist} = \beta_{1} HW_{ist} + \beta_{2} Hispanic_i + \beta_{3} HW_{ist} \cdot Hispanic_i + X_{ist} \pi + \gamma_{t} + \lambda_s + \phi_{ist}
-\end{equation}
+These concerns are mitigated by assortative matching in marriages.
+A substantial body of literature indicates strong selection in marriages based on traits \citep{averettBetterWorseRelationship2008, beckerTreatiseFamily1993,duncanIntermarriageIntergenerationalTransmission2011}.
+This suggests that inter-ethnic WH and HW marriages likely consist of partners with similar characteristics, supporting the comparability of their children in terms of labor market outcomes.
 
-The coefficients of interest from Eq.
-\eqref{eq:iden} are $\beta_{1}$, $\beta_{2}$ and $\beta_{3}$.
-The coefficients $\beta_{1}$ and $\beta_{3}$ capture the effects of Spanish-sounding last names for $HW_{ist}$ that identify as Hispanic \textit{versus} those who do not.
-%table7
-\input{tables/tab06-regression}
+A potential measurement error arises from using parents' birthplace in the CPS data as a proxy for ethnicity and last name.
+However, this concern is minimal as most Hispanics from 1960 to 2000 were first-generation immigrants.
+Census data shows only 3\% of native-born Americans identified as Hispanic, making it highly improbable for an inter-ethnic child to have a native-born, second-generation+ Hispanic father.\footnote{Self-reported Hispanic identity among first-generation Hispanic immigrants and native-borns are based on the author’s calculations from the 1960-2000 Census.
+The sample includes Whites, who are married, and are between the ages 25 and 40.} Empirical evidence from the CPS and US Census will be used to evaluate these concerns and support the validity of the identification strategy.
+
+\section{From the Data: The Differences Between HW and WH Couples}\label{sec:hw-wh-couples-data}
+
+In this section, I examine the data to validate my empirical strategy, focusing on educational and economic outcomes for four ethnic groupings—White White (WW), White Hispanic (WH), Hispanic White (HW), and Hispanic Hispanic (HH)—as shown in Table \ref{tab:combined-sum}.
+The data indicates marriage selection, with smaller differences between HW and WH children compared to the other groups, making them a more suitable comparison for analyzing labor market discrimination against Hispanics.
+Using the synthetic parents sample constructed from Census data, we observe that WW couples have the highest household education levels (24.95 years), while HH couples have the lowest (17.13 years).
+Inter-ethnic couples fall in between, with WH households totaling 22.68 years of schooling and HW households slightly behind at 21.50 years.
+Notably, HW wives are more educated than their WH peers, which could have implications for the educational attainment of HW children, as mothers play a crucial role in human capital accumulation.
+
+Regarding labor market outcomes, WW households have the highest log total family income (10.75), followed by WH (10.65), HW (10.60), and HH (10.42) households.
+The earnings gap between HW and WH men is marginal, with HW men earning 4\% less.
+However, HW women slightly out-earn WH women, with log hourly earnings of 1.75 and 1.73, respectively.
+This reversal in typical gender earning patterns suggests that HW women contribute more to their families' economic stability, potentially benefiting their children's development and future labor market prospects.
+Despite HW mothers' higher education and income levels, HW children complete, on average, 0.4 fewer years of education than their WH peers (Table \ref{tab:combined-sum}), potentially pointing to discrimination or barriers in educational access for HW children.
+
 \section{Results}\label{sec:results}
 
-In this section, I present the results from estimating the two specifications in Eqs.
-\eqref{eq:1a} and \eqref{eq:iden}.
-The results are shown in Tables \ref{tab:lastnamereg} and \ref{tab:identreg}.
-All estimations are done on a sample of White, native-born men between the ages of 25 and 40 employed for full-time full-year (FTFY) and are wage and salary workers.
+\subsection{Effect of Hispanic Last Name on Educational Outcomes}
 
-I find that an inter-ethnic person with a Spanish-sounding last names earns less than an inter-ethnic who has a White last name.
-A person with a Spanish-sounding last name earns 5 percentage points less than a person with a White last name.
-In other words, by comparing inter-ethnic children, a person with a Hispanic last name earns 5 percentage points less than someone with a White last name.
-However, more than half of the earnings gaps could be explained by educational differences.
-When I control for education, the last name effect decreases to a statistically insignificant 2 percentage points earnings gap.
-I also find a significant earnings gap between those that identify as Hispanic.
+Analysis of White, U.S.-born Hispanics ages 25-40 (Table \ref{tab:lastname-ed-reg}) reveals that HW children receive 0.2 years less education than WH children.
+While there is no significant difference in high school dropout rates, HW children are 2 percentage points less likely to earn an associate degree and 3 percentage points less likely to earn a bachelor's degree.
+These gaps are larger for HW women in higher education.
+These disparities suggest potential barriers or discrimination in access to higher education for HW children, particularly women.
 
-\subsection{The Effect of Having a Hispanic Last Name and Hispanic Identity on Labor Market Outcomes}
+\subsection{Effect on Labor Market Outcomes}
 
-I provide the results to the estimation of Eq.
-\eqref{eq:1a} in table \ref{tab:lastnamereg}.
-The sample in this analysis includes full-time, full-year, and wage and salary men, and I control for hours worked and age and include year fixed effects (FE).
-I also restrict the sample to children of HW and WH (omitted) parents.
-Column one in table \ref{tab:lastnamereg} is the results without controlling for education.
-Overall, an inter-ethnic with a Hispanic last name earns 5 percentage points less than an inter-ethnic with a White last name.
-This gap, however, could be entirely explained by educational differences.
-An inter-ethnic with a Hispanic last name earns 2 percentage points less than an inter-ethnic with a White last name, but the result is statistically insignificant.
-
-I provide the results to the estimation of Eq.
-\eqref{eq:iden} in table \ref{tab:identreg}.
-The sample in this analysis includes full-time, full-year, and wage and salary men, and I control for hours worked and age and include year fixed effects (FE).
-I also restrict the sample to children of HW and WH parents.
-The omitted group is children of WH parents that do not identify as Hispanic.
-Column one in table \ref{tab:identreg} is the results without controlling for education.
-An inter-ethnic $HW_{ist}$ that identifies as Hispanic \textit{versus} $HW_{ist}$ who do not---i.e.
-$HW_{ist} + HW_{ist} \cdot Hispanic_i$---earns 5 percentage points less than an inter-ethnic that does not, but the result could be entirely explained by educational differences.
+Results from Table \ref{tab:lastnamereg} indicate a 1 percentage point employment gap between HW and WH workers, and a 5 percentage point crude earnings gap.
+However, when controlling for education, the earnings gap reduces to a statistically insignificant 1 percentage point.
 
@@
 
-As the Hispanic population grows in the United States, studying discrimination against this group becomes increasingly important.
-In this paper, I examine discrimination against Hispanics in the labor market.
-More specifically, I examine the impact of Hispanic last names and Hispanic identification on annual log earnings.
+This study examines discrimination against Hispanics in the U.S.
+labor market, focusing on the impact of Hispanic last names and identification on education and earnings.
+The research finds that inter-ethnic individuals with Spanish-sounding last names receive 0.2 years less education and are more likely to be unemployed.
+HW children earn 5 percentage points less than WH children, but this gap becomes statistically insignificant when controlling for education.
 
-I compare the children of inter-ethnic marriages to study the labor market impact of having a Hispanic last name.
-When I compare the earnings of HW and WH children, which captures the effect of having a Hispanic last name, HW children earn  5 percentage points less than WH children.
-Thus, by comparing inter-ethnic children, a person with a Hispanic last name makes 5 percentage points less than someone with a White last name.
-When I control for education, the last name effect decreases to a statistically insignificant 2 percentage points earnings gap.
-Moreover, I estimate the effect of identifying as Hispanic and having a Hispanic last name.
-I find that an inter-ethnic child with a Hispanic last name and identifies as Hispanic earns 5 percentage points less that an inter-ethnic child with native sounding last name and identifies as Hispanic.
-The earnings gap, however, could be entirely explained by educational differences.
+These results align with existing literature on disparities in educational access \citep{bergman2018education,gaddis2024racial}.
+While the earnings gap diminishes when controlling for education, this doesn't necessarily indicate an absence of discrimination, as education itself can be influenced by bias.
 
-While the earnings gap between children of inter-ethnic parents with and without a Hispanic last name disappears when controlling for education, it does not necessarily indicate the absence of discrimination.
-Education itself can be influenced by bias, potentially resulting in divergent outcomes.
-Consequently, further research is essential to comprehensively understand the earnings gaps between Hispanics and Whites.
+Further research is needed to comprehensively understand the earnings gaps between Hispanics and Whites, particularly considering that mothers of HW children have higher education and earnings, which theoretically should lead to better outcomes for their children.
 
@@
 
+\newpage
+\clearpage
+
 \bibliographystyle{cas-model2-names}
 \bibliography{bibliography}
+\newpage
+\clearpage
+%table1
+\input{tables/tab16-tab1-and-tab2}
+
+\newpage
+\clearpage
+
+%table3
+\input{tables/tab17-all-sums}
+
+\newpage
+\clearpage
+
+\input{tables/tab13-regression}
+
+\newpage
+\clearpage
+
+\input{tables/tab15-emp-earn-regression.tex}
 

### 2024-08-28 manuscripts/06-labour-v5/manuscript_main.tex
....

--- 2024-08-11 manuscripts/labour-v5/manuscript_main.tex
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\section{Introduction}
 
@@
 
-\subsection*{Children of the four parental types}
+\subsection{Children of the four parental types}
 
@@
 
-\subsection*{Synthetic parents}
+\subsection{Synthetic parents}
 
@@
 
-\subsection*{Estimating the effect of having a Hispanic last name}
+\subsection{Estimating the effect of having a Hispanic last name}
 
@@
 
-\subsection*{Estimating the effect of identifying as Hispanic}
+\subsection{Estimating the effect of identifying as Hispanic}
 
@@
 
-\subsection*{Threats to Identification}
+\subsection{Threats to Identification}
 
@@
 
-\section*{From the Data: The Differences Between HW and WH Couples}\label{sec:hw-wh-couples-data}
+\section{From the Data: The Differences Between HW and WH Couples}\label{sec:hw-wh-couples-data}
 
@@
 
-\subsection*{The Effect of Having a Hispanic Last Name on Educational Outcomes}
+\subsection{The Effect of Having a Hispanic Last Name on Educational Outcomes}
 
@@
 HW children are 2 percentage points less likely to earn an associate degree and 3 percentage points less likely to earn a bachelor's degree compared to their WH peers.
+HW children are 2 percentage points less likely to earn an associate degree compared to their WH peers, representing a 13.3\% reduction relative to the WH associate degree rate of 15\%.
+HW children are 3 percentage points less likely to earn a bachelor's degree compared to their WH peers, which is a 13\% reduction relative to the WH bachelor's degree rate of 23\%.
 These gaps are slightly larger for HW women, who are 3 percentage points less likely to earn an associate degree and 4 percentage points less likely to earn a bachelor's degree.
@@
 These results suggest that while the overall educational attainment gap between HW and WH children is small in terms of years of education, the disparities become more pronounced when considering higher education milestones, particularly for HW women.
-However, given the fact that HWs have ,more educated mothers than WHs, we would expect them to have higher levels of educational attainment \autocite{kimball2009risk, gould2020does}.
+However, given the fact that HWs have more educated mothers than WHs, we would expect them to have higher levels of educational attainment \autocite{kimball2009risk, gould2020does}.
 This could indicate potential barriers or discrimination in access to higher education for HW children, particularly.
 
-\subsection*{The Effect of Having a Hispanic Last Name and Hispanic Identity on Labor Market Outcomes}
+\subsection{The Effect of Having a Hispanic Last Name and Hispanic Identity on Labor Market Outcomes}
 

### 2024-08-28 manuscripts/06-labour-v5/tab08-regression-weekearn-mex.tex
....

--- 2023-11-30 my_paper/tables/tab08-regression-weekearn-mex.tex
@@
 \begin{table}[H]
-
+\centering\centering
 \caption{Effect of Having Hispanic Last Name: Hispanics with Mexican Ancestry  \label{tab:lastnamereg-weekearm-mex}}
@@
 \midrule
-$HW_{ist}$ & \num{-0.01} & \num{-0.03}* & \num{0.00} & \num{-0.01}\\
- & (\num{0.02}) & (\num{0.02}) & (\num{0.01}) & (\num{0.01})\\
+$HW_{ist}$ & \num{-0.01} & \num{-0.04}** & \num{-0.01} & \num{-0.01}\\
+ & (\num{0.02}) & (\num{0.02}) & (\num{0.01}) & (\num{0.02})\\
 Constant & \num{6.53}*** &  &  & \\
@@
 Year FE &  & X & X & X\\
-Age FE &  & X & X & X\\
-Education FE &  &  & X & X\\
+State-Year FE &  & X & X & X\\
+Age &  & X & X & X\\
+Education &  &  & X & X\\
 Parental Background &  &  &  & X\\
-Observations & \num{4535} & \num{4535} & \num{4535} & \num{4535}\\
+Observations & \num{4515} & \num{4515} & \num{4515} & \num{4515}\\
 \bottomrule

### 2024-08-28 manuscripts/07-economic-letters-v5/Hadah-Hispanic-Last-Names.tex
....

--- 2024-08-28 manuscripts/05-Hussain-Hadah/Hadah-Hispanic-Last-Names.tex
@@
 This study analyzes the impact of Hispanic-sounding surnames on wages among inter-ethnic children with one White and one Hispanic parent.
-I find that individuals with Hispanic surnames often earn less, with a notable wage gap favoring those with White surnames.
+I find that individuals with Hispanic surnames receive less education, more likely to be unemployed, and often earn less, with a notable wage gap favoring those with White surnames.
 People born to Hispanic fathers and White mothers receive 0.3 years of education less than those born to White fathers and Hispanic mothers.
-Males born to Hispanic fathers and White mothers earn 5 percentage points less than those born to White fathers and Hispanic mothers, largely due to educational differences.
-Additionally, men with Spanish-sounding last names who identify as Hispanic earn significantly less, again due to educational disparities.
+Males born to Hispanic fathers and White mothers are 1 percentage point more likely to be unemployed, and they also earn 5 percentage points less than those born to White fathers and Hispanic mothers.
+The earnings gap is largely due to educational differences.
 \end{abstract}
@@
 Hispanics constitute a large and growing portion of the population in the United States.
-As the number of Hispanics increases, determining whether ethnic discrimination affects their labor market outcomes becomes increasingly crucial \citep{chettyUnitedStatesStill2014, chettyEffectsExposureBetter2016,chettyFadingAmericanDream2017,abramitzkyImmigrantsAssimilateMore2020a, abramitzkyNationImmigrantsAssimilation2014,abramitzkyCulturalAssimilationAge2016,chettyWhereLandOpportunity2014}.
-Thus, it is important to understand whether a person's ethnicity affects their labor market outcomes.
+As the number of Hispanics increases, determining whether ethnic discrimination influences their labor market outcomes becomes increasingly crucial \citep{chettyUnitedStatesStill2014, chettyEffectsExposureBetter2016,chettyFadingAmericanDream2017,abramitzkyImmigrantsAssimilateMore2020a, abramitzkyNationImmigrantsAssimilation2014,abramitzkyCulturalAssimilationAge2016,chettyWhereLandOpportunity2014}.
+Thus, it is important to understand how a person's ethnicity affects their labor market outcomes.
 Assimilation and mobility are crucial because they reflect how well Hispanics can integrate into society and move up the socioeconomic ladder.
 
-In this paper, I answer the following question.
-Does having a Hispanic last name affect education and labor market outcomes?
+In this paper, I answer the following question: does having a Hispanic last name affect education and labor market outcomes?
 Moreover, I aim to show that comparing Hispanic Whites to non-Hispanic Whites might create an artificially higher earnings gap since the two groups differ on many observable characteristics.
-\footnote{Observable characteristics refer to factors that can be measured and quantified, such as education level, work experience, and immigration status.} Others have attempted to compare how native-born White Hispanics fare to non-Hispanic Whites and foreign-born Hispanics.
+\footnote{Observable characteristics refer to factors that can be measured and quantified, such as education level, work experience, and immigration status.} Others have attempted to compare how native-born White Hispanics fare compared to non-Hispanic Whites and foreign-born Hispanics.
 In \citet{antman2020ethnic,antmanEthnicAttritionObserved2016,antmanEthnicAttritionObserved2016a,antmanEthnicAttritionAssimilation2020}, the authors compare the health and educational outcomes of Hispanic Whites to non-Hispanic Whites and native-born Hispanics to foreign-born Hispanics.
@@
 For example, discrimination against Hispanics can lead to reduced job opportunities, lower wages, and hinder assimilation.
-In this paper, I examine the role of having a Hispanic last name on labor market and educational outcomes outcomes.
+In this paper, I examine the role of having a Hispanic last name on labor market and educational outcomes.
 
-Identifying discrimination is difficult because of factors that affect labor market outcomes that are unobservable to economists---such as unobserved skills and separating it from prejudice and stereotypes.
+Identifying discrimination is difficult due to factors affecting labor market outcomes that are unobservable to economists, such as unobserved skills, and the challenge of separating these factors from prejudice and stereotypes.
 One strategy used by researchers is audit or resume studies.
@@
 
+I find that a person with a Hispanic last name receives less education, is more likely to be unemployed, and earns less.
+The results suggest that individuals with Hispanic last names receive 0.2 years less education and are more likely to be unemployed.
+The gap in education is equivalent to a 1.9\% reduction, and the unemployment gap is equal to a 14\% increase.
+The earnings gap between individuals with Hispanic last names and those with White last names is 5 percentage points.
+The earnings gap between individuals with Hispanic last names and those with White last names is 5 percentage points.
+However, when controlling for education, the earnings gap reduces to 1 percentage point.
+The results suggest that the earnings gap is largely due to differences in education.
+
 \section{Data}\label{sec:data}
@@
 
-The CPS sample is restricted to White, US-born citizens aged 25-40, born between 1960-2000.
-Parents are classified as Hispanic if born in a Spanish-speaking country or Puerto Rico, and White if US-born.
+The CPS sample is restricted to White, U.S.-born citizens aged 25-40, born between 1960 and 2000.
+Parents are classified as Hispanic if born in a Spanish-speaking country or Puerto Rico, and White if they are US-born.
 Four parental types are identified: (1) White father and White mother (WW), (2) White father and Hispanic mother (WH), (3) Hispanic father and White mother (HW), and (4) Hispanic father and Hispanic mother (HH).
@@
 
-The difference in means between Hispanics and non-Hispanic Whites could result from discrimination.
+The difference in means between Hispanics and non-Hispanic Whites could result from either discrimination or differences in innate abilities, skills, and parental investments.
 It can also be caused by differences in innate abilities, skills, and parental investments.
@@
 Let $Y_{ist}$ be the outcome of interest for person $i$ in state $s$ at time $t$.
-$HW_{ist}$ is an indicator variables for the type of parents of person $i$ has.
+$HW_{ist}$ is an indicator variable for the type of parents person $i$ has.
 $X_{ist}$ is a vector of controls that includes age and numbers of hours worked, $\gamma_{st}$ are state-year fixed effects (FE), and $\phi_{ist}$ represents the error term.
@@
 
-A potential measurement error arises from using parents' birthplace in the CPS data as a proxy for ethnicity and last name.
+A potential measurement error arises from using parents' birthplace in the CPS data as a proxy for both ethnicity and last name.
 However, this concern is minimal as most Hispanics from 1960 to 2000 were first-generation immigrants.
@@
 In this section, I examine the data to validate my empirical strategy, focusing on educational and economic outcomes for four ethnic groupings—White White (WW), White Hispanic (WH), Hispanic White (HW), and Hispanic Hispanic (HH)—as shown in Table \ref{tab:combined-sum}.
-The data indicates marriage selection, with smaller differences between HW and WH children compared to the other groups, making them a more suitable comparison for analyzing labor market discrimination against Hispanics.
+The data indicates marriage selection, with smaller differences between HW and WH children compared to the other groups.
+This similarity makes HW and WH children a more suitable comparison for analyzing labor market discrimination against Hispanics.
 Using the synthetic parents sample constructed from Census data, we observe that WW couples have the highest household education levels (24.95 years), while HH couples have the lowest (17.13 years).
 Inter-ethnic couples fall in between, with WH households totaling 22.68 years of schooling and HW households slightly behind at 21.50 years.
-Notably, HW wives are more educated than their WH peers, which could have implications for the educational attainment of HW children, as mothers play a crucial role in human capital accumulation.
+Notably, HW wives are more educated than their WH peers, which could have implications for the educational attainment of HW children, as mothers play a crucial role in human capital accumulation \citep{kimball2009risk,magruder2010intergenerational}.
 
@@
 
-Analysis of White, U.S.-born Hispanics ages 25-40 (Table \ref{tab:lastname-ed-reg}) reveals that HW children receive 0.2 years less education than WH children.
-While there is no significant difference in high school dropout rates, HW children are 2 percentage points less likely to earn an associate degree and 3 percentage points less likely to earn a bachelor's degree.
+Analysis of White, U.S.-born Hispanics ages 25-40 (Table \ref{tab:lastname-ed-reg}) reveals that HW children receive 0.2 years less education than WH children---that represents a 1.5\% reduction.
+While there is no significant difference in high school dropout rates, HW children are 2 percentage points less likely to earn an associate degree and 3 percentage points less likely to earn a bachelor's degree, which is equal to a 15\% reduction.
 These gaps are larger for HW women in higher education.
@@
 However, when controlling for education, the earnings gap reduces to a statistically insignificant 1 percentage point.
+I also find that HW children are 1 percentage point more likely to be unemployed than WH children.
+The gap in unemployment is equivalent to a 14\% increase and is statistically significant even after controlling for education.
+These results suggest that having a Hispanic last name negatively impacts labor market outcomes, particularly in terms of employment.
 
@@
 labor market, focusing on the impact of Hispanic last names and identification on education and earnings.
-The research finds that inter-ethnic individuals with Spanish-sounding last names receive 0.2 years less education and are more likely to be unemployed.
+This research finds that inter-ethnic individuals with Spanish-sounding last names receive 0.2 years less education and are more likely to be unemployed.
 HW children earn 5 percentage points less than WH children, but this gap becomes statistically insignificant when controlling for education.
@@
 While the earnings gap diminishes when controlling for education, this doesn't necessarily indicate an absence of discrimination, as education itself can be influenced by bias.
-
+Moreover, educational differences cannot explain the gaps in unemployment, suggesting that discrimination may play a role in labor market outcomes of individuals with Hispanic last names.
 Further research is needed to comprehensively understand the earnings gaps between Hispanics and Whites, particularly considering that mothers of HW children have higher education and earnings, which theoretically should lead to better outcomes for their children.

### 2024-08-28 manuscripts/07-economic-letters-v5/manuscript_main.tex
....

--- 2024-08-28 manuscripts/05-Hussain-Hadah/Hadah-Hispanic-Last-Names.tex
@@
-%%
-%% Copyright 2019-2021 Elsevier Ltd
-%%
-%% This file is part of the 'CAS Bundle'.
-%% --------------------------------------
-%%
-%% It may be distributed under the conditions of the LaTeX Project Public
-%% License, either version 1.2 of this license or (at your option) any
-%% later version.  The latest version of this license is in
-%%    http://www.latex-project.org/lppl.txt
-%% and version 1.2 or later is part of all distributions of LaTeX
-%% version 1999/12/01 or later.
-%%
-%% The list of all files belonging to the 'CAS Bundle' is
-%% given in the file `manifest.txt'.
-%%
-%% Template article for cas-sc documentclass for
-%% single column output.
-
-\documentclass[a4paper,fleqn]{cas-sc}
-\usepackage{setspace} % Load the setspace package for line spacing
-
-% Set line spacing to double spaced
-\doublespacing
-
-% \usepackage{mathptmx} % To set the font to Times New Roman
-% \let\cdot\relax
-\usepackage{amssymb}
-
-\usepackage{graphicx}
-\usepackage{caption}
-
-% Define a custom command to set the font for all tables
-\newcommand{\tablefont}{\fontfamily{ptm}\selectfont}
-
-% If the frontmatter runs over more than one page
-% use the longmktitle option.
-
-%\documentclass[a4paper,fleqn,longmktitle]{cas-sc}
-
-%\usepackage[numbers]{natbib}
-\usepackage[authoryear]{natbib}
-%\usepackage[authoryear,longnamesfirst]{natbib}
-
-%%%Author macros
-\def\tsc#1{\csdef{#1}{\textsc{\lowercase{#1}}\xspace}}
-\tsc{WGM}
-\tsc{QE}
-%%%
-
-% Uncomment and use as if needed
-%\newtheorem{theorem}{Theorem}
-%\newtheorem{lemma}[theorem]{Lemma}
-%\newdefinition{rmk}{Remark}
-%\newproof{pf}{Proof}
-%\newproof{pot}{Proof of Theorem \ref{thm}}
-
-\usepackage{lscape}
-\usepackage{threeparttablex}
-\usepackage{longtable}
-% Allow line breaks with \\ in specialcells
-\newcommand{\specialcell}[2][c]{%
-    \begin{tabular}[#1]{@{}c@{}}#2\end{tabular}
-}
-
-\usepackage{dcolumn}    % aligning decimals
-    \newcolumntype{d}[1]{D{.}{.}{#1}}
-
-\usepackage{siunitx}
-%%%%%%%%%%% TiKz %%%%%%%%%%%%%%%%%%%%%
-\usepackage{tikz}
-\usetikzlibrary{shapes.geometric, arrows}
-
-\tikzstyle{startstop} = [rectangle, rounded corners, minimum width=3cm, minimum height=1cm,text centered, draw=black, fill=red!30]
-
-\tikzstyle{io} = [trapezium, trapezium left angle=70, trapezium right angle=110, minimum width=3cm, minimum height=1cm, text centered, text width =5cm, draw=black, fill=blue!30]
-
-\tikzstyle{process} = [rectangle, minimum width=3cm, minimum height=1cm, text centered, text width =5cm, draw=black, fill=orange!30]
-
-\tikzstyle{decision} = [diamond, minimum width=3cm, minimum height=1cm, text centered, draw=black, fill=green!30]
-
-\tikzstyle{arrow} = [thick,->,>=stealth]
-\begin{document}
-
-\let\WriteBookmarks\relax
-\def\floatpagepagefraction{1}
-\def\textpagefraction{.001}
-
-% Short title
-\shorttitle{The Impact of Hispanic Last Names and Identity on Labor Market Outcomes}
-
-% Short author
-\shortauthors{Hadah}
-
-% Main title of the paper
-\title [mode = title]{The Impact of Hispanic Last Names and Identity on Labor Market Outcomes}
-
-\author[]{Hussain Hadah}[orcid=https://orcid.org/0000-0002-8705-6386]
-\ead{hhadah@tulane.edu}
-
-\affiliation[]{organization={Department of Economics, Tulane University},
-            addressline={6823 St. Charles Ave.},
-            city={New Orleans},
-            % citysep={}, % Uncomment if no comma needed between city and postcode
-            % postcode={70118},
-            state={LA 70118},
-            country={United States}}
-
-\begin{abstract}
-\singlespacing % Set single spacing for the abstract
-Do individuals with Hispanic names face labor market discrimination?
-This study analyzes the impact of Hispanic-sounding surnames on wages among inter-ethnic children with one White and one Hispanic parent.
-I find that individuals with Hispanic surnames often earn less, with a notable wage gap favoring those with White surnames.
-People born to Hispanic fathers and White mothers receive 0.3 years of education less than those born to White fathers and Hispanic mothers.
-Males born to Hispanic fathers and White mothers earn 5 percentage points less than those born to White fathers and Hispanic mothers, largely due to educational differences.
-Additionally, men with Spanish-sounding last names who identify as Hispanic earn significantly less, again due to educational disparities.
-\end{abstract}
-
-\begin{keywords}
-\singlespacing % Set single spacing for keywords
-
- Economics of Minorities, Race, and Immigrants\sep Discrimination and Prejudice\\
- \JEL{J15, J64, J71}
-\end{keywords}
-
-\maketitle
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-A large literature documents substantial earnings gaps across race and ethnicity \citep{bayer2018divergent, charles2008prejudice, card1992school, fryer2004causes, rubinstein2014pride, bertrand2004emily, juhn1991accounting}.
+A large literature documents substantial earnings gaps across race and ethnicity \autocite{bayer2018divergent, charles2008prejudice, card1992school, fryer2004causes, rubinstein2014pride, bertrand2004emily, juhn1991accounting}.
 Hispanics constitute a large and growing portion of the population in the United States.
-As the number of Hispanics increases, determining whether ethnic discrimination affects their labor market outcomes becomes increasingly crucial \citep{chettyUnitedStatesStill2014, chettyEffectsExposureBetter2016,chettyFadingAmericanDream2017,abramitzkyImmigrantsAssimilateMore2020a, abramitzkyNationImmigrantsAssimilation2014,abramitzkyCulturalAssimilationAge2016,chettyWhereLandOpportunity2014}.
+As the number of Hispanics increases, determining whether ethnic discrimination affects their labor market outcomes becomes increasingly crucial \autocite{chettyUnitedStatesStill2014, chettyEffectsExposureBetter2016,chettyFadingAmericanDream2017,abramitzkyImmigrantsAssimilateMore2020a, abramitzkyNationImmigrantsAssimilation2014,abramitzkyCulturalAssimilationAge2016,chettyWhereLandOpportunity2014}.
 Thus, it is important to understand whether a person's ethnicity affects their labor market outcomes.
@@
 \footnote{Observable characteristics refer to factors that can be measured and quantified, such as education level, work experience, and immigration status.} Others have attempted to compare how native-born White Hispanics fare to non-Hispanic Whites and foreign-born Hispanics.
-In \citet{antman2020ethnic,antmanEthnicAttritionObserved2016,antmanEthnicAttritionObserved2016a,antmanEthnicAttritionAssimilation2020}, the authors compare the health and educational outcomes of Hispanic Whites to non-Hispanic Whites and native-born Hispanics to foreign-born Hispanics.
+In \textcite{antman2020ethnic,antmanEthnicAttritionObserved2016,antmanEthnicAttritionObserved2016a,antmanEthnicAttritionAssimilation2020}, the authors compare the health and educational outcomes of Hispanic Whites to non-Hispanic Whites and native-born Hispanics to foreign-born Hispanics.
 They find gaps in education and health between Hispanics and Whites.
 They also find that native-born Hispanics are more likely than their foreign-born counterparts to report poor health.
-\citet{davilaChangesRelativeEarnings2008} documents many gaps in labor market outcomes between Hispanics and Whites.
+\textcite{davilaChangesRelativeEarnings2008} documents many gaps in labor market outcomes between Hispanics and Whites.
 They attribute a big part of this gap to differences in education, experience, immigration status, and regional differences.
@@
 
-I use two datasets: the IPUMS Current Population Survey (CPS) Annual Social and Economic (ASEC) \citep{cps2019} and the 1960 to 2000 US censuses \citep{acs2019}.
+I use two datasets: the IPUMS Current Population Survey (CPS) Annual Social and Economic (ASEC) \autocite{cps2019} and the 1960 to 2000 US censuses \autocite{acs2019}.
 
 The CPS data (1994-2019) is used to study the effect of Hispanic last names on labor market outcomes.
-The census data is employed to construct synthetic parents using the method developed by \citet{rubinstein2014pride}.
+The census data is employed to construct synthetic parents using the method developed by \textcite{rubinstein2014pride}.
 
@@
 
-Table \ref{tab:combined-mat} shows the distribution of these types.
+Table \ref{tab:mat1} shows the distribution of these types.
 WW children comprise 96\% of the sample, HH 3\%, and inter-ethnic children (WH and HW) 1.35\%.
-Summary statistics for each group are presented in Table \ref{tab:combined-sum}.
+Summary statistics for each group are presented in Table \ref{tab:c&p1}.
 
@@
 
-Table \ref{tab:combined-mat} shows the distribution of couple types: WW (96\%), HH (2\%), WH (<1\%), and HW (1\%). Summary statistics for the parents are presented in Table \ref{tab:combined-sum}.
+Table \ref{tab:mat2} shows the distribution of couple types: WW (96\%), HH (2\%), WH (<1\%), and HW (1\%). Summary statistics for the parents are presented in Table \ref{tab:synth}.
 
@@
 First, Selection in the marriage market: Given the potential societal penalties associated with Hispanic identity, the characteristics of White women who marry Hispanic men may be unique.
-Second, Differential parental influence: Fathers and mothers may impact human capital accumulation differently \citep{kimball2009risk,magruder2010intergenerational}.
+Second, Differential parental influence: Fathers and mothers may impact human capital accumulation differently \autocite{kimball2009risk,magruder2010intergenerational}.
 
 These concerns are mitigated by assortative matching in marriages.
-A substantial body of literature indicates strong selection in marriages based on traits \citep{averettBetterWorseRelationship2008, beckerTreatiseFamily1993,duncanIntermarriageIntergenerationalTransmission2011}.
+A substantial body of literature indicates strong selection in marriages based on traits \autocite{averettBetterWorseRelationship2008, beckerTreatiseFamily1993,duncanIntermarriageIntergenerationalTransmission2011}.
 This suggests that inter-ethnic WH and HW marriages likely consist of partners with similar characteristics, supporting the comparability of their children in terms of labor market outcomes.
@@
 However, this concern is minimal as most Hispanics from 1960 to 2000 were first-generation immigrants.
-Census data shows only 3\% of native-born Americans identified as Hispanic, making it highly improbable for an inter-ethnic child to have a native-born, second-generation+ Hispanic father.\footnote{Self-reported Hispanic identity among first-generation Hispanic immigrants and native-borns are based on the author’s calculations from the 1960-2000 Census.
-The sample includes Whites, who are married, and are between the ages 25 and 40.} Empirical evidence from the CPS and US Census will be used to evaluate these concerns and support the validity of the identification strategy.
+Census data shows only 3\% of native-born Americans identified as Hispanic (Table \ref{tab:mat3}), making it highly improbable for an inter-ethnic child to have a native-born, second-generation+ Hispanic father.
+Empirical evidence from the CPS and US Census will be used to evaluate these concerns and support the validity of the identification strategy.
 
@@
 
-In this section, I examine the data to validate my empirical strategy, focusing on educational and economic outcomes for four ethnic groupings—White White (WW), White Hispanic (WH), Hispanic White (HW), and Hispanic Hispanic (HH)—as shown in Table \ref{tab:combined-sum}.
+In this section, I examine the data to validate my empirical strategy, focusing on educational and economic outcomes for four ethnic groupings—White White (WW), White Hispanic (WH), Hispanic White (HW), and Hispanic Hispanic (HH)—as shown in Table \ref{tab:synth}.
 The data indicates marriage selection, with smaller differences between HW and WH children compared to the other groups, making them a more suitable comparison for analyzing labor market discrimination against Hispanics.
@@
 This reversal in typical gender earning patterns suggests that HW women contribute more to their families' economic stability, potentially benefiting their children's development and future labor market prospects.
-Despite HW mothers' higher education and income levels, HW children complete, on average, 0.4 fewer years of education than their WH peers (Table \ref{tab:combined-sum}), potentially pointing to discrimination or barriers in educational access for HW children.
+Despite HW mothers' higher education and income levels, HW children complete, on average, 0.4 fewer years of education than their WH peers (Table \ref{tab:c&p1}), potentially pointing to discrimination or barriers in educational access for HW children.
 
@@
 
-Results from Table \ref{tab:lastnamereg} indicate a 1 percentage point employment gap between HW and WH workers, and a 5 percentage point crude earnings gap.
+Results from Tables \ref{tab:lastnamereg-emp} and \ref{tab:lastnamereg} indicate a 1 percentage point employment gap between HW and WH workers, and a 5 percentage point crude earnings gap.
 However, when controlling for education, the earnings gap reduces to a statistically insignificant 1 percentage point.
@@
 
-These results align with existing literature on disparities in educational access \citep{bergman2018education,gaddis2024racial}.
+These results align with existing literature on disparities in educational access \autocite{bergman2018education,gaddis2024racial}.
 While the earnings gap diminishes when controlling for education, this doesn't necessarily indicate an absence of discrimination, as education itself can be influenced by bias.
@@
 Further research is needed to comprehensively understand the earnings gaps between Hispanics and Whites, particularly considering that mothers of HW children have higher education and earnings, which theoretically should lead to better outcomes for their children.
-
-\section*{Declaration of Competing Interest}
-The author declares that he has no relevant or material financial interests that relate to the research described in this paper.
-
-\section*{Data availability}
-
-Data is available \href{https://github.com/hhadah/hispanic-last-names/tree/main/data/datasets}{here}.
-
-\newpage
-\clearpage
-
-\bibliographystyle{cas-model2-names}
-\bibliography{bibliography}
-\newpage
-\clearpage
-%table1
-\input{tables/tab16-tab1-and-tab2}
-
-\newpage
-\clearpage
-
-%table3
-\input{tables/tab17-all-sums}
-
-\newpage
-\clearpage
-
-\input{tables/tab13-regression}
-
-\newpage
-\clearpage
-
-\input{tables/tab15-emp-earn-regression.tex}
-
-\end{document}

### 2024-08-28 manuscripts/07-economic-letters-v5/manuscript_title_etc.tex
....

--- 2024-08-11 manuscripts/labour-v5/manuscript_title_etc.tex
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\newcommand{\PAPERKEYWORDS}{\textbf{Keywords}: Economics of Minorities, Race, and Immigrants; Discrimination and Prejudice}
+\newcommand{\PAPERKEYWORDS}{\hspace{1em}\makebox{\textbf{Keywords}: Economics of Minorities, Race, and Immigrants; Discrimination and Prejudice}}
 \newcommand{\PAPERJEL}{\textbf{JEL}: J71; J64; J15}
@@
 \newcommand{\ACKNOWLEDGMENTS}{
-I thank Professors Willa Friedman, Chinhui Juhn, Vikram Maheshri, and Yona Rubinstein for their support and advice.
+I thank Patrick Button, Willa Friedman, Chinhui Juhn, Vikram Maheshri, and Yona Rubinstein for their support and advice.
 I also thank Aimee Chin, Steven Craig, German Cubas, Elaine Liu, Fan Wang, and the participants of the Applied Microeconomics Workshop at the University of Houston, and European Society for Population Economics (ESPE) for helpful feedback.}

### 2024-08-28 manuscripts/07-economic-letters-v5/tab14-emp-regression.tex
....

\begin{table}[H]
\centering\centering
\caption{Effect of Having Hispanic Last Name on Employment \label{tab:lastnamereg-emp}}
\centering
\begin{threeparttable}
\begin{tabular}[t]{lcccc}
\toprule
  & \specialcell{(1) \\ Unemployment} & \specialcell{(2) \\ Unemployment} & \specialcell{(3) \\  Unemployment} & \specialcell{(4) \\  Unemployment}\\
\midrule
$HW_{ist}$ & \num{0.01}*** & \num{0.01}** & \num{0.01}* & \num{0.01}\\
 & (\num{0.00}) & (\num{0.00}) & (\num{0.00}) & (\num{0.00})\\
Constant & \num{0.05}*** &  &  & \\
 & (\num{0.00}) &  &  & \\
\midrule
\textit{Controlling for:} &  &  &  & \\
Age &  & X & X & X\\
State FE &  & X & X & X\\
Year FE &  & X & X & X\\
State-Year FE &  & X & X & X\\
Education &  &  & X & X\\
Parental Background &  &  &  & X\\
\specialcell{HW's Unemployment} & 0.07 & 0.07 & 0.07 & 0.07\\
Observations & \num{38090} & \num{38090} & \num{38090} & \num{38090}\\
\bottomrule
\multicolumn{5}{l}{\rule{0pt}{1em}* p $<$ 0.1, ** p $<$ 0.05, *** p $<$ 0.01}\\
\end{tabular}
\begin{tablenotes}
\item[1] {\setstretch{1.0}\footnotesize{This table includes the estimation results of equation (\ref{eq:1a}).}}
\item[2] {\setstretch{1.0}\footnotesize{HW is an indicator variable that is equal to 1 if a person is the child of a Hispanic-father and White-mother.}}
\item[3] {\setstretch{1.0}\footnotesize{The sample is restricted to prime-age men.}}
\item[4] {\setstretch{1.0}\footnotesize{Standard errors are clustered on the state level.}}
\end{tablenotes}
\end{threeparttable}
\end{table}

### 2024-08-28 manuscripts/07-economic-letters-v5/tab16-tab1-and-tab2.tex
....

\begin{table}[H]
\centering
\caption{Number of Children and Couples by Parental and Couples' Type}
\label{tab:combined}
\resizebox{\linewidth}{!}{
\begin{threeparttable}
\begin{tabular}{lcccc}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{4}{c}{Parental/Couples' Type} \\
\cmidrule(l{3pt}r{3pt}){2-5}
 & \specialcell{White Father/ \\ White Mother \\ White Husband/ \\ White Wife \\ (WW)} 
 & \specialcell{White Father/ \\ Hispanic Mother \\ White Husband/ \\ Hispanic Wife \\ (WH)} 
 & \specialcell{Hispanic Father/ \\ White Mother \\ Hispanic Husband/ \\ White Wife \\ (HW)} 
 & \specialcell{Hispanic Father/ \\ Hispanic Mother \\ Hispanic Husband/ \\ Hispanic Wife \\ (HH)}\\
\midrule
\textbf{\specialcell{Observations \\ (Children)}} 
& \specialcell{6,421,328\\0.96} & \specialcell{39,048\\0.01} & \specialcell{51,277\\0.01} & \specialcell{179,827\\0.03} \\
\midrule
\textbf{\specialcell{Observations \\ (Couples)}} 
& \specialcell{5,141,737\\0.96} & \specialcell{33,097\\0.01} & \specialcell{37,847\\0.01} & \specialcell{119,749\\0.02} \\
\bottomrule
\end{tabular}
\begin{tablenotes}
\item[1] Source (Children): Current Population Surveys (CPS) 1994-2019.
\item[2] Source (Couples): 1960-2000 Census.
\item[3] The sample includes Whites, who are married and between the ages 25 and 40. Ethnicity of a person's parents are identified by the parent's place of birth. A parent is Hispanic if she/he was born in a Spanish-speaking country. A parent is White if she/he was born in the United States.
\end{tablenotes}
\end{threeparttable}}
\end{table}

### 2024-08-28 manuscripts/07-economic-letters-v5/tables/tab13-regression.tex
....

--- 2024-08-11 output/tables/tab13-regression.tex
@@
 \centering\centering
-\caption{Effect of Having Hispanic Last Name \label{tab:lastname-ed-reg}}
+\caption{Effect of Having Hispanic Last Name on Educational Outcomes\label{tab:lastname-ed-reg}}
 \centering
@@
 \toprule
-  & \specialcell{(1) \\ Years of \\ Education} & \specialcell{(2) \\ Highschool \\ Dropout} & \specialcell{(3) \\ Associate Degree} & \specialcell{(4) \\ Bachelor Degree}\\
+  & \specialcell{(1) \\ Years of \\ Education} & \specialcell{(2) \\ High School \\ Dropout} & \specialcell{(3) \\ Associate Degree} & \specialcell{(4) \\ Bachelor Degree}\\
 \midrule
@@
 \hspace{1em} & (0.07) & (0.01) & (0.01) & (0.01)\\
-\hspace{1em}Full Sample's Mean & 13.48 & 0.43 & 0.15 & 0.26\\
 \hspace{1em}Observations & 41861 & 42725 & 32593 & 42725\\
+Full Sample's Mean & 13.48 & 0.43 & 0.15 & 0.26\\
 Women's Mean & 13.58 & 0.44 & 0.16 & 0.27\\

### 2024-08-28 manuscripts/07-economic-letters-v5/tables/tab15-emp-earn-regression.tex
....

\begin{table}[H]
\centering
\caption{Effect of Having Hispanic Last Name on Employment and Log Annual Earnings}
\label{tab:lastnamereg}
\begin{threeparttable}
\begin{tabular}{lcccc}
\toprule
  & \specialcell{(3) \\ Log Annual \\ Earnings} & \specialcell{(4) \\ Log Annual \\ Earnings} & \specialcell{(2) \\ Unemployment} & \specialcell{(3) \\ Unemployment} \\
\midrule
$HW_{ist}$ & \num{-0.05}* & \num{-0.02} & \num{0.01}** & \num{0.01}* \\
 & (\num{0.02}) & (\num{0.02}) & (\num{0.00}) & (\num{0.00}) \\
Constant &  &  &  & \\
 &  &  &  & \\
\midrule
\textit{Controlling for:} & & & & \\
Education &  & X &  & X \\
Observations & \num{3621} & \num{3621} & \num{38090} & \num{38090} \\
\bottomrule
\multicolumn{5}{l}{\rule{0pt}{1em}* p $<$ 0.1, ** p $<$ 0.05, *** p $<$ 0.01}\\
\end{tabular}
\begin{tablenotes}
\item[1] {\setstretch{1.0}\footnotesize{This table includes the estimation results of equation (\ref{eq:1a}).}}
\item[2] {\setstretch{1.0}\footnotesize{HW is an indicator variable that is equal to 1 if a person is the child of a Hispanic-father and White-mother.}}
\item[3] {\setstretch{1.0}\footnotesize{The sample is restricted to prime-age men for unemployment and full-time full-year wage and salary workers for earnings.}}
\item[4] {\setstretch{1.0}\footnotesize{I control for hours worked, age, and state-year fixed effects for earnings. I control for age, and state-year fixed effects for earnings.}}

\item[5] {\setstretch{1.0}\footnotesize{Standard errors are clustered at the state level.}}
\end{tablenotes}
\end{threeparttable}
\end{table}

### 2024-08-28 programs/16-table-five-regression.R

@@ -48,47 +48,130 @@ reg1 <- list(
   
 )
 
-# feols(lninctot_1999 ~ .[ParentDummies] + .[ParentControls], vcov = ~statefip,
-#                                                             data = IndividualData |> 
-#                                                           filter(sex == 1 & FTFY == 1 & Self_employed_ASEC == 0 & (HW == 1 | WH == 1)))
+reg2 <- list(
+  "\\specialcell{(1) \\\\ Unemployment}" = feols(1 - Employed ~ HW, vcov = ~statefip,
+                                                            data = IndividualData |> 
+                                                          filter(sex == 1 & (HW == 1 | WH == 1))),
+  "\\specialcell{(2) \\\\ Unemployment}" = feols(1 - Employed ~ HW + age | year*statefip, vcov = ~statefip,
+                                                            data = IndividualData |> 
+                                                          filter(sex == 1 & (HW == 1 | WH == 1))),
+  "\\specialcell{(3) \\\\  Unemployment}" = feols(1 - Employed ~ HW + age + educ | year*statefip, vcov = ~statefip,
+                                                         data = IndividualData |> 
+                                                           filter(sex == 1 & (HW == 1 | WH == 1))),
+  "\\specialcell{(4) \\\\  Unemployment}" = feols(1 - Employed ~ HW + .[ParentControls] + age + educ | year*statefip, vcov = ~statefip,
+                                                         data = IndividualData |> 
+                                                           filter(sex == 1 & (HW == 1 | WH == 1)))
+  
+)
+reg2[[1]]
+reg2[[2]]
+reg2[[3]]
+reg2[[4]]
 
-# feols(lninctot_1999 ~ .[ParentDummies] + .[ParentControls] | year + age + statefip, vcov = ~statefip,
-#                                                             data = IndividualData |> 
-#                                                           filter(sex == 1 & FTFY == 1 & Self_employed_ASEC == 0 & (HW == 1 | WH == 1)))
+reg3 <- list(
+  "\\specialcell{(1) \\\\ In Labor Force}" = feols(InLaborForce ~ HW, vcov = ~statefip,
+                                                            data = IndividualData |> 
+                                                          filter(sex == 1 & (HW == 1 | WH == 1))),
[…]

### 2024-08-28 programs/95-make-all.R

@@ -20,7 +20,7 @@ thesis_tabs  <- paste0(mdir_git,"/my_paper/tables")
 thesis_plots <- paste0(mdir_git,"/my_paper/figure") 
 add_health_data_dir <- "/Users/hhadah/Dropbox/Research/My Research Data and Ideas/Identification_Paper/Data/Datasets"
 ACS_data  <- "/Users/hhadah/Documents/GiT/jmp-decomposition-replication/data/raw"
-manuscript_wd <- paste0(mdir_git,"/manuscripts/labour-v5") 
+manuscript_wd <- paste0(mdir_git,"/manuscripts/07-economic-letters-v5") 
 # pres_tabs <- paste0(mdir,"/presentations/jmp-presentation/tables")
 # pres_plots <- paste0(mdir,"/presentations/jmp-presentation/figures")
 

### 2024-09-06 manuscripts/07-economic-letters-v5/Hadah-Hispanic-Last-Names.tex
paper less than 12 pages and 2000 words

@@
 
-A large literature documents substantial earnings gaps across race and ethnicity \citep{bayer2018divergent, charles2008prejudice, card1992school, fryer2004causes, rubinstein2014pride, bertrand2004emily, juhn1991accounting}.
-Hispanics constitute a large and growing portion of the population in the United States.
-As the number of Hispanics increases, determining whether ethnic discrimination influences their labor market outcomes becomes increasingly crucial \citep{chettyUnitedStatesStill2014, chettyEffectsExposureBetter2016,chettyFadingAmericanDream2017,abramitzkyImmigrantsAssimilateMore2020a, abramitzkyNationImmigrantsAssimilation2014,abramitzkyCulturalAssimilationAge2016,chettyWhereLandOpportunity2014}.
-Thus, it is important to understand how a person's ethnicity affects their labor market outcomes.
-Assimilation and mobility are crucial because they reflect how well Hispanics can integrate into society and move up the socioeconomic ladder.
+A substantial earnings gap exists between native-born White Hispanic men and White men, with Hispanic men earning 21\% less.
+While a significant portion of this gap is attributed to educational differences, discrimination may also play a role, potentially leading to reduced job opportunities, lower wages, and hindered assimilation \citep{duncan2006hispanics, duncan2018identifying, duncan2018socioeconomic}.
+This paper examines the impact of Hispanic last names on labor market and educational outcomes, contributing to the literature by isolating the effect of ethnicity-signaling surnames from other potentially confounding factors.
+By focusing on children of inter-ethnic marriages, this study provides a novel approach to understanding the mechanisms behind Hispanic-White earnings disparities and the potential role of name-based discrimination in the labor market.
 
-In this paper, I answer the following question: does having a Hispanic last name affect education and labor market outcomes?
-Moreover, I aim to show that comparing Hispanic Whites to non-Hispanic Whites might create an artificially higher earnings gap since the two groups differ on many observable characteristics.
-\footnote{Observable characteristics refer to factors that can be measured and quantified, such as education level, work experience, and immigration status.} Others have attempted to compare how native-born White Hispanics fare compared to non-Hispanic Whites and foreign-born Hispanics.
-In \citet{antman2020ethnic,antmanEthnicAttritionObserved2016,antmanEthnicAttritionObserved2016a,antmanEthnicAttritionAssimilation2020}, the authors compare the health and educational outcomes of Hispanic Whites to non-Hispanic Whites and native-born Hispanics to foreign-born Hispanics.
-They find gaps in education and health between Hispanics and Whites.
-They also find that native-born Hispanics are more likely than their foreign-born counterparts to report poor health.
-\citet{davilaChangesRelativeEarnings2008} documents many gaps in labor market outcomes between Hispanics and Whites.
-They attribute a big part of this gap to differences in education, experience, immigration status, and regional differences.
+This study investigates whether having a Hispanic last name affects educational and labor market outcomes.
+Previous research has identified gaps in education, health, and labor market outcomes between Hispanics and Whites \citep{antman2020ethnic, davilaChangesRelativeEarnings2008}.
+However, comparing Hispanic Whites to non-Hispanic Whites might artificially inflate the earnings gap.
+This is due to observable differences in education level, work experience, and immigration status.
 
-The US population is growing in diversity.
-The proportion of non-Whites has increased by more than 10 percentage points from 13 percent in 1995 to 23 percent in 2019.
-The number of Hispanics has grown by 9 percentage points from 9 percent in 1995 to 16 percent in 2019.\footnote{The portion of non-Whites and Hispanics is calculated using the Current Population Survey (CPS).} Native-born White Hispanic men earn 21\% less than White men, although a substantial portion of the earnings gap is due to educational differences between Hispanics and Whites \citep{duncan2006hispanics, duncan2018identifying, duncan2018socioeconomic}.
-Some of the earnings differences may also be due to discrimination which will have negative consequences.
-For example, discrimination against Hispanics can lead to reduced job opportunities, lower wages, and hinder assimilation.
-In this paper, I examine the role of having a Hispanic last name on labor market and educational outcomes.
+Identifying discrimination is challenging due to unobservable factors affecting labor market outcomes.
+While audit studies provide insights, they primarily measure callback rates rather than full labor market outcomes \citep{bertrand2004emily}.
+This study compares children from inter-ethnic marriages: those of Hispanic fathers and White mothers (HW) to those of White fathers and Hispanic mothers (WH) \citep{rubinstein2014pride}.
+This approach leverages marriage selection based on observable characteristics \citep{averettBetterWorseRelationship2008}, allowing for the investigation of how ethnic signals, such as Hispanic last names, affect earnings.
 
-Identifying discrimination is difficult due to factors affecting labor market outcomes that are unobservable to economists, such as unobserved skills, and the challenge of separating these factors from prejudice and stereotypes.
-One strategy used by researchers is audit or resume studies.
-\citet{bertrand2004emily} conducted an audit study where identical resumes were sent to employers with White and Black-sounding names.
-This approach, however, has its drawbacks.
-Audit studies only observe callbacks, not wages.
-
-This study utilizes a method developed by \citet{rubinstein2014pride}.
-I compare children from inter-ethnic marriages.
-More precisely, I compare children of Hispanic fathers and White mothers (henceforth HW) to children of White fathers and Hispanic mothers (henceforth  WH).
-This approach stems from the fact that there is a strong selection among many characteristics, and thus, marriages are not random.
-Couples match on several observable characteristics like income, schooling, socio-economic background, etc.
-\citep{averettBetterWorseRelationship2008, averettEconomicRealityBeauty1996, beckerTheoryMarriagePart1973, beckerTheoryMarriagePart1974, beckerTreatiseFamily1993, browningCollectiveUnitaryModels2006, chiapporiFatterAttractionAnthropometric2012}.
-Children of HW and WH marriages have more similar observable characteristics than children of endogamous/homogamous marriages--- i.e., White fathers-White mothers and Hispanic fathers-Hispanic mothers.
-Moreover, children from a Hispanic father and White mother household will have a Hispanic last name from their fathers, enabling the investigation of how ethnic signals, such as having a Hispanic last name, affect annual log earnings.
-
-The main identifying assumption of my empirical strategy depends on the assumption that people born to HW parents are similar to their WH peer in all observable---and unobservable---aspects and characteristics that are important in the labor market.
-Consequently, the only difference between the two groups is the variation in which group is more likely to have a Hispanic sounding last name.
-Children from mixed ethnic backgrounds may appear physically similar to those from single ethnicity backgrounds, but the influence of family dynamics and upbringing, crucial in developing skills and personal characteristics, varies with the pattern of mixed ethnic marriages.
-Particularly in a society where Hispanic ancestry is perceived negatively, it raises the question of what kind of White women would choose Hispanic men.
-Furthermore, even if such unions were formed randomly, children from White-Hispanic homes might benefit from more favorable family conditions than those from Hispanic-White homes, considering Whites generally have stronger socio-economic backgrounds than Hispanics.
-These factors introduce doubts regarding whether children from Hispanic-White families receive comparable familial support and influences, either genetically or environmentally, as those from White-Hispanic families.
-
-I find that a person with a Hispanic last name receives less education, is more likely to be unemployed, and earns less.
-The results suggest that individuals with Hispanic last names receive 0.2 years less education and are more likely to be unemployed.
+I find that individuals with Hispanic-sounding last names receive 0.2 years less education and are more likely to be unemployed.
 The gap in education is equivalent to a 1.9\% reduction, and the unemployment gap is equal to a 14\% increase.
-The earnings gap between individuals with Hispanic last names and those with White last names is 5 percentage points.
-The earnings gap between individuals with Hispanic last names and those with White last names is 5 percentage points.
-However, when controlling for education, the earnings gap reduces to 1 percentage point.
-The results suggest that the earnings gap is largely due to differences in education.
+The earnings gap between individuals with Hispanic-sounding last names and those with White last names is 5 percentage points.
+However, when controlling for education, the earnings gap reduces to 1 percentage point, suggesting that the earnings gap is largely due to differences in education.
 
@@
 I use two datasets: the IPUMS Current Population Survey (CPS) Annual Social and Economic (ASEC) \citep{cps2019} and the 1960 to 2000 US censuses \citep{acs2019}.
-
-The CPS data (1994-2019) is used to study the effect of Hispanic last names on labor market outcomes.
+The CPS data (1994-2019) is used to study the effect of Hispanic-sounding last name on labor market outcomes.
 The census data is employed to construct synthetic parents using the method developed by \citet{rubinstein2014pride}.
-
-\subsection{Children of the four parental types}
 
@@
 Four parental types are identified: (1) White father and White mother (WW), (2) White father and Hispanic mother (WH), (3) Hispanic father and White mother (HW), and (4) Hispanic father and Hispanic mother (HH).
-
 Table \ref{tab:combined-mat} shows the distribution of these types.
@@
 Summary statistics for each group are presented in Table \ref{tab:combined-sum}.
-
-\subsection{Synthetic parents}
-
 Using the 1960-2000 censuses, I constructed a dataset of synthetic parents, including married White men and women born between 1920-1975.
 Hispanic individuals are defined as White and born in a Spanish-speaking country, while Whites are native-born.
-
-Table \ref{tab:combined-mat} shows the distribution of couple types: WW (96\%), HH (2\%), WH (<1\%), and HW (1\%). Summary statistics for the parents are presented in Table \ref{tab:combined-sum}.
+Table \ref{tab:combined-mat} shows the distribution of couple types: WW (96\%), HH (2\%), WH (<1\%), and HW (1\%).
+Summary statistics for the parents are presented in Table \ref{tab:combined-sum}.
 
@@
 
-In this section, I present two empirical strategies. The first empirical strategy estimates the effect of having a Hispanic last name on educational outcomes and earnings.
+In this section, I present my empirical strategy. The empirical strategy estimates the effect of having a Hispanic last name on educational outcomes and earnings.
 
@@
 
-\subsection{Effect of Hispanic Last Name on Educational Outcomes}
-
-Analysis of White, U.S.-born Hispanics ages 25-40 (Table \ref{tab:lastname-ed-reg}) reveals that HW children receive 0.2 years less education than WH children---that represents a 1.5\% reduction.
+My analysis of White, U.S.-born Hispanics ages 25-40 (Table \ref{tab:lastname-ed-reg}) reveals that HW children receive 0.2 years less education than WH children---that represents a 1.5\% reduction.
 While there is no significant difference in high school dropout rates, HW children are 2 percentage points less likely to earn an associate degree and 3 percentage points less likely to earn a bachelor's degree, which is equal to a 15\% reduction.
@@
 
-\subsection{Effect on Labor Market Outcomes}
-
-Results from Table \ref{tab:lastnamereg} indicate a 1 percentage point employment gap between HW and WH workers, and a 5 percentage point crude earnings gap.
+I show in Table \ref{tab:lastnamereg} a 1 percentage point employment gap between HW and WH workers, and a 5 percentage point crude earnings gap.
 However, when controlling for education, the earnings gap reduces to a statistically insignificant 1 percentage point.
@@
 This study examines discrimination against Hispanics in the U.S.
-labor market, focusing on the impact of Hispanic last names and identification on education and earnings.
+labor market, focusing on the impact of Hispanic-sounding last name and identification on education and earnings.
 This research finds that inter-ethnic individuals with Spanish-sounding last names receive 0.2 years less education and are more likely to be unemployed.
@@
 While the earnings gap diminishes when controlling for education, this doesn't necessarily indicate an absence of discrimination, as education itself can be influenced by bias.
-Moreover, educational differences cannot explain the gaps in unemployment, suggesting that discrimination may play a role in labor market outcomes of individuals with Hispanic last names.
+Moreover, educational differences cannot explain the gaps in unemployment, suggesting that discrimination may play a role in labor market outcomes of individuals with Hispanic-sounding last name.
 Further research is needed to comprehensively understand the earnings gaps between Hispanics and Whites, particularly considering that mothers of HW children have higher education and earnings, which theoretically should lead to better outcomes for their children.

### 2024-09-09 manuscripts/07-economic-letters-v5/Hadah-Hispanic-Last-Names.tex
new edits

@@
 
-Identifying discrimination is challenging due to unobservable factors affecting labor market outcomes.
-While audit studies provide insights, they primarily measure callback rates rather than full labor market outcomes \citep{bertrand2004emily}.
-This study compares children from inter-ethnic marriages: those of Hispanic fathers and White mothers (HW) to those of White fathers and Hispanic mothers (WH) \citep{rubinstein2014pride}.
-This approach leverages marriage selection based on observable characteristics \citep{averettBetterWorseRelationship2008}, allowing for the investigation of how ethnic signals, such as Hispanic last names, affect earnings.
+While these observable differences contribute to measurable disparities, identifying discrimination in the labor market presents further challenges, particularly due to unobservable factors.
+Audit studies, which offer insights into potential bias, primarily focus on callback rates rather than capturing the full scope of labor market outcomes \citep{bertrand2004emily}.
+To address these complexities, this study compares children from inter-ethnic marriages—those of Hispanic fathers and White mothers (HW) to those of White fathers and Hispanic mothers (WH).\footnote{This method was used by \citet{rubinstein2014pride} to estimate the effect of having a Sephardic last name (as opposed to an Ashkenazi one) in Israel.} This approach leverages marriage selection based on observable characteristics \citep{averettBetterWorseRelationship2008}, allowing for a nuanced investigation of how ethnic signals, such as Hispanic last names, affect earnings.
 
@@
 
-I use two datasets: the IPUMS Current Population Survey (CPS) Annual Social and Economic (ASEC) \citep{cps2019} and the 1960 to 2000 US censuses \citep{acs2019}.
+I use two datasets: the IPUMS Current Population Survey (CPS) Annual Social and Economic (ASEC) \citep{cps2019} and the 1950 to 2000 US censuses \citep{acs2019}.
 The CPS data (1994-2019) is used to study the effect of Hispanic-sounding last name on labor market outcomes.
@@
 
-The CPS sample is restricted to White, U.S.-born citizens aged 25-40, born between 1960 and 2000.
+The CPS sample is restricted to White, U.S.-born citizens aged 25-40, born between 1950 and 2000.
 Parents are classified as Hispanic if born in a Spanish-speaking country or Puerto Rico, and White if they are US-born.
@@
 Summary statistics for each group are presented in Table \ref{tab:combined-sum}.
-Using the 1960-2000 censuses, I constructed a dataset of synthetic parents, including married White men and women born between 1920-1975.
-Hispanic individuals are defined as White and born in a Spanish-speaking country, while Whites are native-born.
+Using the 1950-2000 censuses, I constructed a dataset of synthetic parents, including married White men and women born between 1920-1975.\footnote{To proxy for parental characteristics, I generate `synthetic' parents, assuming mothers were between the ages of 20 and 35 at the time of birth.
+For example, a `synthetic mother' of a 35-year-old in 2019, with a father born in the United States and a mother born in Colombia, would have the characteristics of an average married woman born in Colombia between 1949 and 1964, based on both her and her spouse's birthplaces.
+This family would appear in the 1990 Census with a 6-year-old child, with the mother born in Colombia and the father being U.S.-born.} Hispanic individuals are defined as White and born in a Spanish-speaking country, while Whites are native-born.
 Table \ref{tab:combined-mat} shows the distribution of couple types: WW (96\%), HH (2\%), WH (<1\%), and HW (1\%).
@@
 
-\section{Empirical Strategy}\label{sec:emp_model}
+\section{Empirical Approach}\label{sec:emp_model}
 
-In this section, I present my empirical strategy. The empirical strategy estimates the effect of having a Hispanic last name on educational outcomes and earnings.
-
+I estimate the effect of having a Hispanic last name on educational outcomes and earnings.
 The difference in means between Hispanics and non-Hispanic Whites could result from either discrimination or differences in innate abilities, skills, and parental investments.
-It can also be caused by differences in innate abilities, skills, and parental investments.
 While controlling for observable skill measures, I compare children of inter-ethnic marriages, HW and WH.
@@
 
-In this section I restrict the sample to WH and HW groups.
+In this section I go over my empirical approach.
+I restrict the sample to WH and HW groups.
 Let $Y_{ist}$ be the outcome of interest for person $i$ in state $s$ at time $t$.
@@
 $\beta_{1}$ is the coefficient of interest in this specification.
-$\beta_{1}$ represents the gaps in outcomes between children of inter-ethnic marriages who have a Spanish-sounding last name versus a White last name.
+$\beta_{1}$ represents the gaps in outcomes between children of inter-ethnic marriages who have a Spanish-sounding last name versus a native-sounding last name.
 
@@
 A potential measurement error arises from using parents' birthplace in the CPS data as a proxy for both ethnicity and last name.
-However, this concern is minimal as most Hispanics from 1960 to 2000 were first-generation immigrants.
-Census data shows only 3\% of native-born Americans identified as Hispanic, making it highly improbable for an inter-ethnic child to have a native-born, second-generation+ Hispanic father.\footnote{Self-reported Hispanic identity among first-generation Hispanic immigrants and native-borns are based on the author’s calculations from the 1960-2000 Census.
+However, this concern is minimal as most Hispanics from 1950 to 2000 were first-generation immigrants.
+Census data shows only 3\% of native-born Americans identified as Hispanic, making it highly improbable for an inter-ethnic child to have a native-born, second-generation+ Hispanic father.\footnote{Self-reported Hispanic identity among first-generation Hispanic immigrants and native-borns are based on the author’s calculations from the 1950-2000 Census.
 The sample includes Whites, who are married, and are between the ages 25 and 40.} Empirical evidence from the CPS and US Census will be used to evaluate these concerns and support the validity of the identification strategy.
@@
 These gaps are larger for HW women in higher education.
-These disparities suggest potential barriers or discrimination in access to higher education for HW children, particularly women.
+These disparities suggest potential barriers or discrimination in access to higher education for inter-ethnic people with Hispanic last names, particularly women.
 
@@
 This study examines discrimination against Hispanics in the U.S.
-labor market, focusing on the impact of Hispanic-sounding last name and identification on education and earnings.
+labor market, focusing on the impact of Hispanic-sounding last name on education and earnings.
 This research finds that inter-ethnic individuals with Spanish-sounding last names receive 0.2 years less education and are more likely to be unemployed.

### 2024-09-16 manuscripts/07-economic-letters-v5/Manuscript.tex
submit to economic letters

--- 2024-09-09 manuscripts/07-economic-letters-v5/Hadah-Hispanic-Last-Names.tex
@@
 This study analyzes the impact of Hispanic-sounding surnames on wages among inter-ethnic children with one White and one Hispanic parent.
-I find that individuals with Hispanic surnames receive less education, more likely to be unemployed, and often earn less, with a notable wage gap favoring those with White surnames.
+I find that individuals with Hispanic surnames receive less education, more likely to be unemployed, and often earn less, with a notable wage gap favoring those with native-sounding surnames.
 People born to Hispanic fathers and White mothers receive 0.3 years of education less than those born to White fathers and Hispanic mothers.
@@
 While a significant portion of this gap is attributed to educational differences, discrimination may also play a role, potentially leading to reduced job opportunities, lower wages, and hindered assimilation \citep{duncan2006hispanics, duncan2018identifying, duncan2018socioeconomic}.
-This paper examines the impact of Hispanic last names on labor market and educational outcomes, contributing to the literature by isolating the effect of ethnicity-signaling surnames from other potentially confounding factors.
+This paper examines the impact of Hispanic last names on labor market and educational outcomes, contributing to the literature by isolating the effect of ethnic-sounding surnames from other potentially confounding factors.
 By focusing on children of inter-ethnic marriages, this study provides a novel approach to understanding the mechanisms behind Hispanic-White earnings disparities and the potential role of name-based discrimination in the labor market.
@@
 I find that individuals with Hispanic-sounding last names receive 0.2 years less education and are more likely to be unemployed.
-The gap in education is equivalent to a 1.9\% reduction, and the unemployment gap is equal to a 14\% increase.
+The gap in education is equivalent to a 1.5\% reduction, and the unemployment gap is equal to a 14\% increase.
 The earnings gap between individuals with Hispanic-sounding last names and those with White last names is 5 percentage points.
-However, when controlling for education, the earnings gap reduces to 1 percentage point, suggesting that the earnings gap is largely due to differences in education.
+However, when controlling for education, the earnings gap decreases to 1 percentage point, suggesting that the earnings gap is largely due to differences in education.
 
@@
 The CPS sample is restricted to White, U.S.-born citizens aged 25-40, born between 1950 and 2000.
-Parents are classified as Hispanic if born in a Spanish-speaking country or Puerto Rico, and White if they are US-born.
+Parents are classified as Hispanic if born in a Spanish-speaking country or Puerto Rico, and as White if they are US-born non-Hispanics.
 Four parental types are identified: (1) White father and White mother (WW), (2) White father and Hispanic mother (WH), (3) Hispanic father and White mother (HW), and (4) Hispanic father and Hispanic mother (HH).
@@
 WH and HW children are more similar in characteristics but provide employers, and the labor market, with different signals.\footnote{WH and HW children are both half White, half Hispanic.} WH children will have a non-Hispanic last name, while HW children will have a Hispanic last name.
-This is a method developed by \citet{rubinstein2014pride}.
 
@@
 
-In this section I go over my empirical approach.
+In this section, I discuss my empirical approach.
 I restrict the sample to WH and HW groups.
@@
 $\beta_{1}$ is the coefficient of interest in this specification.
-$\beta_{1}$ represents the gaps in outcomes between children of inter-ethnic marriages who have a Spanish-sounding last name versus a native-sounding last name.
+$\beta_{1}$ represents the gaps in outcomes between children of inter-ethnic marriages who have an ethnic-sounding last name versus a native-sounding last name.
 
@@
 
-I show in Table \ref{tab:lastnamereg} a 1 percentage point employment gap between HW and WH workers, and a 5 percentage point crude earnings gap.
+I show in Table \ref{tab:lastnamereg} a 1 percentage point unemployment gap between HW and WH workers, and a 5 percentage point crude earnings gap.
 However, when controlling for education, the earnings gap reduces to a statistically insignificant 1 percentage point.
@@
 While the earnings gap diminishes when controlling for education, this doesn't necessarily indicate an absence of discrimination, as education itself can be influenced by bias.
-Moreover, educational differences cannot explain the gaps in unemployment, suggesting that discrimination may play a role in labor market outcomes of individuals with Hispanic-sounding last name.
+Moreover, educational differences cannot explain the gaps in unemployment, suggesting that discrimination may play a role in labor market outcomes of individuals with Hispanic-sounding last names.
 Further research is needed to comprehensively understand the earnings gaps between Hispanics and Whites, particularly considering that mothers of HW children have higher education and earnings, which theoretically should lead to better outcomes for their children.
@@
 %table1
-\input{tables/tab16-tab1-and-tab2}
+\input{tab16-tab1-and-tab2}
 
@@
 %table3
-\input{tables/tab17-all-sums}
+\input{tab17-all-sums}
 
@@
 
-\input{tables/tab13-regression}
+\input{tab13-regression}
 
@@
 
-\input{tables/tab15-emp-earn-regression.tex}
+\input{tab15-emp-earn-regression.tex}
 

### 2024-10-11 manuscripts/08-labor-research-v6/manuscript_main.tex
edit

--- 2024-08-28 manuscripts/06-labour-v5/manuscript_main.tex
@@
 In this paper, I answer the following questions.
+Does having a Hispanic last name affect educational outcomes?
 Does having a Hispanic last name affect labor market outcomes?
-What is the effect of identifying as Hispanic on earnings?
-Moreover, I aim to show that comparing Hispanic Whites to non-Hispanic Whites might create an artificially higher earnings gap since the two groups differ on many observable characteristics.\footnote{By identifying as Hispanic, I refer to individuals who self-report their ethnicity as Hispanic on surveys or other data collection instruments.} \footnote{Observable characteristics refer to factors that can be measured and quantified, such as education level, work experience, and immigration status.} Others have attempted to compare how native-born White Hispanics fare to non-Hispanic Whites and foreign-born Hispanics.
+I aim to show that comparing Hispanic Whites to non-Hispanic Whites might create an artificially higher earnings gap since the two groups differ on many observable characteristics.\footnote{By identifying as Hispanic, I refer to individuals who self-report their ethnicity as Hispanic on surveys or other data collection instruments.} \footnote{Observable characteristics refer to factors that can be measured and quantified, such as education level, work experience, and immigration status.} Others have attempted to compare how native-born White Hispanics fare to non-Hispanic Whites and foreign-born Hispanics.
 In \textcite{antman2020ethnic,antmanEthnicAttritionObserved2016,antmanEthnicAttritionObserved2016a,antmanEthnicAttritionAssimilation2020}, the authors compare the health and educational outcomes of Hispanic Whites to non-Hispanic Whites and native-born Hispanics to foreign-born Hispanics.
@@
 The summary statistics for the children of the four types of marriages are presented in Table \ref{tab:c&p1}.
-Children of WW marriages (column 1) do better on every measure while children of HH parents (column 4) do worse than other children on every measure.
-Children of WH (column 2) and HW (column 3) marriages fall in between WW and HH children.
+Children of WW marriages (Column 1) do better on every measure while children of HH parents (Column 4) do worse than other children on every measure.
+Children of WH (Column 2) and HW (Column 3) marriages fall in between WW and HH children.
 Moreover, the summary statistics for those that also identify as Hispanic is reported in in Table \ref{tab:c&p2}.
@@
 
-\section{Empirical Strategy}\label{sec:emp_model}
+\section{Empirical Approach}\label{sec:emp_model}
 
-In this section, I present two empirical strategies.
-The first empirical strategy estimates the effect of having a Hispanic last name on educational outcomes and earnings.
-The second empirical strategy estimates the effect of identifying as Hispanic on log annual earnings.
-
+In this section, I present my empirical Approach.
+The specification estimates the effect of having a Hispanic last name on educational outcomes and earnings.
 The difference in means between Hispanics and non-Hispanic Whites could result from discrimination.
@@
 $\beta_{1}$ represents the gaps in outcomes between children of inter-ethnic marriages who have a Spanish-sounding last name versus a White last name.
-
-\subsection{Estimating the effect of identifying as Hispanic}
-
-In this section, I will present an estimation strategy that would allow me to capture the effect of identifying as a Hispanic.
-
-\begin{equation} \label{eq:iden}
-Y_{ist} = \beta_{1} HW_{ist} + \beta_{2} Hispanic_i + \beta_{3} HW_{ist} \cdot Hispanic_i + X_{ist} \pi + \gamma_{st} + \phi_{ist}
-\end{equation}
-
-The coefficients of interest from equation \ref{eq:iden} are $\beta_{1}$, $\beta_{2}$ and $\beta_{3}$.
-The coefficients $\beta_{1}$ and $\beta_{3}$ capture the effects of Spanish-sounding last names for $HW_{ist}$ that identify as Hispanic \textit{versus} those who do not.
 
@@
 
-In this section, I present the results from estimating the two specifications presented in equations \ref{eq:1a} and \ref{eq:iden}.
-I estimate the mean educational outcomes of White, Native-born, Hispanics aged 25-40.
-I estimate the mean wages of White, Native-born, Hispanic men aged 25-40 who are employed full-time and full-year (FTFY) as waged and salaried workers.
-The results are shown in Tables \ref{tab:lastnamereg} and \ref{tab:identreg}.
+In this section, I present the results from estimating the specification presented in equation \ref{eq:1a}.
+I estimate the mean educational outcomes of White, Native-born (results in Table \ref{tab:lastname-ed-reg}), Hispanics aged 25-40, and mean unemployment of White, Native-born, Hispanics men aged 25-40 (results in Table \ref{tab:lastnamereg-emp}).
+I estimate the mean wages of White, Native-born, Hispanic men aged 25-40 who are employed full-time and full-year (FTFY) as waged and salaried workers (results in Table \ref{tab:lastnamereg}).
 
@@
 
-Second, I find that an inter-ethnic person with a Spanish-sounding last names earns less than an inter-ethnic who has a White last name.
+Second, I find that an inter-ethnic person with a Spanish-sounding last names are 1 percentage point more likely to be unemployed and earn less than an inter-ethnic who has a White last name.
 A person with a Spanish-sounding last name earns 5 percentage points less than a person with a White last name.
@@
 The omitted group is children of WH parents.
-Column one in Table \ref{tab:lastname-ed-reg} is the difference in total years of education between HW children and their WH peers.
-Column two is the difference in the probability of having been a high school dropout.
-Column three is the difference in the probability of having an associate degree.
-Column four is the difference in the probability of having a bachelor's degree.
+Column 1 in Table \ref{tab:lastname-ed-reg} is the difference in total years of education between HW children and their WH peers.
+Column 2 is the difference in the probability of having been a high school dropout.
+Column 3 is the difference in the probability of having an associate degree.
+Column 4 is the difference in the probability of having a bachelor's degree.
 All regressions include controls for age, parental education and income, and state-year fixed effects.
@@
 
-\subsection{The Effect of Having a Hispanic Last Name and Hispanic Identity on Labor Market Outcomes}
+\subsection{The Effect of Having a Hispanic Last Name on Labor Market Outcomes}
 
-I provide the results to the estimation of equation \ref{eq:1a} in Table \ref{tab:lastnamereg}.
+I provide the results to the estimation of equation \ref{eq:1a} in Tables \ref{tab:lastnamereg-emp} and \ref{tab:lastnamereg} on unemployment and log earnings.
+I estimate the mean unemployment of White, Native-born, Hispanics men aged 25-40.
 I estimate the mean wages of White, U.S.-born, Hispanic men aged 25-40 who are employed FTFY as waged and salaried workers.
 I also restrict the sample to children of HW and WH (omitted) parents.
-Column one in Table \ref{tab:lastnamereg} is the average crude earnings gap in log annual earnings between HW workers and their WH peers.
+Column 1 in Table \ref{tab:lastnamereg} is the average crude earnings gap in log annual earnings between HW workers and their WH peers.
 In the next 4 columns, I introduce the results with controls for hours worked, state FE, year FE, age FE, Education FE, and parental background.
+
+In addition to examining earnings, I also analyzed the effect of having a Hispanic last name on unemployment rates.
+Table \ref{tab:lastnamereg-emp} presents the results of this analysis.
+Column 1 shows that individuals with Hispanic last names (HW) have a 1 percentage point higher unemployment rate compared to those with White last names.
+This gap persists even after controlling for age, state fixed effects, year fixed effects, and state-year fixed effects (Column 2), though the significance level drops.
+When education is included as a control (Column 3), the gap remains at 1 percentage point.
+Finally, after controlling for parental background (Column 4), the 1 percentage point difference in unemployment rates becomes statistically insignificant.
+These results suggest that while there is an initial unemployment gap associated with having a Hispanic last name, much of this difference can be explained by factors such as education and parental background.
+The mean unemployment rate for individuals with Hispanic last names (HW) is 7\% across all specifications.
 
@@
 The omitted group is children of WH parents that do not identify as Hispanic.
-Column one in Table \ref{tab:identreg} is the results without controlling for education.
+Column 1 in Table \ref{tab:identreg} is the results without controlling for education.
 An inter-ethnic $HW_{ist}$ that identifies as Hispanic \textit{versus} $HW_{ist}$ who do not---i.e.

### 2024-10-11 manuscripts/08-labor-research-v6/manuscript_title_etc.tex
edit

--- 2024-08-28 manuscripts/07-economic-letters-v5/manuscript_title_etc.tex
@@
 % \newcommand{\PAPERTITLE}{\href{https://hhadah.github.io/hispanic-last-names/my_paper/Hadah-last-names-draft.pdf}{The Impact of Hispanic Last Names and Identity on Labor Market Outcomes}}
-\newcommand{\PAPERTITLE}{The Impact of Hispanic Last Names and Identity on Educational and Labor Market Outcomes}
+\newcommand{\PAPERTITLE}{The Impact of Hispanic Last Names on Educational and Labor Market Outcomes}
 
@@
 People born to Hispanic fathers and White mothers receive 0.3 years of education less than those born to White fathers and Hispanic mothers.
-Males born to Hispanic fathers and White mothers earn 5 percentage points less than those born to White fathers and Hispanic mothers, largely due to educational differences.
-Additionally, men with Spanish-sounding last names who identify as Hispanic earn significantly less, again due to educational disparities.
+Males born to Hispanic fathers and White mothers are 1 percentage point more likely to be unemployed, and they also earn 5 percentage points less than those born to White fathers and Hispanic mothers.
+The earnings gap is largely due to educational differences.
 \PAPERJEL}

### 2024-10-15 manuscripts/06-labour-v5/tab04-summary-stats.tex
edits to manuscript

--- 2024-08-11 manuscripts/labour-v5/tab04-summary-stats.tex
@@
 \begin{longtable}[t]{>{\raggedright\arraybackslash}p{5cm}cccccc}
-\caption{Summary Statistics of Outcomes Using Parent's Place of Birth Only for Those That Self-Identify as Hispanic \label{tab:c&p2}}\\
+\caption{Summary Statistics of Outcomes Using Parent's Place of Birth\label{tab:c&p2}}\\
 \toprule
@@
 \endfirsthead
-\caption[]{Summary Statistics of Outcomes Using Parent's Place of Birth Only for Those That Self-Identify as Hispanic  \textit{(continued)}}\\
+\caption[]{Summary Statistics of Outcomes Using Parent's Place of Birth  \textit{(continued)}}\\
 \toprule

### 2024-10-15 manuscripts/08-labor-research-v6/manuscript_main.tex
edits to manuscript

@@
 
-A large literature documents substantial earnings gaps across race and ethnicity \autocite{bayer2018divergent, charles2008prejudice, card1992school, fryer2004causes, rubinstein2014pride, bertrand2004emily, juhn1991accounting}.
+A large literature documents substantial earnings gaps across race and ethnicity \autocite{bayer2018divergent, charles2008prejudice}.
 Hispanics constitute a large and growing portion of the population in the United States.
-As the number of Hispanics increases, determining whether ethnic discrimination affects their labor market outcomes becomes increasingly crucial \autocite{chettyUnitedStatesStill2014, chettyEffectsExposureBetter2016,chettyFadingAmericanDream2017,abramitzkyImmigrantsAssimilateMore2020a, abramitzkyNationImmigrantsAssimilation2014,abramitzkyCulturalAssimilationAge2016,chettyWhereLandOpportunity2014}.
-Thus, it is important to understand whether a person's ethnicity affects their labor market outcomes.
-Assimilation and mobility are crucial because they reflect how well Hispanics can integrate into society and move up the socioeconomic ladder.
+As the number of Hispanics increases, determining whether ethnic discrimination affects their labor market outcomes becomes increasingly crucial \autocite{chettyUnitedStatesStill2014, chettyEffectsExposureBetter2016,chettyFadingAmericanDream2017}.
+Specifically, it is essential to investigate the extent to which a person's Hispanic ethnicity may influence their employment prospects, wages, and career advancement.
+Understanding these labor market outcomes is critical, as they directly impact broader societal issues such as assimilation and economic mobility.
+These factors serve as key indicators of how successfully Hispanics can integrate into society and ascend the socioeconomic ladder.
 
@@
 
-The US population is growing in diversity.
-The proportion of non-Whites has increased by more than 10 percentage points from 13 percent in 1995 to 23 percent in 2019.
-The number of Hispanics has grown by 9 percentage points from 9 percent in 1995 to 16 percent in 2019.\footnote{The portion of non-Whites and Hispanics is calculated using the Current Population Survey (CPS).} Native-born White Hispanic men earn 21\% less than White men, although a substantial portion of the earnings gap is due to educational differences between Hispanics and Whites \autocite{duncan2006hispanics, duncan2018identifying, duncan2018socioeconomic}.
-Some of the earnings differences may also be due to discrimination which will have negative consequences.
+This study builds upon and improves the traditional Oaxaca-Blinder-Kitagawa decomposition approach often used in labor market discrimination studies \autocite{oaxaca1973male,blinder1973wage,kitagawa1955components}.
+While these decomposition methods have provided valuable insights into earnings gaps between different groups, including Hispanics and non-Hispanic Whites \autocite{davilaChangesRelativeEarnings2008}, they are limited in their ability to account for unobserved differences between the groups.
+My approach, by comparing children of inter-ethnic marriages, provides a more closely matched comparison group, allowing me to better isolate the effect of having a Hispanic last name on labor market outcomes.
+
+This methodological improvement is particularly timely and relevant given the rapidly changing demographics of the United States.
+The US population is growing increasingly diverse, with significant implications for labor market dynamics and potential discrimination.
+The proportion of non-Whites has increased by more than 10 percentage points from 13 percent in 1995 to 23 percent in 2019.\footnote{The proportion of non-Whites and Hispanics is calculated using the Current Population Survey (CPS).} Native-born White Hispanic men earn 21\% less than White men, although a substantial portion of the earnings gap is due to educational differences between Hispanics and Whites \autocite{duncan2006hispanics, duncan2018identifying, duncan2018socioeconomic}.
+Some earnings differences may also be due to discrimination which will have negative consequences.
 For example, discrimination against Hispanics can lead to reduced job opportunities, lower wages, and hinder assimilation.
 In this paper, I examine the role of having a Hispanic last name and identifying as Hispanic on labor market outcomes.
+
+Using this novel approach, I find significant effects of having Hispanic last names on both educational and labor market outcomes.
+The results show that individuals with Hispanic last names face substantial disadvantages in educational attainment and earnings compared to their counterparts with non-Hispanic names.
+Specifically, individuals with Hispanic last names complete 0.2 years of education less than those with non-Hispanic names, even when controlling for family background.
+Individuals with Hispanic last names are also 1 percentage point more likely to be unemployed  and earn 5 percentage points less than those with non-Hispanic names.
+These findings suggest that ethnic discrimination continues to play a role in shaping economic opportunities for Hispanics in the United States, highlighting the need for targeted policies to address these persistent disparities.
 
@@
 Couples match on several observable characteristics like income, schooling, socio-economic background, etc.
-\autocite{averettBetterWorseRelationship2008, averettEconomicRealityBeauty1996, beckerTheoryMarriagePart1973, beckerTheoryMarriagePart1974, beckerTreatiseFamily1993, browningCollectiveUnitaryModels2006, chiapporiFatterAttractionAnthropometric2012}.
-Children of HW and WH marriages have more similar observable characteristics than children of endogamous/homogamous marriages--- i.e., White fathers-White mothers and Hispanic fathers-Hispanic mothers.
+\autocite{averettBetterWorseRelationship2008, averettEconomicRealityBeauty1996}.\footnote{For more on assortative mating see \autocite{beckerTheoryMarriagePart1973, beckerTheoryMarriagePart1974, beckerTreatiseFamily1993, browningCollectiveUnitaryModels2006, chiapporiFatterAttractionAnthropometric2012}} Children of HW and WH marriages have more similar observable characteristics than children of endogamous/homogamous marriages--- i.e., White fathers-White mothers and Hispanic fathers-Hispanic mothers.
 Moreover, children from a Hispanic father and White mother household will have a Hispanic last name from their fathers, enabling the investigation of how ethnic signals, such as having a Hispanic last name, affect annual log earnings.
@@
 Children from mixed ethnic backgrounds may appear physically similar to those from single ethnicity backgrounds, but the influence of family dynamics and upbringing, crucial in developing skills and personal characteristics, varies with the pattern of mixed ethnic marriages.
-Particularly in a society where Hispanic ancestry is perceived negatively, it raises the question of what kind of White women would choose Hispanic men.
+Particularly in a society where Hispanic ancestry could be perceived negatively, it raises the question of what kind of White women would choose Hispanic men.
 Furthermore, even if such unions were formed randomly, children from White-Hispanic homes might benefit from more favorable family conditions than those from Hispanic-White homes, considering Whites generally have stronger socio-economic backgrounds than Hispanics.
@@
 
-The distribution of the four types of children is presented Table \ref{tab:mat1}.
+The distribution of the four types of children is presented in Table \ref{tab:mat1}.
 The majority of the sample (96\%) is WW children.
@@
 Even though WH and HW are only 1.35\% of the sample, I have plenty of observations to carry out an analysis.
-The summary statistics for the children of the four types of marriages are presented in Table \ref{tab:c&p1}.
+The summary statistics for the children of the four types of marriages are presented in Table \ref{tab:c&p2}.
 Children of WW marriages (Column 1) do better on every measure while children of HH parents (Column 4) do worse than other children on every measure.
 Children of WH (Column 2) and HW (Column 3) marriages fall in between WW and HH children.
-Moreover, the summary statistics for those that also identify as Hispanic is reported in in Table \ref{tab:c&p2}.
 
@@
 
-\subsection{Estimating the effect of having a Hispanic last name}
+This approach allows me to control for several key factors that could influence labor market outcomes.
+Importantly, I control for parental education and income using my synthetic parents data, which addresses concerns about intergenerational transmission of human capital.
+By comparing children from inter-ethnic marriages, I also implicitly control for many unobservable family characteristics that might affect labor market outcomes, as these families are likely to be more similar to each other than to families with two Hispanic or two non-Hispanic White parents.
+
+\subsection{Estimating the Effect of Having a Hispanic Last Name}
 
@@
 $HW_{ist}$ is an indicator variables for the type of parents of person $i$ has.
-$X_{ist}$ is a vector of controls that includes age and numbers of hours worked, $\gamma_{st}$ are state-year fixed effects (FE), and $\phi_{ist}$ represents the error term.
-The equation for this strategy is written as follows:
+$X_{ist}$ is a vector of controls that includes age and numbers of hours worked, $\gamma_{st}$ are state-year fixed effects (FE), and $\phi_{ist}$ represents the error term.\footnote{I also include parental characteristics of `synthetic' parents as controls in alternative specifications.
+These controls include income and education of `synthetic' parents.} The equation for this strategy is written as follows:
 \begin{equation} \label{eq:1a}
@@
 $\beta_{1}$ represents the gaps in outcomes between children of inter-ethnic marriages who have a Spanish-sounding last name versus a White last name.
+If $\beta_{1} > 0$, then people with a Hispanic last name have better outcomes than people with a White last name.
+If $\beta_{1} < 0$, then people with a Hispanic last name have worse outcomes than people with a White last name.
 
@@
 Consequently, if marriages were random, the White husband in an WH marriage would have better traits than wife, and the White wife in the HW marriage would have the stronger traits.
-The random matching coupled with the fact that parents are not perfect substitutes in the inter-generational transmission of human capital would render the identification invalid since the parental background will affect labor market outcomes.
+Random matching in marriages would pose a challenge to the identification strategy.
+This is because parents are not perfect substitutes in the inter-generational transmission of human capital.
+Consequently, differences in parental background could significantly affect children's labor market outcomes, potentially invalidating the comparison between HW and WH children.
 
@@
 Despite higher levels of education and income among Hispanic White (HW) mothers compared to White Hispanic (WH) mothers, suggesting an expectation of greater educational attainment for HW children, data reveals a discrepancy.
-Specifically, HW children complete, on average, 0.4 fewer years of education compared to their WH peers (refer to Table \ref{tab:c&p1}).
+Specifically, HW children complete, on average, 0.4 fewer years of education compared to their WH peers (refer to Table \ref{tab:c&p2}).
 This gap may suggest potential discrimination or barriers in educational access for HW children.
@@
 
-First, I find that an inter-ethnic person with a Spanish-sounding last names receive less total years of education than an inter-ethnic who has a White last name.
+First, I find that an inter-ethnic person with a Spanish-sounding last name receives fewer total years of education than an inter-ethnic person who has a White last name.
 A person with a Spanish-sounding last name receives 0.2 years of education less than a person with a White last name, have the same high school dropout rate as a person with a White last name, and 2 percentage points less likely to have an associate degree.
@@
 To increase the sample size of my analysis I estimate equation \ref{eq:1a} using weekly earnings as a dependent variable.
-Weekly earnings are available in all of the monthly CPS surveys and not just the March Supplement.
+Weekly earnings are available in all the monthly CPS surveys and not just the March Supplement.
 I present the results to these estimations in tables \ref{tab:lastnamereg-weekearm}-\ref{tab:lastnamereg-weekearm-cub}.
 
-First, in Table \ref{tab:lastnamereg-weekearm}, I present the results to the estimation of equation \ref{eq:1a} using weekly earnings as a dependent variable on the full sample.
-Similar to the analysis before, I find that a person with a Hispanic last name earns 4 p.p.
-less that a person without a Hispanic last name.
-This gap could be explained by Educational differences.
-Second, in Table \ref{tab:lastnamereg-weekearm-mex}, I present the results to the estimation of equation \ref{eq:1a} using weekly earnings as a dependent variable on a sample of Mexican Hispanics.
-I find that a Mexican with a Hispanic sounding last name earns 3 p.p.
-less than a Mexican with a native sounding last name.
-The gap becomes an imprecise zero after controlling for education and parental background.
-Third, in Table \ref{tab:lastnamereg-weekearm-nonmex}, I present the results to the estimation of equation \ref{eq:1a} using weekly earnings as a dependent variable on a sample of non-Mexican Hispanics.
-The gap between non-Mexicans with Hispanic sounding last names and those with native sounding last names was similarly explained by educational differences.
-I find that a non-Mexican with a Hispanic sounding last name earns 3 p.p.
-less than a Mexican with a native sounding last name.
-The gap becomes an imprecise zero after controlling for education and parental background.
-Finally, in Table \ref{tab:lastnamereg-weekearm-cub}, I present the results to the estimation of equation \ref{eq:1a} using weekly earnings as a dependent variable on a sample of Cubans.
-I find that a Cuban with Hispanic sounding last name earns 4 p.p.
-less than a Cuban with a native sounding last name but the gap was statistically insignificant.
-
-I provide the results to the estimation of equation \ref{eq:iden} in Table \ref{tab:identreg}.
-The sample in this analysis includes full-time, full-year, and wage and salary men, and I control for hours worked and age and include year FE.
-I also restrict the sample to children of HW and WH parents.
-The omitted group is children of WH parents that do not identify as Hispanic.
-Column 1 in Table \ref{tab:identreg} is the results without controlling for education.
-An inter-ethnic $HW_{ist}$ that identifies as Hispanic \textit{versus} $HW_{ist}$ who do not---i.e.
-$HW_{ist} + HW_{ist} \cdot Hispanic_i$---earns 6 percentage points less than an inter-ethnic that does not, but the result could be entirely explained by educational differences.
+First, in Table \ref{tab:lastnamereg-weekearm}, I present the results of estimating equation \ref{eq:1a} using weekly earnings as the dependent variable for the full sample.
+Similar to the previous analysis, I find that a person with a Hispanic last name earns 4 percentage points less than a person without a Hispanic last name.
+This gap could be explained by educational differences.
+Second, in Table \ref{tab:lastnamereg-weekearm-mex}, I present the results for a sample of Mexican Hispanics.
+I find that a Mexican with a Hispanic-sounding last name earns 3 percentage points less than a Mexican with a native-sounding last name.
+This gap becomes an imprecise zero after controlling for education and parental background.
+Third, in Table \ref{tab:lastnamereg-weekearm-nonmex}, I present the results for a sample of non-Mexican Hispanics.
+The gap between non-Mexicans with Hispanic-sounding last names and those with native-sounding last names is similarly explained by educational differences.
+I find that a non-Mexican with a Hispanic-sounding last name earns 3 percentage points less than a non-Mexican with a native-sounding last name.
+This gap also becomes an imprecise zero after controlling for education and parental background.
+Finally, in Table \ref{tab:lastnamereg-weekearm-cub}, I present the results for a sample of Cubans.
+I find that a Cuban with a Hispanic-sounding last name earns 4 percentage points less than a Cuban with a native-sounding last name, but this gap is statistically insignificant.
 
@@
 When I control for education, the last name effect decreases to a statistically insignificant 2 percentage points earnings gap.
-Moreover, I estimate the effect of identifying as Hispanic and having a Hispanic last name.
-I find that an inter-ethnic child with a Hispanic last name and identifies as Hispanic earns 5 percentage points less that an inter-ethnic child with native sounding last name and identifies as Hispanic.
 

### 2024-10-15 manuscripts/08-labor-research-v6/manuscript_title_etc.tex
edits to manuscript

@@
 People born to Hispanic fathers and White mothers receive 0.3 years of education less than those born to White fathers and Hispanic mothers.
-Males born to Hispanic fathers and White mothers are 1 percentage point more likely to be unemployed, and they also earn 5 percentage points less than those born to White fathers and Hispanic mothers.
+Men born to Hispanic fathers and White mothers are 1 percentage point more likely to be unemployed, and earn 5 percentage points less than those born to White fathers and Hispanic mothers.
 The earnings gap is largely due to educational differences.

### 2024-10-15 my_paper/tables/tab04-summary-stats.tex
edits to manuscript

--- 2023-05-09 my_paper/tables/tab04-summary-stats.tex
@@
 \begin{TableNotes}
-\item[1] Source: The 1950-2000 Census for synthetic parents, and 1994-2019 Current Population Surveys (CPS) for children's outcomes
+\item[1] Source: The 1994-2019 Current Population Surveys (CPS) for children's outcomes
 \item[2] The data is restricted to native-born United States citizens between 1994 and 2019 who are also White and between the ages of 25 and 40.
@@
 \begin{longtable}[t]{>{\raggedright\arraybackslash}p{5cm}cccccc}
-\caption{Summary statistics of outcomes using parent's place of birth only for those that self-identify as Hispanic \label{tab:c&p2}}\\
+\caption{Summary Statistics of Outcomes Using Parent's Place of Birth \label{tab:c&p2}}\\
 \toprule
@@
 \endfirsthead
-\caption[]{Summary statistics of outcomes using parent's place of birth only for those that self-identify as Hispanic  \textit{(continued)}}\\
+\caption[]{Summary Statistics of Outcomes Using Parent's Place of Birth  \textit{(continued)}}\\
 \toprule
@@
 \textbf{Panel A: Children's Education} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
-\hspace{1em}Men’s education (Total Years) & \specialcell{13.82\\(2.42)} & \specialcell{13.4\\(2.38)} & \specialcell{13.08\\(2.27)} & \specialcell{12.87\\(2.3)} & \specialcell{-0.94***\\(0.01)} & \specialcell{-0.32**\\(0.03)}\\
-\hspace{1em}Women’s education (Total Years) & \specialcell{14.06\\(2.37)} & \specialcell{13.62\\(2.39)} & \specialcell{13.26\\(2.35)} & \specialcell{13.22\\(2.39)} & \specialcell{-0.84***\\(0.01)} & \specialcell{-0.36**\\(0.02)}\\
+\hspace{1em}Men’s education (Total Years) & \specialcell{13.82\\(2.42)} & \specialcell{13.57\\(2.4)} & \specialcell{13.22\\(2.34)} & \specialcell{12.9\\(2.31)} & \specialcell{-0.92***\\(0.01)} & \specialcell{-0.36**\\(0.02)}\\
+\hspace{1em}Women’s education (Total Years) & \specialcell{14.06\\(2.37)} & \specialcell{13.79\\(2.44)} & \specialcell{13.42\\(2.38)} & \specialcell{13.24\\(2.39)} & \specialcell{-0.82***\\(0.01)} & \specialcell{-0.37**\\(0.02)}\\
+\hspace{1em}Men’s High School Dropout Rate & \specialcell{0.35\\(0.48)} & \specialcell{0.42\\(0.49)} & \specialcell{0.43\\(0.49)} & \specialcell{0.44\\(0.5)} & \specialcell{0.09***\\(0)} & \specialcell{0***\\(0)}\\
+\hspace{1em}Women’s High School Dropout Rate & \specialcell{0.36\\(0.48)} & \specialcell{0.43\\(0.49)} & \specialcell{0.45\\(0.5)} & \specialcell{0.46\\(0.5)} & \specialcell{0.09***\\(0)} & \specialcell{0.02***\\(0)}\\
+\addlinespace
+\hspace{1em}Men’s Associate Degree Rate & \specialcell{0.15\\(0.36)} & \specialcell{0.15\\(0.36)} & \specialcell{0.12\\(0.33)} & \specialcell{0.12\\(0.32)} & \specialcell{-0.03***\\(0)} & \specialcell{-0.03***\\(0)}\\
+\hspace{1em}Women’s Associate Degree Rate & \specialcell{0.19\\(0.39)} & \specialcell{0.17\\(0.38)} & \specialcell{0.14\\(0.35)} & \specialcell{0.15\\(0.36)} & \specialcell{-0.04***\\(0)} & \specialcell{-0.03***\\(0)}\\
+\hspace{1em}Men’s BA Degree Rate & \specialcell{0.32\\(0.47)} & \specialcell{0.26\\(0.44)} & \specialcell{0.22\\(0.41)} & \specialcell{0.17\\(0.38)} & \specialcell{-0.15***\\(0)} & \specialcell{-0.05***\\(0)}\\
+\hspace{1em}Women’s BA Degree Rate & \specialcell{0.36\\(0.48)} & \specialcell{0.31\\(0.46)} & \specialcell{0.25\\(0.43)} & \specialcell{0.22\\(0.42)} & \specialcell{-0.14***\\(0)} & \specialcell{-0.07***\\(0)}\\
+\hspace{1em}Men’s Professional Degree Rate & \specialcell{0.09\\(0.28)} & \specialcell{0.07\\(0.26)} & \specialcell{0.06\\(0.23)} & \specialcell{0.04\\(0.19)} & \specialcell{-0.05***\\(0)} & \specialcell{-0.02***\\(0)}\\
+\addlinespace
+\hspace{1em}Women’s Professional Degree Rate & \specialcell{0.1\\(0.3)} & \specialcell{0.09\\(0.28)} & \specialcell{0.07\\(0.26)} & \specialcell{0.06\\(0.24)} & \specialcell{-0.04***\\(0)} & \specialcell{-0.01***\\(0)}\\
 \textbf{Panel B: Children's Employment and Earnings} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
-\hspace{1em}Men’s Unemployment Rate & \specialcell{0.04\\(0.8)} & \specialcell{0.06\\(0.76)} & \specialcell{0.07\\(0.74)} & \specialcell{0.07\\(0.75)} & \specialcell{0.02***\\(0.00)} & \specialcell{0.01***\\(0.00)}\\
+\hspace{1em}Men’s Unemployment Rate & \specialcell{0.04\\(0.8)} & \specialcell{0.05\\(0.77)} & \specialcell{0.07\\(0.75)} & \specialcell{0.07\\(0.75)} & \specialcell{0.02***\\(0.00)} & \specialcell{0.01***\\(0.00)}\\
+\hspace{1em}Women’s Unemployment Rate & \specialcell{0.04\\(0.81)} & \specialcell{0.05\\(0.22)} & \specialcell{0.06\\(0.76)} & \specialcell{0.06\\(0.76)} & \specialcell{0.02***\\(0.00)} & \specialcell{0.01***\\(0.00)}\\
+\hspace{1em}Men’s Log Hourly Earnings & \specialcell{2.51\\(0.45)} & \specialcell{2.44\\(0.47)} & \specialcell{2.43\\(0.45)} & \specialcell{2.42\\(0.43)} & \specialcell{-0.09***\\(0)} & \specialcell{-0.01**\\(0.01)}\\
 \addlinespace
-\hspace{1em}Women’s Unemployment Rate & \specialcell{0.04\\(0.81)} & \specialcell{0.06\\(0.24)} & \specialcell{0.07\\(0.75)} & \specialcell{0.06\\(0.76)} & \specialcell{0.02***\\(0.00)} & \specialcell{0.01***\\(0.00)}\\
-\hspace{1em}Men’s Log Hourly Earnings & \specialcell{2.51\\(0.45)} & \specialcell{2.43\\(0.46)} & \specialcell{2.42\\(0.44)} & \specialcell{2.42\\(0.43)} & \specialcell{-0.09***\\(0)} & \specialcell{-0.00**\\(0.02)}\\
-\hspace{1em}Women’s Log Hourly Earnings & \specialcell{2.32\\(0.49)} & \specialcell{2.31\\(0.45)} & \specialcell{2.28\\(0.45)} & \specialcell{2.31\\(0.42)} & \specialcell{-0.02***\\(0)} & \specialcell{-0.03**\\(0.02)}\\
-\hspace{1em}Men’s Log Annual Earnings & \specialcell{10.29\\(1.01)} & \specialcell{10.09\\(1.05)} & \specialcell{10.06\\(0.97)} & \specialcell{10.01\\(1.04)} & \specialcell{-0.28***\\(0.01)} & \specialcell{-0.03**\\(0.03)}\\
-\hspace{1em}Women’s Log Annual Earnings & \specialcell{9.43\\(1.77)} & \specialcell{9.52\\(1.59)} & \specialcell{9.47\\(1.56)} & \specialcell{9.53\\(1.49)} & \specialcell{0.1**\\(0.02)} & \specialcell{-0.05*\\(0.05)}\\*
+\hspace{1em}Women’s Log Hourly Earnings & \specialcell{2.32\\(0.49)} & \specialcell{2.32\\(0.46)} & \specialcell{2.28\\(0.46)} & \specialcell{2.31\\(0.42)} & \specialcell{-0.02***\\(0)} & \specialcell{-0.03**\\(0.01)}\\
+\hspace{1em}Men’s Log Annual Earnings & \specialcell{10.29\\(1.01)} & \specialcell{10.12\\(1.05)} & \specialcell{10.08\\(1.01)} & \specialcell{10.01\\(1.04)} & \specialcell{-0.28***\\(0.01)} & \specialcell{-0.04**\\(0.03)}\\
+\hspace{1em}Women’s Log Annual Earnings & \specialcell{10.13\\(0.78)} & \specialcell{10.07\\(0.78)} & \specialcell{10.04\\(0.73)} & \specialcell{10.01\\(0.73)} & \specialcell{-0.12***\\(0.01)} & \specialcell{-0.03**\\(0.02)}\\
+\textbf{Panel C: Children's Hispanic Identity} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
+\hspace{1em}Men & \specialcell{0.04} & \specialcell{0.74} & \specialcell{0.83} & \specialcell{0.96} &  & \\
+\addlinespace
+\hspace{1em}Women & \specialcell{0.05} & \specialcell{0.78} & \specialcell{0.81} & \specialcell{0.97} &  & \\*
 \end{longtable}

### 2024-10-15 programs/95-make-all.R

@@ -20,7 +20,7 @@ thesis_tabs  <- paste0(mdir_git,"/my_paper/tables")
 thesis_plots <- paste0(mdir_git,"/my_paper/figure") 
 add_health_data_dir <- "/Users/hhadah/Dropbox/Research/My Research Data and Ideas/Identification_Paper/Data/Datasets"
 ACS_data  <- "/Users/hhadah/Documents/GiT/jmp-decomposition-replication/data/raw"
-manuscript_wd <- paste0(mdir_git,"/manuscripts/07-economic-letters-v5") 
+manuscript_wd <- paste0(mdir_git,"/manuscripts/08-labor-research-v6") 
 # pres_tabs <- paste0(mdir,"/presentations/jmp-presentation/tables")
 # pres_plots <- paste0(mdir,"/presentations/jmp-presentation/figures")
 

### 2024-10-15 manuscripts/08-JERP-v6/manuscript_main.tex
edits for JERP

--- 2024-10-15 manuscripts/08-labor-research-v6/manuscript_main.tex
@@
 Does having a Hispanic last name affect labor market outcomes?
-I aim to show that comparing Hispanic Whites to non-Hispanic Whites might create an artificially higher earnings gap since the two groups differ on many observable characteristics.\footnote{By identifying as Hispanic, I refer to individuals who self-report their ethnicity as Hispanic on surveys or other data collection instruments.} \footnote{Observable characteristics refer to factors that can be measured and quantified, such as education level, work experience, and immigration status.} Others have attempted to compare how native-born White Hispanics fare to non-Hispanic Whites and foreign-born Hispanics.
+I aim to show that comparing Hispanic Whites to non-Hispanic Whites might create an artificially higher earnings gap since the two groups differ in many observable characteristics.\footnote{By identifying as Hispanic, I refer to individuals who self-report their ethnicity as Hispanic on surveys or other data collection instruments.} \footnote{Observable characteristics refer to factors that can be measured and quantified, such as education level, work experience, and immigration status.} Others have attempted to compare how native-born White Hispanics fare to non-Hispanic Whites and foreign-born Hispanics.
 In \textcite{antman2020ethnic,antmanEthnicAttritionObserved2016,antmanEthnicAttritionObserved2016a,antmanEthnicAttritionAssimilation2020}, the authors compare the health and educational outcomes of Hispanic Whites to non-Hispanic Whites and native-born Hispanics to foreign-born Hispanics.
@@
 
+Moreover, by comparing interethnic children, I can control for many unobservable family characteristics that might affect labor market outcomes.
+These families are likely to be more similar to each other than to families with two Hispanic or two non-Hispanic White parents.
+This approach allows me to isolate the effect of having a Hispanic last name on labor market outcomes, providing a more accurate estimate of the impact.
+This is an improvement over traditional Oaxaca-Blinder-Kitagawa decomposition methods, which are limited in their ability to account for unobserved differences between groups by providing a more suitable comparison group to measure discrimination \autocite{oaxaca1973male,blinder1973wage,kitagawa1955components}.
+
 \subsection{Threats to Identification}
@@
 Such comparisons shed light on the nuanced dynamics of ethnicity, education, and economic outcomes in inter-ethnic marriages.
-\footnote{I present in Tables \ref{tab:synthmex} and \ref{tab:syntnonmex} the summary statistics which detail the educational and economic profiles of parents from four different ethnic groupings—White White (WW), White Hispanic (WH), Hispanic White (HW), and Hispanic Hispanic (HH) on sub-samples of Hispanics of Mexican and non-Mexican ancestries.
+
+% \footnote{I present in Tables \ref{tab:synthmex} and \ref{tab:syntnonmex} the summary statistics which detail the educational and economic profiles of parents from four different ethnic groupings—White White (WW), White Hispanic (WH), Hispanic White (HW), and Hispanic Hispanic (HH) on sub-samples of Hispanics of Mexican and non-Mexican ancestries.
 I find similar results that describe a selection into inter-ethnic marriages among the two groups.}
@@
 
-Since Hispanics are a very heterogeneous group, I conduct a heterogeneity analysis on different samples of Hispanics.
+% Since Hispanics are a very heterogeneous group, I conduct a heterogeneity analysis on different samples of Hispanics.
 To increase the sample size of my analysis I estimate equation \ref{eq:1a} using weekly earnings as a dependent variable.
@@
 
-First, in Table \ref{tab:lastnamereg-weekearm}, I present the results of estimating equation \ref{eq:1a} using weekly earnings as the dependent variable for the full sample.
+% First, in Table \ref{tab:lastnamereg-weekearm}, I present the results of estimating equation \ref{eq:1a} using weekly earnings as the dependent variable for the full sample.
 Similar to the previous analysis, I find that a person with a Hispanic last name earns 4 percentage points less than a person without a Hispanic last name.
@@
 I find that inter-ethnic people with Spanish-sounding last names receive 0.2 years of education less than inter-ethnic counterparts that have a native-sounding last name.
-When I compare the earnings of HW and WH children, which captures the effect of having a Hispanic last name, HW children earn  6 percentage points less than WH children.
+When I compare the earnings of HW and WH children, which captures the effect of having a Hispanic last name, HW children are 1 percentage point more likely to be unemployed than WH children.
+HW children earn 5 percentage points less than WH children.
 Thus, by comparing inter-ethnic children, a person with a Hispanic last name makes 5 percentage points less than someone with a White last name.

### 2024-10-15 manuscripts/08-JERP-v6/submit_journal-main.tex
edits for JERP

--- 2024-05-20 manuscripts/latest-revisions/submit_journal-title-abstract.tex
@@
 % Part I.a.1 Title
-\section*{Title and Author}
-\subsection*{Title}
+\section*{TITLE AND AUTHORS}
+\subsection*{TITLE}
 \begin{itemize}[label={}, leftmargin=*]
@@
 % Part I.a.2 Author info
-\subsection*{Author}
+\subsection*{AUTHORS}
 \begin{itemize}[label={}, leftmargin=*]
@@
 
+% Part I.b Abstract
+\doublespacing
+\section*{ABSTRACT}
+\PAPERABSTRACT
+\clearpage
+
+% Part I.c Keywords
+\doublespacing
+\section*{KEYWORDS}
+\PAPERKEYWORDS
+\clearpage
+
+% statements
+
+\doublespacing
+\section*{Acknowledgement}
+\ACKNOWLEDGMENTS
+\clearpage
+
 \subsection*{Data Availability Statement}
@@
 \subsection*{Conflict of Interest Disclosure}
-The author declare that they have no conflicts of interest regarding the publication of this manuscript.
+The author declares that they have no conflicts of interest regarding the publication of this manuscript.
 
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Part II. Manuscript main text
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\pagenumbering{arabic}
+\setcounter{page}{1}
+\renewcommand*{\thefootnote}{\arabic{footnote}}
+\doublespacing
+\begingroup
+  \centering
+  \Large Manuscript Text\\[1em]
+\endgroup
+% Manuscript main text
+\input{manuscript_main}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Part III. Bibliography
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\newpage
+\begingroup
+\setstretch{1.0}
+\setlength\bibitemsep{5.0pt}
+\printbibliography[title=References for Manuscript]
+\endgroup
+\pagebreak
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Part IV. Tables and figures
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\input{manuscript_tab_fig}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% PART V. Appendix
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% \appendix
+% \begin{refsection}
+
+% % reset page numbers
+% \pagenumbering{arabic}
+% \setcounter{page}{1}
+
+% % Title Page for Appendix
+% % special footnote symbols
+% \renewcommand*{\thefootnote}{\fnsymbol{footnote}}
+% \begingroup
+% \doublespacing
+% \centering
+% \Large Online Appendix \\[1.5em]
+% \LARGE \PAPERTITLE \\[0.75em]
+% \large
+% \AUTHORHADAH\footnote[1]{\AUTHORHADAHINFO}\\[1.0em]
+% \endgroup
+% \clearpage
+
+% % Online appendix
+% \input{manuscript_online_appendix}
+
+% % Appendix Bibliography
+% \begingroup
+% \setstretch{1.0}
+% \setlength\bibitemsep{0pt}
+% \printbibliography[title=References for Online Appendix]
+% \endgroup
+% \pagebreak
+% \end{refsection}
+
 \end{document}

### 2024-10-15 manuscripts/08-JERP-v6/tab06-regression.tex
edits for JERP

--- 2023-05-09 output/tables/tab06-regression.tex
@@
 \begin{table}[H]
-
+\centering\centering
 \caption{Effect of Having Hispanic Last Name by Hispanic Identity\label{tab:identreg}}
 \centering
-\resizebox{\linewidth}{!}{
+\resizebox{\ifdim\width>\linewidth\linewidth\else\width\fi}{!}{
 \begin{threeparttable}
@@
 \midrule
-$HW_{ist} \times Hispanic_{ist}$ & \num{-0.09} & \num{-0.06} & \num{-0.05} & \num{-0.04} & \num{-0.04}\\
- & (\num{0.07}) & (\num{0.07}) & (\num{0.06}) & (\num{0.06}) & (\num{0.06})\\
-$Hispanic_{ist}$ & \num{-0.13}** & \num{-0.11} & \num{-0.09}** & \num{-0.03} & \num{-0.02}\\
- & (\num{0.07}) & (\num{0.07}) & (\num{0.04}) & (\num{0.05}) & (\num{0.04})\\
-$HW_{ist}$ & \num{0.03} & \num{0.01} & \num{0.00} & \num{0.02} & \num{0.02}\\
- & (\num{0.06}) & (\num{0.06}) & (\num{0.06}) & (\num{0.06}) & (\num{0.05})\\
+$HW_{ist} \times Hispanic_{ist}$ & \num{-0.09} & \num{-0.06} & \num{0.01} & \num{0.02} & \num{0.03}\\
+ & (\num{0.07}) & (\num{0.07}) & (\num{0.07}) & (\num{0.07}) & (\num{0.07})\\
+$Hispanic_{ist}$ & \num{-0.13}** & \num{-0.11} & \num{-0.12}*** & \num{-0.05} & \num{-0.03}\\
+ & (\num{0.07}) & (\num{0.07}) & (\num{0.04}) & (\num{0.04}) & (\num{0.04})\\
+$HW_{ist}$ & \num{0.03} & \num{0.01} & \num{-0.05} & \num{-0.04} & \num{-0.04}\\
+ & (\num{0.06}) & (\num{0.06}) & (\num{0.07}) & (\num{0.07}) & (\num{0.06})\\
 Constant & \num{10.53}*** & \num{9.57}*** &  &  & \\
@@
 \midrule
-$HW_{ist} \times Hispanic_{ist} + HW_{ist}$ & -0.06*** & -0.05** & -0.05** & -0.02 & -0.01\\
+$HW_{ist} \times Hispanic_{ist} + HW_{ist}$ & -0.06*** & -0.05** & -0.04* & -0.02 & -0.01\\
  & (0.02) & (0.02) & (0.02) & (0.02) & (0.02)\\
@@
 Year FE &  &  & X & X & X\\
-Age FE &  &  & X & X & X\\
-Education FE &  &  &  & X & X\\
+State-Year FE &  &  & X & X & X\\
+Age &  &  & X & X & X\\
+Education &  &  &  & X & X\\
 Parental Background &  &  &  &  & X\\

### 2024-10-15 manuscripts/08-JERP-v6/tab12-occ.tex
edits for JERP

\begin{table}[H]
\centering\centering
\caption{Descriptive Statistics About Selection Into Occupation by Type \label{tab:occ-sum}}
\centering
\resizebox{\ifdim\width>\linewidth\linewidth\else\width\fi}{!}{
\begin{threeparttable}
\begin{tabular}[t]{>{}lcccc}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{1}{c}{Hispanic Last Name} & \multicolumn{1}{c}{White Last Name} & \multicolumn{2}{c}{Difference-in-means: HW - WH} \\
\cmidrule(l{3pt}r{3pt}){2-2} \cmidrule(l{3pt}r{3pt}){3-3} \cmidrule(l{3pt}r{3pt}){4-5}
Occupation & Hispanic-White & White-Hispanic & Difference & p value\\
\midrule
\textbf{Management and Business} & 0.111 & 0.129 & -0.018 & 0.000\\
\textbf{STEM Occupations} & 0.030 & 0.040 & -0.009 & 0.000\\
\textbf{Healthcare Occupations} & 0.063 & 0.058 & 0.005 & 0.002\\
\textbf{Education and Social Services} & 0.062 & 0.064 & -0.002 & 0.237\\
\textbf{Arts, Media, and Entertainment} & 0.028 & 0.034 & -0.006 & 0.000\\
\addlinespace
\textbf{Service Occupations} & 0.107 & 0.099 & 0.008 & 0.000\\
\textbf{Manual and Industrial Labor} & 0.199 & 0.171 & 0.028 & 0.000\\
\bottomrule
\end{tabular}
\begin{tablenotes}
\item[1] Source: Current Population Survey (CPS) 1994-2019.
\end{tablenotes}
\end{threeparttable}}
\end{table}

### 2024-10-15 manuscripts/submit_journal-cover.tex
edits for JERP

\documentclass[12pt,english]{article}

\input{preamble_package}
% Citation comes loading other packages, so that footnotes follow biblatex-chicago style
\input{preamble_citation}

\begin{document}

\thispagestyle{empty}
\begingroup
  \doublespacing
  \centering
  \LARGE Submission to the Journal of Economics, Race and Policy, \\[0.25em]
  % \LARGE ACCEPTED MANUSCRIPT: MS2020XXX \\[1.0em]
\endgroup

\noindent Dear Editors of the Journal of Economics, Race and Policy,

\bigskip
\noindent I am writing to submit my manuscript titled "The Impact of Hispanic Last Names on Educational and Labor Market Outcomes" for consideration in the Journal of Economics, Race and Policy.

\bigskip
\noindent This paper examines the critical issue of discrimination against Hispanics in the United States labor market, with a particular focus on the impact of Hispanic last names and self-identification on educational attainment and earnings. As the Hispanic population continues to grow in the United States, understanding the extent and mechanisms of discrimination against this group becomes increasingly important for policymakers, educators, and economists alike.

\bigskip
\noindent The research presented in this paper makes several important contributions:

1. It employs a novel methodological approach, comparing children of inter-ethnic marriages (Hispanic-White and White-Hispanic) to isolate the effect of having a Hispanic last name on labor market outcomes.

2. It provides evidence of disparities in both educational attainment and earnings associated with Hispanic last names, even when controlling for family background and other factors.

3. It contributes to the growing literature on discrimination in access to education and its subsequent effects on labor market outcomes for minority groups.

4. It offers insights into the heterogeneity of outcomes within the Hispanic population, examining differences between Mexican and non-Mexican Hispanics.

\bigskip
\noindent The findings of this study have important implications for understanding the persistent economic disparities faced by Hispanics in the United States. By demonstrating that individuals with Hispanic last names complete fewer years of education and earn less than their counterparts with non-Hispanic names, even when controlling for family background, this research highlights the need for targeted policies to address these inequalities.

\bigskip
\noindent I believe this paper aligns well with the scope and aims of the Journal of Economics, Race and Policy, as it addresses critical issues at the intersection of economics, race, and public policy. The interdisciplinary nature of this research, combining economic analysis with insights from sociology and education, makes it particularly suitable for your journal's readership.

\bigskip
\noindent Thank you for your consideration. I look forward to your response and any feedback you may have.

I am including in this document:
\begin{enumerate}
    \item Title page information
    \begin{itemize}
        \item Title and Author
        \item Abstract
        \item Keywords
    \end{itemize}
    \item The manuscript
    \begin{itemize}
        \item Main manuscript text
        \item Main manuscript references
        \item Main manuscript figures
        \item Main manuscript tables
    \end{itemize}    
    \item The online appendix, which is separately paged.
\end{enumerate}

\bigskip
\noindent Sincerely,
\noindent Hussain Hadah

% \section*{Acknowledgment}
% I thank Professors Willa Friedman, Chinhui Juhn, Vikram Maheshri, and Yona Rubinstein for their support and advice. I also thank Aimee Chin, Steven Craig, German Cubas, Elaine Liu, Fan Wang, and the participants of the Applied Microeconomics Workshop at the University of Houston, and European Society for Population Economics (ESPE) for helpful feedback.

% \section*{Data and Programs}

% The data that support the findings of this study will be openly available to all researchers after the review process. For immediate information regarding the data and/or computer programs used for this study, please contact Hussain Hadah at \href{mailto:hhadah@tulane.edu}{hhadah@tulane.edu}.

% \clearpage

\end{document}

### 2024-10-16 manuscripts/08-JERP-v6/manuscript_title_etc.tex
edits for JERP
submission to JERP

--- 2024-10-15 manuscripts/08-labor-research-v6/manuscript_title_etc.tex
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\newcommand{\PAPERKEYWORDS}{\hspace{1em}\makebox{\textbf{Keywords}: Economics of Minorities, Race, and Immigrants; Discrimination and Prejudice}}
-\newcommand{\PAPERJEL}{\textbf{JEL}: J71; J64; J15}
+\newcommand{\PAPERKEYWORDS}{\hspace{1em}\makebox{\textbf{Keywords}: Discrimination and Prejudice; Wage Gap; Inter-ethnic Families}}
+\newcommand{\PAPERJEL}{\textbf{JEL}: J71; J64; J15; J31; I24; J62}
 
@@
 I find that individuals with Hispanic surnames often earn less, with a notable wage gap favoring those with White surnames.
-People born to Hispanic fathers and White mothers receive 0.3 years of education less than those born to White fathers and Hispanic mothers.
+People born to Hispanic fathers and White mothers receive 0.2 years of education less than those born to White fathers and Hispanic mothers.
 Men born to Hispanic fathers and White mothers are 1 percentage point more likely to be unemployed, and earn 5 percentage points less than those born to White fathers and Hispanic mothers.

### 2025-01-11 manuscripts/08-JERP-v6/manuscript_main.tex
submission to JERP
add occ fe table

@@
 They attribute a big part of this gap to differences in education, experience, immigration status, and regional differences.
+
+Understanding discrimination against Hispanics in labor markets has far-reaching implications that extend beyond individual earnings.
+Labor market discrimination can create persistent intergenerational disadvantages by limiting economic mobility, reducing access to quality healthcare and education, and constraining residential choices \autocite{chettyUnitedStatesStill2014, chettyEffectsExposureBetter2016,chettyFadingAmericanDream2017}.
+These barriers may trap Hispanic families in cycles of lower socioeconomic status, as reduced earnings and employment opportunities limit their ability to invest in their children's human capital or build wealth through homeownership and savings.
+While previous research has documented earnings gaps between Hispanics and non-Hispanic Whites, this paper makes several important contributions.
+First, it develops a novel empirical strategy that better isolates the causal effect of Hispanic ethnicity on labor market outcomes by comparing children of inter-ethnic marriages.
+This approach helps control for typically unobservable family background characteristics that might confound traditional analyses.
+Second, it provides new evidence on the specific role of Hispanic surnames in driving discrimination, offering insights into how ethnic signals influence both educational attainment and labor market outcomes.
+Finally, by examining both education and employment outcomes, this study helps illuminate the channels through which ethnic discrimination may perpetuate economic disparities across generations.
 
@@
 
-Moreover, a large literature of audit studies in education economics investigated the effect of racial and ethnic signal on access to education.
+Moreover, audit studies in education economics investigated the effect of racial and ethnic signal on access to education.
 \textcite{bergman2018education} found that students with Hispanic sounding names received 2 percentage points fewer responses from schools than students with White sounding names.
@@
 Finally, \textcite{bourabain2023school} found more evidence of discrimination in access to education against students with underprivileged backgrounds in the Flemish education system.
+Discrimination can affect educational attainment through multiple channels: restricted access to advanced coursework, biased academic counseling, limited encouragement to pursue higher education, and subtle institutional barriers that may cause Hispanic students to feel unwelcome or unsupported in educational settings.
+While Hispanic college enrollment has increased substantially in recent decades, discrimination within educational institutions may still impede degree completion through these various mechanisms.
 This paper contributes to this literature by providing evidence that discrimination in access to education could lead to lower earnings for Hispanics.

### 2025-01-11 manuscripts/08-JERP-v6/tab15-regression-occ-fe.tex
add occ fe table

--- 2024-08-11 output/tables/tab05-regression.tex
@@
 State-Year FE &  &  & X & X & X\\
-Education &  &  &  & X & X\\
+Occupation FE &  &  &  & X & X\\
 Parental Background &  &  &  &  & X\\
-Observations & \num{3621} & \num{3621} & \num{3621} & \num{3621} & \num{3621}\\
+Observations & \num{3621} & \num{3621} & \num{3621} & \num{3487} & \num{3487}\\
 \bottomrule

### 2025-01-11 programs/95-make-all.R

@@ -20,7 +20,7 @@ thesis_tabs  <- paste0(mdir_git,"/my_paper/tables")
 thesis_plots <- paste0(mdir_git,"/my_paper/figure") 
 add_health_data_dir <- "/Users/hhadah/Dropbox/Research/My Research Data and Ideas/Identification_Paper/Data/Datasets"
 ACS_data  <- "/Users/hhadah/Documents/GiT/jmp-decomposition-replication/data/raw"
-manuscript_wd <- paste0(mdir_git,"/manuscripts/08-labor-research-v6") 
+manuscript_wd <- paste0(mdir_git,"/manuscripts/08-JERP-v6") 
 # pres_tabs <- paste0(mdir,"/presentations/jmp-presentation/tables")
 # pres_plots <- paste0(mdir,"/presentations/jmp-presentation/figures")
 

### 2025-01-13 manuscripts/08-JERP-v6/manuscript_main.tex
some edits to address reviwer comments

@@
 Specifically, it is essential to investigate the extent to which a person's Hispanic ethnicity may influence their employment prospects, wages, and career advancement.
-Understanding these labor market outcomes is critical, as they directly impact broader societal issues such as assimilation and economic mobility.
+Understanding these labor market outcomes is critical, as they directly impact broader societal issues such as assimilation and economic mobility \autocite{chettyUnitedStatesStill2014,chettyEffectsExposureBetter2016}.
 These factors serve as key indicators of how successfully Hispanics can integrate into society and ascend the socioeconomic ladder.
@@
 Does having a Hispanic last name affect labor market outcomes?
-I aim to show that comparing Hispanic Whites to non-Hispanic Whites might create an artificially higher earnings gap since the two groups differ in many observable characteristics.\footnote{By identifying as Hispanic, I refer to individuals who self-report their ethnicity as Hispanic on surveys or other data collection instruments.} \footnote{Observable characteristics refer to factors that can be measured and quantified, such as education level, work experience, and immigration status.} Others have attempted to compare how native-born White Hispanics fare to non-Hispanic Whites and foreign-born Hispanics.
+I aim to show that comparing Hispanic Whites to non-Hispanic Whites might create an artificially higher earnings gap since the two groups differ in many observable characteristics.\footnote{By identifying as Hispanic, I refer to individuals who self-report their ethnicity as Hispanic on surveys or other data collection instruments.} \footnote{Observable characteristics refer to factors that can be measured and quantified, such as education level, work experience, and immigration status.} My analysis focuses on US-born children with a foreign-born parent---i.e.
+inter-ethnic.\footnote{This study's focus on US-born children with foreign-born parents means the findings may not generalize to Hispanic children with US-born parents, who likely face different socioeconomic and cultural circumstances.
+Additionally, the analysis does not account for heterogeneity in immigrant characteristics across different Spanish-speaking countries of origin, such as variations in educational attainment, socioeconomic status, and gender-specific migration patterns, which could influence both parental selection into migration and subsequent child outcomes.
+} Others have attempted to compare how native-born White Hispanics fare to non-Hispanic Whites and foreign-born Hispanics.
 In \textcite{antman2020ethnic,antmanEthnicAttritionObserved2016,antmanEthnicAttritionObserved2016a,antmanEthnicAttritionAssimilation2020}, the authors compare the health and educational outcomes of Hispanic Whites to non-Hispanic Whites and native-born Hispanics to foreign-born Hispanics.
@@
 They attribute a big part of this gap to differences in education, experience, immigration status, and regional differences.
+This paper builds on those studies by emphasizing how cultural assimilation and generational status further shape educational and labor market disparities for Hispanics.
 
 Understanding discrimination against Hispanics in labor markets has far-reaching implications that extend beyond individual earnings.
-Labor market discrimination can create persistent intergenerational disadvantages by limiting economic mobility, reducing access to quality healthcare and education, and constraining residential choices \autocite{chettyUnitedStatesStill2014, chettyEffectsExposureBetter2016,chettyFadingAmericanDream2017}.
-These barriers may trap Hispanic families in cycles of lower socioeconomic status, as reduced earnings and employment opportunities limit their ability to invest in their children's human capital or build wealth through homeownership and savings.
-While previous research has documented earnings gaps between Hispanics and non-Hispanic Whites, this paper makes several important contributions.
+Labor market discrimination can create persistent intergenerational disadvantages by limiting economic mobility, reducing access to quality healthcare and education, and constraining residential choices \autocite{chettyUnitedStatesStill2014, chettyEffectsExposureBetter2016,chettyFadingAmericanDream2017,bowles2002inheritance, djajic2003assimilation}.
+Additionally, discrimination may foster occupational sorting and segregation, as shown by recent task-based models of racial wage gaps that identify race-specific barriers in “Contact” tasks and explain the stagnation of these gaps post-1980 \autocite{hurst2024task}.
+These barriers may trap Hispanic families in cycles of lower socioeconomic status, as reduced earnings and employment opportunities limit their ability to invest in their children's human capital or build wealth through homeownership and savings.\footnote{See \textcite{bowles2002inheritance} for an analysis of the intergenerational transmission of economic status .} While previous research has documented earnings gaps between Hispanics and non-Hispanic Whites, this paper makes several important contributions.
 First, it develops a novel empirical strategy that better isolates the causal effect of Hispanic ethnicity on labor market outcomes by comparing children of inter-ethnic marriages.
@@
 Finally, by examining both education and employment outcomes, this study helps illuminate the channels through which ethnic discrimination may perpetuate economic disparities across generations.
+Furthermore, it extends earlier findings by demonstrating how naming cues can compound structural inequalities and reinforce existing barriers faced by Hispanics.
 
@@
 Some earnings differences may also be due to discrimination which will have negative consequences.
-For example, discrimination against Hispanics can lead to reduced job opportunities, lower wages, and hinder assimilation.
-In this paper, I examine the role of having a Hispanic last name and identifying as Hispanic on labor market outcomes.
+For example, discrimination against Hispanics can lead to reduced job opportunities, lower wages, and hinder assimilation \autocite{chettyWhereLandOpportunity2014,bowles2002inheritance, djajic2003assimilation}.\footnote{This study focuses on U.S.-born children of inter-ethnic unions, many of whom have at least one foreign-born parent.
+Assimilation concerns remain relevant because foreign-born parents may pass on cultural norms and language preferences, which can still shape these children's educational and labor market outcomes.}
 
@@
 They found that resumes with Black-sounding names received substantially fewer callbacks than their White counterparts.
+Furthermore, these early innovations in leveraging name-based identification highlight the importance of understanding how such signals can capture both conscious and unconscious bias, thereby situating this study within a broader framework of research on race, names, and economic outcomes.
 
@@
 While Hispanic college enrollment has increased substantially in recent decades, discrimination within educational institutions may still impede degree completion through these various mechanisms.
-This paper contributes to this literature by providing evidence that discrimination in access to education could lead to lower earnings for Hispanics.
+This paper contributes to this literature by providing evidence that discrimination in access to education could lead to lower earnings for Hispanics, and it further shows how parental background and marriage market selection can importantly shape these outcomes in ways not previously explored by earlier studies.
 
@@
 
-I use two datasets:  the Integrated Public Use Microdata Series (IPUMS) Current Population Survey (CPS) Annual Social and Economic (ASEC) \autocite{cps2019} and the 1960 to 2000 US censuses \autocite{acs2019}.
+I use two datasets:  the Integrated Public Use Microdata Series (IPUMS) Current Population Survey (CPS), CPS' Annual Social and Economic (ASEC) supplement, and CPS' outgoing rotation \autocite{cps2019} and the 1960 to 2000 US censuses \autocite{acs2019}.
 
@@
 I then link these synthetic parents using the birth year of child, and parents' places of birth to the year of birth of the "children" and the parents' places of birth in the CPS sample.
+
+For my main analysis, I use different dependent variables.
+I construct different educational variables: years of education, whether the person has a high school diploma, whether the person has an associate degree, and whether the person has a bachelor degree.
+For labor market outcomes, I construct unemployment rate variable using the main CPS, log annual earnings using the ASEC, and log weekly earnings using the outgoing rotation.
+The unemployment indicator is a binary variable that equals 1 if someone is unemployed and 0 if employed, based on civilian labor force status.Log annual earnings is the log of total personal income from the CPS' ASEC supplement.
+Log weekly earnings is the natural logarithm of weekly earnings for hourly workers from the outgoing rotation CPS supplement.
+I use the following controls: age, hours worked, and state-year fixed effects.
+I also include the parents' characteristics from the synthetic parents data as controls in alternative specifications.
+These controls include income and education of the synthetic parents.
 
@@
 
+Moreover, my sample includes inter-ethnic children with Hispanic ancestry aged 25 to 40 that are citizens.
+I do not restrict the sample to the self-reported Hispanic identity and keep both Hispanic White and non-Hispanic White children because ethnic attrition could lead to biases estimates when measuring gaps as shown by \autocite{hadah2024effect}.
+\footnote{A person self-report Hispanic identity by answering the following Census question: ``Is this person Spanish/Hispanic/Latino?''} \footnote{Ethnic attrition happens when a U.S.-born descendant of a Hispanic immigrant fails to self-identify as Hispanic.
+For more discussion of this phenomenon see \autocite{antmanEthnicAttritionObserved2016,antmanEthnicAttritionAssimilation2020}} To maintain analytical clarity, I restricted the sample to Hispanic and non-Hispanic White individuals, excluding other racial groups who also identify as Hispanic, as this allows for a more focused examination of Hispanic ancestry effects without introducing confounding racial factors.
+
 The distribution of the four types of children is presented in Table \ref{tab:mat1}.
@@
 Children of WH (Column 2) and HW (Column 3) marriages fall in between WW and HH children.
+The rates of self-reported Hispanic identity vary significantly across groups.
+Among children of WW marriages (Column 1), only 4\% of men and 5\% of women identify as Hispanic.
+This proportion increases substantially for children of inter-ethnic marriages: in WH families (Column 2), 74\% of men and 78\% of women identify as Hispanic, while in HW families (Column 3), the rates are 83\% for men and 81\% for women.
+Children of HH marriages (Column 4) show the highest rates of Hispanic identification, with 96\% of men and 97\% of women self-reporting as Hispanic.\footnote{The ethnic attrition rates are similar to those found in \textcite{antmanEthnicAttritionObserved2016,antmanEthnicAttritionAssimilation2020, hadah2024effect}.}
 
@@
 I assume that parents have children between the ages of 25 and 40.
-Therefore, my sample consists of married White men and women with children that were born in the 1920 to 1975 cohorts\footnote{The construction of "synthetic parents" follows the method used by \textcite{rubinstein2014pride}.}.
+Therefore, my sample consists of married White men and women with children that were born in the 1920 to 1975 cohorts\footnote{The construction of ``synthetic parents'' follows the method used by \textcite{rubinstein2014pride}.}.
 
+For example, consider someone who was 35 years old in 1999, meaning they were born in 1964.
+If this person's mother was born in Mexico and their father was born in the United States, their ``synthetic parents'' would be identified using the 1970 Census, when this person was 6 years old.
+These ``synthetic parents'' are created by matching the characteristics of Mexican-American couples who had a 6-year-old child in 1970.
+The ``synthetic mother'' would have the average education level, income, and other socioeconomic characteristics of Mexican-born women who were married to U.S.-born men and had children around 1964, when they were between 20 and 35 years old (meaning they were born between 1929-1944).
+Similarly, the ``synthetic father'' would have the average characteristics of U.S.-born men who were married to Mexican-born women and had children in that same year.
+This method preserves the demographic composition of parents while assigning them historically accurate socioeconomic characteristics based on their origins and marriage patterns.
 I show the distribution of the four types of couples in Table \ref{tab:mat2}.
@@
 Such comparisons shed light on the nuanced dynamics of ethnicity, education, and economic outcomes in inter-ethnic marriages.
-
-% \footnote{I present in Tables \ref{tab:synthmex} and \ref{tab:syntnonmex} the summary statistics which detail the educational and economic profiles of parents from four different ethnic groupings—White White (WW), White Hispanic (WH), Hispanic White (HW), and Hispanic Hispanic (HH) on sub-samples of Hispanics of Mexican and non-Mexican ancestries.
+\footnote{I present in Tables \ref{tab:synthmex} and \ref{tab:syntnonmex} the summary statistics which detail the educational and economic profiles of parents from four different ethnic groupings—White White (WW), White Hispanic (WH), Hispanic White (HW), and Hispanic Hispanic (HH) on sub-samples of Hispanics of Mexican and non-Mexican ancestries.
 I find similar results that describe a selection into inter-ethnic marriages among the two groups.}

### 2025-01-14 manuscripts/08-JERP-v6/manuscript_main.tex
add sensetivity analysis tables

@@
 
-% Since Hispanics are a very heterogeneous group, I conduct a heterogeneity analysis on different samples of Hispanics.
+\subsection{Sensitivity Analysis}
+
+Since Hispanics are a very heterogeneous group, I conduct a sensitivity analysis on different samples of Hispanics.
 To increase the sample size of my analysis I estimate equation \ref{eq:1a} using weekly earnings as a dependent variable.
 Weekly earnings are available in all the monthly CPS surveys and not just the March Supplement.
-I present the results to these estimations in tables \ref{tab:lastnamereg-weekearm}-\ref{tab:lastnamereg-weekearm-cub}.
+I present the results to these estimations in tables \ref{tab:lastname-ed-reg-mex}, \ref{tab:lastname-ed-reg-nonmex}, \ref{tab:lastnamereg-weekearm-mex}, and \ref{tab:lastnamereg-weekearm-nonmex}.
 
-% First, in Table \ref{tab:lastnamereg-weekearm}, I present the results of estimating equation \ref{eq:1a} using weekly earnings as the dependent variable for the full sample.
+First, in Table \ref{tab:lastnamereg-weekearm}, I present the results of estimating equation \ref{eq:1a} using weekly earnings as the dependent variable for the full sample.
 Similar to the previous analysis, I find that a person with a Hispanic last name earns 4 percentage points less than a person without a Hispanic last name.
@@
 This gap also becomes an imprecise zero after controlling for education and parental background.
-Finally, in Table \ref{tab:lastnamereg-weekearm-cub}, I present the results for a sample of Cubans.
-I find that a Cuban with a Hispanic-sounding last name earns 4 percentage points less than a Cuban with a native-sounding last name, but this gap is statistically insignificant.
 

### 2025-01-14 manuscripts/08-JERP-v6/submit_journal-main.tex
add sensetivity analysis tables

@@
 
+% Responses to Referees
+\usepackage{blindtext}
+\newcommand{\aptxspc}{1.8}
+\newcommand{\rrquote}{1.1}
+\newcommand{\rrxspc}{1.5}
+
 \begin{document}
+
+% SUBMISSION COVER
+\thispagestyle{empty}
+\begingroup
+  \doublespacing
+  \centering
+  \LARGE REVISE AND RESUBMIT\\[0.25em]
+  \LARGE JOURNAL OF ECONOMICS, RACE, AND POLICY \\[0.25em]
+  \LARGE SUBMISSION ID: 33481294-df2a-4cce-80ee-c20f2b9975ac \\[1.0em]
+\endgroup
+This document and the associated contents in the revise and resubmit version of manuscript 33481294-df2a-4cce-80ee-c20f2b9975ac for the Journal of Economics, Rac
+e, and Policy. I include:
+\begin{enumerate}
+    \item Title page information
+    \begin{itemize}
+        \item Title and Author
+        \item Abstract
+        \item Keywords
+        \item Acknowledgment
+    \end{itemize}
+    \item The manuscript
+    \begin{itemize}
+        \item Main manuscript text
+        \item Main manuscript references
+        \item Main manuscript figures
+        \item Main manuscript tables
+    \end{itemize}  
+    \item The online appendix, which is separately paged. It includes an online appendix title page, appendix A, appendix B, appendix C, and references for citations in the appendix.
+\end{enumerate}
+\clearpage
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% \appendix
-% \begin{refsection}
+\appendix
+\begin{refsection}
 
-% % reset page numbers
-% \pagenumbering{arabic}
-% \setcounter{page}{1}
+% reset page numbers
+\pagenumbering{arabic}
+\setcounter{page}{1}
+\end{refsection}
 
-% % Title Page for Appendix
-% % special footnote symbols
-% \renewcommand*{\thefootnote}{\fnsymbol{footnote}}
-% \begingroup
-% \doublespacing
-% \centering
-% \Large Online Appendix \\[1.5em]
-% \LARGE \PAPERTITLE \\[0.75em]
-% \large
-% \AUTHORHADAH\footnote[1]{\AUTHORHADAHINFO}\\[1.0em]
-% \endgroup
-% \clearpage
+% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% % PART VI. Responses to Editor and Referees
+% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{refsection}
+% Reset section numbering
+\setcounter{section}{0}
+\renewcommand{\thesection}{\Alph{section}}
+\renewcommand{\thesubsection}{\Alph{section}.\arabic{subsection}}
+\renewcommand{\thesubsubsection}{\Alph{section}.\arabic{subsection}.\arabic{subsubsection}}
 
-% % Online appendix
-% \input{manuscript_online_appendix}
+        \section{Responses to Editors and Referee} \label{r&r:responses}
+        \begin{spacing}{\rrxspc}
+        \end{spacing}
 
-% % Appendix Bibliography
-% \begingroup
-% \setstretch{1.0}
-% \setlength\bibitemsep{0pt}
-% \printbibliography[title=References for Online Appendix]
-% \endgroup
-% \pagebreak
-% \end{refsection}
+    \newpage
+
+    \section{Responses to Referee One}
+    \begin{spacing}{\rrquote}
+    \begin{quotation}
+    \textbf{R1: } 1. 
+    \end{quotation}
+    \end{spacing}
+
+    \begin{spacing}{\rrxspc}
+    \end{spacing}
+
+    \clearpage
+    \pagebreak
+    \end{refsection}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% PART VII. Online Appendix
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{refsection}
+% Reset section numbering
+\setcounter{section}{0}
+\renewcommand{\thesection}{\Alph{section}}
+\renewcommand{\thesubsection}{\Alph{section}.\arabic{subsection}}
+\renewcommand{\thesubsubsection}{\Alph{section}.\arabic{subsection}.\arabic{subsubsection}}
+
+% Title Page for Appendix
+% special footnote symbols
+\renewcommand*{\thefootnote}{\fnsymbol{footnote}}
+\begingroup
+\doublespacing
+\centering
+\Large ONLINE APPENDIX \\[1.5em]
+\LARGE \PAPERTITLE \\[0.75em]
+\large
+\AUTHORHADAH\footnote[1]{\AUTHORHADAHINFO}\\[1.0em]
+\endgroup
+\clearpage
+
+% Online appendix
+\input{manuscript_online_appendix}
+
+\newpage
+\pagebreak
+
+% Appendix Bibliography
+\begingroup
+\setstretch{1.0}
+\setlength\bibitemsep{0pt}
+\printbibliography[title=References for Online Appendix]
+\endgroup
+\pagebreak
+\end{refsection}
 

### 2025-01-14 manuscripts/08-JERP-v6/tab16-educ-mex-regression.tex
add sensetivity analysis tables

--- 2024-08-28 manuscripts/07-economic-letters-v5/tables/tab13-regression.tex
@@
 \centering\centering
-\caption{Effect of Having Hispanic Last Name on Educational Outcomes\label{tab:lastname-ed-reg}}
+\caption{Effect of Having Hispanic Last Name: Hispanics with Mexican Ancestry \label{tab:lastname-ed-reg-mex}}
 \centering
@@
 \multicolumn{5}{l}{\textit{Panel A: Full Sample}}\\
-\midrule \hspace{1em}$HW_{ist}$ & -0.20*** & 0.01 & -0.02*** & -0.03***\\
-\hspace{1em} & (0.05) & (0.01) & (0.01) & (0.01)\\
-\hspace{1em}Observations & 88377 & 90027 & 66927 & 90027\\
+\midrule \hspace{1em}$HW_{ist}$ & -0.25*** & 0.01 & -0.02*** & -0.03***\\
+\hspace{1em} & (0.06) & (0.01) & (0.01) & (0.01)\\
+\hspace{1em}Observations & 62483 & 63867 & 52008 & 63867\\
 \addlinespace[0.5em]
 \multicolumn{5}{l}{\textit{Panel B: Women}}\\
-\midrule \hspace{1em}$HW_{ist}$ & -0.25*** & 0.02 & -0.03*** & -0.04***\\
-\hspace{1em} & (0.06) & (0.01) & (0.01) & (0.01)\\
-\hspace{1em}Observations & 46516 & 47302 & 34334 & 47302\\
+\midrule \hspace{1em}$HW_{ist}$ & -0.31*** & 0.01 & -0.02** & -0.04***\\
+\hspace{1em} & (0.07) & (0.01) & (0.01) & (0.01)\\
+\hspace{1em}Observations & 33292 & 33946 & 27072 & 33946\\
 \addlinespace[0.5em]
 \multicolumn{5}{l}{\textit{Panel C: Men}}\\
-\midrule \hspace{1em}$HW_{ist}$ & -0.16** & 0.00 & -0.02** & -0.02\\
-\hspace{1em} & (0.07) & (0.01) & (0.01) & (0.01)\\
-\hspace{1em}Observations & 41861 & 42725 & 32593 & 42725\\
-Full Sample's Mean & 13.48 & 0.43 & 0.15 & 0.26\\
-Women's Mean & 13.58 & 0.44 & 0.16 & 0.27\\
-Men's Mean & 13.38 & 0.43 & 0.14 & 0.24\\
+\midrule \hspace{1em}$HW_{ist}$ & -0.16** & 0.01 & -0.01 & -0.02\\
+\hspace{1em} & (0.08) & (0.02) & (0.01) & (0.02)\\
+\hspace{1em}Full Sample's Mean & 13.07 & 0.47 & 0.13 & 0.19\\
+\hspace{1em}Observations & 29191 & 29921 & 24936 & 29921\\
+Women's Mean & 13.16 & 0.47 & 0.14 & 0.2\\
+Men's Mean & 12.96 & 0.46 & 0.13 & 0.17\\
 \bottomrule

### 2025-01-14 manuscripts/08-JERP-v6/tab17-educ-nonmex-regression.tex
add sensetivity analysis tables

--- 2024-08-28 manuscripts/07-economic-letters-v5/tables/tab13-regression.tex
@@
 \centering\centering
-\caption{Effect of Having Hispanic Last Name on Educational Outcomes\label{tab:lastname-ed-reg}}
+\caption{Effect of Having Hispanic Last Name: Hispanics with non-Mexican Ancestry \label{tab:lastname-ed-reg-nonmex}}
 \centering
@@
 \multicolumn{5}{l}{\textit{Panel A: Full Sample}}\\
-\midrule \hspace{1em}$HW_{ist}$ & -0.20*** & 0.01 & -0.02*** & -0.03***\\
-\hspace{1em} & (0.05) & (0.01) & (0.01) & (0.01)\\
-\hspace{1em}Observations & 88377 & 90027 & 66927 & 90027\\
+\midrule \hspace{1em}$HW_{ist}$ & -0.25*** & 0.01 & -0.04*** & -0.04***\\
+\hspace{1em} & (0.07) & (0.01) & (0.01) & (0.01)\\
+\hspace{1em}Observations & 37893 & 38372 & 24567 & 38372\\
 \addlinespace[0.5em]
 \multicolumn{5}{l}{\textit{Panel B: Women}}\\
-\midrule \hspace{1em}$HW_{ist}$ & -0.25*** & 0.02 & -0.03*** & -0.04***\\
-\hspace{1em} & (0.06) & (0.01) & (0.01) & (0.01)\\
-\hspace{1em}Observations & 46516 & 47302 & 34334 & 47302\\
+\midrule \hspace{1em}$HW_{ist}$ & -0.35*** & 0.03** & -0.04** & -0.06***\\
+\hspace{1em} & (0.11) & (0.02) & (0.02) & (0.02)\\
+\hspace{1em}Observations & 19774 & 20039 & 12476 & 20039\\
 \addlinespace[0.5em]
 \multicolumn{5}{l}{\textit{Panel C: Men}}\\
-\midrule \hspace{1em}$HW_{ist}$ & -0.16** & 0.00 & -0.02** & -0.02\\
-\hspace{1em} & (0.07) & (0.01) & (0.01) & (0.01)\\
-\hspace{1em}Observations & 41861 & 42725 & 32593 & 42725\\
-Full Sample's Mean & 13.48 & 0.43 & 0.15 & 0.26\\
-Women's Mean & 13.58 & 0.44 & 0.16 & 0.27\\
-Men's Mean & 13.38 & 0.43 & 0.14 & 0.24\\
+\midrule \hspace{1em}$HW_{ist}$ & -0.24** & 0.00 & -0.04* & -0.04*\\
+\hspace{1em} & (0.10) & (0.02) & (0.02) & (0.02)\\
+\hspace{1em}Full Sample's Mean & 14.07 & 0.38 & 0.16 & 0.36\\
+\hspace{1em}Observations & 18119 & 18333 & 12091 & 18333\\
+Women's Mean & 14.17 & 0.38 & 0.17 & 0.38\\
+Men's Mean & 13.97 & 0.38 & 0.16 & 0.34\\
 \bottomrule

### 2025-01-14 programs/26-educ-regressions.R

@@ -190,3 +190,256 @@ regression_tab %>%
 
 regression_tab %>%
   save_kable(file.path(manuscript_wd,"tab13-regression.tex"))
+
+#------------------------------------------------------------------------------------
+# Mexican Ancenstry
+#------------------------------------------------------------------------------------
+Mexicans_USA  <- IndividualData |> 
+    filter((HW == 1 | WH == 1) & BirthPlaceMom %in% c("Mexico", "USA") & BirthPlaceDad %in% c("Mexico", "USA"))
+
+regression <- list(
+  "Panel A: Full Sample" = list(
+  "\\specialcell{(1) \\\\ Years of \\\\ Education}"  = feols(Education ~ HW  +  age +.[ParentControls]  | year*statefip, vcov = ~statefip, data = Mexicans_USA),
+  "\\specialcell{(2) \\\\ High School \\\\ Dropout}"  = feols(hs_dropout ~ HW  +  age +.[ParentControls]  | year*statefip, vcov = ~statefip, data = Mexicans_USA),
+  "\\specialcell{(3) \\\\ Associate Degree}"         = feols(associate_degree ~ HW  +  age +.[ParentControls]  | year*statefip, vcov = ~statefip, data = Mexicans_USA),
+  "\\specialcell{(4) \\\\ Bachelor Degree}"          = feols(ba_degree ~ HW  +  age +.[ParentControls]  | year*statefip, vcov = ~statefip, data = Mexicans_USA)
+  ),
+
+  "Panel B: Women" = list(
+  "\\specialcell{(1) \\\\ Years of \\\\ Education}"  = feols(Education ~ HW  +  age +.[ParentControls]  | year*statefip, vcov = ~statefip, data = Mexicans_USA  |> filter(sex == 2)),
+  "\\specialcell{(2) \\\\ High School \\\\ Dropout}"  = feols(hs_dropout ~ HW  +  age +.[ParentControls]  | year*statefip, vcov = ~statefip, data = Mexicans_USA  |> filter(sex == 2)),
+  "\\specialcell{(3) \\\\ Associate Degree}"         = feols(associate_degree ~ HW  +  age +.[ParentControls]  | year*statefip, vcov = ~statefip, data = Mexicans_USA  |> filter(sex == 2)),
[…]

### 2025-01-18 manuscripts/08-JERP-v6/manuscript_main.tex
some more edits

@@
 
-In this section, I present the results from estimating the specification presented in equation \ref{eq:1a}.
+% In this section, I present the results from estimating the specification presented in equation \ref{eq:1a}.
 I estimate the mean educational outcomes of White, Native-born (results in Table \ref{tab:lastname-ed-reg}), Hispanics aged 25-40, and mean unemployment of White, Native-born, Hispanics men aged 25-40 (results in Table \ref{tab:lastnamereg-emp}).
@@
 
-First, I find that an inter-ethnic person with a Spanish-sounding last name receives fewer total years of education than an inter-ethnic person who has a White last name.
+% First, I find that an inter-ethnic person with a Spanish-sounding last name receives fewer total years of education than an inter-ethnic person who has a White last name.
 A person with a Spanish-sounding last name receives 0.2 years of education less than a person with a White last name, have the same high school dropout rate as a person with a White last name, and 2 percentage points less likely to have an associate degree.
@@
 
-Second, I find that an inter-ethnic person with a Spanish-sounding last names are 1 percentage point more likely to be unemployed and earn less than an inter-ethnic who has a White last name.
+% Second, I find that an inter-ethnic person with a Spanish-sounding last names are 1 percentage point more likely to be unemployed and earn less than an inter-ethnic who has a White last name.
 A person with a Spanish-sounding last name earns 5 percentage points less than a person with a White last name.
@@
 
-\subsection{The Effect of Having a Hispanic Last Name on Educational Outcomes}
-
-I provide the results of the estimation of equation \ref{eq:1a} in Table \ref{tab:lastname-ed-reg}.
-I estimate the mean educational outcomes of White, U.S.-born Hispanics ages 25-40.
-I also restrict the sample to children of HW and WH parents.
-The omitted group is children of WH parents.
-Column 1 in Table \ref{tab:lastname-ed-reg} is the difference in total years of education between HW children and their WH peers.
-Column 2 is the difference in the probability of having been a high school dropout.
-Column 3 is the difference in the probability of having an associate degree.
-Column 4 is the difference in the probability of having a bachelor's degree.
-All regressions include controls for age, parental education and income, and state-year fixed effects.
-
-Overall, there is a significant gap in total years of education between HW and WH children.
-HW children receive 0.2 years of education less than WH children.
-The gap between HW and WH women is larger than the gap between HW and WH men.
-Women with a Hispanic last name receive 0.25 years of education less than WH women.
-The gap between HW and WH men is equal to 0.16 years.
-
-Even though there is a significant, albeit small, gap in total years of education between HW and WH children, there is no significant gap in the probability of being a high school dropout.
-The gap between HW and WH high school dropouts is equal to a statistically insignificant 1 percentage point.
-The same is true for HW women (2 percentage points) and men (0 percentage points).
-
-However, there is a notable difference when it comes to higher education outcomes.
-HW children are 2 percentage points less likely to earn an associate degree and 3 percentage points less likely to earn a bachelor's degree compared to their WH peers.
-HW children are 2 percentage points less likely to earn an associate degree compared to their WH peers, representing a 13.3\% reduction relative to the WH associate degree rate of 15\%.
-HW children are 3 percentage points less likely to earn a bachelor's degree compared to their WH peers, which is a 13\% reduction relative to the WH bachelor's degree rate of 23\%.
-These gaps are slightly larger for HW women, who are 3 percentage points less likely to earn an associate degree and 4 percentage points less likely to earn a bachelor's degree.
-For HW men, the gap is 2 percentage points for an associate degree and statistically insignificant for a bachelor's degree.
-
-These results suggest that while the overall educational attainment gap between HW and WH children is small in terms of years of education, the disparities become more pronounced when considering higher education milestones, particularly for HW women.
-However, given the fact that HWs have more educated mothers than WHs, we would expect them to have higher levels of educational attainment \autocite{kimball2009risk, gould2020does}.
-This could indicate potential barriers or discrimination in access to higher education for HW children, particularly.
-
-\subsection{The Effect of Having a Hispanic Last Name on Labor Market Outcomes}
-
-I provide the results to the estimation of equation \ref{eq:1a} in Tables \ref{tab:lastnamereg-emp} and \ref{tab:lastnamereg} on unemployment and log earnings.
-I estimate the mean unemployment of White, Native-born, Hispanics men aged 25-40.
-I estimate the mean wages of White, U.S.-born, Hispanic men aged 25-40 who are employed FTFY as waged and salaried workers.
-I also restrict the sample to children of HW and WH (omitted) parents.
-Column 1 in Table \ref{tab:lastnamereg} is the average crude earnings gap in log annual earnings between HW workers and their WH peers.
-In the next 4 columns, I introduce the results with controls for hours worked, state FE, year FE, age FE, Education FE, and parental background.
-
-In addition to examining earnings, I also analyzed the effect of having a Hispanic last name on unemployment rates.
-Table \ref{tab:lastnamereg-emp} presents the results of this analysis.
-Column 1 shows that individuals with Hispanic last names (HW) have a 1 percentage point higher unemployment rate compared to those with White last names.
-This gap persists even after controlling for age, state fixed effects, year fixed effects, and state-year fixed effects (Column 2), though the significance level drops.
-When education is included as a control (Column 3), the gap remains at 1 percentage point.
-Finally, after controlling for parental background (Column 4), the 1 percentage point difference in unemployment rates becomes statistically insignificant.
-These results suggest that while there is an initial unemployment gap associated with having a Hispanic last name, much of this difference can be explained by factors such as education and parental background.
-The mean unemployment rate for individuals with Hispanic last names (HW) is 7\% across all specifications.
-
-Overall, the crude gap between HW and WH workers is equal to 5 percentage points (Table \ref{tab:lastnamereg} column 1).
-An inter-ethnic with a Hispanic last name earns 5 percentage points less than an inter-ethnic with a White last name.
-Even after controlling for hours worked, and including state, year, and age FEs in the estimation, the gap stays at 5 percentage points.
-This gap, however, could be entirely explained by educational differences.
-An inter-ethnic with a Hispanic last name earns 1 percentage points less than an inter-ethnic with a White last name, but the result is statistically insignificant.
-
+a
 \subsection{Sensitivity Analysis}
@@
 
+I also present the results of the estimation of equation \ref{eq:1a} using log annual earnings as the dependent variable but with occupation fixed effects instead of education fixed effects.
+I present the results of this estimation in Table \ref{tab:lastnamereg-occ}.
+I find that the gap between people with Hispanic-sounding last names and those with native-sounding last names is explained by differences in occupation, similar to my findings with educational controls.
+I find that a person with a Hispanic-sounding last name earns 5 percentage points less than a person with a native-sounding last name.
+This gap becomes an imprecise zero after controlling for occupation.
+
 \section{Conclusion}\label{sec:con1}
@@
 I compare the children of inter-ethnic marriages to study the labor and educational markets impact of having a Hispanic last name.
-I find that inter-ethnic people with Spanish-sounding last names receive 0.2 years of education less than inter-ethnic counterparts that have a native-sounding last name.
-When I compare the earnings of HW and WH children, which captures the effect of having a Hispanic last name, HW children are 1 percentage point more likely to be unemployed than WH children.
-HW children earn 5 percentage points less than WH children.
+I find that inter-ethnic people with Spanish-sounding last names receive 0.2 years of education less than inter-ethnic counterparts that have a native-sounding last name, representing a 1.4\% reduction in total years of education.
+This gap is larger for women (0.25 years, or 1.8\%) than for men (0.16 years, or 1.2\%).
+When I compare the labor market outcomes of HW and WH children, which captures the effect of having a Hispanic last name, HW children are 1 percentage point more likely to be unemployed than WH children, representing a 14.3\% increase from the mean unemployment rate.
+In terms of earnings, HW children earn 5 percentage points (4.9\%) less than WH children.
 Thus, by comparing inter-ethnic children, a person with a Hispanic last name makes 5 percentage points less than someone with a White last name.
-When I control for education, the last name effect decreases to a statistically insignificant 2 percentage points earnings gap.
+When I control for education, the last name effect decreases to a statistically insignificant 1 percentage point earnings gap.
+
+The findings reveal gaps that could be attributed to discrimination against individuals with Hispanic surnames, with effects varying across educational levels and gender.
+While the impact on total years of education appears modest (1.4\% reduction), the more substantial gaps in higher education attainment—particularly the 13\% reduction in bachelor's degree completion and 13.3\% reduction in associate degree attainment—suggest that discrimination may intensify at crucial educational transition points.
+This pattern is especially pronounced for women with Hispanic surnames, who face larger penalties in both total education (1.8\% reduction versus 1.2\% for men) and degree completion (14.8\% reduction in bachelor's degree attainment).
 
@@
 
+In the labor market, the 14.3\% higher unemployment rate and 4.9\% earnings penalty for those with Hispanic surnames indicate persistent discrimination even among individuals who share similar family backgrounds.
+Putting the results in context, \autocite{trejoWhyMexicanAmericans1997} find an earning gap between US-born Hispanics and US-born whites that is equal to 30.7 percentage points.
+My findings could be an indication that out of the 30.7 percentage points gap, at least 16\% of it could be attributed to discrimination against Hispanics.
+However, the fact that the earnings gap becomes statistically insignificant after controlling for education suggests that much of the labor market discrimination operates through educational channels rather than direct wage discrimination.
+These findings leave several important questions unanswered.
+
 While the earnings gap between children of inter-ethnic parents with and without a Hispanic last name disappears when controlling for education, it does not necessarily indicate the absence of discrimination.
@@
 This is especially the case when parental characteristics indicate that people with a Hispanic-sounding last-names should in theory complete more years of education---since mothers of inter-ethnic children with a Hispanic last-names have more years of education and earn more.
+While I observe where discrimination manifests (particularly in higher education access), my analysis cannot identify the specific mechanisms—such as teacher bias, institutional barriers, or social networks—through which surname-based discrimination operates.
 Consequently, further research is needed to comprehensively understand the earnings gaps between Hispanics and Whites.

### 2025-01-18 manuscripts/08-JERP-v6/tab13-regression.tex
edits for JERP
some more edits

--- 2024-08-28 manuscripts/07-economic-letters-v5/tables/tab13-regression.tex
@@
 \centering\centering
-\caption{Effect of Having Hispanic Last Name on Educational Outcomes\label{tab:lastname-ed-reg}}
+\caption{Effect of Having Hispanic Last Name \label{tab:lastname-ed-reg}}
 \centering
@@
 \hspace{1em} & (0.07) & (0.01) & (0.01) & (0.01)\\
+\hspace{1em}Full Sample's Mean & 13.48 & 0.43 & 0.15 & 0.26\\
 \hspace{1em}Observations & 41861 & 42725 & 32593 & 42725\\
-Full Sample's Mean & 13.48 & 0.43 & 0.15 & 0.26\\
 Women's Mean & 13.58 & 0.44 & 0.16 & 0.27\\
 Men's Mean & 13.38 & 0.43 & 0.14 & 0.24\\
+p-value test Women - Men & $p=$0.87 & $p=$0.25 & $p=$0.92 & $p=$0.24\\
 \bottomrule

### 2025-01-18 programs/16-table-five-regression.R

@@ -221,6 +221,20 @@ reg8 <- list(
 # Earnigns
 #--------------------------------------
 
+# calculate mean for all sample
+IndividualData_mean <- IndividualData |> 
+  filter(HW == 1)
+all_sample <- c("\\specialcell{HW's Mean \\\\ Unemployment}", 
+                      round(mean(IndividualData_mean$lninctot_1999, na.rm=T), 2),
+                      round(mean(IndividualData_mean$lninctot_1999, na.rm=T), 2),
+                      round(mean(IndividualData_mean$lninctot_1999, na.rm=T), 2), 
+                      round(mean(IndividualData_mean$lninctot_1999, na.rm=T), 2),
+                      round(mean(IndividualData_mean$lninctot_1999, na.rm=T), 2))
+
+
+dim(all_sample) <- c(1,6)
+all_sample <- as.data.frame(all_sample)
+
 controling_for <-  c("\\textit{Controlling for:}", " ", "", " ", "", "")
 dim(controling_for) <- c(1,6)
 
@@ -247,7 +261,8 @@ all_row <- rbind(
   hoursworked,
   age_cont,
   educ_cont,
-  parentalback)
+  parentalback,
+  all_sample)
 
 attr(all_row, 'position') <- c(#7:8, 
                               5:7, 11:12)

### 2025-01-22 manuscripts/08-JERP-v6/manuscript_main.tex
reviwer comments

--- 2025-01-14 manuscripts/08-JERP-v6/manuscript_main.tex
@@
 Moreover, children from a Hispanic father and White mother household will have a Hispanic last name from their fathers, enabling the investigation of how ethnic signals, such as having a Hispanic last name, affect annual log earnings.
+
+Moreover, when analyzing the effect of having a Hispanic last name on educational outcomes, I provide separate analysis of men and women.
+Separating the results for men and women is motivated by evidence that gender and ethnic discrimination operate through distinct mechanisms.
+As \textcite{bertrand2020gender} discusses, women's labor market outcomes are shaped by persistent gender norms and stereotypes, particularly around motherhood and caregiving.
+\textcite{antecol2002relative} found that young Mexican women earned 9.5\% less than young white women in 1994, with differences in education appearing to be the most important explanation for this gap.
+This contrasts with findings for black women, where differences in labor force attachment were the key driver of wage disparities.
+The way discrimination manifests can also evolve differently by gender over time--- \textcite{goldin2004making} document how women's professional identity and career continuity have become increasingly important factors in labor market outcomes, while \textcite{darity2015tour} argue that a stratification economics approach helps reveal how group-based hierarchies are maintained through both discriminatory practices and intergenerational resource transfers.
+Therefore, disaggregating the results by gender provides important insights into how the intersection of ethnic and gender discrimination shapes educational and economic opportunities.
 
@@
 Moreover, my sample includes inter-ethnic children with Hispanic ancestry aged 25 to 40 that are citizens.
-I do not restrict the sample to the self-reported Hispanic identity and keep both Hispanic White and non-Hispanic White children because ethnic attrition could lead to biases estimates when measuring gaps as shown by \autocite{hadah2024effect}.
+I do not restrict the sample to the self-reported Hispanic identity and keep both Hispanic White and non-Hispanic White children because ethnic attrition could lead to biases estimates when measuring gaps as shown by \textcite{hadah2024effect}.
 \footnote{A person self-report Hispanic identity by answering the following Census question: ``Is this person Spanish/Hispanic/Latino?''} \footnote{Ethnic attrition happens when a U.S.-born descendant of a Hispanic immigrant fails to self-identify as Hispanic.
-For more discussion of this phenomenon see \autocite{antmanEthnicAttritionObserved2016,antmanEthnicAttritionAssimilation2020}} To maintain analytical clarity, I restricted the sample to Hispanic and non-Hispanic White individuals, excluding other racial groups who also identify as Hispanic, as this allows for a more focused examination of Hispanic ancestry effects without introducing confounding racial factors.
+For more discussion of this phenomenon see \textcite{antmanEthnicAttritionObserved2016,antmanEthnicAttritionAssimilation2020}} To maintain analytical clarity, I restricted the sample to Hispanic and non-Hispanic White individuals, excluding other racial groups who also identify as Hispanic, as this allows for a more focused examination of Hispanic ancestry effects without introducing confounding racial factors.
 
@@
 $\beta_{1}$ is the coefficient of interest in this specification.
-$\beta_{1}$ represents the gaps in outcomes between children of inter-ethnic marriages who have a Spanish-sounding last name versus a White last name.
+$\beta_{1}$ represents the gaps in outcomes between children of inter-ethnic marriages who likely have a Spanish-sounding last name versus a White last name.
 If $\beta_{1} > 0$, then people with a Hispanic last name have better outcomes than people with a White last name.
@@
 This is an improvement over traditional Oaxaca-Blinder-Kitagawa decomposition methods, which are limited in their ability to account for unobserved differences between groups by providing a more suitable comparison group to measure discrimination \autocite{oaxaca1973male,blinder1973wage,kitagawa1955components}.
-
-\subsection{Threats to Identification}
 
@@
 These results show that there is selection in marriage, and that the differences between children born to HW and children born to WH are less severe that those between the other two groups.
-Consequently, comparing WH and HW children to each other to analyze discrimination against Hispanics in the labor market.
+Consequently, comparing WH and HW children to each other to analyze discrimination against Hispanics in the labor market provides a better comparison since any gaps that emerge between these otherwise similar groups are more likely to be attributable to discrimination rather than underlying differences in family background or socioeconomic status.
 
@@
 Both Hispanic men and women in inter-ethnic couples marry white spouses that are, on average, more educated the the average HH couple.
-More importantly for human capital accumulation, the mother in the HW marriages---the children that have a Spanish sounding last-name---are more educated than their WH peers.
+More importantly for human capital accumulation, the mother in the HW marriages---the children likely have a Spanish sounding last-name---are more educated than their WH peers.
 Since mothers could be more important in the a child's human capital accumulation and education, this would suggest that a child with a Hispanic last-name should complete more years of education \autocite{gould2020does}.
@@
 \section{Results}\label{sec:results}
-
-In this section, I present the results from estimating the specification presented in equation \ref{eq:1a}.
-I estimate the mean educational outcomes of White, Native-born (results in Table \ref{tab:lastname-ed-reg}), Hispanics aged 25-40, and mean unemployment of White, Native-born, Hispanics men aged 25-40 (results in Table \ref{tab:lastnamereg-emp}).
-I estimate the mean wages of White, Native-born, Hispanic men aged 25-40 who are employed full-time and full-year (FTFY) as waged and salaried workers (results in Table \ref{tab:lastnamereg}).
-
-First, I find that an inter-ethnic person with a Spanish-sounding last name receives fewer total years of education than an inter-ethnic person who has a White last name.
-A person with a Spanish-sounding last name receives 0.2 years of education less than a person with a White last name, have the same high school dropout rate as a person with a White last name, and 2 percentage points less likely to have an associate degree.
-People with a Spanish-sounding last name are 3 percentage points less likely to have a bachelor's degree.
-
-Second, I find that an inter-ethnic person with a Spanish-sounding last names are 1 percentage point more likely to be unemployed and earn less than an inter-ethnic who has a White last name.
-A person with a Spanish-sounding last name earns 5 percentage points less than a person with a White last name.
-In other words, by comparing inter-ethnic children, a person with a Hispanic last name earns 5 percentage points less than someone with a White last name.
-However, more than half of the earnings gaps could be explained by educational differences.
-When I control for education, the last name effect decreases to a statistically insignificant 1 percentage points earnings gap.
-I also find a significant earnings gap between those that identify as Hispanic.
 
@@
 Second, in Table \ref{tab:lastnamereg-weekearm-mex}, I present the results for a sample of Mexican Hispanics.
-I find that a Mexican with a Hispanic-sounding last name earns 3 percentage points less than a Mexican with a native-sounding last name.
+I find that among individuals of Mexican origin, those with that likely have Hispanic surnames earn 3 percentage points less than those with a likely native-sounding surnames.
 This gap becomes an imprecise zero after controlling for education and parental background.
 Third, in Table \ref{tab:lastnamereg-weekearm-nonmex}, I present the results for a sample of non-Mexican Hispanics.
-The gap between non-Mexicans with Hispanic-sounding last names and those with native-sounding last names is similarly explained by educational differences.
-I find that a non-Mexican with a Hispanic-sounding last name earns 3 percentage points less than a non-Mexican with a native-sounding last name.
+The gap between non-Mexicans that likely have a Hispanic-sounding last names and those that likely have a native-sounding last names is similarly explained by educational differences.
+I find that a non-Mexican individual that likely has a Hispanic-sounding last name earns 3 percentage points less than a non-Mexican with a native-sounding last name.
 This gap also becomes an imprecise zero after controlling for education and parental background.
+
+I also present the results of the estimation of equation \ref{eq:1a} using log annual earnings as the dependent variable but with occupation fixed effects instead of education fixed effects.
+I present the results of this estimation in Table \ref{tab:lastnamereg-occ}.
+I find that the gap between people with a likely Hispanic-sounding last names and those with a likely native-sounding last names is explained by differences in occupation, similar to my findings with educational controls.
+I find that a person with a likely Hispanic-sounding last name earns 5 percentage points less than a person with a likely native-sounding last name.
+This gap becomes an imprecise zero after controlling for occupation.
 
@@
 I compare the children of inter-ethnic marriages to study the labor and educational markets impact of having a Hispanic last name.
-I find that inter-ethnic people with Spanish-sounding last names receive 0.2 years of education less than inter-ethnic counterparts that have a native-sounding last name.
-When I compare the earnings of HW and WH children, which captures the effect of having a Hispanic last name, HW children are 1 percentage point more likely to be unemployed than WH children.
-HW children earn 5 percentage points less than WH children.
+I find that inter-ethnic people with a likely Spanish-sounding last names receive 0.2 years of education less than inter-ethnic counterparts that likely have a native-sounding last name, representing a 1.4\% reduction in total years of education.
+This gap is larger for women (0.25 years, or 1.8\%) than for men (0.16 years, or 1.2\%).
+When I compare the labor market outcomes of HW and WH children, which captures the effect of having a Hispanic last name, HW children are 1 percentage point more likely to be unemployed than WH children, representing a 14.3\% increase from the mean unemployment rate.
+In terms of earnings, HW children earn 5 percentage points (4.9\%) less than WH children.
 Thus, by comparing inter-ethnic children, a person with a Hispanic last name makes 5 percentage points less than someone with a White last name.
-When I control for education, the last name effect decreases to a statistically insignificant 2 percentage points earnings gap.
+When I control for education, the last name effect decreases to a statistically insignificant 1 percentage point earnings gap.
+
+The findings reveal gaps that could be attributed to discrimination against individuals with Hispanic surnames, with effects varying across educational levels and gender.
+While the impact on total years of education appears modest (1.4\% reduction), the more substantial gaps in higher education attainment—particularly the 13\% reduction in bachelor's degree completion and 13.3\% reduction in associate degree attainment—suggest that discrimination may intensify at crucial educational transition points.
+This pattern is especially pronounced for women with Hispanic surnames, who face larger penalties in both total education (1.8\% reduction versus 1.2\% for men) and degree completion (14.8\% reduction in bachelor's degree attainment).
 
@@
 In an audit study of charter and traditional public schools, \textcite{bergman2018education,gaddis2024racial} find that students with Hispanic names---compared to students that are presumed White---are less likely to get a response to inquiries from schools.
-This could be an indication that people with Hispanic sounding names could end up lower access to education, including high value-added schools.
-Therefore, the earnings gap between people with Hispanic sounding last names and those with native sounding last names could be due to differences in access to education that are in turn due to discrimination.
+This could be an indication that people with a likely Hispanic sounding names could end up lower access to education, including high value-added schools.
+Therefore, the earnings gap between people with a likely Hispanic sounding last names and those with native sounding last names could be due to differences in access to education that are in turn due to discrimination.
 This paper provides further evidence, using observational data, that discrimination in access to education could lead to lower earnings for Hispanics.
+
+In the labor market, the 14.3\% higher unemployment rate and 4.9\% earnings penalty for those with Hispanic surnames indicate persistent discrimination even among individuals who share similar family backgrounds.
+Putting the results in context, \autocite{trejoWhyMexicanAmericans1997} find an earning gap between US-born Hispanics and US-born whites that is equal to 30.7 percentage points.
+My findings could be an indication that out of the 30.7 percentage points gap, at least 16\% of it could be attributed to discrimination against Hispanics.
+However, the fact that the earnings gap becomes statistically insignificant after controlling for education suggests that much of the labor market discrimination operates through educational channels rather than direct wage discrimination.
+These findings leave several important questions unanswered.
 
@@
 Education itself can be influenced by bias, potentially resulting in divergent outcomes \autocite{bergman2018education,gaddis2024racial}.
-This is especially the case when parental characteristics indicate that people with a Hispanic-sounding last-names should in theory complete more years of education---since mothers of inter-ethnic children with a Hispanic last-names have more years of education and earn more.
+This is especially the case when parental characteristics indicate that people with a likely Hispanic-sounding last-names should in theory complete more years of education---since mothers of inter-ethnic children with a Hispanic last-names have more years of education and earn more.
+While I observe where discrimination manifests (particularly in higher education access), my analysis cannot identify the specific mechanisms—such as teacher bias, institutional barriers, or social networks—through which surname-based discrimination operates.
 Consequently, further research is needed to comprehensively understand the earnings gaps between Hispanics and Whites.

### 2025-01-22 manuscripts/08-JERP-v6/submit_journal-main.tex
reviwer comments

@@
     \end{itemize}  
+    \item \hyperref[r&r:responses]{Responses to the editor and referees comments}
     \item The online appendix, which is separately paged. It includes an online appendix title page, appendix A, appendix B, appendix C, and references for citations in the appendix.
@@
         \begin{spacing}{\rrxspc}
+            I would like to thank the editor and the anonymous referees for their insightful comments, suggestions, and effort and time in reviewing this paper.
+I have addressed all the comments and suggestions in the revised manuscript.
+Below, I provide a summary of the changes made to the manuscript in response to the comments and suggestions.
         \end{spacing}
@@
     \section{Responses to Referee One}
+        I would like to thank referee one for the insightful comments and suggestions. Below is a detailed response to the comments and suggestions.
     \begin{spacing}{\rrquote}
@@
     \textbf{R1: } 1.
+My first concern relates to the threats to identification.
+It is feasible that children born to HW couples may differ systematically to children born to WH couples, particularly with respect to unobservable characteristics important for labor market outcomes.
+For example, HW pairs may exhibit differing parental characteristics related to parenting styles, parental preferences for education, gender norms and beliefs.
+The paper would be improved by including an expanded discussion around the threat to identification.
+In light of these issues, the paper would be improved by providing a more thorough discussion of the relative advantages of the current approach, compared to the conventional Oaxaca-Blinder decomposition.
     \end{quotation}
@@
     \begin{spacing}{\rrxspc}
+        I included an extensive discussion of the differences between HW and WH couples.
+These differences, that includes cultural differences, in couple of sections throughout the manuscript.
+While I acknowledge the concern, the estimation strategy rest of the fact that selection in the marriage market decreases the differences between HW and WH couples.
+This is the reason why comparing children of intermarried couples provides a better comparison than comparing children of endogamous marriages, i.e.
+marriages where both parents are either non-Hispanic White or Hispanic.
+I also included a discussion of the relative advantages of the current approach compared to the conventional Oaxaca-Blinder-Kitagawa decomposition.
+I argue that the current approach provides a more accurate estimate of gaps that are due to discrimination.
+In fact, the concern the reviewer raises is one of the reasons why my approach is preferred over the Oaxaca-Blinder-Kitagawa since children of endogamous marriages are more likely to have different characteristics than children of intermarried couples, including those that are unobservable.
+I argue, and show from the data, that children of intermarried couples are more likely to have similar characteristics than children of endogamous couples, see sections 3.1, and 4.
+    \end{spacing}
+
+    \begin{spacing}{\rrxspc}
+    \begin{quotation}
+        \textbf{R1: } 2.
+The analysis sample is restricted to individuals who self-identify as ‘White’.
+It is likely that this self-identification of ethnicity is endogenous to labor market outcomes.
+Importantly, the unobservable determinants of racial identification are potentially correlated with labor market outcomes.
+The paper would be improved by including a more thorough discussion of this (sample) selection issue.
+Related to this is the depiction of parents born in the United States as ’White’.
+This seems particularly over-simplifying and ignores the long history of immigration.
+A more accurate description would be to refer to this group as US born.
+        \end{quotation}
+        \end{spacing}
+
+    \begin{spacing}{\rrxspc}
+        I thank the reviewer for this comment and concern over the endogeneity of self-identification and labor market outcomes.
+I added some discussion over why I chose those who self-identify as White.
+I argue that inclusion those that are not White could contaminate the estimate of bias against Hispanics due to the racial signal.
+Moreover, \autocite{hadah2024effect} shows that there exists strong correlation between bias against Hispanics and self-reported Hispanic identity.
+Their results show that by not including Hispanics that attrit in the estimation of racial and ethnic gaps, the estimate of bias against Hispanics could be overestimated in the most biased states---more on ethnic attrition of Hispanics could be found in \textcite{antmanEthnicAttritionObserved2016,antmanEthnicAttritionAssimilation2020}.
+Regarding the comment of oversimplifying and ignoring the long history of immigration in the US, the majority of immigrants from Spanish-speaking countries in the US are first- and second-generation (see \autocite{antman2020ethnic}).
+I also show that the probability that parents are second-generation+ immigrants from Spanish-speaking countries is very low using Census data from 1960 to 2000, see Table 5.
+    \end{spacing}
+
+    \begin{spacing}{\rrxspc}
+        \begin{quotation}
+            \textbf{R1: } 3.
+The discussion on the construction of the ‘synthetic parents’ is somewhat brief.
+The paper suggests that the potential parents are matched using the birth year of the child and the parent’s place of birth to the children’s information collected in the CPS sample.
+While not explicitly stated in the paper, I assume that this sample of synthetic parents is used to construct mean parental education and family income at the time of the birth of the child.
+The process for matching the potential parents to the children is somewhat aggregated and their is likely considerable heterogeneity in the mean educational attainment of the potential parents within the set of children with the same birth year with the same parent’s country of birth.
+The paper could be improved by exploring the possibility of improving the quality of the match of potential parents by including further characteristics to match the potential parents to their children.
+    \end{quotation}
+        \end{spacing}
+
+    \begin{spacing}{\rrxspc}
+            I appreciate this insightful suggestion.
+I have expanded the discussion and provided a detailed example of the 'synthetic parents' construction methodology.
+While additional matching dimensions would be valuable, the limitations of publicly available data constrain the potential matching criteria.
+Nevertheless, the current approach provides meaningful estimates while acknowledging these data constraints.
+    \end{spacing}
+
+    \begin{spacing}{\rrxspc}
+        \begin{quotation}
+            \textbf{R1: } 4.
+The description of the estimated model is somewhat brief.
+However, there are two main issues associated with statistical inference on the estimated parameters in model (1).
+First, there is a ‘generated regressor’ issue associated with using estimated group-level parental education.
+A failure to account for this sampling variation will lead to misleadingly small standard errors.
+Second, given this group structure for parental education, it seems reasonable to assume that the model errors are uncorrelated across clusters but correlated within (potential parents) clusters.
+It is well understood that failing to account for this ‘clustering’ problem can lead to misleadingly small standard errors, narrow confidence intervals, and low p-values.
+        \end{quotation}
+    \end{spacing}
+
+    \begin{spacing}{\rrxspc}
+        Thank you for these important methodological points.
+Regarding the generated regressor issue: The parental characteristics are included solely as controls and not for inference purposes.
+I do not interpret these coefficients, as my primary interest lies in $\beta_1$, which captures the gap between Hispanic and White last names.
+Since I am only using these synthetic characteristics as controls and not making inference about their coefficients, this should not affect the standard errors or p-values of my coefficient of interest $\beta_1$, which is not a generated regressor.
+    \end{spacing}
+
+    \begin{spacing}{\rrxspc}
+        \begin{quotation}
+            \textbf{R1: } 5.
+The paper would be improved through an expanded discussion of the impact of measurement error on the reported estimates.
+While there is a well understood result in the measurement error literature that measures of group mean parental education will provide estimates that are more robust to the presence of measurement error in individual level measures of parental education, there is still an issue with the non-random attrition of potential parents.
+Specifically, not all potential parents have children in the CPS sample, nor do all children in the CPS sample have parents in the sample of potential parents.
+Moreover, the non-random attrition of potential parents implies that mean parental education may be systematically higher or lower than the actual parental education of the children in the CPS sample.
+The paper would be improved by providing an expanded discussion of the likely impacts of measurement error on the reported estimates.
+        \end{quotation}
+    \end{spacing}
+
+    \begin{spacing}{\rrxspc}
+        Thank you for these important methodological points.
+Let me clarify that the synthetic parents are constructed using the birth year of the child.
+The place of birth of parents---mothers and fathers---are questions that are asked of all participants in the Current Population Survey (CPS) starting 1994.
+Therefore, I do not use the information of parents that are in the CPS sample at the time of the survey.
+Thus, the concern over the attrition of parents in the sample is not applicable to my analysis.
+Please see the data section for a discussion of the data used in the analysis.
+    \end{spacing}
+    \clearpage
+    \pagebreak
+
+    \section{Responses to Referee Two}
+    I would like to thank referee two for the insightful and constructive comments and suggestions. Below is a detailed response to the comments and suggestions.
+
+    \begin{spacing}{\rrquote}
+        \begin{quotation}
+        \textbf{R2: } 1.
+The paper needs to be more clearly motivated in the introduction.
+There are a lot of reasons why studying discrimination by race and ethnicity is important and the paper would benefit from clearly articulating this, including discussing the implications of this discrimination.
+You mention economic mobility and I think there is more you can say about this.
+I also think clearly stating the contributions of the paper earlier on would be important.
+        \end{quotation}
+        \end{spacing}
+
+        \begin{spacing}{\rrxspc}
+            Thank you for the important suggestion.
+I have expanded the introduction to better motivate the study.
+I have included a discussion of the implications of discrimination on economic mobility and the contributions of the paper.
+I have also included a discussion of the importance of studying discrimination
+    \end{spacing}
+
+    \begin{spacing}{\rrquote}
+        \begin{quotation}
+        \textbf{R2: } 2.
+I would also like to see a clearer theory and review of the literature on the topic.
+For example, why should we expect that discrimination will affect educational attainment?
+Hispanics are one of the fastest growing groups entering college, but the extent to which they complete an associate or BA degree may reflect discrimination that happens when students are in school that ultimately translates into differences in attainment.
+Why might we expect differences in employment and years of education?
+        Why might we expect differences by gender?
+        \end{quotation}
+        \end{spacing}
+
+        \begin{spacing}{\rrxspc}
+            I appreciate this suggestion. Here are my replies to the different comments.
+
+            First, regarding including more on why discrimination could affect educational attainment.
+I have expanded the part of the introduction that discusses differences in education and the channels in which discrimination can affect educational attainment.
+The literature on discrimination and education shows that discrimination can affect educational attainment of minorities through various channels, including differences in school quality and bias from teachers, administrators, counselors, etc.
+These biases could lead to differences in educational attainment by preventing students access to some schools, recommendation letters, or counselors restricting access to more advanced courses, etc.
+Consequently, these biases could lead to differences in educational attainment.
+For example, having access to advanced courses in high school or better recommendation letters could affect the likelihood of attending college.
+
+            Second, if minorities face discrimination in access to education and the labor market, then we would expect differences in employment and years of education.
+The literature shows that discrimination can affect labor market outcomes through various channels, including differences in access to jobs and wage differences.
+
+            Finally, to address the comment here, and in other places, on why we might expect differences between genders.
+I added a discussion of why we might expect differences between men and women in the results section.
+I believe that showing that there might be heterogeneity in gaps between men and women in educational outcomes to be an important contribution of the paper since the literature studying gaps in earnings mainly focuses on the average gap between men.
+Moreover, showing that couples with a Hispanic husband and a White wife do not invest differently in their children than couples with a White husband and a Hispanic wife could be a way to test for cultural differences between the two groups.
+        \end{spacing}
+
+    \begin{spacing}{\rrquote}
+        \begin{quotation}
+        \textbf{R2: } 3.
+Occupational segregation can also reflect discrimination in the labor market with important implications for economic mobility and other outcomes, and it would be important to mention this in the literature review.
+In general, the review of the literature needs to provide more details about the different studies and how your paper contributes to that literature above and beyond using a cleaner comparison.
+        \end{quotation}
+        \end{spacing}
+
+        \begin{spacing}{\rrxspc}
+            Thank you for this suggestion.
+I have included a discussion of occupational segregation in the literature review and added how my paper contributes to the different strands of the literature.
+    \end{spacing}
+
+    \begin{spacing}{\rrquote}
+        \begin{quotation}
+        \textbf{R2: } 4.
+There are places throughout the paper where additional references are needed.
+For example, when you state in the introduction on page 3 that discrimination can lead to lower wages, reduced opportunities, and hinder assimilation this statement needs references.
+        \end{quotation}
+        \end{spacing}
+
+        \begin{spacing}{\rrxspc}
+            Thank you for pointing this out. I have added references to such statements. 
+    \end{spacing}
+
+    \begin{spacing}{\rrquote}
+        \begin{quotation}
+        \textbf{R2: } 5.
+Can you explain why you discuss assimilation and in what ways this is connected to your theory given that you are focusing on U.S.
+born children?
+If this matters because you are focusing on U.S.
+born Hispanic children with one foreign-born parent, then you need to clearly state this.
+        \end{quotation}
+        \end{spacing}
+
+        \begin{spacing}{\rrxspc}
+            Thank you for the comment. I added a footnote to the introduction to address it. 
+    \end{spacing}
+
+    \begin{spacing}{\rrquote}
+        \begin{quotation}
+        \textbf{R2: } 6.
+You also need to state clearly and early in the paper that your study focuses on children with a foreign-born parent.
+Children with U.S.
+born Hispanic parents may be different than those with foreign-born parents and you could more explicitly discuss this in the paper.
+        \end{quotation}
+        \end{spacing}
+
+        \begin{spacing}{\rrxspc}
+            I appreciate this comment. I made it more clear throughout the paper that I focus on children with a foreign-born parent. 
+    \end{spacing}
+
+    \begin{spacing}{\rrquote}
+        \begin{quotation}
+        \textbf{R2: } 7.
+Also given the differences in the characteristics of immigrants from different Spanish-speaking countries living in the U.S.
+(differences in socioeconomic status, education, etc.), you probably want to mention that your study is not capturing this.
+Further, to the extent that men and women migrate from different countries and have different pre- and post-migration characteristics then this might affect your results.
+
+        One possible way to address this is to conduct sensitivity analyses limited to Children whose parents were likely born in Mexico or who respond themselves that they are Mexican.
+Given that immigrants from Mexico are the largest immigrant group from a Spanish speaking country in the U.S., this may offer an even cleaner comparison.
+There could still be differences in the characteristics of Mexican mothers and fathers, but you could potentially check this in the data.
+        \end{quotation}
+        \end{spacing}
+
+        \begin{spacing}{\rrxspc}
+            Thank you for this suggestion.
+I made it clear in the paper that I do not capture differences in the characteristics of immigrants from different Spanish-speaking countries.
+I also added some results and discussion as a sensitivity analysis breaking down the results for Hispanic children with Mexican parents versus Hispanic children with non-Mexican parents.
+    \end{spacing}
+
+    \begin{spacing}{\rrquote}
+        \begin{quotation}
+        \textbf{R2: } 8. There are several sections that are repeated in the paper and I would suggest you streamline the text. For example, you review results twice, but this isn’t necessary.
+
+        Similarly, I think you can more systematically organize the section describing your empirical strategy and clearly explaining your identification strategy, the concerns that it helps you overcome, your assumptions, and how you are testing whether these assumptions likely hold.
+Currently this is explained in multiple sections throughout the paper.
+        \end{quotation}
+        \end{spacing}
+
+        \begin{spacing}{\rrxspc}
+            I appreciate this comment. I have streamlined the text and reorganized the paper to address these concerns. 
+    \end{spacing}
+
+    \begin{spacing}{\rrquote}
+        \begin{quotation}
+        \textbf{R2: } 9. Can you provide an example that illustrates how you link individuals in the CPS to the synthetic parents?
+
+        \end{quotation}
+        \end{spacing}
+            I added an example to the data section that illustrates how I link individuals in the CPS to the synthetic parents.
+        \begin{spacing}{\rrxspc}
+
+    \end{spacing}
+
+    \begin{spacing}{\rrquote}
+        \begin{quotation}
+        \textbf{R2: } 10.
+I would like more details about the sample and the decisions you make.
+For example, you identify your sample as U.S.
+born children who identify as White in the CPS but some of these may also identify as Hispanic no?
+Do you restrict your sample to respondents who identify as White Hispanic or not?
+How many individuals in your sample identify as White non-Hispanic even if they have a parent born in a Spanish speaking country and how many do not?
+If someone has a Hispanic parent but do not identify as Hispanic those may be different than people with a Hispanic parent who identify as Hispanic.
+
+        If there are people who identify as White non Hispanic in your sample even though they have a parent born in a Spanish speaking country, can you do a sensitivity analysis removing them?
+
+        For example, on page 18 you say “I also find a significant earnings gap between those that identify as Hispanic.” Can you explain this statement? Is this related to the point I made above?
+
+        \end{quotation}
+        \end{spacing}
+
+        \begin{spacing}{\rrxspc}
+            Thank you for the comment. I added more description of the sample to the data section and the reasoning of why I chose those who self-identify as White. 
+    \end{spacing}
+
+    \begin{spacing}{\rrquote}
+        \begin{quotation}
+        \textbf{R2: } 11.
+You report results for men and women and, I am sorry if I missed it, but I would like to see this motivated in the paper as there are many reasons for doing the analyses separately.
+See my earlier point on the lit review/theory.
+
+        \end{quotation}
+        \end{spacing}
+            Please see my response to the earlier comment on why we might expect differences between men and women.
+        \begin{spacing}{\rrxspc}
+
+    \end{spacing}
+
+    \begin{spacing}{\rrquote}
+        \begin{quotation}
+        \textbf{R2: } 12.
+It would also be important to discuss effect sizes.
+Some of the findings seem to be small and you should discuss whether they are economically meaningful.
+Further, there should be more discussion about what your findings mean to understand discrimination for Hispanics and how it plays out, including by gender and in what ways they fall short in answering this question.
+
+        \end{quotation}
+        \end{spacing}
+
+        \begin{spacing}{\rrxspc}
+            Thank you for the suggestion.
+I have included a discussion of the effect sizes and the economic meaning of the results.
+I have also included a discussion of what the findings mean for understanding discrimination against Hispanics and how it plays out.
+    \end{spacing}
+
+    \begin{spacing}{\rrquote}
+        \begin{quotation}
+        \textbf{R2: } 13. How are you measuring the different outcomes? I don’t believe you discuss in the paper. This may be obvious, but I think it is still important to mention for clarity.
+
+        \end{quotation}
+        \end{spacing}
+
+        \begin{spacing}{\rrxspc}
+            Thank you for the comment. I have included a discussion of how I measure the different outcomes in the data section.        
+    \end{spacing}
+
+    \begin{spacing}{\rrquote}
+        \begin{quotation}
+        \textbf{R2: } 14.
+When you control for education in models that examine earnings, the effect becomes statistically insignificant, what happens if you include industry or occupation fixed effects?
+Can you do that?
+Then you'd be comparing people within the same industry or occupation.
+
+        \end{quotation}
+        \end{spacing}
+
+        \begin{spacing}{\rrxspc}
+            I appreciate this comment. I have added a discussion of the results when I include occupation fixed effects to the sensitivity analysis section.
+    \end{spacing}
+
+    \begin{spacing}{\rrquote}
+        \begin{quotation}
+        \textbf{R2: } 15.
+The conclusion repeats much of what was said in the body of the paper, and I would like to see more discussion about the implications of your findings to understand disparities in education and labor outcomes between people with different ethnic background, and to what extent you can conclude these differences can be attributed to discrimination.
+It would also be important to compare your results with other literature on this topic.
+
+        \end{quotation}
+        \end{spacing}
+
+        \begin{spacing}{\rrxspc}
+            Thank you so much for this comment.
+I have expanded the conclusion to include more discussion about the implications of the findings and how they can be attributed to discrimination.
+I have also included a discussion of how my results compare to the literature on this topic.
+    \end{spacing}
+
+    \begin{spacing}{\rrquote}
+        \begin{quotation}
+        \textbf{R2: } 16. On page 14, this sentence seems to be incomplete: “Consequently, comparing WH and HW children to each other to analyze discrimination against Hispanics in the labor market.”
+
+        \end{quotation}
+        \end{spacing}
+
+        \begin{spacing}{\rrxspc}
+            Thank you for pointing this out. I have corrected this sentence.
+    \end{spacing}
+
+    \begin{spacing}{\rrquote}
+        \begin{quotation}
+        \textbf{R2: } 17.
+Throughout the paper you refer to people with a parent born in a Spanish speaking country as “children who have a Spanish-sounding last name” I would encourage you to soften this language and say “who likely have” a Spanish sounding last name.
+First, you do not know if a person actually does have the last name of their father.
+Second, while it is true that Spanish sounding last names are very common in Spanish-speaking Latin American countries and among Latinos in the U.S., this may not be true across the board and many people born in Spanish speaking countries, who are Spanish speakers and Hispanic, may not have a traditionally sounding Spanish last name.
+
+        \end{quotation}
+        \end{spacing}
+
+        \begin{spacing}{\rrxspc}
+            Thank you for your comment. I changed the mention of "children who have a Spanish-sounding last name" to "children who likely have a Spanish-sounding last name" throughout the paper.
+    \end{spacing}
+    \end{refsection}
+
+    \clearpage
+    \pagebreak
+
+    \section{Responses to Referee three}
+    I would like to thank referee three the comments. Below is a detailed response to the comments and suggestions.
+
+    \begin{spacing}{\rrquote}
+        \begin{quotation}
+        \textbf{R3: } 1.
+As the author states on page 4, the key identifying assumption is that people born to HW parents are similar to their WH peers.
+Only when this assumption is fulfilled can the author attribute the estimated coefficients to evidence of discrimination.
+However, Table 4 clearly shows that the differences between HW and WH synthetic parents’ characteristics, such as the father’s/mother’s education and total family income, are all significantly different.
+More specifically, HW families exhibit lower levels of education and income compared to their WH counterparts.
+Based on this, it is not clear whether the estimated coefficients are indeed evidence of discrimination or they merely reflect the fact that HW children grew up in an environment with less resources/educated parents.
+
+        The manuscript heavily adopts the methodology of Rubinstein and Brenner (2014), which examines the impacts of having a Sephardic-sounding surname on wages by comparing the Israeli-Jewish men born to Sephardic fathers and Ashkenazi mothers (SA) with those born to Ashkenazi fathers and Sephardic mothers (AS).
+They too find that the AS and SA families have statistically different educational and labor market outcomes.
+However, in their case, despite the fact that Sephardic Israelis face tougher labor market conditions, SA parents exhibit better education and labor market outcomes than AS parents.
+Therefore, any evidence indicating that SA offspring have worse labor market outcomes provides convincing (and potentially a lower bound) evidence of discrimination.
+
+        \end{quotation}
+        \end{spacing}
+
+        \begin{spacing}{\rrxspc}
+            Thank you for the comment.
+Even though the synthetic parents have different characteristics, the estimation strategy rests on the fact that selection in the marriage market decreases the differences between HW and WH couples.
+This is the reason why comparing children of intermarried couples provides a better comparison than comparing children of endogamous marriages, i.e.
+marriages where both parents are either non-Hispanic White or Hispanic.
+I also included a discussion of the relative advantages of the current approach compared to the conventional Oaxaca-Blinder-Kitagawa decomposition.
+I argue that the current approach provides a more accurate estimate of gaps that are due to discrimination.
+In fact, the concern the reviewer raises is one of the reasons why my approach is preferred over the Oaxaca-Blinder-Kitagawa since children of endogamous marriages are more likely to have different characteristics than children of intermarried couples, including those that are unobservable.
+I argue, and show from the data, that children of intermarried couples are more likely to have similar characteristics than children of endogamous couples.
     \end{spacing}
@@
 \pagebreak
-    \end{refsection}
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2025-01-24 manuscripts/08-JERP-v6/manuscript_main.tex
after proof reading

@@
 Hispanics constitute a large and growing portion of the population in the United States.
-As the number of Hispanics increases, determining whether ethnic discrimination affects their labor market outcomes becomes increasingly crucial \autocite{chettyUnitedStatesStill2014, chettyEffectsExposureBetter2016,chettyFadingAmericanDream2017}.
+As the number of Hispanics increases, determining whether ethnic discrimination affects their labor market outcomes becomes increasingly crucial.
 Specifically, it is essential to investigate the extent to which a person's Hispanic ethnicity may influence their employment prospects, wages, and career advancement.
-Understanding these labor market outcomes is critical, as they directly impact broader societal issues such as assimilation and economic mobility \autocite{chettyUnitedStatesStill2014,chettyEffectsExposureBetter2016}.
+Understanding these labor market outcomes is critical, as they directly impact broader societal issues such as assimilation and economic mobility \autocite{chettyUnitedStatesStill2014, chettyEffectsExposureBetter2016,chettyFadingAmericanDream2017}.
 These factors serve as key indicators of how successfully Hispanics can integrate into society and ascend the socioeconomic ladder.
@@
 These barriers may trap Hispanic families in cycles of lower socioeconomic status, as reduced earnings and employment opportunities limit their ability to invest in their children's human capital or build wealth through homeownership and savings.\footnote{See \textcite{bowles2002inheritance} for an analysis of the intergenerational transmission of economic status .} While previous research has documented earnings gaps between Hispanics and non-Hispanic Whites, this paper makes several important contributions.
-First, it develops a novel empirical strategy that better isolates the causal effect of Hispanic ethnicity on labor market outcomes by comparing children of inter-ethnic marriages.
+First, this paper develops a novel empirical strategy that better isolates the causal effect of Hispanic ethnicity on educational and labor market outcomes by comparing children of inter-ethnic marriages.
 This approach helps control for typically unobservable family background characteristics that might confound traditional analyses.
@@
 
-This study builds upon and improves the traditional Oaxaca-Blinder-Kitagawa decomposition approach often used in labor market discrimination studies \autocite{oaxaca1973male,blinder1973wage,kitagawa1955components}.
+This study builds upon and improves the traditional Oaxaca-Blinder-Kitagawa decomposition approach often used in labor market discrimination studies \autocite{kitagawa1955components, oaxaca1973male,blinder1973wage}.
 While these decomposition methods have provided valuable insights into earnings gaps between different groups, including Hispanics and non-Hispanic Whites \autocite{davilaChangesRelativeEarnings2008}, they are limited in their ability to account for unobserved differences between the groups.
@@
 
-This methodological improvement is particularly timely and relevant given the rapidly changing demographics of the United States.
+The methodological improvement in this paper is particularly timely and relevant given the rapidly changing demographics of the United States.
 The US population is growing increasingly diverse, with significant implications for labor market dynamics and potential discrimination.
@@
 Using this novel approach, I find significant effects of having Hispanic last names on both educational and labor market outcomes.
-The results show that individuals with Hispanic last names face substantial disadvantages in educational attainment and earnings compared to their counterparts with non-Hispanic names.
-Specifically, individuals with Hispanic last names complete 0.2 years of education less than those with non-Hispanic names, even when controlling for family background.
-Individuals with Hispanic last names are also 1 percentage point more likely to be unemployed  and earn 5 percentage points less than those with non-Hispanic names.
+The results show that individuals that likely have a Hispanic last names face substantial disadvantages in educational attainment and earnings compared to their counterparts with likely non-Hispanic names.
+Specifically, individuals that likely have Hispanic last names complete 0.2 years of education less than those with non-Hispanic names, even when controlling for family background.
+Individuals individuals that likely have Hispanic last names are also 1 percentage point more likely to be unemployed  and earn 5 percentage points less than those with non-Hispanic names.
 These findings suggest that ethnic discrimination continues to play a role in shaping economic opportunities for Hispanics in the United States, highlighting the need for targeted policies to address these persistent disparities.
@@
 
-This study utilizes a method developed by \textcite{rubinstein2014pride}.
+This study uses a method developed by \textcite{rubinstein2014pride}.
 I compare children from inter-ethnic marriages.
@@
 
-Moreover, when analyzing the effect of having a Hispanic last name on educational outcomes, I provide separate analysis of men and women.
+In addition, when analyzing the effect of having a Hispanic last name on educational outcomes, I provide separate analysis of men and women.
 Separating the results for men and women is motivated by evidence that gender and ethnic discrimination operate through distinct mechanisms.
@@
 Children from mixed ethnic backgrounds may appear physically similar to those from single ethnicity backgrounds, but the influence of family dynamics and upbringing, crucial in developing skills and personal characteristics, varies with the pattern of mixed ethnic marriages.
-Particularly in a society where Hispanic ancestry could be perceived negatively, it raises the question of what kind of White women would choose Hispanic men.
+Particularly in a society where Hispanic ancestry could be perceived negatively, it raises the question of the type of White women would choose Hispanic men.
 Furthermore, even if such unions were formed randomly, children from White-Hispanic homes might benefit from more favorable family conditions than those from Hispanic-White homes, considering Whites generally have stronger socio-economic backgrounds than Hispanics.
@@
 
-I use two datasets:  the Integrated Public Use Microdata Series (IPUMS) Current Population Survey (CPS), CPS' Annual Social and Economic (ASEC) supplement, and CPS' outgoing rotation \autocite{cps2019} and the 1960 to 2000 US censuses \autocite{acs2019}.
+I use three datasets:  the Integrated Public Use Microdata Series (IPUMS) Current Population Survey (CPS), CPS' Annual Social and Economic (ASEC) supplement, and CPS' outgoing rotation \autocite{cps2019}, and the 1960 to 2000 US censuses \autocite{acs2019}.
 
@@
 
-In this section, I present my empirical Approach.
+% In this section, I present my empirical Approach.
 The specification estimates the effect of having a Hispanic last name on educational outcomes and earnings.
@@
 This is a method developed by \textcite{rubinstein2014pride}.
+I restrict the sample to WH and HW groups.
 
-This approach allows me to control for several key factors that could influence labor market outcomes.
+% This approach allows me to control for several key factors that could influence labor market outcomes.
 Importantly, I control for parental education and income using my synthetic parents data, which addresses concerns about intergenerational transmission of human capital.
@@
 
-\subsection{Estimating the Effect of Having a Hispanic Last Name}
-
-In this section I restrict the sample to WH and HW groups.
 Let $Y_{ist}$ be the outcome of interest for person $i$ in state $s$ at time $t$.
@@
 $X_{ist}$ is a vector of controls that includes age and numbers of hours worked, $\gamma_{st}$ are state-year fixed effects (FE), and $\phi_{ist}$ represents the error term.\footnote{I also include parental characteristics of `synthetic' parents as controls in alternative specifications.
-These controls include income and education of `synthetic' parents.} The equation for this strategy is written as follows:
+These controls include income and education of `synthetic' parents.} The equation for this strategy is written as follows and the sample is restricted to I restrict the sample to WH and HW groups:
 \begin{equation} \label{eq:1a}
@@
 This approach allows me to isolate the effect of having a Hispanic last name on labor market outcomes, providing a more accurate estimate of the impact.
+The difference in means between Hispanics and non-Hispanic Whites could result from discrimination.
+It can also be caused by differences in innate abilities, skills, and parental investments.
+WH children will have a non-Hispanic last name, while HW children will have a Hispanic last name.
+This is a method developed by \textcite{rubinstein2014pride}.
 This is an improvement over traditional Oaxaca-Blinder-Kitagawa decomposition methods, which are limited in their ability to account for unobserved differences between groups by providing a more suitable comparison group to measure discrimination \autocite{oaxaca1973male,blinder1973wage,kitagawa1955components}.
@@
 The previous two threats to identification could be addressed by the selection of partners among inter-ethnic couples.
-The average White person exhibits better observable characteristics than the Average Hispanic.
+The average White person exhibits better observable characteristics than the average Hispanic.
 Consequently, if marriages were random, the White husband in an WH marriage would have better traits than wife, and the White wife in the HW marriage would have the stronger traits.

### 2025-01-24 manuscripts/08-JERP-v6/submit_journal-main.tex
after proof reading

@@
 \endgroup
-This document and the associated contents in the revise and resubmit version of manuscript 33481294-df2a-4cce-80ee-c20f2b9975ac for the Journal of Economics, Rac
-e, and Policy. I include:
+This document and the associated contents in the revise and resubmit version of manuscript 33481294-df2a-4cce-80ee-c20f2b9975ac for the Journal of Economics, Race, and Policy. I include:
 \begin{enumerate}
@@
         \textbf{R2: } 11.
+If there are people who identify as White non Hispanic in your sample even though they have a parent born in a Spanish speaking country, can you do a sensitivity analysis removing them?
+
+        \end{quotation}
+        \end{spacing}
+
+        \begin{spacing}{\rrxspc}
+            Thank you for the comment. I added more description of the sample to the data section and the reasoning of why I chose those who self-identify as White. 
+    \end{spacing}
+
+    \begin{spacing}{\rrquote}
+        \begin{quotation}
+        \textbf{R2: } 12.
+For example, on page 18 you say “I also find a significant earnings gap between those that identify as Hispanic.” Can you explain this statement?
+Is this related to the point I made above?
+
+        \end{quotation}
+        \end{spacing}
+
+        \begin{spacing}{\rrxspc}
+            Thank you for the comment. I added more description of the sample to the data section and the reasoning of why I chose those who self-identify as White. 
+    \end{spacing}
+
+    \begin{spacing}{\rrquote}
+        \begin{quotation}
+        \textbf{R2: } 13.
 You report results for men and women and, I am sorry if I missed it, but I would like to see this motivated in the paper as there are many reasons for doing the analyses separately.
@@
         \begin{quotation}
-        \textbf{R2: } 12.
+        \textbf{R2: } 14.
 It would also be important to discuss effect sizes.
@@
         \begin{quotation}
-        \textbf{R2: } 13. How are you measuring the different outcomes? I don’t believe you discuss in the paper. This may be obvious, but I think it is still important to mention for clarity.
+        \textbf{R2: } 15. How are you measuring the different outcomes? I don’t believe you discuss in the paper. This may be obvious, but I think it is still important to mention for clarity.
 
@@
         \begin{quotation}
-        \textbf{R2: } 14.
+        \textbf{R2: } 16.
 When you control for education in models that examine earnings, the effect becomes statistically insignificant, what happens if you include industry or occupation fixed effects?
@@
         \begin{quotation}
-        \textbf{R2: } 15.
+        \textbf{R2: } 17.
 The conclusion repeats much of what was said in the body of the paper, and I would like to see more discussion about the implications of your findings to understand disparities in education and labor outcomes between people with different ethnic background, and to what extent you can conclude these differences can be attributed to discrimination.
@@
         \begin{quotation}
-        \textbf{R2: } 16. On page 14, this sentence seems to be incomplete: “Consequently, comparing WH and HW children to each other to analyze discrimination against Hispanics in the labor market.”
+        \textbf{R2: } 18. On page 14, this sentence seems to be incomplete: “Consequently, comparing WH and HW children to each other to analyze discrimination against Hispanics in the labor market.”
 
@@
         \begin{quotation}
-        \textbf{R2: } 17.
+        \textbf{R2: } 19.
 Throughout the paper you refer to people with a parent born in a Spanish speaking country as “children who have a Spanish-sounding last name” I would encourage you to soften this language and say “who likely have” a Spanish sounding last name.

### 2025-03-17 manuscripts/08-JERP-v6/manuscript_main.tex
edits

@@
 
-A large literature documents substantial earnings gaps across race and ethnicity \autocite{bayer2018divergent, charles2008prejudice}.
-Hispanics constitute a large and growing portion of the population in the United States.
-As the number of Hispanics increases, determining whether ethnic discrimination affects their labor market outcomes becomes increasingly crucial.
-Specifically, it is essential to investigate the extent to which a person's Hispanic ethnicity may influence their employment prospects, wages, and career advancement.
-Understanding these labor market outcomes is critical, as they directly impact broader societal issues such as assimilation and economic mobility \autocite{chettyUnitedStatesStill2014, chettyEffectsExposureBetter2016,chettyFadingAmericanDream2017}.
-These factors serve as key indicators of how successfully Hispanics can integrate into society and ascend the socioeconomic ladder.
+A large body of literature provides evidence of substantial earnings gaps across race and ethnicity in the US \autocite{bayer2018divergent, charles2008prejudice}.
+Hispanic individuals constitute a large and growing portion of the US population, and as this number increases, it is increasingly crucial to determine whether ethnic discrimination affects their employment prospects, wages, and career advancement.
+Understanding these labor market outcomes is critical, as they directly affect broader societal issues like assimilation and economic mobility
+\autocite{chettyUnitedStatesStill2014, chettyEffectsExposureBetter2016,chettyFadingAmericanDream2017}.
+These factors serve as key indicators of how successfully Hispanic immigrants can navigate society and ascend the socioeconomic ladder.
 
@@
 Does having a Hispanic last name affect labor market outcomes?
-I aim to show that comparing Hispanic Whites to non-Hispanic Whites might create an artificially higher earnings gap since the two groups differ in many observable characteristics.\footnote{By identifying as Hispanic, I refer to individuals who self-report their ethnicity as Hispanic on surveys or other data collection instruments.} \footnote{Observable characteristics refer to factors that can be measured and quantified, such as education level, work experience, and immigration status.} My analysis focuses on US-born children with a foreign-born parent---i.e.
-inter-ethnic.\footnote{This study's focus on US-born children with foreign-born parents means the findings may not generalize to Hispanic children with US-born parents, who likely face different socioeconomic and cultural circumstances.
-Additionally, the analysis does not account for heterogeneity in immigrant characteristics across different Spanish-speaking countries of origin, such as variations in educational attainment, socioeconomic status, and gender-specific migration patterns, which could influence both parental selection into migration and subsequent child outcomes.
-} Others have attempted to compare how native-born White Hispanics fare to non-Hispanic Whites and foreign-born Hispanics.
-In \textcite{antman2020ethnic,antmanEthnicAttritionObserved2016,antmanEthnicAttritionObserved2016a,antmanEthnicAttritionAssimilation2020}, the authors compare the health and educational outcomes of Hispanic Whites to non-Hispanic Whites and native-born Hispanics to foreign-born Hispanics.
+I aim to show that comparing Hispanic White individuals to non-Hispanic White individuals might create an artificially higher earnings gap since the two groups differ in many observable characteristics.\footnote{By identifying as Hispanic, I refer to individuals who self-report their ethnicity as Hispanic on surveys or other data collection instruments.} \footnote{Observable characteristics refer to factors that can be measured and quantified, such as education level, work experience, and immigration status.} My analysis focuses on US-born children with foreign-born parent (i.e., inter-ethnic).\footnote{This study's focus on US-born children with foreign-born parents means the findings may not generalize to Hispanic children with US-born parents, who likely face different socioeconomic and cultural circumstances.
+Additionally, the analysis does not account for heterogeneity in immigrant characteristics across different Spanish-speaking countries of origin, such as variations in educational attainment, socioeconomic status, and gender-specific migration patterns, which could influence both parental selection into migration and subsequent child outcomes.} Others have attempted to compare how native-born White Hispanics fare compared to non-Hispanic Whites and foreign-born Hispanics.
+\cites{antman2020ethnic,antmanEthnicAttritionObserved2016,antmanEthnicAttritionObserved2016a,antmanEthnicAttritionAssimilation2020}, the authors compare the health and educational outcomes of Hispanic Whites to non-Hispanic Whites and native-born Hispanics to foreign-born Hispanics.
 They find gaps in education and health between Hispanics and Whites.

### 2025-03-17 manuscripts/08-JERP-v6/manuscript_title_etc.tex
after proof reading
edits

@@
 
-Do individuals with Hispanic names face labor market discrimination?
-This study analyzes the impact of Hispanic-sounding surnames on wages among inter-ethnic children with one White and one Hispanic parent.
-I find that individuals with Hispanic surnames often earn less, with a notable wage gap favoring those with White surnames.
-People born to Hispanic fathers and White mothers receive 0.2 years of education less than those born to White fathers and Hispanic mothers.
-Men born to Hispanic fathers and White mothers are 1 percentage point more likely to be unemployed, and earn 5 percentage points less than those born to White fathers and Hispanic mothers.
-The earnings gap is largely due to educational differences.
+Do individuals with Hispanic surnames face labor market discrimination?
+This study analyzes the impact of Hispanic-sounding surnames on wages among inter-ethnic Americans with one White and one Hispanic parent.
+I find that individuals that likely have Hispanic surnames (i.e., those with a Hispanic father) often earn less, with a notable wage gap favoring those with White-sounding surnames.
+People born to Hispanic fathers and White mothers receive 0.2 fewer years of education than men born to White fathers and Hispanic mothers.
+Men born to Hispanic fathers and White mothers are 1 percentage point more likely to be unemployed, and they earn 5 percentage points less than those born to White fathers and Hispanic mothers.
+I also show that this gap can largely be explained by educational differences.
 \PAPERJEL}

### 2025-03-19 manuscripts/08-JERP-v6/manuscript_main.tex
finish edits from CE

@@
 I aim to show that comparing Hispanic White individuals to non-Hispanic White individuals might create an artificially higher earnings gap since the two groups differ in many observable characteristics.\footnote{By identifying as Hispanic, I refer to individuals who self-report their ethnicity as Hispanic on surveys or other data collection instruments.} \footnote{Observable characteristics refer to factors that can be measured and quantified, such as education level, work experience, and immigration status.} My analysis focuses on US-born children with foreign-born parent (i.e., inter-ethnic).\footnote{This study's focus on US-born children with foreign-born parents means the findings may not generalize to Hispanic children with US-born parents, who likely face different socioeconomic and cultural circumstances.
-Additionally, the analysis does not account for heterogeneity in immigrant characteristics across different Spanish-speaking countries of origin, such as variations in educational attainment, socioeconomic status, and gender-specific migration patterns, which could influence both parental selection into migration and subsequent child outcomes.} Others have attempted to compare how native-born White Hispanics fare compared to non-Hispanic Whites and foreign-born Hispanics.
-\cites{antman2020ethnic,antmanEthnicAttritionObserved2016,antmanEthnicAttritionObserved2016a,antmanEthnicAttritionAssimilation2020}, the authors compare the health and educational outcomes of Hispanic Whites to non-Hispanic Whites and native-born Hispanics to foreign-born Hispanics.
+The analysis does not account for heterogeneity in immigrant characteristics across different Spanish-speaking countries of origin, such as variations in educational attainment, socioeconomic status, and gender-specific migration patterns, which could influence both parental selection into migration and subsequent child outcomes.} Others have attempted to compare how native-born White Hispanics fare compared to non-Hispanic Whites and foreign-born Hispanics.
+\cites{antman2020ethnic} have compared the health and educational outcomes of Hispanic Whites to non-Hispanic Whites and native-born Hispanics to foreign-born Hispanics.
 They find gaps in education and health between Hispanics and Whites.
-They also find that native-born Hispanics are more likely than their foreign-born counterparts to report poor health.
-\textcite{davilaChangesRelativeEarnings2008} documents many gaps in labor market outcomes between Hispanics and Whites.
-They attribute a big part of this gap to differences in education, experience, immigration status, and regional differences.
+They also find that native-born Hispanics are more likely than their foreign-born counterparts to report poor health.\footnote{For additional studies examining ethnic identity and outcomes among Hispanic populations, see \textcite{antman2020ethnic,antmanEthnicAttritionObserved2016,antmanEthnicAttritionObserved2016a,antmanEthnicAttritionAssimilation2020}, which collectively document the relationships between ethnic identification, assimilation patterns, and various socioeconomic outcomes.} \textcite{davilaChangesRelativeEarnings2008} found gaps in labor market outcomes between Hispanics and Whites, which they attribute largely to differences in education, experience, immigration status, and regional differences.
 This paper builds on those studies by emphasizing how cultural assimilation and generational status further shape educational and labor market disparities for Hispanics.
@@
 Labor market discrimination can create persistent intergenerational disadvantages by limiting economic mobility, reducing access to quality healthcare and education, and constraining residential choices \autocite{chettyUnitedStatesStill2014, chettyEffectsExposureBetter2016,chettyFadingAmericanDream2017,bowles2002inheritance, djajic2003assimilation}.
-Additionally, discrimination may foster occupational sorting and segregation, as shown by recent task-based models of racial wage gaps that identify race-specific barriers in “Contact” tasks and explain the stagnation of these gaps post-1980 \autocite{hurst2024task}.
-These barriers may trap Hispanic families in cycles of lower socioeconomic status, as reduced earnings and employment opportunities limit their ability to invest in their children's human capital or build wealth through homeownership and savings.\footnote{See \textcite{bowles2002inheritance} for an analysis of the intergenerational transmission of economic status .} While previous research has documented earnings gaps between Hispanics and non-Hispanic Whites, this paper makes several important contributions.
-First, this paper develops a novel empirical strategy that better isolates the causal effect of Hispanic ethnicity on educational and labor market outcomes by comparing children of inter-ethnic marriages.
+Discrimination may foster occupational sorting and segregation, as shown by recent task-based models of racial wage gaps that identify race-specific barriers in “Contact” tasks and explain the stagnation of these gaps post-1980 \autocite{hurst2024task}.
+These barriers may trap Hispanic families in cycles of lower socioeconomic status, as reduced earnings and employment opportunities limit their ability to invest in their children's human capital or build wealth through homeownership and savings.\footnote{See \textcite{bowles2002inheritance} for an analysis of the intergenerational transmission of economic status.} Although previous research has identified earnings gaps between Hispanic and non-Hispanic Whites, this paper makes several important contributions.
+First, it develops a novel empirical strategy that better isolates the causal effect of Hispanic ethnicity on educational and labor market outcomes by comparing children from inter-ethnic families.
 This approach helps control for typically unobservable family background characteristics that might confound traditional analyses.
-Second, it provides new evidence on the specific role of Hispanic surnames in driving discrimination, offering insights into how ethnic signals influence both educational attainment and labor market outcomes.
-Finally, by examining both education and employment outcomes, this study helps illuminate the channels through which ethnic discrimination may perpetuate economic disparities across generations.
-Furthermore, it extends earlier findings by demonstrating how naming cues can compound structural inequalities and reinforce existing barriers faced by Hispanics.
+Second, it provides new evidence on the specific role of Hispanic surnames in driving discrimination, offering insights into how ethnic signals influence educational attainment and labor market outcomes.
+Finally, by examining both education and employment outcomes, this study helps illuminate the channels through which ethnic discrimination may perpetuate economic disparities across generations and extends earlier findings by demonstrating how naming cues can compound structural inequalities and reinforce existing barriers.
 
 This study builds upon and improves the traditional Oaxaca-Blinder-Kitagawa decomposition approach often used in labor market discrimination studies \autocite{kitagawa1955components, oaxaca1973male,blinder1973wage}.
-While these decomposition methods have provided valuable insights into earnings gaps between different groups, including Hispanics and non-Hispanic Whites \autocite{davilaChangesRelativeEarnings2008}, they are limited in their ability to account for unobserved differences between the groups.
-My approach, by comparing children of inter-ethnic marriages, provides a more closely matched comparison group, allowing me to better isolate the effect of having a Hispanic last name on labor market outcomes.
+While these decomposition methods have provided valuable insights into earnings gaps between groups, including Hispanics and non-Hispanic Whites \autocite{davilaChangesRelativeEarnings2008}, they are limited in their ability to account for unobserved differences between the groups.
+My approach comparing children of inter-ethnic marriages, provides a more closely matched comparison group, allowing me to better isolate the effect of having a Hispanic last name on labor market outcomes.
 
@@
 The US population is growing increasingly diverse, with significant implications for labor market dynamics and potential discrimination.
-The proportion of non-Whites has increased by more than 10 percentage points from 13 percent in 1995 to 23 percent in 2019.\footnote{The proportion of non-Whites and Hispanics is calculated using the Current Population Survey (CPS).} Native-born White Hispanic men earn 21\% less than White men, although a substantial portion of the earnings gap is due to educational differences between Hispanics and Whites \autocite{duncan2006hispanics, duncan2018identifying, duncan2018socioeconomic}.
-Some earnings differences may also be due to discrimination which will have negative consequences.
-For example, discrimination against Hispanics can lead to reduced job opportunities, lower wages, and hinder assimilation \autocite{chettyWhereLandOpportunity2014,bowles2002inheritance, djajic2003assimilation}.\footnote{This study focuses on U.S.-born children of inter-ethnic unions, many of whom have at least one foreign-born parent.
+The proportion of non-Whites has increased by more than 10 percentage points from 13 percent in 1995 to 23 percent in 2019.\footnote{The proportion of non-Whites and Hispanics is calculated using the Current Population Survey (CPS).} Native-born White Hispanic men earn 21\% less than White men, although there is evidence that a substantial portion of this gap is due to differences in education \autocite{duncan2006hispanics, duncan2018identifying, duncan2018socioeconomic}.
+Discrepancies may also be due to discrimination against Hispanics, which can lead to reduced job opportunities, lower wages, and poor assimilation \autocite{chettyWhereLandOpportunity2014,bowles2002inheritance, djajic2003assimilation}.\footnote{This study focuses on US-born children of inter-ethnic unions, many of whom have at least one foreign-born parent.
 Assimilation concerns remain relevant because foreign-born parents may pass on cultural norms and language preferences, which can still shape these children's educational and labor market outcomes.}
 
-Using this novel approach, I find significant effects of having Hispanic last names on both educational and labor market outcomes.
-The results show that individuals that likely have a Hispanic last names face substantial disadvantages in educational attainment and earnings compared to their counterparts with likely non-Hispanic names.
-Specifically, individuals that likely have Hispanic last names complete 0.2 years of education less than those with non-Hispanic names, even when controlling for family background.
-Individuals individuals that likely have Hispanic last names are also 1 percentage point more likely to be unemployed  and earn 5 percentage points less than those with non-Hispanic names.
+Using this novel approach, I find significant effects of Hispanic surnames on both educational and labor market outcomes, and that individuals that likely have a Hispanic last names face substantial disadvantages in educational attainment and earnings compared to their counterparts with likely non-Hispanic names.
+Specifically, individuals with presumably Hispanic last names complete 0.2 fewer years of education than those with non-Hispanic names, even when controlling for family background.
+Individuals with Hispanic last names are also 1 percentage point more likely to be unemployed and earn 5 percentage points less than those with non-Hispanic names.
 These findings suggest that ethnic discrimination continues to play a role in shaping economic opportunities for Hispanics in the United States, highlighting the need for targeted policies to address these persistent disparities.
 
-Identifying discrimination is difficult because of factors that affect labor market outcomes that are unobservable to economists---such as unobserved skills and separating it from prejudice and stereotypes.
+Factors that affect labor market outcomes, like skills and stereotypes, are unobservable to economists, so it is challenging to identify discrimination.
 One strategy used by researchers is audit or resume studies.
-\textcite{bertrand2004emily} conducted an audit study where identical resumes were sent to employers with White and Black-sounding names.
-This approach, however, has its drawbacks.
+\textcite{bertrand2004emily} conducted an audit study by sending employers identical resumes that differed only in the ethnic and racial signal of the applicant’s name (Black-sounding versus White-sounding).
+They found that resumes with Black-sounding names received substantially fewer callbacks than their White counterparts.
 Audit studies only observe callbacks, not wages.
+The  current study uses a method developed by \textcite{rubinstein2014pride}.
+The authors compared the children of mixed marriages between Sephardic and Ashkenazi people in Israel.\footnote{Ashkenazi and Sephardic are two distinct Jewish ethnic groups.} They found that workers with Sephardic last names earn substantially less than those with Ashkenazi last names.
+I compare children of Hispanic fathers and White mothers (henceforth HW) to children of White fathers and Hispanic mothers (WH).
+This approach accounts for the fact that  couples are likely to have similar income, schooling, and socioeconomic background, so these factors are largely controlled for \autocite{averettBetterWorseRelationship2008, averettEconomicRealityBeauty1996}.\footnote{For more on assortative mating see \autocite{beckerTheoryMarriagePart1973, beckerTheoryMarriagePart1974, beckerTreatiseFamily1993, browningCollectiveUnitaryModels2006, chiapporiFatterAttractionAnthropometric2012}} Children of HW and WH marriages have more similar observable characteristics than children of endogamous/homogamous marriages (i.e., White fathers-White mothers and Hispanic fathers-Hispanic mothers).
+In the US, children  from households with a Hispanic father and White mother  are overwhelmingly likely to  have their father’s Hispanic last name, allowing us to investigate how this ethnic signal affects annual log earnings.
 
-This study uses a method developed by \textcite{rubinstein2014pride}.
-I compare children from inter-ethnic marriages.
-More precisely, I compare children of Hispanic fathers and White mothers (henceforth HW) to children of White fathers and Hispanic mothers (henceforth  WH).
-This approach stems from the fact that there is a strong selection among many characteristics, and thus, marriages are not random.
-Couples match on several observable characteristics like income, schooling, socio-economic background, etc.
-\autocite{averettBetterWorseRelationship2008, averettEconomicRealityBeauty1996}.\footnote{For more on assortative mating see \autocite{beckerTheoryMarriagePart1973, beckerTheoryMarriagePart1974, beckerTreatiseFamily1993, browningCollectiveUnitaryModels2006, chiapporiFatterAttractionAnthropometric2012}} Children of HW and WH marriages have more similar observable characteristics than children of endogamous/homogamous marriages--- i.e., White fathers-White mothers and Hispanic fathers-Hispanic mothers.
-Moreover, children from a Hispanic father and White mother household will have a Hispanic last name from their fathers, enabling the investigation of how ethnic signals, such as having a Hispanic last name, affect annual log earnings.
+The choice to include separate results for men and women is motivated by evidence that discrimination based on sex/gender and ethnicity operate through distinct mechanisms.
+As \textcite{bertrand2020gender} discusses, women's labor market outcomes are shaped by persistent gender norms and stereotypes, particularly around motherhood and caregiving.
+\textcite{antecol2002relative} found that young Mexican women earned 9.5\% less than young white women in the US in 1994, and like the current study, differences in education appeared to be the most likely explanation for this gap.
+For Black women, differences in labor force attachment appear to be the key driver of wage disparities.
+The way discrimination manifests can also evolve differently by gender over time: \textcite{goldin2004making} found that women's professional identity and career continuity are increasingly important factors in labor market outcomes, while \textcite{darity2015tour} argue that a stratification economics approach helps reveal how group-based hierarchies are maintained through both discriminatory practices and intergenerational resource transfers.
+Disaggregating the results by gender provides important insights into how the intersection of ethnic and gender discrimination shapes educational and economic opportunities.
 
-In addition, when analyzing the effect of having a Hispanic last name on educational outcomes, I provide separate analysis of men and women.
-Separating the results for men and women is motivated by evidence that gender and ethnic discrimination operate through distinct mechanisms.
-As \textcite{bertrand2020gender} discusses, women's labor market outcomes are shaped by persistent gender norms and stereotypes, particularly around motherhood and caregiving.
-\textcite{antecol2002relative} found that young Mexican women earned 9.5\% less than young white women in 1994, with differences in education appearing to be the most important explanation for this gap.
-This contrasts with findings for black women, where differences in labor force attachment were the key driver of wage disparities.
-The way discrimination manifests can also evolve differently by gender over time--- \textcite{goldin2004making} document how women's professional identity and career continuity have become increasingly important factors in labor market outcomes, while \textcite{darity2015tour} argue that a stratification economics approach helps reveal how group-based hierarchies are maintained through both discriminatory practices and intergenerational resource transfers.
-Therefore, disaggregating the results by gender provides important insights into how the intersection of ethnic and gender discrimination shapes educational and economic opportunities.
+The main identifying assumption of my empirical strategy depends on the assumption that people born to HW parents are similar to their WH peer in all observable---and unobservable---aspects and characteristics relevant to the labor market, so surname is the only difference between these groups.
+Children from White-Hispanic homes might benefit from more favorable family conditions than those from Hispanic-White homes, considering Whites statistically have higher socioeconomic status than Hispanics.
+These factors introduce doubts regarding whether children from Hispanic-White families receive comparable familial support and influences, genetically or environmentally, as those from White-Hispanic families.
 
-The main identifying assumption of my empirical strategy depends on the assumption that people born to HW parents are similar to their WH peer in all observable---and unobservable---aspects and characteristics that are important in the labor market.
-Consequently, the only difference between the two groups is the variation in which group is more likely to have a Hispanic sounding last name.
-Children from mixed ethnic backgrounds may appear physically similar to those from single ethnicity backgrounds, but the influence of family dynamics and upbringing, crucial in developing skills and personal characteristics, varies with the pattern of mixed ethnic marriages.
-Particularly in a society where Hispanic ancestry could be perceived negatively, it raises the question of the type of White women would choose Hispanic men.
-Furthermore, even if such unions were formed randomly, children from White-Hispanic homes might benefit from more favorable family conditions than those from Hispanic-White homes, considering Whites generally have stronger socio-economic backgrounds than Hispanics.
-These factors introduce doubts regarding whether children from Hispanic-White families receive comparable familial support and influences, either genetically or environmentally, as those from White-Hispanic families.
+Previous studies have also used names as a proxy for race and ethnicity \autocite{fryer2004causes, bertrand2004emily}.
+\textcite{fryer2004causes} establish that names can be a predictor of a person's race.
+They found that having a Black-sounding name, after controlling for the home environment at birth, does not affect labor market outcomes.
+These early innovations in leveraging name-based identification highlight the importance of understanding how such signals can capture both conscious and unconscious bias, placing the current study within a broader framework of research on race, names, and economic outcomes.
 
-Previous studies have also used names as a proxy for race and ethnicity \autocite{fryer2004causes, rubinstein2014pride, bertrand2004emily}.
-\textcite{fryer2004causes} point out that names can be a predictor of a person's race.
-Specifically, they provide a rising pattern among Blacks having different names than Whites.
-They did, however, find that having a Black name, after controlling for the home environment at birth, does not affect their labor market outcomes.
-\textcite{rubinstein2014pride} compared the children of mixed marriages between Sephardic and Ashkenazi Jews in Israel.\footnote{Ashkenazi and Sephardic are two distinct Jewish ethnic groups.} They found that workers with Sephardic last names earn substantially less than those with Ashkenazi last names.
-\textcite{bertrand2004emily} conducted an audit study by sending employers identical resumes that differ in the ethnic and racial signal of a name (Black sounding name versus a White sounding one).
-They found that resumes with Black-sounding names received substantially fewer callbacks than their White counterparts.
-Furthermore, these early innovations in leveraging name-based identification highlight the importance of understanding how such signals can capture both conscious and unconscious bias, thereby situating this study within a broader framework of research on race, names, and economic outcomes.
-
-Moreover, audit studies in education economics investigated the effect of racial and ethnic signal on access to education.
-\textcite{bergman2018education} found that students with Hispanic sounding names received 2 percentage points fewer responses from schools than students with White sounding names.
+Audit studies in education economics investigated the effect of racial and ethnic signals on access to education.
+\textcite{bergman2018education} found that students with Hispanic-sounding names received 2 percentage points fewer responses from schools than students with White sounding names.
 \textcite{janssen2022guidance} found that guidance counselors restricted Asian students from advanced opportunities.
-\textcite{gaddis2024racial} found significant discrimination in interactions with principals against Hispanic and Chinese American families.
+\textcite{gaddis2024racial} found significant discrimination in interactions between school principals and Hispanic and Chinese American families.
 Finally, \textcite{bourabain2023school} found more evidence of discrimination in access to education against students with underprivileged backgrounds in the Flemish education system.
-Discrimination can affect educational attainment through multiple channels: restricted access to advanced coursework, biased academic counseling, limited encouragement to pursue higher education, and subtle institutional barriers that may cause Hispanic students to feel unwelcome or unsupported in educational settings.
-While Hispanic college enrollment has increased substantially in recent decades, discrimination within educational institutions may still impede degree completion through these various mechanisms.
-This paper contributes to this literature by providing evidence that discrimination in access to education could lead to lower earnings for Hispanics, and it further shows how parental background and marriage market selection can importantly shape these outcomes in ways not previously explored by earlier studies.
+Discrimination can affect educational access and success through multiple channels: restricted access to advanced coursework, biased academic counseling, limited encouragement to pursue higher education, and subtle institutional barriers that may cause Hispanic students to feel unwelcome or unsupported in educational settings.
+While Hispanic college enrollment has increased substantially in recent decades, discrimination within educational institutions may still impede degree completion.
+The current paper contributes to this literature by providing evidence that discrimination in access to education could lead to lower earnings for Hispanics, and how parental background and marriage market selection can shape these outcomes in ways not previously explored by earlier studies.
 
@@
 
-I use the CPS data set to study the effect of having a Hispanic last name on a person's labor market outcomes.
-I take advantage of the fact that the CPS asks parents' place of birth, ethnicity, and race.
-The data spans the period between 1994--- the earliest sample to ask about a parent's place of birth--- to 2019.
-Moreover, since the CPS does not provide data on parents' characteristics, essential to determine the family background, I use the census to construct synthetic parents.
-The census offers a larger sample of potential parents.
-Similar to the CPS, the 1960 to 2000 censuses ask about the person's place of birth and the individual's race and ethnicity.
-I employ this information to construct "synthetic parents" using a method developed by \textcite{rubinstein2014pride}.
-I construct the synthetic parents by linking husbands and wives in the census data to each other.
-I assume that parents have children between the ages of 25 and 40.
-I then link these synthetic parents using the birth year of child, and parents' places of birth to the year of birth of the "children" and the parents' places of birth in the CPS sample.
+I use the 1994–2019 CPS data set to study the effect of Hispanic surnames on labor market outcomes, because that span contains data on parents’ place of birth, ethnicity, and race.
+The CPS does not provide data on parents’ characteristics, essential to determine the family background, but the 1960–2000 US Census data includes parents’ place of birth, race, and ethnicity.
+I employ this information to construct "synthetic parents" using a method developed by \textcite{rubinstein2014pride} linking husbands and wives  in the census data.
+Assuming that parents have children 25–40 years old, I link these “parents” using parents’ places of birth, birth year of "children," and the parents’ places of birth.
 
-For my main analysis, I use different dependent variables.
-I construct different educational variables: years of education, whether the person has a high school diploma, whether the person has an associate degree, and whether the person has a bachelor degree.
-For labor market outcomes, I construct unemployment rate variable using the main CPS, log annual earnings using the ASEC, and log weekly earnings using the outgoing rotation.
-The unemployment indicator is a binary variable that equals 1 if someone is unemployed and 0 if employed, based on civilian labor force status.Log annual earnings is the log of total personal income from the CPS' ASEC supplement.
+For my main analysis, I construct different educational variables: years of education, and whether or not the person received a high school diploma, associate degree, and/or bachelor’s degree.
+For labor market outcomes, I construct an unemployment rate variable using the main CPS, log annual earnings using the ASEC, and log weekly earnings using the outgoing rotation.
+The unemployment indicator is a binary variable that equals 1 if someone is unemployed and 0 if employed, based on civilian labor force status.
+Log annual earnings is the log of total personal income from the CPS’ ASEC supplement.
 Log weekly earnings is the natural logarithm of weekly earnings for hourly workers from the outgoing rotation CPS supplement.
 I use the following controls: age, hours worked, and state-year fixed effects.
-I also include the parents' characteristics from the synthetic parents data as controls in alternative specifications.
+I also include the parents’ characteristics from the synthetic parents data as controls in alternative specifications.
 These controls include income and education of the synthetic parents.
@@
 
-I use the CPS for my primary analysis of the effect of having a Hispanic last name on earnings.
-I restrict my sample to Whites, United States-born citizens aged 25 to 40 and born between 1960 to 2000.
-Taking advantage of data on parents' place of birth, I divide the sample into four groups depending on their parent's ethnicity.
-Mothers or fathers are Hispanic if they were born in a Spanish-speaking country and Puerto Rico, and White if they were born in the United States.\footnote{The list of Spanish Speaking countries include: Argentina, Bolivia, Chile, Colombia, Costa Rica, Cuba, Dominican Republic, Ecuador, El Salvador, Guatemala, Equatorial Guinea, Honduras, Mexico, Nicaragua, Panama, Paraguay, Peru, Spain, Uruguay, Venezuela.} Therefore, an observation can be the product of four types of parents:
+I use the CPS for my primary analysis of the effect of a Hispanic surname on earnings.
+I restrict my sample to US citizens  aged 25–40 born between 1960 and 2000 in the US.
+Taking advantage of data on parents’ place of birth, I divide the sample into four groups depending on their parents’ ethnicity.
+Mothers or fathers are Hispanic if they were born in a Spanish-speaking country or Puerto Rico, and White if they were born in the United States.\footnote{The list of Spanish-speaking countries includes: Argentina, Bolivia, Chile, Colombia, Costa Rica, Cuba, Dominican Republic, Ecuador, El Salvador, Guatemala, Equatorial Guinea, Honduras, Mexico, Nicaragua, Panama, Paraguay, Peru, Spain, Uruguay, Venezuela.} Therefore, an observation can be the product of four types of parents:
 \begin{enumerate}
@@
 
-Moreover, my sample includes inter-ethnic children with Hispanic ancestry aged 25 to 40 that are citizens.
-I do not restrict the sample to the self-reported Hispanic identity and keep both Hispanic White and non-Hispanic White children because ethnic attrition could lead to biases estimates when measuring gaps as shown by \textcite{hadah2024effect}.
-\footnote{A person self-report Hispanic identity by answering the following Census question: ``Is this person Spanish/Hispanic/Latino?''} \footnote{Ethnic attrition happens when a U.S.-born descendant of a Hispanic immigrant fails to self-identify as Hispanic.
-For more discussion of this phenomenon see \textcite{antmanEthnicAttritionObserved2016,antmanEthnicAttritionAssimilation2020}} To maintain analytical clarity, I restricted the sample to Hispanic and non-Hispanic White individuals, excluding other racial groups who also identify as Hispanic, as this allows for a more focused examination of Hispanic ancestry effects without introducing confounding racial factors.
+My sample includes inter-ethnic children with Hispanic ancestry aged 25–40 who are US citizens.
+I include both Hispanic White and non-Hispanic White individuals, since ethnic attrition could lead to biased estimates when measuring gaps as shown  by \textcite{hadah2024effect}.
+\footnote{A person self-reports Hispanic identity by answering the Census question: ``Is this person Spanish/Hispanic/Latino?''} \footnote{Ethnic attrition happens when a US-born descendant of a Hispanic immigrant fails to self-identify as Hispanic.
+For more discussion of this phenomenon see \textcite{antmanEthnicAttritionObserved2016,antmanEthnicAttritionAssimilation2020}} To maintain analytical clarity, I restricted the sample to Hispanic and non-Hispanic White individuals, excluding other racial groups who also identify as Hispanic, to avoid confounding racial factors.
 
-The distribution of the four types of children is presented in Table \ref{tab:mat1}.
-The majority of the sample (96\%) is WW children.
+The CPS data distribution of the four types of parenthood is presented in Table \ref{tab:mat1}.
+The vast majority (96\%) are WW children.
 The second biggest group is HH, which constitutes 3\% of the sample.
@@
 Even though the census asks a person whether they are Hispanic or not, I took advantage of the questions on place of birth to create a proxy for ethnicity.
-I consider a Hispanic person as White and born in a Spanish-speaking country.
-Consequently, Whites in the sample are people who are White and native-born.
+I consider a Hispanic persons that self-report White as their racial group and are born in a Spanish-speaking country.
+Consequently, Whites in the sample are White and native-born.
 Using the information provided in the census, I can link husbands and wives with each other.
-I assume that parents have children between the ages of 25 and 40.
-Therefore, my sample consists of married White men and women with children that were born in the 1920 to 1975 cohorts\footnote{The construction of ``synthetic parents'' follows the method used by \textcite{rubinstein2014pride}.}.
+I assume that parents have children between the ages of 25 and 40, so my sample consists of married White men and women with children born in the 1920 to 1975 cohorts\footnote{The construction of ``synthetic parents'' follows the method used by \textcite{rubinstein2014pride}.}.
 
@@
 These ``synthetic parents'' are created by matching the characteristics of Mexican-American couples who had a 6-year-old child in 1970.
-The ``synthetic mother'' would have the average education level, income, and other socioeconomic characteristics of Mexican-born women who were married to U.S.-born men and had children around 1964, when they were between 20 and 35 years old (meaning they were born between 1929-1944).
-Similarly, the ``synthetic father'' would have the average characteristics of U.S.-born men who were married to Mexican-born women and had children in that same year.
+The ``synthetic mother'' would have the average education level, income, and other socioeconomic characteristics of Mexican-born women who were married to US-born men and had children around 1964, when they were between 20 and 35 years old (meaning they were born between 1929 and 1944).
+The ``synthetic father'' would have the average characteristics of US-born men who were married to Mexican-born women and had children in that same year.
 This method preserves the demographic composition of parents while assigning them historically accurate socioeconomic characteristics based on their origins and marriage patterns.
@@
 White husbands and White wives (WW) make up the majority of couples in the sample, 96\% (5,141,737 couples).
-Hispanic husbands and wives (HH) are the second largest group making up 2\% (119,749 couples) of all couples.
-White husbands and Hispanic wives (WH) couples are less 1\% (33,097 couples) of the sample and Hispanic husbands and White wives (HW) are also 1\% (37,847 couples).
+Hispanic husbands and wives (HH) are the second largest group, coprising 2\% (119,749 couples).
+White husbands and Hispanic wives (WH) $< 1\%$ (33,097 couples) of the sample and Hispanic husbands and White wives (HW) are around 1\% (37,847 couples).
 I present the summary statistics of the parents in Table \ref{tab:synth}.
@@
 
-% In this section, I present my empirical Approach.
-The specification estimates the effect of having a Hispanic last name on educational outcomes and earnings.
-The difference in means between Hispanics and non-Hispanic Whites could result from discrimination.
-It can also be caused by differences in innate abilities, skills, and parental investments.
-While controlling for observable skill measures, I compare children of inter-ethnic marriages, HW and WH.
-WH and HW children are more similar in characteristics but provide employers, and the labor market, with different signals.\footnote{WH and HW children are both half White, half Hispanic.} WH children will have a non-Hispanic last name, while HW children will have a Hispanic last name.
-This is a method developed by \textcite{rubinstein2014pride}.
-I restrict the sample to WH and HW groups.
-
-% This approach allows me to control for several key factors that could influence labor market outcomes.
-Importantly, I control for parental education and income using my synthetic parents data, which addresses concerns about intergenerational transmission of human capital.
-By comparing children from inter-ethnic marriages, I also implicitly control for many unobservable family characteristics that might affect labor market outcomes, as these families are likely to be more similar to each other than to families with two Hispanic or two non-Hispanic White parents.
-
 Let $Y_{ist}$ be the outcome of interest for person $i$ in state $s$ at time $t$.
-$HW_{ist}$ is an indicator variables for the type of parents of person $i$ has.
+$HW_{ist}$ is an indicator variable for the type of parents of person $i$.
 $X_{ist}$ is a vector of controls that includes age and numbers of hours worked, $\gamma_{st}$ are state-year fixed effects (FE), and $\phi_{ist}$ represents the error term.\footnote{I also include parental characteristics of `synthetic' parents as controls in alternative specifications.
@@
 $\beta_{1}$ is the coefficient of interest in this specification.
-$\beta_{1}$ represents the gaps in outcomes between children of inter-ethnic marriages who likely have a Spanish-sounding last name versus a White last name.
+$\beta_{1}$ represents the gaps in outcomes between children of inter-ethnic marriages who likely have a Hispanic-sounding last name versus a White-sounding last name.
 If $\beta_{1} > 0$, then people with a Hispanic last name have better outcomes than people with a White last name.
@@
 
-Moreover, by comparing interethnic children, I can control for many unobservable family characteristics that might affect labor market outcomes.
-These families are likely to be more similar to each other than to families with two Hispanic or two non-Hispanic White parents.
-This approach allows me to isolate the effect of having a Hispanic last name on labor market outcomes, providing a more accurate estimate of the impact.
-The difference in means between Hispanics and non-Hispanic Whites could result from discrimination.
-It can also be caused by differences in innate abilities, skills, and parental investments.
+By comparing interethnic children, I can control for many unobservable family characteristics that may affect labor market outcomes, since these families are likely to be more similar to each other than to families with two Hispanic or two non-Hispanic White parents.
+This approach allows me to isolate the effect of surname, providing a more accurate estimate of its impact.
+The difference in means between Hispanic and non-Hispanic Whites could result from discrimination, but it could also be due to differences in innate abilities, skills, and parental investments .
 WH children will have a non-Hispanic last name, while HW children will have a Hispanic last name.
-This is a method developed by \textcite{rubinstein2014pride}.
-This is an improvement over traditional Oaxaca-Blinder-Kitagawa decomposition methods, which are limited in their ability to account for unobserved differences between groups by providing a more suitable comparison group to measure discrimination \autocite{oaxaca1973male,blinder1973wage,kitagawa1955components}.
+This method, developed by \textcite{rubinstein2014pride}.
+This is an improves traditional Oaxaca-Blinder-Kitagawa decomposition methods, which are limited in their ability to account for unobserved differences between groups, by providing a more suitable comparison group to measure discrimination \autocite{oaxaca1973male,blinder1973wage,kitagawa1955components}.
 
-The central assumption underpinning my estimation strategy is based on the hypothesis that individuals born to a HW person exhibits comparable characteristics to their peers of WH descent, especially in areas vital to the labor market.
-This assumption includes similarities in educational background, skill sets, and work experiences that are significant determinants in employment opportunities, salary levels, and career advancement.
+The central assumption underpinning my estimation strategy is based on the hypothesis that individuals born to a HW person exhibit comparable characteristics to their peers of WH descent, especially in areas like educational background, skill sets, and work experiences that are significant determinants in employment opportunities, salary levels, and career advancement.
 
-Two primary reasons underscore that this assumption.
+Two primary reasons underscore this assumption.
 First, there is a significant selection in the marriage market.
-Since belonging to the Hispanic out-group comes with a negative penalty in society, one might wonder who are the White women that would be willing to give their kids a Hispanic last-name that would potentially negatively influence their futures.
+Since belonging to the Hispanic out-group brings negative societal effects, it is worth exploring factors common between White women that have children with Hispanic men.
 Second, fathers and mothers influence human capital accumulation differently \autocite{kimball2009risk,magruder2010intergenerational}.
-If marriage was random, then WH households might be a better environment for children that HA.
+If marriage were random, then WH households might be a better environment for children.
 Using the CPS and the US Census, I will evaluate these empirical concerns.
 
-The previous two threats to identification could be addressed by the selection of partners among inter-ethnic couples.
-The average White person exhibits better observable characteristics than the average Hispanic.
-Consequently, if marriages were random, the White husband in an WH marriage would have better traits than wife, and the White wife in the HW marriage would have the stronger traits.
+% These issues could be addressed by selecting partners among interethnic couples.
+The average White person exhibits better observable characteristics  than the average Hispanic person.
+If marriages were random, the White husband in a WH marriage would have better traits  than his wife, and the White wife in the HW marriage would have the stronger traits.
 Random matching in marriages would pose a challenge to the identification strategy.
 This is because parents are not perfect substitutes in the inter-generational transmission of human capital.
-Consequently, differences in parental background could significantly affect children's labor market outcomes, potentially invalidating the comparison between HW and WH children.
+Consequently, differences in parental background could significantly affect children’s labor market outcomes, potentially invalidating the comparison between HW and WH children.
 
-Marriages, however, are not random.
-There is large body of empirical and theoretical work that indicates that marriages exhibits strong selection, or what is referred to as assortative matching, on traits \autocite{averettBetterWorseRelationship2008, averettEconomicRealityBeauty1996, beckerTheoryMarriagePart1973, beckerTheoryMarriagePart1974, beckerTreatiseFamily1993, browningCollectiveUnitaryModels2006, chiapporiFatterAttractionAnthropometric2012}.
-Also, \textcite{duncanIntermarriageIntergenerationalTransmission2011} shows a similar pattern of assortative matching among Mexicans in the United States.
-These papers suggest that inter-ethnic WH and HW marriages will, on average, consist of partners sharing similar characteristics.
-Consequently, the literature would predict that both WH and HW parents will be similar in all aspects that are relevant to a child's labor market outcomes.
-I will go over the empirical evidence from the data to show that this is actually the case.
+A large body of empirical and theoretical work indicates that marriages exhibit strong selection, or "assortative matching." \autocite{averettBetterWorseRelationship2008, averettEconomicRealityBeauty1996, beckerTheoryMarriagePart1973, beckerTheoryMarriagePart1974, beckerTreatiseFamily1993, browningCollectiveUnitaryModels2006, chiapporiFatterAttractionAnthropometric2012}.
+Also, \textcite{duncanIntermarriageIntergenerationalTransmission2011} show a similar pattern of assortative matching among Mexicans in the United States.
+Interethnic WH and HW marriages generally consist of partners sharing similar characteristics , suggesting both WH and HW parents are similar in aspects relevant to labor market outcomes.
 
-Another threat to identification could arise from measurement error in using the place of birth of parents in the CPS data as a proxy for a parent's ethnicity and last name.
-The CPS only asks about the place of birth of parents and not their ethnic or racial identity.
-Therefore, it is possible that a native-born father could be a second-generation or later immigrant from a Spanish-speaking country.
-However, this scenario is highly improbable.
-Most Hispanics from 1960 to 2000 were first-generation immigrants, and the number of second-generation or later was very small.
-This is supported by data from the 1960 to 2000 censuses (see Table \ref{tab:mat3}).
-Only 3\% of native-born Americans identified as Hispanic, indicating the small number of second-generation or later Hispanic immigrants in the sample.
-Therefore, the probability that an inter-ethnic child with a native-born father that is also a second generation+ Hispanic immigrant is very small.
+Another threat to identifying a true effect could arise from measurement error in using parents’ place of birth from the CPS data as a proxy for a parent’s ethnicity and last name: The CPS notes place of birth but not ethnic or racial identity.
+It is possible that a native-born father could be a second-generation or later immigrant from a Spanish-speaking country, but this is unlikely.
+Most Hispanics from 1960 to 2000  were first-generation immigrants, and the number of second-generation or later was very small (Table \ref{tab:mat3}).\footnote{Based on the author's calculations using US Census data.} Only 3\% of native-born Americans identified as Hispanic, making it unlikely that an interethnic child with a native-born father is also a second generation Hispanic immigrant.
 
@@
 
-In this section, I explore the empirical data to affirm the validity of my empirical strategy, as presented in Table \ref{tab:synth}, which details the educational and economic profiles of parents from four different ethnic groupings—White White (WW), White Hispanic (WH), Hispanic White (HW), and Hispanic Hispanic (HH).
-This table meticulously lays out the average outcomes and discrepancies for each group, highlighting the impact of inter-ethnic marriages on children's prospects.
-These results show that there is selection in marriage, and that the differences between children born to HW and children born to WH are less severe that those between the other two groups.
-Consequently, comparing WH and HW children to each other to analyze discrimination against Hispanics in the labor market provides a better comparison since any gaps that emerge between these otherwise similar groups are more likely to be attributable to discrimination rather than underlying differences in family background or socioeconomic status.
+In this section, I explore the empirical data to affirm the validity of my empirical strategy.
+Table \ref{tab:synth} details the educational and economic profiles of parents from four different ethnic groups—White White (WW), White Hispanic (WH), Hispanic White (HW), and Hispanic Hispanic (HH), revealing the average outcomes and discrepancies for each group and highlighting the impact of interethnic marriages on their children’s prospects.
+These results show that there is selection in marriage, and that the differences between children born to HW and children born to WH are less extreme than those between the other two groups.
+In comparing WH and HW individuals to each other, any gaps that emerge between these otherwise similar groups are more likely to be attributable to discrimination rather than underlying differences in family background or socioeconomic status.
 
-Using the ``synthetic'' parents sample that I constructed using the Census data, I examine the family background of the different type of children.
-WW couples have higher education, 12.58 years for husbands and 12.36 for wives.
-As a household, WW couples have 24.95 years of schooling.
-Husbands in HH marriages have 8.64 years of education, while women have 8.49 years of schooling.
+Using the “synthetic” parents I constructed using the Census data, I examine the family background of the different types of children.
+WW couples have the highest level of education among the four groups: 12.58 years  for husbands and 12.36 for wives, so a WW household has 24.95 years of schooling.
+Men in HH marriages have 8.64 years of education, while women have 8.49.
 As a household, HH couples have 17.13 years of education.
-As predicted, the inter-ethnic couples exhibit matching in which partners marry people similar to them.
-WH husbands have 11.82 years of education, while wives have an average of 10.71 years.
-WH household attained a total of 22.68 years of schooling in total.
-HW husbands have 10.33 years of education, while wives have an average of 11.01 years.
-HW household attained a total of 21.50 years of schooling in total.
-Both Hispanic men and women in inter-ethnic couples marry white spouses that are, on average, more educated the the average HH couple.
-More importantly for human capital accumulation, the mother in the HW marriages---the children likely have a Spanish sounding last-name---are more educated than their WH peers.
-Since mothers could be more important in the a child's human capital accumulation and education, this would suggest that a child with a Hispanic last-name should complete more years of education \autocite{gould2020does}.
+As predicted, interethnic couples marry people similar to them.
+WH husbands have 11.82 years of education, while wives have an average of 10.71, a household total of 22.68 years.
+HW husbands have 10.33 years, while wives have an average of 11.01, giving HW households a total of 21.50 years.
+Both  Hispanic men and women in interethnic couples marry white spouses that are, on average, more educated than members of the average HH couple.
+More importantly for human capital accumulation, the mother in the HW marriages (those with children most likely to have a Hispanic-sounding last name—are more educated than their WH peers.) Since mothers could be more important in the child’s human capital accumulation and education, this would suggest that a child with a Hispanic last name should complete more years of education  \autocite{gould2020does}.
 
-The data reveals that WW couples have the highest total household education, amounting to 24.95 years, significantly surpassing the 17.69 years of HH couples.
-This stark contrast underlines the substantial educational divide between these groups.
-Inter-ethnic couples, WH and HW, display intermediate educational achievements, with WH households totaling 22.68 years of education and HW households slightly behind at 21.50 years.
+The data reveals that WW couples have the highest total household education, amounting to 24.95 years, significantly surpassing the 17.69 years of HH couples and illustrating the substantial educational divide between the groups.
+Interethnic couples (WH and HW) have intermediate education: WH households total 22.68 years and HW households are close behind at 21.50 years.
 Notably, HW husbands are less educated (10.33 years) compared to WH husbands (11.82 years), yet HW wives surpass their WH counterparts with 11.01 years of education, suggesting a balance in educational attainment within these marriages.
-This factor is particularly pertinent for children with Hispanic last names who might derive greater benefits from their mother's higher education.
+This factor is particularly pertinent for children with Hispanic last names who might derive greater benefits from their mother’s higher education.
 
-In terms of labor market performance, WW households boast the highest log total family income at 10.75, while HH households fall at the lower end with 10.42.
-Among inter-ethnic couples, WH households have a slightly higher total income of 10.65 compared to HW households at 10.60.
-Notably, the difference in husbands' log hourly earnings between HW and WH is marginal, indicating HW men earn 4\% less than their WH counterparts.
-Furthermore, HW women actually surpass WH women in earnings, with HW wives earning 1.75 and WH wives earning 1.73 in log hourly earnings, respectively.
-This reversal in the typical earning pattern not only speaks to the closing economic disparities between these groups but also implies potentially greater economic contributions from HW women to their families, which could be advantageous for children and to the developments of traits that are important in the labor market.
+In terms of labor market performance, WW households boast the highest log total family income at 10.75 , while HH households fall at the lower end with 10.42.
+Among interethnic couples, WH households have a slightly higher total income of 10.65 compared to HW households at 10.60.
+The difference in husbands’ log hourly earnings between HW and WH is marginal, with HW men earning 4\% less than their WH counterparts.
+HW women surpass WH women in earnings, with HW wives earning 1.75 and WH wives earning 1.73 in log hourly earnings, respectively.
+This reversal in the typical earning pattern speaks to the closing economic disparities between these groups and implies potentially greater economic contributions from HW women to their families, which could benefit their children when they enter the labor market.
 
-The table also sheds light on fertility trends, noting that WW couples have less children than HH couples, reflecting broader socio-economic and cultural patterns.
-The difference in fertility between HW and WH couples is positive but significantly lower than the difference in fertility between HH and WW couples.
+The table also reveals that WW couples have fewer children than HH couples, reflecting broader socioeconomic and cultural patterns.
+The difference in number of children between HW and WH couples is positive but significantly lower than the difference between HH and WW couples.
 
-The evidence presented in this section supports a robust empirical strategy, revealing significant selection in marriage and distinctive educational and income patterns among different types of families.
-Inter-ethnic couples exhibit comparable levels of education and earnings.
-This suggests that the children of HW families are likely positioned for better educational outcomes, critical in understanding discrimination in the labor market.
-Notably, the disparities between HW and WH families are considerably less pronounced than those between other groups, emphasizing the importance of comparing these children directly.
-Such comparisons shed light on the nuanced dynamics of ethnicity, education, and economic outcomes in inter-ethnic marriages.
-\footnote{I present in Tables \ref{tab:synthmex} and \ref{tab:syntnonmex} the summary statistics which detail the educational and economic profiles of parents from four different ethnic groupings—White White (WW), White Hispanic (WH), Hispanic White (HW), and Hispanic Hispanic (HH) on sub-samples of Hispanics of Mexican and non-Mexican ancestries.
+The evidence presented in this section supports a robust empirical strategy, revealing
+
+significant selection in marriage and distinctive educational and income patterns among different types of families.
+Interethnic couples have comparable levels of education and earnings, suggesting that the children of HW families are likely positioned for better educational outcomes, a critical factor for understanding discrimination in the labor market.
+The disparities between HW and WH families are considerably less pronounced than those between other groups, emphasizing the importance of comparing these children directly.
+Such comparisons shed light on the nuanced dynamics of ethnicity, education, and economic outcomes in interethnic marriages.\footnote{I present in Tables \ref{tab:synthmex} and \ref{tab:syntnonmex} the summary statistics which detail the educational and economic profiles of parents from four different ethnic groupings—White White (WW), White Hispanic (WH), Hispanic White (HW), and Hispanic Hispanic (HH) on sub-samples of Hispanics of Mexican and non-Mexican ancestries.
 I find similar results that describe a selection into inter-ethnic marriages among the two groups.}
 
-Despite higher levels of education and income among Hispanic White (HW) mothers compared to White Hispanic (WH) mothers, suggesting an expectation of greater educational attainment for HW children, data reveals a discrepancy.
-Specifically, HW children complete, on average, 0.4 fewer years of education compared to their WH peers (refer to Table \ref{tab:c&p2}).
+Despite higher levels of education and income among HW mothers compared to WH mothers, HW children complete an average of 0.4 fewer years of education than their WH peers (Table \ref{tab:c&p2}).
 This gap may suggest potential discrimination or barriers in educational access for HW children.
@@
 I provide the results of the estimation of equation \ref{eq:1a} in Table \ref{tab:lastname-ed-reg}.
-I estimate the mean educational outcomes of White, U.S.-born Hispanics ages 25-40.
+I estimate the mean educational outcomes of White, US-born Hispanics ages 25-40.
 I also restrict the sample to children of HW and WH parents.
@@
 Column 1 in Table \ref{tab:lastname-ed-reg} is the difference in total years of education between HW children and their WH peers.
-Column 2 is the difference in the probability of having been a high school dropout.
+Column 2 is the difference in the probability of not completing high school.
 Column 3 is the difference in the probability of having an associate degree.
-Column 4 is the difference in the probability of having a bachelor's degree.
+Column 4 is the difference in the probability of having a bachelor’s degree.
 All regressions include controls for age, parental education and income, and state-year fixed effects.
 
-Overall, there is a significant gap in total years of education between HW and WH children.
-HW children receive 0.2 years of education less than WH children.
+There is a significant gap in total years of education between HW and WH children.
+HW children receive 0.2 fewer years of education than WH children.
 The gap between HW and WH women is larger than the gap between HW and WH men.
-Women with a Hispanic last name receive 0.25 years of education less than WH women.
-The gap between HW and WH men is equal to 0.16 years.
+Women with a Hispanic last name receive 0.25 fewer years of education than WH women.
+The gap between HW and WH men is 0.16 years.
 
-Even though there is a significant, albeit small, gap in total years of education between HW and WH children, there is no significant gap in the probability of being a high school dropout.
-The gap between HW and WH high school dropouts is equal to a statistically insignificant 1 percentage point.
+Although there is a small but significant gap in total years of education between HW and WH children, there is no significant difference in the probability of dropping out of high school.
+The gap between HW and WH high school dropouts is statistically insignificant (1 percentage point).
 The same is true for HW women (2 percentage points) and men (0 percentage points).
 
-However, there is a notable difference when it comes to higher education outcomes.
-HW children are 2 percentage points less likely to earn an associate degree and 3 percentage points less likely to earn a bachelor's degree compared to their WH peers.
+Notable differences emerge for higher education outcomes.
+HW children are 2 percentage points less likely to earn an associate degree and 3 percentage points less likely to earn a bachelor’s degree compared to their WH peers.
 HW children are 2 percentage points less likely to earn an associate degree compared to their WH peers, representing a 13.3\% reduction relative to the WH associate degree rate of 15\%.
-HW children are 3 percentage points less likely to earn a bachelor's degree compared to their WH peers, which is a 13\% reduction relative to the WH bachelor's degree rate of 23\%.
-These gaps are slightly larger for HW women, who are 3 percentage points less likely to earn an associate degree and 4 percentage points less likely to earn a bachelor's degree.
-For HW men, the gap is 2 percentage points for an associate degree and statistically insignificant for a bachelor's degree.
+HW children are 3 percentage points less likely to earn a bachelor’s degree compared to their WH peers, a 13\% reduction compared to the WH bachelor’s degree rate of 23\%.
+These differences are slightly larger for HW women, who are 3 percentage points less likely to earn an associate degree and 4 percentage points less likely to earn a bachelor’s degree.
+For HW men, the gap is 2 percentage points for an associate degree and statistically insignificant for a bachelor’s degree.
 
-These results suggest that while the overall educational attainment gap between HW and WH children is small in terms of years of education, the disparities become more pronounced when considering higher education milestones, particularly for HW women.
-However, given the fact that HWs have more educated mothers than WHs, we would expect them to have higher levels of educational attainment \autocite{kimball2009risk, gould2020does}.
-This could indicate potential barriers or discrimination in access to higher education for HW children, particularly.
+These results suggest that while the overall educational gap between HW and WH children is small in terms of years of education, the disparities become more pronounced when considering higher education milestones, particularly for HW women.
+Since HWs have more educated mothers than WHs, we may expect higher levels of educational attainment from them \autocite{kimball2009risk, gould2020does}.
+The fact that this does not hold true could indicate potential barriers or discrimination in access to higher education for HW children.
 
@@
 
-I provide the results to the estimation of equation \ref{eq:1a} in Tables \ref{tab:lastnamereg-emp} and \ref{tab:lastnamereg} on unemployment and log earnings.
-I estimate the mean unemployment of White, Native-born, Hispanics men aged 25-40.
-I estimate the mean wages of White, U.S.-born, Hispanic men aged 25-40 who are employed FTFY as waged and salaried workers.
+I provide the results of the estimation of equation \ref{eq:1a} in Tables \ref{tab:lastnamereg-emp} and \ref{tab:lastnamereg} on unemployment and log earnings.
+I estimate the mean unemployment and mean wages of White US-born Hispanic men aged 25-40 who are employed full-time.
 I also restrict the sample to children of HW and WH (omitted) parents.
 Column 1 in Table \ref{tab:lastnamereg} is the average crude earnings gap in log annual earnings between HW workers and their WH peers.
-In the next 4 columns, I introduce the results with controls for hours worked, state FE, year FE, age FE, Education FE, and parental background.
+In the next 4 columns, I introduce the results with controls for hours worked, state fixed effects (FE), year FE, age FE, education FE, and parental background.
 
-In addition to examining earnings, I also analyzed the effect of having a Hispanic last name on unemployment rates.
+I also analyzed the effect of having a Hispanic last name on unemployment rates.
 Table \ref{tab:lastnamereg-emp} presents the results of this analysis.
 Column 1 shows that individuals with Hispanic last names (HW) have a 1 percentage point higher unemployment rate compared to those with White last names.
-This gap persists even after controlling for age, state fixed effects, year fixed effects, and state-year fixed effects (Column 2), though the significance level drops.
+This discrepancy persists even after controlling for age, state FE, year FE, and state-year FE (Column 2), though the significance is lower.
 When education is included as a control (Column 3), the gap remains at 1 percentage point.
 Finally, after controlling for parental background (Column 4), the 1 percentage point difference in unemployment rates becomes statistically insignificant.
-These results suggest that while there is an initial unemployment gap associated with having a Hispanic last name, much of this difference can be explained by factors such as education and parental background.
+These results suggest that while there is an initial unemployment gap associated with having a Hispanic last name, much of this difference can be explained by factors such as education and parental background and does not necessarily reflect discrimination.
 The mean unemployment rate for individuals with Hispanic last names (HW) is 7\% across all specifications.
@@
 Overall, the crude gap between HW and WH workers is equal to 5 percentage points (Table \ref{tab:lastnamereg} column 1).
-An inter-ethnic with a Hispanic last name earns 5 percentage points less than an inter-ethnic with a White last name.
-Even after controlling for hours worked, and including state, year, and age FEs in the estimation, the gap stays at 5 percentage points.
-This gap, however, could be entirely explained by educational differences.
-An inter-ethnic with a Hispanic last name earns 1 percentage points less than an inter-ethnic with a White last name, but the result is statistically insignificant.
+An interethnic worker with a Hispanic last name earns 5 percentage points less than an interethnic worker with a White last name.
+Even after controlling for hours worked, and including state, year, and age FEs in the estimation, the gap stays at 5 percentage points; however, the difference could be attributed to educational differences.
+An interethnic individual with a Hispanic last name earns 1 percentage point less than one with a White last name, but the result is statistically insignificant.
 
@@
 
-Since Hispanics are a very heterogeneous group, I conduct a sensitivity analysis on different samples of Hispanics.
-To increase the sample size of my analysis I estimate equation \ref{eq:1a} using weekly earnings as a dependent variable.
-Weekly earnings are available in all the monthly CPS surveys and not just the March Supplement.
-I present the results to these estimations in tables \ref{tab:lastname-ed-reg-mex}, \ref{tab:lastname-ed-reg-nonmex}, \ref{tab:lastnamereg-weekearm-mex}, and \ref{tab:lastnamereg-weekearm-nonmex}.
+Since Hispanics are very heterogeneous, I conduct a sensitivity analysis on different groups.
+To increase the sample size of my analysis, I estimate equation \ref{eq:1a} using weekly earnings as a dependent variable in Tables \ref{tab:lastname-ed-reg-mex}, \ref{tab:lastname-ed-reg-nonmex}, \ref{tab:lastnamereg-weekearm-mex}, and \ref{tab:lastnamereg-weekearm-nonmex}.
 
 First, in Table \ref{tab:lastnamereg-weekearm}, I present the results of estimating equation \ref{eq:1a} using weekly earnings as the dependent variable for the full sample.
-Similar to the previous analysis, I find that a person with a Hispanic last name earns 4 percentage points less than a person without a Hispanic last name.
-This gap could be explained by educational differences.
+Like the previous analysis, I find that a person with a Hispanic last name earns 4 percentage points less than a person without a Hispanic last name, which could be explained by educational differences.
 Second, in Table \ref{tab:lastnamereg-weekearm-mex}, I present the results for a sample of Mexican Hispanics.
-I find that among individuals of Mexican origin, those with that likely have Hispanic surnames earn 3 percentage points less than those with a likely native-sounding surnames.
+I find that among individuals of Mexican origin, those that likely have Hispanic surnames earn 3 percentage points less than those with White-sounding surnames.
 This gap becomes an imprecise zero after controlling for education and parental background.
 Third, in Table \ref{tab:lastnamereg-weekearm-nonmex}, I present the results for a sample of non-Mexican Hispanics.
-The gap between non-Mexicans that likely have a Hispanic-sounding last names and those that likely have a native-sounding last names is similarly explained by educational differences.
-I find that a non-Mexican individual that likely has a Hispanic-sounding last name earns 3 percentage points less than a non-Mexican with a native-sounding last name.
+The gap between non-Mexicans that likely have a Hispanic-sounding last name and those that likely have a White-sounding last name can also be explained by educational differences.
+I find that a non-Mexican individual that likely has a Hispanic-sounding last name earns 3 percentage points less than a non-Mexican with a White-sounding last name.
 This gap also becomes an imprecise zero after controlling for education and parental background.
 
-I also present the results of the estimation of equation \ref{eq:1a} using log annual earnings as the dependent variable but with occupation fixed effects instead of education fixed effects.
-I present the results of this estimation in Table \ref{tab:lastnamereg-occ}.
-I find that the gap between people with a likely Hispanic-sounding last names and those with a likely native-sounding last names is explained by differences in occupation, similar to my findings with educational controls.
-I find that a person with a likely Hispanic-sounding last name earns 5 percentage points less than a person with a likely native-sounding last name.
+I I also present the results of the estimation of equation 1 using log annual earnings as the dependent variable but with occupation fixed effects instead of education fixed effects (Table \ref{tab:lastnamereg-occ}).
+I find that the gap between people with a likely Hispanic-sounding last names and those with a likely White-sounding last names is explained by differences in occupation, similar to my findings with educational controls.
+I find that a person with a likely Hispanic-sounding last name earns 5 percentage points less than a person with a likely White-sounding last name.
 This gap becomes an imprecise zero after controlling for occupation.
@@
 
-As the Hispanic population grows in the United States, studying discrimination against this group becomes increasingly important.
+As the Hispanic population grows in the United States, studying discrimination becomes increasingly important.
 In this paper, I examine discrimination against Hispanics in the labor market.
 More specifically, I examine the impact of Hispanic last names and Hispanic identification on annual log earnings.
-
-I compare the children of inter-ethnic marriages to study the labor and educational markets impact of having a Hispanic last name.
-I find that inter-ethnic people with a likely Spanish-sounding last names receive 0.2 years of education less than inter-ethnic counterparts that likely have a native-sounding last name, representing a 1.4\% reduction in total years of education.
+I compare the children of interethnic marriages to study the labor and educational market effects of having a Hispanic last name and find that interethnic people with likely Hispanic-sounding last names receive 0.2 fewer years of education than interethnic counterparts that likely have a White-sounding  last name, representing a 1.4\% reduction.
 This gap is larger for women (0.25 years, or 1.8\%) than for men (0.16 years, or 1.2\%).
 When I compare the labor market outcomes of HW and WH children, which captures the effect of having a Hispanic last name, HW children are 1 percentage point more likely to be unemployed than WH children, representing a 14.3\% increase from the mean unemployment rate.
-In terms of earnings, HW children earn 5 percentage points (4.9\%) less than WH children.
-Thus, by comparing inter-ethnic children, a person with a Hispanic last name makes 5 percentage points less than someone with a White last name.
+In terms of earnings, HW children (likely to have a Hispanic-sounding surname earn 5 percentage points (4.9\%) less than WH children (likely to have a non-Hispanic-sounding name).
 When I control for education, the last name effect decreases to a statistically insignificant 1 percentage point earnings gap.
+The findings reveal gaps that could be attributed to discrimination against individuals with Hispanic surnames, with effects varying across educational levels and gender.
+While the impact on total years of education appears modest (1.4\% reduction), the more substantial gaps in higher education attainment—particularly the 13\% reduction in bachelor’s degree completion and 13.3\% reduction in associate degree attainment—suggest that discrimination may intensify at crucial educational transition points.
+This pattern is especially pronounced for women with Hispanic surnames, who face larger discrepancies in both total education (1.8\% reduction versus 1.2\% for men) and degree completion (14.8\% reduction in bachelor’s degree attainment ).
+This paper contributes to a body of education economics literature showing disparities in access to education.
+In an audit study of charter and traditional public schools, \textcite{bergman2018education,gaddis2024racial} find that students  with Hispanic names—compared to students that are presumed White—are less likely to get a response from schools they contact for application information.
+This suggests that the earnings gap between people with a likely Hispanic sounding last names and those with a native  sounding last names could be due
 
-The findings reveal gaps that could be attributed to discrimination against individuals with Hispanic surnames, with effects varying across educational levels and gender.
-While the impact on total years of education appears modest (1.4\% reduction), the more substantial gaps in higher education attainment—particularly the 13\% reduction in bachelor's degree completion and 13.3\% reduction in associate degree attainment—suggest that discrimination may intensify at crucial educational transition points.
-This pattern is especially pronounced for women with Hispanic surnames, who face larger penalties in both total education (1.8\% reduction versus 1.2\% for men) and degree completion (14.8\% reduction in bachelor's degree attainment).
-
-There is a large literature in education economics that shows disparities in access to education, which this paper contributes to.
-In an audit study of charter and traditional public schools, \textcite{bergman2018education,gaddis2024racial} find that students with Hispanic names---compared to students that are presumed White---are less likely to get a response to inquiries from schools.
-This could be an indication that people with a likely Hispanic sounding names could end up lower access to education, including high value-added schools.
-Therefore, the earnings gap between people with a likely Hispanic sounding last names and those with native sounding last names could be due to differences in access to education that are in turn due to discrimination.
-This paper provides further evidence, using observational data, that discrimination in access to education could lead to lower earnings for Hispanics.
-
-In the labor market, the 14.3\% higher unemployment rate and 4.9\% earnings penalty for those with Hispanic surnames indicate persistent discrimination even among individuals who share similar family backgrounds.
-Putting the results in context, \autocite{trejoWhyMexicanAmericans1997} find an earning gap between US-born Hispanics and US-born whites that is equal to 30.7 percentage points.
-My findings could be an indication that out of the 30.7 percentage points gap, at least 16\% of it could be attributed to discrimination against Hispanics.
+To differences in access to education that are in turn due to discrimination.
+This paper provides further evidence, using observational data, that discrimination in access to education could lead to lower earnings for Hispanic workers.
+In the labor market, the 14.3\% higher unemployment rate and 4.9\% lower salaries for those with Hispanic surnames indicate persistent discrimination even among individuals who share similar family backgrounds.
+Putting the results in context, \autocite{trejoWhyMexicanAmericans1997} found an earnings gap between US-born Hispanics and US-born whites that is to 30.7 percentage points.
+My findings could be an indication that 16\% of this gap could be attributed to discrimination against Hispanics.
 However, the fact that the earnings gap becomes statistically insignificant after controlling for education suggests that much of the labor market discrimination operates through educational channels rather than direct wage discrimination.
@@
 
-While the earnings gap between children of inter-ethnic parents with and without a Hispanic last name disappears when controlling for education, it does not necessarily indicate the absence of discrimination.
+The earnings gap between individuals with vs without a Hispanic last name disappears when controlling for education, but this does not necessarily indicate the absence of discrimination.
 Education itself can be influenced by bias, potentially resulting in divergent outcomes \autocite{bergman2018education,gaddis2024racial}.
-This is especially the case when parental characteristics indicate that people with a likely Hispanic-sounding last-names should in theory complete more years of education---since mothers of inter-ethnic children with a Hispanic last-names have more years of education and earn more.
+This is especially the case when parental characteristics indicate that people with a likely Hispanic-sounding last-names should in theory complete more years of education, since mothers of interethnic children with likely Hispanic surnames have more years of education and earn more.
 While I observe where discrimination manifests (particularly in higher education access), my analysis cannot identify the specific mechanisms—such as teacher bias, institutional barriers, or social networks—through which surname-based discrimination operates.
-Consequently, further research is needed to comprehensively understand the earnings gaps between Hispanics and Whites.
+Further research is needed to understand these disparities.

### 2025-05-13 manuscripts/09-JERP-v2/manuscript_main.tex
start working on R&R v2

--- 2025-03-19 manuscripts/08-JERP-v6/manuscript_main.tex
@@
 Discrimination may foster occupational sorting and segregation, as shown by recent task-based models of racial wage gaps that identify race-specific barriers in “Contact” tasks and explain the stagnation of these gaps post-1980 \autocite{hurst2024task}.
-These barriers may trap Hispanic families in cycles of lower socioeconomic status, as reduced earnings and employment opportunities limit their ability to invest in their children's human capital or build wealth through homeownership and savings.\footnote{See \textcite{bowles2002inheritance} for an analysis of the intergenerational transmission of economic status.} Although previous research has identified earnings gaps between Hispanic and non-Hispanic Whites, this paper makes several important contributions.
+These barriers may trap Hispanic families in cycles of lower socioeconomic status, as reduced earnings and employment opportunities limit their ability to invest in their children's human capital or build wealth through homeownership and savings.\footnote{See \textcite{bowles2002inheritance} for an analysis of the intergenerational transmission of economic status.}
+
+Although previous research has identified earnings gaps between Hispanic and non-Hispanic Whites, this paper makes several important contributions.
 First, it develops a novel empirical strategy that better isolates the causal effect of Hispanic ethnicity on educational and labor market outcomes by comparing children from inter-ethnic families.
@@
 While these decomposition methods have provided valuable insights into earnings gaps between groups, including Hispanics and non-Hispanic Whites \autocite{davilaChangesRelativeEarnings2008}, they are limited in their ability to account for unobserved differences between the groups.
-My approach comparing children of inter-ethnic marriages, provides a more closely matched comparison group, allowing me to better isolate the effect of having a Hispanic last name on labor market outcomes.
+My approach comparing children of inter-ethnic marriages, provides a more closely matched comparison group, allowing me to better isolate the effect of likely having a Hispanic last name on labor market outcomes.
 
@@
 Specifically, individuals with presumably Hispanic last names complete 0.2 fewer years of education than those with non-Hispanic names, even when controlling for family background.
-Individuals with Hispanic last names are also 1 percentage point more likely to be unemployed and earn 5 percentage points less than those with non-Hispanic names.
+Individuals with presumably Hispanic last names are also 1 percentage point more likely to be unemployed and earn 5 percentage points less than those with likely non-Hispanic names.
 These findings suggest that ethnic discrimination continues to play a role in shaping economic opportunities for Hispanics in the United States, highlighting the need for targeted policies to address these persistent disparities.
@@
 The authors compared the children of mixed marriages between Sephardic and Ashkenazi people in Israel.\footnote{Ashkenazi and Sephardic are two distinct Jewish ethnic groups.} They found that workers with Sephardic last names earn substantially less than those with Ashkenazi last names.
-I compare children of Hispanic fathers and White mothers (henceforth HW) to children of White fathers and Hispanic mothers (WH).
+I compare children of Hispanic fathers and White mothers (henceforth HW) and likely have a Hispanic-sounding last name to children of White fathers and Hispanic mothers (WH).
 This approach accounts for the fact that  couples are likely to have similar income, schooling, and socioeconomic background, so these factors are largely controlled for \autocite{averettBetterWorseRelationship2008, averettEconomicRealityBeauty1996}.\footnote{For more on assortative mating see \autocite{beckerTheoryMarriagePart1973, beckerTheoryMarriagePart1974, beckerTreatiseFamily1993, browningCollectiveUnitaryModels2006, chiapporiFatterAttractionAnthropometric2012}} Children of HW and WH marriages have more similar observable characteristics than children of endogamous/homogamous marriages (i.e., White fathers-White mothers and Hispanic fathers-Hispanic mothers).
@@
 \item White father and White mother (hereafter WW) 
-\item White father and Hispanic mother (hereafter WH)
-\item Hispanic father and White mother (hereafter HW)
+\item White father and Hispanic mother (hereafter WH)---the comparison group and the group that is likely to have a non-Hispanic-sounding last name
+\item Hispanic father and White mother (hereafter HW)---the focus of this study and the group most likely to have a Hispanic-sounding last name
 \item Hispanic father and Hispanic mother (hereafter HH).
@@
 My sample includes inter-ethnic children with Hispanic ancestry aged 25–40 who are US citizens.
-I include both Hispanic White and non-Hispanic White individuals, since ethnic attrition could lead to biased estimates when measuring gaps as shown  by \textcite{hadah2024effect}.
+I include both Hispanic White and non-Hispanic White individuals in this study.
+For second-generation immigrants, I use parents' place of birth as a proxy for ethnicity.
+This approach helps avoid biased estimates that could result from ethnic attrition, as demonstrated by \textcite{hadah2024effect}.
 \footnote{A person self-reports Hispanic identity by answering the Census question: ``Is this person Spanish/Hispanic/Latino?''} \footnote{Ethnic attrition happens when a US-born descendant of a Hispanic immigrant fails to self-identify as Hispanic.
@@
 WW couples have the highest level of education among the four groups: 12.58 years  for husbands and 12.36 for wives, so a WW household has 24.95 years of schooling.
-Men in HH marriages have 8.64 years of education, while women have 8.49.
-As a household, HH couples have 17.13 years of education.
+Men in HH marriages have 8.91 years of education, while women have 8.68.
+As a household, HH couples have 17.69 years of education.
 As predicted, interethnic couples marry people similar to them.
@@
 
-In terms of labor market performance, WW households boast the highest log total family income at 10.75 , while HH households fall at the lower end with 10.42.
-Among interethnic couples, WH households have a slightly higher total income of 10.65 compared to HW households at 10.60.
+In terms of labor market performance, WW households boast the highest log total family income, while HH households fall at the lower end.
+Among interethnic couples, WH households have a slightly higher total income compared to HW households---WH households earn 5\% more than HW.
 The difference in husbands’ log hourly earnings between HW and WH is marginal, with HW men earning 4\% less than their WH counterparts.
-HW women surpass WH women in earnings, with HW wives earning 1.75 and WH wives earning 1.73 in log hourly earnings, respectively.
+HW women earn 2\% more than WH women.
 This reversal in the typical earning pattern speaks to the closing economic disparities between these groups and implies potentially greater economic contributions from HW women to their families, which could benefit their children when they enter the labor market.
@@
 
-The evidence presented in this section supports a robust empirical strategy, revealing
-
-significant selection in marriage and distinctive educational and income patterns among different types of families.
+The evidence presented in this section supports a robust empirical strategy, revealing significant selection in marriage and distinctive educational and income patterns among different types of families.
 Interethnic couples have comparable levels of education and earnings, suggesting that the children of HW families are likely positioned for better educational outcomes, a critical factor for understanding discrimination in the labor market.
@@
 Even after controlling for hours worked, and including state, year, and age FEs in the estimation, the gap stays at 5 percentage points; however, the difference could be attributed to educational differences.
-An interethnic individual with a Hispanic last name earns 1 percentage point less than one with a White last name, but the result is statistically insignificant.
+An interethnic man with a likely Hispanic last name earns 1 percentage point less than one with a White last name, but the result is statistically insignificant.
 

### 2025-05-13 manuscripts/10-JERP-v3/manuscript_main.tex
start working on R&R v2

--- 2025-05-13 manuscripts/09-JERP-v2/manuscript_main.tex
@@
 Hispanic individuals constitute a large and growing portion of the US population, and as this number increases, it is increasingly crucial to determine whether ethnic discrimination affects their employment prospects, wages, and career advancement.
-Understanding these labor market outcomes is critical, as they directly affect broader societal issues like assimilation and economic mobility
-\autocite{chettyUnitedStatesStill2014, chettyEffectsExposureBetter2016,chettyFadingAmericanDream2017}.
+Understanding these labor market outcomes is critical, as they directly affect broader societal issues like assimilation and economic mobility, \autocite{chettyUnitedStatesStill2014, chettyEffectsExposureBetter2016,chettyFadingAmericanDream2017}.
 These factors serve as key indicators of how successfully Hispanic immigrants can navigate society and ascend the socioeconomic ladder.
@@
 Does having a Hispanic last name affect labor market outcomes?
-I aim to show that comparing Hispanic White individuals to non-Hispanic White individuals might create an artificially higher earnings gap since the two groups differ in many observable characteristics.\footnote{By identifying as Hispanic, I refer to individuals who self-report their ethnicity as Hispanic on surveys or other data collection instruments.} \footnote{Observable characteristics refer to factors that can be measured and quantified, such as education level, work experience, and immigration status.} My analysis focuses on US-born children with foreign-born parent (i.e., inter-ethnic).\footnote{This study's focus on US-born children with foreign-born parents means the findings may not generalize to Hispanic children with US-born parents, who likely face different socioeconomic and cultural circumstances.
+I aim to show that comparing Hispanic White individuals to to non-Hispanic White individuals might create an artificially higher earnings gap since the two groups differ in many observable characteristics.\footnote{I define Hispanic individuals as those who self-report their ethnicity as Hispanic on surveys or other data collection instruments.} \footnote{Observable characteristics refer to factors that can be measured and quantified, such as education level, work experience, and immigration status.} My analysis focuses on US-born children with foreign-born parent (i.e., inter-ethnic).\footnote{This study's focus on US-born children with foreign-born parents means the findings may not generalize to Hispanic children with US-born parents, who likely face different socioeconomic and cultural circumstances.
 The analysis does not account for heterogeneity in immigrant characteristics across different Spanish-speaking countries of origin, such as variations in educational attainment, socioeconomic status, and gender-specific migration patterns, which could influence both parental selection into migration and subsequent child outcomes.} Others have attempted to compare how native-born White Hispanics fare compared to non-Hispanic Whites and foreign-born Hispanics.
@@
 
-This study builds upon and improves the traditional Oaxaca-Blinder-Kitagawa decomposition approach often used in labor market discrimination studies \autocite{kitagawa1955components, oaxaca1973male,blinder1973wage}.
+This study builds upon and improves upon the traditional Oaxaca-Blinder-Kitagawa decomposition approach often used in labor market discrimination studies \autocite{kitagawa1955components, oaxaca1973male,blinder1973wage}.
 While these decomposition methods have provided valuable insights into earnings gaps between groups, including Hispanics and non-Hispanic Whites \autocite{davilaChangesRelativeEarnings2008}, they are limited in their ability to account for unobserved differences between the groups.
@@
 The US population is growing increasingly diverse, with significant implications for labor market dynamics and potential discrimination.
-The proportion of non-Whites has increased by more than 10 percentage points from 13 percent in 1995 to 23 percent in 2019.\footnote{The proportion of non-Whites and Hispanics is calculated using the Current Population Survey (CPS).} Native-born White Hispanic men earn 21\% less than White men, although there is evidence that a substantial portion of this gap is due to differences in education \autocite{duncan2006hispanics, duncan2018identifying, duncan2018socioeconomic}.
+The proportion of non-Whites has increased by more than 10 percentage points from 13 percent in 1995 to 23 percent in 2019.\footnote{The proportion of non-Whites and Hispanics is based on the author's calculations using the Current Population Survey (CPS).} Native-born White Hispanic men earn 21\% less than White men, although there is evidence that a substantial portion of this gap is due to differences in education \autocite{duncan2006hispanics, duncan2018identifying, duncan2018socioeconomic}.
 Discrepancies may also be due to discrimination against Hispanics, which can lead to reduced job opportunities, lower wages, and poor assimilation \autocite{chettyWhereLandOpportunity2014,bowles2002inheritance, djajic2003assimilation}.\footnote{This study focuses on US-born children of inter-ethnic unions, many of whom have at least one foreign-born parent.
@@
 The authors compared the children of mixed marriages between Sephardic and Ashkenazi people in Israel.\footnote{Ashkenazi and Sephardic are two distinct Jewish ethnic groups.} They found that workers with Sephardic last names earn substantially less than those with Ashkenazi last names.
-I compare children of Hispanic fathers and White mothers (henceforth HW) and likely have a Hispanic-sounding last name to children of White fathers and Hispanic mothers (WH).
-This approach accounts for the fact that  couples are likely to have similar income, schooling, and socioeconomic background, so these factors are largely controlled for \autocite{averettBetterWorseRelationship2008, averettEconomicRealityBeauty1996}.\footnote{For more on assortative mating see \autocite{beckerTheoryMarriagePart1973, beckerTheoryMarriagePart1974, beckerTreatiseFamily1993, browningCollectiveUnitaryModels2006, chiapporiFatterAttractionAnthropometric2012}} Children of HW and WH marriages have more similar observable characteristics than children of endogamous/homogamous marriages (i.e., White fathers-White mothers and Hispanic fathers-Hispanic mothers).
+I compare children of Hispanic fathers and US born mothers (henceforth HW) and likely have a Hispanic-sounding last name to children of US born fathers and Hispanic mothers (WH).
+This approach accounts for the fact that  couples are likely to have similar income, schooling, and socioeconomic background, so these factors are largely controlled for \autocite{averettBetterWorseRelationship2008, averettEconomicRealityBeauty1996}.\footnote{For more on assortative mating see \autocite{beckerTheoryMarriagePart1973, beckerTheoryMarriagePart1974, beckerTreatiseFamily1993, browningCollectiveUnitaryModels2006, chiapporiFatterAttractionAnthropometric2012}} Children of HW and WH marriages have more similar observable characteristics than children of endogamous/homogamous marriages (i.e., US born fathers-US born mothers and Hispanic fathers-Hispanic mothers).
 In the US, children  from households with a Hispanic father and White mother  are overwhelmingly likely to  have their father’s Hispanic last name, allowing us to investigate how this ethnic signal affects annual log earnings.
@@
 The main identifying assumption of my empirical strategy depends on the assumption that people born to HW parents are similar to their WH peer in all observable---and unobservable---aspects and characteristics relevant to the labor market, so surname is the only difference between these groups.
-Children from White-Hispanic homes might benefit from more favorable family conditions than those from Hispanic-White homes, considering Whites statistically have higher socioeconomic status than Hispanics.
-These factors introduce doubts regarding whether children from Hispanic-White families receive comparable familial support and influences, genetically or environmentally, as those from White-Hispanic families.
+Children from US born-Hispanic homes might benefit from more favorable family conditions than those from Hispanic-US born homes, considering Whites statistically have higher socioeconomic status than Hispanics.
+These factors introduce doubts regarding whether children from Hispanic-US born families receive comparable familial support and influences, genetically or environmentally, as those from US born-Hispanic families.
 
@@
 Taking advantage of data on parents’ place of birth, I divide the sample into four groups depending on their parents’ ethnicity.
-Mothers or fathers are Hispanic if they were born in a Spanish-speaking country or Puerto Rico, and White if they were born in the United States.\footnote{The list of Spanish-speaking countries includes: Argentina, Bolivia, Chile, Colombia, Costa Rica, Cuba, Dominican Republic, Ecuador, El Salvador, Guatemala, Equatorial Guinea, Honduras, Mexico, Nicaragua, Panama, Paraguay, Peru, Spain, Uruguay, Venezuela.} Therefore, an observation can be the product of four types of parents:
+Mothers or fathers are Hispanic if they were born in a Spanish-speaking country or Puerto Rico, and US born White if they were born in the United States.\footnote{The list of Spanish-speaking countries includes: Argentina, Bolivia, Chile, Colombia, Costa Rica, Cuba, Dominican Republic, Ecuador, El Salvador, Guatemala, Equatorial Guinea, Honduras, Mexico, Nicaragua, Panama, Paraguay, Peru, Spain, Uruguay, Venezuela.} Therefore, an observation can be the product of four types of parents:
 \begin{enumerate}
-\item White father and White mother (hereafter WW) 
-\item White father and Hispanic mother (hereafter WH)---the comparison group and the group that is likely to have a non-Hispanic-sounding last name
-\item Hispanic father and White mother (hereafter HW)---the focus of this study and the group most likely to have a Hispanic-sounding last name
+\item US born father and US born mother (hereafter WW) 
+\item US born father and Hispanic mother (hereafter WH)---the comparison group and the group that is likely to have a non-Hispanic-sounding last name
+\item Hispanic father and US born mother (hereafter HW)---the focus of this study and the group most likely to have a Hispanic-sounding last name
 \item Hispanic father and Hispanic mother (hereafter HH).
@@
 I consider a Hispanic persons that self-report White as their racial group and are born in a Spanish-speaking country.
-Consequently, Whites in the sample are White and native-born.
+Consequently, US born in the sample are those that self-report White as race and are native-born.
 Using the information provided in the census, I can link husbands and wives with each other.
@@
 This method preserves the demographic composition of parents while assigning them historically accurate socioeconomic characteristics based on their origins and marriage patterns.
+
 I show the distribution of the four types of couples in Table \ref{tab:mat2}.
-White husbands and White wives (WW) make up the majority of couples in the sample, 96\% (5,141,737 couples).
-Hispanic husbands and wives (HH) are the second largest group, coprising 2\% (119,749 couples).
-White husbands and Hispanic wives (WH) $< 1\%$ (33,097 couples) of the sample and Hispanic husbands and White wives (HW) are around 1\% (37,847 couples).
+US born husbands and US born wives (WW) make up the majority of couples in the sample, 96\% (5,141,737 couples).
+Hispanic husbands and wives (HH) are the second-largest group, compromising 2\% (119,749 couples).
+US born husbands and Hispanic wives (WH) $< 1\%$ (33,097 couples) of the sample and Hispanic husbands and US born wives (HW) are around 1\% (37,847 couples).
 I present the summary statistics of the parents in Table \ref{tab:synth}.
@@
 
-% These issues could be addressed by selecting partners among interethnic couples.
-The average White person exhibits better observable characteristics  than the average Hispanic person.
-If marriages were random, the White husband in a WH marriage would have better traits  than his wife, and the White wife in the HW marriage would have the stronger traits.
-Random matching in marriages would pose a challenge to the identification strategy.
-This is because parents are not perfect substitutes in the inter-generational transmission of human capital.
-Consequently, differences in parental background could significantly affect children’s labor market outcomes, potentially invalidating the comparison between HW and WH children.
-
 A large body of empirical and theoretical work indicates that marriages exhibit strong selection, or "assortative matching." \autocite{averettBetterWorseRelationship2008, averettEconomicRealityBeauty1996, beckerTheoryMarriagePart1973, beckerTheoryMarriagePart1974, beckerTreatiseFamily1993, browningCollectiveUnitaryModels2006, chiapporiFatterAttractionAnthropometric2012}.
 Also, \textcite{duncanIntermarriageIntergenerationalTransmission2011} show a similar pattern of assortative matching among Mexicans in the United States.
-Interethnic WH and HW marriages generally consist of partners sharing similar characteristics , suggesting both WH and HW parents are similar in aspects relevant to labor market outcomes.
+Interethnic WH and HW marriages generally consist of partners sharing similar characteristics, suggesting both WH and HW parents are similar in aspects relevant to labor market outcomes.
 
@@
 In comparing WH and HW individuals to each other, any gaps that emerge between these otherwise similar groups are more likely to be attributable to discrimination rather than underlying differences in family background or socioeconomic status.
+
+It is important to acknowledge potential threats to identification in this approach.
+While HW and WH couples exhibit similar observable socioeconomic characteristics, they may differ systematically in unobservable dimensions that influence labor market outcomes for their children.
+These unobservable differences could include parenting styles, educational preferences, time investments in children, gender norms, or cultural attitudes that are not captured in conventional socioeconomic metrics.
+However, this approach offers distinct advantages over traditional Oaxaca-Blinder-Kitagawa decomposition methods.
+First, by focusing specifically on children from interethnic marriages with similar parental characteristics but different surname ethnicities, I use a more targeted comparison group that better isolates the effect of perceived ethnicity from family background.
+Traditional decomposition approaches would struggle to separate discrimination effects from the broader socioeconomic gaps between Hispanic and White families overall.
+Second, the smaller educational and income disparities between HW and WH families compared to HH and WW families suggest that selection bias is less severe in this comparison, allowing for a clearer attribution of outcome differences to discrimination rather than unobserved heterogeneity.
+Nevertheless, I acknowledge that residual selection bias remains a concern, particularly regarding cultural and attitudinal factors that may systematically differ between HW and WH households despite similar educational and economic profiles.
 
@@
 
-As the Hispanic population grows in the United States, studying discrimination becomes increasingly important.
+As the Hispanic population grows in the United States, studying discrimination becomes important.
 In this paper, I examine discrimination against Hispanics in the labor market.

### 2025-05-13 manuscripts/10-JERP-v3/tab07-regression-weekearn.tex
start working on R&R v2

--- 2023-11-30 my_paper/tables/tab08-regression-weekearn-mex.tex
@@
 \begin{table}[H]
-
-\caption{Effect of Having Hispanic Last Name: Hispanics with Mexican Ancestry  \label{tab:lastnamereg-weekearm-mex}}
+\centering\centering
+\caption{Effect of Having Hispanic Last Name (Log Weekly Earnings) \label{tab:lastnamereg-weekearm}}
 \centering
@@
 \midrule
-$HW_{ist}$ & \num{-0.01} & \num{-0.03}* & \num{0.00} & \num{-0.01}\\
- & (\num{0.02}) & (\num{0.02}) & (\num{0.01}) & (\num{0.01})\\
-Constant & \num{6.53}*** &  &  & \\
- & (\num{0.04}) &  &  & \\
+$HW_{ist}$ & \num{-0.03}** & \num{-0.05}*** & \num{-0.02}* & \num{-0.01}\\
+ & (\num{0.01}) & (\num{0.01}) & (\num{0.01}) & (\num{0.01})\\
+Constant & \num{6.60}*** &  &  & \\
+ & (\num{0.03}) &  &  & \\
 \midrule
@@
 Year FE &  & X & X & X\\
-Age FE &  & X & X & X\\
-Education FE &  &  & X & X\\
+State-Year FE &  & X & X & X\\
+Age &  & X & X & X\\
+Education &  &  & X & X\\
 Parental Background &  &  &  & X\\
-Observations & \num{4535} & \num{4535} & \num{4535} & \num{4535}\\
+Observations & \num{6568} & \num{6568} & \num{6568} & \num{6568}\\
 \bottomrule

### 2025-05-13 programs/114-table-three-summary-stats.R

@@ -1,920 +1,0 @@
-# summary stats
-
-# table three
-
-library(Hmisc)
-
-# date: January 19th, 2023
-
-IndividualData <- read_dta(file.path(datasets,"BySexAnalysisData.dta")) |> 
-  mutate(Education = as.numeric(Education))
-ParentData     <- read_dta(file.path(datasets,"ParentDataFull.dta")) |> 
-  mutate(Tot_ed = HusbandEducation + WifeEducation)
-
-row1 <- c("Variables", 
-          "\\specialcell{White \\\\ White \\\\ (WW) \\\\ (1)}",
-          "\\specialcell{White \\\\ Hispanic \\\\ (WH) \\\\ (2)}",
-          "\\specialcell{Hispanic \\\\ White \\\\ (HW) \\\\ (3)}",
-          "\\specialcell{Hispanic \\\\ Hispanic \\\\ (HH) \\\\ (4)}",
-          "\\specialcell{HH - WW \\\\ (5)}",
-          "\\specialcell{HW - WH \\\\ (6)}"
-          )
-dim(row1) <- c(1,7)
-row2 <- c("Panel A: Synthetic Parents", " ", " ", " ", " ", " ", " ")
-dim(row2) <- c(1,7)
-
-# parent's panel
-
-## Husband's Education
-HusbandEducation <- ParentData |> 
-  group_by(CoupleType) |> 
-  summarise(HusEducation   = weighted.mean(HusbandEducation, na.rm = T, hhwt),
-            HusEducationSD = weighted.sd(HusbandEducation, na.rm = T, hhwt))
-
-model <- lm(HusbandEducation ~ 0 + WW + WM + MW + MM, 
-    data = ParentData, 
-    weights = hhwt)
-test1 <- tidy(glht(model, linfct = c("MM - WW = 0")))
-pvalues1 <- test1$std.error
-differences1 <- test1$estimate
-
-test2 <- tidy(glht(model, linfct = c("MW - WM = 0")))
-pvalues2 <- test2$std.error
-differences2 <- test2$estimate
-
-row3 <- c(
-  "Husband's education (Total Years)",
-  paste0("\\specialcell{", 
-         round(HusbandEducation[4,2], digits = 2), 
-         "\\\\",
-         "(",
-         round(HusbandEducation[4,3], digits = 2),
-         ")}"
-  ),
-  paste0("\\specialcell{", 
-         round(HusbandEducation[3,2], digits = 2), 
[…]

### 2025-05-17 manuscripts/10-JERP-v3/manuscript_main.tex
some edits

@@
 A large body of literature provides evidence of substantial earnings gaps across race and ethnicity in the US \autocite{bayer2018divergent, charles2008prejudice}.
-Hispanic individuals constitute a large and growing portion of the US population, and as this number increases, it is increasingly crucial to determine whether ethnic discrimination affects their employment prospects, wages, and career advancement.
+Hispanic individuals constitute a large and growing portion of the US population.
+As this population grows, it is crucial to determine whether ethnic discrimination affects their employment prospects, wages, and career advancement.
 Understanding these labor market outcomes is critical, as they directly affect broader societal issues like assimilation and economic mobility, \autocite{chettyUnitedStatesStill2014, chettyEffectsExposureBetter2016,chettyFadingAmericanDream2017}.
@@
 Does having a Hispanic last name affect labor market outcomes?
-I aim to show that comparing Hispanic White individuals to to non-Hispanic White individuals might create an artificially higher earnings gap since the two groups differ in many observable characteristics.\footnote{I define Hispanic individuals as those who self-report their ethnicity as Hispanic on surveys or other data collection instruments.} \footnote{Observable characteristics refer to factors that can be measured and quantified, such as education level, work experience, and immigration status.} My analysis focuses on US-born children with foreign-born parent (i.e., inter-ethnic).\footnote{This study's focus on US-born children with foreign-born parents means the findings may not generalize to Hispanic children with US-born parents, who likely face different socioeconomic and cultural circumstances.
+I aim to show that comparing Hispanic White individuals to non-Hispanic White individuals might create an artificially higher earnings gap since these groups differ in many observable characteristics.\footnote{In this study, I define Hispanic individuals specifically as US-born persons who self-identify as White and have at least one parent born in a Spanish-speaking country.} \footnote{Observable characteristics refer to factors that can be measured and quantified, such as education level, work experience, and immigration status.} My analysis focuses on US-born children with one foreign-born parent (i.e., inter-ethnic).\footnote{This study's focus on US-born children with foreign-born parents means the findings may not generalize to Hispanic children with US-born parents, who likely face different socioeconomic and cultural circumstances.
 The analysis does not account for heterogeneity in immigrant characteristics across different Spanish-speaking countries of origin, such as variations in educational attainment, socioeconomic status, and gender-specific migration patterns, which could influence both parental selection into migration and subsequent child outcomes.} Others have attempted to compare how native-born White Hispanics fare compared to non-Hispanic Whites and foreign-born Hispanics.
 \cites{antman2020ethnic} have compared the health and educational outcomes of Hispanic Whites to non-Hispanic Whites and native-born Hispanics to foreign-born Hispanics.
-They find gaps in education and health between Hispanics and Whites.
-They also find that native-born Hispanics are more likely than their foreign-born counterparts to report poor health.\footnote{For additional studies examining ethnic identity and outcomes among Hispanic populations, see \textcite{antman2020ethnic,antmanEthnicAttritionObserved2016,antmanEthnicAttritionObserved2016a,antmanEthnicAttritionAssimilation2020}, which collectively document the relationships between ethnic identification, assimilation patterns, and various socioeconomic outcomes.} \textcite{davilaChangesRelativeEarnings2008} found gaps in labor market outcomes between Hispanics and Whites, which they attribute largely to differences in education, experience, immigration status, and regional differences.
+They find gaps in education and health between Hispanics and non-Hispanic Whites.
+They also find that native-born Hispanics are more likely than their foreign-born counterparts to report poor health.\footnote{For additional studies examining ethnic identity and outcomes among Hispanic populations, see \textcite{antman2020ethnic,antmanEthnicAttritionObserved2016,antmanEthnicAttritionObserved2016a,antmanEthnicAttritionAssimilation2020}, which collectively document the relationships between ethnic identification, assimilation patterns, and various socioeconomic outcomes.} \textcite{davilaChangesRelativeEarnings2008} found gaps in labor market outcomes between Hispanics and non-Hispanic Whites, which they attribute largely to differences in education, experience, immigration status, and regional differences.
 This paper builds on those studies by emphasizing how cultural assimilation and generational status further shape educational and labor market disparities for Hispanics.
@@
 Understanding discrimination against Hispanics in labor markets has far-reaching implications that extend beyond individual earnings.
-Labor market discrimination can create persistent intergenerational disadvantages by limiting economic mobility, reducing access to quality healthcare and education, and constraining residential choices \autocite{chettyUnitedStatesStill2014, chettyEffectsExposureBetter2016,chettyFadingAmericanDream2017,bowles2002inheritance, djajic2003assimilation}.
+Labor market discrimination can create persistent intergenerational disadvantages by limiting economic mobility, reducing access to quality healthcare and education, and constraining residential choices; \autocite{chettyUnitedStatesStill2014, chettyEffectsExposureBetter2016,chettyFadingAmericanDream2017,bowles2002inheritance, djajic2003assimilation}.
 Discrimination may foster occupational sorting and segregation, as shown by recent task-based models of racial wage gaps that identify race-specific barriers in “Contact” tasks and explain the stagnation of these gaps post-1980 \autocite{hurst2024task}.
@@
 
-This study builds upon and improves upon the traditional Oaxaca-Blinder-Kitagawa decomposition approach often used in labor market discrimination studies \autocite{kitagawa1955components, oaxaca1973male,blinder1973wage}.
+This study builds and improves upon the traditional Oaxaca-Blinder-Kitagawa decomposition approach often used in labor market discrimination studies \autocite{kitagawa1955components, oaxaca1973male,blinder1973wage}.
 While these decomposition methods have provided valuable insights into earnings gaps between groups, including Hispanics and non-Hispanic Whites \autocite{davilaChangesRelativeEarnings2008}, they are limited in their ability to account for unobserved differences between the groups.
@@
 The US population is growing increasingly diverse, with significant implications for labor market dynamics and potential discrimination.
-The proportion of non-Whites has increased by more than 10 percentage points from 13 percent in 1995 to 23 percent in 2019.\footnote{The proportion of non-Whites and Hispanics is based on the author's calculations using the Current Population Survey (CPS).} Native-born White Hispanic men earn 21\% less than White men, although there is evidence that a substantial portion of this gap is due to differences in education \autocite{duncan2006hispanics, duncan2018identifying, duncan2018socioeconomic}.
-Discrepancies may also be due to discrimination against Hispanics, which can lead to reduced job opportunities, lower wages, and poor assimilation \autocite{chettyWhereLandOpportunity2014,bowles2002inheritance, djajic2003assimilation}.\footnote{This study focuses on US-born children of inter-ethnic unions, many of whom have at least one foreign-born parent.
+The proportion of non-Whites has increased by more than 10 percentage points from 13 percent in 1995 to 23 percent in 2019.\footnote{The proportion of non-Whites and Hispanics is based on the author's calculations using the Current Population Survey (CPS).} Native-born White Hispanic men earn 21\% less than non-Hispanic White men, although there is evidence that a substantial portion of this gap is due to differences in education \autocite{duncan2006hispanics, duncan2018identifying, duncan2018socioeconomic}.
+These discrepancies may also stem from discrimination against Hispanics, which can lead to reduced job opportunities, lower wages, and hindered assimilation \autocite{chettyWhereLandOpportunity2014,bowles2002inheritance, djajic2003assimilation}.\footnote{This study focuses on US-born children of inter-ethnic unions, many of whom have at least one foreign-born parent.
 Assimilation concerns remain relevant because foreign-born parents may pass on cultural norms and language preferences, which can still shape these children's educational and labor market outcomes.}
 
-Using this novel approach, I find significant effects of Hispanic surnames on both educational and labor market outcomes, and that individuals that likely have a Hispanic last names face substantial disadvantages in educational attainment and earnings compared to their counterparts with likely non-Hispanic names.
+ Using this novel approach, I find significant effects of Hispanic surnames on both educational and labor market outcomes.
+Individuals who likely have Hispanic last names face substantial disadvantages in educational attainment and earnings compared to their counterparts with likely non-Hispanic names.
 Specifically, individuals with presumably Hispanic last names complete 0.2 fewer years of education than those with non-Hispanic names, even when controlling for family background.
@@
 The main identifying assumption of my empirical strategy depends on the assumption that people born to HW parents are similar to their WH peer in all observable---and unobservable---aspects and characteristics relevant to the labor market, so surname is the only difference between these groups.
-Children from US born-Hispanic homes might benefit from more favorable family conditions than those from Hispanic-US born homes, considering Whites statistically have higher socioeconomic status than Hispanics.
-These factors introduce doubts regarding whether children from Hispanic-US born families receive comparable familial support and influences, genetically or environmentally, as those from US born-Hispanic families.
+Children from US-born-Hispanic homes might benefit from more favorable family conditions than those from Hispanic-US-born homes, considering Whites statistically have higher socioeconomic status than Hispanics.
+These factors introduce doubts regarding whether children from Hispanic-US-born families receive comparable familial support and influences, genetically or environmentally, as those from US-born-Hispanic families.
 
@@
 
-I use three datasets:  the Integrated Public Use Microdata Series (IPUMS) Current Population Survey (CPS), CPS' Annual Social and Economic (ASEC) supplement, and CPS' outgoing rotation \autocite{cps2019}, and the 1960 to 2000 US censuses \autocite{acs2019}.
+I use three datasets: the Integrated Public Use Microdata Series (IPUMS) Current Population Survey (CPS), CPS's Annual Social and Economic (ASEC) supplement, and CPS's outgoing rotation \autocite{cps2019}, and the 1960 to 2000 US censuses \autocite{acs2019}.
 
-I use the 1994–2019 CPS data set to study the effect of Hispanic surnames on labor market outcomes, because that span contains data on parents’ place of birth, ethnicity, and race.
+I use the 1994–2019 CPS data set to study the effect of Hispanic surnames on labor market outcomes, because this time period contains data on parents’ place of birth, ethnicity, and race.
 The CPS does not provide data on parents’ characteristics, essential to determine the family background, but the 1960–2000 US Census data includes parents’ place of birth, race, and ethnicity.
-I employ this information to construct "synthetic parents" using a method developed by \textcite{rubinstein2014pride} linking husbands and wives  in the census data.
+I employ this information to construct 'synthetic parents' using a method developed by \textcite{rubinstein2014pride} linking husbands and wives  in the census data.
 Assuming that parents have children 25–40 years old, I link these “parents” using parents’ places of birth, birth year of "children," and the parents’ places of birth.
@@
 Taking advantage of data on parents’ place of birth, I divide the sample into four groups depending on their parents’ ethnicity.
-Mothers or fathers are Hispanic if they were born in a Spanish-speaking country or Puerto Rico, and US born White if they were born in the United States.\footnote{The list of Spanish-speaking countries includes: Argentina, Bolivia, Chile, Colombia, Costa Rica, Cuba, Dominican Republic, Ecuador, El Salvador, Guatemala, Equatorial Guinea, Honduras, Mexico, Nicaragua, Panama, Paraguay, Peru, Spain, Uruguay, Venezuela.} Therefore, an observation can be the product of four types of parents:
+Mothers or fathers are Hispanic if they were born in a Spanish-speaking country or Puerto Rico, and non-Hispanic White if they were born in the United States.\footnote{The list of Spanish-speaking countries includes: Argentina, Bolivia, Chile, Colombia, Costa Rica, Cuba, Dominican Republic, Ecuador, El Salvador, Guatemala, Equatorial Guinea, Honduras, Mexico, Nicaragua, Panama, Paraguay, Peru, Spain, Uruguay, Venezuela.} Therefore, an observation can be the product of four types of parents:
 \begin{enumerate}
-\item US born father and US born mother (hereafter WW) 
-\item US born father and Hispanic mother (hereafter WH)---the comparison group and the group that is likely to have a non-Hispanic-sounding last name
-\item Hispanic father and US born mother (hereafter HW)---the focus of this study and the group most likely to have a Hispanic-sounding last name
+\item US-born father and US-born mother (hereafter WW) 
+\item US-born father and Hispanic mother (hereafter WH)---the comparison group and the group that is likely to have a non-Hispanic-sounding last name
+\item Hispanic father and US-born mother (hereafter HW)---the focus of this study and the group most likely to have a Hispanic-sounding last name
 \item Hispanic father and Hispanic mother (hereafter HH).
@@
 For more discussion of this phenomenon see \textcite{antmanEthnicAttritionObserved2016,antmanEthnicAttritionAssimilation2020}} To maintain analytical clarity, I restricted the sample to Hispanic and non-Hispanic White individuals, excluding other racial groups who also identify as Hispanic, to avoid confounding racial factors.
+
+Regarding the classification of US-born parents as 'White'.
+While the United States has a long history of immigration, the probability that a US-born parent in this sample is a second-generation or later immigrant from a Spanish-speaking country is very low, as shown in Table \ref{tab:mat3}.
+The majority of Hispanics in the US during this period were first- and second-generation immigrants \autocite{antmanEthnicAttritionAssimilation2020}.
+Only 3\% of native-born Americans identified as Hispanic during this period, making it statistically unlikely that an interethnic child with a native-born parent is also a second-generation Hispanic immigrant.
+While this approach limits the generalizability of the findings to the broader Hispanic population, it provides a cleaner identification of surname effects, which is the primary focus of this study.
 
@@
 Even though the census asks a person whether they are Hispanic or not, I took advantage of the questions on place of birth to create a proxy for ethnicity.
-I consider a Hispanic persons that self-report White as their racial group and are born in a Spanish-speaking country.
-Consequently, US born in the sample are those that self-report White as race and are native-born.
+I consider a Hispanic persons who self-report White as their racial group and are born in a Spanish-speaking country.
+Consequently, US born in the sample are those who self-report White as race and are native-born.
 Using the information provided in the census, I can link husbands and wives with each other.
@@
 
-For example, consider someone who was 35 years old in 1999, meaning they were born in 1964.
-If this person's mother was born in Mexico and their father was born in the United States, their ``synthetic parents'' would be identified using the 1970 Census, when this person was 6 years old.
-These ``synthetic parents'' are created by matching the characteristics of Mexican-American couples who had a 6-year-old child in 1970.
-The ``synthetic mother'' would have the average education level, income, and other socioeconomic characteristics of Mexican-born women who were married to US-born men and had children around 1964, when they were between 20 and 35 years old (meaning they were born between 1929 and 1944).
-The ``synthetic father'' would have the average characteristics of US-born men who were married to Mexican-born women and had children in that same year.
-This method preserves the demographic composition of parents while assigning them historically accurate socioeconomic characteristics based on their origins and marriage patterns.
+To illustrate the construction of synthetic parents more concretely, consider someone who was 35 years old in 1999, meaning they were born in 1964.
+If this person's mother was born in Mexico and their father was born in the United States, their 'synthetic parents' would be identified using the 1970 Census data, when the person was 6 years old.
+The 'synthetic mother' would have the average characteristics (education, income, etc.) of Mexican-born women who were married to US-born men and had children around 1964, when they were between 20 and 35 years old (meaning they were born between 1929 and 1944).
+Similarly, the 'synthetic father' would have the average characteristics of US-born men who were married to Mexican-born women and had children in that same year.
+
+This aggregation process introduces some heterogeneity, as individual parents vary considerably from these group-level measures.
+However, this approach provides valuable information about typical family background characteristics that would otherwise be completely unobserved.
+The synthetic parents' characteristics serve as useful proxies for the socioeconomic environment in which these children were raised, allowing me to control for important background factors that influence educational and labor market outcomes.
+While more refined matching would be ideal, the publicly available data limit the potential matching dimensions.
+Nevertheless, the current approach provides meaningful estimates while acknowledging these data constraints.
 
@@
 Let $Y_{ist}$ be the outcome of interest for person $i$ in state $s$ at time $t$.
-$HW_{ist}$ is an indicator variable for the type of parents of person $i$.
-$X_{ist}$ is a vector of controls that includes age and numbers of hours worked, $\gamma_{st}$ are state-year fixed effects (FE), and $\phi_{ist}$ represents the error term.\footnote{I also include parental characteristics of `synthetic' parents as controls in alternative specifications.
-These controls include income and education of `synthetic' parents.} The equation for this strategy is written as follows and the sample is restricted to I restrict the sample to WH and HW groups:
+$HW_{ist}$ is an indicator variable equal to 1 if person $i$ has a Hispanic father and US-born mother, and 0 if person $i$ has a US-born father and Hispanic mother.
+$X_{ist}$ is a vector of controls that includes age and hours worked, $\gamma_{st}$ represents state-year fixed effects, and $\phi_{ist}$ is the error term.
+The equation for this strategy is written as follows, and the sample is restricted to individuals from WH and HW families:
+
 \begin{equation} \label{eq:1a}
@@
 $\beta_{1}$ is the coefficient of interest in this specification.
-$\beta_{1}$ represents the gaps in outcomes between children of inter-ethnic marriages who likely have a Hispanic-sounding last name versus a White-sounding last name.
-If $\beta_{1} > 0$, then people with a Hispanic last name have better outcomes than people with a White last name.
-If $\beta_{1} < 0$, then people with a Hispanic last name have worse outcomes than people with a White last name.
+$\beta_{1}$ represents the gap in outcomes between children of inter-ethnic marriages who likely have a Hispanic-sounding last name versus those who likely have a White-sounding last name.
+If $\beta_{1} > 0$, then people who likely have a Hispanic last name have better outcomes than people who likely have a White last name.
+If $\beta_{1} < 0$, then people who likely have a Hispanic last name have worse outcomes than people who likely have a White last name.
 
-By comparing interethnic children, I can control for many unobservable family characteristics that may affect labor market outcomes, since these families are likely to be more similar to each other than to families with two Hispanic or two non-Hispanic White parents.
+By comparing children from interethnic marriages, I can control for many unobservable family characteristics that may affect labor market outcomes, since these families are likely to be more similar to each other than to families with two Hispanic or two non-Hispanic White parents.
 This approach allows me to isolate the effect of surname, providing a more accurate estimate of its impact.
 The difference in means between Hispanic and non-Hispanic Whites could result from discrimination, but it could also be due to differences in innate abilities, skills, and parental investments .
-WH children will have a non-Hispanic last name, while HW children will have a Hispanic last name.
-This method, developed by \textcite{rubinstein2014pride}.
-This is an improves traditional Oaxaca-Blinder-Kitagawa decomposition methods, which are limited in their ability to account for unobserved differences between groups, by providing a more suitable comparison group to measure discrimination \autocite{oaxaca1973male,blinder1973wage,kitagawa1955components}.
+WH children will likely have a non-Hispanic last name, while HW children will likely have a Hispanic last name.
+This method was developed by \textcite{rubinstein2014pride}.
+This approach improves upon traditional Oaxaca-Blinder-Kitagawa decomposition methods, which are limited in their ability to account for unobserved differences between groups, by providing a more suitable comparison group to measure discrimination \autocite{oaxaca1973male,blinder1973wage,kitagawa1955components}.
 
@@
 
-Two primary reasons underscore this assumption.
-First, there is a significant selection in the marriage market.
-Since belonging to the Hispanic out-group brings negative societal effects, it is worth exploring factors common between White women that have children with Hispanic men.
-Second, fathers and mothers influence human capital accumulation differently \autocite{kimball2009risk,magruder2010intergenerational}.
+Two primary reasons underscore this assumption: 
+\begin{itemize}
+    \item First, there is significant selection in the marriage market.
+Since belonging to the Hispanic out-group may bring negative societal effects, it is worth exploring factors common among White women who have children with Hispanic men.
+    \item Second, fathers and mothers influence human capital accumulation differently \autocite{kimball2009risk,magruder2010intergenerational}.
 If marriage were random, then WH households might be a better environment for children.
 Using the CPS and the US Census, I will evaluate these empirical concerns.
+\end{itemize}
 
-A large body of empirical and theoretical work indicates that marriages exhibit strong selection, or "assortative matching." \autocite{averettBetterWorseRelationship2008, averettEconomicRealityBeauty1996, beckerTheoryMarriagePart1973, beckerTheoryMarriagePart1974, beckerTreatiseFamily1993, browningCollectiveUnitaryModels2006, chiapporiFatterAttractionAnthropometric2012}.
+A large body of empirical and theoretical work indicates that marriages exhibit strong selection, or `assortative matching' \autocite{averettBetterWorseRelationship2008, averettEconomicRealityBeauty1996, beckerTheoryMarriagePart1973, beckerTheoryMarriagePart1974, beckerTreatiseFamily1993, browningCollectiveUnitaryModels2006, chiapporiFatterAttractionAnthropometric2012}.
 Also, \textcite{duncanIntermarriageIntergenerationalTransmission2011} show a similar pattern of assortative matching among Mexicans in the United States.
@@
 It is possible that a native-born father could be a second-generation or later immigrant from a Spanish-speaking country, but this is unlikely.
-Most Hispanics from 1960 to 2000  were first-generation immigrants, and the number of second-generation or later was very small (Table \ref{tab:mat3}).\footnote{Based on the author's calculations using US Census data.} Only 3\% of native-born Americans identified as Hispanic, making it unlikely that an interethnic child with a native-born father is also a second generation Hispanic immigrant.
+Most Hispanics from 1960 to 2000,  were first-generation immigrants, and the number of second-generation or later was very small (Table \ref{tab:mat3}).\footnote{Based on the author's calculations using US Census data.} Only 3\% of native-born Americans identified as Hispanic, making it unlikely that an interethnic child with a native-born father is also a second generation Hispanic immigrant.
+
+It is important to address potential statistical inference issues related to using synthetic parents' characteristics as control variables.
+Regarding the 'generated regressor' issue: The parental characteristics are included solely as controls and not for inference purposes.
+I do not interpret these coefficients, as my primary interest lies in $\beta_1$, which captures the gap between Hispanic and White last names.
+Since I am only using these synthetic characteristics as controls and not making inference about their coefficients, this should not affect the standard errors or p-values of my coefficient of interest $\beta_1$, which is not a generated regressor.
+
+The synthetic parents are constructed using the birth year of the child and are not based on actual parents in the CPS sample at the time of the survey.
+The place of birth of parents—mothers and fathers—are questions asked of all participants in the Current Population Survey (CPS) starting 1994.
+Therefore, I do not use the information of parents who are in the CPS sample at the time of the survey, but rather construct synthetic parents from historical Census data.
+This approach means that concerns about non-random attrition of parents from the CPS sample do not apply to my analysis.
+
+Nevertheless, measurement error may still arise from the use of group averages that may not perfectly represent the actual parents of each individual in my sample.
+However, using group means typically reduces measurement error compared to individual-level measures.
+Since these variables are used as controls rather than variables of primary interest, any remaining measurement error is unlikely to substantially bias the coefficient of interest ($\beta_1$) on Hispanic last name, though it might reduce the precision of the estimates.
+
+I cluster standard errors at the state level.
+This approach accounts for correlation in the error terms within geographical regions, which is likely to capture much of the potential correlation structure.
+This clustering approach is common in the literature and helps ensure that the statistical significance of the results is not overstated due to correlated errors.
 
@@
 These unobservable differences could include parenting styles, educational preferences, time investments in children, gender norms, or cultural attitudes that are not captured in conventional socioeconomic metrics.
+
 However, this approach offers distinct advantages over traditional Oaxaca-Blinder-Kitagawa decomposition methods.
@@
 HW husbands have 10.33 years, while wives have an average of 11.01, giving HW households a total of 21.50 years.
-Both  Hispanic men and women in interethnic couples marry white spouses that are, on average, more educated than members of the average HH couple.
+Both  Hispanic men and women in interethnic couples marry white spouses who are, on average, more educated than members of the average HH couple.
 More importantly for human capital accumulation, the mother in the HW marriages (those with children most likely to have a Hispanic-sounding last name—are more educated than their WH peers.) Since mothers could be more important in the child’s human capital accumulation and education, this would suggest that a child with a Hispanic last name should complete more years of education  \autocite{gould2020does}.
@@
 
-I provide the results of the estimation of equation \ref{eq:1a} in Table \ref{tab:lastname-ed-reg}.
-I estimate the mean educational outcomes of White, US-born Hispanics ages 25-40.
-I also restrict the sample to children of HW and WH parents.
-The omitted group is children of WH parents.
+I present the results from estimating equation \ref{eq:1a} in Table \ref{tab:lastname-ed-reg}.
+I estimate the mean educational outcomes of White Hispanic, US-born individuals ages 25-40.
+I also restrict the sample to children of HW and WH parents, and the omitted group is children of WH parents.
 Column 1 in Table \ref{tab:lastname-ed-reg} is the difference in total years of education between HW children and their WH peers.
@@
 
-Although there is a small but significant gap in total years of education between HW and WH children, there is no significant difference in the probability of dropping out of high school.
+Although there is a modest yet statistically significant gap in total years of education between HW and WH children, there is no significant difference in the probability of dropping out of high school.
 The gap between HW and WH high school dropouts is statistically insignificant (1 percentage point).
@@
 Notable differences emerge for higher education outcomes.
-HW children are 2 percentage points less likely to earn an associate degree and 3 percentage points less likely to earn a bachelor’s degree compared to their WH peers.
 HW children are 2 percentage points less likely to earn an associate degree compared to their WH peers, representing a 13.3\% reduction relative to the WH associate degree rate of 15\%.
-HW children are 3 percentage points less likely to earn a bachelor’s degree compared to their WH peers, a 13\% reduction compared to the WH bachelor’s degree rate of 23\%.
+Similarly, they are 3 percentage points less likely to earn a bachelor's degree, a 13\% reduction compared to the WH bachelor's degree rate of 23\%.
 These differences are slightly larger for HW women, who are 3 percentage points less likely to earn an associate degree and 4 percentage points less likely to earn a bachelor’s degree.
@@
 
-I provide the results of the estimation of equation \ref{eq:1a} in Tables \ref{tab:lastnamereg-emp} and \ref{tab:lastnamereg} on unemployment and log earnings.
+I provide the results of the estimating equation \ref{eq:1a} in Tables \ref{tab:lastnamereg-emp} and \ref{tab:lastnamereg} on unemployment and log earnings.
 I estimate the mean unemployment and mean wages of White US-born Hispanic men aged 25-40 who are employed full-time.
@@
 
-I also analyzed the effect of having a Hispanic last name on unemployment rates.
+I also analyzed the effect of likely having a Hispanic last name on unemployment rates.
 Table \ref{tab:lastnamereg-emp} presents the results of this analysis.
-Column 1 shows that individuals with Hispanic last names (HW) have a 1 percentage point higher unemployment rate compared to those with White last names.
+Column 1 shows that individuals who likely have Hispanic last names (HW) have a 1 percentage point higher unemployment rate compared to those with White last names.
 This discrepancy persists even after controlling for age, state FE, year FE, and state-year FE (Column 2), though the significance is lower.
@@
 Finally, after controlling for parental background (Column 4), the 1 percentage point difference in unemployment rates becomes statistically insignificant.
-These results suggest that while there is an initial unemployment gap associated with having a Hispanic last name, much of this difference can be explained by factors such as education and parental background and does not necessarily reflect discrimination.
+These results suggest that while there is an initial unemployment gap associated with likely having a Hispanic last name, much of this difference can be explained by factors such as education and parental background and does not necessarily reflect discrimination.
 The mean unemployment rate for individuals with Hispanic last names (HW) is 7\% across all specifications.
@@
 Overall, the crude gap between HW and WH workers is equal to 5 percentage points (Table \ref{tab:lastnamereg} column 1).
-An interethnic worker with a Hispanic last name earns 5 percentage points less than an interethnic worker with a White last name.
+An interethnic worker who likely has a Hispanic last name earns 5 percentage points less than an interethnic worker with a White last name.
 Even after controlling for hours worked, and including state, year, and age FEs in the estimation, the gap stays at 5 percentage points; however, the difference could be attributed to educational differences.
@@
 First, in Table \ref{tab:lastnamereg-weekearm}, I present the results of estimating equation \ref{eq:1a} using weekly earnings as the dependent variable for the full sample.
-Like the previous analysis, I find that a person with a Hispanic last name earns 4 percentage points less than a person without a Hispanic last name, which could be explained by educational differences.
+Like the previous analysis, I find that a person who likely has a Hispanic last name earns 4 percentage points less than a person who likely has a non-Hispanic last name, which could be explained by educational differences.
 Second, in Table \ref{tab:lastnamereg-weekearm-mex}, I present the results for a sample of Mexican Hispanics.
@@
 
-I I also present the results of the estimation of equation 1 using log annual earnings as the dependent variable but with occupation fixed effects instead of education fixed effects (Table \ref{tab:lastnamereg-occ}).
+I also present the results of the estimation of equation 1 using log annual earnings as the dependent variable but with occupation fixed effects instead of education fixed effects (Table \ref{tab:lastnamereg-occ}).
 I find that the gap between people with a likely Hispanic-sounding last names and those with a likely White-sounding last names is explained by differences in occupation, similar to my findings with educational controls.
@@
 
-As the Hispanic population grows in the United States, studying discrimination becomes important.
-In this paper, I examine discrimination against Hispanics in the labor market.
-More specifically, I examine the impact of Hispanic last names and Hispanic identification on annual log earnings.
-I compare the children of interethnic marriages to study the labor and educational market effects of having a Hispanic last name and find that interethnic people with likely Hispanic-sounding last names receive 0.2 fewer years of education than interethnic counterparts that likely have a White-sounding  last name, representing a 1.4\% reduction.
+As the Hispanic population grows in the United States, studying discrimination becomes increasingly important.
+In this paper, I examine discrimination against Hispanics in the labor market, specifically focusing on the impact of Hispanic last names on educational and labor market outcomes.
+
+By comparing children of interethnic marriages, I find that individuals who likely have Hispanic-sounding last names receive 0.2 fewer years of education than their counterparts who likely have White-sounding last names, representing a 1.4\% reduction.
 This gap is larger for women (0.25 years, or 1.8\%) than for men (0.16 years, or 1.2\%).
-When I compare the labor market outcomes of HW and WH children, which captures the effect of having a Hispanic last name, HW children are 1 percentage point more likely to be unemployed than WH children, representing a 14.3\% increase from the mean unemployment rate.
-In terms of earnings, HW children (likely to have a Hispanic-sounding surname earn 5 percentage points (4.9\%) less than WH children (likely to have a non-Hispanic-sounding name).
-When I control for education, the last name effect decreases to a statistically insignificant 1 percentage point earnings gap.
-The findings reveal gaps that could be attributed to discrimination against individuals with Hispanic surnames, with effects varying across educational levels and gender.
-While the impact on total years of education appears modest (1.4\% reduction), the more substantial gaps in higher education attainment—particularly the 13\% reduction in bachelor’s degree completion and 13.3\% reduction in associate degree attainment—suggest that discrimination may intensify at crucial educational transition points.
-This pattern is especially pronounced for women with Hispanic surnames, who face larger discrepancies in both total education (1.8\% reduction versus 1.2\% for men) and degree completion (14.8\% reduction in bachelor’s degree attainment ).
-This paper contributes to a body of education economics literature showing disparities in access to education.
-In an audit study of charter and traditional public schools, \textcite{bergman2018education,gaddis2024racial} find that students  with Hispanic names—compared to students that are presumed White—are less likely to get a response from schools they contact for application information.
-This suggests that the earnings gap between people with a likely Hispanic sounding last names and those with a native  sounding last names could be due
+In labor market outcomes, those with likely Hispanic surnames are 1 percentage point more likely to be unemployed (a 14.3\% increase from the mean rate) and earn 5 percentage points (4.9\%) less than their counterparts.
 
-To differences in access to education that are in turn due to discrimination.
-This paper provides further evidence, using observational data, that discrimination in access to education could lead to lower earnings for Hispanic workers.
-In the labor market, the 14.3\% higher unemployment rate and 4.9\% lower salaries for those with Hispanic surnames indicate persistent discrimination even among individuals who share similar family backgrounds.
-Putting the results in context, \autocite{trejoWhyMexicanAmericans1997} found an earnings gap between US-born Hispanics and US-born whites that is to 30.7 percentage points.
-My findings could be an indication that 16\% of this gap could be attributed to discrimination against Hispanics.
-However, the fact that the earnings gap becomes statistically insignificant after controlling for education suggests that much of the labor market discrimination operates through educational channels rather than direct wage discrimination.
-These findings leave several important questions unanswered.
+While the overall impact on total years of education appears modest, the disparities become more pronounced in higher education attainment—with a 13.3\% reduction in associate degree completion and 13\% reduction in bachelor's degree attainment.
+These patterns are especially pronounced for women with Hispanic surnames, who face larger discrepancies in both total education and degree completion (14.8\% reduction in bachelor's degree attainment).
 
-The earnings gap between individuals with vs without a Hispanic last name disappears when controlling for education, but this does not necessarily indicate the absence of discrimination.
-Education itself can be influenced by bias, potentially resulting in divergent outcomes \autocite{bergman2018education,gaddis2024racial}.
-This is especially the case when parental characteristics indicate that people with a likely Hispanic-sounding last-names should in theory complete more years of education, since mothers of interethnic children with likely Hispanic surnames have more years of education and earn more.
-While I observe where discrimination manifests (particularly in higher education access), my analysis cannot identify the specific mechanisms—such as teacher bias, institutional barriers, or social networks—through which surname-based discrimination operates.
+This paper contributes to the education economics literature by providing evidence of disparities in access to education.
+Previous audit studies by \textcite{bergman2018education} and \textcite{gaddis2024racial} found that students with Hispanic names are less likely to receive responses from schools compared to those presumed White.
+My findings suggest that discrimination in access to education could lead to lower earnings for Hispanic workers.
+
+However, this does not necessarily indicate the absence of discrimination.
+Education itself can be influenced by bias, particularly since mothers of interethnic children with likely Hispanic surnames have higher education levels and earnings than mothers of children with non-Hispanic surnames—suggesting these children should theoretically complete more years of education.
+While my analysis identifies where discrimination manifests (primarily in higher education access), it cannot determine the specific mechanisms—such as teacher bias, institutional barriers, or social networks—through which surname-based discrimination operates.
 Further research is needed to understand these disparities.

### 2025-05-17 manuscripts/10-JERP-v3/manuscript_title_etc.tex
some edits

--- 2025-03-17 manuscripts/08-JERP-v6/manuscript_title_etc.tex
@@
 Do individuals with Hispanic surnames face labor market discrimination?
-This study analyzes the impact of Hispanic-sounding surnames on wages among inter-ethnic Americans with one White and one Hispanic parent.
+This study analyzes the impact of likely Hispanic surnames on wages among Americans with one White and one Hispanic parent.
 I find that individuals that likely have Hispanic surnames (i.e., those with a Hispanic father) often earn less, with a notable wage gap favoring those with White-sounding surnames.
-People born to Hispanic fathers and White mothers receive 0.2 fewer years of education than men born to White fathers and Hispanic mothers.
+People born to Hispanic fathers and White mothers receive 0.2 fewer years of education than those born to White fathers and Hispanic mothers.
 Men born to Hispanic fathers and White mothers are 1 percentage point more likely to be unemployed, and they earn 5 percentage points less than those born to White fathers and Hispanic mothers.

### 2025-06-13 manuscripts/10-JERP-v3/manuscript_main.tex
edit manuscript

@@
 
-A large body of literature provides evidence of substantial earnings gaps across race and ethnicity in the US \autocite{bayer2018divergent, charles2008prejudice}.
-Hispanic individuals constitute a large and growing portion of the US population.
-As this population grows, it is crucial to determine whether ethnic discrimination affects their employment prospects, wages, and career advancement.
-Understanding these labor market outcomes is critical, as they directly affect broader societal issues like assimilation and economic mobility, \autocite{chettyUnitedStatesStill2014, chettyEffectsExposureBetter2016,chettyFadingAmericanDream2017}.
-These factors serve as key indicators of how successfully Hispanic immigrants can navigate society and ascend the socioeconomic ladder.
+Hispanic Americans represent one of the fastest growing demographic groups in the United States, making their labor market experiences increasingly central to understanding the American economy.
+Although extensive research documents persistent earnings gaps between racial and ethnic groups \autocite{bayer2018divergent, charles2008prejudice}, with native-born Hispanic White men earning 21\% less than non-Hispanic White men \autocite{duncan2006hispanics, duncan2018identifying, duncan2018socioeconomic}, the specific mechanisms affecting Hispanic workers' employment prospects, wages, and career advancement require closer examination.
+These labor market outcomes extend far beyond individual earnings—they fundamentally shape patterns of assimilation, economic mobility, and intergenerational progress \autocite{chettyUnitedStatesStill2014, chettyEffectsExposureBetter2016,chettyFadingAmericanDream2017}.
+Understanding how discrimination affects Hispanics offers valuable insight into both the obstacles they face and the possible routes for their socioeconomic advancement.
 
-In this paper, I answer the following questions.
-Does having a Hispanic last name affect educational outcomes?
-Does having a Hispanic last name affect labor market outcomes?
-I aim to show that comparing Hispanic White individuals to non-Hispanic White individuals might create an artificially higher earnings gap since these groups differ in many observable characteristics.\footnote{In this study, I define Hispanic individuals specifically as US-born persons who self-identify as White and have at least one parent born in a Spanish-speaking country.} \footnote{Observable characteristics refer to factors that can be measured and quantified, such as education level, work experience, and immigration status.} My analysis focuses on US-born children with one foreign-born parent (i.e., inter-ethnic).\footnote{This study's focus on US-born children with foreign-born parents means the findings may not generalize to Hispanic children with US-born parents, who likely face different socioeconomic and cultural circumstances.
-The analysis does not account for heterogeneity in immigrant characteristics across different Spanish-speaking countries of origin, such as variations in educational attainment, socioeconomic status, and gender-specific migration patterns, which could influence both parental selection into migration and subsequent child outcomes.} Others have attempted to compare how native-born White Hispanics fare compared to non-Hispanic Whites and foreign-born Hispanics.
-\cites{antman2020ethnic} have compared the health and educational outcomes of Hispanic Whites to non-Hispanic Whites and native-born Hispanics to foreign-born Hispanics.
-They find gaps in education and health between Hispanics and non-Hispanic Whites.
-They also find that native-born Hispanics are more likely than their foreign-born counterparts to report poor health.\footnote{For additional studies examining ethnic identity and outcomes among Hispanic populations, see \textcite{antman2020ethnic,antmanEthnicAttritionObserved2016,antmanEthnicAttritionObserved2016a,antmanEthnicAttritionAssimilation2020}, which collectively document the relationships between ethnic identification, assimilation patterns, and various socioeconomic outcomes.} \textcite{davilaChangesRelativeEarnings2008} found gaps in labor market outcomes between Hispanics and non-Hispanic Whites, which they attribute largely to differences in education, experience, immigration status, and regional differences.
-This paper builds on those studies by emphasizing how cultural assimilation and generational status further shape educational and labor market disparities for Hispanics.
+This paper examines whether having a Hispanic last name affects the educational and labor market outcomes of native-born Hispanics.
+I introduce a novel empirical strategy comparing children from interethnic families—those with likely Hispanic-sounding last names (HW) with those with White-sounding last names (WH).
+Specifically, I focus on US-born children with one foreign-born parent to isolate the effect of surname-based discrimination from other confounding factors.\footnote{Following the approach of \textcite{antman2020ethnic, hadah2024effect}, I define Hispanic individuals as people born in the US who self-identify as White and have at least one parent born in a Spanish-speaking country.
+This definition helps mitigate issues related to ethnic attrition.
+As demonstrated by \textcite{hadah2024effect}, ethnic attrition, particularly when associated with state-level bias, can result in an overestimation of the Hispanic-White gap in the most biased states.
+Additionally, focusing on White Hispanics provides cleaner estimates of discrimination based on Hispanic ethnicity by avoiding the confounding effects of racial discrimination that would complicate the interpretation of results.} This comparison leverages the similarity of family backgrounds in interethnic marriages to identify discrimination's role in shaping economic outcomes.
+Simple comparisons between Hispanic and non-Hispanic groups overstate earnings gaps, because these populations differ systematically in education, work experience, immigration status, and other observable characteristics.\footnote{Observable characteristics refer to factors that can be measured and quantified, such as education level, work experience, and immigration status.} By examining individuals who share similar backgrounds but differ primarily in their surnames, I can better isolate the impact of ethnic signals on educational and labor market outcomes.\footnote{This study's focus on US-born children with foreign-born parents means the findings may not generalize to Hispanic children with US-born parents, who likely face different socioeconomic and cultural circumstances.
+The analysis does not account for heterogeneity in immigrant characteristics in different Spanish-speaking countries of origin, such as variations in educational attainment, socioeconomic status, and gender-specific migration patterns, which could influence both parental selection in migration and subsequent child outcomes.}
 
-Understanding discrimination against Hispanics in labor markets has far-reaching implications that extend beyond individual earnings.
-Labor market discrimination can create persistent intergenerational disadvantages by limiting economic mobility, reducing access to quality healthcare and education, and constraining residential choices; \autocite{chettyUnitedStatesStill2014, chettyEffectsExposureBetter2016,chettyFadingAmericanDream2017,bowles2002inheritance, djajic2003assimilation}.
-Discrimination may foster occupational sorting and segregation, as shown by recent task-based models of racial wage gaps that identify race-specific barriers in “Contact” tasks and explain the stagnation of these gaps post-1980 \autocite{hurst2024task}.
-These barriers may trap Hispanic families in cycles of lower socioeconomic status, as reduced earnings and employment opportunities limit their ability to invest in their children's human capital or build wealth through homeownership and savings.\footnote{See \textcite{bowles2002inheritance} for an analysis of the intergenerational transmission of economic status.}
+This approach builds on \textcite{rubinstein2014pride}, who compared children of mixed Ashkenazi-Sephardic marriages in Israel, finding substantial wage penalties associated with Sephardic surnames.
+Previous work by \textcite{antman2020ethnic} and \textcite{davilaChangesRelativeEarnings2008} has shown that differences in education, experience, and immigration status explain much but not all of the Hispanic-White gap, suggesting discrimination may play a role.
+\textcite{antman2020ethnic} documents significant ethnic attrition among Hispanics and finds persistent gaps in education and health between native-born Hispanics and non-Hispanic Whites, with native-born Hispanics more likely to report poor health than their foreign-born counterparts.\footnote{For additional studies examining ethnic identity and outcomes among Hispanic populations, see \textcite{antman2020ethnic, antmanEthnicAttritionObserved2016, antmanEthnicAttritionObserved2016a, antmanEthnicAttritionAssimilation2020}, which collectively document the relationships between ethnic identification, assimilation patterns, and various socioeconomic outcomes.}
 
-Although previous research has identified earnings gaps between Hispanic and non-Hispanic Whites, this paper makes several important contributions.
-First, it develops a novel empirical strategy that better isolates the causal effect of Hispanic ethnicity on educational and labor market outcomes by comparing children from inter-ethnic families.
-This approach helps control for typically unobservable family background characteristics that might confound traditional analyses.
-Second, it provides new evidence on the specific role of Hispanic surnames in driving discrimination, offering insights into how ethnic signals influence educational attainment and labor market outcomes.
-Finally, by examining both education and employment outcomes, this study helps illuminate the channels through which ethnic discrimination may perpetuate economic disparities across generations and extends earlier findings by demonstrating how naming cues can compound structural inequalities and reinforce existing barriers.
+Discrimination against Hispanics in the US labor market undermines economic mobility and opportunity, reducing earnings and narrowing pathways to educational and professional advancement.
+These effects extend beyond wages to affect access to quality health care, residential choice, and occupational sorting \autocite{chettyUnitedStatesStill2014, hurst2024task}, potentially perpetuating inequality across generations.
+This paper's methodological improvement is particularly timely given the rapidly changing demographics of the United States.
+The proportion of non-Whites has increased by more than 10 percentage points from 13 percent in 1995 to 23 percent in 2019,\footnote{The proportion of non-Whites and Hispanics is based on the author's calculations using the Current Population Survey (CPS).} making it crucial to understand how ethnic signals affect labor market outcomes in an increasingly diverse society.
 
-This study builds and improves upon the traditional Oaxaca-Blinder-Kitagawa decomposition approach often used in labor market discrimination studies \autocite{kitagawa1955components, oaxaca1973male,blinder1973wage}.
-While these decomposition methods have provided valuable insights into earnings gaps between groups, including Hispanics and non-Hispanic Whites \autocite{davilaChangesRelativeEarnings2008}, they are limited in their ability to account for unobserved differences between the groups.
-My approach comparing children of inter-ethnic marriages, provides a more closely matched comparison group, allowing me to better isolate the effect of likely having a Hispanic last name on labor market outcomes.
+My empirical strategy relies on one key identifying assumption: children of Hispanic-father/White-mother couples (HW) and children of White-father/Hispanic-mother couples (WH) are alike on every trait that matters in the labor market, both observable and unobservable.
+If this holds, surname is the only systematic difference between them.
+This approach accounts for the fact that couples are likely to have similar income, schooling, and socioeconomic background, so these factors are largely controlled for \autocite{averettBetterWorseRelationship2008, averettEconomicRealityBeauty1996}.\footnote{For more on assortative mating see \textcite{beckerTheoryMarriagePart1973, beckerTheoryMarriagePart1974, beckerTreatiseFamily1993, browningCollectiveUnitaryModels2006, chiapporiFatterAttractionAnthropometric2012}} Indeed, children of HW and WH marriages have more similar observable characteristics than children of endogamous/homogamous marriages (that is, White-fathers/White-mothers and Hispanic fathers/Hispanic mothers), as I demonstrate in Section \ref{sec:hw-wh-couples-data}.
+It is probably true that offspring from interethnic unions have more comparable parental and cultural backgrounds with one another than those from endogamous Hispanic or White marriages do among themselves.
+In the US, children from households with a Hispanic father and White mother are overwhelmingly likely to have their father's Hispanic last name, allowing me to investigate how this ethnic signal affects labor market and educational outcomes \autocite{davenport2016role}.
 
-The methodological improvement in this paper is particularly timely and relevant given the rapidly changing demographics of the United States.
-The US population is growing increasingly diverse, with significant implications for labor market dynamics and potential discrimination.
-The proportion of non-Whites has increased by more than 10 percentage points from 13 percent in 1995 to 23 percent in 2019.\footnote{The proportion of non-Whites and Hispanics is based on the author's calculations using the Current Population Survey (CPS).} Native-born White Hispanic men earn 21\% less than non-Hispanic White men, although there is evidence that a substantial portion of this gap is due to differences in education \autocite{duncan2006hispanics, duncan2018identifying, duncan2018socioeconomic}.
-These discrepancies may also stem from discrimination against Hispanics, which can lead to reduced job opportunities, lower wages, and hindered assimilation \autocite{chettyWhereLandOpportunity2014,bowles2002inheritance, djajic2003assimilation}.\footnote{This study focuses on US-born children of inter-ethnic unions, many of whom have at least one foreign-born parent.
-Assimilation concerns remain relevant because foreign-born parents may pass on cultural norms and language preferences, which can still shape these children's educational and labor market outcomes.}
+My surname-based approach complements the audit study literature, which has documented discrimination but cannot observe actual wages.
+While \textcite{bertrand2004emily} found that Black-sounding names received fewer callbacks, and education audit studies show 2-11 percentage point penalties for Hispanic names in school responses \autocite{bergman2018education, gaddis2024racial}, my method reveals the downstream consequences of such discrimination on actual earnings and educational attainment.
+\textcite{fryer2004causes} established that names can be a predictor of a person's race, though they found that having a Black-sounding name, after controlling for the home environment at birth, does not affect labor market outcomes.
+These studies highlight the importance of understanding how such signals can capture both conscious and unconscious bias.
+The education audit experiments are particularly relevant, as discrimination can affect educational access through multiple channels: restricted access to advanced coursework as found by \textcite{janssen2022guidance} with Asian students, biased academic counseling, and subtle institutional barriers documented by \textcite{bourabain2023school} in the Flemish education system.
+While Hispanic college enrollment has increased substantially in recent decades, discrimination within educational institutions may still impede degree completion, ultimately affecting labor market outcomes.
 
- Using this novel approach, I find significant effects of Hispanic surnames on both educational and labor market outcomes.
-Individuals who likely have Hispanic last names face substantial disadvantages in educational attainment and earnings compared to their counterparts with likely non-Hispanic names.
-Specifically, individuals with presumably Hispanic last names complete 0.2 fewer years of education than those with non-Hispanic names, even when controlling for family background.
-Individuals with presumably Hispanic last names are also 1 percentage point more likely to be unemployed and earn 5 percentage points less than those with likely non-Hispanic names.
-These findings suggest that ethnic discrimination continues to play a role in shaping economic opportunities for Hispanics in the United States, highlighting the need for targeted policies to address these persistent disparities.
+As gender and ethnic discrimination operate through distinct mechanisms, I separate out a portion of my results based on gender.
+As \textcite{bertrand2020gender} notes, women's outcomes in the labor market are shaped by persistent norms around caregiving and motherhood.
+\textcite{antecol2002relative} found that in 1994, young Mexican women earned 9.5\% less than young White women in the US, with differences in education largely explaining the gap.
+\textcite{goldin2004making} emphasizes the growing importance of professional identity and career continuity for women, while \textcite{darity2015tour} show how group-based hierarchies are perpetuated through both discrimination and intergenerational transfers.
+The breakdown of my results by gender therefore reveals how ethnic and gender biases intersect to shape educational and economic opportunities.
 
-Factors that affect labor market outcomes, like skills and stereotypes, are unobservable to economists, so it is challenging to identify discrimination.
-One strategy used by researchers is audit or resume studies.
-\textcite{bertrand2004emily} conducted an audit study by sending employers identical resumes that differed only in the ethnic and racial signal of the applicant’s name (Black-sounding versus White-sounding).
-They found that resumes with Black-sounding names received substantially fewer callbacks than their White counterparts.
-Audit studies only observe callbacks, not wages.
-The  current study uses a method developed by \textcite{rubinstein2014pride}.
-The authors compared the children of mixed marriages between Sephardic and Ashkenazi people in Israel.\footnote{Ashkenazi and Sephardic are two distinct Jewish ethnic groups.} They found that workers with Sephardic last names earn substantially less than those with Ashkenazi last names.
-I compare children of Hispanic fathers and US born mothers (henceforth HW) and likely have a Hispanic-sounding last name to children of US born fathers and Hispanic mothers (WH).
-This approach accounts for the fact that  couples are likely to have similar income, schooling, and socioeconomic background, so these factors are largely controlled for \autocite{averettBetterWorseRelationship2008, averettEconomicRealityBeauty1996}.\footnote{For more on assortative mating see \autocite{beckerTheoryMarriagePart1973, beckerTheoryMarriagePart1974, beckerTreatiseFamily1993, browningCollectiveUnitaryModels2006, chiapporiFatterAttractionAnthropometric2012}} Children of HW and WH marriages have more similar observable characteristics than children of endogamous/homogamous marriages (i.e., US born fathers-US born mothers and Hispanic fathers-Hispanic mothers).
-In the US, children  from households with a Hispanic father and White mother  are overwhelmingly likely to  have their father’s Hispanic last name, allowing us to investigate how this ethnic signal affects annual log earnings.
-
-The choice to include separate results for men and women is motivated by evidence that discrimination based on sex/gender and ethnicity operate through distinct mechanisms.
-As \textcite{bertrand2020gender} discusses, women's labor market outcomes are shaped by persistent gender norms and stereotypes, particularly around motherhood and caregiving.
-\textcite{antecol2002relative} found that young Mexican women earned 9.5\% less than young white women in the US in 1994, and like the current study, differences in education appeared to be the most likely explanation for this gap.
-For Black women, differences in labor force attachment appear to be the key driver of wage disparities.
-The way discrimination manifests can also evolve differently by gender over time: \textcite{goldin2004making} found that women's professional identity and career continuity are increasingly important factors in labor market outcomes, while \textcite{darity2015tour} argue that a stratification economics approach helps reveal how group-based hierarchies are maintained through both discriminatory practices and intergenerational resource transfers.
-Disaggregating the results by gender provides important insights into how the intersection of ethnic and gender discrimination shapes educational and economic opportunities.
-
-The main identifying assumption of my empirical strategy depends on the assumption that people born to HW parents are similar to their WH peer in all observable---and unobservable---aspects and characteristics relevant to the labor market, so surname is the only difference between these groups.
-Children from US-born-Hispanic homes might benefit from more favorable family conditions than those from Hispanic-US-born homes, considering Whites statistically have higher socioeconomic status than Hispanics.
-These factors introduce doubts regarding whether children from Hispanic-US-born families receive comparable familial support and influences, genetically or environmentally, as those from US-born-Hispanic families.
-
-Previous studies have also used names as a proxy for race and ethnicity \autocite{fryer2004causes, bertrand2004emily}.
-\textcite{fryer2004causes} establish that names can be a predictor of a person's race.
-They found that having a Black-sounding name, after controlling for the home environment at birth, does not affect labor market outcomes.
-These early innovations in leveraging name-based identification highlight the importance of understanding how such signals can capture both conscious and unconscious bias, placing the current study within a broader framework of research on race, names, and economic outcomes.
-
-Audit studies in education economics investigated the effect of racial and ethnic signals on access to education.
-\textcite{bergman2018education} found that students with Hispanic-sounding names received 2 percentage points fewer responses from schools than students with White sounding names.
-\textcite{janssen2022guidance} found that guidance counselors restricted Asian students from advanced opportunities.
-\textcite{gaddis2024racial} found significant discrimination in interactions between school principals and Hispanic and Chinese American families.
-Finally, \textcite{bourabain2023school} found more evidence of discrimination in access to education against students with underprivileged backgrounds in the Flemish education system.
-Discrimination can affect educational access and success through multiple channels: restricted access to advanced coursework, biased academic counseling, limited encouragement to pursue higher education, and subtle institutional barriers that may cause Hispanic students to feel unwelcome or unsupported in educational settings.
-While Hispanic college enrollment has increased substantially in recent decades, discrimination within educational institutions may still impede degree completion.
-The current paper contributes to this literature by providing evidence that discrimination in access to education could lead to lower earnings for Hispanics, and how parental background and marriage market selection can shape these outcomes in ways not previously explored by earlier studies.
-
-The rest of the paper is organized as follows.
+The remainder of the paper is organized as follows.
 In Section \ref{sec:data}, I describe the data used in this paper.
 In Section \ref{sec:emp_model}, I present the empirical strategy.
-In Section \ref{sec:results}, I present the results from the estimation of the two specifications.
+In Section \ref{sec:results}, I present the results of the estimation of the two specifications.
 Finally, in Section \ref{sec:con1}, I conclude.
@@
 
-I use the 1994–2019 CPS data set to study the effect of Hispanic surnames on labor market outcomes, because this time period contains data on parents’ place of birth, ethnicity, and race.
-The CPS does not provide data on parents’ characteristics, essential to determine the family background, but the 1960–2000 US Census data includes parents’ place of birth, race, and ethnicity.
-I employ this information to construct 'synthetic parents' using a method developed by \textcite{rubinstein2014pride} linking husbands and wives  in the census data.
-Assuming that parents have children 25–40 years old, I link these “parents” using parents’ places of birth, birth year of "children," and the parents’ places of birth.
+I use the 1994–2019 CPS data to study the effect of Hispanic surnames on labor market outcomes, as this period contains data on parents' place of birth.
+The place of birth of mothers and fathers is explicitly asked of all participants in the CPS starting 1994, and this information is essential for identifying interethnic marriages.
+Since the CPS lacks direct parental characteristics data, I construct 'synthetic parents' using the 1960–2000 US Census data, following \textcite{rubinstein2014pride}.\footnote{For more on 'synthetic' parents see \textcite{aaronson2008intergenerational}.} This method links married couples in census data based on parents' places of birth and children's birth years, assuming that parents have children between ages 25 and 40.
 
-For my main analysis, I construct different educational variables: years of education, and whether or not the person received a high school diploma, associate degree, and/or bachelor’s degree.
-For labor market outcomes, I construct an unemployment rate variable using the main CPS, log annual earnings using the ASEC, and log weekly earnings using the outgoing rotation.
-The unemployment indicator is a binary variable that equals 1 if someone is unemployed and 0 if employed, based on civilian labor force status.
-Log annual earnings is the log of total personal income from the CPS’ ASEC supplement.
-Log weekly earnings is the natural logarithm of weekly earnings for hourly workers from the outgoing rotation CPS supplement.
-I use the following controls: age, hours worked, and state-year fixed effects.
-I also include the parents’ characteristics from the synthetic parents data as controls in alternative specifications.
-These controls include income and education of the synthetic parents.
+\subsection{Sample Construction}
 
-\subsection{Children of the four parental types}
+My sample includes US citizens aged 25 to 40 years born between 1960 and 2000. Using parents' place of birth data, I classify observations into four parental types:
 
-I use the CPS for my primary analysis of the effect of a Hispanic surname on earnings.
-I restrict my sample to US citizens  aged 25–40 born between 1960 and 2000 in the US.
-Taking advantage of data on parents’ place of birth, I divide the sample into four groups depending on their parents’ ethnicity.
-Mothers or fathers are Hispanic if they were born in a Spanish-speaking country or Puerto Rico, and non-Hispanic White if they were born in the United States.\footnote{The list of Spanish-speaking countries includes: Argentina, Bolivia, Chile, Colombia, Costa Rica, Cuba, Dominican Republic, Ecuador, El Salvador, Guatemala, Equatorial Guinea, Honduras, Mexico, Nicaragua, Panama, Paraguay, Peru, Spain, Uruguay, Venezuela.} Therefore, an observation can be the product of four types of parents:
 \begin{enumerate}
-\item US-born father and US-born mother (hereafter WW) 
-\item US-born father and Hispanic mother (hereafter WH)---the comparison group and the group that is likely to have a non-Hispanic-sounding last name
-\item Hispanic father and US-born mother (hereafter HW)---the focus of this study and the group most likely to have a Hispanic-sounding last name
-\item Hispanic father and Hispanic mother (hereafter HH).
+\item \textbf{WW}: US-born father and US-born mother
+\item \textbf{WH}: US-born father and Hispanic mother (comparison group, likely non-Hispanic surname)
+\item \textbf{HW}: Hispanic father and US-born mother (treatment group, likely Hispanic surname)
+\item \textbf{HH}: Hispanic father and Hispanic mother
 \end{enumerate}
 
-My sample includes inter-ethnic children with Hispanic ancestry aged 25–40 who are US citizens.
-I include both Hispanic White and non-Hispanic White individuals in this study.
-For second-generation immigrants, I use parents' place of birth as a proxy for ethnicity.
-This approach helps avoid biased estimates that could result from ethnic attrition, as demonstrated by \textcite{hadah2024effect}.
-\footnote{A person self-reports Hispanic identity by answering the Census question: ``Is this person Spanish/Hispanic/Latino?''} \footnote{Ethnic attrition happens when a US-born descendant of a Hispanic immigrant fails to self-identify as Hispanic.
-For more discussion of this phenomenon see \textcite{antmanEthnicAttritionObserved2016,antmanEthnicAttritionAssimilation2020}} To maintain analytical clarity, I restricted the sample to Hispanic and non-Hispanic White individuals, excluding other racial groups who also identify as Hispanic, to avoid confounding racial factors.
-
-Regarding the classification of US-born parents as 'White'.
-While the United States has a long history of immigration, the probability that a US-born parent in this sample is a second-generation or later immigrant from a Spanish-speaking country is very low, as shown in Table \ref{tab:mat3}.
-The majority of Hispanics in the US during this period were first- and second-generation immigrants \autocite{antmanEthnicAttritionAssimilation2020}.
-Only 3\% of native-born Americans identified as Hispanic during this period, making it statistically unlikely that an interethnic child with a native-born parent is also a second-generation Hispanic immigrant.
-While this approach limits the generalizability of the findings to the broader Hispanic population, it provides a cleaner identification of surname effects, which is the primary focus of this study.
-
-The CPS data distribution of the four types of parenthood is presented in Table \ref{tab:mat1}.
-The vast majority (96\%) are WW children.
-The second biggest group is HH, which constitutes 3\% of the sample.
-Inter-ethnic children, WH and HW, make up 1.35\% of the sample with 90,325 observations.
-Even though WH and HW are only 1.35\% of the sample, I have plenty of observations to carry out an analysis.
-The summary statistics for the children of the four types of marriages are presented in Table \ref{tab:c&p2}.
-Children of WW marriages (Column 1) do better on every measure while children of HH parents (Column 4) do worse than other children on every measure.
-Children of WH (Column 2) and HW (Column 3) marriages fall in between WW and HH children.
-The rates of self-reported Hispanic identity vary significantly across groups.
-Among children of WW marriages (Column 1), only 4\% of men and 5\% of women identify as Hispanic.
-This proportion increases substantially for children of inter-ethnic marriages: in WH families (Column 2), 74\% of men and 78\% of women identify as Hispanic, while in HW families (Column 3), the rates are 83\% for men and 81\% for women.
-Children of HH marriages (Column 4) show the highest rates of Hispanic identification, with 96\% of men and 97\% of women self-reporting as Hispanic.\footnote{The ethnic attrition rates are similar to those found in \textcite{antmanEthnicAttritionObserved2016,antmanEthnicAttritionAssimilation2020, hadah2024effect}.}
-
-\subsection{Synthetic parents}
-
-Using the 1960 to 2000 censuses, I constructed a data set of synthetic parents.
-The sample includes married White men and women.
-Even though the census asks a person whether they are Hispanic or not, I took advantage of the questions on place of birth to create a proxy for ethnicity.
-I consider a Hispanic persons who self-report White as their racial group and are born in a Spanish-speaking country.
-Consequently, US born in the sample are those who self-report White as race and are native-born.
-Using the information provided in the census, I can link husbands and wives with each other.
-I assume that parents have children between the ages of 25 and 40, so my sample consists of married White men and women with children born in the 1920 to 1975 cohorts\footnote{The construction of ``synthetic parents'' follows the method used by \textcite{rubinstein2014pride}.}.
+Parents are classified as Hispanic if born in a Spanish-speaking country or Puerto Rico, and non-Hispanic White if US-born.\footnote{Spanish-speaking countries include: Argentina, Bolivia, Chile, Colombia, Costa Rica, Cuba, Dominican Republic, Ecuador, El Salvador, Guatemala, Equatorial Guinea, Honduras, Mexico, Nicaragua, Panama, Paraguay, Peru, Spain, Uruguay, Venezuela.} This approach helps avoid biased estimates from ethnic attrition \autocite{hadah2024effect}.\footnote{Ethnic attrition occurs when US-born descendants of Hispanic immigrants fail to self-identify as Hispanic.
+See \textcite{antmanEthnicAttritionObserved2016,antmanEthnicAttritionAssimilation2020}.}
 
@@
 This aggregation process introduces some heterogeneity, as individual parents vary considerably from these group-level measures.
-However, this approach provides valuable information about typical family background characteristics that would otherwise be completely unobserved.
+However, this approach provides valuable information about typical family background characteristics that would otherwise be completely unobserved \autocite{rubinstein2014pride,aaronson2008intergenerational}.
 The synthetic parents' characteristics serve as useful proxies for the socioeconomic environment in which these children were raised, allowing me to control for important background factors that influence educational and labor market outcomes.
@@
 
-I show the distribution of the four types of couples in Table \ref{tab:mat2}.
-US born husbands and US born wives (WW) make up the majority of couples in the sample, 96\% (5,141,737 couples).
-Hispanic husbands and wives (HH) are the second-largest group, compromising 2\% (119,749 couples).
-US born husbands and Hispanic wives (WH) $< 1\%$ (33,097 couples) of the sample and Hispanic husbands and US born wives (HW) are around 1\% (37,847 couples).
-I present the summary statistics of the parents in Table \ref{tab:synth}.
+The sample is restricted to Hispanic and non-Hispanic White individuals to avoid confounding racial factors.
+While this limits generalizability to the broader Hispanic population, it provides cleaner identification of surname effects—the study's primary focus.
+While the United States has a long history of immigration, the probability that a US-born parent in this sample is a second-generation or later immigrant from a Spanish-speaking country is very low, as shown in Table \ref{tab:mat3}.
+The majority of Hispanics in the US during this period were first- and second-generation immigrants.
+Only 3\% of native-born Americans identified as Hispanic during this period, making it statistically unlikely that an interethnic child with a native-born parent is also a second-generation Hispanic immigrant.
+
+\subsection{Variables and Measures}
+
+\textbf{Educational outcomes}: Years of education, high school diploma, associate degree, and bachelor's degree completion.
+
+\textbf{Labor market outcomes}:
+\begin{itemize}
+\item Unemployment (binary indicator from civilian labor force status)
+\item Log annual earnings (from ASEC supplement)
+\item Log weekly earnings (from outgoing rotation for hourly workers)
+\end{itemize}
+\textbf{Controls}: Age, hours worked, state-year fixed effects, and synthetic parents' characteristics (income and education) in alternative specifications.
+
+\subsection{Sample Distribution}
+Table \ref{tab:mat1} shows that WW children comprise 96\% of the sample, HH children 3\%, and interethnic children (WH and HW) make up 1.35\% with 90,325 observations—sufficient for robust analysis.
+Summary statistics and group comparisons are discussed in Section \ref{sec:hw-wh-couples-data}.
 
@@
 Let $Y_{ist}$ be the outcome of interest for person $i$ in state $s$ at time $t$.
-$HW_{ist}$ is an indicator variable equal to 1 if person $i$ has a Hispanic father and US-born mother, and 0 if person $i$ has a US-born father and Hispanic mother.
-$X_{ist}$ is a vector of controls that includes age and hours worked, $\gamma_{st}$ represents state-year fixed effects, and $\phi_{ist}$ is the error term.
+$HW_{ist}$ is an indicator variable equal to one if person $i$ has a Hispanic father and US-born mother, and zero if person $i$ has a US-born father and Hispanic mother.
+$X_{ist}$ is a vector of controls that includes age, hours worked, and synthetic parents' characteristics (income and education), $\gamma_{st}$ represents state-year fixed effects, and $\phi_{ist}$ is the error term.
 The equation for this strategy is written as follows, and the sample is restricted to individuals from WH and HW families:
@@
 If $\beta_{1} < 0$, then people who likely have a Hispanic last name have worse outcomes than people who likely have a White last name.
+This strategy follows \textcite{rubinstein2014pride}, comparing children from WH and HW interethnic families to estimate the effect of surname-based ethnic cues.
+As detailed in the Introduction, this isolates discrimination effects by holding family background constant.
 
-By comparing children from interethnic marriages, I can control for many unobservable family characteristics that may affect labor market outcomes, since these families are likely to be more similar to each other than to families with two Hispanic or two non-Hispanic White parents.
-This approach allows me to isolate the effect of surname, providing a more accurate estimate of its impact.
-The difference in means between Hispanic and non-Hispanic Whites could result from discrimination, but it could also be due to differences in innate abilities, skills, and parental investments .
-WH children will likely have a non-Hispanic last name, while HW children will likely have a Hispanic last name.
-This method was developed by \textcite{rubinstein2014pride}.
-This approach improves upon traditional Oaxaca-Blinder-Kitagawa decomposition methods, which are limited in their ability to account for unobserved differences between groups, by providing a more suitable comparison group to measure discrimination \autocite{oaxaca1973male,blinder1973wage,kitagawa1955components}.
+The central assumption underpinning my estimation strategy is that individuals born to a HW person exhibit characteristics comparable to their peers of WH descent, especially in areas like educational background, skill sets, work experiences, family culture, and parenting styles, that are significant determinants in employment opportunities, salary levels, and career advancement.
 
-The central assumption underpinning my estimation strategy is based on the hypothesis that individuals born to a HW person exhibit comparable characteristics to their peers of WH descent, especially in areas like educational background, skill sets, and work experiences that are significant determinants in employment opportunities, salary levels, and career advancement.
+A potential concern is measurement error in using CPS data to infer parental ethnicity, since it relies on place of birth rather than self-identified Hispanic origin.
+However, during the period studied, second-generation Hispanics were rare: only about 3\% of native-born Americans identified as Hispanic, see Table \ref{tab:mat3}.
+This makes it unlikely that US-born parents in interethnic unions are second-generation+ Hispanic immigrants.
 
-Two primary reasons underscore this assumption: 
-\begin{itemize}
-    \item First, there is significant selection in the marriage market.
-Since belonging to the Hispanic out-group may bring negative societal effects, it is worth exploring factors common among White women who have children with Hispanic men.
-    \item Second, fathers and mothers influence human capital accumulation differently \autocite{kimball2009risk,magruder2010intergenerational}.
-If marriage were random, then WH households might be a better environment for children.
-Using the CPS and the US Census, I will evaluate these empirical concerns.
-\end{itemize}
-
-A large body of empirical and theoretical work indicates that marriages exhibit strong selection, or `assortative matching' \autocite{averettBetterWorseRelationship2008, averettEconomicRealityBeauty1996, beckerTheoryMarriagePart1973, beckerTheoryMarriagePart1974, beckerTreatiseFamily1993, browningCollectiveUnitaryModels2006, chiapporiFatterAttractionAnthropometric2012}.
-Also, \textcite{duncanIntermarriageIntergenerationalTransmission2011} show a similar pattern of assortative matching among Mexicans in the United States.
-Interethnic WH and HW marriages generally consist of partners sharing similar characteristics, suggesting both WH and HW parents are similar in aspects relevant to labor market outcomes.
-
-Another threat to identifying a true effect could arise from measurement error in using parents’ place of birth from the CPS data as a proxy for a parent’s ethnicity and last name: The CPS notes place of birth but not ethnic or racial identity.
-It is possible that a native-born father could be a second-generation or later immigrant from a Spanish-speaking country, but this is unlikely.
-Most Hispanics from 1960 to 2000,  were first-generation immigrants, and the number of second-generation or later was very small (Table \ref{tab:mat3}).\footnote{Based on the author's calculations using US Census data.} Only 3\% of native-born Americans identified as Hispanic, making it unlikely that an interethnic child with a native-born father is also a second generation Hispanic immigrant.
+Prior research documents strong assortative mating in interethnic unions, particularly across education and income lines \autocite{beckerTheoryMarriagePart1973, chiapporiFatterAttractionAnthropometric2012, duncanIntermarriageIntergenerationalTransmission2011}.
+I confirm this pattern using synthetic parent data in Section \ref{sec:hw-wh-couples-data}.
 
 It is important to address potential statistical inference issues related to using synthetic parents' characteristics as control variables.
-Regarding the 'generated regressor' issue: The parental characteristics are included solely as controls and not for inference purposes.
+The parental characteristics are included solely as controls and not for inference purposes.
 I do not interpret these coefficients, as my primary interest lies in $\beta_1$, which captures the gap between Hispanic and White last names.
-Since I am only using these synthetic characteristics as controls and not making inference about their coefficients, this should not affect the standard errors or p-values of my coefficient of interest $\beta_1$, which is not a generated regressor.
+Since I am using these synthetic characteristics purely as controls, without drawing any conclusions from their coefficients, this should not affect the standard errors or p-values of my coefficient of interest $\beta_1$.
 
@@
 The place of birth of parents—mothers and fathers—are questions asked of all participants in the Current Population Survey (CPS) starting 1994.
-Therefore, I do not use the information of parents who are in the CPS sample at the time of the survey, but rather construct synthetic parents from historical Census data.
+Therefore, I do not use the information of parents who are in the CPS sample at the time of the survey, but rather construct synthetic parents from Census data.
 This approach means that concerns about non-random attrition of parents from the CPS sample do not apply to my analysis.
-
-Nevertheless, measurement error may still arise from the use of group averages that may not perfectly represent the actual parents of each individual in my sample.
-However, using group means typically reduces measurement error compared to individual-level measures.
-Since these variables are used as controls rather than variables of primary interest, any remaining measurement error is unlikely to substantially bias the coefficient of interest ($\beta_1$) on Hispanic last name, though it might reduce the precision of the estimates.
 
@@
 In this section, I explore the empirical data to affirm the validity of my empirical strategy.
-Table \ref{tab:synth} details the educational and economic profiles of parents from four different ethnic groups—White White (WW), White Hispanic (WH), Hispanic White (HW), and Hispanic Hispanic (HH), revealing the average outcomes and discrepancies for each group and highlighting the impact of interethnic marriages on their children’s prospects.
-These results show that there is selection in marriage, and that the differences between children born to HW and children born to WH are less extreme than those between the other two groups.
-In comparing WH and HW individuals to each other, any gaps that emerge between these otherwise similar groups are more likely to be attributable to discrimination rather than underlying differences in family background or socioeconomic status.
+Table \ref{tab:synth} details the educational and labor market outcomes of parents from four different ethnic groups—White White (WW), White Hispanic (WH), Hispanic White (HW), and Hispanic Hispanic (HH), revealing the average outcomes for each group and highlighting the impact of interethnic marriages on their children’s prospects.
+These results confirm the empirical strategy: HW and WH families are similar in background, making the outcome gaps between them more indicative of ethnic discrimination.
 
-It is important to acknowledge potential threats to identification in this approach.
-While HW and WH couples exhibit similar observable socioeconomic characteristics, they may differ systematically in unobservable dimensions that influence labor market outcomes for their children.
-These unobservable differences could include parenting styles, educational preferences, time investments in children, gender norms, or cultural attitudes that are not captured in conventional socioeconomic metrics.
+Unlike traditional decomposition methods, this comparison directly isolates the role of surname signaling, reducing the confounding effect of background disparities.
+By focusing specifically on children from interethnic marriages with similar parental characteristics but different surname ethnicities, I use a more targeted comparison group that better isolates the effect of perceived ethnicity from family background.
+Traditional decomposition approaches would struggle to separate discrimination effects from the broader socioeconomic gaps between Hispanic and White families overall.
+The smaller educational and income disparities between HW and WH families compared to HH and WW families, allowing for a clearer attribution of outcome differences to discrimination rather than unobserved heterogeneity.
+Nevertheless, I acknowledge that residual selection bias may remain, particularly regarding cultural and attitudinal factors that may systematically differ between HW and WH households despite similar educational and economic profiles.
 
-However, this approach offers distinct advantages over traditional Oaxaca-Blinder-Kitagawa decomposition methods.
-First, by focusing specifically on children from interethnic marriages with similar parental characteristics but different surname ethnicities, I use a more targeted comparison group that better isolates the effect of perceived ethnicity from family background.
-Traditional decomposition approaches would struggle to separate discrimination effects from the broader socioeconomic gaps between Hispanic and White families overall.
-Second, the smaller educational and income disparities between HW and WH families compared to HH and WW families suggest that selection bias is less severe in this comparison, allowing for a clearer attribution of outcome differences to discrimination rather than unobserved heterogeneity.
-Nevertheless, I acknowledge that residual selection bias remains a concern, particularly regarding cultural and attitudinal factors that may systematically differ between HW and WH households despite similar educational and economic profiles.
-
-Using the “synthetic” parents I constructed using the Census data, I examine the family background of the different types of children.
-WW couples have the highest level of education among the four groups: 12.58 years  for husbands and 12.36 for wives, so a WW household has 24.95 years of schooling.
-Men in HH marriages have 8.91 years of education, while women have 8.68.
-As a household, HH couples have 17.69 years of education.
-As predicted, interethnic couples marry people similar to them.
-WH husbands have 11.82 years of education, while wives have an average of 10.71, a household total of 22.68 years.
-HW husbands have 10.33 years, while wives have an average of 11.01, giving HW households a total of 21.50 years.
-Both  Hispanic men and women in interethnic couples marry white spouses who are, on average, more educated than members of the average HH couple.
-More importantly for human capital accumulation, the mother in the HW marriages (those with children most likely to have a Hispanic-sounding last name—are more educated than their WH peers.) Since mothers could be more important in the child’s human capital accumulation and education, this would suggest that a child with a Hispanic last name should complete more years of education  \autocite{gould2020does}.
-
-The data reveals that WW couples have the highest total household education, amounting to 24.95 years, significantly surpassing the 17.69 years of HH couples and illustrating the substantial educational divide between the groups.
-Interethnic couples (WH and HW) have intermediate education: WH households total 22.68 years and HW households are close behind at 21.50 years.
-Notably, HW husbands are less educated (10.33 years) compared to WH husbands (11.82 years), yet HW wives surpass their WH counterparts with 11.01 years of education, suggesting a balance in educational attainment within these marriages.
-This factor is particularly pertinent for children with Hispanic last names who might derive greater benefits from their mother’s higher education.
+Using synthetic parents constructed from Census data, I compare household educational attainment across ethnic pairings.
+WW households have 24.95 years of schooling, while HH households have 17.69.
+Interethnic couples fall in between: WH at 22.68 and HW at 21.50 years.
+Notably, HW mothers are more educated than WH mothers, which is critical for child outcomes, as maternal education is a strong predictor of human capital \autocite{gould2020does}.
 
@@
 HW women earn 2\% more than WH women.
-This reversal in the typical earning pattern speaks to the closing economic disparities between these groups and implies potentially greater economic contributions from HW women to their families, which could benefit their children when they enter the labor market.
+This reversal in the typical earning pattern highlights the closing economic disparities between these groups and implies potentially greater economic contributions from HW women to their families, which could benefit their children when they enter the labor market.
 
@@
 
-The evidence presented in this section supports a robust empirical strategy, revealing significant selection in marriage and distinctive educational and income patterns among different types of families.
-Interethnic couples have comparable levels of education and earnings, suggesting that the children of HW families are likely positioned for better educational outcomes, a critical factor for understanding discrimination in the labor market.
-The disparities between HW and WH families are considerably less pronounced than those between other groups, emphasizing the importance of comparing these children directly.
-Such comparisons shed light on the nuanced dynamics of ethnicity, education, and economic outcomes in interethnic marriages.\footnote{I present in Tables \ref{tab:synthmex} and \ref{tab:syntnonmex} the summary statistics which detail the educational and economic profiles of parents from four different ethnic groupings—White White (WW), White Hispanic (WH), Hispanic White (HW), and Hispanic Hispanic (HH) on sub-samples of Hispanics of Mexican and non-Mexican ancestries.
-I find similar results that describe a selection into inter-ethnic marriages among the two groups.}
-
-Despite higher levels of education and income among HW mothers compared to WH mothers, HW children complete an average of 0.4 fewer years of education than their WH peers (Table \ref{tab:c&p2}).
+These patterns reinforce the empirical strategy: HW and WH households are sufficiently similar to make surname-based differences informative about discrimination.\footnote{I present in Tables \ref{tab:synthmex} and \ref{tab:syntnonmex} the summary statistics which detail the educational and economic profiles of parents from four different ethnic groupings—White White (WW), White Hispanic (WH), Hispanic White (HW), and Hispanic Hispanic (HH) on sub-samples of Hispanics of Mexican and non-Mexican ancestries.
+I find similar results that describe a selection into inter-ethnic marriages among the two groups.} Despite higher levels of education and income among HW mothers compared to WH mothers, HW children complete an average of 0.4 fewer years of education than their WH peers (Table \ref{tab:c&p2}).
 This gap may suggest potential discrimination or barriers in educational access for HW children.
@@
 Column 1 in Table \ref{tab:lastname-ed-reg} is the difference in total years of education between HW children and their WH peers.
-Column 2 is the difference in the probability of not completing high school.
+Column 2 is the difference in the probability of dropping out of high school.
 Column 3 is the difference in the probability of having an associate degree.
@@
 These results suggest that while the overall educational gap between HW and WH children is small in terms of years of education, the disparities become more pronounced when considering higher education milestones, particularly for HW women.
-Since HWs have more educated mothers than WHs, we may expect higher levels of educational attainment from them \autocite{kimball2009risk, gould2020does}.
-The fact that this does not hold true could indicate potential barriers or discrimination in access to higher education for HW children.
+Given HW mothers are more educated, HW children's lower degree attainment may reflect structural barriers in higher education access---such as lower counselor encouragement, less access to AP courses, or name-based scholarship bias.
 
@@
 
-First, in Table \ref{tab:lastnamereg-weekearm}, I present the results of estimating equation \ref{eq:1a} using weekly earnings as the dependent variable for the full sample.
-Like the previous analysis, I find that a person who likely has a Hispanic last name earns 4 percentage points less than a person who likely has a non-Hispanic last name, which could be explained by educational differences.
-Second, in Table \ref{tab:lastnamereg-weekearm-mex}, I present the results for a sample of Mexican Hispanics.
-I find that among individuals of Mexican origin, those that likely have Hispanic surnames earn 3 percentage points less than those with White-sounding surnames.
-This gap becomes an imprecise zero after controlling for education and parental background.
-Third, in Table \ref{tab:lastnamereg-weekearm-nonmex}, I present the results for a sample of non-Mexican Hispanics.
-The gap between non-Mexicans that likely have a Hispanic-sounding last name and those that likely have a White-sounding last name can also be explained by educational differences.
-I find that a non-Mexican individual that likely has a Hispanic-sounding last name earns 3 percentage points less than a non-Mexican with a White-sounding last name.
-This gap also becomes an imprecise zero after controlling for education and parental background.
+Tables \ref{tab:lastnamereg-weekearm}, \ref{tab:lastnamereg-weekearm-mex}, and \ref{tab:lastnamereg-weekearm-nonmex} present results using weekly earnings as the dependent variable for the full sample, as well as for Mexican and non-Mexican Hispanic subgroups.
+Across all groups, individuals with likely Hispanic surnames earn 3–4 percentage points less than their counterparts with White-sounding names.
+However, these gaps consistently shrink to statistical insignificance once education and parental background are included as controls, suggesting that socioeconomic differences, rather than surname-based discrimination per se, account for most of the observed earnings disparity.
 
@@
 
-As the Hispanic population grows in the United States, studying discrimination becomes increasingly important.
-In this paper, I examine discrimination against Hispanics in the labor market, specifically focusing on the impact of Hispanic last names on educational and labor market outcomes.
-
+This study uses surname-based ethnic cues to show that individuals with Hispanic-sounding names face educational and earnings disadvantages, even when born into similarly situated interethnic families.
 By comparing children of interethnic marriages, I find that individuals who likely have Hispanic-sounding last names receive 0.2 fewer years of education than their counterparts who likely have White-sounding last names, representing a 1.4\% reduction.
@@
 These patterns are especially pronounced for women with Hispanic surnames, who face larger discrepancies in both total education and degree completion (14.8\% reduction in bachelor's degree attainment).
-
-This paper contributes to the education economics literature by providing evidence of disparities in access to education.
-Previous audit studies by \textcite{bergman2018education} and \textcite{gaddis2024racial} found that students with Hispanic names are less likely to receive responses from schools compared to those presumed White.
-My findings suggest that discrimination in access to education could lead to lower earnings for Hispanic workers.
-
-However, this does not necessarily indicate the absence of discrimination.
-Education itself can be influenced by bias, particularly since mothers of interethnic children with likely Hispanic surnames have higher education levels and earnings than mothers of children with non-Hispanic surnames—suggesting these children should theoretically complete more years of education.
-While my analysis identifies where discrimination manifests (primarily in higher education access), it cannot determine the specific mechanisms—such as teacher bias, institutional barriers, or social networks—through which surname-based discrimination operates.
-Further research is needed to understand these disparities.
+This study highlights how surname---especially in interethnic families---can serve as subtle but powerful signals of ethnic identity that continue to shape educational attainment and labor market inequality in the United States.

### 2025-06-13 manuscripts/10-JERP-v3/manuscript_title_etc.tex
edit manuscript

@@
 I thank Patrick Button, Willa Friedman, Chinhui Juhn, Vikram Maheshri, and Yona Rubinstein for their support and advice.
-I also thank Aimee Chin, Steven Craig, German Cubas, Elaine Liu, Fan Wang, and the participants of the Applied Microeconomics Workshop at the University of Houston, and European Society for Population Economics (ESPE) for helpful feedback.}
+I also thank Aimee Chin, Steven Craig, German Cubas, Elaine Liu, Fan Wang, and the participants of the Applied Microeconomics Workshop at the University of Houston, the European Society for Population Economics (ESPE), the anonymous referees, and the editors for helpful feedback.}
 
@@
 \newcommand{\PAPERABSTRACT}{
-% Do individuals with ethnically Hispanic names face labor market discrimination?
-In this study, I analyze the impact of Hispanic-sounding surnames on wages, considering inter-ethnic children and the distinct traits of native and Hispanic surnames.
-I examine the earnings of individuals with one White and one Hispanic parent, finding that those with Hispanic surnames often experience wage disparities.
-My findings reveal a notable wage gap favoring individuals with White surnames.
-Although males born to Hispanic fathers and White mothers earn 5 percentage points less than those born to White fathers and Hispanic mothers, this difference can be attributed to educational variances.
-However, this shouldn't imply an absence of discrimination; it may reflect discrimination in human capital accumulation.
-Furthermore, I investigate the impact of self-identifying as Hispanic on earnings.
-Men with a Spanish-sounding last name who identify as Hispanic earn notably less than those who do not, largely due to educational differences.
 
-Do individuals with Hispanic surnames face labor market discrimination?
+Do individuals with Hispanic surnames face discrimination in access to education and the labor market?
 This study analyzes the impact of likely Hispanic surnames on wages among Americans with one White and one Hispanic parent.
-I find that individuals that likely have Hispanic surnames (i.e., those with a Hispanic father) often earn less, with a notable wage gap favoring those with White-sounding surnames.
+Individuals who likely have Hispanic surnames (i.e.
+those with a Hispanic father) often complete less education and earn less, with a notable wage gap favoring those with white-sounding surnames.
 People born to Hispanic fathers and White mothers receive 0.2 fewer years of education than those born to White fathers and Hispanic mothers.
-Men born to Hispanic fathers and White mothers are 1 percentage point more likely to be unemployed, and they earn 5 percentage points less than those born to White fathers and Hispanic mothers.
-I also show that this gap can largely be explained by educational differences.
+Men born to Hispanic fathers and White mothers earn 5 percentage points less than those born to White fathers and Hispanic mothers, and they are 1 percentage point more likely to be unemployed.
+I also show that this gap can be largely explained by educational differences.
 \PAPERJEL}

### 2025-06-13 manuscripts/10-JERP-v3/submit_journal-main.tex
start working on R&R v2
edit manuscript

--- 2025-01-24 manuscripts/08-JERP-v6/submit_journal-main.tex
@@
 The data that support the findings of this study will be openly available to all researchers after the review process.
-For immediate information regarding the data and/or computer programs used for this study, please contact Hussain Hadah at hhadah@tulane.edu.
+For immediate information on the data and / or computer programs used for this study, please contact Hussain Hadah at hhadah@tulane.edu.
 
@@
 \subsection*{Conflict of Interest Disclosure}
-The author declares that they have no conflicts of interest regarding the publication of this manuscript.
+The author declares that they have no conflict of interest with respect to the publication of this manuscript.
 
@@
 \end{spacing}
+\begin{spacing}{\rrxspc}
+    Thank you for raising these concerns.
+I have significantly expanded the discussion of identification threats and the advantages of my approach compared to traditional methods.
+In Section \ref{sec:hw-wh-couples-data} (``From the Data: The Differences Between HW and WH Couples''), I now explicitly acknowledge that ``Unlike traditional decomposition methods, this comparison directly isolates the role of surname signaling, reducing the confounding effect of background disparities.
+By focusing specifically on children from interethnic marriages with similar parental characteristics but different surname ethnicities, I use a more targeted comparison group that better isolates the effect of perceived ethnicity from family background.
+Traditional decomposition approaches would struggle to separate discrimination effects from the broader socioeconomic gaps between Hispanic and White families overall.
+The smaller educational and income disparities between HW and WH families compared to HH and WW families, allowing for a clearer attribution of outcome differences to discrimination rather than unobserved heterogeneity.
+Nevertheless, I acknowledge that residual selection bias remains a concern, particularly regarding cultural and attitudinal factors that may systematically differ between HW and WH households despite similar educational and economic outcomes.''
+\end{spacing}
 
 \begin{spacing}{\rrxspc}
-        I included an extensive discussion of the differences between HW and WH couples.
-These differences, that includes cultural differences, in couple of sections throughout the manuscript.
-While I acknowledge the concern, the estimation strategy rest of the fact that selection in the marriage market decreases the differences between HW and WH couples.
-This is the reason why comparing children of intermarried couples provides a better comparison than comparing children of endogamous marriages, i.e.
-marriages where both parents are either non-Hispanic White or Hispanic.
-I also included a discussion of the relative advantages of the current approach compared to the conventional Oaxaca-Blinder-Kitagawa decomposition.
-I argue that the current approach provides a more accurate estimate of gaps that are due to discrimination.
-In fact, the concern the reviewer raises is one of the reasons why my approach is preferred over the Oaxaca-Blinder-Kitagawa since children of endogamous marriages are more likely to have different characteristics than children of intermarried couples, including those that are unobservable.
-I argue, and show from the data, that children of intermarried couples are more likely to have similar characteristics than children of endogamous couples, see sections 3.1, and 4.
+    However, I argue that this approach offers distinct advantages over traditional Oaxaca-Blinder-Kitagawa decomposition methods.
+First, by focusing specifically on children from interethnic marriages with similar parental characteristics but different surname ethnicities, I use a more targeted comparison group that better isolates the effect of perceived ethnicity from family background.
+Traditional decomposition approaches would struggle to separate discrimination effects from the broader socioeconomic gaps between Hispanic and White families overall.
+Second, the smaller educational and income disparities between HW and WH families compared to HH and WW families, allowing for a clearer attribution of outcome differences to discrimination rather than unobserved heterogeneity.
+\end{spacing}
+
+\begin{spacing}{\rrxspc}
+    The empirical evidence supports this approach: Table \ref{tab:synth} shows that interethnic couples are much more similar to each other than endogamous couples are to each other, which is exactly why comparing children of interethnic couples provides a better comparison than comparing children of endogamous marriages.
+Nevertheless, I acknowledge that residual selection bias remains a concern, particularly regarding cultural and attitudinal factors that may systematically differ between HW and WH households despite similar educational and economic outcomes.
+Moreover, it is likely that the parenting and cultural backgrounds of interethnic children are more similar to each other than those of children of endogamous marriages.
     \end{spacing}
@@
     \textbf{R1: } 2.
-The analysis sample is restricted to individuals who self-identify as ‘White’.
+The analysis sample is restricted to individuals who self-identify as 'White'.
 It is likely that this self-identification of ethnicity is endogenous to labor market outcomes.
@@
 The paper would be improved by including a more thorough discussion of this (sample) selection issue.
-Related to this is the depiction of parents born in the United States as ’White’.
+Related to this is the depiction of parents born in the United States as 'White'.
 This seems particularly over-simplifying and ignores the long history of immigration.
@@
 \begin{spacing}{\rrxspc}
-        I thank the reviewer for this comment and concern over the endogeneity of self-identification and labor market outcomes.
-I added some discussion over why I chose those who self-identify as White.
-I argue that inclusion those that are not White could contaminate the estimate of bias against Hispanics due to the racial signal.
-Moreover, \autocite{hadah2024effect} shows that there exists strong correlation between bias against Hispanics and self-reported Hispanic identity.
-Their results show that by not including Hispanics that attrit in the estimation of racial and ethnic gaps, the estimate of bias against Hispanics could be overestimated in the most biased states---more on ethnic attrition of Hispanics could be found in \textcite{antmanEthnicAttritionObserved2016,antmanEthnicAttritionAssimilation2020}.
-Regarding the comment of oversimplifying and ignoring the long history of immigration in the US, the majority of immigrants from Spanish-speaking countries in the US are first- and second-generation (see \autocite{antman2020ethnic}).
-I also show that the probability that parents are second-generation+ immigrants from Spanish-speaking countries is very low using Census data from 1960 to 2000, see Table 5.
+    Thank you for your comments and suggestions.
+I have addressed both concerns raised by the reviewer.
+First, regarding the endogeneity of self-identification, I now explicitly discuss this issue in the data section.
+I explain that ``The sample is restricted to Hispanic and non-Hispanic White individuals to avoid confounding racial factors.'' I argue that including those who do not identify as White could contaminate the estimate of bias against Hispanics due to racial signals.
+Moreover, as shown by \textcite{hadah2024effect}, there exists strong correlation between bias against Hispanics and self-reported Hispanic identity, and excluding Hispanics who experience ethnic attrition could lead to overestimated bias in the most discriminatory states.
+To address the comment on the long history of immigration in the US, I updated the manuscript to make clear that the interethnic children I am studying are US born.
+I also added the following: ``This study's focus on US-born children with foreign-born parents means the findings may not generalize to Hispanic children with US-born parents, who likely face different socioeconomic and cultural circumstances.
+The analysis does not account for heterogeneity in immigrant characteristics in different Spanish-speaking countries of origin, such as variations in educational attainment, socioeconomic status, and gender-specific migration patterns, which could influence both parental selection in migration and subsequent child outcomes.''
+\end{spacing}
+
+\begin{spacing}{\rrxspc}
+    Second, regarding the characterization of US-born parents, I have revised the language throughout the manuscript to make it clear that the analysis focuses on US-born individuals.
+I now explicitly address this concern: ``While the United States has a long history of immigration, the probability that a US-born parent in this sample is a second-generation or later immigrant from a Spanish-speaking country is very low, as shown in Table \ref{tab:mat3}.
+The majority of Hispanics in the US during this period were first- and second-generation immigrants.
+Only 3\% of native-born Americans identified as Hispanic during this period, making it statistically unlikely that an interethnic child with a native-born parent is also a second-generation Hispanic immigrant.'' I have also changed terminology throughout to use ``US-born'' rather than ``White'' when referring to native-born parents, and I clarify that this approach ``provides a cleaner identification of surname effects, which is the primary focus of this study.''
 \end{spacing}
@@
         \textbf{R1: } 3.
-The discussion on the construction of the ‘synthetic parents’ is somewhat brief.
-The paper suggests that the potential parents are matched using the birth year of the child and the parent’s place of birth to the children’s information collected in the CPS sample.
+The discussion on the construction of the 'synthetic parents' is somewhat brief.
+The paper suggests that the potential parents are matched using the birth year of the child and the parent's place of birth to the children's information collected in the CPS sample.
 While not explicitly stated in the paper, I assume that this sample of synthetic parents is used to construct mean parental education and family income at the time of the birth of the child.
-The process for matching the potential parents to the children is somewhat aggregated and their is likely considerable heterogeneity in the mean educational attainment of the potential parents within the set of children with the same birth year with the same parent’s country of birth.
+The process for matching the potential parents to the children is somewhat aggregated and their is likely considerable heterogeneity in the mean educational attainment of the potential parents within the set of children with the same birth year with the same parent's country of birth.
 The paper could be improved by exploring the possibility of improving the quality of the match of potential parents by including further characteristics to match the potential parents to their children.
@@
 \begin{spacing}{\rrxspc}
-            I appreciate this insightful suggestion.
-I have expanded the discussion and provided a detailed example of the 'synthetic parents' construction methodology.
-While additional matching dimensions would be valuable, the limitations of publicly available data constrain the potential matching criteria.
-Nevertheless, the current approach provides meaningful estimates while acknowledging these data constraints.
+        Thank you for raising this point.
+I have significantly expanded the discussion of synthetic parents construction and provided a detailed concrete example.
+In the data section, I now include: ``To illustrate the construction of synthetic parents more concretely, consider someone who was 35 years old in 1999, meaning they were born in 1964.
+If this person's mother was born in Mexico and their father was born in the United States, their 'synthetic parents' would be identified using the 1970 Census data, when the person was 6 years old.
+The 'synthetic mother' would have the average characteristics (education, income, etc.) of Mexican-born women who were married to US-born men and had children around 1964, when they were between 20 and 35 years old (meaning they were born between 1929 and 1944).
+Similarly, the 'synthetic father' would have the average characteristics of US-born men who were married to Mexican-born women and had children in that same year.''
+
+        I acknowledge the reviewer's suggestion about including additional characteristics to improve the matching quality.
+While incorporating more matching variables (such as parent's age at birth) could potentially reduce heterogeneity within the synthetic parent groups, the current approach balances precision with sample size considerations.
+Adding more matching criteria would create increasingly small cells in the Census data, potentially leading to unstable estimates or empty cells, particularly for less common parent birthplace combinations.
+Moreover, since the synthetic parent characteristics serve as controls rather than the primary variables of interest, the current level of aggregation is sufficient for the paper's main empirical strategy.
+\end{spacing}
+
+\begin{spacing}{\rrxspc}
+    I also acknowledge the limitations: ``This aggregation process introduces some heterogeneity, as individual parents vary considerably from these group-level measures.
+However, this approach provides valuable information about typical family background characteristics that would otherwise be completely unobserved.
+The synthetic parents' characteristics serve as useful proxies for the socioeconomic environment in which these children were raised, allowing me to control for important background factors that influence educational and labor market outcomes.
+While more refined matching would be ideal, the publicly available data limit the potential matching dimensions.
+Nevertheless, the current approach provides meaningful estimates while acknowledging these data constraints.''
 \end{spacing}
@@
 However, there are two main issues associated with statistical inference on the estimated parameters in model (1).
-First, there is a ‘generated regressor’ issue associated with using estimated group-level parental education.
+First, there is a 'generated regressor' issue associated with using estimated group-level parental education.
 A failure to account for this sampling variation will lead to misleadingly small standard errors.
 Second, given this group structure for parental education, it seems reasonable to assume that the model errors are uncorrelated across clusters but correlated within (potential parents) clusters.
-It is well understood that failing to account for this ‘clustering’ problem can lead to misleadingly small standard errors, narrow confidence intervals, and low p-values.
+It is well understood that failing to account for this 'clustering' problem can lead to misleadingly small standard errors, narrow confidence intervals, and low p-values.
     \end{quotation}
@@
 \begin{spacing}{\rrxspc}
-        Thank you for these important methodological points.
-Regarding the generated regressor issue: The parental characteristics are included solely as controls and not for inference purposes.
+    I am grateful to the referee for raising this important concern.
+I have expanded the discussion of these methodological issues in the empirical approach section.
+Regarding the generated regressor issue: ``The parental characteristics are included solely as controls and not for inference purposes.
 I do not interpret these coefficients, as my primary interest lies in $\beta_1$, which captures the gap between Hispanic and White last names.
-Since I am only using these synthetic characteristics as controls and not making inference about their coefficients, this should not affect the standard errors or p-values of my coefficient of interest $\beta_1$, which is not a generated regressor.
+Since I am only using these synthetic characteristics as controls and not making inference about their coefficients, this should not affect the standard errors or p-values of my coefficient of interest $\beta_1$.''
+
+    Regarding the clustering concern: To address this issue, I re-estimated the log annual earnings model using standard errors clustered at the parental birthplace level (mother's and father's birthplace), which corresponds to the level at which `synthetic' parental characteristics were estimated---see table below.
+The clustered standard errors for $\beta_1$ became slightly smaller (ranging from 0.00 to 0.01 across specifications) compared to the original specification.
+The fact that standard errors decreased rather than increased after clustering indicates that the original standard errors were, if anything, conservative for my main coefficient of interest.
+Moreover, since immigrants from Spanish-speaking countries have strong heterogeneous geographical preferences on where to live, clustering at the state level would further address the concern raised by the referee.
+For example, Cuban immigrants strongly select Florida when they move to the US, while Mexican immigrants self-select to California and Texas.
+This geographical clustering pattern actually strengthens the case for state-level clustering, as it directly addresses the referee's concern about spatial correlation in the error term.
+By clustering standard errors at the state level, I account for this within-state correlation that arises from the non-random geographical distribution of immigrant groups.
+
 \end{spacing}
+
+\newgeometry{bottom=1.0in}
+\input{10-JERP-v3/tab0000-regression-cluster}
+\restoregeometry
 
@@
 \end{spacing}
+\begin{spacing}{\rrxspc}
+    Thank you for your comments.
+I have clarified this methodological point and expanded the discussion of measurement error concerns.
+First, I want to clarify that the concern about attrition does not apply to my analysis: The synthetic parents are constructed using the birth year of the child and are not based on actual parents in the CPS sample at the time of the survey.
+The place of birth of parents---mothers and fathers---are questions asked of all participants in the Current Population Survey (CPS) starting 1994.
+Therefore, I do not use the information of parents who are in the CPS sample at the time of the survey, and I construct synthetic parents from Census data.
+Because this approach uses Census data rather than contemporaneous parent-child pairs, concerns about non-random attrition of parents from the CPS sample do not apply to my analysis.
+    \end{spacing}
 
 \begin{spacing}{\rrxspc}     
-        Thank you for these important methodological points.
-Let me clarify that the synthetic parents are constructed using the birth year of the child.
-The place of birth of parents---mothers and fathers---are questions that are asked of all participants in the Current Population Survey (CPS) starting 1994.
-Therefore, I do not use the information of parents that are in the CPS sample at the time of the survey.
-Thus, the concern over the attrition of parents in the sample is not applicable to my analysis.
-Please see the data section for a discussion of the data used in the analysis.
+    Regarding measurement error more broadly, I have expanded the discussion in the empirical approach section: ``A potential concern is measurement error in using CPS data to infer parental ethnicity, since it relies on place of birth rather than self-identified Hispanic origin.
+However, during the period studied, second-generation Hispanics were rare: only about 3\% of native-born Americans identified as Hispanic (see Table \ref{tab:mat3}, which presents Hispanic identification rates by nativity).
+This makes it unlikely that US-born parents in interethnic unions are second-generation+ Hispanic immigrants.''
     \end{spacing}
+\begin{spacing}{\rrxspc}
+    Additionally, while using census averages for synthetic parent characteristics does introduce some imprecision compared to actual parent data, this approach avoids the selection bias that would arise from using only parent-child pairs who remain in the CPS sample.
+The synthetic parents methodology, used by \textcite{rubinstein2014pride}, provides population-level averages that are more representative than the selected sample of parents who happen to be surveyed alongside their children.
+Moreover, since these synthetic parent controls are not interpreted causally, any measurement error from using group means rather than individual-level data is unlikely to bias the estimated surname effect, which is the key parameter of interest.
+    \end{spacing}
+
 \clearpage
@@
 \renewcommand{\thesubsubsection}{\Alph{section}.\arabic{subsection}.\arabic{subsubsection}}
+\setcounter{page}{1}
 

### 2025-06-13 manuscripts/10-JERP-v3/tab0000-regression-cluster.tex
edit manuscript

--- 2024-08-11 output/tables/tab05-regression.tex
@@
 \centering\centering
-\caption{Effect of Having Hispanic Last Name (Log Annual Earnings) \label{tab:lastnamereg}}
+\caption*{Effect of Having Hispanic Last Name (Log Annual Earnings)}
 \centering
@@
 \midrule
-$HW_{ist}$ & \num{-0.05}*** & \num{-0.05}** & \num{-0.05}* & \num{-0.02} & \num{-0.01}\\
- & (\num{0.02}) & (\num{0.02}) & (\num{0.02}) & (\num{0.02}) & (\num{0.02})\\
+$HW_{ist}$ & \num{-0.05}*** & \num{-0.05}*** & \num{-0.05}*** & \num{-0.02}*** & \num{-0.01}***\\
+ & (\num{0.00}) & (\num{0.00}) & (\num{0.01}) & (\num{0.00}) & (\num{0.00})\\
 Constant & \num{10.42}*** & \num{9.46}*** &  &  & \\
- & (\num{0.04}) & (\num{0.07}) &  &  & \\
+ & (\num{0.00}) & (\num{0.02}) &  &  & \\
 \midrule
@@
 Column two has the results after controlling for education.}}
-\item[5] {\setstretch{1.0}\footnotesize{Standard errors are clustered on the state level.}}
+\item[5] {\setstretch{1.0}\footnotesize{Standard errors are clustered at the mother's and father's birthplace level.}}
 \end{tablenotes}

### 2025-06-13 manuscripts/10-JERP-v3/tab12-occ.tex
edit manuscript

--- 2024-10-15 manuscripts/08-JERP-v6/tab12-occ.tex
@@
 \toprule
-\multicolumn{1}{c}{ } & \multicolumn{1}{c}{Hispanic Last Name} & \multicolumn{1}{c}{White Last Name} & \multicolumn{2}{c}{Difference-in-means: HW - WH} \\
+\multicolumn{1}{c}{ } & \multicolumn{1}{c}{Hispanic Last Name} & \multicolumn{1}{c}{White Last Name} & \multicolumn{2}{c}{Difference-in-means: WH - HW} \\
 \cmidrule(l{3pt}r{3pt}){2-2} \cmidrule(l{3pt}r{3pt}){3-3} \cmidrule(l{3pt}r{3pt}){4-5}

### 2025-06-15 manuscripts/10-JERP-v3/tab08-regression-weekearn-mex.tex
remove parental characteristics  from the controls

--- 2023-11-30 my_paper/tables/tab08-regression-weekearn-mex.tex
@@
 \begin{threeparttable}
-\begin{tabular}[t]{lcccc}
+\begin{tabular}[t]{lccc}
 \toprule
-  & \specialcell{(1) \\ Log weekly \\ earnings} & \specialcell{(2) \\ Log weekly \\ earnings} & \specialcell{(3) \\  Log weekly \\ earnings} & \specialcell{(4) \\  Log weekly \\ earnings}\\
+  & \specialcell{(1) \\ Log weekly \\ earnings} & \specialcell{(2) \\ Log weekly \\ earnings} & \specialcell{(3) \\  Log weekly \\ earnings}\\
 \midrule
-$HW_{ist}$ & \num{-0.01} & \num{-0.03}* & \num{0.00} & \num{-0.01}\\
- & (\num{0.02}) & (\num{0.02}) & (\num{0.01}) & (\num{0.01})\\
-Constant & \num{6.53}*** &  &  & \\
- & (\num{0.04}) &  &  & \\
+$HW_{ist}$ & \num{-0.01} & \num{-0.04}*** & \num{-0.02}\\
+ & (\num{0.02}) & (\num{0.01}) & (\num{0.01})\\
+Constant & \num{6.57}*** &  & \\
+ & (\num{0.03}) &  & \\
 \midrule
-\textit{Controlling for:} &  &  &  & \\
-State FE &  & X & X & X\\
-Year FE &  & X & X & X\\
-Age FE &  & X & X & X\\
-Education FE &  &  & X & X\\
-Parental Background &  &  &  & X\\
-Observations & \num{4535} & \num{4535} & \num{4535} & \num{4535}\\
+\textit{Controlling for:} &  &  & \\
+State FE &  & X & X\\
+Year FE &  & X & X\\
+State-Year FE &  & X & X\\
+Age &  & X & X\\
+Education &  &  & X\\
+Observations & \num{5105} & \num{5105} & \num{5105}\\
 \bottomrule
-\multicolumn{5}{l}{\rule{0pt}{1em}* p $<$ 0.1, ** p $<$ 0.05, *** p $<$ 0.01}\\
+\multicolumn{4}{l}{\rule{0pt}{1em}* p $<$ 0.1, ** p $<$ 0.05, *** p $<$ 0.01}\\
 \end{tabular}

### 2025-06-15 manuscripts/10-JERP-v3/tab09-regression-weekearn-nonmex.tex
remove parental characteristics  from the controls

--- 2025-06-15 manuscripts/10-JERP-v3/tab08-regression-weekearn-mex.tex
@@
 \centering\centering
-\caption{Effect of Having Hispanic Last Name: Hispanics with Mexican Ancestry  \label{tab:lastnamereg-weekearm-mex}}
+\caption{Effect of Having Hispanic Last Name: Hispanics with Non-Mexican Ancestry  \label{tab:lastnamereg-weekearm-nonmex}}
 \centering
@@
 \midrule
-$HW_{ist}$ & \num{-0.01} & \num{-0.04}*** & \num{-0.02}\\
- & (\num{0.02}) & (\num{0.01}) & (\num{0.01})\\
-Constant & \num{6.57}*** &  & \\
+$HW_{ist}$ & \num{-0.05}* & \num{-0.07}*** & \num{-0.04}**\\
+ & (\num{0.03}) & (\num{0.02}) & (\num{0.02})\\
+Constant & \num{6.75}*** &  & \\
  & (\num{0.03}) &  & \\
@@
 Education &  &  & X\\
-Observations & \num{5105} & \num{5105} & \num{5105}\\
+Parental Background &  &  & \\
+Observations & \num{3225} & \num{3225} & \num{3225}\\
 \bottomrule

### 2025-06-15 manuscripts/10-JERP-v3/tab10-regression-weekearn-cuba.tex
remove parental characteristics  from the controls

--- 2025-06-15 manuscripts/10-JERP-v3/tab08-regression-weekearn-mex.tex
@@
 \centering\centering
-\caption{Effect of Having Hispanic Last Name: Hispanics with Mexican Ancestry  \label{tab:lastnamereg-weekearm-mex}}
+\caption{Effect of Having Hispanic Last Name: Hispanics with Cuban Ancestry  \label{tab:lastnamereg-weekearm-cub}}
 \centering
@@
 \midrule
-$HW_{ist}$ & \num{-0.01} & \num{-0.04}*** & \num{-0.02}\\
- & (\num{0.02}) & (\num{0.01}) & (\num{0.01})\\
-Constant & \num{6.57}*** &  & \\
- & (\num{0.03}) &  & \\
+$HW_{ist}$ & \num{-0.09}*** & \num{-0.06} & \num{-0.01}\\
+ & (\num{0.03}) & (\num{0.06}) & (\num{0.04})\\
+Constant & \num{6.73}*** &  & \\
+ & (\num{0.04}) &  & \\
 \midrule
@@
 Education &  &  & X\\
-Observations & \num{5105} & \num{5105} & \num{5105}\\
+Observations & \num{1318} & \num{1318} & \num{1318}\\
 \bottomrule

### 2025-06-15 manuscripts/10-JERP-v3/tab16-educ-mex-regression.tex
remove parental characteristics  from the controls

--- 2024-08-28 manuscripts/07-economic-letters-v5/tables/tab13-regression.tex
@@
 \centering\centering
-\caption{Effect of Having Hispanic Last Name on Educational Outcomes\label{tab:lastname-ed-reg}}
+\caption{Effect of Having Hispanic Last Name on Educational Outcomes: Hispanics with Mexican Ancestry \label{tab:lastname-ed-reg-mex}}
 \centering
@@
 \multicolumn{5}{l}{\textit{Panel A: Full Sample}}\\
-\midrule \hspace{1em}$HW_{ist}$ & -0.20*** & 0.01 & -0.02*** & -0.03***\\
-\hspace{1em} & (0.05) & (0.01) & (0.01) & (0.01)\\
-\hspace{1em}Observations & 88377 & 90027 & 66927 & 90027\\
+\midrule \hspace{1em}$HW_{ist}$ & -0.31*** & 0.00 & -0.03*** & -0.04***\\
+\hspace{1em} & (0.04) & (0.01) & (0.01) & (0.01)\\
+\hspace{1em}Observations & 74926 & 76499 & 61273 & 76499\\
 \addlinespace[0.5em]
 \multicolumn{5}{l}{\textit{Panel B: Women}}\\
-\midrule \hspace{1em}$HW_{ist}$ & -0.25*** & 0.02 & -0.03*** & -0.04***\\
-\hspace{1em} & (0.06) & (0.01) & (0.01) & (0.01)\\
-\hspace{1em}Observations & 46516 & 47302 & 34334 & 47302\\
+\midrule \hspace{1em}$HW_{ist}$ & -0.35*** & 0.01 & -0.03*** & -0.05***\\
+\hspace{1em} & (0.05) & (0.01) & (0.01) & \vphantom{1} (0.01)\\
+\hspace{1em}Observations & 39867 & 40608 & 31761 & 40608\\
 \addlinespace[0.5em]
 \multicolumn{5}{l}{\textit{Panel C: Men}}\\
-\midrule \hspace{1em}$HW_{ist}$ & -0.16** & 0.00 & -0.02** & -0.02\\
-\hspace{1em} & (0.07) & (0.01) & (0.01) & (0.01)\\
-\hspace{1em}Observations & 41861 & 42725 & 32593 & 42725\\
-Full Sample's Mean & 13.48 & 0.43 & 0.15 & 0.26\\
-Women's Mean & 13.58 & 0.44 & 0.16 & 0.27\\
-Men's Mean & 13.38 & 0.43 & 0.14 & 0.24\\
+\midrule \hspace{1em}$HW_{ist}$ & -0.28*** & -0.01 & -0.02*** & -0.04***\\
+\hspace{1em} & (0.05) & (0.01) & (0.01) & (0.01)\\
+\hspace{1em}Full Sample's Mean & 13.15 & 0.46 & 0.14 & 0.2\\
+\hspace{1em}Observations & 35059 & 35891 & 29512 & 35891\\
+Women's Mean & 13.26 & 0.46 & 0.15 & 0.22\\
+Men's Mean & 13.03 & 0.45 & 0.13 & 0.18\\
 \bottomrule

### 2025-06-15 manuscripts/10-JERP-v3/tab17-educ-nonmex-regression.tex
remove parental characteristics  from the controls

--- 2025-01-14 manuscripts/08-JERP-v6/tab17-educ-nonmex-regression.tex
@@
 \centering\centering
-\caption{Effect of Having Hispanic Last Name: Hispanics with non-Mexican Ancestry \label{tab:lastname-ed-reg-nonmex}}
+\caption{Effect of Having Hispanic Last Name on Educational Outcomes: Hispanics with non-Mexican Ancestry \label{tab:lastname-ed-reg-nonmex}}
 \centering
@@
 \multicolumn{5}{l}{\textit{Panel A: Full Sample}}\\
-\midrule \hspace{1em}$HW_{ist}$ & -0.25*** & 0.01 & -0.04*** & -0.04***\\
+\midrule \hspace{1em}$HW_{ist}$ & -0.39*** & 0.03*** & -0.04*** & -0.07***\\
 \hspace{1em} & (0.07) & (0.01) & (0.01) & (0.01)\\
-\hspace{1em}Observations & 37893 & 38372 & 24567 & 38372\\
+\hspace{1em}Observations & 45492 & 46007 & 28517 & 46007\\
 \addlinespace[0.5em]
 \multicolumn{5}{l}{\textit{Panel B: Women}}\\
-\midrule \hspace{1em}$HW_{ist}$ & -0.35*** & 0.03** & -0.04** & -0.06***\\
+\midrule \hspace{1em}$HW_{ist}$ & -0.48*** & 0.05*** & -0.04** & -0.09***\\
 \hspace{1em} & (0.11) & (0.02) & (0.02) & (0.02)\\
-\hspace{1em}Observations & 19774 & 20039 & 12476 & 20039\\
+\hspace{1em}Observations & 23508 & 23797 & 14222 & 23797\\
 \addlinespace[0.5em]
 \multicolumn{5}{l}{\textit{Panel C: Men}}\\
-\midrule \hspace{1em}$HW_{ist}$ & -0.24** & 0.00 & -0.04* & -0.04*\\
-\hspace{1em} & (0.10) & (0.02) & (0.02) & (0.02)\\
-\hspace{1em}Full Sample's Mean & 14.07 & 0.38 & 0.16 & 0.36\\
-\hspace{1em}Observations & 18119 & 18333 & 12091 & 18333\\
-Women's Mean & 14.17 & 0.38 & 0.17 & 0.38\\
-Men's Mean & 13.97 & 0.38 & 0.16 & 0.34\\
+\midrule \hspace{1em}$HW_{ist}$ & -0.38*** & 0.03* & -0.05*** & -0.07***\\
+\hspace{1em} & (0.09) & (0.01) & (0.02) & (0.02)\\
+\hspace{1em}Full Sample's Mean & 14.18 & 0.37 & 0.17 & 0.38\\
+\hspace{1em}Observations & 21984 & 22210 & 14295 & 22210\\
+Women's Mean & 14.29 & 0.37 & 0.18 & 0.4\\
+Men's Mean & 14.06 & 0.37 & 0.16 & 0.36\\
 \bottomrule

### 2025-06-15 my_paper/tables/tab13-regression.tex
remove parental characteristics  from the controls

--- 2024-08-28 manuscripts/07-economic-letters-v5/tables/tab13-regression.tex
@@
 \multicolumn{5}{l}{\textit{Panel A: Full Sample}}\\
-\midrule \hspace{1em}$HW_{ist}$ & -0.20*** & 0.01 & -0.02*** & -0.03***\\
-\hspace{1em} & (0.05) & (0.01) & (0.01) & (0.01)\\
-\hspace{1em}Observations & 88377 & 90027 & 66927 & 90027\\
+\midrule \hspace{1em}$HW_{ist}$ & -0.39*** & 0.01 & -0.03*** & -0.06***\\
+\hspace{1em} & (0.04) & (0.01) & (0.01) & (0.01)\\
+\hspace{1em}Observations & 106905 & 108777 & 79157 & 108777\\
 \addlinespace[0.5em]
 \multicolumn{5}{l}{\textit{Panel B: Women}}\\
-\midrule \hspace{1em}$HW_{ist}$ & -0.25*** & 0.02 & -0.03*** & -0.04***\\
-\hspace{1em} & (0.06) & (0.01) & (0.01) & (0.01)\\
-\hspace{1em}Observations & 46516 & 47302 & 34334 & 47302\\
+\midrule \hspace{1em}$HW_{ist}$ & -0.42*** & 0.02* & -0.04*** & -0.07***\\
+\hspace{1em} & (0.05) & (0.01) & (0.01) & (0.01)\\
+\hspace{1em}Observations & 56115 & 57009 & 40322 & 57009\\
 \addlinespace[0.5em]
 \multicolumn{5}{l}{\textit{Panel C: Men}}\\
-\midrule \hspace{1em}$HW_{ist}$ & -0.16** & 0.00 & -0.02** & -0.02\\
-\hspace{1em} & (0.07) & (0.01) & (0.01) & (0.01)\\
-\hspace{1em}Observations & 41861 & 42725 & 32593 & 42725\\
-Full Sample's Mean & 13.48 & 0.43 & 0.15 & 0.26\\
-Women's Mean & 13.58 & 0.44 & 0.16 & 0.27\\
-Men's Mean & 13.38 & 0.43 & 0.14 & 0.24\\
+\midrule \hspace{1em}$HW_{ist}$ & -0.38*** & 0.01 & -0.03*** & -0.06***\\
+\hspace{1em} & (0.06) & (0.01) & (0.01) & (0.01)\\
+\hspace{1em}Full Sample's Mean & 13.57 & 0.42 & 0.15 & 0.27\\
+\hspace{1em}Observations & 50790 & 51768 & 38835 & 51768\\
+Women's Mean & 13.68 & 0.43 & 0.16 & 0.29\\
+Men's Mean & 13.45 & 0.41 & 0.14 & 0.25\\
+p-value test Women - Men & $p=$0.48 & $p=$0.12 & $p=$0.46 & $p=$0.41\\
 \bottomrule

### 2025-06-15 my_paper/tables/tab14-emp-regression.tex
remove parental characteristics  from the controls

--- 2024-08-28 manuscripts/07-economic-letters-v5/tab14-emp-regression.tex
@@
 \begin{threeparttable}
-\begin{tabular}[t]{lcccc}
+\begin{tabular}[t]{lccc}
 \toprule
-  & \specialcell{(1) \\ Unemployment} & \specialcell{(2) \\ Unemployment} & \specialcell{(3) \\  Unemployment} & \specialcell{(4) \\  Unemployment}\\
+  & \specialcell{(1) \\ Unemployment} & \specialcell{(2) \\ Unemployment} & \specialcell{(3) \\  Unemployment}\\
 \midrule
-$HW_{ist}$ & \num{0.01}*** & \num{0.01}** & \num{0.01}* & \num{0.01}\\
- & (\num{0.00}) & (\num{0.00}) & (\num{0.00}) & (\num{0.00})\\
-Constant & \num{0.05}*** &  &  & \\
- & (\num{0.00}) &  &  & \\
+$HW_{ist}$ & \num{0.01}*** & \num{0.01}** & \num{0.01}*\\
+ & (\num{0.00}) & (\num{0.00}) & (\num{0.00})\\
+Constant & \num{0.05}*** &  & \\
+ & (\num{0.00}) &  & \\
 \midrule
-\textit{Controlling for:} &  &  &  & \\
-Age &  & X & X & X\\
-State FE &  & X & X & X\\
-Year FE &  & X & X & X\\
-State-Year FE &  & X & X & X\\
-Education &  &  & X & X\\
-Parental Background &  &  &  & X\\
-\specialcell{HW's Unemployment} & 0.07 & 0.07 & 0.07 & 0.07\\
-Observations & \num{38090} & \num{38090} & \num{38090} & \num{38090}\\
+\textit{Controlling for:} &  &  & \\
+Age &  & X & X\\
+State FE &  & X & X\\
+Year FE &  & X & X\\
+State-Year FE &  & X & X\\
+Education &  &  & X\\
+\specialcell{HW's Mean \\ Unemployment} & 0.07 & 0.07 & 0.07\\
+Observations & \num{38204} & \num{38204} & \num{38204}\\
 \bottomrule
-\multicolumn{5}{l}{\rule{0pt}{1em}* p $<$ 0.1, ** p $<$ 0.05, *** p $<$ 0.01}\\
+\multicolumn{4}{l}{\rule{0pt}{1em}* p $<$ 0.1, ** p $<$ 0.05, *** p $<$ 0.01}\\
 \end{tabular}

### 2025-06-15 output/tables/tab05-regression.tex
remove parental characteristics  from the controls

--- 2023-04-20 output/tables/tab05-regression.tex
@@
 \begin{table}[H]
-
-\caption{Effect of Having Hispanic Last Name \label{tab:lastnamereg}}
+\centering\centering
+\caption{Effect of Having Hispanic Last Name (Log Annual Earnings) \label{tab:lastnamereg}}
 \centering
@@
 \midrule
-$HW_{ist}$ & \num{-0.06}** & \num{-0.07}** & \num{-0.05}** & \num{-0.02}\\
- & (\num{0.03}) & (\num{0.03}) & (\num{0.02}) & (\num{0.02})\\
-Constant & \num{10.40}*** & \num{9.49}*** &  & \\
+$HW_{ist}$ & \num{-0.05}*** & \num{-0.05}** & \num{-0.05}* & \num{-0.02}\\
+ & (\num{0.02}) & (\num{0.02}) & (\num{0.02}) & (\num{0.02})\\
+Constant & \num{10.42}*** & \num{9.46}*** &  & \\
  & (\num{0.04}) & (\num{0.07}) &  & \\
@@
 Hours Worked &  & X & X & X\\
+Age &  &  & X & X\\
 State FE &  &  & X & X\\
 Year FE &  &  & X & X\\
-Age FE &  &  & X & X\\
-Education FE &  &  &  & X\\
-Observations & \num{2423} & \num{2423} & \num{2423} & \num{2423}\\
+State-Year FE &  &  & X & X\\
+Education &  &  &  & X\\
+Observations & \num{3621} & \num{3621} & \num{3621} & \num{3621}\\
 \bottomrule

### 2025-06-15 output/tables/tab07-regression-weekearn.tex
remove parental characteristics  from the controls

--- 2025-06-15 manuscripts/10-JERP-v3/tab10-regression-weekearn-cuba.tex
@@
 \centering\centering
-\caption{Effect of Having Hispanic Last Name: Hispanics with Cuban Ancestry  \label{tab:lastnamereg-weekearm-cub}}
+\caption{Effect of Having Hispanic Last Name (Log Weekly Earnings) \label{tab:lastnamereg-weekearm}}
 \centering
@@
 \midrule
-$HW_{ist}$ & \num{-0.09}*** & \num{-0.06} & \num{-0.01}\\
- & (\num{0.03}) & (\num{0.06}) & (\num{0.04})\\
-Constant & \num{6.73}*** &  & \\
- & (\num{0.04}) &  & \\
+$HW_{ist}$ & \num{-0.04}*** & \num{-0.06}*** & \num{-0.03}**\\
+ & (\num{0.01}) & (\num{0.01}) & (\num{0.01})\\
+Constant & \num{6.65}*** &  & \\
+ & (\num{0.03}) &  & \\
 \midrule
@@
 Education &  &  & X\\
-Observations & \num{1318} & \num{1318} & \num{1318}\\
+Observations & \num{7444} & \num{7444} & \num{7444}\\
 \bottomrule

### 2025-06-15 output/tables/tab15-regression-occ-fe.tex
remove parental characteristics  from the controls

--- 2025-06-15 output/tables/tab05-regression.tex
@@
 State-Year FE &  &  & X & X\\
-Education &  &  &  & X\\
-Observations & \num{3621} & \num{3621} & \num{3621} & \num{3621}\\
+Occupation FE &  &  &  & X\\
+Observations & \num{3621} & \num{3621} & \num{3621} & \num{3487}\\
 \bottomrule

### 2025-06-15 programs/01-packages-wds.R

@@ -17,7 +17,7 @@ pacman::p_load(tictoc, parallel, pbapply, future,
                tibble, viridis, wesanderson, censReg, 
                rstatix, srvyr, formatR, sysfonts, 
                showtextdb, showtext, thematic, 
-               sampleSelection, textme, paletteer, 
+               sampleSelection, paletteer, 
                wesanderson, patchwork, RStata, car,
                #textme, lodown,
                BiocManager, Polychrome, effects,

### 2025-06-15 programs/26-educ-regressions.R

@@ -27,63 +27,62 @@ IndividualData  <- IndividualData  |>
           )
 ParentDummies <- c("HW", "uhrsworkly")
 
-ParentControls  <- c("IncomeQuint", "EducationQuint")
 # By generation
 feols(Education ~ HW  , vcov = ~statefip, data = IndividualData)
 feols(Education ~ HW  | year + statefip, vcov = ~statefip, data = IndividualData)
-feols(Education ~ HW  +  age +.[ParentControls]  | year + statefip, vcov = ~statefip, data = IndividualData)
-marginaleffects::avg_slopes(feols(Education ~ HW  +  age +.[ParentControls]  | year*statefip, vcov = ~statefip, data = IndividualData))
-feols(Education ~ HW*year  + .[ParentControls]  | year*statefip, vcov = ~statefip, data = IndividualData)
-feols(Education ~ HW*year  + HW*statefip  + .[ParentControls]  | year*statefip, vcov = ~statefip, data = IndividualData)
+feols(Education ~ HW  +  age | year + statefip, vcov = ~statefip, data = IndividualData)
+marginaleffects::avg_slopes(feols(Education ~ HW  +  age | year*statefip, vcov = ~statefip, data = IndividualData))
+feols(Education ~ HW*year  | year*statefip, vcov = ~statefip, data = IndividualData)
+feols(Education ~ HW*year  + HW*statefip  | year*statefip, vcov = ~statefip, data = IndividualData)
 
 feols(hs_dropout ~ HW  , vcov = ~statefip, data = IndividualData)
 feols(hs_dropout ~ HW  | year + statefip, vcov = ~statefip, data = IndividualData)
-feols(hs_dropout ~ HW  +  age +.[ParentControls]  | year + statefip, vcov = ~statefip, data = IndividualData)
-feols(hs_dropout ~ HW  +  age +.[ParentControls]  | year*statefip, vcov = ~statefip, data = IndividualData)
-feols(hs_dropout ~ HW*year  + .[ParentControls]  | year*statefip, vcov = ~statefip, data = IndividualData)
-feols(hs_dropout ~ HW*year  + HW*statefip  + .[ParentControls]  | year*statefip, vcov = ~statefip, data = IndividualData)
+feols(hs_dropout ~ HW  +  age | year + statefip, vcov = ~statefip, data = IndividualData)
[…]

### 2025-06-15 programs/95-make-all.R

@@ -30,8 +30,8 @@ options(modelsummary_factory_latex = "kableExtra")
 # install_github("s3alfisc/fwildclusterboot")
 library(fwildclusterboot)
 
-set.seed(12345)
-dqrng::dqset.seed(12345)
+set.seed(34345)
+dqrng::dqset.seed(123)
 
 # main scripts
 source(file.path(programs,"01-packages-wds.r")) # set up package

### 2025-06-15 manuscripts/10-JERP-v3/manuscript_main.tex
some edits to ref response

@@
 $HW_{ist}$ is an indicator variable equal to one if person $i$ has a Hispanic father and US-born mother, and zero if person $i$ has a US-born father and Hispanic mother.
-$X_{ist}$ is a vector of controls that includes age, hours worked, and synthetic parents' characteristics (income and education), $\gamma_{st}$ represents state-year fixed effects, and $\phi_{ist}$ is the error term.
+$X_{ist}$ is a vector of controls that includes age, and hours worked, $\gamma_{st}$ represents state-year fixed effects, and $\phi_{ist}$ is the error term.
 The equation for this strategy is written as follows, and the sample is restricted to individuals from WH and HW families:
@@
 I confirm this pattern using synthetic parent data in Section \ref{sec:hw-wh-couples-data}.
-
-It is important to address potential statistical inference issues related to using synthetic parents' characteristics as control variables.
-The parental characteristics are included solely as controls and not for inference purposes.
-I do not interpret these coefficients, as my primary interest lies in $\beta_1$, which captures the gap between Hispanic and White last names.
-Since I am using these synthetic characteristics purely as controls, without drawing any conclusions from their coefficients, this should not affect the standard errors or p-values of my coefficient of interest $\beta_1$.
-
-The synthetic parents are constructed using the birth year of the child and are not based on actual parents in the CPS sample at the time of the survey.
-The place of birth of parents—mothers and fathers—are questions asked of all participants in the Current Population Survey (CPS) starting 1994.
-Therefore, I do not use the information of parents who are in the CPS sample at the time of the survey, but rather construct synthetic parents from Census data.
-This approach means that concerns about non-random attrition of parents from the CPS sample do not apply to my analysis.
 
@@
 Column 4 is the difference in the probability of having a bachelor’s degree.
-All regressions include controls for age, parental education and income, and state-year fixed effects.
+All regressions include controls for age and state-year fixed effects.
 
 There is a significant gap in total years of education between HW and WH children.
-HW children receive 0.2 fewer years of education than WH children.
+HW children receive 0.39 fewer years of education than WH children.
 The gap between HW and WH women is larger than the gap between HW and WH men.
-Women with a Hispanic last name receive 0.25 fewer years of education than WH women.
-The gap between HW and WH men is 0.16 years.
+Women with a Hispanic last name receive 0.42 fewer years of education than WH women.
+The gap between HW and WH men is 0.38 years.
 
@@
 The gap between HW and WH high school dropouts is statistically insignificant (1 percentage point).
-The same is true for HW women (2 percentage points) and men (0 percentage points).
+The gap between HW and WH women is statistically significant and equal to 2 percentage points, while the gap between HW and WH men is statistically insignificant.
 
 Notable differences emerge for higher education outcomes.
-HW children are 2 percentage points less likely to earn an associate degree compared to their WH peers, representing a 13.3\% reduction relative to the WH associate degree rate of 15\%.
-Similarly, they are 3 percentage points less likely to earn a bachelor's degree, a 13\% reduction compared to the WH bachelor's degree rate of 23\%.
-These differences are slightly larger for HW women, who are 3 percentage points less likely to earn an associate degree and 4 percentage points less likely to earn a bachelor’s degree.
-For HW men, the gap is 2 percentage points for an associate degree and statistically insignificant for a bachelor’s degree.
+HW children are 3 percentage points less likely to earn an associate degree compared to their WH peers, representing a 20\% reduction relative to the WH associate degree rate of 15\%.
+Similarly, they are 6 percentage points less likely to earn a bachelor's degree, a 22.2\% reduction.
+These differences are slightly larger for HW women, who are 4 percentage points less likely to earn an associate degree and 7 percentage points less likely to earn a bachelor’s degree.
+For HW men, the gap is 3 percentage points for an associate degree and 6 percentage points for a bachelor’s degree.
 
 These results suggest that while the overall educational gap between HW and WH children is small in terms of years of education, the disparities become more pronounced when considering higher education milestones, particularly for HW women.
-Given HW mothers are more educated, HW children's lower degree attainment may reflect structural barriers in higher education access---such as lower counselor encouragement, less access to AP courses, or name-based scholarship bias.
+Given HW mothers are, on average, more educated, HW children's lower degree attainment may reflect structural barriers in higher education access---such as lower counselor encouragement, less access to AP courses, or name-based scholarship bias.
 
@@
 Column 1 in Table \ref{tab:lastnamereg} is the average crude earnings gap in log annual earnings between HW workers and their WH peers.
-In the next 4 columns, I introduce the results with controls for hours worked, state fixed effects (FE), year FE, age FE, education FE, and parental background.
+In the next 4 columns, I introduce the results with controls for hours worked, state fixed effects (FE), year FE, age FE, and education FE.
 
@@
 When education is included as a control (Column 3), the gap remains at 1 percentage point.
-Finally, after controlling for parental background (Column 4), the 1 percentage point difference in unemployment rates becomes statistically insignificant.
-These results suggest that while there is an initial unemployment gap associated with likely having a Hispanic last name, much of this difference can be explained by factors such as education and parental background and does not necessarily reflect discrimination.
+These results suggest that while there is an initial unemployment gap associated with likely having a Hispanic last name, much of this difference can be explained by factors such as educational background and does not necessarily reflect discrimination.
 The mean unemployment rate for individuals with Hispanic last names (HW) is 7\% across all specifications.
@@
 This study uses surname-based ethnic cues to show that individuals with Hispanic-sounding names face educational and earnings disadvantages, even when born into similarly situated interethnic families.
-By comparing children of interethnic marriages, I find that individuals who likely have Hispanic-sounding last names receive 0.2 fewer years of education than their counterparts who likely have White-sounding last names, representing a 1.4\% reduction.
-This gap is larger for women (0.25 years, or 1.8\%) than for men (0.16 years, or 1.2\%).
+By comparing children of interethnic marriages, I find that individuals who likely have Hispanic-sounding last names receive 0.39 fewer years of education than their counterparts who likely have White-sounding last names, representing a 2.9\% reduction.
+This gap is larger for women (0.42 years, or 2.9\%) than for men (0.38 years, or 2.8\%).
 In labor market outcomes, those with likely Hispanic surnames are 1 percentage point more likely to be unemployed (a 14.3\% increase from the mean rate) and earn 5 percentage points (4.9\%) less than their counterparts.
 
-While the overall impact on total years of education appears modest, the disparities become more pronounced in higher education attainment—with a 13.3\% reduction in associate degree completion and 13\% reduction in bachelor's degree attainment.
+While the overall impact on total years of education appears modest, the disparities become more pronounced in higher education attainment—with a 15\% reduction in associate degree completion and 22.2\% reduction in bachelor's degree attainment.
 These patterns are especially pronounced for women with Hispanic surnames, who face larger discrepancies in both total education and degree completion (14.8\% reduction in bachelor's degree attainment).

### 2025-06-15 manuscripts/10-JERP-v3/submit_journal-main.tex
Updates from Overleaf
some edits to ref response

@@
 I have expanded the discussion of these methodological issues in the empirical approach section.
-Regarding the generated regressor issue: ``The parental characteristics are included solely as controls and not for inference purposes.
-I do not interpret these coefficients, as my primary interest lies in $\beta_1$, which captures the gap between Hispanic and White last names.
-Since I am only using these synthetic characteristics as controls and not making inference about their coefficients, this should not affect the standard errors or p-values of my coefficient of interest $\beta_1$.''
-
-    Regarding the clustering concern: To address this issue, I re-estimated the log annual earnings model using standard errors clustered at the parental birthplace level (mother's and father's birthplace), which corresponds to the level at which `synthetic' parental characteristics were estimated---see table below.
-The clustered standard errors for $\beta_1$ became slightly smaller (ranging from 0.00 to 0.01 across specifications) compared to the original specification.
-The fact that standard errors decreased rather than increased after clustering indicates that the original standard errors were, if anything, conservative for my main coefficient of interest.
-Moreover, since immigrants from Spanish-speaking countries have strong heterogeneous geographical preferences on where to live, clustering at the state level would further address the concern raised by the referee.
-For example, Cuban immigrants strongly select Florida when they move to the US, while Mexican immigrants self-select to California and Texas.
-This geographical clustering pattern actually strengthens the case for state-level clustering, as it directly addresses the referee's concern about spatial correlation in the error term.
-By clustering standard errors at the state level, I account for this within-state correlation that arises from the non-random geographical distribution of immigrant groups.
+Regarding the generated regressor issue, I updated the regression results in the manuscript with results that do not include the specification where  `synthetic parents` are included in the controls.
+This alleviates the concern regarding generated, as they are no longer in the regression.
 
 \end{spacing}
-
-\newgeometry{bottom=1.0in}
-\input{10-JERP-v3/tab0000-regression-cluster}
-\restoregeometry
 
@@
 I have clarified this methodological point and expanded the discussion of measurement error concerns.
-First, I want to clarify that the concern about attrition does not apply to my analysis: The synthetic parents are constructed using the birth year of the child and are not based on actual parents in the CPS sample at the time of the survey.
+First, I want to clarify that the concern about attrition does not apply to my analysis: Synthetic parents are constructed using the birth year of the child and are not based on actual parents in the CPS sample at the time of the survey.
 The place of birth of parents---mothers and fathers---are questions asked of all participants in the Current Population Survey (CPS) starting 1994.
@@
 \begin{spacing}{\rrxspc}     
-    Regarding measurement error more broadly, I have expanded the discussion in the empirical approach section: ``A potential concern is measurement error in using CPS data to infer parental ethnicity, since it relies on place of birth rather than self-identified Hispanic origin.
+    Regarding measurement error more broadly, I have expanded the discussion in the empirical approach section to include the following: ``A potential concern is measurement error in using CPS data to infer parental ethnicity, since it relies on place of birth rather than self-identified Hispanic origin.
 However, during the period studied, second-generation Hispanics were rare: only about 3\% of native-born Americans identified as Hispanic (see Table \ref{tab:mat3}, which presents Hispanic identification rates by nativity).
 This makes it unlikely that US-born parents in interethnic unions are second-generation+ Hispanic immigrants.''
-    \end{spacing}
-\begin{spacing}{\rrxspc}
-    Additionally, while using census averages for synthetic parent characteristics does introduce some imprecision compared to actual parent data, this approach avoids the selection bias that would arise from using only parent-child pairs who remain in the CPS sample.
-The synthetic parents methodology, used by \textcite{rubinstein2014pride}, provides population-level averages that are more representative than the selected sample of parents who happen to be surveyed alongside their children.
-Moreover, since these synthetic parent controls are not interpreted causally, any measurement error from using group means rather than individual-level data is unlikely to bias the estimated surname effect, which is the key parameter of interest.
     \end{spacing}

### 2025-06-15 manuscripts/10-JERP-v3/tab09-regression-weekearn-nonmex.tex
some edits to ref response

@@
 Education &  &  & X\\
-Parental Background &  &  & \\
 Observations & \num{3225} & \num{3225} & \num{3225}\\

### 2025-06-15 manuscripts/10-JERP-v3/tab13-regression.tex
edit manuscript
some edits to ref response

--- 2024-08-28 manuscripts/07-economic-letters-v5/tables/tab13-regression.tex
@@
 \multicolumn{5}{l}{\textit{Panel A: Full Sample}}\\
-\midrule \hspace{1em}$HW_{ist}$ & -0.20*** & 0.01 & -0.02*** & -0.03***\\
-\hspace{1em} & (0.05) & (0.01) & (0.01) & (0.01)\\
-\hspace{1em}Observations & 88377 & 90027 & 66927 & 90027\\
+\midrule \hspace{1em}$HW_{ist}$ & -0.39*** & 0.01 & -0.03*** & -0.06***\\
+\hspace{1em} & (0.04) & (0.01) & (0.01) & (0.01)\\
+\hspace{1em}Observations & 106905 & 108777 & 79157 & 108777\\
 \addlinespace[0.5em]
 \multicolumn{5}{l}{\textit{Panel B: Women}}\\
-\midrule \hspace{1em}$HW_{ist}$ & -0.25*** & 0.02 & -0.03*** & -0.04***\\
-\hspace{1em} & (0.06) & (0.01) & (0.01) & (0.01)\\
-\hspace{1em}Observations & 46516 & 47302 & 34334 & 47302\\
+\midrule \hspace{1em}$HW_{ist}$ & -0.42*** & 0.02* & -0.04*** & -0.07***\\
+\hspace{1em} & (0.05) & (0.01) & (0.01) & (0.01)\\
+\hspace{1em}Observations & 56115 & 57009 & 40322 & 57009\\
 \addlinespace[0.5em]
 \multicolumn{5}{l}{\textit{Panel C: Men}}\\
-\midrule \hspace{1em}$HW_{ist}$ & -0.16** & 0.00 & -0.02** & -0.02\\
-\hspace{1em} & (0.07) & (0.01) & (0.01) & (0.01)\\
-\hspace{1em}Observations & 41861 & 42725 & 32593 & 42725\\
-Full Sample's Mean & 13.48 & 0.43 & 0.15 & 0.26\\
-Women's Mean & 13.58 & 0.44 & 0.16 & 0.27\\
-Men's Mean & 13.38 & 0.43 & 0.14 & 0.24\\
+\midrule \hspace{1em}$HW_{ist}$ & -0.38*** & 0.01 & -0.03*** & -0.06***\\
+\hspace{1em} & (0.06) & (0.01) & (0.01) & (0.01)\\
+\hspace{1em}Observations & 50790 & 51768 & 38835 & 51768\\
+Full Sample's Mean & 13.57 & 0.42 & 0.15 & 0.27\\
+Women's Mean & 13.68 & 0.43 & 0.16 & 0.29\\
+Men's Mean & 13.45 & 0.41 & 0.14 & 0.25\\
+p-value test Women - Men & $p=$0.48 & $p=$0.12 & $p=$0.46 & $p=$0.41\\
 \bottomrule

### 2025-06-15 manuscripts/10-JERP-v3/tab16-educ-mex-regression.tex
some edits to ref response

@@
 \hspace{1em} & (0.05) & (0.01) & (0.01) & (0.01)\\
-\hspace{1em}Full Sample's Mean & 13.15 & 0.46 & 0.14 & 0.2\\
 \hspace{1em}Observations & 35059 & 35891 & 29512 & 35891\\
+Full Sample's Mean & 13.15 & 0.46 & 0.14 & 0.2\\
 Women's Mean & 13.26 & 0.46 & 0.15 & 0.22\\

### 2025-06-15 manuscripts/10-JERP-v3/tab17-educ-nonmex-regression.tex
some edits to ref response

@@
 \hspace{1em} & (0.09) & (0.01) & (0.02) & (0.02)\\
-\hspace{1em}Full Sample's Mean & 14.18 & 0.37 & 0.17 & 0.38\\
 \hspace{1em}Observations & 21984 & 22210 & 14295 & 22210\\
+Full Sample's Mean & 14.18 & 0.37 & 0.17 & 0.38\\
 Women's Mean & 14.29 & 0.37 & 0.18 & 0.4\\

### 2025-06-20 manuscripts/09-JERP-v2/manuscript_main.tex
edit synthetic parents

@@
 The CPS does not provide data on parents’ characteristics, essential to determine the family background, but the 1960–2000 US Census data includes parents’ place of birth, race, and ethnicity.
-I employ this information to construct "synthetic parents" using a method developed by \textcite{rubinstein2014pride} linking husbands and wives  in the census data.
-Assuming that parents have children 25–40 years old, I link these “parents” using parents’ places of birth, birth year of "children," and the parents’ places of birth.
+I employ this information to construct ``synthetic parents'' using a method developed by \textcite{rubinstein2014pride} linking husbands and wives  in the census data.
+Assuming that parents have children 25–40 years old, I link these “parents” using parents’ places of birth, birth year of ``children,'' and the parents’ places of birth.
 

### 2025-06-20 manuscripts/10-JERP-v3/manuscript_main.tex
edit synthetic parents

@@
 
+Importantly, these synthetic parent measures are constructed to show that interethnic couples are more comparable to each other in terms of observable characteristics, thereby reducing potential bias in comparisons.
+The synthetic parent variables are used exclusively for descriptive summary statistics and are not included as control variables in the main regression analyses.
+
 This aggregation process introduces some heterogeneity, as individual parents vary considerably from these group-level measures.
-However, this approach provides valuable information about typical family background characteristics that would otherwise be completely unobserved \autocite{rubinstein2014pride,aaronson2008intergenerational}.
+However, this approach provides valuable information on typical characteristics of family history that would otherwise be completely unobserved \autocite{rubinstein2014pride,aaronson2008intergenerational}.
 The synthetic parents' characteristics serve as useful proxies for the socioeconomic environment in which these children were raised, allowing me to control for important background factors that influence educational and labor market outcomes.
-While more refined matching would be ideal, the publicly available data limit the potential matching dimensions.
-Nevertheless, the current approach provides meaningful estimates while acknowledging these data constraints.
+While the synthetic parents' characteristics provide useful information about the socioeconomic environment in which these children were raised, they are not used as control variables in the regression analyses.
+More refined matching would be ideal, the publicly available data limit the potential matching dimensions.
+Nevertheless, the current approach provides meaningful information from the data on the different type of couples, i.e.
+interethnic versus endogamous.
 
@@
 \end{itemize}
-\textbf{Controls}: Age, hours worked, state-year fixed effects, and synthetic parents' characteristics (income and education) in alternative specifications.
+\textbf{Controls}: Age, hours worked, and state-year fixed effects.
 

### 2025-06-20 manuscripts/10-JERP-v3/submit_journal-main.tex
edit synthetic parents

@@
 Adding more matching criteria would create increasingly small cells in the Census data, potentially leading to unstable estimates or empty cells, particularly for less common parent birthplace combinations.
-Moreover, since the synthetic parent characteristics serve as controls rather than the primary variables of interest, the current level of aggregation is sufficient for the paper's main empirical strategy.
-\end{spacing}
-
-\begin{spacing}{\rrxspc}
-    I also acknowledge the limitations: ``This aggregation process introduces some heterogeneity, as individual parents vary considerably from these group-level measures.
-However, this approach provides valuable information about typical family background characteristics that would otherwise be completely unobserved.
-The synthetic parents' characteristics serve as useful proxies for the socioeconomic environment in which these children were raised, allowing me to control for important background factors that influence educational and labor market outcomes.
-While more refined matching would be ideal, the publicly available data limit the potential matching dimensions.
-Nevertheless, the current approach provides meaningful estimates while acknowledging these data constraints.''
+Moreover, since the synthetic parent characteristics are used exclusively for descriptive summary statistics and not as control variables in the main regression analyses, the current level of aggregation is appropriate for the paper's empirical strategy.
 \end{spacing}
@@
 This alleviates the concern regarding generated, as they are no longer in the regression.
+I also updated the manuscript to make sure that it is clear that `synthetic parents` parents are only used for descriptive summary statistics to enhance comparability between interethnic families and are not included as control variables in any of the main regressions.
 

### 2025-06-20 manuscripts/09-JERP-v2/submit_journal-main.tex
.....

--- 2025-01-24 manuscripts/08-JERP-v6/submit_journal-main.tex
@@
     \begin{spacing}{\rrxspc}
-        Thank you for these important methodological points.
-Regarding the generated regressor issue: The parental characteristics are included solely as controls and not for inference purposes.
-I do not interpret these coefficients, as my primary interest lies in $\beta_1$, which captures the gap between Hispanic and White last names.
-Since I am only using these synthetic characteristics as controls and not making inference about their coefficients, this should not affect the standard errors or p-values of my coefficient of interest $\beta_1$, which is not a generated regressor.
+        I am grateful to the referee for raising this important concern.
+I have expanded the discussion of these methodological issues in the empirical approach section.
+Regarding the generated regressor issue, I updated the regression results in the manuscript with results that do not include the specification where ‘synthetic parents‘ are included in the controls.
+This alleviates the concern regarding generated, as they are no longer in the regression.
     \end{spacing}

### 2025-06-24 manuscripts/10-JERP-v3/manuscript_main.tex
Updates from Overleaf

@@
 The analysis does not account for heterogeneity in immigrant characteristics in different Spanish-speaking countries of origin, such as variations in educational attainment, socioeconomic status, and gender-specific migration patterns, which could influence both parental selection in migration and subsequent child outcomes.}
+
+My findings reveal significant penalties associated with Hispanic surnames.
+Native-born individuals who likely have Hispanic-sounding last names receive 0.39 fewer years of education than their counterparts with White-sounding last names, with particularly pronounced effects in higher education: they are 20\% less likely to earn an associate degree and 22\% less likely to earn a bachelor's degree.
+In the labor market, those with Hispanic surnames face a 14\% higher unemployment rate and earn 5\% less annually than their peers with White surnames.
+These disparities are especially pronounced for women, who experience larger gaps in both educational attainment and labor market outcomes, highlighting how ethnic and gender discrimination intersect to compound disadvantages.
 
